### 2022-03-25 Manuscript/main.tex

\documentclass[fleqn,10pt]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\title{NMRlipids Databank}

\author[1,*]{Anne Kiirikki}
\author[1,*]{O. H. Samuli Ollila}
%\author[1,2,+]{Christine Author}
%\author[2,+]{Derek Author}
\affil[1]{University of Helsinki, Institute of Biotechonology, Helsinki, Finland}
%\affil[2]{Affiliation, department, city, postcode, country}

\affil[*]{samuli.ollila@helsinki.fi}

%\affil[+]{these authors contributed equally to this work}

%\keywords{Keyword1, Keyword2, Keyword3}

\begin{abstract}
We present a databank of lipid bilayer simulations from the NMRlipids open collaboration project.
\end{abstract}
\begin{document}

\flushbottom
\maketitle
% * <john.hammersley@gmail.com> 2015-02-09T12:07:31.197Z:
%
%  Click the title above to edit the author information and abstract
%
\thispagestyle{empty}

%\noindent Please note: Abbreviations should be introduced at the first mention in the main text – no abbreviations lists. Suggested structure of main text (not enforced) is provided below.

\section{Introduction}

%The demand for sharing and reusing of MD simulation data is increasing, but practical solution remains unclear due to unsolved issues in data storage and indexing.

The importance of sharing MD simulation data following the FAIR principles~\cite{wilkinson16} has been widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20}, and databanks are emerging~\cite{meyer10,kamp10,hospital16,mixcoha16,newport19,bekker20,espigares20,leston22}.
The relevance of quality evaluation of simulation trajectories in databanks regarding technical details of simulations and accuracy of the underlying physical description of the system (force field) has become evident~\cite{tai04,meyer10,hospital20} and such quality evaluation has in some cases also been implemented~\cite{meyer10,hospital16}. However, straightforward quality comparisons between individual simulations or force fields within these databanks remain challenging. 
While importance of such databanks for MD simulations is widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} and different kinds of approaches are emerging~\cite{meyer10,kamp10,hospital16,mixcoha16,newport19,bekker20,espigares20,leston22}, generally accepted protocols and best practices are still under active development.

Here we present a solution for lipid bilayers based on overlay databank structure illustrated in Fig.~\ref{overlay}.

%The Introduction section, of referenced text\cite{Figueredo:2009dg} expands on the background of the work (some overlap with the Abstract is acceptable). The introduction should not include subheadings.

\section{Results}

%Up to three levels of \textbf{subheading} are permitted. Subheadings should not be numbered.

\subsection{Quality evaluation of force fields}

The quality of each simulation is evaluated against NMR order parameters.
Indexing of experimental data to automatize evaluation is currently in progress by Anne Kiirikki.

\subsection{Water permeability across membranes}

Averaging water density profiles over all simulations in the databank enables us to calculate the barrier for the water penetration through lipid bilayers.

\subsection{Spin relaxation rates of confined water close to membranes}

Usefulness of the databank beyond MD simulation experts is demonstrated by analysing water spin relaxation times close to bilayers which are used in MRI imaging.

%Example text under a subsection. Bulleted lists may be used where appropriate, e.g.

%\begin{itemize}
%\item First item
%\item Second item
%\end{itemize}

%\subsubsection*{Third-level section}

%Topical subheadings are allowed.

\section{Discussion}

%The Discussion should be succinct and must not contain subheadings.

\section{Methods}

%Topical subheadings are allowed. Authors must ensure that their Methods section includes adequate experimental and characterization data necessary for others in the field to reproduce their work.

\subsection{Structure of the databank}

\begin{figure}
  \includegraphics[width=\textwidth]{figures/DataBankStructure.pdf}
  \caption{Structure of the NMRlipids databank}\label{DatabankStructure}
\end{figure}

NMRlipids databank is a overlay databank composed of files that contain all the essential information
on the molecular dynamics simulation trajectories that are listed in the databank.
The location publicly avaiable raw MD simulation data is flexible.
The NMRlipids databank files contain all the essential information on the
original data that enable wide range of applications.
%and all essential information on simulations needed in further use and automatic analysis.

Each simulation system is identified using the hash of original trajectory and topology file. The index files are stored in folder structure based on the identity codes of the simulations.

\subsection{Quality evaluation}

For quality evaluation, each simulation is connected to a experimental dataset is similar conditions. Currently experimental and simulation datasets are paired if molar concentrations of all molecules are within $\pm$5 percentage units, charged lipids have the same counterions, and temperature is within $\pm$2 degrees. If lipid to water ratio is above ??, the systems are considered fully hydrated. Otherwise also hydration level is considered.

Quality of lipid conformational ensembles are evaluated against C-H bond order parameters from NMR experiments~\cite{ollila16}. As a first step, the quality of each order parameter is evaluated using the negative logarithm of probability to find the simulated order parameter within the experimental error bars
\begin{equation}
    S_q = -\log_{10}(P),    
\end{equation}
where the probability is calculated from the normal distribution
\begin{equation}\label{gaussian}
    P = \int_{S_{\rm exp}-\Delta S_{\rm exp}}^{S_{\rm exp}+
    \Delta S_{\rm exp}}  \frac{1}{\sigma \sqrt{2\pi}} e^{-\frac{1}{2}(\frac{x-\mu}{\sigma})^2} {\rm d}x.
\end{equation}
$S_{\rm exp}$ and $\Delta S_{\rm exp}$ are experimental order parameter and its error, and $\mu$ and $\sigma$ are the mean order parameter and its standard deviation from simulations. The quality measure $S_q$ approaches to zero when probability for agreement between simulation and experimental results approach one, and increases when simulated and experimental values diverge. The accuracy of $\pm$0.02 is currently assumed for all experimental order parameters~\cite{ollila16}. Because phospholipids sample their conformational ensemble within nanosecond timescale~\cite{ferreira15}, all simulations in the databank would be sufficiently long to sample the realistic conformational phase of individual lipids. However, some force fields exhibit too slow dynamics which leads to large error bars in order parameter values~\cite{antila21a}. Because large error bars widen the gaussian distribution in Eq.~\ref{gaussian} thereby artificially increasing the probability to find the simulated value within experimental error bars, the order parameters with error bars larger than the experimental error 0.02 are not included in the quality evaluation.

The quality of separate fragments in each lipid type within a simulation are then evaluated by averaging individual order parameter qualities over C-H bond belonging to that fragment and dividing this with the percentage of order parameters for which the quality is available within the fragment, $p$
\begin{equation}
    S_q^{\rm frag}[{\rm lipid}] = \frac{\langle S_q[{\rm lipid}]\rangle_{\rm frag}}{p_{\rm frag}[{\rm lipid}]},
\end{equation}
 where frag can be {\it sn}-1, {\it sn}-2, headgroup or total (all order parameters within a molecule). The overall quality of different fragments in a simulation are then defined as a molar fraction weighted average over different lipid components
\begin{equation}
    S_q^{\rm frag} = \sum_{\rm lipid} \chi_{\rm lipid} \langle S_q^{\rm frag}[{\rm lipid}]\rangle_{\rm lipid} ,
\end{equation}
where $\chi_{\rm lipid}$ is the molar fraction of a lipid in the bilayer.

Qualities of form factors were evaluated with the same approach that was used in SIMtoEXP program~\cite{kucerka10}. Because experiments give form factors only in relative scale those were scaled to the simulation data with absolute scale using equation
\begin{equation}
    k_e = \frac{\sum_{i=1}^{N_q} \frac{|F_s(q_i)||F_e(q_i)|}{(\Delta F_e(q_i))^2}}{\sum_{i=1}^{N_q} \frac{|F_e(q_i)|^2}{(\Delta F_e(q_i))^2}}.
\end{equation}
The quality of each form factor was then calculated from the equation
\begin{equation}
    \chi^2 = \frac{\sqrt{\sum_{i=1}^{N_q}(|F_s(q_i)|-k_e|F_e(q_i)|)^2/(\Delta F_e(q_i))^2}}{\sqrt{N_q-1}},
\end{equation}
where $F_s$ and $F_e$ are form factors from a simulation and experiment, respectively, and summation goes over the experimentally available $N_q$ points.

\subsection{Analysis of the data in databank}

Index files in the datanbank contain all the essential information to perform analyses from the simulations. Systems under interest can be selected automatically browsing the index files and filtering the desired properties.

Analysis results can be most conveniently stored to a new datanbank with identical indexing as the original databank. The results can be then easily shared and browsed indentically to the original databank.

\subsection{Indexing the simulation data}

%\subsubsection*{AddData.py}
AddData.py is a script that builds a database that contains a dictionary file and analysis data of each simulation. The dictionary file contains information about the simulation. The script also calculates order parameters of all CH bonds of the lipids in the simulation. To add a simulation it must be first uploaded to Zenodo (www.zenodo.org). The trajectory and topology files of the simulation are downloaded to the working directory from Zenodo but these are not saved into the database. To add a simulation to the database the user has to give some essential information about the simulation. This is done by writing a info file (*INFO.yaml) which is passed to AddData.py. 
\newline \\
AddData.py requires GROMACS and MDAnalysis library to be installed.
\newline \\
Run AddData.py as follows:
\newline \\
python3 AddData.py -f exampleINFO.yaml
\newline \\
There is also a script runAddData.sh that can be used to loop over several info files to add many simulations at one go.

%\subsubsection*{Simulation dictionary}
A simulation dictionary contains information about the simulation. Some of the information is provided by the user. The numbers of lipid molecules, solvent and ions are automatically read from the files and so are the simulation temperature and trajectory length. All this information is saved to a file named README.yaml.

\subsubsection{User input}
Dictionary variables defined by the user are listed here. 
The necessary parameters required for the analyses are described first and are marked as ''compulsory''. Then the parameters used to describe the data for further usage and analyses are described.
While only compulsory parameters are required for the databank entry, strong recommendation is that all the possible parameters would be given to maximize the possibilities to upcycle the contributed data.

\subsubsection*{DOI (compulsory)}
Give DOI identity from where the simulation files are located. Current databank works only for the data in Zenodo, but other potential sources are may be implemented in the future. Note that the DOI must point to a specific version of dataset in Zenodo, i.e., DOI pointing to all versions of certain dataset does not work.

\subsubsection*{SOFTWARE (compulsory)}
Give the name of software used to run the simulation. The options are GROMACS, AMBER, NAMD, CHARMM and OPENMM. So far, only simulations run with GROMACS are accepted by the script.

\subsubsection*{TRJ (compulsory)}
Give the name of the trajectory file that is found from the DOI given above.

\subsubsection*{TPR (compulsory)}
Give the name of the file with topology information (tpr file in the case of Gromacs) that is found from the DOI given above.

\subsubsection*{PREEQTIME (compulsory)}
Give the time simulated before the uploaded trajectory in nanoseconds. For example, if you upload 100-200~ns part of total 200~ns simulation, this should value should be 100.

\subsubsection*{TIMELEFTOUT (compulsory)}
Give the time that should be considered as an equilibration period in the uploaded trajectory. Frames before the give time will be discarded in the analysis. For example, if you upload 0-200~ns part of total 200~ns simulation where the first 100~ns should be considered as an equilibration, this value should be 100.

\subsubsection*{Molecule names (compulsory)}
In the databank, each molecule has a unique abbreviation (see table~\ref{tab:abbreviations}). 
You need to give the residue names of these molecules corresponding your simulation.
%The user must also provide the names of the molecules, ions and solvent that are used in the simulation to match the names used by the databank. The names provided by the user must be the same as in the tpr file. 
If atoms in a lipid belong to different residues (typical situation in Amber force fields), give the name of the head group residue here, and add the residue name of each atom to the third column in the mapping file (see below). If your simulation contains molecules that are not yet in the databank, you need to define the abbreviation and add molecules to the lipids\_dict, molecules\_dict, molecule\_numbers\_dict and molecule\_ff\_dict in the AddData.py script, as well as to table~\ref{tab:abbreviations}.

\begin{table}[h]
    \centering
    \begin{tabular}{c|c}
        Abbreviation & Molecule name \\
        \hline
        POPC &  1-palmitoyl-2-oleoyl-sn-glycero-3-phosphocholine\\
        POPG &  1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoglycerol \\
        POPS & 1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-L-serine \\
        POPE & 1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoethanolamine \\
        CHOL & cholesterol \\
        DHMDMAB & dihexadecyldimethylammonium \\
        \hline
        POT & potassium ion \\
        SOD & sodium ion \\
        CLA & chloride ion \\
        CAL & calcium ion \\
        SOL & water \\
    \end{tabular}
    \caption{Abbreviations used in the databank}
    \label{tab:abbreviations}
\end{table}

\subsubsection*{MAPPING\_DICT (compulsory)}
For the analysis, we need to know the names of atoms in your system.
These are defined using the mapping file convention introduced in \url{http://nmrlipids.blogspot.com/2015/03/mapping-scheme-for-lipid-atom-names-for.html}, where first column gives the universal atom name, second gives the atom name in force field, and third gives the residue name if not the same for all atoms.
The name of the mapping file for each molecule needs to be given in dictionary format. The dictionary the key is the molecule name (abbreviation listed in table \ref{tab:abbreviations}) and the value is the name of the mapping file. The already existing mapping files can be found from the directory named "mapping\_files". If a mapping file for your molecule(s) do not exist, you need to construct one and add to the "mapping\_files" directory.

%\newline \\
%The purpose of a mapping file is to circumvent the problem caused by different atom naming conventions used by different force fields. The first column of a mapping file contains general atom names. The second column contains the name of the atom as it is in the force field. If the lipid consists of several residues which is the case in some AMBER force fields, then a third column is needed which contains the name of the residue to which each atom belongs to.

\subsubsection*{DIR\_WRK (compulsory)}
Give the path of the working directory in your local computer. The trajectory and topology files will be downloaded to this trajectory, and temporary files created during processing will be stored here.

\subsubsection*{UNITEDATOM\_DICT (compulsory for united atom trajectories)}
Order parameters from united atom simulations are calculated using buildH code (\url{https://github.com/patrickfuchs/buildH}). For united atom simulations, you need to tell how hydrogens are added based on definitions in the dic\_lipids.py dictionary. This is done by giving a dictionary where the key is the molecule name (abbreviation listed in table \ref{tab:abbreviations}) and the value is the correct dictionary key in dic\_lipids.py. If correct dictionary key is not yet found, you need to add to dic\_lipids.py. 
In the case of an all atom simulation, UNITEDATOM is left empty.

\subsubsection*{PUBLICATION}
Give reference to a publication(s) related to the data.

\subsubsection*{AUTHORS\_CONTACT}
Give the name and email of the main author(s) of the data.

\subsubsection*{SYSTEM}
Give description of system in free format. For example ''POPC with cholesterol at 301K''.

\subsubsection*{SOFTWARE\_VERSION}
Give the version of the software used.

\subsubsection*{FF}
Give the name of the force field used used in the simulation.

\subsubsection*{FF\_SOURCE}
Describe the source of the force field parameters. For example, CHARMM-GUI, link to webpage where parameters were downloaded, or citation to a paper.

\subsubsection*{FF\_DATE}
Give the date when parameters were accessed or created. The format is day/month/year.

\subsubsection*{Individual force field names for molecules}
In some cases special force fields are used for certain molecules. For example, non-standard parameters for ions or other molecules have been used. These can be specified giving forcefield names separately for individual molecules. These can be given as parameters named as
FF+[abbreviation from table \ref{tab:abbreviations}], i.e., FFPOPC, FFPOT, FFSOL etc.

\subsubsection*{CPT (Gromacs)}
Give the name of the Gromacs checkpoint file that is found from the DOI given above.
CPT stands for the name of the cpt file.

\subsubsection*{LOG (Gromacs)}
Give the name of the Gromacs log file that is found from the DOI given above.

\subsubsection*{TOP (Gromacs)}
Give the name of the Gromacs top file that is found from the DOI given above.

\subsubsection{Automatically analyzed parameters}
The following parameters are read automatically from the trajectory and topology files.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%automatically analyzed parameters%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{Molecule numbers}
Numbers of lipid molecules (NPOPC, NPOPG, etc.) per membrane leaflet are calculated by determining on which side of the center of mass of the membrane the center of mass of the head group of each lipid molecule is located.
\newline \\
\noindent Numbers of other molecules such as solvent and ions (NSOL, NPOT, NSOD, etc.) are read from the topology file.

\subsubsection*{Temperature}
Temperature of the simulation is read from the topology file.

\subsubsection*{Trajectory length}
The length of a trajectory is read from the trajectory file.

\subsubsection*{Date of running}
The date of running the script is saved to the README.yaml file.

\bibliography{refs.bib}

%\noindent LaTeX formats citations and references automatically using the bibliography records in your .bib file, which you can edit via the project menu. Use the cite command for an inline citation, e.g.  \cite{Hao:gidmaps:2014}.

%For data citations of datasets uploaded to e.g. \emph{figshare}, please use the \verb|howpublished| option in the bib entry to specify the platform and the link, as in the \verb|Hao:gidmaps:2014| example in the sample bibliography file.

\section*{Acknowledgements}

%Acknowledgements should be brief, and should not include thanks to anonymous referees and editors, or effusive comments. Grant or contribution numbers may be acknowledged.

\section*{Author contributions statement}

Must include all authors, identified by initials, for example:
A.A. conceived the experiment(s),  A.A. and B.A. conducted the experiment(s), C.A. and D.A. analysed the results.  All authors reviewed the manuscript.

\section*{Additional information}

%To include, in this order: \textbf{Accession codes} (where applicable); \textbf{Competing interests} (mandatory statement).

%The corresponding author is responsible for submitting a \href{http://www.nature.com/srep/policies/index.html#competing}{competing interests statement} on behalf of all authors of the paper. This statement must be included in the submitted article file.

%\begin{figure}[ht]
%\centering
%\includegraphics[width=\linewidth]{stream}
%%\caption{Legend (350 words max). Example legend text.}
%\label{fig:stream}
%\end{figure}

%\begin{table}[ht]
%\centering
%\begin{tabular}{|l|l|l|}
%\hline
%Condition & n & p \\
%\hline
%A & 5 & 0.1 \\
%\hline
%B & 10 & 0.01 \\
%\hline
%\end{tabular}
%\caption{\label{tab:example}Legend (350 words max). Example legend text.}
%\end{table}

%Figures and tables can be referenced in LaTeX using the ref command, e.g. Figure \ref{fig:stream} and Table \ref{tab:example}.

\end{document}

### 2022-03-28 Manuscript/main.tex
Updates from Overleaf

@@
 \usepackage{graphicx}
-\title{NMRlipids Databank}
+\title{NMRlipids Databank: Overlay Databank of Lipid Membrane Simulations Arising from Open Collaboration}
 
@@
 
-Here we present a solution for lipid bilayers based on overlay databank structure illustrated in Fig.~\ref{overlay}.
+\begin{figure}
+    \centering
+    \includegraphics[width = 89mm]{Figures/overlay.pdf}
+    \caption{Structure of a overlay databank illustrated. 
+    Raw data is available from a publicly accessible repository (layer 1).
+    The core of the databank contains information on the location of the raw data and indexing of molecular compositions and simulation details described with universal naming conventions (layer 2).
+    The content of databank can be viewed and analysed using external programs that automatically read the information and data from layers 1 and 2 (layer 3).
+Examples of such applications are the GitHub repository generating the results presented in this work and the graphical access to the databank.
+    More detailed description of the databank is illustrated in Fig.~\ref{DatabankStructure} in the SI.
+    }
+    \label{fig:overlay}
+\end{figure}
+
+Here we present a solution for lipid bilayers based on overlay databank structure illustrated in Fig.~\ref{fig:overlay}.
+The key idea of the overlay databank is that the storage of raw data in layer 1 is distributed in publicly available repositories or other servers with long term stability and permanent links such as digital object identifiers.
+The core of the databank, layer 2, contains only information on the location and content of the raw data, thereby not requiring large resources to handle and maintain.
+This lowers the barrier for starting such databank as well as for long term storage.
+The NMRlipids databank is essentially a git repository containing information on the location of raw data and its indexing with universal naming convention.
+In addition to all computers where the databank is developed and used, the NMRlipids databank git is stored to Zenodo server, thereby enabling a very cost effective long term storage for the databank.
+The databank can be used in layer 3 by accessing the raw data and information stored in the databank by employing the universal naming conventions for atoms, molecules and simulation details.
+The applications can be linked to the core databank (layer 2) without actually including them, thereby enabling flexible extension of the data without compromising the simplicity and lightness of the core databank.
+The concept of overaly databank is developed here to solve the practical challenges in generating databanks of MD simulation data enabling flexible analyses over large sets of simulation data, but it pontentially used for wide range of situations, particularly when storage of raw data requires significant resources and final outcomes or best practices are not yet clear, overlay databank approach lowers the barrier to start without compromizing the long term stability or scalability.
+
+The practical relevance of the NMRlipids databank is exemplified by automatic quality evaluation and ranking of large amount of MD simulation data, data-driven analysis detecting correlations between properties of model cell membranes and analyses of rare phenomena that are beyond the scope of standard MD simulation studies.
+The NMRlipids databank provides new tools for researchers in wide range of fields in academia and industry from cell membrane biology to lipid nanoparticle formulations and data-driven computational chemistry and machine learning.
 
@@
 \subsection{Quality evaluation of force fields}
+Even though the quality of force fields has been widely evaluated against experimental data during parameterization and in separate comparison studies~\cite{botan15,ollila16,catte16,pluhackova16,perez17,leonard19}, defining unambiguously their overall quality still remains unclear, and several controversial results are reported for each in the literature~\cite{antila22b}.
 
-The quality of each simulation is evaluated against NMR order parameters.
-Indexing of experimental data to automatize evaluation is currently in progress by Anne Kiirikki.
+The quality of MD simulations in the NMRlipids databank are evaluated against C-H bond order parameters from NMR and form factors from x-ray scattering, which can be robustly measured from lipid bilayers and directly connected to the simulation data~\cite{ollila16}.
+The first one is related to the conformational ensemble of individual lipids and latter to the overall structure of lipid bilayer.
+NMRlipids databank searches available experimental data for each simulations and, when found, defines the quality of a simualtion based on comparison against C-H bond order parameters and x-ray scattering form factors.
+For more details see the method section.
 

### 2022-04-03 Manuscript/main.tex
Updates from Overleaf

@@
 For more details see the method section.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=180mm]{Figures/quality.pdf}
+    \caption{Caption}
+    \label{fig:quality}
+\end{figure}
+
+The top ranking simulations according to the NMRlipids measures are shown in Fig.~\ref{fig:quality}.
+The overall best ranking simulation, POPC lipid bilayer simulated with Slipids force field, gives a good agreement with experiments for all C-H bond order parameters except for glycerol backbone.
+The best simulation for headroup and glycerol backbone, CHARMM36 POPC, predicts too ordered acyl chain, thereby not being within the top 10 simulations in the overall ranking.
+The best ranking for x-ray form factor is given by the Slipids POPE simulation, however, because the quantitative value for form factor quality depends on the quality of experimental data, form factor qualities can be compared only between simulations evaluated against the same experimental dataset.
 

### 2022-05-02 Manuscript/main.tex
Updates from Overleaf

--- 2022-03-28 Manuscript/main.tex
@@
 \usepackage{graphicx}
+\usepackage{tablefootnote}
+
 \title{NMRlipids Databank: Overlay Databank of Lipid Membrane Simulations Arising from Open Collaboration}
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width = 89mm]{Figures/overlay.pdf}
-    \caption{Structure of a overlay databank illustrated. 
-    Raw data is available from a publicly accessible repository (layer 1).
-    The core of the databank contains information on the location of the raw data and indexing of molecular compositions and simulation details described with universal naming conventions (layer 2).
-    The content of databank can be viewed and analysed using external programs that automatically read the information and data from layers 1 and 2 (layer 3).
-Examples of such applications are the GitHub repository generating the results presented in this work and the graphical access to the databank.
-    More detailed description of the databank is illustrated in Fig.~\ref{DatabankStructure} in the SI.
-    }
-    \label{fig:overlay}
-\end{figure}
-
 Here we present a solution for lipid bilayers based on overlay databank structure illustrated in Fig.~\ref{fig:overlay}.
-The key idea of the overlay databank is that the storage of raw data in layer 1 is distributed in publicly available repositories or other servers with long term stability and permanent links such as digital object identifiers.
-The core of the databank, layer 2, contains only information on the location and content of the raw data, thereby not requiring large resources to handle and maintain.
-This lowers the barrier for starting such databank as well as for long term storage.
-The NMRlipids databank is essentially a git repository containing information on the location of raw data and its indexing with universal naming convention.
-In addition to all computers where the databank is developed and used, the NMRlipids databank git is stored to Zenodo server, thereby enabling a very cost effective long term storage for the databank.
-The databank can be used in layer 3 by accessing the raw data and information stored in the databank by employing the universal naming conventions for atoms, molecules and simulation details.
-The applications can be linked to the core databank (layer 2) without actually including them, thereby enabling flexible extension of the data without compromising the simplicity and lightness of the core databank.
 The concept of overaly databank is developed here to solve the practical challenges in generating databanks of MD simulation data enabling flexible analyses over large sets of simulation data, but it pontentially used for wide range of situations, particularly when storage of raw data requires significant resources and final outcomes or best practices are not yet clear, overlay databank approach lowers the barrier to start without compromizing the long term stability or scalability.
@@
 %Up to three levels of \textbf{subheading} are permitted. Subheadings should not be numbered.
+\subsection{Design of the NMRlipids databank}
+The key idea of the overlay databank is that the storage of raw data in layer 1 is distributed in publicly available repositories or other servers with long term stability and permanent links such as digital object identifiers.
+The core of the databank, layer 2, contains only information on the location and content of the raw data, thereby not requiring large resources to handle and maintain.
+This lowers the barrier for starting such databank as well as for long term storage.
+The NMRlipids databank is essentially a git repository containing information on the location of raw data and its indexing with universal naming convention.
+In addition to all computers where the databank is developed and used, the NMRlipids databank git is stored to Zenodo server, thereby enabling a very cost effective long term storage for the databank.
+The databank can be used in layer 3 by accessing the raw data and information stored in the databank by employing the universal naming conventions for atoms, molecules and simulation details.
+The applications can be linked to the core databank (layer 2) without actually including them, thereby enabling flexible extension of the data without compromising the simplicity and lightness of the core databank.
+
+\begin{figure}
+    \centering
+    \includegraphics[width = 180mm]{Figures/overlay.pdf}
+    \caption{a) Structure of a overlay databank. 
+    %Raw data is available from a publicly accessible repository (layer 1).
+    %The core of the databank contains information on the location of the raw data and indexing of molecular %compositions and simulation details described with universal naming conventions (layer 2).
+    %The content of databank can be viewed and analysed using external programs that automatically read the information and data from layers 1 and 2 (layer 3).
+Examples of such applications are the GitHub repository generating the results presented in this work and the graphical access to the databank.
+    More detailed structure of the layer 2 is illustrated in Fig.~\ref{DatabankStructure} in the SI.
+    b) Distribution of lipids that are present in more than two trajectories in the databank.
+    Lipids occuring in 1-2 trajectories are: DCHOL(2), DDOPC(2), DOPE(2), DEPC(1), DMPC(1), SAPI(1), POPI(1), SLPI(1), DLIPC(2), DSPC(2), DPPG(2), DPPE(1), and DLPC(1).
+    c) Distribution of force fields used in the simulations.
+    d) Distribution of the lengths of the trajectories, total number of trajectories and total lenght of the simulations.
+    e) Currently available binary mixtures in the databank.}
+    \label{fig:overlay}
+\end{figure}
+
+Currently the databank is composed of approximately 500 trajectories with the total length of approximately 231 microseconds of which most are contributed for the previous publications from the NMRlipids open collaboration~\cite{botan15,catte16,antila19,bacle21}.
+The distribution of lipids, force fields, length of the trjajectories and available binary mixtures are shown in Fig.~\ref{fig:overlay}.
 
 \subsection{Quality evaluation of force fields}
-Even though the quality of force fields has been widely evaluated against experimental data during parameterization and in separate comparison studies~\cite{botan15,ollila16,catte16,pluhackova16,perez17,leonard19}, defining unambiguously their overall quality still remains unclear, and several controversial results are reported for each in the literature~\cite{antila22b}.
+Quality of membrane simulations with different force fields have been evaluated against experimental data during parameterization and in separate comparison studies~\cite{botan15,ollila16,catte16,pluhackova16,perez17,leonard19}, but universal quality measure for membrane simulations is not defined and controversial results are often reported from simulations~\cite{antila22b}.
+The unclarities in simulation quality complicate the selection of proper simulation models for specific applications and estimation of reliability of reported simulation results, thereby being a major obstacle in many applications of membrane MD simulations.
 
-The quality of MD simulations in the NMRlipids databank are evaluated against C-H bond order parameters from NMR and form factors from x-ray scattering, which can be robustly measured from lipid bilayers and directly connected to the simulation data~\cite{ollila16}.
-The first one is related to the conformational ensemble of individual lipids and latter to the overall structure of lipid bilayer.
-NMRlipids databank searches available experimental data for each simulations and, when found, defines the quality of a simualtion based on comparison against C-H bond order parameters and x-ray scattering form factors.
-For more details see the method section.
+To enable rapid evaluation and comparison of membrane MD simulation qualities, we define the quality measure for membrane MD simulations using the C-H bond order parameters from NMR experiments and form factors from x-ray scattering, which are robust experimental measurables that can be directly connected to the simulation data~\cite{ollila16}.
+C-H bond order parameters are related to the conformational ensemble of individual lipid molecules, while the form factor to arises from electron density thereby connecting to the overall structure of lipid bilayers.
+This is demonstrated in Fig.~\ref{fig:quality} a) showing correlations between membrane lateral density (area per lipid), thickness, minima of form factor and acyl chain order analyzed from all apporimately 500 trajectories in the NMRlipids databank.
+Membrane area per lipid has negative correlations with membrane order and thickness with Pearson coefficient of -0.78 and -0.49, respectively, because decreased area leads to more ordered, tightly packed and stretched acyl chains leading to thicker membranes.
+On the other hand, locations of form factor minima have positive correlation with the membrane area and negative correlation with the thickness.
+In conclusion, both x-ray scattering form factors and acyl chain order parameters from NMR are on average good proxies for membrane packing and thickness, while NMR order parameters give segmental resolution information on the quality of conformational ensembles of lipids.
 
-\subsection{Water permeability across membranes}
+\begin{figure}[tbp]
+    \centering
+    \includegraphics[width=180mm]{Figures/quality.pdf}
+    \caption{a) Correlations between membrane area per lipid, thickness, second minima of form factor and average order parameter of the {\it sn}-1 acyl chain extracted from the NMRlipids databank.
+All Pearson correlation coefficients shown in bottom left corner have p-value below 0.001.
+    b) Evaluation against experimental data exemplified for the simulation giving the best qualities in overall, for headgroup and x-ray scattering form factor.
+    c) Change in area per lipid upon addition of POPE to POPC in different simulation extracted from the NMRlipids databank and quality evaluation table of POPE simulations.
+    d) Quality evaluation table showing the best 13 simulations according to the overall order parameter quality.}
+    \label{fig:quality}
+\end{figure}
 
-Averaging water density profiles over all simulations in the databank enables us to calculate the barrier for the water penetration through lipid bilayers.
+NMRlipids databank contains also experimental order parameter and x-ray form factor data that is connected to corresponding  simulations in order to define the simulation qualities and rank them to select the suitable simulation models for particular applications.
+The ranking of all simulations based on estimated average probability of order parameters to locate within experimental errors are shown in Fig.~\ref{fig:quality} b) and the comparison for the best models for overall quality, headgroup and glycerol backbone, and form factor are exemplified in Fig.~\ref{fig:quality} d).
+Full details of the NMRlipids quality measure for lipid bilayer simulations are given in the methods section and supplementary information.
+The top ranking simulations in Figs.~\ref{fig:quality} b) and d) demonstrate also the current complexity in lipid bilayer simulation quality.
+Simulation of POPC lipid bilayer with Slipids force field ranks first in overall score, glycerol backbone is not correcetly captured in this model.
+On the othere, simulation with the best quality for headroup and glycerol backbone, CHARMM36 POPC, predicts too ordered acyl chain, thereby not being within the top 10 simulations in the overall ranking.
+The best ranking for x-ray form factor is given by the Slipids POPE simulation, however, because the quantitative value for form factor quality depends on the quality of experimental data, form factor qualities can be compared only between simulations evaluated against the same experimental dataset.
 
-\subsection{Spin relaxation rates of confined water close to membranes}
+The power of NMRlipids quality metrics in selecting the best model for particular application is demontrated in Fig.~\ref{fig:quality} c) where the area per lipids of POPC:POPE mixtures from different force fields are shown and quality of POPE simulation in different force fields are evaluated.
+Among the simulations available in the NMRlipids databank, the Slipids force field gives the best quality in terms of acyl chain order parameters and form factor, and predicts the largest area per molecule for POPE and smallest difference with POPC.
+In conclusion, simulations with Slipids is the most reliable force field for membrane packing in POPC:POPE mixtures althought its glycerol backbone is not accurately modelled.
+Similar comparisons utilizing the preliminary data from the NMRlipids databank have concluded that Slipids is relatively good model also for mixtures of charged POPC:POPS membranes, although models with better counterion binding have higher quality, but CHARMM36 is the best to study differences in headgroup conformational ensembles between different lipid types.
+Understanding such complex picture of lipid bilayer MD simulation quality would not be possible without automatic quality evaluation of large sets of simulations enabled by the NMRlipids databank.
 
-Usefulness of the databank beyond MD simulation experts is demonstrated by analysing water spin relaxation times close to bilayers which are used in MRI imaging.
+\subsection{NMRlipids databank reveals correlations between membrane properties}
+The power of NMRlipids databank to find correlation between membrane properties is exemplified in Fig.~\ref{??} showing how water diffusion through and along the membrane correlate with different properties of systems.
+While NMRlipids databank enables similar analyses for any membrane properties, these are selected due to their relevance in the development of MRI imaging methods~\cite{??}.
+
+%\subsection{Spin relaxation rates of confined water close to membranes}
+
+%Usefulness of the databank beyond MD simulation experts is demonstrated by analysing water spin relaxation times close to bilayers which are used in MRI imaging.
 
@@
 
+Quality measure and automatic quality evaluation of lipid bilayer MD simulations introduced in the NMRlipids databank enables rapid ranking of available simulation models agains experimental NMR and x-ray scattering data.
+This provides a tool for researches to rapidly evaluate the credibility of MD simulations of their own and reported by other groups.
+Because such tool and quality measure has not been available, this will elaborate the quality of published MD simulations of lipid bilyers and reduce potentially misleading results~\cite{antila22b}.
+The power of NMRlipids databank to select the best models for particular applications has been demontrated for PC/PE lipid mixtures (Fig.~\ref{fig:quality}), PC/PS lipid mixtures~\cite{antila22b}, and lipid headgroups~\cite{bacle21}.
+
+The increasing amount of MD simulation data with programmatic access in the NMRlipids databank opens up possibilities for wide range of applications utilizing the large set of accessible data.
+Extend of the data in the NMRlipids databank in terms of quantity (e.g., simulation length and number of conformations), content (e.g, lipid compositions and ion concentrations) and quality enables analyses that are not possible to conduct from MD simulation data produced by a single research group.
+Applications of the NMRlipids databank to understand how diffusion of water through and along cellular membrane depends on its physical properties are demonstrated in Fig.~\ref{??}.
+Permeation of water through membranes resembles the permeation of also other hydrophilic molecules, such as drugs, and detailed understanding of water dynamics through and along membranes is potentially useful for the development of MRI imaging methods~\cite{topgaard20}.
+These examples demonstrate the practical applications of NMRlipids databank on problems in the biological and biomedical sciences.
+
+Building accessible databanks of molecular dynamics simulation data has been challenging due to the required long term support for hardware and software maintenance.
+In the overlay model used in the NMRlipids databank, the demand of hardware can be distributed and open collaboration model reduces the risk for ending software maintenance.
+Furthermore, the open collaboration model used in the NMRlipids project credits contributors by offering authorship in published articles, thereby creating an incentive for contributions.
+This model could be extended also to other fields where similar barriers to establish publicly accessible databanks exist.
+Emerging applications of machine learning are increasing the impact of such databanks.
+For example, the existing Protein Databank (PDB)~\cite{??} containing experimentally determined protein structures with programmatic access has enabled the development of machine learning based tools building on the data collected in the databank over the years~\cite{??}.
+NMRlipids and other databanks with open programmatic access have potential to lead similar unforeseen applications in the future.
+
 \section{Methods}
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{figures/DataBankStructure.pdf}
-  \caption{Structure of the NMRlipids databank}\label{DatabankStructure}
+  \includegraphics[width=\textwidth]{Figures/DataBankChart.pdf}
+  \caption{Structure of the NMRlipids databank.
+User give information (blue boxes) includes basic information on the simulation, including the location of raw data, as well as experimental data if available.
+The databank entries (red box) and analysis results (green boxes) then automatically generated by the computer programs included in the NMRlipids databank (yellow boxes).}\label{DatabankStructure}
 \end{figure}
 
-NMRlipids databank is a overlay databank composed of files that contain all the essential information
-on the molecular dynamics simulation trajectories that are listed in the databank.
-The location publicly avaiable raw MD simulation data is flexible.
-The NMRlipids databank files contain all the essential information on the
-original data that enable wide range of applications.
-%and all essential information on simulations needed in further use and automatic analysis.
+Structure of the NMRlipids databank is illustrated in Fig.~\ref{DatabankStructure}.
+The core of the databank is composed of README.yaml files locating at \url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations} which contain all the essential information on each simulation entry needed to further applications.
+As the NMRlipids databank is a overlay databank, the raw simulation data is not stored there but the README.yaml files contain a permanent links to the raw data which can be then accessed when needed.
+The raw MD simulation data can locate in any stable publicly available repository, although all the current data locates in Zenodo \url{www.zenodo.org}.
 
-Each simulation system is identified using the hash of original trajectory and topology file. The index files are stored in folder structure based on the identity codes of the simulations.
+A README.yaml file is made using AddData.py script for each simulation stored in the databank based on the information given by the author of the simulation data.
+In addition to this information, automatically extracted information about the simulation are stored in the README.yaml files (see table~\ref{tab:READMEkeys}).
+A README.yaml file for each simulation is stored to a folder named according to the hashes of original trajectory and topology file locating in \url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations}.
+These README.yaml files give access to all the essential information on the simulations in the databank either directly, or pointing to a original file from where the information can be extracted.
+
+\begin{table}[]
+    \centering
+    \begin{tabular}{  p{3.5cm}  p{9.5cm}  p{4.0cm} }
+    \toprule
+    key & description & type  \\
+    \midrule
+        %\hline
+    DOI & DOI from where the raw data is found & user given (compulsory) \\
+    SOFTWARE & Software used to run the simulation (e.g. Gromacs, Amber, NAMD, etc.) & \\
+    TRJ & Name of the trajectory file found from DOI & \\
+    TPR & Name of the topology file found from DOI (trp file in the case of Gromacs) & \\
+    PREEQTIME & Pre-equilibrate time simulated before the uploaded trajectory in nanoseconds.
+\tablefootnote{For example, if you upload 100-200 ns part of total 200 ns simulation, this should value should be 100.} & \\
+    TIMELEFTOUT & Equilibration period in the uploaded trajectory that should be discarded in analyses.
+\tablefootnote{For example, if you upload 0-200 ns part of total 200 ns simulation where the first 100 ns should be considered as an equilibration, this value should be 100.} \\
+    COMPOSITION & Molecules names used in the simulation and corresponding mapping files (see ??) & \\
+    DIR\_WRK & Temporary working directory in your local computer. \\
+    UNITEDATOM\_DICT & Information for constucting hydrogens for united atom simulations, empty for all atom simulations & \\
+    TYPEOFSYSTEM & Lipid bilayer or something else & \\
+    \hline
+    PUBLICATION & Give reference to a publication(s) related to the data. \\
+    AUTHORS\_CONTACT & Name and email of the main author(s) of the data. & \\
+    SYSTEM & System description on free text format & \\
+    SOFTWARE\_VERSION & Version of the used software & \\
+    FF & Name of the used force field & \\
+    FF\_SOURCE & Source of the force field parameters, e.g, CHARMM-GUI, webpage, citation to a publication, etc. & \\
+    FF\_DATE &  Date when force field parameters were accessed on the gives source (day/month/year). & \\
+    FF{molename} & Molecule specific force field information, e.g., water model with FFSOL and sodium parameters with FFSOD. & \\
+    CPT & Name of the Gromacs checkpoint file. & \\
+    LOG & Name of the Gromacs log file. & \\
+    TOP & Name of the Gromacs top file. & \\
+    GRO & Name of the Gromacs gro file. & \\
+    \hline
+    TRAJECTORY\_SIZE & Size of the trajectory file in bytes & automatically extracted data. \\
+    TRJLENGTH & Lenght of the trajectory (ps). & \\
+    TEMPERATURE & Temperature of the simulation. & \\
+    NUMBER\_OF\_ATOMS & Number of atoms in the simulation. & \\
+    DATEOFRUNNIG & Date when added into the databank & \\
+    EXPERIMENT & Potentially connected experimental data & \\
+    COMPOSTION & Number of molecules (number of lipids in both leaflets separartely) extracted from the simulation data & \\ 
+    \end{tabular}
+    \caption{Keys stored in the README.yaml files of simulations.}
+    \label{tab:READMEkeys}
+\end{table}
+
+Each simulation is connected to the experimental C-H bond order parameter and x-ray scattering form factor data using the searchDATABANK.py script.
+Available experimental data is collected to \url{https://github.com/NMRLipids/Databank/tree/main/Data/experiments} where each dataset is saved into a folder with the name based on DOI of the related publication and experimental conditions are described in README.yaml files with the keys listed in table~\ref{tab:READMEkeysEXP}.
+If molar concentrations of all molecules are within $\pm$5 percentage units, charged lipids have the same counterions, and temperature is within $\pm$2 degrees, a pair will be formed between simulation and experimental datasets by adding the link to the experimental data to the simulation README.yaml file.
+%If lipid to water ratio is above ??, the systems are considered fully hydrated. Otherwise also hydration level is considered.
+
+\begin{table}[]
+    \centering
+    \begin{tabular}{  p{5.0cm}  p{10.0cm}}
+    \toprule
+    key & description \\
+    \midrule
+    DOI & DOI of the publication related to the experimental data. \\
+    TEMPERATURE & Temperature of the experiment. \\
+    MOLAR\_FRACTIONS & Dictionary of molar fractions of bilayer components \\
+    ION\_CONCENTRATIONS & Dictionary of ion concentrations of the system (defined as ??) \\
+    TOTAL\_LIPID\_CONCENTRATION & Total concentration of lipid components. If exact concentration is not known, but experiments are performed in excess water, 'full hydration' can be given. \\
+    COUNTER\_IONS & Type of counter ions if present.
+\end{tabular}
+    \caption{Keys stored in the README.yaml files of experiments.}
+    \label{tab:READMEkeysEXP}
+\end{table}
 
 \subsection{Quality evaluation}
-
-For quality evaluation, each simulation is connected to a experimental dataset is similar conditions.
-Currently experimental and simulation datasets are paired if molar concentrations of all molecules are within $\pm$5 percentage units, charged lipids have the same counterions, and temperature is within $\pm$2 degrees.
-If lipid to water ratio is above ??, the systems are considered fully hydrated.
-Otherwise also hydration level is considered.
 

### 2022-05-09 Manuscript/main.tex
Updates from Overleaf

@@
     \includegraphics[width = 180mm]{Figures/overlay.pdf}
-    \caption{a) Structure of a overlay databank. 
+    \caption{a) Structure of an overlay databank. 
     %Raw data is available from a publicly accessible repository (layer 1).
@@
 Examples of such applications are the GitHub repository generating the results presented in this work and the graphical access to the databank.
-    More detailed structure of the layer 2 is illustrated in Fig.~\ref{DatabankStructure} in the SI.
-    b) Distribution of lipids that are present in more than two trajectories in the databank.
-    Lipids occuring in 1-2 trajectories are: DCHOL(2), DDOPC(2), DOPE(2), DEPC(1), DMPC(1), SAPI(1), POPI(1), SLPI(1), DLIPC(2), DSPC(2), DPPG(2), DPPE(1), and DLPC(1).
-    c) Distribution of force fields used in the simulations.
-    d) Distribution of the lengths of the trajectories, total number of trajectories and total lenght of the simulations.
-    e) Currently available binary mixtures in the databank.}
+    More detailed structure of the layer 2 in the NMRlipids databank is illustrated in Fig.~\ref{DatabankStructure} in the SI.
+    b) Distribution of the lengths of the trajectories, total number of trajectories and total lenght of the simulations in the NMRlipids databank.
+    c) Distribution of lipids present in the trajectories in the NMRlipids databank. Lipids occuring in five or less simulations ('others') are listed in the right. 
+    d) Currently available binary mixtures in the NMRlipids databank. 
+    e) Distribution of force fields in the simulations in the NMRlipids databank.
+    The figures and numbers are created on 9th of May 2022.}
     \label{fig:overlay}
@@
   \caption{Structure of the NMRlipids databank.
-User give information (blue boxes) includes basic information on the simulation, including the location of raw data, as well as experimental data if available.
-The databank entries (red box) and analysis results (green boxes) then automatically generated by the computer programs included in the NMRlipids databank (yellow boxes).}\label{DatabankStructure}
+Manually added input data (blue boxes) includes basic information on the simulation, permanent links to the raw data, and experimental data if available.
+The databank entries (red box) and analysis results (green boxes), locating at \url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations} are automatically generated by the computer programs included in the NMRlipids databank (yellow boxes).
+Because raw data are not permanently stored but can be accessed based on the information in the databank, this connection is marked with the dashed line.
+}\label{DatabankStructure}
 \end{figure}
@@
 Structure of the NMRlipids databank is illustrated in Fig.~\ref{DatabankStructure}.
-The core of the databank is composed of README.yaml files locating at \url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations} which contain all the essential information on each simulation entry needed to further applications.
-As the NMRlipids databank is a overlay databank, the raw simulation data is not stored there but the README.yaml files contain a permanent links to the raw data which can be then accessed when needed.
+The required input information to create an entry into the NMRlipids databank are listed in table~\ref{tab:READMEkeys}.
+While the raw simulation data is not directly stored in the NMRlipids databank, permanent links from where the raw data can be accessed have to be given and are then stored in the README.yaml files at \url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations}.
+These files contain all the essential information listed in table~\ref{tab:READMEkeys} on each simulation entry that are needed for further applications.
 The raw MD simulation data can locate in any stable publicly available repository, although all the current data locates in Zenodo \url{www.zenodo.org}.
 
-A README.yaml file is made using AddData.py script for each simulation stored in the databank based on the information given by the author of the simulation data.
+%A README.yaml file is made using AddData.py script for each simulation stored in the databank based on the information given by the author of the simulation data.
 In addition to this information, automatically extracted information about the simulation are stored in the README.yaml files (see table~\ref{tab:READMEkeys}).
@@
     \hline
-    PUBLICATION & Give reference to a publication(s) related to the data. \\
+    PUBLICATION & Give reference to a publication(s) related to the data. & User given (optional)\\
     AUTHORS\_CONTACT & Name and email of the main author(s) of the data. & \\
@@
     EXPERIMENT & Potentially connected experimental data & \\
-    COMPOSTION & Number of molecules (number of lipids in both leaflets separartely) extracted from the simulation data & \\ 
+    COMPOSITION & Numbers of lipid molecules (NPOPC, NPOPG, etc.) per membrane leaflet are calculated by determining on which side of the center of mass of the membrane the center of mass of the head group of each lipid molecule is located.
+    Numbers of other molecules such as solvent and ions (NSOL, NPOT, NSOD, etc.) are read from the topology file. & \\
     \end{tabular}
@@
 
-Each simulation is connected to the experimental C-H bond order parameter and x-ray scattering form factor data using the searchDATABANK.py script.
-Available experimental data is collected to \url{https://github.com/NMRLipids/Databank/tree/main/Data/experiments} where each dataset is saved into a folder with the name based on DOI of the related publication and experimental conditions are described in README.yaml files with the keys listed in table~\ref{tab:READMEkeysEXP}.
-If molar concentrations of all molecules are within $\pm$5 percentage units, charged lipids have the same counterions, and temperature is within $\pm$2 degrees, a pair will be formed between simulation and experimental datasets by adding the link to the experimental data to the simulation README.yaml file.
+In order to evaluate the quality of simulations, sets of C-H bond order parameters from NMR and from factors from x-ray scattering are included in the NMRlipids databank (\url{https://github.com/NMRLipids/Databank/tree/main/Data/experiments}).
+The required information for an experimental dataset are listed in table~\ref{tab:READMEkeysEXP}.
+A simulation is connected to a experimental data set
+%C-H bond order parameter and x-ray scattering form factor data using the searchDATABANK.py script.
+Available experimental data is collected to  where each dataset is saved into a folder with the name based on DOI of the related publication and experimental conditions are described in README.yaml files with the keys listed in table~\ref{tab:READMEkeysEXP}.
+if molar concentrations of all molecules are within $\pm$5 percentage units, charged lipids have the same counterions, and temperature is within $\pm$2 degrees.
+In such cases, a simulation and experimental data are paired by adding the path to the experimental data into the simulation README.yaml file.
 %If lipid to water ratio is above ??, the systems are considered fully hydrated. Otherwise also hydration level is considered.
@@
 
+Because README.yaml contains all the essential information on each simulation, arbitrary analyses can be automatically performed over all the simualtions in the NMRlipids databank.
+In addition to the order parameters for each C-bonds and x-ray scattering form factors used in the quality evaluation, the NMRlipids databank contains the area per lipid and thickness calculated from all simulations in the databank.
+These results are stored in the same folders as the README.yaml files in \url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations}.
+
+\subsection{Molecule and atom naming convention}
+
+Because universal convention for lipid molecules and atoms therein has not been defined, the naming conventions vary between authors and force fields.
+To enable automatic analyses over large sets of simulation in the NMRlipids databank, we have defined unique naming conventions for lipid molecules and atoms.
+The abbreviations of molecule names used in the NMRlipids databank are listed in table~\ref{tab:abbreviations}.
+The unique atom names for each molecule and corresponding names in each simulation are defined using mapping files introduced in the NMRlipids project (\url{https://nmrlipids.blogspot.com/2022/04/new-yaml-format-of-mapping-files.html}).
+
+Molecule and atom names in each simulation are connected to the unique naming convention with the COMPOSITION dictionary in the README.yaml file.
+Upon addition of a new entry in the databank, the molecule names in the simulation corresponding the unique names and mapping files (available at \url{https://github.com/NMRLipids/Databank/tree/main/Scripts/BuildDatabank/mapping_files}) are defined in the dictionary.
+The numbers of molecules in the system are then automatically calculated by the NMRlipids databank codes (AddData.py in Fig.~\ref{DatabankStructure}) and stored in the README.yaml together with other content of the COMPOSITION dictionary.
+This information can be then used to find each molecule and atom from each simulation in the analysis codes.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{c|c}
+        Abbreviation & Molecule name \\
+        \hline
+        POPC &  1-palmitoyl-2-oleoyl-sn-glycero-3-phosphocholine\\
+        POPG &  1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoglycerol \\
+        POPS & 1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-L-serine \\
+        POPE & 1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoethanolamine \\
+        CHOL & cholesterol \\
+        DHMDMAB & dihexadecyldimethylammonium \\
+        \hline
+        POT & potassium ion \\
+        SOD & sodium ion \\
+        CLA & chloride ion \\
+        CAL & calcium ion \\
+        SOL & water \\
+    \end{tabular}
+    \caption{Abbreviations used in the databank}
+    \label{tab:abbreviations}
+\end{table}
+
 \subsection{Quality evaluation}
 
-Quality of lipid conformational ensembles are evaluated against C-H bond order parameters from NMR experiments~\cite{ollila16}.
-As a first step, the quality of each order parameter is evaluated using the negative logarithm of probability to find the simulated order parameter within the experimental error bars
+Quality of conformational ensembles of lipid molecules in simulations are evaluated by estimating the average probability of C-H bond order parameters to agree with the values from NMR experiments. 
+%As a first step, the quality of each order parameter is evaluated using the negative logarithm of probability to find the simulated order parameter within the experimental error bars
+%\begin{equation}
+%    S_q = -\log_{10}(P),    
+%\end{equation}
+Because each C-H bond order parameter is calculated as an average over individual lipids in a simulation and conformational ensembles of lipids are assumed to be independent, the random variable
 \begin{equation}
-    S_q = -\log_{10}(P),    
+    \frac{\bar{X}-\mu}{S/\sqrt{n}},
 \end{equation}
+where $\bar{X}$ is the sample mean, $S$ is the sample variance and $n$ is the sample size (number of lipids), has a Student's t-distribution with $n-1$ degrees of freedom and the mean of $\mu$.
+
 where the probability is calculated from the normal distribution
@@
 
-\subsection{Analysis of the data in databank}
+\subsection{Analysing simulations in the NMRlipids databank}
 
-Index files in the datanbank contain all the essential information to perform analyses from the simulations.
-Systems under interest can be selected automatically browsing the index files and filtering the desired properties.
+The README.yaml files contain all the essential information to perform arbitrary analyses of simulations in the NMRlipids databank, i.e., the permanent location of the original data and naming convention for all atoms and molecules in each system.
+In practise, the analyse codes contains a loop over all README.yaml files (i.e., simulations in the NMRlipids databank) which first downloads the raw simulation to a local computer and then uses the information about the atom and molecule naming conventions in README.yaml and mapping files to perform the desired analyses.
+For example, the code that calculates all C-H bond order parameters of all systems is available at \url{https://github.com/NMRLipids/Databank/blob/main/Scripts/AnalyzeDatabank/calcOrderParameters.py} and minimal example of a analysis code is available at \url{https://github.com/NMRLipids/Databank/blob/main/Scripts/AnalyzeDatabank/template.ipynb}.
 
-Analysis results can be most conveniently stored to a new datanbank with identical indexing as the original databank.
-The results can be then easily shared and browsed indentically to the original databank.
+While the order parameters, form factors, area per lipid and thickness are stored within the NMRlipids databank (\url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations}), further analyses can be conventiently stored in separate repositories with the same folder structure based on hash identities of trajectory and topology files.
+For example, results from further analyses performed here are stored in folders at \url{https://github.com/NMRLipids/DataBankManuscript/tree/main/Data}.
+Such organization of the data enables further upcycling of the analyzed data as similarly to the original NMRlipids databank repository.
 
-\subsection{Indexing the simulation data}
+%\subsection{Indexing the simulation data}
 
 %\subsubsection*{AddData.py}
-AddData.py is a script that builds a database that contains a dictionary file and analysis data of each simulation.
+%AddData.py is a script that builds a database that contains a dictionary file and analysis data of each simulation.
 The dictionary file contains information about the simulation.
@@
 This is done by writing a info file (*INFO.yaml) which is passed to AddData.py.
-\newline \\
-AddData.py requires GROMACS and MDAnalysis library to be installed.
-\newline \\
-Run AddData.py as follows:
-\newline \\
-python3 AddData.py -f exampleINFO.yaml
-\newline \\
-There is also a script runAddData.sh that can be used to loop over several info files to add many simulations at one go.
+%\newline \\
+%AddData.py requires GROMACS and MDAnalysis library to be installed.
+%\newline \\
+%Run AddData.py as follows:
+%\newline \\
+%python3 AddData.py -f exampleINFO.yaml
+%\newline \\
+%There is also a script runAddData.sh that can be used to loop over several info files to add many simulations at one go.
 
 %\subsubsection*{Simulation dictionary}
-A simulation dictionary contains information about the simulation.
+%A simulation dictionary contains information about the simulation.
 Some of the information is provided by the user.
@@
 
-\subsubsection{User input}
-Dictionary variables defined by the user are listed here. 
-The necessary parameters required for the analyses are described first and are marked as ''compulsory''. Then the parameters used to describe the data for further usage and analyses are described.
-While only compulsory parameters are required for the databank entry, strong recommendation is that all the possible parameters would be given to maximize the possibilities to upcycle the contributed data.
+%\subsubsection{User input}
+%Dictionary variables defined by the user are listed here. 
+%The necessary parameters required for the analyses are described first and are marked as ''compulsory''. Then the parameters used to describe the data for further usage and analyses are described.
+%While only compulsory parameters are required for the databank entry, strong recommendation is that all the possible parameters would be given to maximize the possibilities to upcycle the contributed data.
 
-\subsubsection*{DOI (compulsory)}
-Give DOI identity from where the simulation files are located.
+%\subsubsection*{DOI (compulsory)}
+%Give DOI identity from where the simulation files are located.
 Current databank works only for the data in Zenodo, but other potential sources are may be implemented in the future.
@@
 
-\subsubsection*{SOFTWARE (compulsory)}
-Give the name of software used to run the simulation. The options are GROMACS, AMBER, NAMD, CHARMM and OPENMM. So far, only simulations run with GROMACS are accepted by the script.
+%\subsubsection*{SOFTWARE (compulsory)}
+%Give the name of software used to run the simulation. The options are GROMACS, AMBER, NAMD, CHARMM and OPENMM. So far, only simulations run with GROMACS are accepted by the script.
 
-\subsubsection*{TRJ (compulsory)}
-Give the name of the trajectory file that is found from the DOI given above.
+%\subsubsection*{TRJ (compulsory)}
+%Give the name of the trajectory file that is found from the DOI given above.
 
-\subsubsection*{TPR (compulsory)}
-Give the name of the file with topology information (tpr file in the case of Gromacs) that is found from the DOI given above.
+%\subsubsection*{TPR (compulsory)}
+%Give the name of the file with topology information (tpr file in the case of Gromacs) that is found from the DOI given above.
 
-\subsubsection*{PREEQTIME (compulsory)}
-Give the time simulated before the uploaded trajectory in nanoseconds. For example, if you upload 100-200~ns part of total 200~ns simulation, this should value should be 100.
+%\subsubsection*{PREEQTIME (compulsory)}
+%Give the time simulated before the uploaded trajectory in nanoseconds. For example, if you upload 100-200~ns part of total 200~ns simulation, this should value should be 100.
 
-\subsubsection*{TIMELEFTOUT (compulsory)}
-Give the time that should be considered as an equilibration period in the uploaded trajectory.
+%\subsubsection*{TIMELEFTOUT (compulsory)}
+%Give the time that should be considered as an equilibration period in the uploaded trajectory.
 Frames before the give time will be discarded in the analysis.
@@
 
-\subsubsection*{Molecule names (compulsory)}
-In the databank, each molecule has a unique abbreviation (see table~\ref{tab:abbreviations}). 
-You need to give the residue names of these molecules corresponding your simulation.
+%\subsubsection*{Molecule names (compulsory)}
+%In the databank, each molecule has a unique abbreviation (see table~\ref{tab:abbreviations}). 
+%You need to give the residue names of these molecules corresponding your simulation.
 %The user must also provide the names of the molecules, ions and solvent that are used in the simulation to match the names used by the databank.
 The names provided by the user must be the same as in the tpr file.
-If atoms in a lipid belong to different residues (typical situation in Amber force fields), give the name of the head group residue here, and add the residue name of each atom to the third column in the mapping file (see below).
+%If atoms in a lipid belong to different residues (typical situation in Amber force fields), give the name of the head group residue here, and add the residue name of each atom to the third column in the mapping file (see below).
 If your simulation contains molecules that are not yet in the databank, you need to define the abbreviation and add molecules to the lipids\_dict, molecules\_dict, molecule\_numbers\_dict and molecule\_ff\_dict in the AddData.py script, as well as to table~\ref{tab:abbreviations}.
 
-\begin{table}[h]
-    \centering
-    \begin{tabular}{c|c}
-        Abbreviation & Molecule name \\
-        \hline
-        POPC &  1-palmitoyl-2-oleoyl-sn-glycero-3-phosphocholine\\
-        POPG &  1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoglycerol \\
-        POPS & 1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-L-serine \\
-        POPE & 1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoethanolamine \\
-        CHOL & cholesterol \\
-        DHMDMAB & dihexadecyldimethylammonium \\
-        \hline
-        POT & potassium ion \\
-        SOD & sodium ion \\
-        CLA & chloride ion \\
-        CAL & calcium ion \\
-        SOL & water \\
-    \end{tabular}
-    \caption{Abbreviations used in the databank}
-    \label{tab:abbreviations}
-\end{table}
-
-\subsubsection*{MAPPING\_DICT (compulsory)}
-For the analysis, we need to know the names of atoms in your system.
-These are defined using the mapping file convention introduced in \url{http://nmrlipids.blogspot.com/2015/03/mapping-scheme-for-lipid-atom-names-for.html}, where first column gives the universal atom name, second gives the atom name in force field, and third gives the residue name if not the same for all atoms.
-The name of the mapping file for each molecule needs to be given in dictionary format.
+%\subsubsection*{MAPPING\_DICT (compulsory)}
+%For the analysis, we need to know the names of atoms in your system.
+%These are defined using the mapping file convention introduced in \url{http://nmrlipids.blogspot.com/2015/03/mapping-scheme-for-lipid-atom-names-for.html}, where first column gives the universal atom name, second gives the atom name in force field, and third gives the residue name if not the same for all atoms.
+%The name of the mapping file for each molecule needs to be given in dictionary format.
 The dictionary the key is the molecule name (abbreviation listed in table \ref{tab:abbreviations}) and the value is the name of the mapping file.
@@
 
-\subsubsection*{DIR\_WRK (compulsory)}
-Give the path of the working directory in your local computer.
+%\subsubsection*{DIR\_WRK (compulsory)}
+%Give the path of the working directory in your local computer.
 The trajectory and topology files will be downloaded to this trajectory, and temporary files created during processing will be stored here.
 
-\subsubsection*{UNITEDATOM\_DICT (compulsory for united atom trajectories)}
-Order parameters from united atom simulations are calculated using buildH code (\url{https://github.com/patrickfuchs/buildH}).
+%\subsubsection*{UNITEDATOM\_DICT (compulsory for united atom trajectories)}
+%Order parameters from united atom simulations are calculated using buildH code (\url{https://github.com/patrickfuchs/buildH}).
 For united atom simulations, you need to tell how hydrogens are added based on definitions in the dic\_lipids.py dictionary.
@@
 If correct dictionary key is not yet found, you need to add to dic\_lipids.py.
-In the case of an all atom simulation, UNITEDATOM is left empty.
+%In the case of an all atom simulation, UNITEDATOM is left empty.
 
-\subsubsection*{PUBLICATION}
-Give reference to a publication(s) related to the data.
+%\subsubsection*{PUBLICATION}
+%Give reference to a publication(s) related to the data.
 
-\subsubsection*{AUTHORS\_CONTACT}
-Give the name and email of the main author(s) of the data.
+%\subsubsection*{AUTHORS\_CONTACT}
+%Give the name and email of the main author(s) of the data.
 
-\subsubsection*{SYSTEM}
-Give description of system in free format. For example ''POPC with cholesterol at 301K''.
+%\subsubsection*{SYSTEM}
+%Give description of system in free format. For example ''POPC with cholesterol at 301K''.
 
-\subsubsection*{SOFTWARE\_VERSION}
-Give the version of the software used.
+%\subsubsection*{SOFTWARE\_VERSION}
+%Give the version of the software used.
 
-\subsubsection*{FF}
-Give the name of the force field used used in the simulation.
+%\subsubsection*{FF}
+%Give the name of the force field used used in the simulation.
 
-\subsubsection*{FF\_SOURCE}
-Describe the source of the force field parameters. For example, CHARMM-GUI, link to webpage where parameters were downloaded, or citation to a paper.
+%\subsubsection*{FF\_SOURCE}
+%Describe the source of the force field parameters. For example, CHARMM-GUI, link to webpage where parameters were downloaded, or citation to a paper.
 
-\subsubsection*{FF\_DATE}
-Give the date when parameters were accessed or created. The format is day/month/year.
+%\subsubsection*{FF\_DATE}
+%Give the date when parameters were accessed or created. The format is day/month/year.
 
-\subsubsection*{Individual force field names for molecules}
-In some cases special force fields are used for certain molecules.
+%%\subsubsection*{Individual force field names for molecules}
+%In some cases special force fields are used for certain molecules.
 For example, non-standard parameters for ions or other molecules have been used.
@@
 These can be given as parameters named as
-FF+[abbreviation from table \ref{tab:abbreviations}], i.e., FFPOPC, FFPOT, FFSOL etc.
+%FF+[abbreviation from table \ref{tab:abbreviations}], i.e., FFPOPC, FFPOT, FFSOL etc.
 
-\subsubsection*{CPT (Gromacs)}
-Give the name of the Gromacs checkpoint file that is found from the DOI given above.
-CPT stands for the name of the cpt file.
+%\subsubsection*{CPT (Gromacs)}
+%Give the name of the Gromacs checkpoint file that is found from the DOI given above.
+%CPT stands for the name of the cpt file.
 
-\subsubsection*{LOG (Gromacs)}
-Give the name of the Gromacs log file that is found from the DOI given above.
+%\subsubsection*{LOG (Gromacs)}
+%Give the name of the Gromacs log file that is found from the DOI given above.
 
-\subsubsection*{TOP (Gromacs)}
-Give the name of the Gromacs top file that is found from the DOI given above.
+%\subsubsection*{TOP (Gromacs)}
+%Give the name of the Gromacs top file that is found from the DOI given above.
 
-\subsubsection{Automatically analyzed parameters}
-The following parameters are read automatically from the trajectory and topology files.
+%\subsubsection{Automatically analyzed parameters}
+%The following parameters are read automatically from the trajectory and topology files.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%automatically analyzed parameters%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection*{Molecule numbers}
-Numbers of lipid molecules (NPOPC, NPOPG, etc.) per membrane leaflet are calculated by determining on which side of the center of mass of the membrane the center of mass of the head group of each lipid molecule is located.
-\newline \\
-\noindent Numbers of other molecules such as solvent and ions (NSOL, NPOT, NSOD, etc.) are read from the topology file.
 
-\subsubsection*{Temperature}
-Temperature of the simulation is read from the topology file.
+%\subsubsection*{Molecule numbers}
+%Numbers of lipid molecules (NPOPC, NPOPG, etc.) per membrane leaflet are calculated by determining on which side of the center of mass of the membrane the center of mass of the head group of each lipid molecule is located.
+%\newline \\
+%\noindent Numbers of other molecules such as solvent and ions (NSOL, NPOT, NSOD, etc.) are read from the topology file.
 
-\subsubsection*{Trajectory length}
-The length of a trajectory is read from the trajectory file.
+%\subsubsection*{Temperature}
+%Temperature of the simulation is read from the topology file.
 
-\subsubsection*{Date of running}
-The date of running the script is saved to the README.yaml file.
+%\subsubsection*{Trajectory length}
+%The length of a trajectory is read from the trajectory file.
+
+%\subsubsection*{Date of running}
+%The date of running the script is saved to the README.yaml file.
 

### 2022-05-10 main.tex
Update authors
Update manuscript

--- 2022-05-09 Manuscript/main.tex
@@
 
-\author[1,*]{Anne Kiirikki}
+\author[1]{Anne Kiirikki}
+\author[2]{...}
 \author[1,*]{O. H. Samuli Ollila}
@@
 \affil[1]{University of Helsinki, Institute of Biotechonology, Helsinki, Finland}
-%\affil[2]{Affiliation, department, city, postcode, country}
+\affil[2]{Affiliation, department, city, postcode, country}
 
@@
 
-Quality of conformational ensembles of lipid molecules in simulations are evaluated by estimating the average probability of C-H bond order parameters to agree with the values from NMR experiments. 
+Quality of conformational ensembles of lipid molecules in simulations are evaluated by estimating
+the average probability of C-H bond order parameters to agree with the values from NMR experiments. 
 %As a first step, the quality of each order parameter is evaluated using the negative logarithm of probability to find the simulated order parameter within the experimental error bars
@@
 %\end{equation}
-Because each C-H bond order parameter is calculated as an average over individual lipids in a simulation and conformational ensembles of lipids are assumed to be independent, the random variable
-\begin{equation}
-    \frac{\bar{X}-\mu}{S/\sqrt{n}},
+Because each C-H bond order parameter is calculated as an average over individual lipids in a simulation
+and conformational ensembles of individual lipids can be assumed to be independent, the random variable
+\begin{equation}\label{studentT}
+    \frac{\bar{X}-\mu}{s/\sqrt{n}},
 \end{equation}
-where $\bar{X}$ is the sample mean, $S$ is the sample variance and $n$ is the sample size (number of lipids), has a Student's t-distribution with $n-1$ degrees of freedom and the mean of $\mu$.
+where $\bar{X}$ is the sample mean (order parameter calculated from simulations),
+$s$ is the sample variance (variance of order parameters in different lipids)
+and $n$ is the sample size (number of lipids),
+has a Student's t-distribution with $n-1$ degrees of freedom and the mean of $\mu$.
+Because lipid bilayer simulations contain at least dozens of lipids,
+the Student's t-distribution could be safely approximated with a normal distribution in here.
+However, normal distribution gives values below the numerical accuracy of computers when values
+are far away from the mean which leads to problems when evaluating simulations with poor performance.
+To avoid such numerical instability, we use the first order Student's t-distribution for the equation \ref{studentT}.
 
-where the probability is calculated from the normal distribution
-\begin{equation}\label{gaussian}
-    P = \int_{S_{\rm exp}-\Delta S_{\rm exp}}^{S_{\rm exp}+
-    \Delta S_{\rm exp}}  \frac{1}{\sigma \sqrt{2\pi}} e^{-\frac{1}{2}(\frac{x-\mu}{\sigma})^2} {\rm d}x.
+We estimate the probability for the simulation order parameter to locate within experimental
+error bars from equation
+\begin{equation}\label{probability}
+  P = T(S_{\rm exp}+\Delta S_{\rm exp}) - T(S_{\rm exp}-\Delta S_{\rm exp}),
 \end{equation}
-$S_{\rm exp}$ and $\Delta S_{\rm exp}$ are experimental order parameter and its error, and $\mu$ and $\sigma$ are the mean order parameter and its standard deviation from simulations.
+where T is the first order Student's t-distribution shifted to $\bar{X}$ and
+scaled with $s$. 
+%where the probability is calculated from the normal distribution
+%\begin{equation}\label{gaussian}
+%    P = \int_{S_{\rm exp}-\Delta S_{\rm exp}}^{S_{\rm exp}+
+%    \Delta S_{\rm exp}}  \frac{1}{\sigma \sqrt{2\pi}} e^{-\frac{1}{2}(\frac{x-\mu}{\sigma})^2} {\rm d}x.
+%\end{equation}
+%$S_{\rm exp}$ and $\Delta S_{\rm exp}$ are experimental order parameter and its error, and $\mu$ and $\sigma$ are the mean order parameter and its standard deviation from simulations.
 The quality measure $S_q$ approaches to zero when probability for agreement between simulation and experimental results approach one, and increases when simulated and experimental values diverge.
-The accuracy of $\pm$0.02 is currently assumed for all experimental order parameters~\cite{ollila16}.
+The error of $\Delta S_{\rm exp} = 0.02$ is currently assumed for all experimental order parameters~\cite{ollila16}.
 Because phospholipids sample their conformational ensemble within nanosecond timescale~\cite{ferreira15}, all simulations in the databank would be sufficiently long to sample the realistic conformational phase of individual lipids.
 However, some force fields exhibit too slow dynamics which leads to large error bars in order parameter values~\cite{antila21a}.
-Because large error bars widen the gaussian distribution in Eq.~\ref{gaussian} thereby artificially increasing the probability to find the simulated value within experimental error bars, the order parameters with error bars larger than the experimental error 0.02 are not included in the quality evaluation.
+Because large error bars widen the Student's t-distribution in Eq.~\ref{probability} thereby artificially increasing the probability to find the simulated value within experimental error bars, the order parameters with error bars larger than the experimental error 0.02 are not included in the quality evaluation.
 
-The quality of separate fragments in each lipid type within a simulation are then evaluated by averaging individual order parameter qualities over C-H bond belonging to that fragment and dividing this with the percentage of order parameters for which the quality is available within the fragment, $p$
+The quality of separate fragments in each lipid type within a simulation
+are then estimated by averaging the propabilities for individual C-H bond order parameters
+to locate within experimental errors
+%qualities over  belonging to that fragment
+and dividing this with the percentage of order parameters for which the quality is available within the fragment, $p$
 \begin{equation}

### 2022-05-11 main.tex
Update quality evaluation.

@@
 
-Quality of conformational ensembles of lipid molecules in simulations are evaluated by estimating
-the average probability of C-H bond order parameters to agree with the values from NMR experiments. 
+The quality of conformational ensembles of individual lipid molecules in simulations are evaluated using the C-H bond order parameters which can be directly compared with robust experimental data~\cite{ollila16}.
+%The C--H bond order parameters were calculated directly
+%from the carbon and hydrogen positions using the 
+%definition
+The order parameters are defined as
+\begin{equation}\label{OP}
+S_{\rm CH}=\frac{1}{2}\left\langle 3\cos^2\theta -1 \right\rangle,
+\end{equation}
+where $\theta$ is the angle between the C--H bond and the membrane normal
+%(taken to align with $z$, with bilayer periodicity in the $xy$-plane).
+and angular brackets denote the ensemble average, i.e. average over all sampled configurations of all lipids in a simulation.
+%by estimating
+%the average probability of C-H bond order parameters to agree with the values from NMR experiments. 
 %As a first step, the quality of each order parameter is evaluated using the negative logarithm of probability to find the simulated order parameter within the experimental error bars
@@
 %\end{equation}
-Because each C-H bond order parameter is calculated as an average over individual lipids in a simulation
-and conformational ensembles of individual lipids can be assumed to be independent, the random variable
-\begin{equation}\label{studentT}
-    \frac{\bar{X}-\mu}{s/\sqrt{n}},
+Because conformational ensembles of individual lipids are independent in a simulation of a fluid lipid bilayer, the number of independent sample points for each C-H bond, $n$, equals the number of lipids in a simulation.
+Denoting the $S_{\rm CH}$ from Eq.~\ref{OP} as the sample mean and $s$ its variance calculated over individual lipids,
+%the probability the true mean of the order parameter is given by the %Student's t-distribution 
+%Because each C-H bond order parameter is calculated as an average over individual lipids in a simulation
+%\begin{equation}\label{studentT}
+$\frac{S_{\rm CH}-\mu}{s/\sqrt{n}}$
+%\end{equation}
+%where $\bar{X}$ is the sample mean (order parameter calculated from simulations),
+%$s$ is the sample variance (variance of order parameters in different lipids)
+%and 
+%has a 
+has a Student's t-distribution with $n-1$ degrees of freedom, where $\mu$ is the real mean of the order parameter. 
+%shifted by $S_{\rm CH}$ and scaled by $s/\sqrt{n}$.
+Because lipid bilayer simulations contain at least dozens of lipids, the Student's t-distribution could be safely approximated with a normal distribution.
+However, when calculating the probability for a simulation to locate within experimental order parameters, the normal distribution gives values below the numerical accuracy of computers for simulations values far from experiments.
+To avoid such numerical instability, we use the first order Student's t-distribution.% for the equation \ref{studentT}.
+Therefore, we estimate the probability for the simulation order parameter to locate within experimental error bars from equation
+\begin{equation}\label{probability}
+  P = f \left( \frac{S_{\rm CH} - (S_{\rm exp}+\Delta S_{\rm exp})}{s/\sqrt{n}} \right) - f \left( \frac{S_{\rm CH} - (S_{\rm exp}-\Delta S_{\rm exp})}{s/\sqrt{n}} \right),
 \end{equation}
-where $\bar{X}$ is the sample mean (order parameter calculated from simulations),
-$s$ is the sample variance (variance of order parameters in different lipids)
-and $n$ is the sample size (number of lipids),
-has a Student's t-distribution with $n-1$ degrees of freedom and the mean of $\mu$.
-Because lipid bilayer simulations contain at least dozens of lipids,
-the Student's t-distribution could be safely approximated with a normal distribution in here.
-However, normal distribution gives values below the numerical accuracy of computers when values
-are far away from the mean which leads to problems when evaluating simulations with poor performance.
-To avoid such numerical instability, we use the first order Student's t-distribution for the equation \ref{studentT}.
+where f(t) is the first order Student's t-distribution.
+% shifted to $\bar{X}$ and scaled with $s$.
 
-We estimate the probability for the simulation order parameter to locate within experimental
-error bars from equation
-\begin{equation}\label{probability}
-  P = T(S_{\rm exp}+\Delta S_{\rm exp}) - T(S_{\rm exp}-\Delta S_{\rm exp}),
-\end{equation}
-where T is the first order Student's t-distribution shifted to $\bar{X}$ and
-scaled with $s$. 
+%We estimate the probability for the simulation order parameter to locate within experimental
+%error bars from equation
+%\begin{equation}\label{probability}
+%  P = T(S_{\rm exp}+\Delta S_{\rm exp}) - T(S_{\rm exp}-\Delta S_{\rm exp}),
+%\end{equation}
+%where T is the first order Student's t-distribution shifted to $\bar{X}$ and
+%scaled with $s$. 
 %where the probability is calculated from the normal distribution
@@
 
-The quality of separate fragments in each lipid type within a simulation
-are then estimated by averaging the propabilities for individual C-H bond order parameters
-to locate within experimental errors
-%qualities over  belonging to that fragment
-and dividing this with the percentage of order parameters for which the quality is available within the fragment, $p$
+The qualities of different fragments within each lipid type in a simulation, $P^{\rm frag}[{\rm lipid}]$, are then estimated by averaging the probabilities for individual C-H bond order parameters to locate within experimental errors according to Eq.~\ref{probability}, and dividing the average with the percentage of order parameters for which the quality is available within the fragment, $p_{\rm frag}[{\rm lipid}]$,
 \begin{equation}
-    S_q^{\rm frag}[{\rm lipid}] = \frac{\langle S_q[{\rm lipid}]\rangle_{\rm frag}}{p_{\rm frag}[{\rm lipid}]},
+    P^{\rm frag}[{\rm lipid}] = \frac{\langle P[{\rm lipid}]\rangle_{\rm frag}}{p_{\rm frag}[{\rm lipid}]},
 \end{equation}
@@
 \begin{equation}
-    S_q^{\rm frag} = \sum_{\rm lipid} \chi_{\rm lipid} \langle S_q^{\rm frag}[{\rm lipid}]\rangle_{\rm lipid} ,
+    P^{\rm frag} = \sum_{\rm lipid} \chi_{\rm lipid} P^{\rm frag}[{\rm lipid}],
 \end{equation}
@@
 
-Qualities of form factors were evaluated with the same approach that was used in SIMtoEXP program~\cite{kucerka10}.
-Because experiments give form factors only in relative scale those were scaled to the simulation data with absolute scale using equation
+Qualities of form factors were evaluated using the same approach as in SIMtoEXP program~\cite{kucerka10}.
+Because experiments give form factors only in relative intensity scale, they were scaled to the simulation data using equation
 \begin{equation}

### 2022-08-09 main.tex
Update text

@@
 
+For the permeation of water through membranes, the number of permeation events in each trajectory was first calculated using the code by Camilo et al.~\cite{camilo2022}, available at~\url{https://github.com/crobertocamilo/MD-permeation}.
+The permeation was then calculated from equation $P=r/2c_w$, where $r$ is the rate per time and area, and $c_w$ is the concentration of water in bulk with the value of 33.3679\,(nm)$^{-3}$~\cite{venable19}.
+
 %\subsection{Indexing the simulation data}

### 2022-08-11 main.tex
Add water permeation to manuscript

@@
 
-\subsection{NMRlipids databank reveals correlations between membrane properties}
-The power of NMRlipids databank to find correlation between membrane properties is exemplified in Fig.~\ref{??} showing how water diffusion through and along the membrane correlate with different properties of systems.
+\subsection{Water permeation through membranes}
+Dependence of water permeation rate through membranes on the membrane properties has been under great interest because
+it serves as a proxy for permeation of other hydrophilic molecules such as drugs \cite{??}, and is also relevant in understanding
+MRI imaging \cite{??}. Several models for water permeation are formulated based on experiments and simulations predicting
+different dependencies on membrane properties but concensus on how water permeability depends on basic membrane properties,
+such as area per lipid and thickness, or molecular composition has not yet been reached \cite{shinoda16,venable19,??}.
+This is partially because collection of experimental and simulation data for water permeability from membranes with
+wide range of properties has been tedious.
+Here we demonstrate the power of the NMRlipids databank by analysing how the water permeability through lipid bilayers depend
+on area per lipid, thickness and temperature. The results shown in Fig.~\ref{fig:permeability} are calculated by averaging over
+all the simulations in the databank with the same variable under interest (area per lipid, thickness or temperature).
+%to find correlation between membrane properties is exemplified in Fig.~\ref{??} showing how water diffusion through and along the membrane correlate with different properties of systems.
 While NMRlipids databank enables similar analyses for any membrane properties, these are selected due to their relevance in the development of MRI imaging methods~\cite{??}.
+
+\begin{figure}[tbp]
+    \centering
+    \includegraphics[width=88mm]{Figures/permeation.pdf}
+    \caption{Water permeation through lipid membranes analyzed from the databank as a function of 
+      a) area per lipid,
+      b) thickness,
+      and c) temperature. 
+      Inserts in a) and b) show the region where the dependence could be considered approximately linear.
+      Insert in c) shows the Arrhenius plot of permeation ($\ln$ (P) vs. 1/T) that gives 17\,$\pm$\,4\,k$_b$T for the average activation energy for water permeation through lipid bilayer.}
+    \label{fig:permeability}
+\end{figure}
+
+The water permeabilities observed in the databank vary between 0.3 and 322\,$\mu$m/s with the mean and median of 14\,$\mu$m/s and 8\,$\mu$m/s, respectively.
+These values are within the range of experimentally reported values and the mean is the same order of magnitude as the diffusive permeability which is 1-2 orders of 
+magnitude slower than osmotic permeability \cite{jansen95,??}. While linear dependence of permeability on both area per lipid and thickness have been proposed based on
+simulations or experiments \cite{mathai08,frallicciardi22}, our results suggest significantly stronger dependence for thicknesses below 3.9\,nm and areas above 69\,Å$^2$.
+Nevertheless, the dependence may appear linear above and below these values as shown in the insets of Figs. \label{fig:permeability} a) and b).
+As expected, the permeability increases with increasing temperature in Figs. \label{fig:permeability} c).
+The Arrhenius plot gives 17\,$\pm$\,4\,k$_b$T for the average energy barrier for the water permeation.
 

### 2022-08-15 main.tex
Update text

--- 2022-05-11 main.tex
@@
 \tablefootnote{For example, if you upload 0-200 ns part of total 200 ns simulation where the first 100 ns should be considered as an equilibration, this value should be 100.} \\
-    COMPOSITION & Molecules names used in the simulation and corresponding mapping files (see ??) & \\
+    COMPOSITION & Molecules names used in the simulation and corresponding mapping files (see section \ref{naming}) & \\
     DIR\_WRK & Temporary working directory in your local computer. \\
@@
     MOLAR\_FRACTIONS & Dictionary of molar fractions of bilayer components \\
-    ION\_CONCENTRATIONS & Dictionary of ion concentrations of the system (defined as ??) \\
+    ION\_CONCENTRATIONS & Dictionary of ion concentrations of the system \\% (defined as ??) \\
     TOTAL\_LIPID\_CONCENTRATION & Total concentration of lipid components. If exact concentration is not known, but experiments are performed in excess water, 'full hydration' can be given. \\
@@
 
-\subsection{Molecule and atom naming convention}
+\subsection{Molecule and atom naming convention} \label{naming}
 

### 2022-08-25 main.tex

@@
 
-\subsection{Water permeation through membranes}
-Dependence of water permeation rate through membranes on the membrane properties has been under great interest because
-it serves as a proxy for permeation of other hydrophilic molecules such as drugs \cite{??}, and is also relevant in understanding
-MRI imaging \cite{??}. Several models for water permeation are formulated based on experiments and simulations predicting
-different dependencies on membrane properties but concensus on how water permeability depends on basic membrane properties,
-such as area per lipid and thickness, or molecular composition has not yet been reached \cite{shinoda16,venable19,??}.
+\subsection{Water diffusion anisotropy in membrane systems}
+The anisotropy of water diffusion in the direction of membrane surface and through the membrane is utilized in MRI imaging \cite{??} and regulates the drug translocation particularly in skin \cite{hansen13,wen18,nitsche19,roberts21,??}.
+The importance of membrane properties on the anisotropic diffusion has been recognized and several models linking the water permeation with the membrane properties have been formulated \cite{??}, yet the consensus on how water permeability depends on basic membrane properties, such as area per lipid and thickness, or molecular composition has not yet been reached \cite{nitsche13,nitsche16,shinoda16,venable19,??}.
 This is partially because collection of experimental and simulation data for water permeability from membranes with

### 2022-08-25 main.tex
Updates from Overleaf

--- 2022-08-15 main.tex
@@
 
-\subsection{NMRlipids databank reveals correlations between membrane properties}
-The power of NMRlipids databank to find correlation between membrane properties is exemplified in Fig.~\ref{??} showing how water diffusion through and along the membrane correlate with different properties of systems.
-While NMRlipids databank enables similar analyses for any membrane properties, these are selected due to their relevance in the development of MRI imaging methods~\cite{??}.
+\subsection{Water diffusion anisotropy in membrane systems}
+The anisotropic water diffusion through and along of membranes is utilized in MRI imaging \cite{??} and plays a role in the drug translocation through biological material, particularly in skin \cite{hansen13,wen18,nitsche19,roberts21,??}.
+Dependencies between membrane properties, water permeablity, and anisotropic water diffusion have been recognized \cite{??} but consensus on how water permeability depends on membrane properties, such as area per lipid and thickness, or molecular composition, has not yet been reached despite of several available models~\cite{nitsche13,nitsche16,shinoda16,venable19,??}.
+MD simulations can be used to elucidate such dependencies, but systematic collection of sufficient amount of data has been challenging because few water permeation events are typically observed in a single MD simulation trajectory \cite{??}.
+Nevertheless, averaging over large amounts of data available in the NMRlipids databank enables to study the trends of water permeation upon changing memebrane properties.
 
-%\subsection{Spin relaxation rates of confined water close to membranes}
+%averaging Here we demonstrate the power of the NMRlipids databank by analysing how 
+Water permeabilities calculated from all the simulations in the NMRlipids databank as a function of area per lipids, thickness, and acyl chain order are shown in  Fig.~\ref{fig:permeability} (a-c).
+To illustrate how permeability depends these properties, average over systems with fixed values of area per lipids, thickness, and acyl chain order are also shown in the figure.
+Permeability increases with increasing area per lipid and decreases with increasing membrane thickness and acyl chain order.
+Previous simulation and experimental studies have reported linear dependence of permeability on both area per lipid and thickness for limited set of lipid mixtures but not always for all types of lipid mixtures \cite{mathai08,frallicciardi22}.
+Our results suggest stronger dependence for thicknesses below 3.9\,nm and areas above 69\,Å$^2$, but the dependence may appear linear above and below these values as shown in the insets of Figs.
+\label{fig:permeability} a) and b).
+As expected, the permeability increases with increasing temperature in Figs.
+\label{fig:permeability} d).
+The Arrhenius plot gives 17\,$\pm$\,4\,k$_b$T for the average energy barrier for the water permeation.
+The values for water permeabilities from simulations in the databank vary between 0.3 and 322\,$\mu$m/s with the mean and median of 14\,$\mu$m/s and 8\,$\mu$m/s, respectively.
+These values are within the range of experimentally reported values and the mean is the same order of magnitude as the diffusive permeability which is 1-2 orders of magnitude slower than osmotic permeability \cite{jansen95,??}.
+Clear dependencies of permeability on hydration level, charged lipid fraction, cholesterol fraction or POPE fraction were not observed, although weak decrease for the latter two may be visible.
 
-%Usefulness of the databank beyond MD simulation experts is demonstrated by analysing water spin relaxation times close to bilayers which are used in MRI imaging.
+\begin{figure}[tbp]
+    \centering
+    \includegraphics[width=180mm]{Figures/permeation2.pdf}
+    \caption{Water permeation through lipid membranes analyzed from the databank as a function of 
+      a) area per lipid,
+      b) thickness,
+      and c) temperature. 
+      Inserts in a) and b) show the region where the dependence could be considered approximately linear.
+      Insert in c) shows the Arrhenius plot of permeation ($\ln$ (P) vs. 1/T) that gives 17\,$\pm$\,4\,k$_b$T for the average activation energy for water permeation through lipid bilayer.}
+    \label{fig:permeability}
+\end{figure}
 
-%Example text under a subsection. Bulleted lists may be used where appropriate, e.g.
-
-%\begin{itemize}
-%\item First item
-%\item Second item
-%\end{itemize}
-
-%\subsubsection*{Third-level section}
-
-%Topical subheadings are allowed.
+To analyze the anisotropic water diffusion, we also calculated the water diffusion along the membrane.
+The results at different hydration levels are shown in Fig.~\ref{fig:permeability} e) together with the experimental data~\cite{??}.
+The experimental water diffusion increases with increasing hydration level toward the value for bulk water (??
+in ??K).
+Simulations overestimate the experimental bulk value approximately with the factor of 2 at high hydration levels which is not surprising as the most common water model used in membrane simulations, TIP3P, overestimates the bulk water diffusion.
+However, simulation results are closer to experiments with low hydration levels.
+To estimate the diffusion anistoropy, we translated the water permation to diffusion coefficent through multilamellar stack using the Tanner equation~\cite{??}.
+The results in
 

### 2022-08-31 SI.tex

--- 2022-08-15 main.tex
@@
 
-\title{NMRlipids Databank: Overlay Databank of Lipid Membrane Simulations Arising from Open Collaboration}
+\title{SUPPLEMENTARY INFORMATION: NMRlipids Databank: Overlay Databank of Lipid Membrane Simulations Arising from Open Collaboration}
 
@@
 \begin{abstract}
-We present a databank of lipid bilayer simulations from the NMRlipids open collaboration project.
+%We present a databank of lipid bilayer simulations from the NMRlipids open collaboration project.
 \end{abstract}
+
 \begin{document}
@@
 
-\section{Introduction}
+\pagebreak
+\section{Water permeation through membranes}
 
-%The demand for sharing and reusing of MD simulation data is increasing, but practical solution remains unclear due to unsolved issues in data storage and indexing.
-
-The importance of sharing MD simulation data following the FAIR principles~\cite{wilkinson16} has been widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20}, and databanks are emerging~\cite{meyer10,kamp10,hospital16,mixcoha16,newport19,bekker20,espigares20,leston22}.
-The relevance of quality evaluation of simulation trajectories in databanks regarding technical details of simulations and accuracy of the underlying physical description of the system (force field) has become evident~\cite{tai04,meyer10,hospital20} and such quality evaluation has in some cases also been implemented~\cite{meyer10,hospital16}.
-However, straightforward quality comparisons between individual simulations or force fields within these databanks remain challenging.
-While importance of such databanks for MD simulations is widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} and different kinds of approaches are emerging~\cite{meyer10,kamp10,hospital16,mixcoha16,newport19,bekker20,espigares20,leston22}, generally accepted protocols and best practices are still under active development.
-
-Here we present a solution for lipid bilayers based on overlay databank structure illustrated in Fig.~\ref{fig:overlay}.
-The concept of overaly databank is developed here to solve the practical challenges in generating databanks of MD simulation data enabling flexible analyses over large sets of simulation data, but it pontentially used for wide range of situations, particularly when storage of raw data requires significant resources and final outcomes or best practices are not yet clear, overlay databank approach lowers the barrier to start without compromizing the long term stability or scalability.
-
-The practical relevance of the NMRlipids databank is exemplified by automatic quality evaluation and ranking of large amount of MD simulation data, data-driven analysis detecting correlations between properties of model cell membranes and analyses of rare phenomena that are beyond the scope of standard MD simulation studies.
-The NMRlipids databank provides new tools for researchers in wide range of fields in academia and industry from cell membrane biology to lipid nanoparticle formulations and data-driven computational chemistry and machine learning.
-
-%The Introduction section, of referenced text\cite{Figueredo:2009dg} expands on the background of the work (some overlap with the Abstract is acceptable).
-The introduction should not include subheadings.
-
-\section{Results}
-
-%Up to three levels of \textbf{subheading} are permitted. Subheadings should not be numbered.
-\subsection{Design of the NMRlipids databank}
-The key idea of the overlay databank is that the storage of raw data in layer 1 is distributed in publicly available repositories or other servers with long term stability and permanent links such as digital object identifiers.
-The core of the databank, layer 2, contains only information on the location and content of the raw data, thereby not requiring large resources to handle and maintain.
-This lowers the barrier for starting such databank as well as for long term storage.
-The NMRlipids databank is essentially a git repository containing information on the location of raw data and its indexing with universal naming convention.
-In addition to all computers where the databank is developed and used, the NMRlipids databank git is stored to Zenodo server, thereby enabling a very cost effective long term storage for the databank.
-The databank can be used in layer 3 by accessing the raw data and information stored in the databank by employing the universal naming conventions for atoms, molecules and simulation details.
-The applications can be linked to the core databank (layer 2) without actually including them, thereby enabling flexible extension of the data without compromising the simplicity and lightness of the core databank.
-
-\begin{figure}
+\begin{figure}[!h]
     \centering
-    \includegraphics[width = 180mm]{Figures/overlay.pdf}
+    \includegraphics[width = 180mm]{Figures/permeationSI.pdf}
     \caption{a) Structure of an overlay databank. 
-    %Raw data is available from a publicly accessible repository (layer 1).
-    %The core of the databank contains information on the location of the raw data and indexing of molecular %compositions and simulation details described with universal naming conventions (layer 2).
-    %The content of databank can be viewed and analysed using external programs that automatically read the information and data from layers 1 and 2 (layer 3).
-Examples of such applications are the GitHub repository generating the results presented in this work and the graphical access to the databank.
     More detailed structure of the layer 2 in the NMRlipids databank is illustrated in Fig.~\ref{DatabankStructure} in the SI.
@@
 
-Currently the databank is composed of approximately 500 trajectories with the total length of approximately 231 microseconds of which most are contributed for the previous publications from the NMRlipids open collaboration~\cite{botan15,catte16,antila19,bacle21}.
-The distribution of lipids, force fields, length of the trjajectories and available binary mixtures are shown in Fig.~\ref{fig:overlay}.
+\pagebreak
+\section{Water diffusion along membranes}
 
-\subsection{Quality evaluation of force fields}
-Quality of membrane simulations with different force fields have been evaluated against experimental data during parameterization and in separate comparison studies~\cite{botan15,ollila16,catte16,pluhackova16,perez17,leonard19}, but universal quality measure for membrane simulations is not defined and controversial results are often reported from simulations~\cite{antila22b}.
-The unclarities in simulation quality complicate the selection of proper simulation models for specific applications and estimation of reliability of reported simulation results, thereby being a major obstacle in many applications of membrane MD simulations.
-
-To enable rapid evaluation and comparison of membrane MD simulation qualities, we define the quality measure for membrane MD simulations using the C-H bond order parameters from NMR experiments and form factors from x-ray scattering, which are robust experimental measurables that can be directly connected to the simulation data~\cite{ollila16}.
-C-H bond order parameters are related to the conformational ensemble of individual lipid molecules, while the form factor to arises from electron density thereby connecting to the overall structure of lipid bilayers.
-This is demonstrated in Fig.~\ref{fig:quality} a) showing correlations between membrane lateral density (area per lipid), thickness, minima of form factor and acyl chain order analyzed from all apporimately 500 trajectories in the NMRlipids databank.
-Membrane area per lipid has negative correlations with membrane order and thickness with Pearson coefficient of -0.78 and -0.49, respectively, because decreased area leads to more ordered, tightly packed and stretched acyl chains leading to thicker membranes.
-On the other hand, locations of form factor minima have positive correlation with the membrane area and negative correlation with the thickness.
-In conclusion, both x-ray scattering form factors and acyl chain order parameters from NMR are on average good proxies for membrane packing and thickness, while NMR order parameters give segmental resolution information on the quality of conformational ensembles of lipids.
-
-\begin{figure}[tbp]
+\begin{figure}[!h]
     \centering
-    \includegraphics[width=180mm]{Figures/quality.pdf}
-    \caption{a) Correlations between membrane area per lipid, thickness, second minima of form factor and average order parameter of the {\it sn}-1 acyl chain extracted from the NMRlipids databank.
-All Pearson correlation coefficients shown in bottom left corner have p-value below 0.001.
-    b) Evaluation against experimental data exemplified for the simulation giving the best qualities in overall, for headgroup and x-ray scattering form factor.
-    c) Change in area per lipid upon addition of POPE to POPC in different simulation extracted from the NMRlipids databank and quality evaluation table of POPE simulations.
-    d) Quality evaluation table showing the best 13 simulations according to the overall order parameter quality.}
-    \label{fig:quality}
+    \includegraphics[width = 180mm]{Figures/LateralDiffusionSI.pdf}
+    \caption{a) Structure of an overlay databank. 
+    More detailed structure of the layer 2 in the NMRlipids databank is illustrated in Fig.~\ref{DatabankStructure} in the SI.
+    b) Distribution of the lengths of the trajectories, total number of trajectories and total lenght of the simulations in the NMRlipids databank.
+    c) Distribution of lipids present in the trajectories in the NMRlipids databank. Lipids occuring in five or less simulations ('others') are listed in the right. 
+    d) Currently available binary mixtures in the NMRlipids databank. 
+    e) Distribution of force fields in the simulations in the NMRlipids databank.
+    The figures and numbers are created on 9th of May 2022.}
+    \label{fig:overlay}
 \end{figure}
-
-NMRlipids databank contains also experimental order parameter and x-ray form factor data that is connected to corresponding  simulations in order to define the simulation qualities and rank them to select the suitable simulation models for particular applications.
-The ranking of all simulations based on estimated average probability of order parameters to locate within experimental errors are shown in Fig.~\ref{fig:quality} b) and the comparison for the best models for overall quality, headgroup and glycerol backbone, and form factor are exemplified in Fig.~\ref{fig:quality} d).
-Full details of the NMRlipids quality measure for lipid bilayer simulations are given in the methods section and supplementary information.
-The top ranking simulations in Figs.~\ref{fig:quality} b) and d) demonstrate also the current complexity in lipid bilayer simulation quality.
-Simulation of POPC lipid bilayer with Slipids force field ranks first in overall score, glycerol backbone is not correcetly captured in this model.
-On the othere, simulation with the best quality for headroup and glycerol backbone, CHARMM36 POPC, predicts too ordered acyl chain, thereby not being within the top 10 simulations in the overall ranking.
-The best ranking for x-ray form factor is given by the Slipids POPE simulation, however, because the quantitative value for form factor quality depends on the quality of experimental data, form factor qualities can be compared only between simulations evaluated against the same experimental dataset.
-
-The power of NMRlipids quality metrics in selecting the best model for particular application is demontrated in Fig.~\ref{fig:quality} c) where the area per lipids of POPC:POPE mixtures from different force fields are shown and quality of POPE simulation in different force fields are evaluated.
-Among the simulations available in the NMRlipids databank, the Slipids force field gives the best quality in terms of acyl chain order parameters and form factor, and predicts the largest area per molecule for POPE and smallest difference with POPC.
-In conclusion, simulations with Slipids is the most reliable force field for membrane packing in POPC:POPE mixtures althought its glycerol backbone is not accurately modelled.
-Similar comparisons utilizing the preliminary data from the NMRlipids databank have concluded that Slipids is relatively good model also for mixtures of charged POPC:POPS membranes, although models with better counterion binding have higher quality, but CHARMM36 is the best to study differences in headgroup conformational ensembles between different lipid types.
-Understanding such complex picture of lipid bilayer MD simulation quality would not be possible without automatic quality evaluation of large sets of simulations enabled by the NMRlipids databank.
-
-\subsection{Water permeation through membranes}
-Dependence of water permeation rate through membranes on the membrane properties has been under great interest because
-it serves as a proxy for permeation of other hydrophilic molecules such as drugs \cite{??}, and is also relevant in understanding
-MRI imaging \cite{??}. Several models for water permeation are formulated based on experiments and simulations predicting
-different dependencies on membrane properties but concensus on how water permeability depends on basic membrane properties,
-such as area per lipid and thickness, or molecular composition has not yet been reached \cite{shinoda16,venable19,??}.
-This is partially because collection of experimental and simulation data for water permeability from membranes with
-wide range of properties has been tedious.
-Here we demonstrate the power of the NMRlipids databank by analysing how the water permeability through lipid bilayers depend
-on area per lipid, thickness and temperature. The results shown in Fig.~\ref{fig:permeability} are calculated by averaging over
-all the simulations in the databank with the same variable under interest (area per lipid, thickness or temperature).
-%to find correlation between membrane properties is exemplified in Fig.~\ref{??} showing how water diffusion through and along the membrane correlate with different properties of systems.
-While NMRlipids databank enables similar analyses for any membrane properties, these are selected due to their relevance in the development of MRI imaging methods~\cite{??}.
-
-\begin{figure}[tbp]
-    \centering
-    \includegraphics[width=88mm]{Figures/permeation.pdf}
-    \caption{Water permeation through lipid membranes analyzed from the databank as a function of 
-      a) area per lipid,
-      b) thickness,
-      and c) temperature. 
-      Inserts in a) and b) show the region where the dependence could be considered approximately linear.
-      Insert in c) shows the Arrhenius plot of permeation ($\ln$ (P) vs. 1/T) that gives 17\,$\pm$\,4\,k$_b$T for the average activation energy for water permeation through lipid bilayer.}
-    \label{fig:permeability}
-\end{figure}
-
-The water permeabilities observed in the databank vary between 0.3 and 322\,$\mu$m/s with the mean and median of 14\,$\mu$m/s and 8\,$\mu$m/s, respectively.
-These values are within the range of experimentally reported values and the mean is the same order of magnitude as the diffusive permeability which is 1-2 orders of 
-magnitude slower than osmotic permeability \cite{jansen95,??}. While linear dependence of permeability on both area per lipid and thickness have been proposed based on
-simulations or experiments \cite{mathai08,frallicciardi22}, our results suggest significantly stronger dependence for thicknesses below 3.9\,nm and areas above 69\,Å$^2$.
-Nevertheless, the dependence may appear linear above and below these values as shown in the insets of Figs. \label{fig:permeability} a) and b).
-As expected, the permeability increases with increasing temperature in Figs. \label{fig:permeability} c).
-The Arrhenius plot gives 17\,$\pm$\,4\,k$_b$T for the average energy barrier for the water permeation.
-
-%\subsection{Spin relaxation rates of confined water close to membranes}
-
-%Usefulness of the databank beyond MD simulation experts is demonstrated by analysing water spin relaxation times close to bilayers which are used in MRI imaging.
-
-%Example text under a subsection. Bulleted lists may be used where appropriate, e.g.
-
-%\begin{itemize}
-%\item First item
-%\item Second item
-%\end{itemize}
-
-%\subsubsection*{Third-level section}
-
-%Topical subheadings are allowed.
-
-\section{Discussion}
-
-%The Discussion should be succinct and must not contain subheadings.
-
-Quality measure and automatic quality evaluation of lipid bilayer MD simulations introduced in the NMRlipids databank enables rapid ranking of available simulation models agains experimental NMR and x-ray scattering data.
-This provides a tool for researches to rapidly evaluate the credibility of MD simulations of their own and reported by other groups.
-Because such tool and quality measure has not been available, this will elaborate the quality of published MD simulations of lipid bilyers and reduce potentially misleading results~\cite{antila22b}.
-The power of NMRlipids databank to select the best models for particular applications has been demontrated for PC/PE lipid mixtures (Fig.~\ref{fig:quality}), PC/PS lipid mixtures~\cite{antila22b}, and lipid headgroups~\cite{bacle21}.
-
-The increasing amount of MD simulation data with programmatic access in the NMRlipids databank opens up possibilities for wide range of applications utilizing the large set of accessible data.
-Extend of the data in the NMRlipids databank in terms of quantity (e.g., simulation length and number of conformations), content (e.g, lipid compositions and ion concentrations) and quality enables analyses that are not possible to conduct from MD simulation data produced by a single research group.
-Applications of the NMRlipids databank to understand how diffusion of water through and along cellular membrane depends on its physical properties are demonstrated in Fig.~\ref{??}.
-Permeation of water through membranes resembles the permeation of also other hydrophilic molecules, such as drugs, and detailed understanding of water dynamics through and along membranes is potentially useful for the development of MRI imaging methods~\cite{topgaard20}.
-These examples demonstrate the practical applications of NMRlipids databank on problems in the biological and biomedical sciences.
-
-Building accessible databanks of molecular dynamics simulation data has been challenging due to the required long term support for hardware and software maintenance.
-In the overlay model used in the NMRlipids databank, the demand of hardware can be distributed and open collaboration model reduces the risk for ending software maintenance.
-Furthermore, the open collaboration model used in the NMRlipids project credits contributors by offering authorship in published articles, thereby creating an incentive for contributions.
-This model could be extended also to other fields where similar barriers to establish publicly accessible databanks exist.
-Emerging applications of machine learning are increasing the impact of such databanks.
-For example, the existing Protein Databank (PDB)~\cite{??} containing experimentally determined protein structures with programmatic access has enabled the development of machine learning based tools building on the data collected in the databank over the years~\cite{??}.
-NMRlipids and other databanks with open programmatic access have potential to lead similar unforeseen applications in the future.
-
-\section{Methods}
-
-%Topical subheadings are allowed. Authors must ensure that their Methods section includes adequate experimental and characterization data necessary for others in the field to reproduce their work.
-
-\subsection{Structure of the databank}
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{Figures/DataBankChart.pdf}
-  \caption{Structure of the NMRlipids databank.
-Manually added input data (blue boxes) includes basic information on the simulation, permanent links to the raw data, and experimental data if available.
-The databank entries (red box) and analysis results (green boxes), locating at \url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations} are automatically generated by the computer programs included in the NMRlipids databank (yellow boxes).
-Because raw data are not permanently stored but can be accessed based on the information in the databank, this connection is marked with the dashed line.
-}\label{DatabankStructure}
-\end{figure}
-
-Structure of the NMRlipids databank is illustrated in Fig.~\ref{DatabankStructure}.
-The required input information to create an entry into the NMRlipids databank are listed in table~\ref{tab:READMEkeys}.
-While the raw simulation data is not directly stored in the NMRlipids databank, permanent links from where the raw data can be accessed have to be given and are then stored in the README.yaml files at \url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations}.
-These files contain all the essential information listed in table~\ref{tab:READMEkeys} on each simulation entry that are needed for further applications.
-The raw MD simulation data can locate in any stable publicly available repository, although all the current data locates in Zenodo \url{www.zenodo.org}.
-
-%A README.yaml file is made using AddData.py script for each simulation stored in the databank based on the information given by the author of the simulation data.
-In addition to this information, automatically extracted information about the simulation are stored in the README.yaml files (see table~\ref{tab:READMEkeys}).
-A README.yaml file for each simulation is stored to a folder named according to the hashes of original trajectory and topology file locating in \url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations}.
-These README.yaml files give access to all the essential information on the simulations in the databank either directly, or pointing to a original file from where the information can be extracted.
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{  p{3.5cm}  p{9.5cm}  p{4.0cm} }
-    \toprule
-    key & description & type  \\
-    \midrule
-        %\hline
-    DOI & DOI from where the raw data is found & user given (compulsory) \\
-    SOFTWARE & Software used to run the simulation (e.g. Gromacs, Amber, NAMD, etc.) & \\
-    TRJ & Name of the trajectory file found from DOI & \\
-    TPR & Name of the topology file found from DOI (trp file in the case of Gromacs) & \\
-    PREEQTIME & Pre-equilibrate time simulated before the uploaded trajectory in nanoseconds.
-\tablefootnote{For example, if you upload 100-200 ns part of total 200 ns simulation, this should value should be 100.} & \\
-    TIMELEFTOUT & Equilibration period in the uploaded trajectory that should be discarded in analyses.
-\tablefootnote{For example, if you upload 0-200 ns part of total 200 ns simulation where the first 100 ns should be considered as an equilibration, this value should be 100.} \\
-    COMPOSITION & Molecules names used in the simulation and corresponding mapping files (see section \ref{naming}) & \\
-    DIR\_WRK & Temporary working directory in your local computer. \\
-    UNITEDATOM\_DICT & Information for constucting hydrogens for united atom simulations, empty for all atom simulations & \\
-    TYPEOFSYSTEM & Lipid bilayer or something else & \\
-    \hline
-    PUBLICATION & Give reference to a publication(s) related to the data. & User given (optional)\\
-    AUTHORS\_CONTACT & Name and email of the main author(s) of the data. & \\
-    SYSTEM & System description on free text format & \\
-    SOFTWARE\_VERSION & Version of the used software & \\
-    FF & Name of the used force field & \\
-    FF\_SOURCE & Source of the force field parameters, e.g, CHARMM-GUI, webpage, citation to a publication, etc. & \\
-    FF\_DATE &  Date when force field parameters were accessed on the gives source (day/month/year). & \\
-    FF{molename} & Molecule specific force field information, e.g., water model with FFSOL and sodium parameters with FFSOD. & \\
-    CPT & Name of the Gromacs checkpoint file. & \\
-    LOG & Name of the Gromacs log file. & \\
-    TOP & Name of the Gromacs top file. & \\
-    GRO & Name of the Gromacs gro file. & \\
-    \hline
-    TRAJECTORY\_SIZE & Size of the trajectory file in bytes & automatically extracted data. \\
-    TRJLENGTH & Lenght of the trajectory (ps). & \\
-    TEMPERATURE & Temperature of the simulation. & \\
-    NUMBER\_OF\_ATOMS & Number of atoms in the simulation. & \\
-    DATEOFRUNNIG & Date when added into the databank & \\
-    EXPERIMENT & Potentially connected experimental data & \\
-    COMPOSITION & Numbers of lipid molecules (NPOPC, NPOPG, etc.) per membrane leaflet are calculated by determining on which side of the center of mass of the membrane the center of mass of the head group of each lipid molecule is located.
-    Numbers of other molecules such as solvent and ions (NSOL, NPOT, NSOD, etc.) are read from the topology file. & \\
-    \end{tabular}
-    \caption{Keys stored in the README.yaml files of simulations.}
-    \label{tab:READMEkeys}
-\end{table}
-
-In order to evaluate the quality of simulations, sets of C-H bond order parameters from NMR and from factors from x-ray scattering are included in the NMRlipids databank (\url{https://github.com/NMRLipids/Databank/tree/main/Data/experiments}).
-The required information for an experimental dataset are listed in table~\ref{tab:READMEkeysEXP}.
-A simulation is connected to a experimental data set
-%C-H bond order parameter and x-ray scattering form factor data using the searchDATABANK.py script.
-Available experimental data is collected to  where each dataset is saved into a folder with the name based on DOI of the related publication and experimental conditions are described in README.yaml files with the keys listed in table~\ref{tab:READMEkeysEXP}.
-if molar concentrations of all molecules are within $\pm$5 percentage units, charged lipids have the same counterions, and temperature is within $\pm$2 degrees.
-In such cases, a simulation and experimental data are paired by adding the path to the experimental data into the simulation README.yaml file.
-%If lipid to water ratio is above ??, the systems are considered fully hydrated. Otherwise also hydration level is considered.
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{  p{5.0cm}  p{10.0cm}}
-    \toprule
-    key & description \\
-    \midrule
-    DOI & DOI of the publication related to the experimental data. \\
-    TEMPERATURE & Temperature of the experiment. \\
-    MOLAR\_FRACTIONS & Dictionary of molar fractions of bilayer components \\
-    ION\_CONCENTRATIONS & Dictionary of ion concentrations of the system \\% (defined as ??) \\
-    TOTAL\_LIPID\_CONCENTRATION & Total concentration of lipid components. If exact concentration is not known, but experiments are performed in excess water, 'full hydration' can be given. \\
-    COUNTER\_IONS & Type of counter ions if present.
-\end{tabular}
-    \caption{Keys stored in the README.yaml files of experiments.}
-    \label{tab:READMEkeysEXP}
-\end{table}
-
-Because README.yaml contains all the essential information on each simulation, arbitrary analyses can be automatically performed over all the simualtions in the NMRlipids databank.
-In addition to the order parameters for each C-bonds and x-ray scattering form factors used in the quality evaluation, the NMRlipids databank contains the area per lipid and thickness calculated from all simulations in the databank.
-These results are stored in the same folders as the README.yaml files in \url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations}.
-
-\subsection{Molecule and atom naming convention} \label{naming}
-
-Because universal convention for lipid molecules and atoms therein has not been defined, the naming conventions vary between authors and force fields.
-To enable automatic analyses over large sets of simulation in the NMRlipids databank, we have defined unique naming conventions for lipid molecules and atoms.
-The abbreviations of molecule names used in the NMRlipids databank are listed in table~\ref{tab:abbreviations}.
-The unique atom names for each molecule and corresponding names in each simulation are defined using mapping files introduced in the NMRlipids project (\url{https://nmrlipids.blogspot.com/2022/04/new-yaml-format-of-mapping-files.html}).
-
-Molecule and atom names in each simulation are connected to the unique naming convention with the COMPOSITION dictionary in the README.yaml file.
-Upon addition of a new entry in the databank, the molecule names in the simulation corresponding the unique names and mapping files (available at \url{https://github.com/NMRLipids/Databank/tree/main/Scripts/BuildDatabank/mapping_files}) are defined in the dictionary.
-The numbers of molecules in the system are then automatically calculated by the NMRlipids databank codes (AddData.py in Fig.~\ref{DatabankStructure}) and stored in the README.yaml together with other content of the COMPOSITION dictionary.
-This information can be then used to find each molecule and atom from each simulation in the analysis codes.
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{c|c}
-        Abbreviation & Molecule name \\
-        \hline
-        POPC &  1-palmitoyl-2-oleoyl-sn-glycero-3-phosphocholine\\
-        POPG &  1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoglycerol \\
-        POPS & 1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-L-serine \\
-        POPE & 1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoethanolamine \\
-        CHOL & cholesterol \\
-        DHMDMAB & dihexadecyldimethylammonium \\
-        \hline
-        POT & potassium ion \\
-        SOD & sodium ion \\
-        CLA & chloride ion \\
-        CAL & calcium ion \\
-        SOL & water \\
-    \end{tabular}
-    \caption{Abbreviations used in the databank}
-    \label{tab:abbreviations}
-\end{table}
-
-\subsection{Quality evaluation}
-
-The quality of conformational ensembles of individual lipid molecules in simulations are evaluated using the C-H bond order parameters which can be directly compared with robust experimental data~\cite{ollila16}.
-%The C--H bond order parameters were calculated directly
-%from the carbon and hydrogen positions using the 
-%definition
-The order parameters are defined as
-\begin{equation}\label{OP}
-S_{\rm CH}=\frac{1}{2}\left\langle 3\cos^2\theta -1 \right\rangle,
-\end{equation}
-where $\theta$ is the angle between the C--H bond and the membrane normal
-%(taken to align with $z$, with bilayer periodicity in the $xy$-plane).
-and angular brackets denote the ensemble average, i.e. average over all sampled configurations of all lipids in a simulation.
-%by estimating
-%the average probability of C-H bond order parameters to agree with the values from NMR experiments. 
-%As a first step, the quality of each order parameter is evaluated using the negative logarithm of probability to find the simulated order parameter within the experimental error bars
-%\begin{equation}
-%    S_q = -\log_{10}(P),    
-%\end{equation}
-Because conformational ensembles of individual lipids are independent in a simulation of a fluid lipid bilayer, the number of independent sample points for each C-H bond, $n$, equals the number of lipids in a simulation.
-Denoting the $S_{\rm CH}$ from Eq.~\ref{OP} as the sample mean and $s$ its variance calculated over individual lipids,
-%the probability the true mean of the order parameter is given by the %Student's t-distribution 
-%Because each C-H bond order parameter is calculated as an average over individual lipids in a simulation
-%\begin{equation}\label{studentT}
-$\frac{S_{\rm CH}-\mu}{s/\sqrt{n}}$
-%\end{equation}
-%where $\bar{X}$ is the sample mean (order parameter calculated from simulations),
-%$s$ is the sample variance (variance of order parameters in different lipids)
-%and 
-%has a 
-has a Student's t-distribution with $n-1$ degrees of freedom, where $\mu$ is the real mean of the order parameter. 
-%shifted by $S_{\rm CH}$ and scaled by $s/\sqrt{n}$.
-Because lipid bilayer simulations contain at least dozens of lipids, the Student's t-distribution could be safely approximated with a normal distribution.
-However, when calculating the probability for a simulation to locate within experimental order parameters, the normal distribution gives values below the numerical accuracy of computers for simulations values far from experiments.
-To avoid such numerical instability, we use the first order Student's t-distribution.% for the equation \ref{studentT}.
-Therefore, we estimate the probability for the simulation order parameter to locate within experimental error bars from equation
-\begin{equation}\label{probability}
-  P = f \left( \frac{S_{\rm CH} - (S_{\rm exp}+\Delta S_{\rm exp})}{s/\sqrt{n}} \right) - f \left( \frac{S_{\rm CH} - (S_{\rm exp}-\Delta S_{\rm exp})}{s/\sqrt{n}} \right),
-\end{equation}
-where f(t) is the first order Student's t-distribution.
-% shifted to $\bar{X}$ and scaled with $s$.
-
-%We estimate the probability for the simulation order parameter to locate within experimental
-%error bars from equation
-%\begin{equation}\label{probability}
-%  P = T(S_{\rm exp}+\Delta S_{\rm exp}) - T(S_{\rm exp}-\Delta S_{\rm exp}),
-%\end{equation}
-%where T is the first order Student's t-distribution shifted to $\bar{X}$ and
-%scaled with $s$. 
-%where the probability is calculated from the normal distribution
-%\begin{equation}\label{gaussian}
-%    P = \int_{S_{\rm exp}-\Delta S_{\rm exp}}^{S_{\rm exp}+
-%    \Delta S_{\rm exp}}  \frac{1}{\sigma \sqrt{2\pi}} e^{-\frac{1}{2}(\frac{x-\mu}{\sigma})^2} {\rm d}x.
-%\end{equation}
-%$S_{\rm exp}$ and $\Delta S_{\rm exp}$ are experimental order parameter and its error, and $\mu$ and $\sigma$ are the mean order parameter and its standard deviation from simulations.
-The quality measure $S_q$ approaches to zero when probability for agreement between simulation and experimental results approach one, and increases when simulated and experimental values diverge.
-The error of $\Delta S_{\rm exp} = 0.02$ is currently assumed for all experimental order parameters~\cite{ollila16}.
-Because phospholipids sample their conformational ensemble within nanosecond timescale~\cite{ferreira15}, all simulations in the databank would be sufficiently long to sample the realistic conformational phase of individual lipids.
-However, some force fields exhibit too slow dynamics which leads to large error bars in order parameter values~\cite{antila21a}.
-Because large error bars widen the Student's t-distribution in Eq.~\ref{probability} thereby artificially increasing the probability to find the simulated value within experimental error bars, the order parameters with error bars larger than the experimental error 0.02 are not included in the quality evaluation.
-
-The qualities of different fragments within each lipid type in a simulation, $P^{\rm frag}[{\rm lipid}]$, are then estimated by averaging the probabilities for individual C-H bond order parameters to locate within experimental errors according to Eq.~\ref{probability}, and dividing the average with the percentage of order parameters for which the quality is available within the fragment, $p_{\rm frag}[{\rm lipid}]$,
-\begin{equation}
-    P^{\rm frag}[{\rm lipid}] = \frac{\langle P[{\rm lipid}]\rangle_{\rm frag}}{p_{\rm frag}[{\rm lipid}]},
-\end{equation}
- where frag can be {\it sn}-1, {\it sn}-2, headgroup or total (all order parameters within a molecule).
-The overall quality of different fragments in a simulation are then defined as a molar fraction weighted average over different lipid components
-\begin{equation}
-    P^{\rm frag} = \sum_{\rm lipid} \chi_{\rm lipid} P^{\rm frag}[{\rm lipid}],
-\end{equation}
-where $\chi_{\rm lipid}$ is the molar fraction of a lipid in the bilayer.
-
-Qualities of form factors were evaluated using the same approach as in SIMtoEXP program~\cite{kucerka10}.
-Because experiments give form factors only in relative intensity scale, they were scaled to the simulation data using equation
-\begin{equation}
-    k_e = \frac{\sum_{i=1}^{N_q} \frac{|F_s(q_i)||F_e(q_i)|}{(\Delta F_e(q_i))^2}}{\sum_{i=1}^{N_q} \frac{|F_e(q_i)|^2}{(\Delta F_e(q_i))^2}}.
-\end{equation}
-The quality of each form factor was then calculated from the equation
-\begin{equation}
-    \chi^2 = \frac{\sqrt{\sum_{i=1}^{N_q}(|F_s(q_i)|-k_e|F_e(q_i)|)^2/(\Delta F_e(q_i))^2}}{\sqrt{N_q-1}},
-\end{equation}
-where $F_s$ and $F_e$ are form factors from a simulation and experiment, respectively, and summation goes over the experimentally available $N_q$ points.
-
-\subsection{Analysing simulations in the NMRlipids databank}
-
-The README.yaml files contain all the essential information to perform arbitrary analyses of simulations in the NMRlipids databank, i.e., the permanent location of the original data and naming convention for all atoms and molecules in each system.
-In practise, the analyse codes contains a loop over all README.yaml files (i.e., simulations in the NMRlipids databank) which first downloads the raw simulation to a local computer and then uses the information about the atom and molecule naming conventions in README.yaml and mapping files to perform the desired analyses.
-For example, the code that calculates all C-H bond order parameters of all systems is available at \url{https://github.com/NMRLipids/Databank/blob/main/Scripts/AnalyzeDatabank/calcOrderParameters.py} and minimal example of a analysis code is available at \url{https://github.com/NMRLipids/Databank/blob/main/Scripts/AnalyzeDatabank/template.ipynb}.
-
-While the order parameters, form factors, area per lipid and thickness are stored within the NMRlipids databank (\url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations}), further analyses can be conventiently stored in separate repositories with the same folder structure based on hash identities of trajectory and topology files.
-For example, results from further analyses performed here are stored in folders at \url{https://github.com/NMRLipids/DataBankManuscript/tree/main/Data}.
-Such organization of the data enables further upcycling of the analyzed data as similarly to the original NMRlipids databank repository.
-
-For the permeation of water through membranes, the number of permeation events in each trajectory was first calculated using the code by Camilo et al.~\cite{camilo2022}, available at~\url{https://github.com/crobertocamilo/MD-permeation}.
-The permeation was then calculated from equation $P=r/2c_w$, where $r$ is the rate per time and area, and $c_w$ is the concentration of water in bulk with the value of 33.3679\,(nm)$^{-3}$~\cite{venable19}.
-
-%\subsection{Indexing the simulation data}
-
-%\subsubsection*{AddData.py}
-%AddData.py is a script that builds a database that contains a dictionary file and analysis data of each simulation.
-The dictionary file contains information about the simulation.
-The script also calculates order parameters of all CH bonds of the lipids in the simulation.
-To add a simulation it must be first uploaded to Zenodo (www.zenodo.org).
-The trajectory and topology files of the simulation are downloaded to the working directory from Zenodo but these are not saved into the database.
-To add a simulation to the database the user has to give some essential information about the simulation.
-This is done by writing a info file (*INFO.yaml) which is passed to AddData.py.
-%\newline \\
-%AddData.py requires GROMACS and MDAnalysis library to be installed.
-%\newline \\
-%Run AddData.py as follows:
-%\newline \\
-%python3 AddData.py -f exampleINFO.yaml
-%\newline \\
-%There is also a script runAddData.sh that can be used to loop over several info files to add many simulations at one go.
-
-%\subsubsection*{Simulation dictionary}
-%A simulation dictionary contains information about the simulation.
-Some of the information is provided by the user.
-The numbers of lipid molecules, solvent and ions are automatically read from the files and so are the simulation temperature and trajectory length.
-All this information is saved to a file named README.yaml.
-
-%\subsubsection{User input}
-%Dictionary variables defined by the user are listed here. 
-%The necessary parameters required for the analyses are described first and are marked as ''compulsory''. Then the parameters used to describe the data for further usage and analyses are described.
-%While only compulsory parameters are required for the databank entry, strong recommendation is that all the possible parameters would be given to maximize the possibilities to upcycle the contributed data.
-
-%\subsubsection*{DOI (compulsory)}
-%Give DOI identity from where the simulation files are located.
-Current databank works only for the data in Zenodo, but other potential sources are may be implemented in the future.
-Note that the DOI must point to a specific version of dataset in Zenodo, i.e., DOI pointing to all versions of certain dataset does not work.
-
-%\subsubsection*{SOFTWARE (compulsory)}
-%Give the name of software used to run the simulation. The options are GROMACS, AMBER, NAMD, CHARMM and OPENMM. So far, only simulations run with GROMACS are accepted by the script.
-
-%\subsubsection*{TRJ (compulsory)}
-%Give the name of the trajectory file that is found from the DOI given above.
-
-%\subsubsection*{TPR (compulsory)}
-%Give the name of the file with topology information (tpr file in the case of Gromacs) that is found from the DOI given above.
-
-%\subsubsection*{PREEQTIME (compulsory)}
-%Give the time simulated before the uploaded trajectory in nanoseconds. For example, if you upload 100-200~ns part of total 200~ns simulation, this should value should be 100.
-
-%\subsubsection*{TIMELEFTOUT (compulsory)}
-%Give the time that should be considered as an equilibration period in the uploaded trajectory.
-Frames before the give time will be discarded in the analysis.
-For example, if you upload 0-200~ns part of total 200~ns simulation where the first 100~ns should be considered as an equilibration, this value should be 100.
-
-%\subsubsection*{Molecule names (compulsory)}
-%In the databank, each molecule has a unique abbreviation (see table~\ref{tab:abbreviations}). 
-%You need to give the residue names of these molecules corresponding your simulation.
-%The user must also provide the names of the molecules, ions and solvent that are used in the simulation to match the names used by the databank.
-The names provided by the user must be the same as in the tpr file.
-%If atoms in a lipid belong to different residues (typical situation in Amber force fields), give the name of the head group residue here, and add the residue name of each atom to the third column in the mapping file (see below).
-If your simulation contains molecules that are not yet in the databank, you need to define the abbreviation and add molecules to the lipids\_dict, molecules\_dict, molecule\_numbers\_dict and molecule\_ff\_dict in the AddData.py script, as well as to table~\ref{tab:abbreviations}.
-
-%\subsubsection*{MAPPING\_DICT (compulsory)}
-%For the analysis, we need to know the names of atoms in your system.
-%These are defined using the mapping file convention introduced in \url{http://nmrlipids.blogspot.com/2015/03/mapping-scheme-for-lipid-atom-names-for.html}, where first column gives the universal atom name, second gives the atom name in force field, and third gives the residue name if not the same for all atoms.
-%The name of the mapping file for each molecule needs to be given in dictionary format.
-The dictionary the key is the molecule name (abbreviation listed in table \ref{tab:abbreviations}) and the value is the name of the mapping file.
-The already existing mapping files can be found from the directory named "mapping\_files".
-If a mapping file for your molecule(s) do not exist, you need to construct one and add to the "mapping\_files" directory.
-
-%\newline \\
-%The purpose of a mapping file is to circumvent the problem caused by different atom naming conventions used by different force fields.
-The first column of a mapping file contains general atom names.
-The second column contains the name of the atom as it is in the force field.
-If the lipid consists of several residues which is the case in some AMBER force fields, then a third column is needed which contains the name of the residue to which each atom belongs to.
-
-%\subsubsection*{DIR\_WRK (compulsory)}
-%Give the path of the working directory in your local computer.
-The trajectory and topology files will be downloaded to this trajectory, and temporary files created during processing will be stored here.
-
-%\subsubsection*{UNITEDATOM\_DICT (compulsory for united atom trajectories)}
-%Order parameters from united atom simulations are calculated using buildH code (\url{https://github.com/patrickfuchs/buildH}).
-For united atom simulations, you need to tell how hydrogens are added based on definitions in the dic\_lipids.py dictionary.
-This is done by giving a dictionary where the key is the molecule name (abbreviation listed in table \ref{tab:abbreviations}) and the value is the correct dictionary key in dic\_lipids.py.
-If correct dictionary key is not yet found, you need to add to dic\_lipids.py.
-%In the case of an all atom simulation, UNITEDATOM is left empty.
-
-%\subsubsection*{PUBLICATION}
-%Give reference to a publication(s) related to the data.
-
-%\subsubsection*{AUTHORS\_CONTACT}
-%Give the name and email of the main author(s) of the data.
-
-%\subsubsection*{SYSTEM}
-%Give description of system in free format. For example ''POPC with cholesterol at 301K''.
-
-%\subsubsection*{SOFTWARE\_VERSION}
-%Give the version of the software used.
-
-%\subsubsection*{FF}
-%Give the name of the force field used used in the simulation.
-
-%\subsubsection*{FF\_SOURCE}
-%Describe the source of the force field parameters. For example, CHARMM-GUI, link to webpage where parameters were downloaded, or citation to a paper.
-
-%\subsubsection*{FF\_DATE}
-%Give the date when parameters were accessed or created. The format is day/month/year.
-
-%%\subsubsection*{Individual force field names for molecules}
-%In some cases special force fields are used for certain molecules.
-For example, non-standard parameters for ions or other molecules have been used.
-These can be specified giving forcefield names separately for individual molecules.
-These can be given as parameters named as
-%FF+[abbreviation from table \ref{tab:abbreviations}], i.e., FFPOPC, FFPOT, FFSOL etc.
-
-%\subsubsection*{CPT (Gromacs)}
-%Give the name of the Gromacs checkpoint file that is found from the DOI given above.
-%CPT stands for the name of the cpt file.
-
-%\subsubsection*{LOG (Gromacs)}
-%Give the name of the Gromacs log file that is found from the DOI given above.
-
-%\subsubsection*{TOP (Gromacs)}
-%Give the name of the Gromacs top file that is found from the DOI given above.
-
-%\subsubsection{Automatically analyzed parameters}
-%The following parameters are read automatically from the trajectory and topology files.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%automatically analyzed parameters%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\subsubsection*{Molecule numbers}
-%Numbers of lipid molecules (NPOPC, NPOPG, etc.) per membrane leaflet are calculated by determining on which side of the center of mass of the membrane the center of mass of the head group of each lipid molecule is located.
-%\newline \\
-%\noindent Numbers of other molecules such as solvent and ions (NSOL, NPOT, NSOD, etc.) are read from the topology file.
-
-%\subsubsection*{Temperature}
-%Temperature of the simulation is read from the topology file.
-
-%\subsubsection*{Trajectory length}
-%The length of a trajectory is read from the trajectory file.
-
-%\subsubsection*{Date of running}
-%The date of running the script is saved to the README.yaml file.
 
@@
 
-\section*{Acknowledgements}
+%\section*{Acknowledgements}
 
@@
 
-\section*{Author contributions statement}
+%\section*{Author contributions statement}
 
-Must include all authors, identified by initials, for example:
-A.A. conceived the experiment(s),  A.A. and B.A. conducted the experiment(s), C.A. and D.A. analysed the results.  All authors reviewed the manuscript.
+%Must include all authors, identified by initials, for example:
+%A.A. conceived the experiment(s),  A.A. and B.A. conducted the experiment(s), C.A. and D.A. analysed the results.  All authors reviewed the manuscript.
 
-\section*{Additional information}
+%\section*{Additional information}
 

### 2022-08-31 main.tex
Updates from Overleaf

@@
 The values for water permeabilities from simulations in the databank vary between 0.3 and 322\,$\mu$m/s with the mean and median of 14\,$\mu$m/s and 8\,$\mu$m/s, respectively.
-These values are within the range of experimentally reported values and the mean is the same order of magnitude as the diffusive permeability which is 1-2 orders of magnitude slower than osmotic permeability \cite{jansen95,??}.
+These values have the same order of magnitude as the experimentally determined diffusive permeability coefficients, but are on average below the values reported for PC lipids in liquid crystalline phase,~190-330\,$\mu$m/s~\cite{jansen95}.
+%within the range of experimentally reported values  permeability which is 1-2 orders of magnitude slower than osmotic permeability \cite{jansen95,??}. 
 Clear dependencies of permeability on hydration level, charged lipid fraction, cholesterol fraction or POPE fraction were not observed, although weak decrease for the latter two may be visible.
@@
     \includegraphics[width=180mm]{Figures/permeation2.pdf}
-    \caption{Water permeation through lipid membranes analyzed from the databank as a function of 
-      a) area per lipid,
-      b) thickness,
-      and c) temperature. 
-      Inserts in a) and b) show the region where the dependence could be considered approximately linear.
-      Insert in c) shows the Arrhenius plot of permeation ($\ln$ (P) vs. 1/T) that gives 17\,$\pm$\,4\,k$_b$T for the average activation energy for water permeation through lipid bilayer.}
+    \caption{(b-c) Water permeation through lipid membranes analyzed from the databank as a function of temperature, thickness, area per lipid, and acyl chain order. 
+      Inserts in c) and d) show the region where the dependence could be considered approximately linear.
+      Insert in a) shows the Arrhenius plot of permeation ($\ln$ (P) vs. 1/T) that gives 17\,$\pm$\,4\,k$_b$T for the average activation energy for water permeation through lipid bilayer.
+      (f-g) Diffusion anistoropy of water as a function of thickness and area per lipid.
+      (h) Lateral diffusion of water as a function of hydration level. Experimental points for DMPC bilayers at 313\,K at different hydration levels are shown~\cite{rudakova04}.}
     \label{fig:permeability}
@@
 To analyze the anisotropic water diffusion, we also calculated the water diffusion along the membrane.
-The results at different hydration levels are shown in Fig.~\ref{fig:permeability} e) together with the experimental data~\cite{??}.
-The experimental water diffusion increases with increasing hydration level toward the value for bulk water (??
-in ??K).
+The results at different hydration levels are shown in Fig.~\ref{fig:permeability} h) together with the experimental data~\cite{rudakova04}.
+The experimental water diffusion increases with increasing hydration level toward the value for bulk water (3.1$\cdot$\,10$^{-9}$\,m$^2$/s at 313\,K)\cite{khakimov08}.
 Simulations overestimate the experimental bulk value approximately with the factor of 2 at high hydration levels which is not surprising as the most common water model used in membrane simulations, TIP3P, overestimates the bulk water diffusion.
 However, simulation results are closer to experiments with low hydration levels.
-To estimate the diffusion anistoropy, we translated the water permation to diffusion coefficent through multilamellar stack using the Tanner equation~\cite{??}.
-The results in
+To estimate the diffusion anistoropy, we translated the water permeation to diffusion coefficient through multilamellar stack using the Tanner equation~\cite{??}.
+The resulting diffusion coefficients trough the membranes are approximately five orders of magnitude slower than along the membrane which is at the upper limits of the anisotropic estimated from the experimental data~\cite{nitsche19}.
+Relatively high anisotropy is understandable as simulations give slightly slower permeation rates and higher lateral diffusion rates than experiments.
+Significant dependence of diffusion anisotropy on membrane thickness and area per lipid are observed in Figs.~\ref{fig:permeability} f) and g).
+The anisotropy becomes linearly stronger with decreasing area per lipid while the dependency is stronger with increasing thickness.
+These results can be explained by the increasing and decreasing permeabilities with thickness and are per lipid (Figs \ref{fig:permeability} c) and d), respectively, while lateral diffusion remains approximately constant (Fig.
+\ref{??} a) and c)).
 

### 2022-09-07 SI.tex
Update manuscript

@@
 \usepackage{tablefootnote}
+
+\makeatletter
+\newcommand*{\centerfloat}{%
+  \parindent \z@
+  \leftskip \z@ \@plus 1fil \@minus \textwidth
+  \rightskip\leftskip
+  \parfillskip \z@skip}
+\makeatother
 
@@
 %\noindent Please note: Abbreviations should be introduced at the first mention in the main text – no abbreviations lists. Suggested structure of main text (not enforced) is provided below.
+
+\pagebreak
+\section{Correlations between area per lipid and thickness with order parameters and form factors}
+
+\begin{figure}[!h]
+    \centerfloat
+    \includegraphics[width = 180mm]{Figures/QualityCorrelationsSI.pdf}
+    \caption{Dependence of form factor, electron density profiles along membrane normal, and C-H bond orderparameters (from top to bottom) on simulation box size with different cholesterol concentration.
+}
+    \label{fig:sizedependence}
+\end{figure}
+
+\pagebreak
+\section{Dependence of form factor and order parameters in the size of simulation}
+
+\begin{figure}[!h]
+    \centerfloat
+    \includegraphics[width = 240mm]{Figures/SizeDependence.pdf}
+    \caption{Dependence of form factor, electron density profiles along membrane normal, and C-H bond orderparameters (from top to bottom) on simulation box size with different cholesterol concentration.
+}
+    \label{fig:sizedependence}
+\end{figure}
 

### 2022-09-07 main.tex
Update manuscript

@@
     \centering
-    \includegraphics[width=180mm]{Figures/quality.pdf}
+    \includegraphics[width=180mm]{Figures/quality3.pdf}
     \caption{a) Correlations between membrane area per lipid, thickness, second minima of form factor and average order parameter of the {\it sn}-1 acyl chain extracted from the NMRlipids databank.
@@
 
+\subsubsection{Conformational ensembles using C-H bond order parameters}
 The quality of conformational ensembles of individual lipid molecules in simulations are evaluated using the C-H bond order parameters which can be directly compared with robust experimental data~\cite{ollila16}.
@@
 
-Qualities of form factors were evaluated using the same approach as in SIMtoEXP program~\cite{kucerka10}.
-Because experiments give form factors only in relative intensity scale, they were scaled to the simulation data using equation
+\subsubsection{Membrane dimensions with x-ray scattering form factors}
+While C-H bond order parameters relate to the conformational ensembles of individual lipids,
+the x-ray scattering form factors depend on membrane dimensions and density distribution~\cite{ollila16}.
+Because experiments give form factors only in relative intensity scale, they are typically scaled to the simulation data.
+Here we use the same scaling method as in SIMtoEXP program~\cite{kucerka10} where experimental form factor intensities
+are scaled by a factor defined as
 \begin{equation}
-    k_e = \frac{\sum_{i=1}^{N_q} \frac{|F_s(q_i)||F_e(q_i)|}{(\Delta F_e(q_i))^2}}{\sum_{i=1}^{N_q} \frac{|F_e(q_i)|^2}{(\Delta F_e(q_i))^2}}.
+    k_e = \frac{\sum_{i=1}^{N_q} \frac{|F_s(q_i)||F_e(q_i)|}{(\Delta F_e(q_i))^2}}{\sum_{i=1}^{N_q} \frac{|F_e(q_i)|^2}{(\Delta F_e(q_i))^2}},
 \end{equation}
-The quality of each form factor was then calculated from the equation
-\begin{equation}
-    \chi^2 = \frac{\sqrt{\sum_{i=1}^{N_q}(|F_s(q_i)|-k_e|F_e(q_i)|)^2/(\Delta F_e(q_i))^2}}{\sqrt{N_q-1}},
-\end{equation}
+%The quality of each form factor was then calculated from the equation
+%\begin{equation}
+%    \chi^2 = \frac{\sqrt{\sum_{i=1}^{N_q}(|F_s(q_i)|-k_e|F_e(q_i)|)^2/(\Delta F_e(q_i))^2}}{\sqrt{N_q-1}},
+%\end{equation}
 where $F_s$ and $F_e$ are form factors from a simulation and experiment, respectively, and summation goes over the experimentally available $N_q$ points. 
+Quality measure based on differences in simulated and experimental form factors accross available q-range is also defined in the SIMtoEXP program.
+However, as shown in Fig.~\ref{??}, the lobe heigths in the simulated form factors depend on the simulation box size,
+but locations of minima does not. Therefore, quality measure depending on lobe heights is also dependent on simulation box size.
+In order to define a quality measure which is independent on simulation box size, we measure the form factor quality based on the
+location of the first minima observed in experiments. As shown in Fig.~\ref{??}, this minima correlates well with the thickness of a membrane.
+Although the correlation in second minima would be even stronger, we have chosen to use the first minima because automatic definition
+of second minima is inaccurate in some experimental data. In practise, we first filter the fluctuations from the form factor data using
+Savitzky-Golay filter (window lenght 30 and polynomial order 1) and locate the first minima above 0.1\,$AA^2$ from both simulation and experimental data.
+The quality of a form factor is then defined as Eucledian distance between the minima in simulated and experimental form factors, $FF_q = |FF_{\rm{min}}^{\rm{sim}}-FF_{\rm{min}}^{\rm{exp}}|$.
 

### 2022-09-09 main.tex
Updates from Overleaf

@@
 \usepackage{tablefootnote}
+\usepackage{xr}
+
+\makeatletter
+\newcommand*{\addFileDependency}[1]{% argument=file name and extension
+  \typeout{(#1)}
+  \@addtofilelist{#1}
+  \IfFileExists{#1}{}{\typeout{No file #1.}}
+}
+\makeatother
+
+\newcommand*{\myexternaldocument}[1]{%
+    \externaldocument{#1}%
+    \addFileDependency{#1.tex}%
+    \addFileDependency{#1.aux}%
+}
+
+\myexternaldocument{SI}
 
@@
 \subsection{Design of the NMRlipids databank}
+The practical challenges related to the distribution of large MD simulation trajectories are solved in the NMRlipids databank using the overlay structure. 
 The key idea of the overlay databank is that the storage of raw data in layer 1 is distributed in publicly available repositories or other servers with long term stability and permanent links such as digital object identifiers.
@@
 Quality of membrane simulations with different force fields have been evaluated against experimental data during parameterization and in separate comparison studies~\cite{botan15,ollila16,catte16,pluhackova16,perez17,leonard19}, but universal quality measure for membrane simulations is not defined and controversial results are often reported from simulations~\cite{antila22b}.
-The unclarities in simulation quality complicate the selection of proper simulation models for specific applications and estimation of reliability of reported simulation results, thereby being a major obstacle in many applications of membrane MD simulations.
+The ambiguity in defining the quality of simulations complicates the selection of proper simulation parameters and estimation of the reliability of simulation results, thereby being a major obstacle in many applications of membrane MD simulations.
+The NMRlipids
+databank enables rapid quality evaluation of MD simulations using the quantitative quality measure for membrane simulations.
+The conformational ensembles of individual lipid molecules are evaluated against C-H bond order parameters from NMR experiments, which is complement by x-ray scattering form factors arising from overall membrane dimensions via electron density profile.
+Both of these are robust experimental measurables that can be directly connected to the simulation data~\cite{ollila16}.
 
-To enable rapid evaluation and comparison of membrane MD simulation qualities, we define the quality measure for membrane MD simulations using the C-H bond order parameters from NMR experiments and form factors from x-ray scattering, which are robust experimental measurables that can be directly connected to the simulation data~\cite{ollila16}.
-C-H bond order parameters are related to the conformational ensemble of individual lipid molecules, while the form factor to arises from electron density thereby connecting to the overall structure of lipid bilayers.
-This is demonstrated in Fig.~\ref{fig:quality} a) showing correlations between membrane lateral density (area per lipid), thickness, minima of form factor and acyl chain order analyzed from all apporimately 500 trajectories in the NMRlipids databank.
+%NMRlipids databank contains also experimental order parameter and x-ray form factor data that is connected to corresponding  simulations in order to define the simulation qualities and rank them to select the suitable simulation models for particular applications.
+The top quality simulations based on average order parameter quality are shown in Fig.~\ref{fig:quality} A and the direct comparison between the best simulation (POPC with OPLS3e) and experiments in shown in Fig.~\ref{fig:quality} D.
+This model reproduces the experimental data very well with the exception of third carbon in the $\it{sn}$-1 chain and carbons 8 and 9 close to double bond in $\it{sn}$-2 chain.
+%Full details of the NMRlipids quality measure for lipid bilayer simulations are given in the methods section and supplementary information.
+The top ranking simulation for the order parameter quality of POPE lipid are shown in Fig.~\ref{fig:quality} B.
+In this case the best overall quality is found in GROMOS-CKP simulation (direct comparison shown in Fig.~\ref{fig:quality} E, but clear discrepancies with experiments are present in this model for glycerol backbone and $\it{sn}$-2 chain.
+On the other hand, glycerol backbone and headgroup for POPE is best described in CHARMM36 model (direct comparison shown in Fig.~\ref{fig:quality} F), but its overall quality suffers from overestimated order in acyl chains.
+Scatter plots and Pearson correlation coefficients in Fig.~\ref{fig:quality} C reveal strong correlation between membrane area per lipid and thickness.
+The area per lipid is also strongly correlated with the average order of $\it{sn}$-1 chain, suggesting that its order parameters can used as a proxy for the area per lipid in addition to the conformational ensemble.
+On the other hand, the first form factor minima is correlated stronger with membrane thickness (Fig.~\ref{fig:quality} C) than with the area per lipid (Fig.~\ref{fig:QualityCorrelationsSI}).
+
+ %This is demonstrated in Fig.~\ref{fig:quality} a) showing correlations between membrane lateral density (area per lipid), thickness, minima of form factor and acyl chain order analyzed from all apporimately 500 trajectories in the NMRlipids databank.
 Membrane area per lipid has negative correlations with membrane order and thickness with Pearson coefficient of -0.78 and -0.49, respectively, because decreased area leads to more ordered, tightly packed and stretched acyl chains leading to thicker membranes.
@@
     \includegraphics[width=180mm]{Figures/quality3.pdf}
-    \caption{a) Correlations between membrane area per lipid, thickness, second minima of form factor and average order parameter of the {\it sn}-1 acyl chain extracted from the NMRlipids databank.
-All Pearson correlation coefficients shown in bottom left corner have p-value below 0.001.
-    b) Evaluation against experimental data exemplified for the simulation giving the best qualities in overall, for headgroup and x-ray scattering form factor.
-    c) Change in area per lipid upon addition of POPE to POPC in different simulation extracted from the NMRlipids databank and quality evaluation table of POPE simulations.
-    d) Quality evaluation table showing the best 13 simulations according to the overall order parameter quality.}
+    \caption{ A) Best simulations ranked based on overall order parameter quality.
+    B) Best simulations ranked based on the overall order parameter quality of POPE lipid. 
+    C) Scatter plots and Pearson correlation coefficients for the area per lipid, thickness, and experimental observables. 
+    %second minima of form factor and average order parameter of the {\it sn}-1 acyl chain extracted from the NMRlipids databank. 
+    All correlation coefficients have p-value below 0.001. For more correlations see Fig.~\ref{fig:QualityCorrelationsSI}.
+    D)-F) Evaluation against experimental data exemplified for a simulation with the best overall order parameter quality (D), the best quality for POPE lipid (E), and the headgroup quality for POPE (F).
+    %c) Change in area per lipid upon addition of POPE to POPC in different simulation extracted from the NMRlipids databank and quality evaluation table of POPE simulations.
+    }
     \label{fig:quality}
 \end{figure}
-
-NMRlipids databank contains also experimental order parameter and x-ray form factor data that is connected to corresponding  simulations in order to define the simulation qualities and rank them to select the suitable simulation models for particular applications.
-The ranking of all simulations based on estimated average probability of order parameters to locate within experimental errors are shown in Fig.~\ref{fig:quality} b) and the comparison for the best models for overall quality, headgroup and glycerol backbone, and form factor are exemplified in Fig.~\ref{fig:quality} d).
-Full details of the NMRlipids quality measure for lipid bilayer simulations are given in the methods section and supplementary information.
-The top ranking simulations in Figs.~\ref{fig:quality} b) and d) demonstrate also the current complexity in lipid bilayer simulation quality.
-Simulation of POPC lipid bilayer with Slipids force field ranks first in overall score, glycerol backbone is not correcetly captured in this model.
-On the othere, simulation with the best quality for headroup and glycerol backbone, CHARMM36 POPC, predicts too ordered acyl chain, thereby not being within the top 10 simulations in the overall ranking.
-The best ranking for x-ray form factor is given by the Slipids POPE simulation, however, because the quantitative value for form factor quality depends on the quality of experimental data, form factor qualities can be compared only between simulations evaluated against the same experimental dataset.
 
@@
 Quality measure based on differences in simulated and experimental form factors accross available q-range is also defined in the SIMtoEXP program.
-However, as shown in Fig.~\ref{??}, the lobe heigths in the simulated form factors depend on the simulation box size,
-but locations of minima does not. Therefore, quality measure depending on lobe heights is also dependent on simulation box size.
-In order to define a quality measure which is independent on simulation box size, we measure the form factor quality based on the
-location of the first minima observed in experiments. As shown in Fig.~\ref{??}, this minima correlates well with the thickness of a membrane.
-Although the correlation in second minima would be even stronger, we have chosen to use the first minima because automatic definition
-of second minima is inaccurate in some experimental data. In practise, we first filter the fluctuations from the form factor data using
-Savitzky-Golay filter (window lenght 30 and polynomial order 1) and locate the first minima above 0.1\,$AA^2$ from both simulation and experimental data.
+However, as shown in Fig.~\ref{fig:sizedependence}, the lobe heigths in the simulated form factors depend on the simulation box size, but locations of minima does not.
+Therefore, quality measure depending on lobe heights is also dependent on simulation box size.
+In order to define a quality measure which is independent on simulation box size, we measure the form factor quality based on the location of the first minima observed in experiments.
+As shown in Fig.~\ref{fig:quality}, this minima correlates well with the thickness of a membrane.
+Although the correlation in second minima would be even stronger (Fig.~\ref{fig:QualityCorrelationsSI}), we have chosen to use the first minima because automatic definition of second minima is inaccurate in some experimental data.
+In practise, we first filter the fluctuations from the form factor data using Savitzky-Golay filter (window lenght 30 and polynomial order 1) and locate the first minima above 0.1\,$AA^2$ from both simulation and experimental data.
 The quality of a form factor is then defined as Eucledian distance between the minima in simulated and experimental form factors, $FF_q = |FF_{\rm{min}}^{\rm{sim}}-FF_{\rm{min}}^{\rm{exp}}|$.

### 2022-10-05 main.tex
Update text

@@
 Here we present a solution for lipid bilayers based on overlay databank structure illustrated in Fig.~\ref{fig:overlay}.
-The concept of overaly databank is developed here to solve the practical challenges in generating databanks of MD simulation data enabling flexible analyses over large sets of simulation data, but it pontentially used for wide range of situations, particularly when storage of raw data requires significant resources and final outcomes or best practices are not yet clear, overlay databank approach lowers the barrier to start without compromizing the long term stability or scalability.
+The concept of overlay databank is developed here to solve the practical challenges in generating databanks of MD simulation data enabling flexible analyses over large sets of simulation data, but it can be  potentially used for wide range of situations.
+Particularly when storage of raw data requires significant resources and final outcomes or best practices are not yet clear, overlay databank approach lowers the barrier to start without compromizing the long term stability or scalability.
 
@@
 %Up to three levels of \textbf{subheading} are permitted. Subheadings should not be numbered.
-\subsection{Design of the NMRlipids databank}
-The practical challenges related to the distribution of large MD simulation trajectories are solved in the NMRlipids databank using the overlay structure. 
-The key idea of the overlay databank is that the storage of raw data in layer 1 is distributed in publicly available repositories or other servers with long term stability and permanent links such as digital object identifiers.
-The core of the databank, layer 2, contains only information on the location and content of the raw data, thereby not requiring large resources to handle and maintain.
-This lowers the barrier for starting such databank as well as for long term storage.
-The NMRlipids databank is essentially a git repository containing information on the location of raw data and its indexing with universal naming convention.
-In addition to all computers where the databank is developed and used, the NMRlipids databank git is stored to Zenodo server, thereby enabling a very cost effective long term storage for the databank.
-The databank can be used in layer 3 by accessing the raw data and information stored in the databank by employing the universal naming conventions for atoms, molecules and simulation details.
+\subsection{Overlay structure of the NMRlipids databank}
+The three layer structure of the NMRlipids databank, illustrated in Fig.~\ref{fig:overlay} a, enables an efficient way to share and upcycle large MD simulation trajectories.
+Layer 1 contains raw simulation data that can locate in any publicly available repository or server offering long term stability and permanent links to the data, such as digital object identifiers.
+Layer 2 contains all the relevant information on the simulations, including links to the raw data in layer 1 and relevant metadata describing the systems.
+Layer 3 is the application layer containing outputs from the databank, such as graphical user interface (\url{http://www.databank.nmrlipids.fi/}) and results from analyses desribed in sections below.
+
+Layer 2 is the core of the databank.
+In addition to the metadata and links to raw simulation data, it contains universal naming conventions for the molecules and atoms in the databank, computer programs to generate and analyse the databank, basic properties calculated from all simulations (area per lipid, C-H bond order parameters, x-ray scattering form factors and membrane thickness), and quality evaluation of simulations against experimental data.
+In practise, this information is stored is the git repository which is currently available at~\url{https://github.com/NMRLipids/Databank}.
+Detailed description on its content is in the supplementary information.
+Applications in layer 3 uses the information in layer 2 to access raw data in layer 1 to perform automatic analyses over large sets of simulations in the databank.
+
+%are enabled by the 
+%The databank can be used in layer 3 by accessing the raw data and information stored in the databank by employing the universal naming conventions for atoms, molecules and simulation details.
 The applications can be linked to the core databank (layer 2) without actually including them, thereby enabling flexible extension of the data without compromising the simplicity and lightness of the core databank.
@@
 
+\subsection{Content of the databank}
 Currently the databank is composed of approximately 500 trajectories with the total length of approximately 231 microseconds of which most are contributed for the previous publications from the NMRlipids open collaboration~\cite{botan15,catte16,antila19,bacle21}.
@@
 Quality of membrane simulations with different force fields have been evaluated against experimental data during parameterization and in separate comparison studies~\cite{botan15,ollila16,catte16,pluhackova16,perez17,leonard19}, but universal quality measure for membrane simulations is not defined and controversial results are often reported from simulations~\cite{antila22b}.
-The ambiguity in defining the quality of simulations complicates the selection of proper simulation parameters and estimation of the reliability of simulation results, thereby being a major obstacle in many applications of membrane MD simulations.
-The NMRlipids
-databank enables rapid quality evaluation of MD simulations using the quantitative quality measure for membrane simulations.
-The conformational ensembles of individual lipid molecules are evaluated against C-H bond order parameters from NMR experiments, which is complement by x-ray scattering form factors arising from overall membrane dimensions via electron density profile.
+The lack of universal quality measure for membrane simulations complicates the selection of proper simulation parameters and the estimation of reliability in simulations, thereby being a major obstacle in many applications of membrane MD simulations.
+To enable the rapid quality evaluation of membrane MD simulations, a quantitative quality measure based on C-H bond order parameters from NMR experiments is defined in the NMRlipids databank.
+%Conformational ensembles of individual lipid molecules are evaluated against C-H bond order parameters from NMR experiments.
+This is complemented by comparing the locations of x-ray scattering form factor minima against experiments. 
+%that are related to the overall membrane dimensions via electron density profile. 
 Both of these are robust experimental measurables that can be directly connected to the simulation data~\cite{ollila16}.
+The used quality measures are defined in detail in the supplementary information.
 
 %NMRlipids databank contains also experimental order parameter and x-ray form factor data that is connected to corresponding  simulations in order to define the simulation qualities and rank them to select the suitable simulation models for particular applications.
-The top quality simulations based on average order parameter quality are shown in Fig.~\ref{fig:quality} A and the direct comparison between the best simulation (POPC with OPLS3e) and experiments in shown in Fig.~\ref{fig:quality} D.
-This model reproduces the experimental data very well with the exception of third carbon in the $\it{sn}$-1 chain and carbons 8 and 9 close to double bond in $\it{sn}$-2 chain.
-%Full details of the NMRlipids quality measure for lipid bilayer simulations are given in the methods section and supplementary information.
-The top ranking simulation for the order parameter quality of POPE lipid are shown in Fig.~\ref{fig:quality} B.
-In this case the best overall quality is found in GROMOS-CKP simulation (direct comparison shown in Fig.~\ref{fig:quality} E, but clear discrepancies with experiments are present in this model for glycerol backbone and $\it{sn}$-2 chain.
-On the other hand, glycerol backbone and headgroup for POPE is best described in CHARMM36 model (direct comparison shown in Fig.~\ref{fig:quality} F), but its overall quality suffers from overestimated order in acyl chains.
-Scatter plots and Pearson correlation coefficients in Fig.~\ref{fig:quality} C reveal strong correlation between membrane area per lipid and thickness.
-The area per lipid is also strongly correlated with the average order of $\it{sn}$-1 chain, suggesting that its order parameters can used as a proxy for the area per lipid in addition to the conformational ensemble.
-On the other hand, the first form factor minima is correlated stronger with membrane thickness (Fig.~\ref{fig:quality} C) than with the area per lipid (Fig.~\ref{fig:QualityCorrelationsSI}).
-
- %This is demonstrated in Fig.~\ref{fig:quality} a) showing correlations between membrane lateral density (area per lipid), thickness, minima of form factor and acyl chain order analyzed from all apporimately 500 trajectories in the NMRlipids databank.
-Membrane area per lipid has negative correlations with membrane order and thickness with Pearson coefficient of -0.78 and -0.49, respectively, because decreased area leads to more ordered, tightly packed and stretched acyl chains leading to thicker membranes.
-On the other hand, locations of form factor minima have positive correlation with the membrane area and negative correlation with the thickness.
-In conclusion, both x-ray scattering form factors and acyl chain order parameters from NMR are on average good proxies for membrane packing and thickness, while NMR order parameters give segmental resolution information on the quality of conformational ensembles of lipids.
+Simulations with the highest scores based on order parameter evaluation for all simulations are shown in Fig.~\ref{fig:quality} A and for POPE lipids in Fig.~\ref{fig:quality} B.
+The direct comparison with experiments for the best simulation, POPC with OPLS3e, is shown in Fig.~\ref{fig:quality} C, where discrepancies with experiments are observed only for the third carbon in the $\it{sn}$-1 chain, and carbons 8 and 9 close to the double bond in $\it{sn}$-2 chain.
+The best overall quality for POPE is found in GROMOS-CKP simulation, for which the direct comparison in Fig.~\ref{fig:quality} D show minor differences in acyl chains, but major discrepancies in glycerol backbone and in the first carbon of $\it{sn}$-2 chain.
+These parts are better described for POPE by CHARMM36 parameters (direct comparison shown in Fig.~\ref{fig:quality} E), but its overall quality suffers from the overestimated order in acyl chains.
 
@@
 
-The power of NMRlipids quality metrics in selecting the best model for particular application is demontrated in Fig.~\ref{fig:quality} c) where the area per lipids of POPC:POPE mixtures from different force fields are shown and quality of POPE simulation in different force fields are evaluated.
+To exemplify the finding of optimal model for a certain lipid, the top ranking simulation for the order parameter quality of POPE lipid are shown in Fig.~\ref{fig:quality} B.
+
+Differences in the predictions of area per lipids for POPC and POPE bilayers between different force fields are demonstrated in Fig.~\ref{fig:POPCPOPEapls}.
+Because area per lipid is strongly correlated with the average order of $\it{sn}$-1 chain ((Fig.~\ref{fig:quality} F)), its order parameters can used as a proxy for the area per lipid to evaluate its quality against experiments.
+Simulations with the largest area per lipids for POPE in Fig.~\ref{fig:POPCPOPEapls}, GROMOS-CKP and Slipids, have significantly higher qualities in Fig.~\ref{fig:quality} B than lipid17 and CHARMM36 with smaller are per lipids.
 Among the simulations available in the NMRlipids databank, the Slipids force field gives the best quality in terms of acyl chain order parameters and form factor, and predicts the largest area per molecule for POPE and smallest difference with POPC.
@@
 Similar comparisons utilizing the preliminary data from the NMRlipids databank have concluded that Slipids is relatively good model also for mixtures of charged POPC:POPS membranes, although models with better counterion binding have higher quality, but CHARMM36 is the best to study differences in headgroup conformational ensembles between different lipid types.
+
+On the other hand, the first form factor minima is correlated stronger with membrane thickness (Fig.~\ref{fig:quality} C) than with the area per lipid (Fig.~\ref{fig:QualityCorrelationsSI}). 
+Scatter plots and Pearson correlation coefficients in Fig.~\ref{fig:quality} C reveal strong correlation between membrane area per lipid and thickness.
+
+\begin{figure}[tbp]
+    \centering
+    \includegraphics[width=88mm]{Figures/APL.pdf}
+    \caption{Differences in area per lipids of POPC and POPE bilayers between different force fields.  
+    }
+    \label{fig:POPCPOPEapls}
+\end{figure}
+
 Understanding such complex picture of lipid bilayer MD simulation quality would not be possible without automatic quality evaluation of large sets of simulations enabled by the NMRlipids databank.
+
+%The power of NMRlipids quality metrics in selecting the best model for particular application is demontrated in Fig.~\ref{fig:quality} c) where the area per lipids of POPC:POPE mixtures from different force fields are shown and quality of POPE simulation in different force fields are evaluated.
+
+ %This is demonstrated in Fig.~\ref{fig:quality} a) showing correlations between membrane lateral density (area per lipid), thickness, minima of form factor and acyl chain order analyzed from all apporimately 500 trajectories in the NMRlipids databank.
+Membrane area per lipid has negative correlations with membrane order and thickness with Pearson coefficient of -0.78 and -0.49, respectively, because decreased area leads to more ordered, tightly packed and stretched acyl chains leading to thicker membranes.
+On the other hand, locations of form factor minima have positive correlation with the membrane area and negative correlation with the thickness.
+In conclusion, both x-ray scattering form factors and acyl chain order parameters from NMR are on average good proxies for membrane packing and thickness, while NMR order parameters give segmental resolution information on the quality of conformational ensembles of lipids.
 
@@
 %The Discussion should be succinct and must not contain subheadings.
+
+, thereby not requiring large resources to handle and maintain.
+This lowers the barrier for starting such databank as well as for long term storage.
+The NMRlipids databank is essentially a git repository containing information on the location of raw data and its indexing with universal naming convention.
+In addition to all computers where the databank is developed and used, the NMRlipids databank git is stored to Zenodo server, thereby enabling a very cost effective long term storage for the databank.
 

### 2022-10-05 SI.tex
Updates from Overleaf
Add figure to SI and update pdf

@@
 \usepackage{tablefootnote}
+\usepackage{xr}
 
@@
 \makeatother
+
+\makeatletter
+\newcommand*{\addFileDependency}[1]{% argument=file name and extension
+  \typeout{(#1)}
+  \@addtofilelist{#1}
+  \IfFileExists{#1}{}{\typeout{No file #1.}}
+}
+\makeatother
+
+\newcommand*{\myexternaldocument}[1]{%
+    \externaldocument{#1}%
+    \addFileDependency{#1.tex}%
+    \addFileDependency{#1.aux}%
+}
+
+\myexternaldocument{main}
+
+\renewcommand{\thetable}{S\arabic{table}}%
+\renewcommand{\thefigure}{S\arabic{figure}}%
+\renewcommand{\thesection}{S\arabic{section}}%
+\renewcommand{\thepage}{S\arabic{page}}%
 
@@
 }
-    \label{fig:sizedependence}
+    \label{fig:QualityCorrelationsSI}
 \end{figure}
@@
     \caption{Dependence of form factor, electron density profiles along membrane normal, and C-H bond orderparameters (from top to bottom) on simulation box size with different cholesterol concentration.
+Simulations with 64, 256 or 1024 POPC lipids taken from Ref.~\citenum{javanainen_matti_2021_7035350}.
 }
     \label{fig:sizedependence}
+\end{figure}
+
+\pagebreak
+\section{Finding the best models for PC and PE mixtures}
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width = 140mm]{Figures/POPC_POPE_dataSI.pdf}
+    \caption{Simulations with the data for both POPC (top) and POPE (bottom) directly compared with the experimental data.
+      The area per lipid in simulations increase from left to right.
+      Simulations with the best overall quality for POPC and POPE order parameters are circled by a square.
+  }
+    \label{fig:POPC_POPE_dataSI}
 \end{figure}

### 2022-10-06 main.tex
Update text

@@
 
-To exemplify the finding of optimal model for a certain lipid, the top ranking simulation for the order parameter quality of POPE lipid are shown in Fig.~\ref{fig:quality} B.
-
-Differences in the predictions of area per lipids for POPC and POPE bilayers between different force fields are demonstrated in Fig.~\ref{fig:POPCPOPEapls}.
-Because area per lipid is strongly correlated with the average order of $\it{sn}$-1 chain ((Fig.~\ref{fig:quality} F)), its order parameters can used as a proxy for the area per lipid to evaluate its quality against experiments.
-Simulations with the largest area per lipids for POPE in Fig.~\ref{fig:POPCPOPEapls}, GROMOS-CKP and Slipids, have significantly higher qualities in Fig.~\ref{fig:quality} B than lipid17 and CHARMM36 with smaller are per lipids.
-Among the simulations available in the NMRlipids databank, the Slipids force field gives the best quality in terms of acyl chain order parameters and form factor, and predicts the largest area per molecule for POPE and smallest difference with POPC.
-In conclusion, simulations with Slipids is the most reliable force field for membrane packing in POPC:POPE mixtures althought its glycerol backbone is not accurately modelled.
-Similar comparisons utilizing the preliminary data from the NMRlipids databank have concluded that Slipids is relatively good model also for mixtures of charged POPC:POPS membranes, although models with better counterion binding have higher quality, but CHARMM36 is the best to study differences in headgroup conformational ensembles between different lipid types.
-
-On the other hand, the first form factor minima is correlated stronger with membrane thickness (Fig.~\ref{fig:quality} C) than with the area per lipid (Fig.~\ref{fig:QualityCorrelationsSI}). 
-Scatter plots and Pearson correlation coefficients in Fig.~\ref{fig:quality} C reveal strong correlation between membrane area per lipid and thickness.
-
+\subsection{Finding the best models for PC and PE lipid mixtures}
+The power of the NMRlipids databank in finding the best models for a certain lipid bilayer system is exemplified here for mixtures of PE and PC lipids.
+Area per lipids from available POPC and POPE bilayer simulations in the databank from different force fields at 310\,K are shown in Fig.~\ref{fig:POPCPOPEapls}.
+Because the area per lipid strongly correlates with the average order of $\it{sn}$-1 chain (Fig.~\ref{fig:quality} F), we use order  parameter qualities to evaluate which simulations predict the best area per lipids for POPC and POPE.
+Among the force fields with area per lipids shown in Fig.~\ref{fig:POPCPOPEapls}, Slipids ranks 1st for POPC with a clear difference to others, and 2nd for POPE with only marginally lower quality than GROMOS-CKP (Figs.~\ref{fig:quality} A and B).
+The direct comparison with experiments in Fig.~\ref{fig:POPC_POPE_dataSI} in the supplementary information reveals that GROMOS-CKP and CHARMM36 predicts too ordered (too low area per lipid) bilayer for POPC, while the order parameters in Lipid17 are lower than in experiments indicating overestimated area per lipid.
+For POPE, order parameters in  Slipids and GROMOS-CKP are close to experiments suggesting that their are per lipids are most reasonable, while CHARMM36 and Lipid17 are too packed thereby overestimating the order parameters.
 \begin{figure}[tbp]
@@
 
-Understanding such complex picture of lipid bilayer MD simulation quality would not be possible without automatic quality evaluation of large sets of simulations enabled by the NMRlipids databank.
+In conclusion, the quality evaluation based on the NMRlipids databank suggests that the Slipids parameters would give the most reliable description for membrane packing of POPC:POPE mixtures.
+Slipids was a relatively good model also for POPC:POPS mixtures in similar comparisons utilizing the preliminary version of the NMRlipids databank~\cite{antila22b}, although charged membranes are more complicated as the counterion binding affinity plays a significant role in membrane packing~\cite{antila19}.
+On the other hand, CHARMM36 overcomes Slipids when the focus is on conformational ensembles of headgroups  in different types of lipids because the accuracy of glycerol backbone is low in Slipids~\cite{bacle21}.
+Altogether, these examples demonstrate how the NMRlipids databank can be used to navigate in the complex landscape of lipid force field quality.
+
+%On the other hand, the first form factor minima is correlated stronger with membrane thickness (Fig.~\ref{fig:quality} C) than with the area per lipid (Fig.~\ref{fig:QualityCorrelationsSI}). 
+%Scatter plots and Pearson correlation coefficients in Fig.~\ref{fig:quality} C reveal strong correlation between membrane area per lipid and thickness.
+
+%Understanding such complex picture of lipid bilayer MD simulation quality would not be possible without automatic quality evaluation of large sets of simulations enabled by the NMRlipids databank.
 
@@
 \subsection{Water diffusion anisotropy in membrane systems}
-The anisotropic water diffusion through and along of membranes is utilized in MRI imaging \cite{??} and plays a role in the drug translocation through biological material, particularly in skin \cite{hansen13,wen18,nitsche19,roberts21,??}.
-Dependencies between membrane properties, water permeablity, and anisotropic water diffusion have been recognized \cite{??} but consensus on how water permeability depends on membrane properties, such as area per lipid and thickness, or molecular composition, has not yet been reached despite of several available models~\cite{nitsche13,nitsche16,shinoda16,venable19,??}.
-MD simulations can be used to elucidate such dependencies, but systematic collection of sufficient amount of data has been challenging because few water permeation events are typically observed in a single MD simulation trajectory \cite{??}.
-Nevertheless, averaging over large amounts of data available in the NMRlipids databank enables to study the trends of water permeation upon changing memebrane properties.
+The anisotropic water diffusion through and along of membranes plays a role in the drug translocation through biological material, particularly in skin \cite{hansen13,wen18,nitsche19,roberts21,??}, and in MRI imaging~\cite{??}.
+Therefore, dependence of water permeation through membranes and translational diffusion at membrane surfaces on membrane properties has been studied with different methods, yet consensus
+%on how water permeability depends on membrane properties, such as area per lipid and thickness, or molecular composition, 
+has not been reached despite of several available models~\cite{nitsche13,nitsche16,shinoda16,venable19,??}.
+MD have been also used in these endeavours, but systematic studies are limited by the accessible amount of data as only few water permeation events are typically observed in a single MD simulation trajectory~\cite{??}.
+Therefore, this serves as an excellent example where the large amount of available MD simulations in the NMRlipids databank can be utilized to answer important biological questions.
 
-%averaging Here we demonstrate the power of the NMRlipids databank by analysing how 
-Water permeabilities calculated from all the simulations in the NMRlipids databank as a function of area per lipids, thickness, and acyl chain order are shown in  Fig.~\ref{fig:permeability} (a-c).
-To illustrate how permeability depends these properties, average over systems with fixed values of area per lipids, thickness, and acyl chain order are also shown in the figure.
-Permeability increases with increasing area per lipid and decreases with increasing membrane thickness and acyl chain order.
-Previous simulation and experimental studies have reported linear dependence of permeability on both area per lipid and thickness for limited set of lipid mixtures but not always for all types of lipid mixtures \cite{mathai08,frallicciardi22}.
-Our results suggest stronger dependence for thicknesses below 3.9\,nm and areas above 69\,Å$^2$, but the dependence may appear linear above and below these values as shown in the insets of Figs.
-\label{fig:permeability} a) and b).
-As expected, the permeability increases with increasing temperature in Figs.
-\label{fig:permeability} d).
-The Arrhenius plot gives 17\,$\pm$\,4\,k$_b$T for the average energy barrier for the water permeation.
-The values for water permeabilities from simulations in the databank vary between 0.3 and 322\,$\mu$m/s with the mean and median of 14\,$\mu$m/s and 8\,$\mu$m/s, respectively.
+To this end, we first calculated the water permeability through membranes from all the simulations in the NMRlipids databank.
+Resulting permeabilities are shown as a function of temperature, membrane thickness, area per lipid, and acyl chain order in Figs.~\ref{fig:permeability} B-E.
+The values for water permeabilities from simulations vary between 0.3 and 322\,$\mu$m/s with the mean and median of 14\,$\mu$m/s and 8\,$\mu$m/s, respectively.
 These values have the same order of magnitude as the experimentally determined diffusive permeability coefficients, but are on average below the values reported for PC lipids in liquid crystalline phase,~190-330\,$\mu$m/s~\cite{jansen95}.
+To analyze how water permeability through membranes depend on these properties on average, we calculated the average permeabilities over all systems with the value of temperature, thickness, area per lipid, or acyl chain order within a fixed range.
+These averaged permeabilities also shown in Figs.~\ref{fig:permeability} with black dots.
+As expected, the permeability increases with the temperature, giving 17\,$\pm$\,4\,k$_b$T for the average energy barrier for the water permeation from the Arrhenius plot in Fig.~\label{fig:permeability} B.
+Water permeability decreases with increasing membrane packing, i.e., with decreasing area per lipid and increasing thickness and acyl chain order.
+Linear dependence of permeability on area per lipid and thickness is previously reported from simulations and experiments for some systems, but not for all~\cite{mathai08,frallicciardi22}.
+Our results suggest on average linear dependence for thicknesses above $\sim$\,3.9\,nm and area per lipids below $\sim$69\,Å$^2$ (insets in Figs.~\ref{fig:permeability} C-D), yet stronger dependence for more loosely packed membranes.
 %within the range of experimentally reported values  permeability which is 1-2 orders of magnitude slower than osmotic permeability \cite{jansen95,??}. 
-Clear dependencies of permeability on hydration level, charged lipid fraction, cholesterol fraction or POPE fraction were not observed, although weak decrease for the latter two may be visible.
+Clear dependencies of permeability on charged lipids, cholesterol, POPE, or hydration level were not observed in Fig.~\ref{fig:permeationSI} in the supplementary information.
+%, although weak decrease for the latter two may be visible.
 
-\begin{figure}[tbp]
+\begin{figure}[tb]
     \centering
     \includegraphics[width=180mm]{Figures/permeation2.pdf}
-    \caption{(b-c) Water permeation through lipid membranes analyzed from the databank as a function of temperature, thickness, area per lipid, and acyl chain order. 
-      Inserts in c) and d) show the region where the dependence could be considered approximately linear.
-      Insert in a) shows the Arrhenius plot of permeation ($\ln$ (P) vs. 1/T) that gives 17\,$\pm$\,4\,k$_b$T for the average activation energy for water permeation through lipid bilayer.
-      (f-g) Diffusion anistoropy of water as a function of thickness and area per lipid.
-      (h) Lateral diffusion of water as a function of hydration level. Experimental points for DMPC bilayers at 313\,K at different hydration levels are shown~\cite{rudakova04}.}
+    \caption{A 
+    {\bf B-E} Water permeation through membranes analyzed from the databank as a function of temperature, thickness, area per lipid, and acyl chain order.
+Values from simulations with non-zero permeation values are shown with blue dots.
+Averages over fixed range of x-axis values are shown with black dots.
+Insert in B) shows the Arrhenius plot of permeation ($\ln$ (P) vs.
+1/T) that gives 17\,$\pm$\,4\,k$_b$T for the average activation energy for water permeation through lipid bilayer.
+Inserts in C) and D) show the region where the dependence could be considered approximately linear.
+    {\bf F-G} Diffusion anistoropy of water as a function of thickness and area per lipid.
+    {\bf H} Lateral diffusion of water as a function of hydration level. Experimental points for DMPC bilayers at 313\,K at different hydration levels are shown~\cite{rudakova04}.}
     \label{fig:permeability}
@@
 
-To analyze the anisotropic water diffusion, we also calculated the water diffusion along the membrane.
-The results at different hydration levels are shown in Fig.~\ref{fig:permeability} h) together with the experimental data~\cite{rudakova04}.
-The experimental water diffusion increases with increasing hydration level toward the value for bulk water (3.1$\cdot$\,10$^{-9}$\,m$^2$/s at 313\,K)\cite{khakimov08}.
-Simulations overestimate the experimental bulk value approximately with the factor of 2 at high hydration levels which is not surprising as the most common water model used in membrane simulations, TIP3P, overestimates the bulk water diffusion.
-However, simulation results are closer to experiments with low hydration levels.
-To estimate the diffusion anistoropy, we translated the water permeation to diffusion coefficient through multilamellar stack using the Tanner equation~\cite{??}.
-The resulting diffusion coefficients trough the membranes are approximately five orders of magnitude slower than along the membrane which is at the upper limits of the anisotropic estimated from the experimental data~\cite{nitsche19}.
+To analyze the anisotropic diffusion of water in membrane environment, we also calculated the water translational diffusion along the membrane surface from all simulations in the NMRlipids databank.
+The results at different hydration levels are shown in Fig.~\ref{fig:permeability} H together with the experimental data~\cite{rudakova04}.
+%The experimental water diffusion increases with increasing hydration level toward the value for bulk water . 
+%As expected, water translational diffusion increase in the diffusion is observed with the increased temperature and hydration level (Figs.~\ref{fig:permeability} H and~\ref{fig:diffusionSI} B). 
+Simulation results are close to the experimental values at low hydration levels, but increase to the values of approximately two times higher than experimental bulk water diffusion value (3.1$\cdot$\,10$^{-9}$\,m$^2$/s at 313\,K~\cite{khakimov08}) at high hydration levels.
+This is not surprising as the most common water model used in membrane simulations, TIP3P, overestimates the bulk water diffusion~\cite{??}.
+The water translational diffusion incereased with the temperature, but dependencies on membrane area per lipid, thickness, or fraction of charged lipids were not observed in Fig.~\ref{fig:diffusionSI} in the supplementary information.
+In order to estimate the diffusion anistoropy of water in multilamellar membrane system, we translated the water permeation coefficients through membranes to diffusion coefficients through multilamellar stack using the Tanner equation~\cite{??}.
+The resulting diffusion coefficients through the membranes are approximately five orders of magnitude slower than along the membrane (Figs.~\ref{fig:permeability} F and G), which is at the upper limits of the anisotropic estimated from the experimental data~\cite{nitsche19}.
 Relatively high anisotropy is understandable as simulations give slightly slower permeation rates and higher lateral diffusion rates than experiments.
-Significant dependence of diffusion anisotropy on membrane thickness and area per lipid are observed in Figs.~\ref{fig:permeability} f) and g).
-The anisotropy becomes linearly stronger with decreasing area per lipid while the dependency is stronger with increasing thickness.
-These results can be explained by the increasing and decreasing permeabilities with thickness and are per lipid (Figs \ref{fig:permeability} c) and d), respectively, while lateral diffusion remains approximately constant (Fig.
-\ref{??} a) and c)).
+Significant increase in the diffusion anisotropy with membrane packing is observed, as $D_{\perp}/D_{||}$ drifts away from one with decreasing area per lipid and increasing thickness in Figs.~\ref{fig:permeability} F and G.
+This follows from decreasing water permeability with membrane packing (Figs \ref{fig:permeability} C and D), while lateral diffusion remains approximately constant (Fig.
+\ref{fig:diffusionSI} A and C).
 

### 2022-10-06 main.tex
Update figures and text

@@
 Inserts in C) and D) show the region where the dependence could be considered approximately linear.
-    {\bf F-G} Diffusion anistoropy of water as a function of thickness and area per lipid.
-    {\bf H} Lateral diffusion of water as a function of hydration level. Experimental points for DMPC bilayers at 313\,K at different hydration levels are shown~\cite{rudakova04}.}
+    {\bf F} Lateral diffusion of water as a function of hydration level. Experimental points for DMPC bilayers at 313\,K at different hydration levels are shown~\cite{rudakova04}.
+    {\bf G-H} Diffusion anistoropy of water as a function of thickness and area per lipid. }
     \label{fig:permeability}
@@
 To analyze the anisotropic diffusion of water in membrane environment, we also calculated the water translational diffusion along the membrane surface from all simulations in the NMRlipids databank.
-The results at different hydration levels are shown in Fig.~\ref{fig:permeability} H together with the experimental data~\cite{rudakova04}.
+The results at different hydration levels are shown in Fig.~\ref{fig:permeability} F together with the experimental data~\cite{rudakova04}.
 %The experimental water diffusion increases with increasing hydration level toward the value for bulk water . 
@@
 In order to estimate the diffusion anistoropy of water in multilamellar membrane system, we translated the water permeation coefficients through membranes to diffusion coefficients through multilamellar stack using the Tanner equation~\cite{??}.
-The resulting diffusion coefficients through the membranes are approximately five orders of magnitude slower than along the membrane (Figs.~\ref{fig:permeability} F and G), which is at the upper limits of the anisotropic estimated from the experimental data~\cite{nitsche19}.
+The resulting diffusion coefficients through the membranes are approximately five orders of magnitude slower than along the membrane (Figs.~\ref{fig:permeability} G and H), which is at the upper limits of the anisotropic estimated from the experimental data~\cite{nitsche19}.
 Relatively high anisotropy is understandable as simulations give slightly slower permeation rates and higher lateral diffusion rates than experiments.
-Significant increase in the diffusion anisotropy with membrane packing is observed, as $D_{\perp}/D_{||}$ drifts away from one with decreasing area per lipid and increasing thickness in Figs.~\ref{fig:permeability} F and G.
+Significant increase in the diffusion anisotropy with membrane packing is observed, as $D_{\perp}/D_{||}$ drifts away from one with decreasing area per lipid and increasing thickness in Figs.~\ref{fig:permeability} G and H.
 This follows from decreasing water permeability with membrane packing (Figs \ref{fig:permeability} C and D), while lateral diffusion remains approximately constant (Fig.

### 2022-10-12 main.tex
Update text

--- 2022-10-06 main.tex
@@
 %The demand for sharing and reusing of MD simulation data is increasing, but practical solution remains unclear due to unsolved issues in data storage and indexing.
+
+While MD simulations of systems with few biomolecules continue having major contributions in our understanding on biology and drug design, researches are now searching new avenues by building models not only for individual molecules but whole cells or organelles using interdisciplinary approaches~\cite{johnson15,thornburg22,gupta22}.
 
@@
 \subsection{Water diffusion anisotropy in membrane systems}
-The anisotropic water diffusion through and along of membranes plays a role in the drug translocation through biological material, particularly in skin \cite{hansen13,wen18,nitsche19,roberts21,??}, and in MRI imaging~\cite{??}.
-Therefore, dependence of water permeation through membranes and translational diffusion at membrane surfaces on membrane properties has been studied with different methods, yet consensus
+The anisotropy in water diffusion in parallel and perpendicular directions with respect to membranes plays a role in the drug translocation through biological material, particularly in skin \cite{hansen13,wen18,nitsche19,roberts21}, and in MRI imaging~\cite{topgaard20}.
+Several models on how water dynamics depends on membrane properties have been proposed based on different experimental and theoretical methods, yet consensus
 %on how water permeability depends on membrane properties, such as area per lipid and thickness, or molecular composition, 
-has not been reached despite of several available models~\cite{nitsche13,nitsche16,shinoda16,venable19,??}.
-MD have been also used in these endeavours, but systematic studies are limited by the accessible amount of data as only few water permeation events are typically observed in a single MD simulation trajectory~\cite{??}.
+has not been reached~\cite{mathai07,nitsche13,nitsche16,shinoda16,venable19,frallicciardi22}.
+MD simulations have been also used in these endeavours, but systematic studies are limited by the accessible amount of data as only few water permeation events are typically observed in a single MD simulation trajectory~\cite{venable19,camilo2022}.
 Therefore, this serves as an excellent example where the large amount of available MD simulations in the NMRlipids databank can be utilized to answer important biological questions.
@@
 These values have the same order of magnitude as the experimentally determined diffusive permeability coefficients, but are on average below the values reported for PC lipids in liquid crystalline phase,~190-330\,$\mu$m/s~\cite{jansen95}.
-To analyze how water permeability through membranes depend on these properties on average, we calculated the average permeabilities over all systems with the value of temperature, thickness, area per lipid, or acyl chain order within a fixed range.
+To analyze how water permeability through membranes depend on bilayer properties on average, we calculated the average permeabilities over all systems with the value of temperature, thickness, area per lipid, or acyl chain order within a fixed range.
 These averaged permeabilities also shown in Figs.~\ref{fig:permeability} with black dots.
-As expected, the permeability increases with the temperature, giving 17\,$\pm$\,4\,k$_b$T for the average energy barrier for the water permeation from the Arrhenius plot in Fig.~\label{fig:permeability} B.
+As expected, the permeability increases with the temperature, giving 17\,$\pm$\,4\,k$_b$T for the average energy barrier for the water permeation from the Arrhenius plot in Fig.~\ref{fig:permeability} B.
 Water permeability decreases with increasing membrane packing, i.e., with decreasing area per lipid and increasing thickness and acyl chain order.
-Linear dependence of permeability on area per lipid and thickness is previously reported from simulations and experiments for some systems, but not for all~\cite{mathai08,frallicciardi22}.
-Our results suggest on average linear dependence for thicknesses above $\sim$\,3.9\,nm and area per lipids below $\sim$69\,Å$^2$ (insets in Figs.~\ref{fig:permeability} C-D), yet stronger dependence for more loosely packed membranes.
+Linear dependence of permeability on area per lipid and thickness is previously reported from simulations and experiments for some systems, but not for all~\cite{mathai07,venable19,frallicciardi22}.
+Our results suggest on average linear dependence for thicknesses above $\sim$\,3.9\,nm and area per lipids below $\sim$69\,Å$^2$ (insets in Figs.~\ref{fig:permeability} C-D), yet stronger dependence is observed for more loosely packed membranes.
 %within the range of experimentally reported values  permeability which is 1-2 orders of magnitude slower than osmotic permeability \cite{jansen95,??}. 
@@
 
-To analyze the anisotropic diffusion of water in membrane environment, we also calculated the water translational diffusion along the membrane surface from all simulations in the NMRlipids databank.
+To analyze the anisotropic diffusion of water in membrane environment, we also calculated the water diffusion parallel to the membrane surface from all simulations in the NMRlipids databank.
 The results at different hydration levels are shown in Fig.~\ref{fig:permeability} H together with the experimental data~\cite{rudakova04}.
@@
 %As expected, water translational diffusion increase in the diffusion is observed with the increased temperature and hydration level (Figs.~\ref{fig:permeability} H and~\ref{fig:diffusionSI} B). 
-Simulation results are close to the experimental values at low hydration levels, but increase to the values of approximately two times higher than experimental bulk water diffusion value (3.1$\cdot$\,10$^{-9}$\,m$^2$/s at 313\,K~\cite{khakimov08}) at high hydration levels.
-This is not surprising as the most common water model used in membrane simulations, TIP3P, overestimates the bulk water diffusion~\cite{??}.
-The water translational diffusion incereased with the temperature, but dependencies on membrane area per lipid, thickness, or fraction of charged lipids were not observed in Fig.~\ref{fig:diffusionSI} in the supplementary information.
-In order to estimate the diffusion anistoropy of water in multilamellar membrane system, we translated the water permeation coefficients through membranes to diffusion coefficients through multilamellar stack using the Tanner equation~\cite{??}.
+Simulation results are close to the experimental values with low hydration levels, but increase to the values of approximately two times higher than experimental bulk water diffusion value (3.1$\cdot$\,10$^{-9}$\,m$^2$/s at 313\,K~\cite{khakimov08}) with high hydration levels.
+This is not surprising as the most common water model used in membrane simulations, TIP3P, overestimates the bulk water diffusion~\cite{pathirannahalage21}.
+The water diffusion parallel to membrane surface increased with the temperature, but dependencies on membrane area per lipid, thickness, or fraction of charged lipids were not observed in Fig.~\ref{fig:diffusionSI} in the supplementary information.
+In order to estimate the diffusion anistoropy of water in multilamellar membrane system, we translated the water permeation coefficients through membranes to diffusion coefficients through multilamellar stack using the Tanner equation~\cite{tanner78,wasterby02}.
 The resulting diffusion coefficients through the membranes are approximately five orders of magnitude slower than along the membrane (Figs.~\ref{fig:permeability} F and G), which is at the upper limits of the anisotropic estimated from the experimental data~\cite{nitsche19}.
@@
 
+\subsection{Flip-flops}
+Lipid flip-flops from one bilayer leaflet to another play an important role in lipid trafficking and regulating membrane properties~\cite{vanmeer08}.
+Phospholipid flip-flops are slow, occuring with the timescales of hours or days, while cholesterol, diacylglycerol and ceramide flip-flops are faster, yet the reported timescales range between minutes to sub-millisecods~\cite{vanmeer08,steck12,parisio16,gu19}.
+Flip-flops have been previously studied mainly with coarse-grained simulations and free energy calculations due to time scale limitations in atomistic simulations~\cite{parisio16}, but direct analyses on cholesterol flip-flops in atomistic simulations have been also emerging in recent years~\cite{gu19,javanainen18,baral20}.
+However, systematic studies on how flip-flop rates depend on membrane properties are still challenging even with the state of the art computational resources.
+The NMRlipids databank can greatly advance such studies by giving an easy access to large amounts of data simultaneously.
+
 \section{Discussion}
@@
 %The Discussion should be succinct and must not contain subheadings.
+
+Building accessible databanks of molecular dynamics simulation data has been challenging due to the required long term support for hardware and software maintenance.
+In the overlay model used in the NMRlipids databank, the demand of hardware can be distributed and open collaboration model reduces the risk for ending software maintenance.
+Furthermore, the open collaboration model used in the NMRlipids project credits contributors by offering authorship in published articles, thereby creating an incentive for contributions.
+This model could be extended also to other fields where similar barriers to establish publicly accessible databanks exist.
 
@@
 
-Building accessible databanks of molecular dynamics simulation data has been challenging due to the required long term support for hardware and software maintenance.
-In the overlay model used in the NMRlipids databank, the demand of hardware can be distributed and open collaboration model reduces the risk for ending software maintenance.
-Furthermore, the open collaboration model used in the NMRlipids project credits contributors by offering authorship in published articles, thereby creating an incentive for contributions.
-This model could be extended also to other fields where similar barriers to establish publicly accessible databanks exist.
 Emerging applications of machine learning are increasing the impact of such databanks.
@@
 \subsection{Structure of the databank}
+Structure of the NMRlipids databank is illustrated more detailed in Fig.~\ref{DatabankStructure} in the supplementary information.
+The databank locates as a git repository at \url{https://github.com/NMRLipids/Databank/} and is permanently stored in Zenodo repository (\url{www.zenodo.org})~\cite{??}.
+Whenever a specific file is referred here, the file path within the NMRlipids databank repository is given.
+The scripts in the NMRlipids databank are mainly written in Python and many of them utilize the MDAnalysis module~\cite{gowers2019mdanalysis,michaud2011mdanalysis}.
 
-\begin{figure}
-  \includegraphics[width=\textwidth]{Figures/DataBankChart.pdf}
-  \caption{Structure of the NMRlipids databank.
-Manually added input data (blue boxes) includes basic information on the simulation, permanent links to the raw data, and experimental data if available.
-The databank entries (red box) and analysis results (green boxes), locating at \url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations} are automatically generated by the computer programs included in the NMRlipids databank (yellow boxes).
-Because raw data are not permanently stored but can be accessed based on the information in the databank, this connection is marked with the dashed line.
-}\label{DatabankStructure}
-\end{figure}
-
-Structure of the NMRlipids databank is illustrated in Fig.~\ref{DatabankStructure}.
-The required input information to create an entry into the NMRlipids databank are listed in table~\ref{tab:READMEkeys}.
-While the raw simulation data is not directly stored in the NMRlipids databank, permanent links from where the raw data can be accessed have to be given and are then stored in the README.yaml files at \url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations}.
-These files contain all the essential information listed in table~\ref{tab:READMEkeys} on each simulation entry that are needed for further applications.
-The raw MD simulation data can locate in any stable publicly available repository, although all the current data locates in Zenodo \url{www.zenodo.org}.
-
-%A README.yaml file is made using AddData.py script for each simulation stored in the databank based on the information given by the author of the simulation data.
-In addition to this information, automatically extracted information about the simulation are stored in the README.yaml files (see table~\ref{tab:READMEkeys}).
-A README.yaml file for each simulation is stored to a folder named according to the hashes of original trajectory and topology file locating in \url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations}.
-These README.yaml files give access to all the essential information on the simulations in the databank either directly, or pointing to a original file from where the information can be extracted.
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{  p{3.5cm}  p{9.5cm}  p{4.0cm} }
-    \toprule
-    key & description & type  \\
-    \midrule
-        %\hline
-    DOI & DOI from where the raw data is found & user given (compulsory) \\
-    SOFTWARE & Software used to run the simulation (e.g. Gromacs, Amber, NAMD, etc.) & \\
-    TRJ & Name of the trajectory file found from DOI & \\
-    TPR & Name of the topology file found from DOI (trp file in the case of Gromacs) & \\
-    PREEQTIME & Pre-equilibrate time simulated before the uploaded trajectory in nanoseconds.
-\tablefootnote{For example, if you upload 100-200 ns part of total 200 ns simulation, this should value should be 100.} & \\
-    TIMELEFTOUT & Equilibration period in the uploaded trajectory that should be discarded in analyses.
-\tablefootnote{For example, if you upload 0-200 ns part of total 200 ns simulation where the first 100 ns should be considered as an equilibration, this value should be 100.} \\
-    COMPOSITION & Molecules names used in the simulation and corresponding mapping files (see section \ref{naming}) & \\
-    DIR\_WRK & Temporary working directory in your local computer. \\
-    UNITEDATOM\_DICT & Information for constucting hydrogens for united atom simulations, empty for all atom simulations & \\
-    TYPEOFSYSTEM & Lipid bilayer or something else & \\
-    \hline
-    PUBLICATION & Give reference to a publication(s) related to the data. & User given (optional)\\
-    AUTHORS\_CONTACT & Name and email of the main author(s) of the data. & \\
-    SYSTEM & System description on free text format & \\
-    SOFTWARE\_VERSION & Version of the used software & \\
-    FF & Name of the used force field & \\
-    FF\_SOURCE & Source of the force field parameters, e.g, CHARMM-GUI, webpage, citation to a publication, etc. & \\
-    FF\_DATE &  Date when force field parameters were accessed on the gives source (day/month/year). & \\
-    FF{molename} & Molecule specific force field information, e.g., water model with FFSOL and sodium parameters with FFSOD. & \\
-    CPT & Name of the Gromacs checkpoint file. & \\
-    LOG & Name of the Gromacs log file. & \\
-    TOP & Name of the Gromacs top file. & \\
-    GRO & Name of the Gromacs gro file. & \\
-    \hline
-    TRAJECTORY\_SIZE & Size of the trajectory file in bytes & automatically extracted data. \\
-    TRJLENGTH & Lenght of the trajectory (ps). & \\
-    TEMPERATURE & Temperature of the simulation. & \\
-    NUMBER\_OF\_ATOMS & Number of atoms in the simulation. & \\
-    DATEOFRUNNIG & Date when added into the databank & \\
-    EXPERIMENT & Potentially connected experimental data & \\
-    COMPOSITION & Numbers of lipid molecules (NPOPC, NPOPG, etc.) per membrane leaflet are calculated by determining on which side of the center of mass of the membrane the center of mass of the head group of each lipid molecule is located.
-    Numbers of other molecules such as solvent and ions (NSOL, NPOT, NSOD, etc.) are read from the topology file. & \\
-    \end{tabular}
-    \caption{Keys stored in the README.yaml files of simulations.}
-    \label{tab:READMEkeys}
-\end{table}
-
-In order to evaluate the quality of simulations, sets of C-H bond order parameters from NMR and from factors from x-ray scattering are included in the NMRlipids databank (\url{https://github.com/NMRLipids/Databank/tree/main/Data/experiments}).
-The required information for an experimental dataset are listed in table~\ref{tab:READMEkeysEXP}.
-A simulation is connected to a experimental data set
-%C-H bond order parameter and x-ray scattering form factor data using the searchDATABANK.py script.
-Available experimental data is collected to  where each dataset is saved into a folder with the name based on DOI of the related publication and experimental conditions are described in README.yaml files with the keys listed in table~\ref{tab:READMEkeysEXP}.
-if molar concentrations of all molecules are within $\pm$5 percentage units, charged lipids have the same counterions, and temperature is within $\pm$2 degrees.
-In such cases, a simulation and experimental data are paired by adding the path to the experimental data into the simulation README.yaml file.
-%If lipid to water ratio is above ??, the systems are considered fully hydrated. Otherwise also hydration level is considered.
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{  p{5.0cm}  p{10.0cm}}
-    \toprule
-    key & description \\
-    \midrule
-    DOI & DOI of the publication related to the experimental data. \\
-    TEMPERATURE & Temperature of the experiment. \\
-    MOLAR\_FRACTIONS & Dictionary of molar fractions of bilayer components \\
-    ION\_CONCENTRATIONS & Dictionary of ion concentrations of the system \\% (defined as ??) \\
-    TOTAL\_LIPID\_CONCENTRATION & Total concentration of lipid components. If exact concentration is not known, but experiments are performed in excess water, 'full hydration' can be given. \\
-    COUNTER\_IONS & Type of counter ions if present.
-\end{tabular}
-    \caption{Keys stored in the README.yaml files of experiments.}
-    \label{tab:READMEkeysEXP}
-\end{table}
-
-Because README.yaml contains all the essential information on each simulation, arbitrary analyses can be automatically performed over all the simualtions in the NMRlipids databank.
-In addition to the order parameters for each C-bonds and x-ray scattering form factors used in the quality evaluation, the NMRlipids databank contains the area per lipid and thickness calculated from all simulations in the databank.
-These results are stored in the same folders as the README.yaml files in \url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations}.
+The core content of the NMRlipids databank is stored in README.yaml files located at {\it /Data/Simulations} in the NMRlipids databank repository.
+Essential information for each simulation entry in the databank is stored in these files.
+These yaml files are both human and machine readable, and contain access to all information that are needed for further analyses of the simulations.
+The keys available in these files and their explanations are listed in table~\ref{tab:READMEkeys} in the supplementary information.
+To keep the databank light, the raw MD simulation data is stored in an external location which can be any stable publicly available repository, although all the current data locates in Zenodo \url{www.zenodo.org}.
+Whenever needed, the raw data can be downloaded through the links stored in the README.yaml files.
 
 \subsection{Molecule and atom naming convention} \label{naming}
+Molecule and atom names used in a simulation are often needed when analyzing the simulation trajectories.
+However, the naming conventions of lipid molecules and atoms therein vary between authors and force fields because universal naming convention has not been defined.
+Because such convention is necessary for automatic analyses over simulations, we have defined the unique naming conventions for lipid molecules and atoms in the NMRlipids databank.
+For each molecule we have defined an unique abbreviation which are listed in table~\ref{tab:abbreviations} in the supplementary information.
+On the other hand, unique names for each atom in a molecule are defined and connected to corresponding names in a simulation using mapping files.
+Mapping files are in yaml format where unique atom names are connected to force field specific names.
+Also residue names and fractions lipids (headgroup, glycerol bakcbone, and acyl chains) can be defined in the mapping file.
+For example, the mapping files used in the NMRlipids databank are located at {\it /Scripts/BuildDatabank/mapping\_files} in the NMRlipids repository.
+For more details about mapping files, see~\url{https://nmrlipids.blogspot.com/2022/04/new-yaml-format-of-mapping-files.html}).
+In practise, information about molecule and atom names are given in the COMPOSITION dictionary in README.yaml files, where force field specific molecule name corresponding the unique abbreviation (table~\ref{tab:abbreviations}) and the name of the mapping file are given.
 
-Because universal convention for lipid molecules and atoms therein has not been defined, the naming conventions vary between authors and force fields.
-To enable automatic analyses over large sets of simulation in the NMRlipids databank, we have defined unique naming conventions for lipid molecules and atoms.
-The abbreviations of molecule names used in the NMRlipids databank are listed in table~\ref{tab:abbreviations}.
-The unique atom names for each molecule and corresponding names in each simulation are defined using mapping files introduced in the NMRlipids project (\url{https://nmrlipids.blogspot.com/2022/04/new-yaml-format-of-mapping-files.html}).
+%Upon addition of a new entry in the databank, the molecule names in the simulation corresponding the unique names and mapping files (available at \url{https://github.com/NMRLipids/Databank/tree/main/Scripts/BuildDatabank/mapping_files}) are defined in the dictionary.
 
-Molecule and atom names in each simulation are connected to the unique naming convention with the COMPOSITION dictionary in the README.yaml file.
-Upon addition of a new entry in the databank, the molecule names in the simulation corresponding the unique names and mapping files (available at \url{https://github.com/NMRLipids/Databank/tree/main/Scripts/BuildDatabank/mapping_files}) are defined in the dictionary.
-The numbers of molecules in the system are then automatically calculated by the NMRlipids databank codes (AddData.py in Fig.~\ref{DatabankStructure}) and stored in the README.yaml together with other content of the COMPOSITION dictionary.
-This information can be then used to find each molecule and atom from each simulation in the analysis codes.
+\subsection{Adding data into the NMRlipids databank}
+In order to add a simulation entry into the NMRlipids databank, an initial info.yaml file containing user given information listed in table~\ref{tab:READMEkeys} needs to be created.
+For examples of these files, see {\it /Scripts/BuildDatabank/info\_files} in the NMRlipids databank repository.
+Using this file as an input, the {\it /Scripts/BuildDatabank/AddData.py} script then calculates the rest of the information to be stored in the README.yaml files, also listed in table~\ref{tab:READMEkeys}.
+The databank is open for submission of these info.yaml files via pull requests to the git repository.
+Large fraction of the current content is composed of the data uploaded to Zenodo in previous NMRlipids projects~\cite{botan15,catte16,antila19,bacle21}, yet also other simulation data from the Zenodo repository are added.
 
-\begin{table}[h]
-    \centering
-    \begin{tabular}{c|c}
-        Abbreviation & Molecule name \\
-        \hline
-        POPC &  1-palmitoyl-2-oleoyl-sn-glycero-3-phosphocholine\\
-        POPG &  1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoglycerol \\
-        POPS & 1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-L-serine \\
-        POPE & 1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoethanolamine \\
-        CHOL & cholesterol \\
-        DHMDMAB & dihexadecyldimethylammonium \\
-        \hline
-        POT & potassium ion \\
-        SOD & sodium ion \\
-        CLA & chloride ion \\
-        CAL & calcium ion \\
-        SOL & water \\
-    \end{tabular}
-    \caption{Abbreviations used in the databank}
-    \label{tab:abbreviations}
-\end{table}
+\subsection{Experimental data}
+Experimental data used in the quality evaluation, currently composed of C-H bond order parameters and x-ray scattering form factors, are stored in {\it /Data/experiments} in the NMRlipids databank repository.
+Similarly to simulations, each experimental data set has a README.yaml file containing all the relevant information about the experiment.
+The keys and their descriptions for the experimental data are given in table~\ref{tab:READMEkeysEXP}.
+NMR data currently in the NMRlipids databank are taken from Refs.~\citenum{??} and x-ray scattering data from Refs.~\citenum{??}.
+In addition, some previously unpublished NMR and x-ray scattering data are used.
+These are measured with previously established methods as described in the supplementary information.
 
-\subsection{Quality evaluation}
-
-\subsubsection{Conformational ensembles using C-H bond order parameters}
-The quality of conformational ensembles of individual lipid molecules in simulations are evaluated using the C-H bond order parameters which can be directly compared with robust experimental data~\cite{ollila16}.
-%The C--H bond order parameters were calculated directly
-%from the carbon and hydrogen positions using the 
-%definition
-The order parameters are defined as
+\subsection{Calculation of C-H bond order parameters}
+The C--H bond order parameters were calculated directly from the carbon and hydrogen positions using the definition
+%The order parameters are defined as
 \begin{equation}\label{OP}
@@
 %(taken to align with $z$, with bilayer periodicity in the $xy$-plane).
-and angular brackets denote the ensemble average, i.e. average over all sampled configurations of all lipids in a simulation.
-%by estimating
-%the average probability of C-H bond order parameters to agree with the values from NMR experiments. 
-%As a first step, the quality of each order parameter is evaluated using the negative logarithm of probability to find the simulated order parameter within the experimental error bars
-%\begin{equation}
-%    S_q = -\log_{10}(P),    
-%\end{equation}
-Because conformational ensembles of individual lipids are independent in a simulation of a fluid lipid bilayer, the number of independent sample points for each C-H bond, $n$, equals the number of lipids in a simulation.
-Denoting the $S_{\rm CH}$ from Eq.~\ref{OP} as the sample mean and $s$ its variance calculated over individual lipids,
-%the probability the true mean of the order parameter is given by the %Student's t-distribution 
-%Because each C-H bond order parameter is calculated as an average over individual lipids in a simulation
-%\begin{equation}\label{studentT}
-$\frac{S_{\rm CH}-\mu}{s/\sqrt{n}}$
-%\end{equation}
-%where $\bar{X}$ is the sample mean (order parameter calculated from simulations),
-%$s$ is the sample variance (variance of order parameters in different lipids)
-%and 
-%has a 
-has a Student's t-distribution with $n-1$ degrees of freedom, where $\mu$ is the real mean of the order parameter. 
-%shifted by $S_{\rm CH}$ and scaled by $s/\sqrt{n}$.
-Because lipid bilayer simulations contain at least dozens of lipids, the Student's t-distribution could be safely approximated with a normal distribution.
-However, when calculating the probability for a simulation to locate within experimental order parameters, the normal distribution gives values below the numerical accuracy of computers for simulations values far from experiments.
-To avoid such numerical instability, we use the first order Student's t-distribution.% for the equation \ref{studentT}.
-Therefore, we estimate the probability for the simulation order parameter to locate within experimental error bars from equation
+and angular brackets denote the ensemble average, i.e., average over all sampled configurations of all lipids in a simulation.
+As in previous NMRlipids publications, the order parameters were first calculated separately for each lipid and the standard error of the mean over different lipids was used as the error bar~\cite{botan15}.
+
+In practise, the analysis code loops over all README.yaml files (i.e., simulations in the NMRlipids databank), downloads the raw simulation to a local computer, uses the information about the atom and molecule naming conventions in README.yaml to determine the location of carbon and hydrogen atoms, and then calculates the order parameters from Eq.~\ref{OP}.
+The script to calculate all C-H bond order parameters from the NMRlipids databank is available at {\it /Scripts/AnalyzeDatabank/calcOrderParameters.py} in the NMRlipids databank repository.
+The resulting order parameters are stored for all simulations in files named \newline
+[{\it lipid\_name}]{\it OrderParameters.json} at folders in {\it /Data/Simulations} in the NMRlipids databank repository.
+
+\subsection{Calculation of x-ray scattering form factors}
+X-ray scattering form factors were calculated with standard equation for symmetric lipid bilayers~\cite{ollila16}
+\begin{equation}\label{FFsimpl}
+F(q)=\int_{-D/2}^{D/2}\Delta \rho_e(z) \cos(zq_z) {\rm d}z,
+\end{equation}
+where $\Delta \rho_e(z)$ is the difference between total and solvent electron densities, and $D$ is the simulation box size in the z-direction.
+For the calculation of density profiles, atom coordinates were centred around the centre of mass of the POPC lipid molecules for every time frame, and a histogram of these centred positions was calculated with the bin width of $1/3$~Å.
+Similarly to order parameters, the script loops over all simulations, downloads the data on local computer, and uses the information in README.yaml file to calculate form factors.
+The script to calculate form factors for all simulations in the NMRlipids databank is available at {\it Scripts/AnalyzeDatabank/calc\_FormFactors.py}.
+The resulting form factors are stored for all simulations in files named {\it FormFactor.json} at folders in {\it /Data/Simulations} in the NMRlipids databank repository.
+
+\subsection{Calculation of membrane thickness, and area per lipids}
+Lipid bilayer thicknesses in all simulations are calculated from the intersections of lipid and water electron densities using the script {\it Scripts/AnalyzeDatabank/calc\_thickness.py} at the NMRlipids databank repository.
+The resulting thicknesses are stored for all simulations in files named {\it thickness.json} at folders in {\it /Data/Simulations} in the NMRlipids databank repository.
+Lipid bilayer area per lipids in all simulations are calculated by dividing the simulation box area with the total number of molecules considered as lipids using the script {\it Scripts/AnalyzeDatabank/calcAPL.py} at the NMRlipids databank repository.
+The resulting area per lipids are stored for all simulations in files named {\it apl.json} at folders in {\it /Data/Simulations} in the NMRlipids databank repository.
+
+\subsection{Quality evaluation of C-H bond order parameters}
+As the first step to evaluate simulation qualities against experimental data, each simulation is connected to an experimental data if molar concentrations of all molecules are within $\pm$5 percentage units, charged lipids have the same counterions, and temperatures are within $\pm$2 degrees.
+For molar concentrations of water, the exact hydration level is considered only for systems with molar water to lipid ratio below 25, otherwise the systems are considered as fully hydrated.
+In practise, this is implemented by adding the path of the experimental data into the simulation README.yaml file using the {\it /Scripts/BuildDatabank/searchDATABANK.py} script in the NMRlipids databank repository.
+
+%\subsubsection{Conformational ensembles using C-H bond order parameters}
+%
+%The quality of conformational ensembles of individual lipid molecules in simulations are evaluated using the C-H bond order parameters which can be directly compared with robust experimental data~\cite{ollila16}.
+
+The quality of each C--H bond order parameter is estimated by calculating the probability of a simulated value to locate within the error bars of the experimental value.
+Because conformational ensembles of individual lipids are independent in a fluid lipid bilayer,
+$\frac{S_{\rm CH}-\mu}{s/\sqrt{n}}$ has a Student's t-distribution with $n-1$ degrees of freedom and the probability for a order parameter in simulation to locate within experimental error bars can be estimated from equation
 \begin{equation}\label{probability}
@@
 \end{equation}
-where f(t) is the first order Student's t-distribution.
-% shifted to $\bar{X}$ and scaled with $s$.
+where f(t) is the %first order 
+Student's t-distribution, $\mu$ is the real mean of the order parameter, $n$ is the number of independent sample points for each C-H bond which equals the number of lipids in a simulation, $S_{\rm CH}$ is the sample mean from Eq.~\ref{OP}, $s$ is the variance of $S_{\rm CH}$ calculated over individual lipids, $S_{\rm exp}$ is the experimental value, and $\Delta S_{\rm exp}$ its error.
+The error of $\Delta S_{\rm exp} = 0.02$ is currently assumed for all experimental order parameters~\cite{ollila16}.
+Because lipid bilayer simulations contain at least dozens of lipids, the Student's t-distribution could be safely approximated with a normal distribution.
+However, the normal distribution gives probability values that are below the numerical accuracy of computers when simulation values are far from experiments.
+To avoid such numerical instabilities, we use the first order Student's t-distribution having slightly higher probabilities for values far away from the mean.
+%Because phospholipids sample their conformational ensemble within nanosecond timescale~\cite{ferreira15}, all simulations in the databank would be sufficiently long to sample the realistic conformational phase of individual lipids.
+On the other hand, some force fields exhibit too slow dynamics which leads to large error bars in order parameter values~\cite{antila21a}.
+Such large error bars widen the Student's t-distribution in Eq.~\ref{probability} thereby artificially increasing the probability to find the simulated value within experimental error bars.
+Therefore, the order parameters with simulation error bars above the experimental error 0.02 are not included in the quality evaluation.
 
@@
 The quality measure $S_q$ approaches to zero when probability for agreement between simulation and experimental results approach one, and increases when simulated and experimental values diverge.
-The error of $\Delta S_{\rm exp} = 0.02$ is currently assumed for all experimental order parameters~\cite{ollila16}.
-Because phospholipids sample their conformational ensemble within nanosecond timescale~\cite{ferreira15}, all simulations in the databank would be sufficiently long to sample the realistic conformational phase of individual lipids.
-However, some force fields exhibit too slow dynamics which leads to large error bars in order parameter values~\cite{antila21a}.
-Because large error bars widen the Student's t-distribution in Eq.~\ref{probability} thereby artificially increasing the probability to find the simulated value within experimental error bars, the order parameters with error bars larger than the experimental error 0.02 are not included in the quality evaluation.
 
-The qualities of different fragments within each lipid type in a simulation, $P^{\rm frag}[{\rm lipid}]$, are then estimated by averaging the probabilities for individual C-H bond order parameters to locate within experimental errors according to Eq.~\ref{probability}, and dividing the average with the percentage of order parameters for which the quality is available within the fragment, $p_{\rm frag}[{\rm lipid}]$,
+To streamline the comparison between simulations, we define the qualities of different fragments (headgroup, acyl chains or total lipid) within each lipid type in a simulation as
 \begin{equation}
@@
 \end{equation}
- where frag can be {\it sn}-1, {\it sn}-2, headgroup or total (all order parameters within a molecule).
+where $P^{\rm frag}[{\rm lipid}]$ is the average of individual C-H bond order parameters qualities within the fragment, $p_{\rm frag}[{\rm lipid}]$ is the percentage of order parameters for which the quality is available within the fragment, and frag can be {\it sn}-1, {\it sn}-2, headgroup or total (all order parameters within a molecule).
 The overall quality of different fragments in a simulation are then defined as a molar fraction weighted average over different lipid components
@@
 
-\subsubsection{Membrane dimensions with x-ray scattering form factors}
-While C-H bond order parameters relate to the conformational ensembles of individual lipids,
-the x-ray scattering form factors depend on membrane dimensions and density distribution~\cite{ollila16}.
-Because experiments give form factors only in relative intensity scale, they are typically scaled to the simulation data.
-Here we use the same scaling method as in SIMtoEXP program~\cite{kucerka10} where experimental form factor intensities
-are scaled by a factor defined as
+The quality evaluation of order parameters is implemented in {\it /Scripts/BuildDatabank/QualityEvaluation.py} in the NMRlipids databank repository.
+The resulting qualities for each order parameters are stored in files named  [{\it lipid\_name}]{\it \_OrderParameters\_quality.json}, for individual lipids in files named [{\it lipid\_name}]{\it \_FragmentQuality.json}, and for overall quality for fragments in files named {\it system\_quality.json} at folders in {\it /Data/Simulations} in the NMRlipids databank repository.
+
+\subsection{Quality evaluation of x-ray scattering form factors}
+%While C-H bond order parameters relate to the conformational ensembles of individual lipids, the x-ray scattering form factors depend on membrane dimensions and density distribution~\cite{ollila16}.
+Because experiments give form factors only in relative intensity scale, they should scaled before comparing with the simulation data.
+Here we use the scaling coefficient for experimental intensities defined in the SIMtoEXP program~\cite{kucerka10}
 \begin{equation}
@@
 %\end{equation}
-where $F_s$ and $F_e$ are form factors from a simulation and experiment, respectively, and summation goes over the experimentally available $N_q$ points.
-Quality measure based on differences in simulated and experimental form factors accross available q-range is also defined in the SIMtoEXP program.
-However, as shown in Fig.~\ref{fig:sizedependence}, the lobe heigths in the simulated form factors depend on the simulation box size, but locations of minima does not.
-Therefore, quality measure depending on lobe heights is also dependent on simulation box size.
-In order to define a quality measure which is independent on simulation box size, we measure the form factor quality based on the location of the first minima observed in experiments.
-As shown in Fig.~\ref{fig:quality}, this minima correlates well with the thickness of a membrane.
-Although the correlation in second minima would be even stronger (Fig.~\ref{fig:QualityCorrelationsSI}), we have chosen to use the first minima because automatic definition of second minima is inaccurate in some experimental data.
+where $F_s(q)$ and $F_e(q)$ are form factors from a simulation and experiment, respectively, $\Delta F_e(q)$ is the error of the experimental form factor, and summation goes over the experimentally available $N_q$ points.
+
+Also a quality measure based on differences in simulated and experimental form factors accross available q-range is defined in the SIMtoEXP program~\cite{kucerka10}.
+However, the lobe heigths in simulated form factors depend on the simulation box size as shown in Fig.~\ref{fig:sizedependence}.
+Therefore, the quality measure defined in SIMtoEXP  would also depend on the simulation box size.
+Nevertheless, locations of form factor minima are independent on simulation box size in Fig.~\ref{fig:sizedependence}.
+Here we use the location of the first form factor minima for the quality evaluation because automatic detection of the location of second minima is inaccurate for some experimental data due to fluctuations, such as for the POPE data in Figs.~\ref{fig:quality} D and E.
+%In order to define a quality measure which is independent on simulation box size, we measure the form factor quality based on the location of the first minima observed in experiments. 
+The first minimum correlates well with the thickness of a membrane (Fig.~\ref{fig:quality} F), although the correlation of the second minima would be even stronger (Fig.~\ref{fig:QualityCorrelationsSI}).
 In practise, we first filter the fluctuations from the form factor data using Savitzky-Golay filter (window lenght 30 and polynomial order 1) and locate the first minima above 0.1\,$AA^2$ from both simulation and experimental data.
@@
 
-\subsection{Analysing simulations in the NMRlipids databank}
+The quality evaluation of form factors is implemented in {\it /Scripts/BuildDatabank/QualityEvaluation.py} in the NMRlipids databank repository.
+The resulting form factor qualities are stored in files named {\it FormFactorQuality.json} at folders in {\it /Data/Simulations} in the NMRlipids databank repository.
 
-The README.yaml files contain all the essential information to perform arbitrary analyses of simulations in the NMRlipids databank, i.e., the permanent location of the original data and naming convention for all atoms and molecules in each system.
-In practise, the analyse codes contains a loop over all README.yaml files (i.e., simulations in the NMRlipids databank) which first downloads the raw simulation to a local computer and then uses the information about the atom and molecule naming conventions in README.yaml and mapping files to perform the desired analyses.
-For example, the code that calculates all C-H bond order parameters of all systems is available at \url{https://github.com/NMRLipids/Databank/blob/main/Scripts/AnalyzeDatabank/calcOrderParameters.py} and minimal example of a analysis code is available at \url{https://github.com/NMRLipids/Databank/blob/main/Scripts/AnalyzeDatabank/template.ipynb}.
+\subsection{Analysing anisotropic diffusion of water in membranes environment from the NMRlipids databank}
 
-While the order parameters, form factors, area per lipid and thickness are stored within the NMRlipids databank (\url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations}), further analyses can be conventiently stored in separate repositories with the same folder structure based on hash identities of trajectory and topology files.
+%The README.yaml files contain all the essential information to perform arbitrary analyses of simulations in the NMRlipids databank, i.e., the permanent location of the original data and naming convention for all atoms and molecules in each system.
+%In practise, the analyse codes contains a loop over all README.yaml files (i.e., simulations in the NMRlipids databank) which first downloads the raw simulation to a local computer and then uses the information about the atom and molecule naming conventions in README.yaml and mapping files to perform the desired analyses.
+
+%A minimal example of a analysis code is available at \url{https://github.com/NMRLipids/Databank/blob/main/Scripts/AnalyzeDatabank/template.ipynb}.
+
+Water permeability through membranes was calculated from equation $P=r/2c_w$, where $r$ is the rate of permeation events per time and area, and $c_w$ is the concentration of water in bulk with the value of 33.3679\,(nm)$^{-3}$~\cite{venable19}.
+The number of permeation events in each trajectory was calculated using the code by Camilo et al.~\cite{camilo2022}, available at~\url{https://github.com/crobertocamilo/MD-permeation}.
+This code was used within the loop that goes over all simulations in the NMRlipids databank and extracts the required information using the information in README.yaml files.
+The code is available at {\it /scripts/calcMD-PERMEATION.py} and resulting permeabilities are stored at {\it /Data/MD-PERMEATION} in the repository containing all analyses specific for this publication at~\url{https://github.com/NMRLipids/DataBankManuscript/}.
+%While the order parameters, form factors, area per lipid and thickness are stored within the NMRlipids databank (\url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations}), further analyses can be conventiently stored in separate repositories with the same folder structure based on hash identities of trajectory and topology files.
 For example, results from further analyses performed here are stored in folders at \url{https://github.com/NMRLipids/DataBankManuscript/tree/main/Data}.
-Such organization of the data enables further upcycling of the analyzed data as similarly to the original NMRlipids databank repository.
+This repository is organized similarly to the NMRlipids databank repository, enabling the upcycling also of the analyzed data without overloading the main NMRlipids databank repository.
 
-For the permeation of water through membranes, the number of permeation events in each trajectory was first calculated using the code by Camilo et al.~\cite{camilo2022}, available at~\url{https://github.com/crobertocamilo/MD-permeation}.
-The permeation was then calculated from equation $P=r/2c_w$, where $r$ is the rate per time and area, and $c_w$ is the concentration of water in bulk with the value of 33.3679\,(nm)$^{-3}$~\cite{venable19}.
+The parallel diffusion along membrane surface was calculated using Einstein's equation with {\it gmx msd} program with the {\it -lateral} option from Gromacs softwater package~\cite{gromacsMANUAL}.
+The code is available at {\it /scripts/calcWATERdiffusion.py} and resulting diffusion coefficients are stored at {\it /Data/WATERdiffusion} in the repository at~\url{https://github.com/NMRLipids/DataBankManuscript/}.
 
-%\subsection{Indexing the simulation data}
+Water diffusion in the perpendicular direction of lipid bilayers in a multilamellar stack was estimated from the Tanner equation $D_\perp = \frac{D_{||} P z_w}{D_{||} + P z_w}$~\cite{tanner78,wasterby02}, where the water layer thickness, $z_w$, was estimated by subtracting bilayer thickness from the size of the simulation box in membrane normal direction.
 
-%\subsubsection*{AddData.py}
-%AddData.py is a script that builds a database that contains a dictionary file and analysis data of each simulation.
-The dictionary file contains information about the simulation.
-The script also calculates order parameters of all CH bonds of the lipids in the simulation.
-To add a simulation it must be first uploaded to Zenodo (www.zenodo.org).
-The trajectory and topology files of the simulation are downloaded to the working directory from Zenodo but these are not saved into the database.
-To add a simulation to the database the user has to give some essential information about the simulation.
-This is done by writing a info file (*INFO.yaml) which is passed to AddData.py.
-%\newline \\
-%AddData.py requires GROMACS and MDAnalysis library to be installed.
-%\newline \\
-%Run AddData.py as follows:
-%\newline \\
-%python3 AddData.py -f exampleINFO.yaml
-%\newline \\
-%There is also a script runAddData.sh that can be used to loop over several info files to add many simulations at one go.
-
-%\subsubsection*{Simulation dictionary}
-%A simulation dictionary contains information about the simulation.
-Some of the information is provided by the user.
-The numbers of lipid molecules, solvent and ions are automatically read from the files and so are the simulation temperature and trajectory length.
-All this information is saved to a file named README.yaml.
-
-%\subsubsection{User input}
-%Dictionary variables defined by the user are listed here. 
-%The necessary parameters required for the analyses are described first and are marked as ''compulsory''. Then the parameters used to describe the data for further usage and analyses are described.
-%While only compulsory parameters are required for the databank entry, strong recommendation is that all the possible parameters would be given to maximize the possibilities to upcycle the contributed data.
-
-%\subsubsection*{DOI (compulsory)}
-%Give DOI identity from where the simulation files are located.
-Current databank works only for the data in Zenodo, but other potential sources are may be implemented in the future.
-Note that the DOI must point to a specific version of dataset in Zenodo, i.e., DOI pointing to all versions of certain dataset does not work.
-
-%\subsubsection*{SOFTWARE (compulsory)}
-%Give the name of software used to run the simulation. The options are GROMACS, AMBER, NAMD, CHARMM and OPENMM. So far, only simulations run with GROMACS are accepted by the script.
-
-%\subsubsection*{TRJ (compulsory)}
-%Give the name of the trajectory file that is found from the DOI given above.
-
-%\subsubsection*{TPR (compulsory)}
-%Give the name of the file with topology information (tpr file in the case of Gromacs) that is found from the DOI given above.
-
-%\subsubsection*{PREEQTIME (compulsory)}
-%Give the time simulated before the uploaded trajectory in nanoseconds. For example, if you upload 100-200~ns part of total 200~ns simulation, this should value should be 100.
-
-%\subsubsection*{TIMELEFTOUT (compulsory)}
-%Give the time that should be considered as an equilibration period in the uploaded trajectory.
-Frames before the give time will be discarded in the analysis.
-For example, if you upload 0-200~ns part of total 200~ns simulation where the first 100~ns should be considered as an equilibration, this value should be 100.
-
-%\subsubsection*{Molecule names (compulsory)}
-%In the databank, each molecule has a unique abbreviation (see table~\ref{tab:abbreviations}). 
-%You need to give the residue names of these molecules corresponding your simulation.
-%The user must also provide the names of the molecules, ions and solvent that are used in the simulation to match the names used by the databank.
-The names provided by the user must be the same as in the tpr file.
-%If atoms in a lipid belong to different residues (typical situation in Amber force fields), give the name of the head group residue here, and add the residue name of each atom to the third column in the mapping file (see below).
-If your simulation contains molecules that are not yet in the databank, you need to define the abbreviation and add molecules to the lipids\_dict, molecules\_dict, molecule\_numbers\_dict and molecule\_ff\_dict in the AddData.py script, as well as to table~\ref{tab:abbreviations}.
-
-%\subsubsection*{MAPPING\_DICT (compulsory)}
-%For the analysis, we need to know the names of atoms in your system.
-%These are defined using the mapping file convention introduced in \url{http://nmrlipids.blogspot.com/2015/03/mapping-scheme-for-lipid-atom-names-for.html}, where first column gives the universal atom name, second gives the atom name in force field, and third gives the residue name if not the same for all atoms.
-%The name of the mapping file for each molecule needs to be given in dictionary format.
-The dictionary the key is the molecule name (abbreviation listed in table \ref{tab:abbreviations}) and the value is the name of the mapping file.
-The already existing mapping files can be found from the directory named "mapping\_files".
-If a mapping file for your molecule(s) do not exist, you need to construct one and add to the "mapping\_files" directory.
+\subsection{Calculation of lipid flip-flops}
 
@@
 If the lipid consists of several residues which is the case in some AMBER force fields, then a third column is needed which contains the name of the residue to which each atom belongs to.
-
-%\subsubsection*{DIR\_WRK (compulsory)}
-%Give the path of the working directory in your local computer.
-The trajectory and topology files will be downloaded to this trajectory, and temporary files created during processing will be stored here.
-
-%\subsubsection*{UNITEDATOM\_DICT (compulsory for united atom trajectories)}
-%Order parameters from united atom simulations are calculated using buildH code (\url{https://github.com/patrickfuchs/buildH}).
-For united atom simulations, you need to tell how hydrogens are added based on definitions in the dic\_lipids.py dictionary.
-This is done by giving a dictionary where the key is the molecule name (abbreviation listed in table \ref{tab:abbreviations}) and the value is the correct dictionary key in dic\_lipids.py.
-If correct dictionary key is not yet found, you need to add to dic\_lipids.py.
-%In the case of an all atom simulation, UNITEDATOM is left empty.
-
-%\subsubsection*{PUBLICATION}
-%Give reference to a publication(s) related to the data.
-
-%\subsubsection*{AUTHORS\_CONTACT}
-%Give the name and email of the main author(s) of the data.
-
-%\subsubsection*{SYSTEM}
-%Give description of system in free format. For example ''POPC with cholesterol at 301K''.
-
-%\subsubsection*{SOFTWARE\_VERSION}
-%Give the version of the software used.
-
-%\subsubsection*{FF}
-%Give the name of the force field used used in the simulation.
-
-%\subsubsection*{FF\_SOURCE}
-%Describe the source of the force field parameters. For example, CHARMM-GUI, link to webpage where parameters were downloaded, or citation to a paper.
-
-%\subsubsection*{FF\_DATE}
-%Give the date when parameters were accessed or created. The format is day/month/year.
-
-%%\subsubsection*{Individual force field names for molecules}
-%In some cases special force fields are used for certain molecules.
-For example, non-standard parameters for ions or other molecules have been used.
-These can be specified giving forcefield names separately for individual molecules.
-These can be given as parameters named as
-%FF+[abbreviation from table \ref{tab:abbreviations}], i.e., FFPOPC, FFPOT, FFSOL etc.
-
-%\subsubsection*{CPT (Gromacs)}
-%Give the name of the Gromacs checkpoint file that is found from the DOI given above.
-%CPT stands for the name of the cpt file.
-
-%\subsubsection*{LOG (Gromacs)}
-%Give the name of the Gromacs log file that is found from the DOI given above.
-
-%\subsubsection*{TOP (Gromacs)}
-%Give the name of the Gromacs top file that is found from the DOI given above.
-
-%\subsubsection{Automatically analyzed parameters}
-%The following parameters are read automatically from the trajectory and topology files.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%automatically analyzed parameters%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\subsubsection*{Molecule numbers}
-%Numbers of lipid molecules (NPOPC, NPOPG, etc.) per membrane leaflet are calculated by determining on which side of the center of mass of the membrane the center of mass of the head group of each lipid molecule is located.
-%\newline \\
-%\noindent Numbers of other molecules such as solvent and ions (NSOL, NPOT, NSOD, etc.) are read from the topology file.
-
-%\subsubsection*{Temperature}
-%Temperature of the simulation is read from the topology file.
-
-%\subsubsection*{Trajectory length}
-%The length of a trajectory is read from the trajectory file.
-
-%\subsubsection*{Date of running}
-%The date of running the script is saved to the README.yaml file.
 

### 2022-10-12 SI.tex
Update text

@@
     The figures and numbers are created on 9th of May 2022.}
-    \label{fig:overlay}
+    \label{fig:permeationSI}
 \end{figure}
@@
     The figures and numbers are created on 9th of May 2022.}
-    \label{fig:overlay}
+    \label{fig:diffusionSI}
 \end{figure}
+
+\section{Databank content}
+
+\begin{figure}
+  \includegraphics[width=\textwidth]{Figures/DataBankChart.pdf}
+  \caption{Structure of the NMRlipids databank.
+Manually added input data (blue boxes) includes basic information on the simulation, permanent links to the raw data, and experimental data if available.
+The databank entries (red box) and analysis results (green boxes), locating at \url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations} are automatically generated by the computer programs included in the NMRlipids databank (yellow boxes).
+Because raw data are not permanently stored but can be accessed based on the information in the databank, this connection is marked with the dashed line.
+}\label{DatabankStructure}
+\end{figure}
+
+\begin{table}[]
+    \centering
+    \begin{tabular}{  p{3.5cm}  p{9.5cm}  p{4.0cm} }
+    \toprule
+    key & description & type  \\
+    \midrule
+        %\hline
+    DOI & DOI from where the raw data is found & user given (compulsory) \\
+    SOFTWARE & Software used to run the simulation (e.g. Gromacs, Amber, NAMD, etc.) & \\
+    TRJ & Name of the trajectory file found from DOI & \\
+    TPR & Name of the topology file found from DOI (trp file in the case of Gromacs) & \\
+    PREEQTIME & Pre-equilibrate time simulated before the uploaded trajectory in nanoseconds.
+\tablefootnote{For example, if you upload 100-200 ns part of total 200 ns simulation, this should value should be 100.} & \\
+    TIMELEFTOUT & Equilibration period in the uploaded trajectory that should be discarded in analyses.
+\tablefootnote{For example, if you upload 0-200 ns part of total 200 ns simulation where the first 100 ns should be considered as an equilibration, this value should be 100.} \\
+    COMPOSITION & Molecules names used in the simulation and corresponding mapping files (see section \ref{naming}) & \\
+    DIR\_WRK & Temporary working directory in your local computer. \\
+    UNITEDATOM\_DICT & Information for constucting hydrogens for united atom simulations, empty for all atom simulations & \\
+    TYPEOFSYSTEM & Lipid bilayer or something else & \\
+    \hline
+    PUBLICATION & Give reference to a publication(s) related to the data. & User given (optional)\\
+    AUTHORS\_CONTACT & Name and email of the main author(s) of the data. & \\
+    SYSTEM & System description on free text format & \\
+    SOFTWARE\_VERSION & Version of the used software & \\
+    FF & Name of the used force field & \\
+    FF\_SOURCE & Source of the force field parameters, e.g, CHARMM-GUI, webpage, citation to a publication, etc. & \\
+    FF\_DATE &  Date when force field parameters were accessed on the gives source (day/month/year). & \\
+    FF{molename} & Molecule specific force field information, e.g., water model with FFSOL and sodium parameters with FFSOD. & \\
+    CPT & Name of the Gromacs checkpoint file. & \\
+    LOG & Name of the Gromacs log file. & \\
+    TOP & Name of the Gromacs top file. & \\
+    GRO & Name of the Gromacs gro file. & \\
+    \hline
+    TRAJECTORY\_SIZE & Size of the trajectory file in bytes & automatically extracted data. \\
+    TRJLENGTH & Lenght of the trajectory (ps). & \\
+    TEMPERATURE & Temperature of the simulation. & \\
+    NUMBER\_OF\_ATOMS & Number of atoms in the simulation. & \\
+    DATEOFRUNNIG & Date when added into the databank & \\
+    EXPERIMENT & Potentially connected experimental data & \\
+    COMPOSITION & Numbers of lipid molecules (NPOPC, NPOPG, etc.) per membrane leaflet are calculated by determining on which side of the center of mass of the membrane the center of mass of the head group of each lipid molecule is located.
+    Numbers of other molecules such as solvent and ions (NSOL, NPOT, NSOD, etc.) are read from the topology file. & \\
+    \end{tabular}
+    \caption{Keys stored in the README.yaml files of simulations.}
+    \label{tab:READMEkeys}
+\end{table}
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{c|c}
+        Abbreviation & Molecule name \\
+        \hline
+        POPC &  1-palmitoyl-2-oleoyl-sn-glycero-3-phosphocholine\\
+        POPG &  1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoglycerol \\
+        POPS & 1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-L-serine \\
+        POPE & 1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoethanolamine \\
+        CHOL & cholesterol \\
+        DHMDMAB & dihexadecyldimethylammonium \\
+        \hline
+        POT & potassium ion \\
+        SOD & sodium ion \\
+        CLA & chloride ion \\
+        CAL & calcium ion \\
+        SOL & water \\
+    \end{tabular}
+    \caption{Abbreviations used in the databank}
+    \label{tab:abbreviations}
+\end{table}
+
+\begin{table}[]
+    \centering
+    \begin{tabular}{  p{5.0cm}  p{10.0cm}}
+    \toprule
+    key & description \\
+    \midrule
+    DOI & DOI of the publication related to the experimental data. \\
+    TEMPERATURE & Temperature of the experiment. \\
+    MOLAR\_FRACTIONS & Dictionary of molar fractions of bilayer components \\
+    ION\_CONCENTRATIONS & Dictionary of ion concentrations of the system \\% (defined as ??) \\
+    TOTAL\_LIPID\_CONCENTRATION & Total concentration of lipid components. If exact concentration is not known, but experiments are performed in excess water, 'full hydration' can be given. \\
+    COUNTER\_IONS & Type of counter ions if present.
+\end{tabular}
+    \caption{Keys stored in the README.yaml files of experiments.}
+    \label{tab:READMEkeysEXP}
+\end{table}
 

### 2022-10-12 main.tex

@@
 To analyze the anisotropic diffusion of water in membrane environment, we also calculated the water diffusion parallel to the membrane surface from all simulations in the NMRlipids databank.
-The results at different hydration levels are shown in Fig.~\ref{fig:permeability} H together with the experimental data~\cite{rudakova04}.
+The results at different hydration levels are shown in Fig.~\ref{fig:permeability} F together with the experimental data~\cite{rudakova04}.
 %The experimental water diffusion increases with increasing hydration level toward the value for bulk water . 
@@
 In order to estimate the diffusion anistoropy of water in multilamellar membrane system, we translated the water permeation coefficients through membranes to diffusion coefficients through multilamellar stack using the Tanner equation~\cite{tanner78,wasterby02}.
-The resulting diffusion coefficients through the membranes are approximately five orders of magnitude slower than along the membrane (Figs.~\ref{fig:permeability} F and G), which is at the upper limits of the anisotropic estimated from the experimental data~\cite{nitsche19}.
+The resulting diffusion coefficients through the membranes are approximately five orders of magnitude slower than along the membrane (Figs.~\ref{fig:permeability} G and H), which is at the upper limits of the anisotropic estimated from the experimental data~\cite{nitsche19}.
 Relatively high anisotropy is understandable as simulations give slightly slower permeation rates and higher lateral diffusion rates than experiments.
-Significant increase in the diffusion anisotropy with membrane packing is observed, as $D_{\perp}/D_{||}$ drifts away from one with decreasing area per lipid and increasing thickness in Figs.~\ref{fig:permeability} F and G.
+Significant increase in the diffusion anisotropy with membrane packing is observed, as $D_{\perp}/D_{||}$ drifts away from one with decreasing area per lipid and increasing thickness in Figs.~\ref{fig:permeability} G and H.
 This follows from decreasing water permeability with membrane packing (Figs \ref{fig:permeability} C and D), while lateral diffusion remains approximately constant (Fig.

### 2022-10-31 main.tex
Updates from Overleaf

@@
 
-Building accessible databanks of molecular dynamics simulation data has been challenging due to the required long term support for hardware and software maintenance.
-In the overlay model used in the NMRlipids databank, the demand of hardware can be distributed and open collaboration model reduces the risk for ending software maintenance.
-Furthermore, the open collaboration model used in the NMRlipids project credits contributors by offering authorship in published articles, thereby creating an incentive for contributions.
-This model could be extended also to other fields where similar barriers to establish publicly accessible databanks exist.
+Applications of MD simulations in understanding biomolecular behaviour are steadily increasing, but protocols for quality evaluation and sharing simulation data fall behind the ones in more elaborated fields, such as structural biology.
+Development of such protocols has been limited by practical challenges, such as lack of universally defined quality measures for MD simulations, incompatible naming conventions in different models and simulation softwares, and high technical requirements for hardware to share large trajectories from MD simulations, as well as incentives for scientist to evaluate and share their MD simulation data.
 
-, thereby not requiring large resources to handle and maintain.
-This lowers the barrier for starting such databank as well as for long term storage.
-The NMRlipids databank is essentially a git repository containing information on the location of raw data and its indexing with universal naming convention.
-In addition to all computers where the databank is developed and used, the NMRlipids databank git is stored to Zenodo server, thereby enabling a very cost effective long term storage for the databank.
+The NMRlipids databank circumvents these challenges by utilizing the overlay databank design and open collaboration approach developed in the NMRlipids project~\cite{botan15}.
+In the overlay databank, the storage of large simulation files is outsourced to existing publicly available repositories while databank itself is essentially a git repository with open access licence containing relevant information about the trajectories, including the location of the raw data.
+This architecture streamlines the construction and maintenance of the databank
+%Building accessible databanks of molecular dynamics simulation data has been challenging due to the required long term support for hardware and software maintenance. 
+%
+%, the demand of hardware can be distributed and open collaboration model reduces the risk for ending software maintenance. 
+%Furthermore, the 
+by distributing the long term commitment for hardware and software maintenance.
+Incentive to share the data is created in the NMRlipids open collaboration by offering authorship in published articles to the contributors.
+%, thereby creating an incentive for contributions. 
+Universal quality measures and naming conventions for molecules and atoms are defined in the NMRlipids databank to enable the automatic quality evaluation and other analyses.
+%of lipid bilayer MD simulations introduced in the NMRlipids databank enables rapid ranking of available simulation models agains experimental NMR and x-ray scattering data. 
+%
+The current NMRlipids databank contains only lipid bilayer simulations, but the concept combining open collaboration and overlay databank structure can be applied also in other fields where similar barriers to establish publicly accessible databanks exist.
+%
+%, thereby not requiring large resources to handle and maintain. 
+%This lowers the barrier for starting such databank as well as for long term storage. 
+%The NMRlipids databank  
+%In addition to all computers where the databank is developed and used, the NMRlipids databank git is stored to Zenodo server, thereby enabling a very cost effective long term storage for the databank.
 
-Quality measure and automatic quality evaluation of lipid bilayer MD simulations introduced in the NMRlipids databank enables rapid ranking of available simulation models agains experimental NMR and x-ray scattering data.
-This provides a tool for researches to rapidly evaluate the credibility of MD simulations of their own and reported by other groups.
-Because such tool and quality measure has not been available, this will elaborate the quality of published MD simulations of lipid bilyers and reduce potentially misleading results~\cite{antila22b}.
-The power of NMRlipids databank to select the best models for particular applications has been demontrated for PC/PE lipid mixtures (Fig.~\ref{fig:quality}), PC/PS lipid mixtures~\cite{antila22b}, and lipid headgroups~\cite{bacle21}.
+The NMRlipids databank will be of great benefit to scientists involved in MD simulations.
+They will be able to rapidly evaluate the quality of MD simulations in order to filter out potentially misleading results or facilitate force field parameter development~\cite{antila22b}.
+The selection of best models for a specific application is demonstrated for PC/PE lipid mixtures in Fig.~\ref{fig:quality}, and previously for PC/PS lipid mixtures~\cite{antila22b}, and lipid headgroups~\cite{bacle21}.
+On the other hand, programmatic access in the NMRlipids databank enables automatic analyses over large sets of simulation data containing more data in terms of quantity (e.g., simulation length and number of conformations) with broader range in terms of content (e.g, lipid compositions and ion concentrations) that can be accessible for a single research group.
+Added value of the access to broad range of systems is exemplified in Figs.~\ref{fig:quality}~and~\ref{fig:permeability} revealing correlations between membrane properties, while large quantity of data enables the detection of rare phenomena with good statistics, as exemplified for water permeability in Fig.~\ref{fig:permeability} and cholesterol flip-flop events in Fig.~\ref{??}.
 
-The increasing amount of MD simulation data with programmatic access in the NMRlipids databank opens up possibilities for wide range of applications utilizing the large set of accessible data.
-Extend of the data in the NMRlipids databank in terms of quantity (e.g., simulation length and number of conformations), content (e.g, lipid compositions and ion concentrations) and quality enables analyses that are not possible to conduct from MD simulation data produced by a single research group.
-Applications of the NMRlipids databank to understand how diffusion of water through and along cellular membrane depends on its physical properties are demonstrated in Fig.~\ref{??}.
-Permeation of water through membranes resembles the permeation of also other hydrophilic molecules, such as drugs, and detailed understanding of water dynamics through and along membranes is potentially useful for the development of MRI imaging methods~\cite{topgaard20}.
-These examples demonstrate the practical applications of NMRlipids databank on problems in the biological and biomedical sciences.
+Providing programmatic access to large scale MD simulation data can lead to applications in unprecedented directions.
+For example, water anisotropic diffusion analyzed in Fig.~\ref{fig:permeability} is relevant not only in pharmacokinetics but also in the development of MRI imaging methods where MD simulations are not yet widely used~\cite{topgaard20}.
+We expect the access provided by the NMRlipids databank to facilitate novel applications also in other fields where MD simulations are less commonly applied, such as design of lipid nanoparticles for drug delivery or other applications, and material science.
+The increasing amount of data in terms of quantity and content will also increase the scope of potential applications of the NMRlipids databank in time.
 
-Emerging applications of machine learning are increasing the impact of such databanks.
-For example, the existing Protein Databank (PDB)~\cite{??} containing experimentally determined protein structures with programmatic access has enabled the development of machine learning based tools building on the data collected in the databank over the years~\cite{??}.
-NMRlipids and other databanks with open programmatic access have potential to lead similar unforeseen applications in the future.
+The NMRlipids databank can also serve as a training set for diverse machine learning applications.
+Most straightforward applications include models that would predict connections between membrane properties that are already stored in the databank.
+For example, a machine learning model predicting electron density profiles from form factors would be highly useful in interpretation of scattering experiments.
+On the other hand, more elaborated models could be trained to predict arbitrary membrane properties using the data from the NMRlipids databank.
+Such models could be used to connect membrane composition to relevant membrane properties in wide range of fields from molecular biology to drug development and biomolecular imaging.
+The overlay databank design can open such avenues also for other biomolecules than lipids, such as disordered proteins and membrane-protein systems.
+
+Different types of applications enabled by the NMRlipids databank in wide range of fields are exemplified in Table~\ref{tab:applications}.
+
+\begin{table}[]
+    \centering
+    \begin{tabular}{p{5.0cm}  p{5.0cm}  p{4.0cm}}
+    Type of application     & Practical examples & Target group \\
+    \hline
+    Analyses of rare phenomena               & Lipid flip-flops, water permeation & Membrane scientists \\
+    \\
+    Correlations between membrane properties & 
+    Membrane structural properties, water dynamics (Figs.~\ref{fig:quality} and~\ref{fig:permeability}) & 
+    Membrane scientists \\
+    \\
+    Applications that are outside typical scope of MD simulations & 
+    Anisotropic water diffusion for pharmagokinetics and MRI imaging applications & 
+    Scientist in fields where MD simulations are not usually applied \\
+    \\
+    Selection of the best simulation model for a specific application & 
+    Best model for POPC lipids (Fig.~\ref{fig:quality}), headgroup conformations~\cite{bacle21}, 
+    packing of PS~\cite{antila22b} and PE (Fig.~\ref{fig:POPCPOPEapls}) containing membranes. &
+    Scientists using MD simulations \\
+    \\
+    Guidance for force field development & 
+    Improvements in ion binding to lipids~\cite{melcr18,melcr20} and lipid headgroup conformational ensembles~\cite{??} &
+    Scientists developing parameters for MD simulations \\
+    \\
+    Training and target data for coarse grained models & 
+    Optimizing parameters of coarse grained models against NMRlipids databank, extracting continuum parameters for membranes. &
+    Scientists developing and using coarse grained MD simulations \\
+    \\
+    Training set for machine learning applicatons &
+    programmatic access to the data and results enables training of machine learning type of models for various applications, such as predictions of membrane properties from composition & Scientists building and using machine learning applications for biomolecules.
+    \end{tabular}
+    \caption{Examples on applications of the NMRlipids databank.}
+    \label{tab:applications}
+\end{table}
 
@@
 \subsection{Quality evaluation of C-H bond order parameters}
-As the first step to evaluate simulation qualities against experimental data, each simulation is connected to an experimental data if molar concentrations of all molecules are within $\pm$5 percentage units, charged lipids have the same counterions, and temperatures are within $\pm$2 degrees.
+As the first step to evaluate simulation qualities against experimental data, each simulation is connected to an experimental data if molar concentrations of all molecules are within $\pm$3 percentage units, charged lipids have the same counterions, and temperatures are within $\pm$2 degrees.
 For molar concentrations of water, the exact hydration level is considered only for systems with molar water to lipid ratio below 25, otherwise the systems are considered as fully hydrated.
@@
 \end{equation}
-where $P^{\rm frag}[{\rm lipid}]$ is the average of individual C-H bond order parameters qualities within the fragment, $p_{\rm frag}[{\rm lipid}]$ is the percentage of order parameters for which the quality is available within the fragment, and frag can be {\it sn}-1, {\it sn}-2, headgroup or total (all order parameters within a molecule).
+where $\langle P[{\rm lipid}]\rangle_{\rm frag}$ is the average of individual C-H bond order parameters qualities within the fragment, $p_{\rm frag}[{\rm lipid}]$ is the percentage of order parameters for which the quality is available within the fragment, and frag can be {\it sn}-1, {\it sn}-2, headgroup or total (all order parameters within a molecule).
 The overall quality of different fragments in a simulation are then defined as a molar fraction weighted average over different lipid components

### 2022-11-02 main.tex
Update text and figure
Updates from Overleaf

@@
 %Up to three levels of \textbf{subheading} are permitted. Subheadings should not be numbered.
-\subsection{Overlay structure of the NMRlipids databank}
+\subsection{Programmatic access to MD simulation data of membranes composed of most common biologically abundant lipids}
+NMRlipids databank is a community driven databank containing atomistic MD simulations of biologically relevant lipid membranes emerging from the NMRlipids open collaboration~\cite{botan15,ollila16,catte16,antila19,bacle21}.
+Nearly thousand simulation trajectories with the total length approaching half a millisecond can be accessed programmatically using Python API or graphical user interface (\url{http://www.databank.nmrlipids.fi/}).
+NMRlipids databank is the only databank of MD simulations with open programmatic access to its content enabling fully flexible analyses of its content.
+Following the FAIR principles and NMRlipids project protocol, not only the content and computer programs related to the databank but also the whole construction process of the databank are open access~\cite{botan15}.
+The NMRlipids databank will be always open for data and other contributions.
+
+The distribution of different lipid types in simulations in the NMRlipids databank (Fig.~\ref{fig:overlay}C) resembles the biological abundancies, the PC being the most common lipid type followed by PE, PS, PG, PI, and other lipids depending on organism and organelle~\cite{vanmeer08}.
+Currently the databank is composed of approximately 500 trajectories with the total length of approximately 231 microseconds of which most are contributed for the previous publications from the NMRlipids open collaboration~\cite{botan15,catte16,antila19,bacle21}.
+The distribution of lipids, force fields, length of the trjajectories and available binary mixtures are shown in Fig.~\ref{fig:overlay}.
+
 The three layer structure of the NMRlipids databank, illustrated in Fig.~\ref{fig:overlay} a, enables an efficient way to share and upcycle large MD simulation trajectories.
@@
     \centering
-    \includegraphics[width = 180mm]{Figures/overlay.pdf}
+    \includegraphics[width = 180mm]{Figures/overlay2.pdf}
     \caption{a) Structure of an overlay databank. 
@@
 \end{figure}
-
-\subsection{Content of the databank}
-Currently the databank is composed of approximately 500 trajectories with the total length of approximately 231 microseconds of which most are contributed for the previous publications from the NMRlipids open collaboration~\cite{botan15,catte16,antila19,bacle21}.
-The distribution of lipids, force fields, length of the trjajectories and available binary mixtures are shown in Fig.~\ref{fig:overlay}.
 

### 2022-11-02 main.tex
Add flip-flop figure

@@
     \includegraphics[width=180mm]{Figures/permeation2.pdf}
-    \caption{A 
+    \caption{{\bf A} Water diffusion, $D_\perp$, and permeability, $P$, through membranes, and lateral diffusion along the membrane, $D_{||}$, illustrated in a multilamellar stack of lipid bilayers. 
     {\bf B-E} Water permeation through membranes analyzed from the databank as a function of temperature, thickness, area per lipid, and acyl chain order.
@@
 The NMRlipids databank can greatly advance such studies by giving an easy access to large amounts of data simultaneously.
+
+\begin{figure}[htb]
+    \centering
+    \includegraphics[width=88mm]{Figures/CholFlipFlops.pdf}
+    \caption{{\bf A} Illustration of cholesterol flip-flop.  
+      {\bf B-D} Cholesterol flip-flops analyzed from the databank as a function of membrane thickness, area per lipid, and acyl chain order.
+      Values from simulations with non-zero flip-flop rates are shown with blue dots. Averages over fixed range of x-axis values are shown with black dots.
+    }
+    \label{fig:permeability}
+\end{figure}
 

### 2022-11-22 SI.tex
Updates from Overleaf

@@
 
+%\newpage
+\pagebreak
 \section{Databank content}
 
-\begin{figure}
+\begin{figure}[!h]
   \includegraphics[width=\textwidth]{Figures/DataBankChart.pdf}
@@
 The databank entries (red box) and analysis results (green boxes), locating at \url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations} are automatically generated by the computer programs included in the NMRlipids databank (yellow boxes).
-Because raw data are not permanently stored but can be accessed based on the information in the databank, this connection is marked with the dashed line.
-}\label{DatabankStructure}
+Because raw data are not permanently stored but can be accessed based on the information in the databank, this connection is marked with the dashed line.}
+  \label{DatabankStructure}
 \end{figure}
@@
     SOFTWARE & Software used to run the simulation (e.g. Gromacs, Amber, NAMD, etc.) & \\
-    TRJ & Name of the trajectory file found from DOI & \\
-    TPR & Name of the topology file found from DOI (trp file in the case of Gromacs) & \\
+    TRJ & Name of the trajectory file found from DOI (trr or xtc for Gromacs, dcd for OpenMM) & \\
+    TPR (Gromacs) & Name of the tpr topology file found from DOI for Gromacs simulations & \\
+    PDB (OpenMM) & Name of the pdb file found from DOI for OpenMM simulations & \\
     PREEQTIME & Pre-equilibrate time simulated before the uploaded trajectory in nanoseconds. 
-\tablefootnote{For example, if you upload 100-200 ns part of total 200 ns simulation, this should value should be 100.} & \\
+    %For example, if you upload 100-200 ns part of total 200 ns simulation, this should value should be 100. 
+    & \\
     TIMELEFTOUT & Equilibration period in the uploaded trajectory that should be discarded in analyses. 
-\tablefootnote{For example, if you upload 0-200 ns part of total 200 ns simulation where the first 100 ns should be considered as an equilibration, this value should be 100.} \\
-    COMPOSITION & Molecules names used in the simulation and corresponding mapping files (see section \ref{naming}) & \\
+    %For example, if you upload 0-200 ns part of total 200 ns simulation where the first 100 ns should be considered as an equilibration, this value should be 100. 
+    & \\
+    COMPOSITION & Dictionary connecting universal molecule and atom names to the ones used in simulation & \\
     DIR\_WRK & Temporary working directory in your local computer. \\
-    UNITEDATOM\_DICT & Information for constucting hydrogens for united atom simulations, empty for all atom simulations & \\
+    UNITEDATOM\_DICT & Information for constucting hydrogens for united atom simulations using buildH program~\cite{santuz21}. Empty for all atom simulations. & \\
     TYPEOFSYSTEM & Lipid bilayer or something else & \\
@@
     LOG & Name of the Gromacs log file. & \\
-    TOP & Name of the Gromacs top file. & \\
     GRO & Name of the Gromacs gro file. & \\
+    TOP & Name of top file for Gromacs or psf file for OpenMM. & \\
+    CRD & Name of crd file for OpenMM. & \\
+    WARNINGS & Dictionary containing information about unusual features in the trajectory, such as ambiguous atom names, membrane normal not oriented in z-direction, old Gromacs version used, etc.
+& \\
     \hline
@@
     EXPERIMENT & Potentially connected experimental data & \\
-    COMPOSITION & Numbers of lipid molecules (NPOPC, NPOPG, etc.) per membrane leaflet are calculated by determining on which side of the center of mass of the membrane the center of mass of the head group of each lipid molecule is located.
-    Numbers of other molecules such as solvent and ions (NSOL, NPOT, NSOD, etc.) are read from the topology file. & \\
+    COMPOSITION & Numbers of lipid molecules in both leaflets and numbers of other molecules are added to the dictionary. & \\
+    ID & Unique ID number to ease the analyses. & \\
     \end{tabular}
@@
 
+%\begin{table}[h]
+%    \centering
+%    \begin{tabular}{c|c}
+%        Abbreviation & Molecule name \\
+%        \hline
+%        POPC &  1-palmitoyl-2-oleoyl-sn-glycero-3-phosphocholine\\
+%        POPG &  1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoglycerol \\
+%        POPS & 1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-L-serine \\
+%        POPE & 1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoethanolamine \\
+%        CHOL & cholesterol \\
+%        DHMDMAB & dihexadecyldimethylammonium \\
+%        \hline
+%        POT & potassium ion \\
+%%        SOD & sodium ion \\
+ %       CLA & chloride ion \\
+%        CAL & calcium ion \\
+%%    \end{tabular}
+ %   \caption{Abbreviations used in the databank}
+%    \label{tab:abbreviations}
+%\end{table}
+
 \begin{table}[h]
     \centering
-    \begin{tabular}{c|c}
+    \begin{tabular}{c&c}
         Abbreviation & Molecule name \\
         \hline
-        POPC &  1-palmitoyl-2-oleoyl-sn-glycero-3-phosphocholine\\
+POPC &  1-palmitoyl-2-oleoyl-sn-glycero-3-phosphocholine  \\
 POPG &  1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoglycerol \\
@@
 POPE & 1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoethanolamine \\
+PYPC & 1-(16:0)-2-(16:1$^\Delta9$)-sn-glycero-3-phosphocholine \\
+PAzePCprot & 1-palmitoyl-2-azelaoyl-sn-glycero-3-phosphocholine protonated \\
+PAzePCdeprot & 1-palmitoyl-2-azelaoyl-sn-glycero-3-phosphocholine deprotonated \\
+DMPC & 1,2-dimyristoyl-sn-glycero-3-phosphocholine \\
+DPPC & 1,2-dipalmitoyl-sn-glycero-3-phosphocholine \\
+DPPE & 1,2-dipalmitoyl-sn-glycero-3-phosphoethanolamine \\
+DPPG & 1,2-dipalmitoyl-sn-glycero-3-phospho-(1'-rac-glycerol) (sodium salt) \\
+DEPC & 1,2-dierucoyl-sn-glycero-3-phosphocholine \\
+DRPC & 1,2-(14:1$^\Delta9$)-sn-glycero-3-phosphocholine \\
+DYPC & 1,2-(16:1$^\Delta9$)-sn-glycero-3-phosphocholine \\
+DLPC & 1,2-dilauroyl-sn-glycero-3-phosphocholine \\
+DLIPC& 1,2-dilinoleoyl-sn-glycero-3-phosphocholine \\
+DOG  & 1,2-dioleoyl-sn-glycerol \\
+DOPC & 1,2-dioleoyl-sn-glycero-3-phosphocholine \\
+DOPE & 1,2-dioleoyl-sn-glycero-3-phosphoethanolamine \\
+DDOPC& 1,2-didocosahexaenoyl-sn-glycero-3-phosphocholine \\
+DOPS & 1,2-dioleoyl-sn-glycero-3-phospho-L-serine \\
+DSPC & 1,2-distearoyl-sn-glycero-3-phosphocholine \\
+DAPC & 1,2-diarachidonoyl-sn-glycero-3-phosphocholine \\
+SLiPC & 1-(18:0)-2-(18:2 $^{\Delta9,12}$)-sn-glycero-3-phosphocholine  \\
+DMTAP & 1,2-dimyristoyl-3-trimethylammonium-propane \\
+SOPC & 1-stearoyl-2-oleoyl-sn-glycero-3-phosphocholine \\
+POPI & \\ 
+SAPI & \\
+SLPI &  \\
+SDG & 1-stearoyl-2-docosahexaenoyl-sn-glycerol \\
+SDPE & 1-stearoyl-2-docosahexaenoyl-sn-glycero-3-phosphoethanolamine \\
+CER  & N-palmitoyl-D-erythro-sphingosine \\
 CHOL & cholesterol  \\
+DCHOL & 18,19-di-nor-cholesterol \\
 DHMDMAB & dihexadecyldimethylammonium  \\
@@
 CAL & calcium ion  \\
+CES & caesium ion \\
 SOL & water  \\
     \end{tabular}
-    \caption{Abbreviations used in the databank}
+    \caption{Abbreviations for molecules used in the databank}
     \label{tab:abbreviations}
@@
 \end{table}
+
+\pagebreak
+\section{Experimental data}
+\subsection{NMR experiments}
+Acyl chain order parameters of POPE and POPG were analyzed from the experiments that were used to determine headgroup order parameters in Ref.~\citenum{bacle21}.
+
+\subsection{X-ray scattering experiments}
+X-ray scattering form factors contributed to the NMRlipids III project are included in the databank (\url{http://nmrlipids.blogspot.com/2022/09/nmrlipids-iii-including-lipid-lateral.html}).
 

### 2022-11-22 main.tex
Updates from Overleaf

@@
 
-\title{NMRlipids Databank: Overlay Databank of Lipid Membrane Simulations Arising from Open Collaboration}
+\title{NMRlipids Databank: Making data-driven analyses of membrane properties accessible for all}
+%Overlay Databank of Lipid Membrane Simulations Arising from Open Collaboration}
 
@@
 
-While MD simulations of systems with few biomolecules continue having major contributions in our understanding on biology and drug design, researches are now searching new avenues by building models not only for individual molecules but whole cells or organelles using interdisciplinary approaches~\cite{johnson15,thornburg22,gupta22}.
+Cellular membranes are composed of hundreds of different types of lipid molecules which regulate membrane properties, morphology and biological functions~\cite{vanmeer08}.
+Membrane lipid composition can be also related to diseases, such as cancer and neurodegenerative disorders, and potential therapeutics that affect membrane compositions have been proposed~\cite{torres21}.
+However, membrane containing systems are often difficult to study experimentally because membranes are in disordered fluid state in biological conditions and sample preparation can be highly complicated due to rich phase behaviour and complex interactions between lipids and other biomolecules.
+Therefore, connections between complex lipid interactions and biological functions are still poorly understood.
+%why such variety of lipids is needed and how their. 
+Molecular dynamics (MD) simulations have been particularly useful in understanding membrane systems, yet their accuracy is often compromised by the quality of model parameters and other artefacts~\cite{antila22b,gupta22}.
+%While MD simulations of systems with few biomolecules continue having major contributions in our understanding on biology and drug design, 
+On the  other hand, the accuracy of models is becoming increasingly important as researches are progressing from the simulations of 
+%now searching new avenues by building models not only 
+individual molecules to the models of whole cells or organelles using interdisciplinary approaches~\cite{johnson15,thornburg22,gupta22}.
+In such systems exhibiting very complex behaviour, inaccuracies are more difficult to detect and multiplication of small errors may have major effects the conclusions.
 
-The importance of sharing MD simulation data following the FAIR principles~\cite{wilkinson16} has been widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20}, and databanks are emerging~\cite{meyer10,kamp10,hospital16,mixcoha16,newport19,bekker20,espigares20,leston22}.
-The relevance of quality evaluation of simulation trajectories in databanks regarding technical details of simulations and accuracy of the underlying physical description of the system (force field) has become evident~\cite{tai04,meyer10,hospital20} and such quality evaluation has in some cases also been implemented~\cite{meyer10,hospital16}.
-However, straightforward quality comparisons between individual simulations or force fields within these databanks remain challenging.
-While importance of such databanks for MD simulations is widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} and different kinds of approaches are emerging~\cite{meyer10,kamp10,hospital16,mixcoha16,newport19,bekker20,espigares20,leston22}, generally accepted protocols and best practices are still under active development.
+In contrast to experimental structural biology where standard protocols to share and quality evaluate the resolved structures are established in the PDB databank~\cite{montelione13}, such best practices are yet to be defined for MD simulations.
+Importance of such approaches is widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} and solutions for sharing data are emerging for proteins in solution~\cite{meyer10,kamp10} or membranes~\cite{newport19,espigares20,leston22}, nucleic acids~\cite{hospital16}, cyclodextrins~\cite{mixcoha16}, and arbitrary biomolecules~\cite{bekker20}.
+However,
+%The relevance of quality evaluation of simulation trajectories in databanks regarding technical details of simulations and accuracy of the underlying physical description of the system (force field) has become evident~\cite{tai04,meyer10,hospital20} and such
+automatic quality evaluation~\cite{meyer10,hospital16} and programmatic access are still rare. 
+%However, straightforward quality comparisons between individual simulations or force fields within these databanks remain challenging. 
+Therefore, tools for automatic quality evaluation of membrane simulations or training sets for machine learning models of membrane containing systems are not yet available.
 
-Here we present a solution for lipid bilayers based on overlay databank structure illustrated in Fig.~\ref{fig:overlay}.
-The concept of overlay databank is developed here to solve the practical challenges in generating databanks of MD simulation data enabling flexible analyses over large sets of simulation data, but it can be  potentially used for wide range of situations.
-Particularly when storage of raw data requires significant resources and final outcomes or best practices are not yet clear, overlay databank approach lowers the barrier to start without compromizing the long term stability or scalability.
+%Applications of MD simulations in understanding biomolecular behaviour are steadily increasing, but protocols for quality evaluation and sharing simulation data fall behind the ones in more elaborated fields, such as structural biology.
+Development of such protocols has been limited by practical challenges, such as lack of universally defined quality measures for MD simulations, incompatible naming conventions in different models and simulation softwares, and high technical requirements for hardware to share large trajectories from MD simulations, as well as incentives for scientist to evaluate and share their MD simulation data.
 
-The practical relevance of the NMRlipids databank is exemplified by automatic quality evaluation and ranking of large amount of MD simulation data, data-driven analysis detecting correlations between properties of model cell membranes and analyses of rare phenomena that are beyond the scope of standard MD simulation studies.
-The NMRlipids databank provides new tools for researchers in wide range of fields in academia and industry from cell membrane biology to lipid nanoparticle formulations and data-driven computational chemistry and machine learning.
+Here we present the NMRlipids databank, a community driven open access databank with programmatic access containing atomistic resolution MD simulations of lipid bilayers.
+The programmatic access enables various data-driven approaches, providing new tools for researchers in wide range of fields in academia and industry from cell membrane biology to lipid nanoparticle formulations and data-driven computational chemistry and machine learning.
+Here we demonstrate how data-driven analysis of water anisotropic diffusion in membrane systems from the NMRlipids databank can extend the scope of MD simulations to MRI imaging and pharmacokinetics, and how the NMRlipids databank can be utilized to analyse rare phenomena that are beyond the scope of standard MD simulation studies.
+Furthermore, we perform automatic quality evaluation of membrane simulations which guides the selection of best models for specific applications and development of simulation parameters and methodology.
+%The practical relevance of the NMRlipids databank is exemplified by automatic quality evaluation and ranking of large amount of MD simulation data,
+
+While the NMRlipids databank currently contains only bilayer systems, it can be readily expanded to contain other molecules than lipids, such as disordered proteins or sugars. 
+%a solution for lipid bilayers based on overlay databank structure illustrated in Fig.~\ref{fig:overlay}.  
+The combination of overlay databank structure and open collaboration introduced here can be applied also when building databanks in other fields,  
+%is developed here to solve the practical challenges in generating databanks of MD simulation data enabling flexible analyses over large sets of simulation data, but it can be  potentially used for wide range of situations.
+particularly when storage of raw data requires significant resources, best practices in the field are not defined, and incentives to share data do not exist.
+%overlay databank approach lowers the barrier to start without compromizing the long term stability or scalability.
+Therefore, the approach opens up possibilities to create databanks enabling data-driven and machine learning applications in new fields.
 
@@
 Nearly thousand simulation trajectories with the total length approaching half a millisecond can be accessed programmatically using Python API or graphical user interface (\url{http://www.databank.nmrlipids.fi/}).
-NMRlipids databank is the only databank of MD simulations with open programmatic access to its content enabling fully flexible analyses of its content.
-Following the FAIR principles and NMRlipids project protocol, not only the content and computer programs related to the databank but also the whole construction process of the databank are open access~\cite{botan15}.
-The NMRlipids databank will be always open for data and other contributions.
+%NMRlipids databank is the only databank of MD simulations with open programmatic access to its content enabling fully flexible analyses of its content. 
+Following the FAIR principles~\cite{wilkinson16} and NMRlipids project protocol, not only the content and computer programs related to the databank but also the whole construction process of the databank are open access~\cite{botan15}.
+%The NMRlipids databank will be always open for data and other contributions.
 
-The distribution of different lipid types in simulations in the NMRlipids databank (Fig.~\ref{fig:overlay}C) resembles the biological abundancies, the PC being the most common lipid type followed by PE, PS, PG, PI, and other lipids depending on organism and organelle~\cite{vanmeer08}.
-Currently the databank is composed of approximately 500 trajectories with the total length of approximately 231 microseconds of which most are contributed for the previous publications from the NMRlipids open collaboration~\cite{botan15,catte16,antila19,bacle21}.
+Currently available single component lipid membranes and binary mixtures in the NMRlipids databank are illustrated in Fig.~\ref{fig:overlay}D.
+While these compose the majority of available trajectories, also mixtures up to five lipids are available.
+The distribution of available simulations containing a specific lipid is shown in Fig.~\ref{fig:overlay}C.
+This roughly resembles the biological abundancies of different lipid types, PC being the most common  followed by cholesterol, PE, PS, PG, PI, and other lipids depending on organism and organelle~\cite{vanmeer08}.
+The force fields used in the simulations cover all the essential parameters used in lipid simulations, ranging from highly popular CHARMM36 parameters to united atom and polarizable force fields.
+Therefore, the averages calculated over the databank can be considered as mean predictions from available lipid models for an average cell membrane.
+%Currently the databank is composed of approximately 500 trajectories with the total length of approximately 231 microseconds of which most are contributed for the previous publications from the NMRlipids open collaboration~\cite{botan15,catte16,antila19,bacle21}.
 The distribution of lipids, force fields, length of the trjajectories and available binary mixtures are shown in Fig.~\ref{fig:overlay}.
 
-The three layer structure of the NMRlipids databank, illustrated in Fig.~\ref{fig:overlay} a, enables an efficient way to share and upcycle large MD simulation trajectories.
-Layer 1 contains raw simulation data that can locate in any publicly available repository or server offering long term stability and permanent links to the data, such as digital object identifiers.
-Layer 2 contains all the relevant information on the simulations, including links to the raw data in layer 1 and relevant metadata describing the systems.
-Layer 3 is the application layer containing outputs from the databank, such as graphical user interface (\url{http://www.databank.nmrlipids.fi/}) and results from analyses desribed in sections below.
-
-Layer 2 is the core of the databank.
-In addition to the metadata and links to raw simulation data, it contains universal naming conventions for the molecules and atoms in the databank, computer programs to generate and analyse the databank, basic properties calculated from all simulations (area per lipid, C-H bond order parameters, x-ray scattering form factors and membrane thickness), and quality evaluation of simulations against experimental data.
+The efficient upcycling of large MD simulation trajectories in the NMRlipids databank is enabled by the overlay structure illustrated in Fig.~\ref{fig:overlay}A.
+Raw simulation data in the {\it data layer} can be stored in any publicly available location with long term stability and permanent links to the data, such as digital object identifiers.
+The {\it Databank layer} is the core of the databank containing all the relevant information about the simulations, including links to the raw data, relevant metadata describing the systems, universal naming conventions for lipids and their atoms, computer programs to create and analyse the entries, basic properties calculated from all simulations (area per lipid, C-H bond order parameters, x-ray scattering form factors and membrane thickness), and quality evaluation of simulations against experimental data.
 In practise, this information is stored is the git repository which is currently available at~\url{https://github.com/NMRLipids/Databank}.
-Detailed description on its content is in the supplementary information.
-Applications in layer 3 uses the information in layer 2 to access raw data in layer 1 to perform automatic analyses over large sets of simulations in the databank.
+The {\it application layer} contains outputs from the databank, such as graphical user interface (\url{http://www.databank.nmrlipids.fi/}) and results from analyses described in sections below.
+More detailed description on the NMRlipids databank structure is available the supplementary information.
 
@@
     e) Distribution of force fields in the simulations in the NMRlipids databank.
-    The figures and numbers are created on 9th of May 2022.}
+    }
     \label{fig:overlay}
@@
 
-\subsection{Quality evaluation of force fields}
-Quality of membrane simulations with different force fields have been evaluated against experimental data during parameterization and in separate comparison studies~\cite{botan15,ollila16,catte16,pluhackova16,perez17,leonard19}, but universal quality measure for membrane simulations is not defined and controversial results are often reported from simulations~\cite{antila22b}.
-The lack of universal quality measure for membrane simulations complicates the selection of proper simulation parameters and the estimation of reliability in simulations, thereby being a major obstacle in many applications of membrane MD simulations.
-To enable the rapid quality evaluation of membrane MD simulations, a quantitative quality measure based on C-H bond order parameters from NMR experiments is defined in the NMRlipids databank.
-%Conformational ensembles of individual lipid molecules are evaluated against C-H bond order parameters from NMR experiments.
-This is complemented by comparing the locations of x-ray scattering form factor minima against experiments. 
-%that are related to the overall membrane dimensions via electron density profile. 
-Both of these are robust experimental measurables that can be directly connected to the simulation data~\cite{ollila16}.
-The used quality measures are defined in detail in the supplementary information.
+\subsection{Selection of best simulation parameters for applications using the NMRlipids databank}
+%Quality of membrane simulations with different force fields have been evaluated against experimental data during parameterization and in separate comparison studies~\cite{botan15,ollila16,catte16,pluhackova16,perez17,leonard19,??}, but the lack of universal quality measure for membrane simulations complicates the estimations of reliability of simulations.
+Understanding the accuracy of used force field parameters is crucial when estimating the reliability and significance MD simulation results.
+However, the lack of universal quality measures for MD simulations and complicated landscape of force field quality for membranes hamper the quality of simulations which has resulted to many controversial results~\cite{antila22b}.
+For example, previous quality evaluations of membrane simulations have concluded that CHARMM36 parameters give the best description for lipid headgroup conformational ensembles~\cite{bacle21}, GROMOS-CKP parameters best capture the membrane packing in POPS bilayers~\cite{antila22b}, and that OPLS3e parameters overcome CHARMM36 in structural quality for POPC but predicts overestimated ion binding~\cite{kurki22}.
+%, thereby being a major obstacle in many applications of membrane MD simulations. 
+To enable the navigation in this complex force field quality landscape, we define here quantitative quality measures that can be used to rapidly find the best available parameters for a specific application or to guide force field development.
+For detailed definitions of quality measures, see the supplementary information.
 
-%NMRlipids databank contains also experimental order parameter and x-ray form factor data that is connected to corresponding  simulations in order to define the simulation qualities and rank them to select the suitable simulation models for particular applications.
-Simulations with the highest scores based on order parameter evaluation for all simulations are shown in Fig.~\ref{fig:quality} A and for POPE lipids in Fig.~\ref{fig:quality} B.
-The direct comparison with experiments for the best simulation, POPC with OPLS3e, is shown in Fig.~\ref{fig:quality} C, where discrepancies with experiments are observed only for the third carbon in the $\it{sn}$-1 chain, and carbons 8 and 9 close to the double bond in $\it{sn}$-2 chain.
-The best overall quality for POPE is found in GROMOS-CKP simulation, for which the direct comparison in Fig.~\ref{fig:quality} D show minor differences in acyl chains, but major discrepancies in glycerol backbone and in the first carbon of $\it{sn}$-2 chain.
-These parts are better described for POPE by CHARMM36 parameters (direct comparison shown in Fig.~\ref{fig:quality} E), but its overall quality suffers from the overestimated order in acyl chains.
-
-\begin{figure}[tbp]
+\begin{figure}[!t]
     \centering
@@
     B) Best simulations ranked based on the overall order parameter quality of POPE lipid. 
-    C) Scatter plots and Pearson correlation coefficients for the area per lipid, thickness, and experimental observables. 
-    %second minima of form factor and average order parameter of the {\it sn}-1 acyl chain extracted from the NMRlipids databank. 
+    C)-E) Evaluation against experimental data exemplified for a simulation with the best overall order parameter quality (C), the best quality for POPE lipid (D), and the headgroup quality for POPE (E).
+    F) Scatter plots and Pearson correlation coefficients for the membrane area per lipid, thickness, first minima of x-ray scattering form factor and average order parameter of the {\it sn}-1 acyl chain extracted from the NMRlipids databank.
     All correlation coefficients have p-value below 0.001. For more correlations see Fig.~\ref{fig:QualityCorrelationsSI}.
-    D)-F) Evaluation against experimental data exemplified for a simulation with the best overall order parameter quality (D), the best quality for POPE lipid (E), and the headgroup quality for POPE (F).
-    %c) Change in area per lipid upon addition of POPE to POPC in different simulation extracted from the NMRlipids databank and quality evaluation table of POPE simulations.
     }
@@
 
-\subsection{Finding the best models for PC and PE lipid mixtures}
-The power of the NMRlipids databank in finding the best models for a certain lipid bilayer system is exemplified here for mixtures of PE and PC lipids.
-Area per lipids from available POPC and POPE bilayer simulations in the databank from different force fields at 310\,K are shown in Fig.~\ref{fig:POPCPOPEapls}.
-Because the area per lipid strongly correlates with the average order of $\it{sn}$-1 chain (Fig.~\ref{fig:quality} F), we use order  parameter qualities to evaluate which simulations predict the best area per lipids for POPC and POPE.
-Among the force fields with area per lipids shown in Fig.~\ref{fig:POPCPOPEapls}, Slipids ranks 1st for POPC with a clear difference to others, and 2nd for POPE with only marginally lower quality than GROMOS-CKP (Figs.~\ref{fig:quality} A and B).
-The direct comparison with experiments in Fig.~\ref{fig:POPC_POPE_dataSI} in the supplementary information reveals that GROMOS-CKP and CHARMM36 predicts too ordered (too low area per lipid) bilayer for POPC, while the order parameters in Lipid17 are lower than in experiments indicating overestimated area per lipid.
-For POPE, order parameters in  Slipids and GROMOS-CKP are close to experiments suggesting that their are per lipids are most reasonable, while CHARMM36 and Lipid17 are too packed thereby overestimating the order parameters.
+Ranking of all simulations with the highest scores based on evaluation against C-H bond order parameters from NMR experiments is shown in Fig.~\ref{fig:quality} A and ranking of best POPE lipid bilayer simulations in Fig.~\ref{fig:quality} B.
+Direct comparisons with experiments for selected simulations are shown in Figs.~\ref{fig:quality} C-D.
+C-H bond order parameters are sensitive to conformational ensembles of individual lipid molecules~\cite{ollila16}, but they correlate well also with the membrane thickness and lateral packing as shown in Fig.~\ref{fig:quality} F.
+This is complemented by comparing the absolute values of x-ray scattering form factors between simulations and experiments that are related to the overall membrane dimensions via electron density profile.
+Particularly, the location of minima in these form factors are correlated with the membrane dimensions as shown in Figs.~\ref{fig:quality} F,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}.
+
+%best simulation, POPC with OPLS3e, is shown in
+%based on C-H bond order parameters from NMR experiments and x-ray scattering form factors that enable the rapid quality evaluation of membrane MD simulations in the NMRlipids databank.
+
+%Order parameters and scattering form factors are robust experimental measurables that can be directly connected to the simulation data~\cite{ollila16}.
+
+%NMRlipids databank contains also experimental order parameter and x-ray form factor data that is connected to corresponding  simulations in order to define the simulation qualities and rank them to select the suitable simulation models for particular applications.
+
+%Fig.~\ref{fig:quality} C, where discrepancies with experiments are observed only for the third carbon in the $\it{sn}$-1 chain, and carbons 8 and 9 close to the double bond in $\it{sn}$-2 chain.
+The best overall quality for POPE is found in GROMOS-CKP simulation, for which the direct comparison in Fig.~\ref{fig:quality} D show minor differences in acyl chains, but major discrepancies in glycerol backbone and in the first carbon of $\it{sn}$-2 chain.
+These parts are better described for POPE by CHARMM36 parameters (direct comparison shown in Fig.~\ref{fig:quality} E), but its overall quality suffers from the overestimated order in acyl chains.
+
+%\subsection{Finding the best models for PC and PE lipid mixtures}
+The power of the NMRlipids databank in selecting the best simulation parameters for a specific application is demonstrated for mixtures of PE and PC lipids in Fig.~\ref{fig:POPCPOPEapls} showing area per lipids from available POPC and POPE bilayer simulations in the databank from different force fields at 310\,K.
+Predictions from different force fields deviate significantly in terms of absolute values and the slope of decrease upon addition of POPE.
+Because the area per lipid correlates with the average order parameter of $\it{sn}$-1 chain (Fig.~\ref{fig:quality} F), simulations with the best predictions for area per lipids can be selected based on order parameter quality evaluation.
+Quality evaluation in Fig.~\ref{fig:quality} and direct comparison in Fig.~\ref{fig:POPC_POPE_dataSI} reveal that Slipids ranks 1st for POPC with a clear difference to others, and 2nd for POPE with only marginally lower quality than GROMOS-CKP, therefore being the best selection for studies where packing effects of PE lipids are relevant.
+%The direct comparison with experiments in Fig.~\ref{fig:POPC_POPE_dataSI} in the supplementary information reveals that 
+Simulations with CHARMM36 and GROMOS-CKP predict too packed bilayer for POPC (overestimated order in Fig.~\ref{fig:POPC_POPE_dataSI}), while area per lipid for POPC is overstimated in Lipid17.
+For POPE,
+%order parameters in  Slipids and GROMOS-CKP are close to experiments suggesting that their are per lipids are most reasonable, while
+CHARMM36 and Lipid17 predict too packed membranes.
+
+In conclusion, 
+%the quality evaluation based on the NMRlipids databank suggests that the Slipids parameters would give the most reliable description for membrane packing of POPC:POPE mixtures.
+Slipids was a relatively good model also for POPC:POPS mixtures in similar comparisons utilizing the preliminary version of the NMRlipids databank~\cite{antila22b}, although charged membranes are more complicated as the counterion binding affinity plays a significant role in membrane packing~\cite{antila19}.
+On the other hand, CHARMM36 overcomes Slipids when the focus is on conformational ensembles of headgroups  in different types of lipids because the accuracy of glycerol backbone is low in Slipids~\cite{bacle21}.
+Altogether, these examples demonstrate how
+the automatic quality evaluation in the NMRlipids databank enables rapid selection of the best simulation parameters to specific applications without extensive and tedious manual force field evaluation.
+This possibility will promote more reliable simulation results for wide range of applications, support force field development and parametrization of coarse grained force fields against the best atomistic MD simulations.
+The automatically extracted quantitative quality measures will be particularly useful to guide automatic parametrization procedures.
+
 \begin{figure}[tbp]
@@
 \end{figure}
-
-In conclusion, the quality evaluation based on the NMRlipids databank suggests that the Slipids parameters would give the most reliable description for membrane packing of POPC:POPE mixtures.
-Slipids was a relatively good model also for POPC:POPS mixtures in similar comparisons utilizing the preliminary version of the NMRlipids databank~\cite{antila22b}, although charged membranes are more complicated as the counterion binding affinity plays a significant role in membrane packing~\cite{antila19}.
-On the other hand, CHARMM36 overcomes Slipids when the focus is on conformational ensembles of headgroups  in different types of lipids because the accuracy of glycerol backbone is low in Slipids~\cite{bacle21}.
-Altogether, these examples demonstrate how the NMRlipids databank can be used to navigate in the complex landscape of lipid force field quality.
 
@@
 
-\subsection{Water diffusion anisotropy in membrane systems}
+\subsection{Detection rare phenomena using NMRlipids databank: Cholesterol flip-flops}
+Lipid flip-flops from one bilayer leaflet to another play an important role in lipid trafficking and regulating membrane properties~\cite{vanmeer08}.
+Phospholipid flip-flops are slow, occuring with the timescales of hours or days, while cholesterol, diacylglycerol and ceramide flip-flops are faster, yet the reported timescales range between minutes to sub-millisecods~\cite{vanmeer08,steck12,parisio16,gu19}.
+These timescales were previously accessible only by coarse-grained simulations or free energy calculations~\cite{parisio16}, yet atomistic simulations reporting cholesterol flip-flop events have been published recently~\cite{gu19,javanainen19,baral20}.
+%However, systematic studies on how cholesterol flip-flop rate depends on membrane properties are still beyond the current computational resources available for a single research group. 
+These studies report increase in cholesterol flip-flop rates with increasing unsaturation level and decreasing cholesterol concentration~\cite{gu19,javanainen19}, but correlations between cholesterol flip-flop rates and membrane properties have not been systematically studied.
+Here we demonstrate that the large amount of simulation data available in the NMRlipids databank with programmatic access makes the analysis of these correlations accessible for all.
+
+Flip-flops were observed for cholesterol, % (634 events in 77 different simulations), 
+DCHOL (18,19-di-nor-cholesterol), %, 16 flip-flops in 1 simulation) 
+DOG (1,2-dioleoyl-sn-glycerol), %, 44 flip-flops in 4 simulations) 
+and SDG (1-stearoyl-2-docosahexaenoyl-sn-glycerol) in the NMRlipids databank.
+Cholesterol flip-flop rates range between 0.001-1.6\,$\mu$s$^{-1}$ with the mean of 0.16\,$\mu$s$^{-1}$ and median of 0.07\,$\mu$s$^{-1}$.
+These values are in line with the previously reported values from atomistic MD simulations~\cite{gu19,javanainen19,baral20}.
+Flip-flop rates of DCHOL, 0.2\,$\mu$s$^{-1}$, was close to the average value of cholesterol, while average rates for diacylglycerols DOG and SDG were higher than for cholesterol, 0.4\,$\mu$s$^{-1}$ and 0.5\,$\mu$s$^{-1}$, respectively.
+Flip-flops were not observed for other lipids, giving the upper limits for PC lipid flip-flop rate of 9\,s$^{-1}$ and for ceramide (N-palmitoyl-D-erythro-sphingosine) of 0.002\,$\mu$s$^{-1}$.
+Therefore, data in the NMRlipids databank suggests that the lipid flip-flip rate decreases in the order of diacylglycerols $>$ cholesterol $>$ other lipids including ceramides.
+However, amount of data for diacylgycerols (8 simulations with Lipid17 force field) and ceramide (3 simulations with CHARMM force field) is less than for cholesterol (83 simulations), thus we cannot fully exclude the effect of force field or composition on this comparison.
+
+Nevertheless, we use the wide range of available simulations to analyse how cholesterol flip-flop rate depends on membrane properties.
+Cholesterol flip-flop rates and their averages over fixed ranges of x-axis values are plotted as a function of membrane thickness, lateral density and order in Figs.~\ref{fig:flip-flops} B-D.
+The results reveal a non-linear correlation between decreasing cholesterol flip-flop rate and membrane packing.
+Flip-flop rates increase with an order of magnitude when membrane packing density decreases and a major jump is observed at low membrane packing.
+The order of magnitude changes in cholesterol flip-flop rate with the membrane composition may have major implications in understanding lipid trafficking and membrane biochemistry~\cite{gu19,baral20}.
+Because the results from the NMRlipids databank are averaged over large range of membrane compositions and force fields, they show that the strong dependence of cholesterol flip-flop rate on membrane properties is not limited to certain lipid compositions or force fields used in previous studies~\cite{gu19,javanainen19,baral20}.
+
+\begin{figure}[htb]
+    \centering
+    \includegraphics[width=88mm]{Figures/CholFlipFlops.pdf}
+    \caption{{\bf A} Illustration of cholesterol flip-flop.  
+      {\bf B-D} Cholesterol flip-flops analyzed from the databank as a function of membrane thickness, area per lipid, and acyl chain order.
+Values from simulations with non-zero flip-flop rates are shown with blue dots.
+Averages over fixed range of x-axis values are shown with black dots.
+    }
+    \label{fig:flip-flops}
+\end{figure}
+
+Cholesterol flip-flops are phenomena whose analyses from MD simulations has only recently became possible for groups with substantial resources and special expertise available~\cite{gu19,javanainen19,baral20}.
+Our analysis demonstrates how the NMRlipids databank makes analyses of such rare phenomena accessible to wide range of scientists in different fields of science and industry who do not have the access to required resources and expertise to produce the large amounts of MD simulation data.
+
+\subsection{Extending the scope of MD simulations to new fields using the NMRlipids databank: Water diffusion anisotropy in membrane systems}
 The anisotropy in water diffusion in parallel and perpendicular directions with respect to membranes plays a role in the drug translocation through biological material, particularly in skin \cite{hansen13,wen18,nitsche19,roberts21}, and in MRI imaging~\cite{topgaard20}.
-Several models on how water dynamics depends on membrane properties have been proposed based on different experimental and theoretical methods, yet consensus
-%on how water permeability depends on membrane properties, such as area per lipid and thickness, or molecular composition, 
-has not been reached~\cite{mathai07,nitsche13,nitsche16,shinoda16,venable19,frallicciardi22}.
-MD simulations have been also used in these endeavours, but systematic studies are limited by the accessible amount of data as only few water permeation events are typically observed in a single MD simulation trajectory~\cite{venable19,camilo2022}.
-Therefore, this serves as an excellent example where the large amount of available MD simulations in the NMRlipids databank can be utilized to answer important biological questions.
-
-To this end, we first calculated the water permeability through membranes from all the simulations in the NMRlipids databank.
-Resulting permeabilities are shown as a function of temperature, membrane thickness, area per lipid, and acyl chain order in Figs.~\ref{fig:permeability} B-E.
-The values for water permeabilities from simulations vary between 0.3 and 322\,$\mu$m/s with the mean and median of 14\,$\mu$m/s and 8\,$\mu$m/s, respectively.
-These values have the same order of magnitude as the experimentally determined diffusive permeability coefficients, but are on average below the values reported for PC lipids in liquid crystalline phase,~190-330\,$\mu$m/s~\cite{jansen95}.
-To analyze how water permeability through membranes depend on bilayer properties on average, we calculated the average permeabilities over all systems with the value of temperature, thickness, area per lipid, or acyl chain order within a fixed range.
-These averaged permeabilities also shown in Figs.~\ref{fig:permeability} with black dots.
-As expected, the permeability increases with the temperature, giving 17\,$\pm$\,4\,k$_b$T for the average energy barrier for the water permeation from the Arrhenius plot in Fig.~\ref{fig:permeability} B.
-Water permeability decreases with increasing membrane packing, i.e., with decreasing area per lipid and increasing thickness and acyl chain order.
-Linear dependence of permeability on area per lipid and thickness is previously reported from simulations and experiments for some systems, but not for all~\cite{mathai07,venable19,frallicciardi22}.
-Our results suggest on average linear dependence for thicknesses above $\sim$\,3.9\,nm and area per lipids below $\sim$69\,Å$^2$ (insets in Figs.~\ref{fig:permeability} C-D), yet stronger dependence is observed for more loosely packed membranes.
-%within the range of experimentally reported values  permeability which is 1-2 orders of magnitude slower than osmotic permeability \cite{jansen95,??}. 
-Clear dependencies of permeability on charged lipids, cholesterol, POPE, or hydration level were not observed in Fig.~\ref{fig:permeationSI} in the supplementary information.
-%, although weak decrease for the latter two may be visible.
+MD simulations are rarely used to analyze anistropic diffusion of water as only few permeation events for water are typically observed in a single MD simulation trajectory~\cite{venable19,camilo2022}, thereby making the collection of sufficient amount data challenging.
+Here we show that the NMRlipids databank can be used to analyze how anisotropic water diffusion in a multilamellar membrane systems depends on membrane properties.
 
@@
 
-To analyze the anisotropic diffusion of water in membrane environment, we also calculated the water diffusion parallel to the membrane surface from all simulations in the NMRlipids databank.
-The results at different hydration levels are shown in Fig.~\ref{fig:permeability} F together with the experimental data~\cite{rudakova04}.
+To this end, we first calculated the water permeability through membranes from all simulations in the NMRlipids databank.
+The resulting non-zero values range between 0.3 and 322\,$\mu$m/s with the mean and median of 14\,$\mu$m/s and 8\,$\mu$m/s, respectively.
+These values agree with the previously reported simulation results~\cite{venable19,camilo2022} and have the same order of magnitude as experimentally determined diffusive permeability coefficients, but are on average larger than the values reported for PC lipids in liquid crystalline phase,~0.19-0.33\,$\mu$m/s~\cite{jansen95}.
+Observed permeabilities and their averages over fixed ranges of values at x-axis are shown in Figs.~\ref{fig:permeability} B-E as a function of temperature, membrane thickness, area per lipid, and acyl chain order.
+%on average, we calculated the average permeabilities over all systems with the value of temperature, thickness, area per lipid, or acyl chain order within a fixed range.
+These averaged permeabilities also shown in Figs.~\ref{fig:permeability} with black dots.
+As expected, the permeability increases with the temperature, giving 17\,$\pm$\,4\,k$_b$T for the average energy barrier for the water permeation from the Arrhenius plot in Fig.~\ref{fig:permeability} B.
+On the other hand, the permeability of water decreases on average when membranes become more packed, i.e., with decreasing area per lipid and increasing thickness and acyl chain order (Figs.~\ref{fig:permeability} C-E).
+Permeation of water through bilayers depends on membrane properties also according previous studies, but it is not agreed whether the area per lipid~\cite{nagle08} or bilayer thickness~\cite{frallicciardi22} would be the main parameter determining the permeability.
+%is previously reported from simulations and experiments for some systems, but not for all~\cite{mathai07,venable19,frallicciardi22}. 
+Our analysis over the NMRlipids databank, containing significantly more data than that was available in previous studies, suggest non-linear dependence on both of these parameters, yet the linear correlation would be a good approximation for thicknesses above $\sim$\,3.9\,nm and area per lipids below $\sim$69\,Å$^2$ (insets in Figs.~\ref{fig:permeability} C-D).
+%within the range of experimentally reported values  permeability which is 1-2 orders of magnitude slower than osmotic permeability \cite{jansen95,??}. 
+Clear dependencies of permeability on charged lipids, cholesterol, POPE, or hydration level were not observed in Fig.~\ref{fig:permeationSI} in the supplementary information.
+%, although weak decrease for the latter two may be visible.
+%
+%
+%
+%%Several models on how water dynamics depends on membrane properties have been proposed based on different experimental and %%theoretical methods, yet consensus 
+%%%on how water permeability depends on membrane properties, such as area per lipid and thickness, or molecular composition, 
+%%has not been reached~\cite{mathai07,nitsche13,nitsche16,shinoda16,venable19,frallicciardi22}.
+
+To analyze how water diffusion anisotropy depends on membrane properties in a multi-lamellar lipid bilayer system, we calculated also the water diffusion parallel to the membrane surface from all simulations in the NMRlipids databank.
+The parallel diffusion coefficient of water, $D_{||}$, decreases with reduced hydration and increases with the temperature, but dependencies on membrane area per lipid, thickness, or fraction of charged lipids were not observed in Figs.~\ref{fig:permeability} and~\ref{fig:diffusionSI}.
 %The experimental water diffusion increases with increasing hydration level toward the value for bulk water . 
 %As expected, water translational diffusion increase in the diffusion is observed with the increased temperature and hydration level (Figs.~\ref{fig:permeability} H and~\ref{fig:diffusionSI} B). 
-Simulation results are close to the experimental values with low hydration levels, but increase to the values of approximately two times higher than experimental bulk water diffusion value (3.1$\cdot$\,10$^{-9}$\,m$^2$/s at 313\,K~\cite{khakimov08}) with high hydration levels.
+%The results at different hydration levels are shown in Fig.~\ref{fig:permeability} F together with the experimental data~\cite{rudakova04}. 
+Simulation results are close to the experimental values with low hydration levels in Fig.~\ref{fig:permeability} F, but increase to the values of approximately two times higher than experimental bulk water diffusion value (3.1$\cdot$\,10$^{-9}$\,m$^2$/s at 313\,K~\cite{khakimov08}) with high hydration levels.
 This is not surprising as the most common water model used in membrane simulations, TIP3P, overestimates the bulk water diffusion~\cite{pathirannahalage21}.
-The water diffusion parallel to membrane surface increased with the temperature, but dependencies on membrane area per lipid, thickness, or fraction of charged lipids were not observed in Fig.~\ref{fig:diffusionSI} in the supplementary information.
-In order to estimate the diffusion anistoropy of water in multilamellar membrane system, we translated the water permeation coefficients through membranes to diffusion coefficients through multilamellar stack using the Tanner equation~\cite{tanner78,wasterby02}.
-The resulting diffusion coefficients through the membranes are approximately five orders of magnitude slower than along the membrane (Figs.~\ref{fig:permeability} G and H), which is at the upper limits of the anisotropic estimated from the experimental data~\cite{nitsche19}.
-Relatively high anisotropy is understandable as simulations give slightly slower permeation rates and higher lateral diffusion rates than experiments.
+To estimate the diffusion anisotropy of water, $D_{\perp}/D_{||}$, in multi-lamellar membrane system, the permeability coeffients of water through membranes were translated to perpendicular diffusion coefficients, $D_{\perp}$, using the Tanner equation~\cite{tanner78,wasterby02}.
+The resulting perpendicular diffusion coefficients are approximately five orders of magnitude smaller than lateral diffusion coefficients of water (Figs.~\ref{fig:permeability} G and H), which is at the upper limit of the anisotropic estimated from the experimental data~\cite{nitsche19}.
+%Large anisotropy values are understandable as simulations give slightly slower permeation rates and higher lateral diffusion rates than experiments. 
 Significant increase in the diffusion anisotropy with membrane packing is observed, as $D_{\perp}/D_{||}$ drifts away from one with decreasing area per lipid and increasing thickness in Figs.~\ref{fig:permeability} G and H.
@@
 
-\subsection{Flip-flops}
-Lipid flip-flops from one bilayer leaflet to another play an important role in lipid trafficking and regulating membrane properties~\cite{vanmeer08}.
-Phospholipid flip-flops are slow, occuring with the timescales of hours or days, while cholesterol, diacylglycerol and ceramide flip-flops are faster, yet the reported timescales range between minutes to sub-millisecods~\cite{vanmeer08,steck12,parisio16,gu19}.
-Flip-flops have been previously studied mainly with coarse-grained simulations and free energy calculations due to time scale limitations in atomistic simulations~\cite{parisio16}, but direct analyses on cholesterol flip-flops in atomistic simulations have been also emerging in recent years~\cite{gu19,javanainen18,baral20}.
-However, systematic studies on how flip-flop rates depend on membrane properties are still challenging even with the state of the art computational resources.
-The NMRlipids databank can greatly advance such studies by giving an easy access to large amounts of data simultaneously.
-
-\begin{figure}[htb]
-    \centering
-    \includegraphics[width=88mm]{Figures/CholFlipFlops.pdf}
-    \caption{{\bf A} Illustration of cholesterol flip-flop.  
-      {\bf B-D} Cholesterol flip-flops analyzed from the databank as a function of membrane thickness, area per lipid, and acyl chain order.
-      Values from simulations with non-zero flip-flop rates are shown with blue dots. Averages over fixed range of x-axis values are shown with black dots.
-    }
-    \label{fig:permeability}
-\end{figure}
+In conclusion, our results suggest that the bilayer packing has a substantial effect on anisotropic water diffusion in multi-membrane lipid systems.
+Several folds larger anisotropy in membranes with higher lateral density is expected to play a role in pharmagokinetic models not only for water but also for other hydrophilic molecules~\cite{nitsche19}.
+Furthermore, the understanding of this anisotropy may help in developing new MRI imaging methods~\cite{topgaard20}.
 
@@
 
-Applications of MD simulations in understanding biomolecular behaviour are steadily increasing, but protocols for quality evaluation and sharing simulation data fall behind the ones in more elaborated fields, such as structural biology.
-Development of such protocols has been limited by practical challenges, such as lack of universally defined quality measures for MD simulations, incompatible naming conventions in different models and simulation softwares, and high technical requirements for hardware to share large trajectories from MD simulations, as well as incentives for scientist to evaluate and share their MD simulation data.
+The focus of biomolecular simulations is moving from studies of individual molecules to larger complexes and even whole cells and organelles~\cite{johnson15,thornburg22,gupta22}.
+On the other hand, machine learning based models predicting behaviour of biomolecules and automatic approaches to parametrize models are emerging~\cite{jumper21,antila22b}.
+The NMRlipids databank will support the development in all these directions.
+The automatic quality evaluation and ranking of simulations in the NMRlipids databank guides the optimization of new force field parameters and the selection of best parameters for large biomolecular complexes where simulation quality becomes increasingly important due the multiplication of small errors in large simulations.
+On the other hand, the NMRlipids databank serves as a training set for diverse machine learning applications.
+%Most straightforward applications include models that would predict connections between membrane properties that are already stored in the databank. 
+For example, a machine learning model predicting electron density profiles from form factors would be highly useful in interpretation of scattering experiments.
+On the other hand, more elaborated models can be trained to predict arbitrary membrane properties using the data from the NMRlipids databank.
 
-The NMRlipids databank circumvents these challenges by utilizing the overlay databank design and open collaboration approach developed in the NMRlipids project~\cite{botan15}.
-In the overlay databank, the storage of large simulation files is outsourced to existing publicly available repositories while databank itself is essentially a git repository with open access licence containing relevant information about the trajectories, including the location of the raw data.
-This architecture streamlines the construction and maintenance of the databank
-%Building accessible databanks of molecular dynamics simulation data has been challenging due to the required long term support for hardware and software maintenance. 
-%
-%, the demand of hardware can be distributed and open collaboration model reduces the risk for ending software maintenance. 
-%Furthermore, the 
-by distributing the long term commitment for hardware and software maintenance.
-Incentive to share the data is created in the NMRlipids open collaboration by offering authorship in published articles to the contributors.
-%, thereby creating an incentive for contributions. 
-Universal quality measures and naming conventions for molecules and atoms are defined in the NMRlipids databank to enable the automatic quality evaluation and other analyses.
-%of lipid bilayer MD simulations introduced in the NMRlipids databank enables rapid ranking of available simulation models agains experimental NMR and x-ray scattering data. 
-%
-The current NMRlipids databank contains only lipid bilayer simulations, but the concept combining open collaboration and overlay databank structure can be applied also in other fields where similar barriers to establish publicly accessible databanks exist.
-%
-%, thereby not requiring large resources to handle and maintain. 
-%This lowers the barrier for starting such databank as well as for long term storage. 
-%The NMRlipids databank  
-%In addition to all computers where the databank is developed and used, the NMRlipids databank git is stored to Zenodo server, thereby enabling a very cost effective long term storage for the databank.
-
-The NMRlipids databank will be of great benefit to scientists involved in MD simulations.
+Providing programmatic access to large scale MD simulation data in the NMRlipids databank can lead to applications in unprecedented directions in fields where MD simulations are less commonly utilized.
+Here we exemplified this by analysing the anisotropic diffusion of water (Fig.~\ref{fig:permeability}), which is relevant in pharmacokinetic modeling and in MRI imaging where MD simulations are not yet commonly used~\cite{nitsche19,topgaard20}.
+%The NMRlipids databank will be of great benefit to scientists involved in MD simulations.
 They will be able to rapidly evaluate the quality of MD simulations in order to filter out potentially misleading results or facilitate force field parameter development~\cite{antila22b}.
 The selection of best models for a specific application is demonstrated for PC/PE lipid mixtures in Fig.~\ref{fig:quality}, and previously for PC/PS lipid mixtures~\cite{antila22b}, and lipid headgroups~\cite{bacle21}.
-On the other hand, programmatic access in the NMRlipids databank enables automatic analyses over large sets of simulation data containing more data in terms of quantity (e.g., simulation length and number of conformations) with broader range in terms of content (e.g, lipid compositions and ion concentrations) that can be accessible for a single research group.
-Added value of the access to broad range of systems is exemplified in Figs.~\ref{fig:quality}~and~\ref{fig:permeability} revealing correlations between membrane properties, while large quantity of data enables the detection of rare phenomena with good statistics, as exemplified for water permeability in Fig.~\ref{fig:permeability} and cholesterol flip-flop events in Fig.~\ref{??}.
+On the other hand, programmatic access in the NMRlipids databank enables automatic analyses over larger sets of simulation data in terms of quantity (e.g., simulation length and number of conformations) and content (e.g, lipid compositions and ion concentrations) that is currently possible in a single research group.
+Added value of such access to simulation data is demonstrated here for the analysis of simulation qualities (Fig.~\ref{fig:quality}), water permeation through membranes (Fig.~\ref{fig:permeability}) and cholesterol flip-flop events (Fig.~\ref{fig:flip-flops}).
+%Such models could be used to connect membrane composition to relevant membrane properties in wide range of fields from molecular biology to drug development and biomolecular imaging. 
+Different types of applications enabled by the NMRlipids databank in wide range of fields are exemplified in Table~\ref{tab:applications}.
+The increasing amount of data is expected to further increase the scope of potential applications of the NMRlipids databank in wide range of fields ranging from molecular biology and biotechnology to
+%the design of lipid nanoparticles for drug delivery or other applications, and 
+material science and biomolecular imaging.
 
-Providing programmatic access to large scale MD simulation data can lead to applications in unprecedented directions.
-For example, water anisotropic diffusion analyzed in Fig.~\ref{fig:permeability} is relevant not only in pharmacokinetics but also in the development of MRI imaging methods where MD simulations are not yet widely used~\cite{topgaard20}.
-We expect the access provided by the NMRlipids databank to facilitate novel applications also in other fields where MD simulations are less commonly applied, such as design of lipid nanoparticles for drug delivery or other applications, and material science.
-The increasing amount of data in terms of quantity and content will also increase the scope of potential applications of the NMRlipids databank in time.
-
-The NMRlipids databank can also serve as a training set for diverse machine learning applications.
-Most straightforward applications include models that would predict connections between membrane properties that are already stored in the databank.
-For example, a machine learning model predicting electron density profiles from form factors would be highly useful in interpretation of scattering experiments.
-On the other hand, more elaborated models could be trained to predict arbitrary membrane properties using the data from the NMRlipids databank.
-Such models could be used to connect membrane composition to relevant membrane properties in wide range of fields from molecular biology to drug development and biomolecular imaging.
-The overlay databank design can open such avenues also for other biomolecules than lipids, such as disordered proteins and membrane-protein systems.
-
-Different types of applications enabled by the NMRlipids databank in wide range of fields are exemplified in Table~\ref{tab:applications}.
-
-\begin{table}[]
+\begin{table}[t]
     \centering
@@
     Guidance for force field development & 
-    Improvements in ion binding to lipids~\cite{melcr18,melcr20} and lipid headgroup conformational ensembles~\cite{??} &
+    Improvements in ion binding to lipids~\cite{melcr18,melcr20} and lipid headgroup conformational ensembles~\cite{yu21,dickson22,grote20} &
     Scientists developing parameters for MD simulations \\
@@
 
+Main practical limitations in building open access databanks of molecular dynamics simulations have been the required commitment in long term support for hardware and software maintenance, and the lack of incentives for researchers to share the data.
+The NMRlipids databank circumvents these challenges with the overlay databank design and open collaboration approach developed in the NMRlipids project~\cite{botan15}.
+In this model, the file storage is distributed to publicly available stable repositories ({\it Data layer} in Fig.~\ref{fig:overlay} A) and maintenance of the databank does not depend on individual scientists or groups because all its version controlled content is available with an open access licence ({\it Databank layer} in Fig.~\ref{fig:overlay} A).
+%containing relevant information about the trajectories, including the location of the raw data. This architecture streamlines the construction and maintenance of the databank
+%
+%, the demand of hardware can be distributed and open collaboration model reduces the risk for ending software maintenance. 
+%Furthermore, the 
+%by distributing the long term commitment for hardware and software maintenance. 
+Incentive to share the data is created in the NMRlipids open collaboration by offering authorship in published articles to the contributors following the NMRlipids project protocol~\cite{botan15}.
+%, thereby creating an incentive for contributions. 
+%
+%Universal quality measures and naming conventions for molecules and atoms are defined in the NMRlipids databank to enable the automatic quality evaluation and other analyses.
+%
+%of lipid bilayer MD simulations introduced in the NMRlipids databank enables rapid ranking of available simulation models agains experimental NMR and x-ray scattering data. 
+%
+The current NMRlipids databank contains only lipid bilayer simulations, but the concept can be applied also to other biomolecules, such as disordered proteins and membrane-protein systems,
+%combining open collaboration and overlay databank structure can be applied also 
+or other fields where similar barriers to establish publicly accessible databanks exist. 
+%The overlay databank design can open such avenues also for other
+
+%
+%, thereby not requiring large resources to handle and maintain. 
+%This lowers the barrier for starting such databank as well as for long term storage. 
+%The NMRlipids databank  
+%In addition to all computers where the databank is developed and used, the NMRlipids databank git is stored to Zenodo server, thereby enabling a very cost effective long term storage for the databank.
+
+%\newpage
+
 \section{Methods}
@@
 Structure of the NMRlipids databank is illustrated more detailed in Fig.~\ref{DatabankStructure} in the supplementary information.
-The databank locates as a git repository at \url{https://github.com/NMRLipids/Databank/} and is permanently stored in Zenodo repository (\url{www.zenodo.org})~\cite{??}.
+The core content of the databank ({\it Databank layer} in Fig.~\ref{fig:overlay}) locates as a git repository at \url{https://github.com/NMRLipids/Databank/} and is permanently stored in Zenodo repository (\url{www.zenodo.org})~\cite{??}.
 Whenever a specific file is referred here, the file path within the NMRlipids databank repository is given.
@@
 
-The core content of the NMRlipids databank is stored in README.yaml files located at {\it /Data/Simulations} in the NMRlipids databank repository.
-Essential information for each simulation entry in the databank is stored in these files.
-These yaml files are both human and machine readable, and contain access to all information that are needed for further analyses of the simulations.
-The keys available in these files and their explanations are listed in table~\ref{tab:READMEkeys} in the supplementary information.
-To keep the databank light, the raw MD simulation data is stored in an external location which can be any stable publicly available repository, although all the current data locates in Zenodo \url{www.zenodo.org}.
-Whenever needed, the raw data can be downloaded through the links stored in the README.yaml files.
+Essential information of each simulation in the NMRlipids databank is stored in a human and machine readable README.yaml file located at {\it /Data/Simulations} in the NMRlipids databank repository.
+These files contain access to all information that are needed for further analyses of simulations.
+The content of these files is described in detail in table~\ref{tab:READMEkeys} in the supplementary information.
+%To keep the databank light, 
+Raw MD simulation data are stored in external publicly available and stable repositories ({\it Data layer} in Fig.~\ref{fig:overlay}), such as Zenodo (\url{www.zenodo.org}), from where it can be downloaded whenever needed using the links in README.yaml files.
+
+\ref{DatabankStructure}
+\ref{fig:permeationSI}
 
 \subsection{Molecule and atom naming convention} \label{naming}
-Molecule and atom names used in a simulation are often needed when analyzing the simulation trajectories.
-However, the naming conventions of lipid molecules and atoms therein vary between authors and force fields because universal naming convention has not been defined.
-Because such convention is necessary for automatic analyses over simulations, we have defined the unique naming conventions for lipid molecules and atoms in the NMRlipids databank.
-For each molecule we have defined an unique abbreviation which are listed in table~\ref{tab:abbreviations} in the supplementary information.
-On the other hand, unique names for each atom in a molecule are defined and connected to corresponding names in a simulation using mapping files.
-Mapping files are in yaml format where unique atom names are connected to force field specific names.
-Also residue names and fractions lipids (headgroup, glycerol bakcbone, and acyl chains) can be defined in the mapping file.
-For example, the mapping files used in the NMRlipids databank are located at {\it /Scripts/BuildDatabank/mapping\_files} in the NMRlipids repository.
+When analysing simulations, atoms and molecules needs to be often called by their names defined in the simulation trajectory.
+However, these names often vary between the force fields because the universal naming convention has not been defined for lipids.
+To enable automatic analyses over simulations with different atom and molecule names in the NMRlipids databank, we have defined unique naming conventions for molecules and atoms therein.
+Unique abbreviations used in the NMRlipids databank for each molecule are listed in table~\ref{tab:abbreviations} in the supplementary information.
+Atom names in a simulation are connected to unique atom names defined for each molecule using mapping files located at {\it /Scripts/BuildDatabank/mapping\_files} in the NMRlipids repository.
+%Mapping files are in yaml format where unique atom names are connected to force field specific names. 
+Also residue names for each atom and whether they belong to headgroup, glycerol bakcbone, or acyl chain region are defined in the mapping files.
 For more details about mapping files, see~\url{https://nmrlipids.blogspot.com/2022/04/new-yaml-format-of-mapping-files.html}).
-In practise, information about molecule and atom names are given in the COMPOSITION dictionary in README.yaml files, where force field specific molecule name corresponding the unique abbreviation (table~\ref{tab:abbreviations}) and the name of the mapping file are given.
-
+In practise,
+%information about molecule and atom names are given, where 
+mapping file names and force field specific molecule names corresponding the unique abbreviation (table~\ref{tab:abbreviations}) are given in the COMPOSITION dictionary in README.yaml files. 
 %Upon addition of a new entry in the databank, the molecule names in the simulation corresponding the unique names and mapping files (available at \url{https://github.com/NMRLipids/Databank/tree/main/Scripts/BuildDatabank/mapping_files}) are defined in the dictionary.
@@
 \subsection{Adding data into the NMRlipids databank}
-In order to add a simulation entry into the NMRlipids databank, an initial info.yaml file containing user given information listed in table~\ref{tab:READMEkeys} needs to be created.
-For examples of these files, see {\it /Scripts/BuildDatabank/info\_files} in the NMRlipids databank repository.
-Using this file as an input, the {\it /Scripts/BuildDatabank/AddData.py} script then calculates the rest of the information to be stored in the README.yaml files, also listed in table~\ref{tab:READMEkeys}.
-The databank is open for submission of these info.yaml files via pull requests to the git repository.
-Large fraction of the current content is composed of the data uploaded to Zenodo in previous NMRlipids projects~\cite{botan15,catte16,antila19,bacle21}, yet also other simulation data from the Zenodo repository are added.
+Anyone can add a new simulation into the NMRlipids databank by first creating an info.yaml file containing the manually entered information listed in table~\ref{tab:READMEkeys} and then adding this file to {\it /Scripts/BuildDatabank/info\_files} folder in the NMRlipids databank repository via pull requests.
+Pull requests are manually accepted and automatically extracted data for README.yaml files,  listed in table~\ref{tab:READMEkeys}, will be calculated using the {\it /Scripts/BuildDatabank/AddData.py} script.
+Current content of the NMRlipids databank is composed of the simulations found from Zenodo repository with an appropriate licence.
+Most, but not all, of these trajectories originate from previous NMRlipids projects~\cite{botan15,catte16,antila19,bacle21}.
 
@@
 The keys and their descriptions for the experimental data are given in table~\ref{tab:READMEkeysEXP}.
-NMR data currently in the NMRlipids databank are taken from Refs.~\citenum{??} and x-ray scattering data from Refs.~\citenum{??}.
+NMR data currently in the NMRlipids databank are taken from Refs.~\citenum{scherer87,ferreira13,antila19,melcr20,bacle21} and x-ray scattering data from Refs.~\citenum{Kucerka08a,kucerka11,pan12b,pan14,kucerka15}.
 In addition, some previously unpublished NMR and x-ray scattering data are used.
 These are measured with previously established methods as described in the supplementary information.
+
+\subsection{Analyzing the databank}
+Programs that analyze large data sets from the NMRlipids databank loop over README.yaml files of all simulations, download the raw simulation data to the local computer, and perform the analysis from the downloaded trajectories.
+When molecule or atom names are needed in the analysis, they will be read from the README.yaml file and mapping files defined therein.
 
@@
 \end{equation}
-where $\theta$ is the angle between the C--H bond and the membrane normal
-%(taken to align with $z$, with bilayer periodicity in the $xy$-plane).
-and angular brackets denote the ensemble average, i.e., average over all sampled configurations of all lipids in a simulation.
+where angular brackets denote the ensemble average, i.e., average over all sampled configurations of all lipids in a simulation, and $\theta$ is the angle between the C--H bond and the membrane normal.
 As in previous NMRlipids publications, the order parameters were first calculated separately for each lipid and the standard error of the mean over different lipids was used as the error bar~\cite{botan15}.
-
-In practise, the analysis code loops over all README.yaml files (i.e., simulations in the NMRlipids databank), downloads the raw simulation to a local computer, uses the information about the atom and molecule naming conventions in README.yaml to determine the location of carbon and hydrogen atoms, and then calculates the order parameters from Eq.~\ref{OP}.
-The script to calculate all C-H bond order parameters from the NMRlipids databank is available at {\it /Scripts/AnalyzeDatabank/calcOrderParameters.py} in the NMRlipids databank repository.
-The resulting order parameters are stored for all simulations in files named \newline
+The script that calculates C-H bond order parameters from all simulations in the NMRlipids databank is available at {\it /Scripts/AnalyzeDatabank/calcOrderParameters.py} in the NMRlipids databank repository.
+The resulting order parameters are stored for all simulations in files named %\newline
 [{\it lipid\_name}]{\it OrderParameters.json} at folders in {\it /Data/Simulations} in the NMRlipids databank repository.
@@
 where $\Delta \rho_e(z)$ is the difference between total and solvent electron densities, and $D$ is the simulation box size in the z-direction.
-For the calculation of density profiles, atom coordinates were centred around the centre of mass of the POPC lipid molecules for every time frame, and a histogram of these centred positions was calculated with the bin width of $1/3$~Å.
-Similarly to order parameters, the script loops over all simulations, downloads the data on local computer, and uses the information in README.yaml file to calculate form factors.
+For the calculation of density profiles, atom coordinates were first centred around the centre of mass of lipid molecules for every time frame, and a histogram of these centred positions weighted with the number of electrons in each atom was then calculated with the bin width of $1/3$~Å.
+%Similarly to order parameters, the script loops over all simulations, downloads the data on local computer, and uses the information in README.yaml file to calculate form factors. 
 The script to calculate form factors for all simulations in the NMRlipids databank is available at {\it Scripts/AnalyzeDatabank/calc\_FormFactors.py}.
@@
 
-\subsection{Calculation of membrane thickness, and area per lipids}
-Lipid bilayer thicknesses in all simulations are calculated from the intersections of lipid and water electron densities using the script {\it Scripts/AnalyzeDatabank/calc\_thickness.py} at the NMRlipids databank repository.
-The resulting thicknesses are stored for all simulations in files named {\it thickness.json} at folders in {\it /Data/Simulations} in the NMRlipids databank repository.
-Lipid bilayer area per lipids in all simulations are calculated by dividing the simulation box area with the total number of molecules considered as lipids using the script {\it Scripts/AnalyzeDatabank/calcAPL.py} at the NMRlipids databank repository.
-The resulting area per lipids are stored for all simulations in files named {\it apl.json} at folders in {\it /Data/Simulations} in the NMRlipids databank repository.
+\subsection{Calculation of a bilayer area per lipid and thickness}
+Area per lipids of bilayers were calculated by dividing the area of the simulation box with the total number of lipid molecules in the simulation.
+The script that calculates area per lipids from all simulations in the NMRlipids databank repository is available at {\it Scripts/AnalyzeDatabank/calcAPL.py} in the NMRlipids databank repository.
+The resulting area per lipids are stored for all simulations in files named {\it apl.json} at folders in {\it /Data/Simulations}.
+
+Thicknesses of lipid bilayers were calculated from the intersections of lipid and water electron densities.
+The script that calculates thickness of all simulations in the NMRlipids databank is available at {\it Scripts/AnalyzeDatabank/calc\_thickness.py} in the NMRlipids databank repository.
+The resulting thicknesses are stored in files named {\it thickness.json} at folders in {\it /Data/Simulations} in the NMRlipids databank repository.
 
@@
 Student's t-distribution, $\mu$ is the real mean of the order parameter, $n$ is the number of independent sample points for each C-H bond which equals the number of lipids in a simulation, $S_{\rm CH}$ is the sample mean from Eq.~\ref{OP}, $s$ is the variance of $S_{\rm CH}$ calculated over individual lipids, $S_{\rm exp}$ is the experimental value, and $\Delta S_{\rm exp}$ its error.
-The error of $\Delta S_{\rm exp} = 0.02$ is currently assumed for all experimental order parameters~\cite{ollila16}.
+The error of $\Delta S_{\rm exp} = 0.02$ is currently assumed for all experimental order parameters~\cite{ollila16}, yet more accurate may be available in the future~\cite{wurl22}.
 Because lipid bilayer simulations contain at least dozens of lipids, the Student's t-distribution could be safely approximated with a normal distribution.
@@
 
-Water permeability through membranes was calculated from equation $P=r/2c_w$, where $r$ is the rate of permeation events per time and area, and $c_w$ is the concentration of water in bulk with the value of 33.3679\,(nm)$^{-3}$~\cite{venable19}.
+Water permeability through membranes was calculated from equation $P=r/2c_w$, where $r$ is the rate of permeation events per time and area, and $c_w$=33.3679\,(nm)$^{-3}$ which is the concentration of water in bulk~\cite{venable19}.
 The number of permeation events in each trajectory was calculated using the code by Camilo et al.~\cite{camilo2022}, available at~\url{https://github.com/crobertocamilo/MD-permeation}.
-This code was used within the loop that goes over all simulations in the NMRlipids databank and extracts the required information using the information in README.yaml files.
-The code is available at {\it /scripts/calcMD-PERMEATION.py} and resulting permeabilities are stored at {\it /Data/MD-PERMEATION} in the repository containing all analyses specific for this publication at~\url{https://github.com/NMRLipids/DataBankManuscript/}.
+%This code was used within the loop that goes over all simulations in the NMRlipids databank and extracts the required information using the information in README.yaml files. 
+The code that calculates permeabilities for all simulations in the NMRlipids databank is available at {\it /scripts/calcMD-PERMEATION.py} and the resulting permeabilities are stored at {\it /Data/MD-PERMEATION} in the repository containing all analyses specific for this publication at~\url{https://github.com/NMRLipids/DataBankManuscript/}.
 %While the order parameters, form factors, area per lipid and thickness are stored within the NMRlipids databank (\url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations}), further analyses can be conventiently stored in separate repositories with the same folder structure based on hash identities of trajectory and topology files.
 For example, results from further analyses performed here are stored in folders at \url{https://github.com/NMRLipids/DataBankManuscript/tree/main/Data}.
-This repository is organized similarly to the NMRlipids databank repository, enabling the upcycling also of the analyzed data without overloading the main NMRlipids databank repository.
+This repository is organized similarly to the NMRlipids databank repository, enabling the upcycling of also the analyzed data without overloading the main NMRlipids databank repository.
 
-The parallel diffusion along membrane surface was calculated using Einstein's equation with {\it gmx msd} program with the {\it -lateral} option from Gromacs softwater package~\cite{gromacsMANUAL}.
-The code is available at {\it /scripts/calcWATERdiffusion.py} and resulting diffusion coefficients are stored at {\it /Data/WATERdiffusion} in the repository at~\url{https://github.com/NMRLipids/DataBankManuscript/}.
+The lateral diffusion of water along the membrane surface, $D_{||}$, was calculated with Einstein's equation using {\it -lateral} option in {\it gmx msd} program within the Gromacs software package~\cite{gromacsMANUAL}.
+The code that calculates $D_{||}$ for water from all simulations in the NMRlipids databank is available at {\it /scripts/calcWATERdiffusion.py}, and the resulting diffusion coefficients are stored at {\it /Data/WATERdiffusion} in the repository at~\url{https://github.com/NMRLipids/DataBankManuscript/}.
 
@@
 \subsection{Calculation of lipid flip-flops}
+Flip-flop rates were calculated using {\it AssignLeaflets} and {\it FlipFlop} tools from LiPyphilic package~\cite{LiPyphilic2021}.
+Headgroup atoms of each molecule defined in the mapping file were used to determine in which leaflet the molecule locates.
+The midplane cut-off defining the region between leaflets was 1\,nm and frame cut-off was 100.
+This means that if the headgroup of a molecule entered within the distance of 1\,nm from the bilayer midplane and was found from the opposing leaflet after 100 steps, this event was considered as a successfull flip-flop event.
+A code that finds flip-flop rates from all simulations in the databank is available at {\it scripts/FlipFlop.py} and the resulting flip-flop events at {\it Data/Flipflops/} in the repository at~\url{https://github.com/NMRLipids/DataBankManuscript/}.
 

### 2022-11-23 main.tex
Updates from Overleaf

@@
 
-%\newpage
+\newpage
 
@@
 
-\ref{DatabankStructure}
-\ref{fig:permeationSI}
-
 \subsection{Molecule and atom naming convention} \label{naming}
 When analysing simulations, atoms and molecules needs to be often called by their names defined in the simulation trajectory.
-However, these names often vary between the force fields because the universal naming convention has not been defined for lipids.
+However, these names typically vary between force fields because the universal naming convention has not been defined for lipids.
 To enable automatic analyses over simulations with different atom and molecule names in the NMRlipids databank, we have defined unique naming conventions for molecules and atoms therein.
 Unique abbreviations used in the NMRlipids databank for each molecule are listed in table~\ref{tab:abbreviations} in the supplementary information.
-Atom names in a simulation are connected to unique atom names defined for each molecule using mapping files located at {\it /Scripts/BuildDatabank/mapping\_files} in the NMRlipids repository.
-%Mapping files are in yaml format where unique atom names are connected to force field specific names. 
-Also residue names for each atom and whether they belong to headgroup, glycerol bakcbone, or acyl chain region are defined in the mapping files.
-For more details about mapping files, see~\url{https://nmrlipids.blogspot.com/2022/04/new-yaml-format-of-mapping-files.html}).
-In practise,
-%information about molecule and atom names are given, where 
-mapping file names and force field specific molecule names corresponding the unique abbreviation (table~\ref{tab:abbreviations}) are given in the COMPOSITION dictionary in README.yaml files. 
-%Upon addition of a new entry in the databank, the molecule names in the simulation corresponding the unique names and mapping files (available at \url{https://github.com/NMRLipids/Databank/tree/main/Scripts/BuildDatabank/mapping_files}) are defined in the dictionary.
+Atom names used in simulation trajectories are connected to unique atom names using mapping files that are defined in the NMRlipids project (\url{https://nmrlipids.blogspot.com/2022/04/new-yaml-format-of-mapping-files.html}).
+These files are located at {\it /Scripts/BuildDatabank/mapping\_files} in the NMRlipids databank repository.
+These files also define whether an atom belongs to headgroup, glycerol backbone, or acyl chain region in a lipid.
+In practise, force field specific molecule names and mapping files
+names are given in the COMPOSITION dictionary in README.yaml files for each molecule in each simulation in the NMRlipids databank.
 
 \subsection{Adding data into the NMRlipids databank}
-Anyone can add a new simulation into the NMRlipids databank by first creating an info.yaml file containing the manually entered information listed in table~\ref{tab:READMEkeys} and then adding this file to {\it /Scripts/BuildDatabank/info\_files} folder in the NMRlipids databank repository via pull requests.
-Pull requests are manually accepted and automatically extracted data for README.yaml files,  listed in table~\ref{tab:READMEkeys}, will be calculated using the {\it /Scripts/BuildDatabank/AddData.py} script.
-Current content of the NMRlipids databank is composed of the simulations found from Zenodo repository with an appropriate licence.
+The NMRlipids databank is open for additions of simulation data by anyone.
+The first step is to create an info.yaml file containing the information that needs to be manually entered as listed in table~\ref{tab:READMEkeys}.
+This file can be then added into {\it /Scripts/BuildDatabank/info\_files} folder in the NMRlipids databank repository via pull requests.
+After the pull request is manually accepted, the rest of the information for the README.yaml file, listed in table~\ref{tab:READMEkeys}, will be automatically extracted using the {\it /Scripts/BuildDatabank/AddData.py} script.
+Currently the NMRlipids databank is composed of simulations found from Zenodo repository with an appropriate licence.
 Most, but not all, of these trajectories originate from previous NMRlipids projects~\cite{botan15,catte16,antila19,bacle21}.
@@
 
-The quality of each C--H bond order parameter is estimated by calculating the probability of a simulated value to locate within the error bars of the experimental value.
-Because conformational ensembles of individual lipids are independent in a fluid lipid bilayer,
-$\frac{S_{\rm CH}-\mu}{s/\sqrt{n}}$ has a Student's t-distribution with $n-1$ degrees of freedom and the probability for a order parameter in simulation to locate within experimental error bars can be estimated from equation
+The quality of each C--H bond order parameter is estimated by calculating the probability for a simulated value to locate within the error bars of the experimental value.
+Because conformational ensembles of individual lipids are independent in a fluid lipid bilayer, $\frac{S_{\rm CH}-\mu}{s/\sqrt{n}}$ has a Student's t-distribution with $n-1$ degrees of freedom, and the probability for an order parameter from simulation to locate within experimental error bars can be estimated from equation
 \begin{equation}\label{probability}
@@
 The error of $\Delta S_{\rm exp} = 0.02$ is currently assumed for all experimental order parameters~\cite{ollila16}, yet more accurate may be available in the future~\cite{wurl22}.
-Because lipid bilayer simulations contain at least dozens of lipids, the Student's t-distribution could be safely approximated with a normal distribution.
+Because a lipid bilayer simulation contains at least dozens of lipids, the Student's t-distribution could be safely approximated with a normal distribution.
 However, the normal distribution gives probability values that are below the numerical accuracy of computers when simulation values are far from experiments.
@@
 The first minimum correlates well with the thickness of a membrane (Fig.~\ref{fig:quality} F), although the correlation of the second minima would be even stronger (Fig.~\ref{fig:QualityCorrelationsSI}).
-In practise, we first filter the fluctuations from the form factor data using Savitzky-Golay filter (window lenght 30 and polynomial order 1) and locate the first minima above 0.1\,$AA^2$ from both simulation and experimental data.
+In practise, we first filter the fluctuations from the form factor data using Savitzky-Golay filter (window lenght 30 and polynomial order 1) and locate the first minima above 0.1\,\AA$^{-1}$ from both simulation and experimental data.
 The quality of a form factor is then defined as Eucledian distance between the minima in simulated and experimental form factors, $FF_q = |FF_{\rm{min}}^{\rm{sim}}-FF_{\rm{min}}^{\rm{exp}}|$.
@@
 
-\subsection{Analysing anisotropic diffusion of water in membranes environment from the NMRlipids databank}
+\subsection{Calculation of lipid flip-flops}
+Flip-flop rates were calculated using {\it AssignLeaflets} and {\it FlipFlop} tools from LiPyphilic package~\cite{LiPyphilic2021}.
+Headgroup atoms of each molecule as defined in the mapping file were used to determine in which leaflet the molecule locates.
+The midplane cut-off defining the region between leaflets was 1\,nm and frame cut-off was 100.
+This means that if the headgroup of a molecule entered within the distance of 1\,nm from the bilayer midplane and was found from the opposing leaflet after 100 steps, this event was considered as a successfull flip-flop event.
+The code that finds flip-flop events from all simulations in the NMRlipids databank is available at {\it scripts/FlipFlop.py} and the results at {\it Data/Flipflops/} in the repository at~\url{https://github.com/NMRLipids/DataBankManuscript/}.
+
+\subsection{Analysing anisotropic diffusion of water in membrane environment from the NMRlipids databank}
 
@@
 
-Water permeability through membranes was calculated from equation $P=r/2c_w$, where $r$ is the rate of permeation events per time and area, and $c_w$=33.3679\,(nm)$^{-3}$ which is the concentration of water in bulk~\cite{venable19}.
+Water permeability through membranes was calculated from equation $P=r/2c_w$, where $r$ is the rate of permeation events per time and area, and $c_w$=33.3679\,(nm)$^{-3}$ is the concentration of water in bulk~\cite{venable19}.
 The number of permeation events in each trajectory was calculated using the code by Camilo et al.~\cite{camilo2022}, available at~\url{https://github.com/crobertocamilo/MD-permeation}.
@@
 Water diffusion in the perpendicular direction of lipid bilayers in a multilamellar stack was estimated from the Tanner equation $D_\perp = \frac{D_{||} P z_w}{D_{||} + P z_w}$~\cite{tanner78,wasterby02}, where the water layer thickness, $z_w$, was estimated by subtracting bilayer thickness from the size of the simulation box in membrane normal direction.
-
-\subsection{Calculation of lipid flip-flops}
-Flip-flop rates were calculated using {\it AssignLeaflets} and {\it FlipFlop} tools from LiPyphilic package~\cite{LiPyphilic2021}.
-Headgroup atoms of each molecule defined in the mapping file were used to determine in which leaflet the molecule locates.
-The midplane cut-off defining the region between leaflets was 1\,nm and frame cut-off was 100.
-This means that if the headgroup of a molecule entered within the distance of 1\,nm from the bilayer midplane and was found from the opposing leaflet after 100 steps, this event was considered as a successfull flip-flop event.
-A code that finds flip-flop rates from all simulations in the databank is available at {\it scripts/FlipFlop.py} and the resulting flip-flop events at {\it Data/Flipflops/} in the repository at~\url{https://github.com/NMRLipids/DataBankManuscript/}.
 

### 2022-12-06 main.tex
Update main.tex

@@
 Membrane lipid composition can be also related to diseases, such as cancer and neurodegenerative disorders, and potential therapeutics that affect membrane compositions have been proposed~\cite{torres21}.
-However, membrane containing systems are often difficult to study experimentally because membranes are in disordered fluid state in biological conditions and sample preparation can be highly complicated due to rich phase behaviour and complex interactions between lipids and other biomolecules.
-Therefore, connections between complex lipid interactions and biological functions are still poorly understood.
+However, membrane-containing systems are often difficult to study experimentally because the membranes are in a disordered fluid state under biological conditions and sample preparation can be arduous due to the complicated phase behaviour and rich interactions between lipids and other biomolecules.
+For those reasons, the detailed connections between lipid interactions and biological functions taking place in or around membranes are still poorly understood.
 %why such variety of lipids is needed and how their. 
-Molecular dynamics (MD) simulations have been particularly useful in understanding membrane systems, yet their accuracy is often compromised by the quality of model parameters and other artefacts~\cite{antila22b,gupta22}.
+Molecular dynamics (MD) simulations have been particularly useful in understanding membrane systems, yet their accuracy is often compromised by the quality of model parameters, methodological inadequacy, and other artefacts~\cite{antila22b,gupta22}.
 %While MD simulations of systems with few biomolecules continue having major contributions in our understanding on biology and drug design, 
-On the  other hand, the accuracy of models is becoming increasingly important as researches are progressing from the simulations of 
+On the  other hand, the accuracy of models is becoming increasingly important as the scale (in both size and time) of the systems studied continuously increases 
+%researches are progressing from the simulations of 
 %now searching new avenues by building models not only 
-individual molecules to the models of whole cells or organelles using interdisciplinary approaches~\cite{johnson15,thornburg22,gupta22}.
-In such systems exhibiting very complex behaviour, inaccuracies are more difficult to detect and multiplication of small errors may have major effects the conclusions.
+from the level of individual molecules to models of whole cells or organelles using interdisciplinary approaches~\cite{johnson15,thornburg22,gupta22}.
+Such systems exhibit complex emergent behaviour making inaccuracies more difficult to detect and accumulation of even modest errors may have major effects the conclusions.
 
-In contrast to experimental structural biology where standard protocols to share and quality evaluate the resolved structures are established in the PDB databank~\cite{montelione13}, such best practices are yet to be defined for MD simulations.
-Importance of such approaches is widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} and solutions for sharing data are emerging for proteins in solution~\cite{meyer10,kamp10} or membranes~\cite{newport19,espigares20,leston22}, nucleic acids~\cite{hospital16}, cyclodextrins~\cite{mixcoha16}, and arbitrary biomolecules~\cite{bekker20}.
+In contrast to experimental structural biology, where standard protocols to share and quality evaluate the resolved structures are established in the PDB databank~\cite{montelione13}, the equivalent best practices are yet to be defined for MD simulations.
+The importance of such approaches is widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} and solutions for sharing data are emerging for proteins in solution~\cite{meyer10,kamp10}, proteins in membranes~\cite{newport19,espigares20,leston22}, nucleic acids~\cite{hospital16}, cyclodextrins~\cite{mixcoha16}, and arbitrary biomolecules~\cite{bekker20}.
 However,
 %The relevance of quality evaluation of simulation trajectories in databanks regarding technical details of simulations and accuracy of the underlying physical description of the system (force field) has become evident~\cite{tai04,meyer10,hospital20} and such
-automatic quality evaluation~\cite{meyer10,hospital16} and programmatic access are still rare. 
+a streamlined workflow with automatic quality evaluation~\cite{meyer10,hospital16} and programmatic access is still rare. 
 %However, straightforward quality comparisons between individual simulations or force fields within these databanks remain challenging. 
-Therefore, tools for automatic quality evaluation of membrane simulations or training sets for machine learning models of membrane containing systems are not yet available.
+Therefore, tools for automatic quality evaluation of membrane simulations or training sets for machine learning models of membrane-containing systems are not yet available.
 
@@
 
-Here we present the NMRlipids databank, a community driven open access databank with programmatic access containing atomistic resolution MD simulations of lipid bilayers.
-The programmatic access enables various data-driven approaches, providing new tools for researchers in wide range of fields in academia and industry from cell membrane biology to lipid nanoparticle formulations and data-driven computational chemistry and machine learning.
-Here we demonstrate how data-driven analysis of water anisotropic diffusion in membrane systems from the NMRlipids databank can extend the scope of MD simulations to MRI imaging and pharmacokinetics, and how the NMRlipids databank can be utilized to analyse rare phenomena that are beyond the scope of standard MD simulation studies.
-Furthermore, we perform automatic quality evaluation of membrane simulations which guides the selection of best models for specific applications and development of simulation parameters and methodology.
+Here we present the NMRlipids databank, a community-driven open access databank with programmatic access containing atomistic resolution MD simulations of lipid bilayers.
+The programmatic access enables various data-driven approaches, providing new tools for researchers in a wide range of fields covering academia and industry from cell membrane biology to lipid nanoparticle formulations, data-driven computational chemistry, and machine learning.
+Here, we demonstrate how data-driven analysis of water anisotropic diffusion in membrane systems from the NMRlipids databank can extend the scope of MD simulations to MRI imaging and pharmacokinetics as well as how the NMRlipids databank can be utilized to analyse rare phenomena that are beyond the scope of standard MD simulation investigations.
+Furthermore, we perform automatic quality evaluation of membrane simulations, which guides the selection of best models for specific applications and the improvement of simulation parameters and methodology.
 %The practical relevance of the NMRlipids databank is exemplified by automatic quality evaluation and ranking of large amount of MD simulation data,
 
-While the NMRlipids databank currently contains only bilayer systems, it can be readily expanded to contain other molecules than lipids, such as disordered proteins or sugars. 
+While the NMRlipids databank currently contains only bilayer systems, it can readily be expanded to contain other molecules than lipids, such as disordered proteins or sugars. 
 %a solution for lipid bilayers based on overlay databank structure illustrated in Fig.~\ref{fig:overlay}.  
-The combination of overlay databank structure and open collaboration introduced here can be applied also when building databanks in other fields,  
+The combination of an overlay databank structure and a community-wide open collaboration used here has been applied successfully when building databanks in other fields,  
 %is developed here to solve the practical challenges in generating databanks of MD simulation data enabling flexible analyses over large sets of simulation data, but it can be  potentially used for wide range of situations.
@@
 %overlay databank approach lowers the barrier to start without compromizing the long term stability or scalability.
-Therefore, the approach opens up possibilities to create databanks enabling data-driven and machine learning applications in new fields.
+Thus, the approach has proven itself in enabling data-driven and machine learning applications in new fields.
 

### 2022-12-06 main.tex
Update main.tex

@@
 %overlay databank approach lowers the barrier to start without compromizing the long term stability or scalability.
-Thus, the approach has proven itself in enabling data-driven and machine learning applications in new fields.
+Thus, the approach has already proven itself most useful in enabling data-driven and machine learning applications in new fields.
 
@@
 \subsection{Programmatic access to MD simulation data of membranes composed of most common biologically abundant lipids}
-NMRlipids databank is a community driven databank containing atomistic MD simulations of biologically relevant lipid membranes emerging from the NMRlipids open collaboration~\cite{botan15,ollila16,catte16,antila19,bacle21}.
-Nearly thousand simulation trajectories with the total length approaching half a millisecond can be accessed programmatically using Python API or graphical user interface (\url{http://www.databank.nmrlipids.fi/}).
+NMRlipids databank is a community-driven databank containing atomistic MD simulations of biologically relevant lipid membranes emerging from the NMRlipids open collaboration~\cite{botan15,ollila16,catte16,antila19,bacle21}.
+Nearly a thousand simulation trajectories with the total length approaching half a millisecond can be accessed programmatically using the Python application programming interface (API) or a graphical user interface (GUI) (\url{http://www.databank.nmrlipids.fi/}).
 %NMRlipids databank is the only databank of MD simulations with open programmatic access to its content enabling fully flexible analyses of its content. 
-Following the FAIR principles~\cite{wilkinson16} and NMRlipids project protocol, not only the content and computer programs related to the databank but also the whole construction process of the databank are open access~\cite{botan15}.
+Following the FAIR principles~\cite{wilkinson16} and NMRlipids project protocol, not only the content and computer programs related to the databank, but also the entire construction process of the databank, are open access~\cite{botan15}.
 %The NMRlipids databank will be always open for data and other contributions.
@@
 Currently available single component lipid membranes and binary mixtures in the NMRlipids databank are illustrated in Fig.~\ref{fig:overlay}D.
-While these compose the majority of available trajectories, also mixtures up to five lipids are available.
+While these comprise the majority of available trajectories, mixtures with up to five lipid types are available.
 The distribution of available simulations containing a specific lipid is shown in Fig.~\ref{fig:overlay}C.
-This roughly resembles the biological abundancies of different lipid types, PC being the most common  followed by cholesterol, PE, PS, PG, PI, and other lipids depending on organism and organelle~\cite{vanmeer08}.
-The force fields used in the simulations cover all the essential parameters used in lipid simulations, ranging from highly popular CHARMM36 parameters to united atom and polarizable force fields.
+This roughly resembles the biological abundance of different lipid types with phosphatidylcholine (PC) being the most common followed by cholesterol, phosphatidylethanolamine (PE), phosphatidylserine (PS), phosphatidylglycerol (PG), phosphatidylinositol (PI), and other lipids depending on organism and organelle~\cite{vanmeer08}.
+The force fields used in the simulations cover all the essential parameters commonly used in lipid simulations, ranging from highly popular CHARMM36 parameters to united atom and polarizable force fields.
 Therefore, the averages calculated over the databank can be considered as mean predictions from available lipid models for an average cell membrane.
@@
 The efficient upcycling of large MD simulation trajectories in the NMRlipids databank is enabled by the overlay structure illustrated in Fig.~\ref{fig:overlay}A.
-Raw simulation data in the {\it data layer} can be stored in any publicly available location with long term stability and permanent links to the data, such as digital object identifiers.
-The {\it Databank layer} is the core of the databank containing all the relevant information about the simulations, including links to the raw data, relevant metadata describing the systems, universal naming conventions for lipids and their atoms, computer programs to create and analyse the entries, basic properties calculated from all simulations (area per lipid, C-H bond order parameters, x-ray scattering form factors and membrane thickness), and quality evaluation of simulations against experimental data.
+Raw simulation data in the {\it data layer} can be stored in any publicly available location with long term stability and permanent links to the data (e.g., digital object identifiers (DOIs)).
+The {\it Databank layer} is the core of the databank containing all the relevant information about the simulations, including links to the raw data, relevant metadata describing the systems, universal naming conventions for the lipids and their atoms, computer programs to create and analyse the entries, basic properties calculated from all simulations (area per lipid, C-H bond order parameters, X-ray scattering form factors and membrane thickness), and quality evaluation of simulations against experimental data.
 In practise, this information is stored is the git repository which is currently available at~\url{https://github.com/NMRLipids/Databank}.
-The {\it application layer} contains outputs from the databank, such as graphical user interface (\url{http://www.databank.nmrlipids.fi/}) and results from analyses described in sections below.
-More detailed description on the NMRlipids databank structure is available the supplementary information.
+The {\it application layer} contains outputs from the databank, such as graphical user interface (\url{http://www.databank.nmrlipids.fi/}), and results from analyses described in sections below.
+A more detailed description on the NMRlipids databank structure is available the supplementary information.
 
@@
 %Quality of membrane simulations with different force fields have been evaluated against experimental data during parameterization and in separate comparison studies~\cite{botan15,ollila16,catte16,pluhackova16,perez17,leonard19,??}, but the lack of universal quality measure for membrane simulations complicates the estimations of reliability of simulations.
-Understanding the accuracy of used force field parameters is crucial when estimating the reliability and significance MD simulation results.
-However, the lack of universal quality measures for MD simulations and complicated landscape of force field quality for membranes hamper the quality of simulations which has resulted to many controversial results~\cite{antila22b}.
+Understanding the accuracy of used force field parameters is crucial when assessing the reliability and significance MD simulation results.
+However, the lack of universal quality measures for MD simulations and the complicated landscape of force field quality for membranes hamper the applicability such simulations, resulting in substantial controversies~\cite{antila22b}.
 For example, previous quality evaluations of membrane simulations have concluded that CHARMM36 parameters give the best description for lipid headgroup conformational ensembles~\cite{bacle21}, GROMOS-CKP parameters best capture the membrane packing in POPS bilayers~\cite{antila22b}, and that OPLS3e parameters overcome CHARMM36 in structural quality for POPC but predicts overestimated ion binding~\cite{kurki22}.
 %, thereby being a major obstacle in many applications of membrane MD simulations. 
-To enable the navigation in this complex force field quality landscape, we define here quantitative quality measures that can be used to rapidly find the best available parameters for a specific application or to guide force field development.
+To enable the navigation in this complex force field milieu, we here define quantitative quality measures that can be used to rapidly find the best available parameters for a specific application or to guide the development of improved force fields.
 For detailed definitions of quality measures, see the supplementary information.
@@
 Direct comparisons with experiments for selected simulations are shown in Figs.~\ref{fig:quality} C-D.
-C-H bond order parameters are sensitive to conformational ensembles of individual lipid molecules~\cite{ollila16}, but they correlate well also with the membrane thickness and lateral packing as shown in Fig.~\ref{fig:quality} F.
-This is complemented by comparing the absolute values of x-ray scattering form factors between simulations and experiments that are related to the overall membrane dimensions via electron density profile.
-Particularly, the location of minima in these form factors are correlated with the membrane dimensions as shown in Figs.~\ref{fig:quality} F,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}.
+C-H bond order parameters are sensitive to conformational ensembles of individual lipid molecules~\cite{ollila16} while also correlating well with the membrane thickness and lateral packing as shown in Fig.~\ref{fig:quality} F.
+This is complemented by comparing the absolute values of X-ray scattering form factors between simulations and experiments that are related to the overall membrane dimensions via the electron density profile.
+In particular, the location of minima in these form factors are correlated with the membrane dimensions as shown in Figs.~\ref{fig:quality} F,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}.
 

### 2022-12-06 main.tex
Update main.tex

@@
 %\subsection{Finding the best models for PC and PE lipid mixtures}
-The power of the NMRlipids databank in selecting the best simulation parameters for a specific application is demonstrated for mixtures of PE and PC lipids in Fig.~\ref{fig:POPCPOPEapls} showing area per lipids from available POPC and POPE bilayer simulations in the databank from different force fields at 310\,K.
+The power of the NMRlipids databank in selecting the best simulation parameters for a specific application is demonstrated for mixtures of PE and PC lipids in Fig.~\ref{fig:POPCPOPEapls}, showing areas per lipid from available POPC and POPE bilayer simulations in the databank from different force fields at 310\,K.
 Predictions from different force fields deviate significantly in terms of absolute values and the slope of decrease upon addition of POPE.
-Because the area per lipid correlates with the average order parameter of $\it{sn}$-1 chain (Fig.~\ref{fig:quality} F), simulations with the best predictions for area per lipids can be selected based on order parameter quality evaluation.
-Quality evaluation in Fig.~\ref{fig:quality} and direct comparison in Fig.~\ref{fig:POPC_POPE_dataSI} reveal that Slipids ranks 1st for POPC with a clear difference to others, and 2nd for POPE with only marginally lower quality than GROMOS-CKP, therefore being the best selection for studies where packing effects of PE lipids are relevant.
+Because the area per lipid correlates with the average order parameter of the $\it{sn}$-1 chain (Fig.~\ref{fig:quality} F), simulations with the best predictions for area per lipids can be selected based on order parameter quality evaluation.
+Quality evaluation in Fig.~\ref{fig:quality} and direct comparison in Fig.~\ref{fig:POPC_POPE_dataSI} reveal that Slipids ranks 1st for POPC with a clear advantage to the others, and 2nd for POPE with only marginally lower quality score than GROMOS-CKP and is therefore the best avilable choice for studies where packing effects of PE lipids are important.
 %The direct comparison with experiments in Fig.~\ref{fig:POPC_POPE_dataSI} in the supplementary information reveals that 

### 2022-12-06 main.tex
Update main.tex

@@
 %The direct comparison with experiments in Fig.~\ref{fig:POPC_POPE_dataSI} in the supplementary information reveals that 
-Simulations with CHARMM36 and GROMOS-CKP predict too packed bilayer for POPC (overestimated order in Fig.~\ref{fig:POPC_POPE_dataSI}), while area per lipid for POPC is overstimated in Lipid17.
+Simulations with CHARMM36 and GROMOS-CKP predict overly packed bilayer for POPC (overestimated order in Fig.~\ref{fig:POPC_POPE_dataSI}), while the area per lipid for POPC is overstimated in Lipid17.
 For POPE,
 %order parameters in  Slipids and GROMOS-CKP are close to experiments suggesting that their are per lipids are most reasonable, while
-CHARMM36 and Lipid17 predict too packed membranes.
+CHARMM36 and Lipid17 predict membranes that are too packed.
 
@@
 Altogether, these examples demonstrate how
-the automatic quality evaluation in the NMRlipids databank enables rapid selection of the best simulation parameters to specific applications without extensive and tedious manual force field evaluation.
+the automatic quality evaluation in the NMRlipids databank enables rapid exploration and selection of the best simulation parameters for specific applications without extensive and tedious manual force field evaluation.
 This possibility will promote more reliable simulation results for wide range of applications, support force field development and parametrization of coarse grained force fields against the best atomistic MD simulations.
@@
 Lipid flip-flops from one bilayer leaflet to another play an important role in lipid trafficking and regulating membrane properties~\cite{vanmeer08}.
-Phospholipid flip-flops are slow, occuring with the timescales of hours or days, while cholesterol, diacylglycerol and ceramide flip-flops are faster, yet the reported timescales range between minutes to sub-millisecods~\cite{vanmeer08,steck12,parisio16,gu19}.
+Phospholipid flip-flops are slow when not facilitated by proteins, occuring spontaneously on the timescale of hours or days, while cholesterol, diacylglycerol and ceramide flip-flops are faster.
+Still, the reported timescales range between minutes to sub-millisecods~\cite{vanmeer08,steck12,parisio16,gu19}.
 These timescales were previously accessible only by coarse-grained simulations or free energy calculations~\cite{parisio16}, yet atomistic simulations reporting cholesterol flip-flop events have been published recently~\cite{gu19,javanainen19,baral20}.
 %However, systematic studies on how cholesterol flip-flop rate depends on membrane properties are still beyond the current computational resources available for a single research group. 
-These studies report increase in cholesterol flip-flop rates with increasing unsaturation level and decreasing cholesterol concentration~\cite{gu19,javanainen19}, but correlations between cholesterol flip-flop rates and membrane properties have not been systematically studied.
-Here we demonstrate that the large amount of simulation data available in the NMRlipids databank with programmatic access makes the analysis of these correlations accessible for all.
+These studies report an increase in cholesterol flip-flop rates with increasing unsaturation level and decreasing cholesterol concentration~\cite{gu19,javanainen19}, but correlations between cholesterol flip-flop rates and membrane properties have not been systematically studied.
+Here, we demonstrate that the large amount of simulation data available in the NMRlipids databank with programmatic access makes the analysis of these correlations easily accessible for all.
 
@@
 Cholesterol flip-flop rates range between 0.001-1.6\,$\mu$s$^{-1}$ with the mean of 0.16\,$\mu$s$^{-1}$ and median of 0.07\,$\mu$s$^{-1}$.
-These values are in line with the previously reported values from atomistic MD simulations~\cite{gu19,javanainen19,baral20}.
+These values are consistent with the previously reported values from atomistic MD simulations~\cite{gu19,javanainen19,baral20}.
 Flip-flop rates of DCHOL, 0.2\,$\mu$s$^{-1}$, was close to the average value of cholesterol, while average rates for diacylglycerols DOG and SDG were higher than for cholesterol, 0.4\,$\mu$s$^{-1}$ and 0.5\,$\mu$s$^{-1}$, respectively.
 Flip-flops were not observed for other lipids, giving the upper limits for PC lipid flip-flop rate of 9\,s$^{-1}$ and for ceramide (N-palmitoyl-D-erythro-sphingosine) of 0.002\,$\mu$s$^{-1}$.
-Therefore, data in the NMRlipids databank suggests that the lipid flip-flip rate decreases in the order of diacylglycerols $>$ cholesterol $>$ other lipids including ceramides.
-However, amount of data for diacylgycerols (8 simulations with Lipid17 force field) and ceramide (3 simulations with CHARMM force field) is less than for cholesterol (83 simulations), thus we cannot fully exclude the effect of force field or composition on this comparison.
+Thus, avilable data in the NMRlipids databank suggest that the lipid flip-flip rate decreases in the order of diacylglycerols $>$ cholesterol $>$ other lipids including ceramides.
+However, the amount of data for diacylgycerols (8 simulations with Lipid17 force field) and ceramide (3 simulations with CHARMM force field) is less than for cholesterol (83 simulations) so we cannot fully exclude the effect of force field or composition on this comparison.
 
 Nevertheless, we use the wide range of available simulations to analyse how cholesterol flip-flop rate depends on membrane properties.
-Cholesterol flip-flop rates and their averages over fixed ranges of x-axis values are plotted as a function of membrane thickness, lateral density and order in Figs.~\ref{fig:flip-flops} B-D.
+Cholesterol flip-flop rates and their averages over fixed ranges of $x$-axis values are plotted as a function of membrane thickness, lateral density and order in Figs.~\ref{fig:flip-flops} B-D.
 The results reveal a non-linear correlation between decreasing cholesterol flip-flop rate and membrane packing.
 Flip-flop rates increase with an order of magnitude when membrane packing density decreases and a major jump is observed at low membrane packing.
-The order of magnitude changes in cholesterol flip-flop rate with the membrane composition may have major implications in understanding lipid trafficking and membrane biochemistry~\cite{gu19,baral20}.
+The orders of magnitude change in cholesterol flip-flop rate with the membrane composition may have major implications in understanding lipid trafficking and membrane biochemistry~\cite{gu19,baral20}.
 Because the results from the NMRlipids databank are averaged over large range of membrane compositions and force fields, they show that the strong dependence of cholesterol flip-flop rate on membrane properties is not limited to certain lipid compositions or force fields used in previous studies~\cite{gu19,javanainen19,baral20}.
@@
 
-Cholesterol flip-flops are phenomena whose analyses from MD simulations has only recently became possible for groups with substantial resources and special expertise available~\cite{gu19,javanainen19,baral20}.
-Our analysis demonstrates how the NMRlipids databank makes analyses of such rare phenomena accessible to wide range of scientists in different fields of science and industry who do not have the access to required resources and expertise to produce the large amounts of MD simulation data.
+MD simulation and exploration of the cholesterol flip-flops phenomenon has only recently become possible for research groups with substantial resources and specialized expertise~\cite{gu19,javanainen19,baral20}.
+Our analysis demonstrates how the NMRlipids databank makes analyses of such rare phenomena accessible to wide range of scientists in different fields of science and industry who do not have the access to required resources and expertise to produce the large amounts of MD simulation data required.
 
 \subsection{Extending the scope of MD simulations to new fields using the NMRlipids databank: Water diffusion anisotropy in membrane systems}
-The anisotropy in water diffusion in parallel and perpendicular directions with respect to membranes plays a role in the drug translocation through biological material, particularly in skin \cite{hansen13,wen18,nitsche19,roberts21}, and in MRI imaging~\cite{topgaard20}.
-MD simulations are rarely used to analyze anistropic diffusion of water as only few permeation events for water are typically observed in a single MD simulation trajectory~\cite{venable19,camilo2022}, thereby making the collection of sufficient amount data challenging.
-Here we show that the NMRlipids databank can be used to analyze how anisotropic water diffusion in a multilamellar membrane systems depends on membrane properties.
+The anisotropy in water diffusion in parallel and perpendicular directions with respect to membranes plays a role in the drug translocation through biological material, particularly in skin \cite{hansen13,wen18,nitsche19,roberts21} and in MRI imaging~\cite{topgaard20}.
+MD simulations are rarely used to analyze anistropic diffusion of water since only few permeation events for water are typically observed in a single MD simulation trajectory~\cite{venable19,camilo2022}, making the collection of sufficient data challenging or impossible.
+Here, we show that the NMRlipids databank can be used to systematically analyze how anisotropic water diffusion in a multilamellar membrane systems depends on membrane properties.
 
@@
 
-To this end, we first calculated the water permeability through membranes from all simulations in the NMRlipids databank.
-The resulting non-zero values range between 0.3 and 322\,$\mu$m/s with the mean and median of 14\,$\mu$m/s and 8\,$\mu$m/s, respectively.
-These values agree with the previously reported simulation results~\cite{venable19,camilo2022} and have the same order of magnitude as experimentally determined diffusive permeability coefficients, but are on average larger than the values reported for PC lipids in liquid crystalline phase,~0.19-0.33\,$\mu$m/s~\cite{jansen95}.
-Observed permeabilities and their averages over fixed ranges of values at x-axis are shown in Figs.~\ref{fig:permeability} B-E as a function of temperature, membrane thickness, area per lipid, and acyl chain order.
+To this end, we calculated the water permeability through membranes from all simulations in the NMRlipids databank.
+The resulting non-zero values range between 0.3 and 322\,$\mu$m/s with a mean and median of 14\,$\mu$m/s and 8\,$\mu$m/s, respectively.
+These values agree with the previously reported simulation results~\cite{venable19,camilo2022} and have the same order of magnitude as experimentally determined diffusive permeability coefficients, but are on average larger than the values reported for PC lipids in the liquid crystalline phase,~0.19-0.33\,$\mu$m/s~\cite{jansen95}.
+Observed permeabilities and their averages over fixed ranges of values at $x$-axis are shown in Figs.~\ref{fig:permeability} B-E as a function of temperature, membrane thickness, area per lipid, and acyl chain order.
 %on average, we calculated the average permeabilities over all systems with the value of temperature, thickness, area per lipid, or acyl chain order within a fixed range.
 These averaged permeabilities also shown in Figs.~\ref{fig:permeability} with black dots.
-As expected, the permeability increases with the temperature, giving 17\,$\pm$\,4\,k$_b$T for the average energy barrier for the water permeation from the Arrhenius plot in Fig.~\ref{fig:permeability} B.
+As expected, the permeability increases with the temperature, giving an average energy barrier of 17\,$\pm$\,4\,k$_b$T for the water permeation from the Arrhenius plot in Fig.~\ref{fig:permeability} B.
 On the other hand, the permeability of water decreases on average when membranes become more packed, i.e., with decreasing area per lipid and increasing thickness and acyl chain order (Figs.~\ref{fig:permeability} C-E).
-Permeation of water through bilayers depends on membrane properties also according previous studies, but it is not agreed whether the area per lipid~\cite{nagle08} or bilayer thickness~\cite{frallicciardi22} would be the main parameter determining the permeability.
+Permeation of water through bilayers also depends on membrane properties according previous studies, but there is no established consensus on whether the area per lipid~\cite{nagle08} or bilayer thickness~\cite{frallicciardi22} is the main parameter determining the permeability.
 %is previously reported from simulations and experiments for some systems, but not for all~\cite{mathai07,venable19,frallicciardi22}. 
@@
 
-To analyze how water diffusion anisotropy depends on membrane properties in a multi-lamellar lipid bilayer system, we calculated also the water diffusion parallel to the membrane surface from all simulations in the NMRlipids databank.
+To analyze how water diffusion anisotropy depends on membrane properties in a multi-lamellar lipid bilayer system, we calculated the water diffusion parallel to the membrane surface from all simulations in the NMRlipids databank.
 The parallel diffusion coefficient of water, $D_{||}$, decreases with reduced hydration and increases with the temperature, but dependencies on membrane area per lipid, thickness, or fraction of charged lipids were not observed in Figs.~\ref{fig:permeability} and~\ref{fig:diffusionSI}.
@@
 %The results at different hydration levels are shown in Fig.~\ref{fig:permeability} F together with the experimental data~\cite{rudakova04}. 
-Simulation results are close to the experimental values with low hydration levels in Fig.~\ref{fig:permeability} F, but increase to the values of approximately two times higher than experimental bulk water diffusion value (3.1$\cdot$\,10$^{-9}$\,m$^2$/s at 313\,K~\cite{khakimov08}) with high hydration levels.
-This is not surprising as the most common water model used in membrane simulations, TIP3P, overestimates the bulk water diffusion~\cite{pathirannahalage21}.
+Simulation results are close to the experimental values with low hydration levels in Fig.~\ref{fig:permeability} F, but increase to approximately two times higher than the experimental value for bulk water diffusion (3.1$\cdot$\,10$^{-9}$\,m$^2$/s at 313\,K~\cite{khakimov08}) with high hydration levels.
+This is not surprising as the most common water model used in membrane simulations, TIP3P, is known to overestimate the bulk water diffusion~\cite{pathirannahalage21}.
 To estimate the diffusion anisotropy of water, $D_{\perp}/D_{||}$, in multi-lamellar membrane system, the permeability coeffients of water through membranes were translated to perpendicular diffusion coefficients, $D_{\perp}$, using the Tanner equation~\cite{tanner78,wasterby02}.
-The resulting perpendicular diffusion coefficients are approximately five orders of magnitude smaller than lateral diffusion coefficients of water (Figs.~\ref{fig:permeability} G and H), which is at the upper limit of the anisotropic estimated from the experimental data~\cite{nitsche19}.
+The resulting perpendicular diffusion coefficients are approximately five orders of magnitude smaller than lateral diffusion coefficients of water (Figs.~\ref{fig:permeability} G and H), which is at the upper limit of the anisotropic diffusion as determined by experiments~\cite{nitsche19}.
 %Large anisotropy values are understandable as simulations give slightly slower permeation rates and higher lateral diffusion rates than experiments. 
@@
 In conclusion, our results suggest that the bilayer packing has a substantial effect on anisotropic water diffusion in multi-membrane lipid systems.
-Several folds larger anisotropy in membranes with higher lateral density is expected to play a role in pharmagokinetic models not only for water but also for other hydrophilic molecules~\cite{nitsche19}.
-Furthermore, the understanding of this anisotropy may help in developing new MRI imaging methods~\cite{topgaard20}.
+Several folds larger anisotropy in membranes with higher lateral density is expected to play a role in pharmagokinetic models, not only for water but also for other hydrophilic molecules~\cite{nitsche19}.
+Furthermore, the enhanced understanding of this anisotropy may help in developing new MRI imaging methods~\cite{topgaard20}.
 

### 2022-12-06 main.tex
Update main.tex

@@
     C)-E) Evaluation against experimental data exemplified for a simulation with the best overall order parameter quality (C), the best quality for POPE lipid (D), and the headgroup quality for POPE (E).
-    F) Scatter plots and Pearson correlation coefficients for the membrane area per lipid, thickness, first minima of x-ray scattering form factor and average order parameter of the {\it sn}-1 acyl chain extracted from the NMRlipids databank.
+    F) Scatter plots and Pearson correlation coefficients for the membrane area per lipid, thickness, first minima of X-ray scattering form factor and average order parameter of the {\it sn}-1 acyl chain extracted from the NMRlipids databank.
     All correlation coefficients have p-value below 0.001. For more correlations see Fig.~\ref{fig:QualityCorrelationsSI}.
@@
 %best simulation, POPC with OPLS3e, is shown in
-%based on C-H bond order parameters from NMR experiments and x-ray scattering form factors that enable the rapid quality evaluation of membrane MD simulations in the NMRlipids databank.
+%based on C-H bond order parameters from NMR experiments and X-ray scattering form factors that enable the rapid quality evaluation of membrane MD simulations in the NMRlipids databank.
 
@@
 
-%NMRlipids databank contains also experimental order parameter and x-ray form factor data that is connected to corresponding  simulations in order to define the simulation qualities and rank them to select the suitable simulation models for particular applications.
+%NMRlipids databank contains also experimental order parameter and X-ray form factor data that is connected to corresponding  simulations in order to define the simulation qualities and rank them to select the suitable simulation models for particular applications.
 
@@
 On the other hand, locations of form factor minima have positive correlation with the membrane area and negative correlation with the thickness.
-In conclusion, both x-ray scattering form factors and acyl chain order parameters from NMR are on average good proxies for membrane packing and thickness, while NMR order parameters give segmental resolution information on the quality of conformational ensembles of lipids.
+In conclusion, both X-ray scattering form factors and acyl chain order parameters from NMR are on average good proxies for membrane packing and thickness, while NMR order parameters give segmental resolution information on the quality of conformational ensembles of lipids.
 
@@
 The focus of biomolecular simulations is moving from studies of individual molecules to larger complexes and even whole cells and organelles~\cite{johnson15,thornburg22,gupta22}.
-On the other hand, machine learning based models predicting behaviour of biomolecules and automatic approaches to parametrize models are emerging~\cite{jumper21,antila22b}.
+Simultaneously, machine learning-based models predicting behaviour of biomolecules and automatic approaches to parametrize models are emerging~\cite{jumper21,antila22b}.
 The NMRlipids databank will support the development in all these directions.
-The automatic quality evaluation and ranking of simulations in the NMRlipids databank guides the optimization of new force field parameters and the selection of best parameters for large biomolecular complexes where simulation quality becomes increasingly important due the multiplication of small errors in large simulations.
-On the other hand, the NMRlipids databank serves as a training set for diverse machine learning applications.
+The automatic quality evaluation and ranking of simulations in the NMRlipids databank guides the optimization of new force field parameters and the selection of best parameters for large biomolecular complexes where simulation quality becomes increasingly important due the accumulation of small errors in large simulations.
+Furthermore, the NMRlipids databank serves as a training set for diverse machine learning applications.
 %Most straightforward applications include models that would predict connections between membrane properties that are already stored in the databank. 
 For example, a machine learning model predicting electron density profiles from form factors would be highly useful in interpretation of scattering experiments.
-On the other hand, more elaborated models can be trained to predict arbitrary membrane properties using the data from the NMRlipids databank.
+In addition, more elaborated models can be trained to predict arbitrary membrane properties using the data from the NMRlipids databank.
 
 Providing programmatic access to large scale MD simulation data in the NMRlipids databank can lead to applications in unprecedented directions in fields where MD simulations are less commonly utilized.
-Here we exemplified this by analysing the anisotropic diffusion of water (Fig.~\ref{fig:permeability}), which is relevant in pharmacokinetic modeling and in MRI imaging where MD simulations are not yet commonly used~\cite{nitsche19,topgaard20}.
+Here, we exemplified this by analysing the anisotropic diffusion of water (Fig.~\ref{fig:permeability}), which is relevant in pharmacokinetic modeling and in MRI imaging, both areas where MD simulations are not yet commonly used~\cite{nitsche19,topgaard20}.
 %The NMRlipids databank will be of great benefit to scientists involved in MD simulations.
@@
 The selection of best models for a specific application is demonstrated for PC/PE lipid mixtures in Fig.~\ref{fig:quality}, and previously for PC/PS lipid mixtures~\cite{antila22b}, and lipid headgroups~\cite{bacle21}.
-On the other hand, programmatic access in the NMRlipids databank enables automatic analyses over larger sets of simulation data in terms of quantity (e.g., simulation length and number of conformations) and content (e.g, lipid compositions and ion concentrations) that is currently possible in a single research group.
-Added value of such access to simulation data is demonstrated here for the analysis of simulation qualities (Fig.~\ref{fig:quality}), water permeation through membranes (Fig.~\ref{fig:permeability}) and cholesterol flip-flop events (Fig.~\ref{fig:flip-flops}).
+The programmatic access in the NMRlipids databank enables automatic analyses over larger sets of simulation data in terms of quantity (e.g., simulation length and number of conformations) and content (e.g, lipid compositions and ion concentrations) than is currently possible by a single research group.
+The added value that access to this simulation data provides is demonstrated here for the analysis of simulation qualities (Fig.~\ref{fig:quality}), water permeation through membranes (Fig.~\ref{fig:permeability}) and cholesterol flip-flop events (Fig.~\ref{fig:flip-flops}).
 %Such models could be used to connect membrane composition to relevant membrane properties in wide range of fields from molecular biology to drug development and biomolecular imaging. 
-Different types of applications enabled by the NMRlipids databank in wide range of fields are exemplified in Table~\ref{tab:applications}.
-The increasing amount of data is expected to further increase the scope of potential applications of the NMRlipids databank in wide range of fields ranging from molecular biology and biotechnology to
+Different types of applications enabled by the NMRlipids databank across a wide range of fields are exemplified in Table~\ref{tab:applications}.
+The ever growing amount of data is expected to further increase the scope of potential applications of the NMRlipids databank in molecular biology, biotechnology,
 %the design of lipid nanoparticles for drug delivery or other applications, and 
-material science and biomolecular imaging.
+material science, biomolecular imaging and beyond.
 
@@
 
-Main practical limitations in building open access databanks of molecular dynamics simulations have been the required commitment in long term support for hardware and software maintenance, and the lack of incentives for researchers to share the data.
+The main practical limitations in building open access databanks of molecular dynamics simulations have been the required commitment in long term support for hardware and software maintenance, and the lack of incentives for researchers to share the data.
 The NMRlipids databank circumvents these challenges with the overlay databank design and open collaboration approach developed in the NMRlipids project~\cite{botan15}.
-In this model, the file storage is distributed to publicly available stable repositories ({\it Data layer} in Fig.~\ref{fig:overlay} A) and maintenance of the databank does not depend on individual scientists or groups because all its version controlled content is available with an open access licence ({\it Databank layer} in Fig.~\ref{fig:overlay} A).
+In this model, the file storage is distributed to publicly available stable repositories ({\it Data layer} in Fig.~\ref{fig:overlay} A) and maintenance of the databank does not depend on individual scientists or groups because all its version-controlled content is available with an open access licence ({\it Databank layer} in Fig.~\ref{fig:overlay} A).
 %containing relevant information about the trajectories, including the location of the raw data. This architecture streamlines the construction and maintenance of the databank
@@
 %
-%of lipid bilayer MD simulations introduced in the NMRlipids databank enables rapid ranking of available simulation models agains experimental NMR and x-ray scattering data. 
+%of lipid bilayer MD simulations introduced in the NMRlipids databank enables rapid ranking of available simulation models agains experimental NMR and X-ray scattering data. 
 %
-The current NMRlipids databank contains only lipid bilayer simulations, but the concept can be applied also to other biomolecules, such as disordered proteins and membrane-protein systems,
+The current NMRlipids databank contains only lipid bilayer simulations, but the concept can also be applied to other biomolecules, such as disordered proteins and membrane-protein systems,
 %combining open collaboration and overlay databank structure can be applied also 
@@
 After the pull request is manually accepted, the rest of the information for the README.yaml file, listed in table~\ref{tab:READMEkeys}, will be automatically extracted using the {\it /Scripts/BuildDatabank/AddData.py} script.
-Currently the NMRlipids databank is composed of simulations found from Zenodo repository with an appropriate licence.
+Currently, the NMRlipids databank is composed of simulations found from Zenodo repository with an appropriate licence.
 Most, but not all, of these trajectories originate from previous NMRlipids projects~\cite{botan15,catte16,antila19,bacle21}.
@@
 \subsection{Experimental data}
-Experimental data used in the quality evaluation, currently composed of C-H bond order parameters and x-ray scattering form factors, are stored in {\it /Data/experiments} in the NMRlipids databank repository.
+Experimental data used in the quality evaluation, currently composed of C-H bond order parameters and X-ray scattering form factors, are stored in {\it /Data/experiments} in the NMRlipids databank repository.
 Similarly to simulations, each experimental data set has a README.yaml file containing all the relevant information about the experiment.
 The keys and their descriptions for the experimental data are given in table~\ref{tab:READMEkeysEXP}.
-NMR data currently in the NMRlipids databank are taken from Refs.~\citenum{scherer87,ferreira13,antila19,melcr20,bacle21} and x-ray scattering data from Refs.~\citenum{Kucerka08a,kucerka11,pan12b,pan14,kucerka15}.
-In addition, some previously unpublished NMR and x-ray scattering data are used.
+NMR data currently in the NMRlipids databank are taken from Refs.~\citenum{scherer87,ferreira13,antila19,melcr20,bacle21} and X-ray scattering data from Refs.~\citenum{Kucerka08a,kucerka11,pan12b,pan14,kucerka15}.
+In addition, some previously unpublished NMR and X-ray scattering data are used.
 These are measured with previously established methods as described in the supplementary information.
@@
 
-\subsection{Calculation of x-ray scattering form factors}
+\subsection{Calculation of X-ray scattering form factors}
 X-ray scattering form factors were calculated with standard equation for symmetric lipid bilayers~\cite{ollila16}
@@
 
-\subsection{Quality evaluation of x-ray scattering form factors}
-%While C-H bond order parameters relate to the conformational ensembles of individual lipids, the x-ray scattering form factors depend on membrane dimensions and density distribution~\cite{ollila16}.
+\subsection{Quality evaluation of X-ray scattering form factors}
+%While C-H bond order parameters relate to the conformational ensembles of individual lipids, the X-ray scattering form factors depend on membrane dimensions and density distribution~\cite{ollila16}.
 Because experiments give form factors only in relative intensity scale, they should scaled before comparing with the simulation data.
@@
 
-Also a quality measure based on differences in simulated and experimental form factors accross available q-range is defined in the SIMtoEXP program~\cite{kucerka10}.
+Also, a quality measure based on differences in simulated and experimental form factors accross available q-range is defined in the SIMtoEXP program~\cite{kucerka10}.
 However, the lobe heigths in simulated form factors depend on the simulation box size as shown in Fig.~\ref{fig:sizedependence}.
@@
 Headgroup atoms of each molecule as defined in the mapping file were used to determine in which leaflet the molecule locates.
-The midplane cut-off defining the region between leaflets was 1\,nm and frame cut-off was 100.
+The midplane cut-off defining the region between leaflets was 1\,nm and frame cutoff was 100.
 This means that if the headgroup of a molecule entered within the distance of 1\,nm from the bilayer midplane and was found from the opposing leaflet after 100 steps, this event was considered as a successfull flip-flop event.

### 2023-01-03 SI.tex
Updates from Overleaf

@@
 
-\begin{table}[]
+\begin{table}[!h]
     \centering
@@
 
-\begin{table}[h]
+\begin{table}[!h]
     \centering
-    \begin{tabular}{c&c}
+    \begin{tabular}{c c}
     Abbreviation & Molecule name \\
@@
 
-\pagebreak
+\newpage
 \section{Experimental data}
@@
 Acyl chain order parameters of POPE and POPG were analyzed from the experiments that were used to determine headgroup order parameters in Ref.~\citenum{bacle21}.
+The analysis of the crowded spectral region at 29-31 ppm was based on the previous assignment reported for POPC membranes~\cite{ferreira13}.
 
+\begin{figure}
+  \includegraphics[width=\textwidth]{Figures/POPE_INEPT_contour_SCH.pdf}
+  \caption{Determination of the POPE acyl chain order parameters from a R-PDLF spectrum measured at a magic angle spinning frequency of 5.15 kHz.
+(A) $^{13}$C rINEPT spectrum with peak assignment.
+The labels used are shown in the chemical structure of POPE depicted.
+The chemical shift of the methyl groups was defined as 13.8 ppm.
+(B) Contour plot of the R-PDLF spectrum for the crowded spectral region.
+The assignment was based on a previous assignment reported for POPC membranes~\cite{ferreira13}.
+(C) C-H order parameter profile for the acyl chains of POPE.
+The splittings used for calculating the order parameters are shown in the next figure.
+The unassigned peaks belong to the headgroup and glycerol backbone carbons.
+A detailed assignment and order parameter analysis of these carbons was shown previously~\cite{bacle21}.
+}
+  \label{POPEexp1}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\textwidth]{Figures/POPE_slices.pdf}
+  \caption{Dipolar spectra obtained from the 2D R-PDLF spectrum described in the previous figure.
+The number at the top left of each plot denotes the corresponding chemical shift.
+The carbon label for each splitting is displayed on the top right.
+The labels are the same as shown in the previous figure.
+}
+  \label{POPEexp1}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\textwidth]{Figures/POPG_INEPT_contour_SCH.pdf}
+  \caption{Determination of the POPG acyl chain order parameters from a R-PDLF spectrum measured at a magic angle spinning frequency of 5.15 kHz.
+(A) $^{13}$C rINEPT spectrum with peak assignment.
+The labels used are shown in the chemical structure of POPE depicted.
+The chemical shift of the methyl groups was defined as 13.8 ppm.
+(B) Contour plot of the R-PDLF spectrum for the crowded spectral region.
+The assignment was based on a previous assignment reported for POPC membranes~\cite{ferreira13}.
+(C) C-H order parameter profile for the acyl chains of POPE.
+The splittings used for calculating the order parameters are shown in the next figure.
+The unassigned peaks belong to the headgroup and glycerol backbone carbons.
+A detailed assignment and order parameter analysis of these carbons was shown previously~\cite{bacle21}.}
+  \label{POPEexp1}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\textwidth]{Figures/POPG_slices.pdf}
+  \caption{Dipolar spectra obtained from the 2D R-PDLF spectrum described in the previous figure.
+The number at the top left of each plot denotes the corresponding chemical shift.
+The carbon label for each splitting is displayed on the top right.
+The labels are the same as shown in the previous figure.
+}
+  \label{POPEexp1}
+\end{figure}
+
+\pagebreak
 \subsection{X-ray scattering experiments}
@@
 
+\pagebreak
 \bibliography{refs.bib}

### 2023-01-03 main.tex

--- 2022-12-06 main.tex
@@
 \usepackage[T1]{fontenc}
+\usepackage{textcomp}
 \usepackage{graphicx}
@@
 
-\title{NMRlipids Databank: Making data-driven analyses of membrane properties accessible for all}
-%Overlay Databank of Lipid Membrane Simulations Arising from Open Collaboration}
+\title{NMRlipids Databank makes data-driven analysis of biomembrane properties accessible for all}
+%NMRlipids Databank: Making data-driven analyses of membrane properties accessible for all^
+%Overlay Databank of Lipid Membrane Simulations Arising from Open Collaboration
 
-\author[1]{Anne Kiirikki}
-\author[2]{...}
-\author[1,*]{O. H. Samuli Ollila}
+\author[1]{Anne Kiirikki}         %ORCID: ?
+\author[2]{Hanne Antila}          %ORCID: 0000-0002-2474-5053
+\author[2]{Lara Bort}             %ORCID: ?
+\author[2]{Pavel Buslaev}         %ORCID: ?
+%\author[2]{Matti Javanainen}      %ORCID: 0000-0003-4858-364X
+\author[2]{Batuhan Kav}           %ORCID: ?
+\author[2]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
+\author[2]{...}                   %ORCID: ?
+\author[1,*]{O. H. Samuli Ollila} %ORCID: 0000-0002- 8728-1006
 %\author[1,2,+]{Christine Author}
 %\author[2,+]{Derek Author}
-\affil[1]{University of Helsinki, Institute of Biotechonology, Helsinki, Finland}
+\affil[1]{University of Helsinki, Institute of Biotechnology, Helsinki, Finland}
 \affil[2]{Affiliation, department, city, postcode, country}
@@
 
-Cellular membranes are composed of hundreds of different types of lipid molecules which regulate membrane properties, morphology and biological functions~\cite{vanmeer08}.
-Membrane lipid composition can be also related to diseases, such as cancer and neurodegenerative disorders, and potential therapeutics that affect membrane compositions have been proposed~\cite{torres21}.
-However, membrane-containing systems are often difficult to study experimentally because the membranes are in a disordered fluid state under biological conditions and sample preparation can be arduous due to the complicated phase behaviour and rich interactions between lipids and other biomolecules.
-For those reasons, the detailed connections between lipid interactions and biological functions taking place in or around membranes are still poorly understood.
+Cellular membranes contain hundreds of different types of lipid molecules that regulate the membrane properties, morphology, and biological functions~\cite{vanmeer08,Lorent:2020a,Slatter:2016a}.
+Membrane lipid composition is implicated in diseases, such as cancer and neurodegenerative disorders, and therapeutics that affect membrane compositions are emerging~\cite{torres21}.
+However, biomembranes are often difficult to study experimentally because at biological conditions they are in disordered fluid state, and sample preparation can be highly complicated due to the rich phase behaviour and complex interactions between lipids and other biomolecules.
+Consequently, the connections between complex lipid interactions and biological functions remain poorly understood.
 %why such variety of lipids is needed and how their. 
-Molecular dynamics (MD) simulations have been particularly useful in understanding membrane systems, yet their accuracy is often compromised by the quality of model parameters, methodological inadequacy, and other artefacts~\cite{antila22b,gupta22}.
+Molecular dynamics (MD) simulations have been particularly useful in understanding membrane systems, although their accuracy has often been compromised by artefacts such as the quality of model parameters~\cite{antila22b,gupta22}.
 %While MD simulations of systems with few biomolecules continue having major contributions in our understanding on biology and drug design, 
-On the  other hand, the accuracy of models is becoming increasingly important as the scale (in both size and time) of the systems studied continuously increases 
-%researches are progressing from the simulations of 
+Presently, the accuracy of models is becoming increasingly important as researches are progressing from simulations of 
 %now searching new avenues by building models not only 
-from the level of individual molecules to models of whole cells or organelles using interdisciplinary approaches~\cite{johnson15,thornburg22,gupta22}.
-Such systems exhibit complex emergent behaviour making inaccuracies more difficult to detect and accumulation of even modest errors may have major effects the conclusions.
+individual molecules to simulating whole organelles or even cells using interdisciplinary approaches~\cite{johnson15,thornburg22,gupta22}. In such systems exhibiting exceedingly complex behaviour,
+model inaccuracies become increasingly difficult to detect and multiplication of small errors may have major impact on the conclusions.
 
-In contrast to experimental structural biology, where standard protocols to share and quality evaluate the resolved structures are established in the PDB databank~\cite{montelione13}, the equivalent best practices are yet to be defined for MD simulations.
-The importance of such approaches is widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} and solutions for sharing data are emerging for proteins in solution~\cite{meyer10,kamp10}, proteins in membranes~\cite{newport19,espigares20,leston22}, nucleic acids~\cite{hospital16}, cyclodextrins~\cite{mixcoha16}, and arbitrary biomolecules~\cite{bekker20}.
+In contrast to experimental structural biology, where standard protocols to share and quality-evaluate the resolved structures are firmly established \textit{via} the Protein Data Bank (PDB)~\cite{montelione13}, such best practices are yet to be defined for MD simulations.
+Importance of such approaches is widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} and data-sharing solutions are emerging for proteins in solution~\cite{meyer10,kamp10} and in membranes~\cite{newport19,espigares20,leston22}, as well as for nucleic acids~\cite{hospital16}, cyclodextrins~\cite{mixcoha16}, COVID-19--involved macromolecules~(\url{bioexcel-cv19.bsc.es}) and arbitrary biomolecules~\cite{bekker20}.
 However,
 %The relevance of quality evaluation of simulation trajectories in databanks regarding technical details of simulations and accuracy of the underlying physical description of the system (force field) has become evident~\cite{tai04,meyer10,hospital20} and such
-a streamlined workflow with automatic quality evaluation~\cite{meyer10,hospital16} and programmatic access is still rare. 
+automatic quality evaluation of simulation data~\cite{meyer10,hospital16} and programmatic access are still rare. 
 %However, straightforward quality comparisons between individual simulations or force fields within these databanks remain challenging. 
-Therefore, tools for automatic quality evaluation of membrane simulations or training sets for machine learning models of membrane-containing systems are not yet available.
+In particular, tools for automatic quality evaluation of membrane simulations, or training sets for machine learning models of membrane-containing systems, are not yet available.
+The profound, and sometimes unexpected, effect that such open data sharing and quality evaluation can have on the field cannot be understated.
+At the very least, open access data will reduce the man and computational hours originating from double work, reducing the environmental impact.
+Beyond that, by providing the training data, the PDB laid the foundation for the machine learning algorithms now closing into solving the protein folding problem and the strides those machine learning tools have facilitated in the biomedical sciences since.
 
@@
 
-Here we present the NMRlipids databank, a community-driven open access databank with programmatic access containing atomistic resolution MD simulations of lipid bilayers.
-The programmatic access enables various data-driven approaches, providing new tools for researchers in a wide range of fields covering academia and industry from cell membrane biology to lipid nanoparticle formulations, data-driven computational chemistry, and machine learning.
-Here, we demonstrate how data-driven analysis of water anisotropic diffusion in membrane systems from the NMRlipids databank can extend the scope of MD simulations to MRI imaging and pharmacokinetics as well as how the NMRlipids databank can be utilized to analyse rare phenomena that are beyond the scope of standard MD simulation investigations.
-Furthermore, we perform automatic quality evaluation of membrane simulations, which guides the selection of best models for specific applications and the improvement of simulation parameters and methodology.
+Here we present the NMRlipids Databank---a community-driven, open-for-all database featuring programmatic access to atom-resolution MD simulations of lipid bilayers.
+The Databank performs automatic quality evaluation of membrane simulations, which facilitates the selection of best models for specific applications and accelerates the development of simulation parameters and methodology.
+The programmatic access enables users to apply data-driven approaches on the Databank, thus facilitating the creation of new tools for (and by) researchers in a wide range of fields, both in academia and industry, from cell membrane biology and lipid nanoparticle formulations to computational chemistry and machine learning.
+As two usage examples of what is already possible, we demonstrate here (i) how data-driven analysis of water anisotropic diffusion in all the membrane systems available in the Databank can extend the scope of MD simulations to magnetic resonance imaging (MRI) and pharmacokinetics, and (ii) how the Databank allows its users to analyse rare phenomena that are beyond the scope of standard MD simulation studies.
+Wider adaptation of the NMRlipids Databank will open even more possibities.
 %The practical relevance of the NMRlipids databank is exemplified by automatic quality evaluation and ranking of large amount of MD simulation data,
 
-While the NMRlipids databank currently contains only bilayer systems, it can readily be expanded to contain other molecules than lipids, such as disordered proteins or sugars. 
+While the NMRlipids Databank currently contains only lipid bilayer systems, it can be readily expanded to contain other molecules, such as disordered proteins or sugars. 
 %a solution for lipid bilayers based on overlay databank structure illustrated in Fig.~\ref{fig:overlay}.  
-The combination of an overlay databank structure and a community-wide open collaboration used here has been applied successfully when building databanks in other fields,  
+The powerful combination of the overlay databank structure and open collaboration can be applied also when building databases in other fields,  
 %is developed here to solve the practical challenges in generating databanks of MD simulation data enabling flexible analyses over large sets of simulation data, but it can be  potentially used for wide range of situations.
@@
 %overlay databank approach lowers the barrier to start without compromizing the long term stability or scalability.
-Thus, the approach has already proven itself most useful in enabling data-driven and machine learning applications in new fields.
+Therefore, the approach introduced here opens up possibilities to create databases enabling data-driven and machine learning applications in new fields.
 
@@
 \subsection{Programmatic access to MD simulation data of membranes composed of most common biologically abundant lipids}
-NMRlipids databank is a community-driven databank containing atomistic MD simulations of biologically relevant lipid membranes emerging from the NMRlipids open collaboration~\cite{botan15,ollila16,catte16,antila19,bacle21}.
-Nearly a thousand simulation trajectories with the total length approaching half a millisecond can be accessed programmatically using the Python application programming interface (API) or a graphical user interface (GUI) (\url{http://www.databank.nmrlipids.fi/}).
+
+NMRlipids Databank is a community driven database containing atomistic MD simulations of biologically relevant lipid membranes emerging from the NMRlipids open collaboration~\cite{botan15,ollila16,catte16,antila19,bacle21}.
+Nearly thousand simulation trajectories with the total length approaching half a millisecond can be accessed programmatically using Python API or graphical user interface (\url{http://www.databank.nmrlipids.fi/}).
 %NMRlipids databank is the only databank of MD simulations with open programmatic access to its content enabling fully flexible analyses of its content. 
-Following the FAIR principles~\cite{wilkinson16} and NMRlipids project protocol, not only the content and computer programs related to the databank, but also the entire construction process of the databank, are open access~\cite{botan15}.
+Following the FAIR principles~\cite{wilkinson16} and NMRlipids project protocol, not only the content and computer programs related to the databank but also the whole construction process of the databank are open access~\cite{botan15}.
 %The NMRlipids databank will be always open for data and other contributions.
@@
 Currently available single component lipid membranes and binary mixtures in the NMRlipids databank are illustrated in Fig.~\ref{fig:overlay}D.
-While these comprise the majority of available trajectories, mixtures with up to five lipid types are available.
+While these compose the majority of available trajectories, also mixtures up to five lipids are available.
 The distribution of available simulations containing a specific lipid is shown in Fig.~\ref{fig:overlay}C.
-This roughly resembles the biological abundance of different lipid types with phosphatidylcholine (PC) being the most common followed by cholesterol, phosphatidylethanolamine (PE), phosphatidylserine (PS), phosphatidylglycerol (PG), phosphatidylinositol (PI), and other lipids depending on organism and organelle~\cite{vanmeer08}.
-The force fields used in the simulations cover all the essential parameters commonly used in lipid simulations, ranging from highly popular CHARMM36 parameters to united atom and polarizable force fields.
+This roughly resembles the biological abundancies of different lipid types, PC being the most common followed by cholesterol, PE, PS, PG, PI, and other lipids depending on organism and organelle~\cite{vanmeer08}.
+The force fields used in the simulations cover all the essential parameters used in lipid simulations, ranging from highly popular CHARMM36 parameters to united atom and polarizable force fields.
 Therefore, the averages calculated over the databank can be considered as mean predictions from available lipid models for an average cell membrane.
@@
 The efficient upcycling of large MD simulation trajectories in the NMRlipids databank is enabled by the overlay structure illustrated in Fig.~\ref{fig:overlay}A.
-Raw simulation data in the {\it data layer} can be stored in any publicly available location with long term stability and permanent links to the data (e.g., digital object identifiers (DOIs)).
-The {\it Databank layer} is the core of the databank containing all the relevant information about the simulations, including links to the raw data, relevant metadata describing the systems, universal naming conventions for the lipids and their atoms, computer programs to create and analyse the entries, basic properties calculated from all simulations (area per lipid, C-H bond order parameters, X-ray scattering form factors and membrane thickness), and quality evaluation of simulations against experimental data.
-In practise, this information is stored is the git repository which is currently available at~\url{https://github.com/NMRLipids/Databank}.
-The {\it application layer} contains outputs from the databank, such as graphical user interface (\url{http://www.databank.nmrlipids.fi/}), and results from analyses described in sections below.
-A more detailed description on the NMRlipids databank structure is available the supplementary information.
+Raw simulation data in the {\it data layer} can be stored in any publicly available location with long term stability and permanent links to the data, such as digital object identifiers (DOIs) or uniform resource names (URNs).
+The {\it Databank layer} is the core of the databank containing all the relevant information about the simulations, including links to the raw data, relevant metadata describing the systems, universal naming conventions for lipids and their atoms, computer programs to create and analyse the entries, basic properties calculated from all simulations (area per lipid, C--H bond order parameters, X-ray scattering form factors and membrane thickness), and quality evaluation of simulations against experimental data.
+In practise, this information is stored in the git repository which is currently available at~\url{https://github.com/NMRLipids/Databank}.
+The {\it application layer} contains outputs from the databank, such as graphical user interface (\url{http://www.databank.nmrlipids.fi/}) and results from analyses described in sections below.
+More detailed description on the NMRlipids databank structure is available the supplementary information.
 
@@
     \centering
-    \includegraphics[width = 180mm]{Figures/overlay2.pdf}
+    \includegraphics[width=\linewidth]{Figures/overlay2.pdf}
     \caption{a) Structure of an overlay databank. 
@@
 Examples of such applications are the GitHub repository generating the results presented in this work and the graphical access to the databank.
-    More detailed structure of the layer 2 in the NMRlipids databank is illustrated in Fig.~\ref{DatabankStructure} in the SI.
-    b) Distribution of the lengths of the trajectories, total number of trajectories and total lenght of the simulations in the NMRlipids databank.
-    c) Distribution of lipids present in the trajectories in the NMRlipids databank. Lipids occuring in five or less simulations ('others') are listed in the right. 
+    A more detailed structure of the databank layer in the NMRlipids databank is illustrated in Fig.~\ref{DatabankStructure} in the SI.
+    b) Distribution of the lengths of the trajectories, total number of trajectories and total length of the simulations in the NMRlipids databank.
+    c) Distribution of lipids present in the trajectories in the NMRlipids databank. Lipids occuring in five or fewer simulations ('others') are listed on the right. 
     d) Currently available binary mixtures in the NMRlipids databank. 
@@
 %Quality of membrane simulations with different force fields have been evaluated against experimental data during parameterization and in separate comparison studies~\cite{botan15,ollila16,catte16,pluhackova16,perez17,leonard19,??}, but the lack of universal quality measure for membrane simulations complicates the estimations of reliability of simulations.
-Understanding the accuracy of used force field parameters is crucial when assessing the reliability and significance MD simulation results.
-However, the lack of universal quality measures for MD simulations and the complicated landscape of force field quality for membranes hamper the applicability such simulations, resulting in substantial controversies~\cite{antila22b}.
+Understanding the accuracy of the used force field parameters is crucial when estimating the reliability and significance of MD simulation results.
+However, the lack of universal quality measures for MD simulations and complicated landscape of force field quality for membranes hamper the quality of simulations which has resulted to many controversial results~\cite{antila22b}.
 For example, previous quality evaluations of membrane simulations have concluded that CHARMM36 parameters give the best description for lipid headgroup conformational ensembles~\cite{bacle21}, GROMOS-CKP parameters best capture the membrane packing in POPS bilayers~\cite{antila22b}, and that OPLS3e parameters overcome CHARMM36 in structural quality for POPC but predicts overestimated ion binding~\cite{kurki22}.
 %, thereby being a major obstacle in many applications of membrane MD simulations. 
-To enable the navigation in this complex force field milieu, we here define quantitative quality measures that can be used to rapidly find the best available parameters for a specific application or to guide the development of improved force fields.
+To enable the navigation in this complex force field quality landscape, we define here quantitative quality measures that can be used to rapidly find the best available parameters for a specific application or to guide force field development.
 For detailed definitions of quality measures, see the supplementary information.
@@
     \centering
-    \includegraphics[width=180mm]{Figures/quality3.pdf}
+    \includegraphics[width=\linewidth]{Figures/quality3.pdf}
     \caption{ A) Best simulations ranked based on overall order parameter quality.
     B) Best simulations ranked based on the overall order parameter quality of POPE lipid. 
-    C)-E) Evaluation against experimental data exemplified for a simulation with the best overall order parameter quality (C), the best quality for POPE lipid (D), and the headgroup quality for POPE (E).
-    F) Scatter plots and Pearson correlation coefficients for the membrane area per lipid, thickness, first minima of x-ray scattering form factor and average order parameter of the {\it sn}-1 acyl chain extracted from the NMRlipids databank.
+    C)--E) Evaluation against experimental (X-ray scattering  NMR order parameter) data exemplified for a simulation with the best overall order parameter quality (C), the best quality for POPE lipid (D), and the headgroup quality for POPE (E).
+    F) Scatter plots and Pearson correlation coefficients for the membrane area per lipid, thickness, first minima of x-ray scattering form factor and average order parameter of the {\textit sn}-1 acyl chain extracted from the NMRlipids databank.
     All correlation coefficients have p-value below 0.001. For more correlations see Fig.~\ref{fig:QualityCorrelationsSI}.
@@
 
-Ranking of all simulations with the highest scores based on evaluation against C-H bond order parameters from NMR experiments is shown in Fig.~\ref{fig:quality} A and ranking of best POPE lipid bilayer simulations in Fig.~\ref{fig:quality} B.
-Direct comparisons with experiments for selected simulations are shown in Figs.~\ref{fig:quality} C-D.
-C-H bond order parameters are sensitive to conformational ensembles of individual lipid molecules~\cite{ollila16} while also correlating well with the membrane thickness and lateral packing as shown in Fig.~\ref{fig:quality} F.
-This is complemented by comparing the absolute values of X-ray scattering form factors between simulations and experiments that are related to the overall membrane dimensions via the electron density profile.
-In particular, the location of minima in these form factors are correlated with the membrane dimensions as shown in Figs.~\ref{fig:quality} F,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}.
+Ranking of all simulations with the highest scores based on evaluation against C--H bond order parameters from NMR experiments is shown in Fig.~\ref{fig:quality} A and ranking of best POPE lipid bilayer simulations in Fig.~\ref{fig:quality} B.
+Direct comparisons with experiments for selected simulations are shown in Figs.~\ref{fig:quality} C--D.
+C--H bond order parameters are sensitive to conformational ensembles of individual lipid molecules~\cite{ollila16}, but they correlate well also with the membrane thickness and lateral packing as shown in Fig.~\ref{fig:quality} F.
+This is complemented by comparing the absolute values of X-ray scattering form factors between simulations and experiments that are related to the overall membrane dimensions \textit{via} electron density profile.
+Particularly, the location of minima in these form factors are correlated with the membrane dimensions as shown in Figs.~\ref{fig:quality} F,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}.
 
@@
 %\subsection{Finding the best models for PC and PE lipid mixtures}
-The power of the NMRlipids databank in selecting the best simulation parameters for a specific application is demonstrated for mixtures of PE and PC lipids in Fig.~\ref{fig:POPCPOPEapls}, showing areas per lipid from available POPC and POPE bilayer simulations in the databank from different force fields at 310\,K.
+The power of the NMRlipids databank in selecting the best simulation parameters for a specific application is demonstrated for mixtures of PE and PC lipids in Fig.~\ref{fig:POPCPOPEapls} showing area per lipids from available POPC and POPE bilayer simulations in the databank from different force fields at 310\,K.
 Predictions from different force fields deviate significantly in terms of absolute values and the slope of decrease upon addition of POPE.
-Because the area per lipid correlates with the average order parameter of the $\it{sn}$-1 chain (Fig.~\ref{fig:quality} F), simulations with the best predictions for area per lipids can be selected based on order parameter quality evaluation.
-Quality evaluation in Fig.~\ref{fig:quality} and direct comparison in Fig.~\ref{fig:POPC_POPE_dataSI} reveal that Slipids ranks 1st for POPC with a clear advantage to the others, and 2nd for POPE with only marginally lower quality score than GROMOS-CKP and is therefore the best avilable choice for studies where packing effects of PE lipids are important.
+Because the area per lipid correlates with the average order parameter of $\textit{sn}$-1 chain (Fig.~\ref{fig:quality} F), simulations with the best predictions for area per lipids can be selected based on order parameter quality evaluation.
+Quality evaluation in Fig.~\ref{fig:quality} and direct comparison in Fig.~\ref{fig:POPC_POPE_dataSI} reveal that Slipids ranks 1\textsuperscript{st} for POPC with a clear difference to others, and 2\textsuperscript{nd} for POPE with only marginally lower quality than GROMOS-CKP, therefore being the best selection for studies where packing effects of PE lipids are relevant.
 %The direct comparison with experiments in Fig.~\ref{fig:POPC_POPE_dataSI} in the supplementary information reveals that 
-Simulations with CHARMM36 and GROMOS-CKP predict too packed bilayer for POPC (overestimated order in Fig.~\ref{fig:POPC_POPE_dataSI}), while area per lipid for POPC is overstimated in Lipid17.
+Simulations with CHARMM36 and GROMOS-CKP predict too packed bilayer for POPC (overestimated order in Fig.~\ref{fig:POPC_POPE_dataSI}), while area per lipid for POPC is overestimated in Lipid17.
 For POPE,
@@
 
-In conclusion, 
+In summary, 
 %the quality evaluation based on the NMRlipids databank suggests that the Slipids parameters would give the most reliable description for membrane packing of POPC:POPE mixtures.
@@
 Altogether, these examples demonstrate how
-the automatic quality evaluation in the NMRlipids databank enables rapid selection of the best simulation parameters to specific applications without extensive and tedious manual force field evaluation.
-This possibility will promote more reliable simulation results for wide range of applications, support force field development and parametrization of coarse grained force fields against the best atomistic MD simulations.
+the automatic quality evaluation of the NMRlipids databank enables the rapid selection of the best simulation parameters to specific applications without extensive and tedious manual force field evaluation.
+This possibility will promote more reliable simulation results for wide range of applications, support force field development, and streamline the parametrization of coarse grained force fields against the best atomistic MD simulations.
 The automatically extracted quantitative quality measures will be particularly useful to guide automatic parametrization procedures.
@@
 
-\subsection{Detection rare phenomena using NMRlipids databank: Cholesterol flip-flops}
-Lipid flip-flops from one bilayer leaflet to another play an important role in lipid trafficking and regulating membrane properties~\cite{vanmeer08}.
-Phospholipid flip-flops are slow, occuring with the timescales of hours or days, while cholesterol, diacylglycerol and ceramide flip-flops are faster, yet the reported timescales range between minutes to sub-millisecods~\cite{vanmeer08,steck12,parisio16,gu19}.
-These timescales were previously accessible only by coarse-grained simulations or free energy calculations~\cite{parisio16}, yet atomistic simulations reporting cholesterol flip-flop events have been published recently~\cite{gu19,javanainen19,baral20}.
+\subsection{Detecting rare phenomena using NMRlipids databank: Cholesterol flip--flops}
+Lipid flip--flops from one bilayer leaflet to another play an important role in lipid trafficking and regulating membrane properties~\cite{vanmeer08}.
+Phospholipid flip--flops are slow, occurring with the timescales of hours or days, while cholesterol, diacylglycerol and ceramide flip-flops are faster, yet the reported timescales range from minutes to sub-millisecods~\cite{vanmeer08,steck12,parisio16,gu19}.
+These timescales were previously accessible only by coarse-grained simulations or free energy calculations~\cite{parisio16}, yet atomistic simulations reporting cholesterol flip--flop events have been published recently~\cite{gu19,javanainen19,baral20}.
 %However, systematic studies on how cholesterol flip-flop rate depends on membrane properties are still beyond the current computational resources available for a single research group. 
-These studies report increase in cholesterol flip-flop rates with increasing unsaturation level and decreasing cholesterol concentration~\cite{gu19,javanainen19}, but correlations between cholesterol flip-flop rates and membrane properties have not been systematically studied.
-Here we demonstrate that the large amount of simulation data available in the NMRlipids databank with programmatic access makes the analysis of these correlations accessible for all.
+These studies report increase in cholesterol flip--flop rates with increasing acyl chain unsaturation level and decreasing cholesterol concentration~\cite{gu19,javanainen19}, but correlations between cholesterol flip--flop rates and membrane properties have not been systematically studied.
+Here we demonstrate that the large amount of simulation data available in the NMRlipids databank, along with programmatic access providing the option to flexibly implement varied queries of the data, enables anybody to perform such analyses.
 
-Flip-flops were observed for cholesterol, % (634 events in 77 different simulations), 
+Flip--flops were observed for cholesterol, % (634 events in 77 different simulations), 
 DCHOL (18,19-di-nor-cholesterol), %, 16 flip-flops in 1 simulation) 
-DOG (1,2-dioleoyl-sn-glycerol), %, 44 flip-flops in 4 simulations) 
-and SDG (1-stearoyl-2-docosahexaenoyl-sn-glycerol) in the NMRlipids databank.
-Cholesterol flip-flop rates range between 0.001-1.6\,$\mu$s$^{-1}$ with the mean of 0.16\,$\mu$s$^{-1}$ and median of 0.07\,$\mu$s$^{-1}$.
+DOG (1,2-dioleoyl-\textit{sn}-glycerol), %, 44 flip-flops in 4 simulations) 
+and SDG (1-stearoyl-2-docosahexaenoyl-\textit{sn}-glycerol) in the NMRlipids databank.
+Cholesterol flip--flop rates range between 0.001--1.6\,\textmu{}s$^{-1}$ with the mean of 0.16\,\textmu{}s$^{-1}$ and median of 0.07\,\textmu{}s$^{-1}$.
 These values are in line with the previously reported values from atomistic MD simulations~\cite{gu19,javanainen19,baral20}.
-Flip-flop rates of DCHOL, 0.2\,$\mu$s$^{-1}$, was close to the average value of cholesterol, while average rates for diacylglycerols DOG and SDG were higher than for cholesterol, 0.4\,$\mu$s$^{-1}$ and 0.5\,$\mu$s$^{-1}$, respectively.
-Flip-flops were not observed for other lipids, giving the upper limits for PC lipid flip-flop rate of 9\,s$^{-1}$ and for ceramide (N-palmitoyl-D-erythro-sphingosine) of 0.002\,$\mu$s$^{-1}$.
-Therefore, data in the NMRlipids databank suggests that the lipid flip-flip rate decreases in the order of diacylglycerols $>$ cholesterol $>$ other lipids including ceramides.
-However, amount of data for diacylgycerols (8 simulations with Lipid17 force field) and ceramide (3 simulations with CHARMM force field) is less than for cholesterol (83 simulations), thus we cannot fully exclude the effect of force field or composition on this comparison.
+The flip--flop rate of DCHOL, 0.2\,\textmu{}s$^{-1}$, was close to the average value of cholesterol, while average rates for diacylglycerols DOG and SDG were higher than for cholesterol, 0.4\,\textmu{}s$^{-1}$ and 0.5\,\textmu{}s$^{-1}$, respectively.
+Flip--flops were not observed for other lipids, giving the upper limits for PC lipid flip--flop rate of 9\,s$^{-1}$ and for ceramide (\textit{N}-palmitoyl-\textsc{D}-\textit{erythro}-sphingosine) of 0.002\,\textmu{}s$^{-1}$.
+Therefore, data in the NMRlipids databank suggests that the lipid flip--flip rate decreases in the order of diacylglycerols $>$ cholesterol $>$ other lipids including ceramides.
+However, the amount of data for diacylgycerols (8 simulations with Lipid17 force field) and ceramide (3 simulations with CHARMM36 force field) is less than for cholesterol (83 simulations), thus we cannot fully exclude the effect of force field or composition on this comparison.
 
-Nevertheless, we use the wide range of available simulations to analyse how cholesterol flip-flop rate depends on membrane properties.
-Cholesterol flip-flop rates and their averages over fixed ranges of x-axis values are plotted as a function of membrane thickness, lateral density and order in Figs.~\ref{fig:flip-flops} B-D.
-The results reveal a non-linear correlation between decreasing cholesterol flip-flop rate and membrane packing.
-Flip-flop rates increase with an order of magnitude when membrane packing density decreases and a major jump is observed at low membrane packing.
-The order of magnitude changes in cholesterol flip-flop rate with the membrane composition may have major implications in understanding lipid trafficking and membrane biochemistry~\cite{gu19,baral20}.
-Because the results from the NMRlipids databank are averaged over large range of membrane compositions and force fields, they show that the strong dependence of cholesterol flip-flop rate on membrane properties is not limited to certain lipid compositions or force fields used in previous studies~\cite{gu19,javanainen19,baral20}.
+Nevertheless, we use the wide range of available simulations to analyse how cholesterol flip--flop rate depends on membrane properties.
+Cholesterol flip--flop rates and their histograms are plotted as a function of membrane thickness, lateral density and order in Figs.~\ref{fig:flip-flops} B--D.
+The results reveal a non-linear correlation between cholesterol flip--flop rate and membrane packing.
+Flip--flop rates increase by an order of magnitude when membrane packing density decreases and a major jump is observed at low membrane packing.
+The order of magnitude changes in cholesterol flip.-flop rate with the membrane composition may have major implications in understanding lipid trafficking and membrane biochemistry~\cite{gu19,baral20}.
+Because the results from the NMRlipids databank are averaged over large range of membrane compositions and force fields, they show that the strong dependence of cholesterol flip--flop rate on membrane properties is not limited to certain lipid compositions or force fields used in previous studies~\cite{gu19,javanainen19,baral20}.
 
@@
     \includegraphics[width=88mm]{Figures/CholFlipFlops.pdf}
-    \caption{{\bf A} Illustration of cholesterol flip-flop.  
-      {\bf B-D} Cholesterol flip-flops analyzed from the databank as a function of membrane thickness, area per lipid, and acyl chain order.
-Values from simulations with non-zero flip-flop rates are shown with blue dots.
-Averages over fixed range of x-axis values are shown with black dots.
+    \caption{\textbf{A} Illustration of cholesterol flip-flop.  
+      \textbf{B--D} Cholesterol flip--flops analyzed from the databank as a function of membrane thickness, area per lipid, and acyl chain order.
+Values from simulations with non-zero flip--flop rates are shown with blue dots.
+Histogrammed values are shown with black dots.
     }
@@
 
-Cholesterol flip-flops are phenomena whose analyses from MD simulations has only recently became possible for groups with substantial resources and special expertise available~\cite{gu19,javanainen19,baral20}.
-Our analysis demonstrates how the NMRlipids databank makes analyses of such rare phenomena accessible to wide range of scientists in different fields of science and industry who do not have the access to required resources and expertise to produce the large amounts of MD simulation data.
+Cholesterol flip--flops are phenomena whose analyses from MD simulations has only recently became possible for research groups with substantial resources and special expertise~\cite{gu19,javanainen19,baral20}.
+Our analysis demonstrates how the NMRlipids databank makes analyses of such rare events accessible to wide range of scientists in different fields of science and industry who do not have the access to required resources and expertise to produce the large amounts of MD simulation data.
 
@@
 The anisotropy in water diffusion in parallel and perpendicular directions with respect to membranes plays a role in the drug translocation through biological material, particularly in skin \cite{hansen13,wen18,nitsche19,roberts21}, and in MRI imaging~\cite{topgaard20}.
-MD simulations are rarely used to analyze anistropic diffusion of water as only few permeation events for water are typically observed in a single MD simulation trajectory~\cite{venable19,camilo2022}, thereby making the collection of sufficient amount data challenging.
-Here we show that the NMRlipids databank can be used to analyze how anisotropic water diffusion in a multilamellar membrane systems depends on membrane properties.
+MD simulations are rarely used to analyze anistropic diffusion of water as only few membrane permeation events for water are typically observed in a single MD simulation trajectory~\cite{venable19,camilo2022}, thereby making the collection of sufficient amount of data challenging.
+Here we show that the NMRlipids databank can be used to analyze how the anisotropic water diffusion in multilamellar membrane systems depends on membrane properties.
 
@@
     \centering
-    \includegraphics[width=180mm]{Figures/permeation2.pdf}
-    \caption{{\bf A} Water diffusion, $D_\perp$, and permeability, $P$, through membranes, and lateral diffusion along the membrane, $D_{||}$, illustrated in a multilamellar stack of lipid bilayers. 
-    {\bf B-E} Water permeation through membranes analyzed from the databank as a function of temperature, thickness, area per lipid, and acyl chain order.
+    \includegraphics[width=\linewidth]{Figures/permeation2.pdf}
+    \caption{\textbf{A} Water diffusion, $D_\mathrm{perp}$, and permeability, $P$, through membranes, and lateral diffusion along the membrane, $D_{||}$, illustrated in a multilamellar stack of lipid bilayers.
+    \textbf{B--E} Water permeation through membranes analyzed from the databank as a function of temperature, thickness, area per lipid, and acyl chain order.
 Values from simulations with non-zero permeation values are shown with blue dots.
-Averages over fixed range of x-axis values are shown with black dots.
-Insert in B) shows the Arrhenius plot of permeation ($\ln$ (P) vs.
-1/T) that gives 17\,$\pm$\,4\,k$_b$T for the average activation energy for water permeation through lipid bilayer.
-Inserts in C) and D) show the region where the dependence could be considered approximately linear.
-    {\bf F} Lateral diffusion of water as a function of hydration level. Experimental points for DMPC bilayers at 313\,K at different hydration levels are shown~\cite{rudakova04}.
-    {\bf G-H} Diffusion anistoropy of water as a function of thickness and area per lipid. }
+Histogrammed values are shown with black dots.
+Inset in B) shows the Arrhenius plot of permeation ($\ln(P)$ vs.
+$1/T$) that gives 17\,$\pm$\,4\,$k_\mathrm{B}T$ for the average activation energy for water permeation through lipid bilayer.
+Insets in C) and D) show the region where the dependence could be considered approximately linear.
+    \textbf{F} Lateral diffusion of water as a function of hydration level. Experimental points for DMPC bilayers at 313\,K at different hydration levels are shown~\cite{rudakova04}.
+    \textbf{G--H} Diffusion anistoropy of water as a function of thickness and area per lipid. }
     \label{fig:permeability}
@@
 To this end, we first calculated the water permeability through membranes from all simulations in the NMRlipids databank.
-The resulting non-zero values range between 0.3 and 322\,$\mu$m/s with the mean and median of 14\,$\mu$m/s and 8\,$\mu$m/s, respectively.
-These values agree with the previously reported simulation results~\cite{venable19,camilo2022} and have the same order of magnitude as experimentally determined diffusive permeability coefficients, but are on average larger than the values reported for PC lipids in liquid crystalline phase,~0.19-0.33\,$\mu$m/s~\cite{jansen95}.
-Observed permeabilities and their averages over fixed ranges of values at x-axis are shown in Figs.~\ref{fig:permeability} B-E as a function of temperature, membrane thickness, area per lipid, and acyl chain order.
+The resulting non-zero values range between 0.3 and 322\,\textmu{}m/s with the mean and median of 14\,\textmu{}m/s and 8\,\textmu{}m/s, respectively.
+These values agree with the previously reported simulation results~\cite{venable19,camilo2022} and have the same order of magnitude as experimentally determined diffusive permeability coefficients, but are on average larger than the values reported for PC lipids in liquid crystalline phase,~0.19--0.33\,\textmu{}m/s~\cite{jansen95}.
+Observed permeabilities and their histogrammed values are shown in Figs.~\ref{fig:permeability} B--E as a function of temperature, membrane thickness, area per lipid, and acyl chain order.
 %on average, we calculated the average permeabilities over all systems with the value of temperature, thickness, area per lipid, or acyl chain order within a fixed range.
 These averaged permeabilities also shown in Figs.~\ref{fig:permeability} with black dots.
-As expected, the permeability increases with the temperature, giving 17\,$\pm$\,4\,k$_b$T for the average energy barrier for the water permeation from the Arrhenius plot in Fig.~\ref{fig:permeability} B.
-On the other hand, the permeability of water decreases on average when membranes become more packed, i.e., with decreasing area per lipid and increasing thickness and acyl chain order (Figs.~\ref{fig:permeability} C-E).
+As expected, the permeability increases with the temperature, giving 17\,$\pm$\,4\,$k_\mathrm{B}T$ for the average energy barrier for the water permeation from the Arrhenius plot in Fig.~\ref{fig:permeability} B.
+On the other hand, the permeability of water decreases on average when membranes become more packed, \textit{i.e.}, with decreasing area per lipid and increasing thickness and acyl chain order (Figs.~\ref{fig:permeability} C--E).
 Permeation of water through bilayers depends on membrane properties also according previous studies, but it is not agreed whether the area per lipid~\cite{nagle08} or bilayer thickness~\cite{frallicciardi22} would be the main parameter determining the permeability.
 %is previously reported from simulations and experiments for some systems, but not for all~\cite{mathai07,venable19,frallicciardi22}. 
-Our analysis over the NMRlipids databank, containing significantly more data than that was available in previous studies, suggest non-linear dependence on both of these parameters, yet the linear correlation would be a good approximation for thicknesses above $\sim$\,3.9\,nm and area per lipids below $\sim$69\,Å$^2$ (insets in Figs.~\ref{fig:permeability} C-D).
+Our analysis over the NMRlipids databank, containing significantly more data than what was available in previous studies, suggest non-linear dependencies on both of these parameters, yet the linear correlation would be a good approximation for thicknesses above $\sim$\,3.9\,nm and area per lipids below $\sim$69\,\AA{}$^2$ (insets in Figs.~\ref{fig:permeability} C--D).
 %within the range of experimentally reported values  permeability which is 1-2 orders of magnitude slower than osmotic permeability \cite{jansen95,??}. 
@@
 This is not surprising as the most common water model used in membrane simulations, TIP3P, overestimates the bulk water diffusion~\cite{pathirannahalage21}.
-To estimate the diffusion anisotropy of water, $D_{\perp}/D_{||}$, in multi-lamellar membrane system, the permeability coeffients of water through membranes were translated to perpendicular diffusion coefficients, $D_{\perp}$, using the Tanner equation~\cite{tanner78,wasterby02}.
+To estimate the diffusion anisotropy of water, $D_\mathrm{\perp}/D_{||}$, in multilamellar membrane system, the permeability coefficients of water through membranes were translated to perpendicular diffusion coefficients, $D_\mathrm{\perp}$, using the Tanner equation~\cite{tanner78,wasterby02}.
 The resulting perpendicular diffusion coefficients are approximately five orders of magnitude smaller than lateral diffusion coefficients of water (Figs.~\ref{fig:permeability} G and H), which is at the upper limit of the anisotropic estimated from the experimental data~\cite{nitsche19}.
 %Large anisotropy values are understandable as simulations give slightly slower permeation rates and higher lateral diffusion rates than experiments. 
-Significant increase in the diffusion anisotropy with membrane packing is observed, as $D_{\perp}/D_{||}$ drifts away from one with decreasing area per lipid and increasing thickness in Figs.~\ref{fig:permeability} G and H.
-This follows from decreasing water permeability with membrane packing (Figs \ref{fig:permeability} C and D), while lateral diffusion remains approximately constant (Fig.
-\ref{fig:diffusionSI} A and C).
+Significant increase in the diffusion anisotropy with membrane packing is observed, as $D_\mathrm{\perp}/D_{||}$ drifts away from one with decreasing area per lipid and increasing thickness in Figs.~\ref{fig:permeability} G and H.
+This follows from decreasing water permeability with membrane packing (Figs.~\ref{fig:permeability} C and D), while lateral diffusion remains approximately constant (Figs.~\ref{fig:diffusionSI} A and C).
 
-In conclusion, our results suggest that the bilayer packing has a substantial effect on anisotropic water diffusion in multi-membrane lipid systems.
-Several folds larger anisotropy in membranes with higher lateral density is expected to play a role in pharmagokinetic models not only for water but also for other hydrophilic molecules~\cite{nitsche19}.
-Furthermore, the understanding of this anisotropy may help in developing new MRI imaging methods~\cite{topgaard20}.
+In summary, our results suggest that the bilayer packing has a substantial effect on anisotropic water diffusion in multi-membrane lipid systems.
+Several folds larger anisotropy in membranes with higher lateral density is expected to play a role in pharmacokinetic models not only for water but also for other hydrophilic molecules~\cite{nitsche19}.
+Furthermore, the understanding of this anisotropy may help in developing new (functional) MRI imaging methods~\cite{topgaard20}.
 
@@
 For example, a machine learning model predicting electron density profiles from form factors would be highly useful in interpretation of scattering experiments.
-On the other hand, more elaborated models can be trained to predict arbitrary membrane properties using the data from the NMRlipids databank.
+Furthermore, the data contained in the NMRlipids databank will enable training of more elaborated models to predict arbitrary membrane properties.
 
@@
 The selection of best models for a specific application is demonstrated for PC/PE lipid mixtures in Fig.~\ref{fig:quality}, and previously for PC/PS lipid mixtures~\cite{antila22b}, and lipid headgroups~\cite{bacle21}.
-On the other hand, programmatic access in the NMRlipids databank enables automatic analyses over larger sets of simulation data in terms of quantity (e.g., simulation length and number of conformations) and content (e.g, lipid compositions and ion concentrations) that is currently possible in a single research group.
-Added value of such access to simulation data is demonstrated here for the analysis of simulation qualities (Fig.~\ref{fig:quality}), water permeation through membranes (Fig.~\ref{fig:permeability}) and cholesterol flip-flop events (Fig.~\ref{fig:flip-flops}).
+On the other hand, programmatic access in the NMRlipids databank enables automatic analyses over larger sets of simulation data in terms of quantity (\textit{e.g.}, simulation length and number of conformations) and content (\textit{e.g.}, lipid compositions and ion concentrations) that is currently possible in a single research group.
+Added value of such access to simulation data is demonstrated here for the analysis of simulation qualities (Fig.~\ref{fig:quality}), water permeation through membranes (Fig.~\ref{fig:permeability}) and cholesterol flip--flop events (Fig.~\ref{fig:flip-flops}).
 %Such models could be used to connect membrane composition to relevant membrane properties in wide range of fields from molecular biology to drug development and biomolecular imaging. 
@@
 material science and biomolecular imaging. 
-
+cide
 \begin{table}[t]
@@
     \hline
-    Analyses of rare phenomena               & Lipid flip-flops, water permeation & Membrane scientists \\
+    Analyses of rare phenomena               & Lipid flip--flops, water permeation & Membrane scientists \\
     \\
@@
     Applications that are outside typical scope of MD simulations & 
-    Anisotropic water diffusion for pharmagokinetics and MRI imaging applications & 
+    Anisotropic water diffusion for pharmacokinetics and MRI imaging applications & 
     Scientist in fields where MD simulations are not usually applied \\
@@
 The NMRlipids databank circumvents these challenges with the overlay databank design and open collaboration approach developed in the NMRlipids project~\cite{botan15}.
-In this model, the file storage is distributed to publicly available stable repositories ({\it Data layer} in Fig.~\ref{fig:overlay} A) and maintenance of the databank does not depend on individual scientists or groups because all its version controlled content is available with an open access licence ({\it Databank layer} in Fig.~\ref{fig:overlay} A).
+In this model, the file storage is distributed to publicly available stable repositories (\textit{Data layer} in Fig.~\ref{fig:overlay} A) and maintenance of the databank does not depend on individual scientists or groups because all its version controlled content is available with an open access licence (\textit{Databank layer} in Fig.~\ref{fig:overlay} A).
 %containing relevant information about the trajectories, including the location of the raw data. This architecture streamlines the construction and maintenance of the databank
@@
 Structure of the NMRlipids databank is illustrated more detailed in Fig.~\ref{DatabankStructure} in the supplementary information.
-The core content of the databank ({\it Databank layer} in Fig.~\ref{fig:overlay}) locates as a git repository at \url{https://github.com/NMRLipids/Databank/} and is permanently stored in Zenodo repository (\url{www.zenodo.org})~\cite{??}.
+The core content of the databank (\textit{Databank layer} in Fig.~\ref{fig:overlay}) locates as a git repository at \url{https://github.com/NMRLipids/Databank/} and is permanently stored in Zenodo repository (\url{www.zenodo.org})~\cite{??}.
 Whenever a specific file is referred here, the file path within the NMRlipids databank repository is given.
@@
 
-Essential information of each simulation in the NMRlipids databank is stored in a human and machine readable README.yaml file located at {\it /Data/Simulations} in the NMRlipids databank repository.
+Essential information of each simulation in the NMRlipids databank is stored in a human and machine readable README.yaml file located at \texttt{/Data/Simulations} in the NMRlipids databank repository.
 These files contain access to all information that are needed for further analyses of simulations.
-The content of these files is described in detail in table~\ref{tab:READMEkeys} in the supplementary information.
+The content of these files is described in detail in Table~\ref{tab:READMEkeys} in the supplementary information.
 %To keep the databank light, 
-Raw MD simulation data are stored in external publicly available and stable repositories ({\it Data layer} in Fig.~\ref{fig:overlay}), such as Zenodo (\url{www.zenodo.org}), from where it can be downloaded whenever needed using the links in README.yaml files.
+Raw MD simulation data are stored in external publicly available and stable repositories (\textit{Data layer} in Fig.~\ref{fig:overlay}), such as Zenodo (\url{www.zenodo.org}), from where it can be downloaded whenever needed using the links in README.yaml files.
 
@@
 Atom names used in simulation trajectories are connected to unique atom names using mapping files that are defined in the NMRlipids project (\url{https://nmrlipids.blogspot.com/2022/04/new-yaml-format-of-mapping-files.html}).
-These files are located at {\it /Scripts/BuildDatabank/mapping\_files} in the NMRlipids databank repository.
+These files are located at \texttt{/Scripts/BuildDatabank/mapping\_files} in the NMRlipids databank repository.
 These files also define whether an atom belongs to headgroup, glycerol backbone, or acyl chain region in a lipid.
@@
 The NMRlipids databank is open for additions of simulation data by anyone.
-The first step is to create an info.yaml file containing the information that needs to be manually entered as listed in table~\ref{tab:READMEkeys}.
-This file can be then added into {\it /Scripts/BuildDatabank/info\_files} folder in the NMRlipids databank repository via pull requests.
-After the pull request is manually accepted, the rest of the information for the README.yaml file, listed in table~\ref{tab:READMEkeys}, will be automatically extracted using the {\it /Scripts/BuildDatabank/AddData.py} script.
+The first step is to create an info.yaml file containing the information that needs to be manually entered as listed in Table~\ref{tab:READMEkeys}.
+This file can be then added into \texttt{/Scripts/BuildDatabank/info\_files} folder in the NMRlipids databank repository via pull requests.
+After the pull request is manually accepted, the rest of the information for the README.yaml file, listed in Table~\ref{tab:READMEkeys}, will be automatically extracted using the \texttt{/Scripts/BuildDatabank/AddData.py} script.
 Currently the NMRlipids databank is composed of simulations found from Zenodo repository with an appropriate licence.
@@
 \subsection{Experimental data}
-Experimental data used in the quality evaluation, currently composed of C-H bond order parameters and x-ray scattering form factors, are stored in {\it /Data/experiments} in the NMRlipids databank repository.
+Experimental data used in the quality evaluation, currently composed of C--H bond order parameters and X-ray scattering form factors, are stored in \texttt{/Data/experiments} in the NMRlipids databank repository.
 Similarly to simulations, each experimental data set has a README.yaml file containing all the relevant information about the experiment.
-The keys and their descriptions for the experimental data are given in table~\ref{tab:READMEkeysEXP}.
-NMR data currently in the NMRlipids databank are taken from Refs.~\citenum{scherer87,ferreira13,antila19,melcr20,bacle21} and x-ray scattering data from Refs.~\citenum{Kucerka08a,kucerka11,pan12b,pan14,kucerka15}.
-In addition, some previously unpublished NMR and x-ray scattering data are used.
+The keys and their descriptions for the experimental data are given in Table~\ref{tab:READMEkeysEXP}.
+NMR data currently in the NMRlipids databank are taken from Refs.~\citenum{scherer87,ferreira13,antila19,melcr20,bacle21} and X-ray scattering data from Refs.~\citenum{Kucerka08a,kucerka11,pan12b,pan14,kucerka15}.
+In addition, some previously unpublished NMR and X-ray scattering data are used.
 These are measured with previously established methods as described in the supplementary information.
@@
 
-\subsection{Calculation of C-H bond order parameters}
+\subsection{Calculation of C--H bond order parameters}
 The C--H bond order parameters were calculated directly from the carbon and hydrogen positions using the definition
@@
 \end{equation}
-where angular brackets denote the ensemble average, i.e., average over all sampled configurations of all lipids in a simulation, and $\theta$ is the angle between the C--H bond and the membrane normal.
-As in previous NMRlipids publications, the order parameters were first calculated separately for each lipid and the standard error of the mean over different lipids was used as the error bar~\cite{botan15}.
-The script that calculates C-H bond order parameters from all simulations in the NMRlipids databank is available at {\it /Scripts/AnalyzeDatabank/calcOrderParameters.py} in the NMRlipids databank repository.
+where angular brackets denote the ensemble average, \textit{i.e.}, average over all sampled configurations of all lipids in a simulation, and $\theta$ is the angle between the C--H bond and the membrane normal.
+As in previous NMRlipids publications, the order parameters were first calculated separately for each lipid and the standard error of the mean over different lipids was used as the error estimate~\cite{botan15}.
+The script that calculates C--H bond order parameters from all simulations in the NMRlipids databank is available at \texttt{/Scripts/AnalyzeDatabank/calcOrderParameters.py} in the NMRlipids databank repository.
 The resulting order parameters are stored for all simulations in files named %\newline
-[{\it lipid\_name}]{\it OrderParameters.json} at folders in {\it /Data/Simulations} in the NMRlipids databank repository.
+[\texttt{lipid\_name}]\texttt{OrderParameters.json} at folders in \texttt{/Data/Simulations} in the NMRlipids databank repository.
 
-\subsection{Calculation of x-ray scattering form factors}
+\subsection{Calculation of X-ray scattering form factors}
 X-ray scattering form factors were calculated with standard equation for symmetric lipid bilayers~\cite{ollila16}
@@
 \end{equation}
-where $\Delta \rho_e(z)$ is the difference between total and solvent electron densities, and $D$ is the simulation box size in the z-direction.
-For the calculation of density profiles, atom coordinates were first centred around the centre of mass of lipid molecules for every time frame, and a histogram of these centred positions weighted with the number of electrons in each atom was then calculated with the bin width of $1/3$~Å.
+where $\Delta \rho_e(z)$ is the difference between total and solvent electron densities, and $D$ is the simulation box size in the $z$-direction (normal to the membrane).
+For the calculation of density profiles, atom coordinates were first centred around the centre of mass of lipid molecules for every time frame, and a histogram of these centred positions weighted with the number of electrons in each atom was then calculated with the bin width of $1/3$~\AA{}.
 %Similarly to order parameters, the script loops over all simulations, downloads the data on local computer, and uses the information in README.yaml file to calculate form factors. 
-The script to calculate form factors for all simulations in the NMRlipids databank is available at {\it Scripts/AnalyzeDatabank/calc\_FormFactors.py}.
-The resulting form factors are stored for all simulations in files named {\it FormFactor.json} at folders in {\it /Data/Simulations} in the NMRlipids databank repository.
+The script to calculate form factors for all simulations in the NMRlipids databank is available at \texttt{Scripts/AnalyzeDatabank/calc\_FormFactors.py}.
+The resulting form factors are stored for all simulations in files named \texttt{FormFactor.json} at folders in \texttt{/Data/Simulations} in the NMRlipids databank repository.
 
 \subsection{Calculation of a bilayer area per lipid and thickness}
-Area per lipids of bilayers were calculated by dividing the area of the simulation box with the total number of lipid molecules in the simulation.
-The script that calculates area per lipids from all simulations in the NMRlipids databank repository is available at {\it Scripts/AnalyzeDatabank/calcAPL.py} in the NMRlipids databank repository.
-The resulting area per lipids are stored for all simulations in files named {\it apl.json} at folders in {\it /Data/Simulations}.
+Area per lipids of bilayers were calculated by dividing the time-averaged area of the simulation box with the total number of lipid molecules in the simulation.
+The script that calculates area per lipids from all simulations in the NMRlipids databank repository is available at \texttt{Scripts/AnalyzeDatabank/calcAPL.py} in the NMRlipids databank repository.
+The resulting area per lipids are stored for all simulations in files named \texttt{apl.json} at folders in \texttt{/Data/Simulations}.
 
 Thicknesses of lipid bilayers were calculated from the intersections of lipid and water electron densities.
-The script that calculates thickness of all simulations in the NMRlipids databank is available at {\it Scripts/AnalyzeDatabank/calc\_thickness.py} in the NMRlipids databank repository.
-The resulting thicknesses are stored in files named {\it thickness.json} at folders in {\it /Data/Simulations} in the NMRlipids databank repository.
+The script that calculates thickness of all simulations in the NMRlipids databank is available at \texttt{Scripts/AnalyzeDatabank/calc\_thickness.py} in the NMRlipids databank repository.
+The resulting thicknesses are stored in files named \texttt{thickness.json} at folders in \texttt{/Data/Simulations} in the NMRlipids databank repository.
 
-\subsection{Quality evaluation of C-H bond order parameters}
+\subsection{Quality evaluation of C--H bond order parameters}
 As the first step to evaluate simulation qualities against experimental data, each simulation is connected to an experimental data if molar concentrations of all molecules are within $\pm$3 percentage units, charged lipids have the same counterions, and temperatures are within $\pm$2 degrees.
 For molar concentrations of water, the exact hydration level is considered only for systems with molar water to lipid ratio below 25, otherwise the systems are considered as fully hydrated.
-In practise, this is implemented by adding the path of the experimental data into the simulation README.yaml file using the {\it /Scripts/BuildDatabank/searchDATABANK.py} script in the NMRlipids databank repository.
+In practise, this is implemented by adding the path of the experimental data into the simulation README.yaml file using the \texttt{/Scripts/BuildDatabank/searchDATABANK.py} script in the NMRlipids databank repository.
 
@@
 The quality of each C--H bond order parameter is estimated by calculating the probability for a simulated value to locate within the error bars of the experimental value.
-Because conformational ensembles of individual lipids are independent in a fluid lipid bilayer, $\frac{S_{\rm CH}-\mu}{s/\sqrt{n}}$ has a Student's t-distribution with $n-1$ degrees of freedom, and the probability for an order parameter from simulation to locate within experimental error bars can be estimated from equation
+Because conformational ensembles of individual lipids are independent in a fluid lipid bilayer, $\frac{S_\mathrm{CH}-\mu}{s/\sqrt{n}}$ has a Student's t-distribution with $n-1$ degrees of freedom, and the probability for an order parameter from simulation to locate within experimental error bars can be estimated from equation
 \begin{equation}\label{probability}
-  P = f \left( \frac{S_{\rm CH} - (S_{\rm exp}+\Delta S_{\rm exp})}{s/\sqrt{n}} \right) - f \left( \frac{S_{\rm CH} - (S_{\rm exp}-\Delta S_{\rm exp})}{s/\sqrt{n}} \right),
+  P = f \left( \frac{S_\mathrm{CH} - (S_\mathrm{exp}+\Delta S_\mathrm{exp})}{s/\sqrt{n}} \right) - f \left( \frac{S_\mathrm{CH} - (S_\mathrm{exp}-\Delta S_\mathrm{exp})}{s/\sqrt{n}} \right),
 \end{equation}
-where f(t) is the %first order 
-Student's t-distribution, $\mu$ is the real mean of the order parameter, $n$ is the number of independent sample points for each C-H bond which equals the number of lipids in a simulation, $S_{\rm CH}$ is the sample mean from Eq.~\ref{OP}, $s$ is the variance of $S_{\rm CH}$ calculated over individual lipids, $S_{\rm exp}$ is the experimental value, and $\Delta S_{\rm exp}$ its error.
-The error of $\Delta S_{\rm exp} = 0.02$ is currently assumed for all experimental order parameters~\cite{ollila16}, yet more accurate may be available in the future~\cite{wurl22}.
+where $f(t)$ is the %first order 
+Student's t-distribution, $\mu$ is the real mean of the order parameter, $n$ is the number of independent sample points for each C--H bond which equals the number of lipids in a simulation, $S_\mathrm{CH}$ is the sample mean from Eq.~\ref{OP}, $s$ is the variance of $S_\mathrm {CH}$ calculated over individual lipids, $S_\mathrm{exp}$ is the experimental value, and $\Delta S_\mathrm{exp}$ its error.
+The error of $\Delta S_\mathrm{exp} = 0.02$ is currently assumed for all experimental order parameters~\cite{ollila16}, yet more accurate ones may be available in the future~\cite{wurl22}.
 Because a lipid bilayer simulation contains at least dozens of lipids, the Student's t-distribution could be safely approximated with a normal distribution.
@@
 \begin{equation}
-    P^{\rm frag}[{\rm lipid}] = \frac{\langle P[{\rm lipid}]\rangle_{\rm frag}}{p_{\rm frag}[{\rm lipid}]},
+    P^\mathrm{frag}[\mathrm{lipid}] = \frac{\langle P[\mathrm{lipid}]\rangle_\mathrm{frag}}{p_\mathrm{frag}[\mathrm{lipid}]},
 \end{equation}
-where $\langle P[{\rm lipid}]\rangle_{\rm frag}$ is the average of individual C-H bond order parameters qualities within the fragment, $p_{\rm frag}[{\rm lipid}]$ is the percentage of order parameters for which the quality is available within the fragment, and frag can be {\it sn}-1, {\it sn}-2, headgroup or total (all order parameters within a molecule).
+where $\langle P[\mathrm{lipid}]\rangle_\mathrm{frag}$ is the average of individual C-H bond order parameters qualities within the fragment, $p_\mathrm{frag}[\mathrm{lipid}]$ is the percentage of order parameters for which the quality is available within the fragment, and frag can be \textit{sn}-1, \textit{sn}-2, headgroup or total (all order parameters within a molecule).
 The overall quality of different fragments in a simulation are then defined as a molar fraction weighted average over different lipid components
 \begin{equation}
-    P^{\rm frag} = \sum_{\rm lipid} \chi_{\rm lipid} P^{\rm frag}[{\rm lipid}],
+    P^\mathrm{frag} = \sum_\mathrm{lipid} \chi_\mathrm{lipid} P^\mathrm{frag}[\mathrm{lipid}],
 \end{equation}
-where $\chi_{\rm lipid}$ is the molar fraction of a lipid in the bilayer.
+where $\chi_\mathrm{lipid}$ is the molar fraction of a lipid in the bilayer.
 
-The quality evaluation of order parameters is implemented in {\it /Scripts/BuildDatabank/QualityEvaluation.py} in the NMRlipids databank repository.
-The resulting qualities for each order parameters are stored in files named  [{\it lipid\_name}]{\it \_OrderParameters\_quality.json}, for individual lipids in files named [{\it lipid\_name}]{\it \_FragmentQuality.json}, and for overall quality for fragments in files named {\it system\_quality.json} at folders in {\it /Data/Simulations} in the NMRlipids databank repository.
+The quality evaluation of order parameters is implemented in \texttt{/Scripts/BuildDatabank/QualityEvaluation.py} in the NMRlipids databank repository.
+The resulting qualities for each order parameters are stored in files named  [\texttt{lipid\_name}]\texttt{ \_OrderParameters\_quality.json}, for individual lipids in files named [\texttt{lipid\_name}]\texttt{\_FragmentQuality.json}, and for overall quality for fragments in files named \texttt{system\_quality.json} at folders in \texttt{/Data/Simulations} in the NMRlipids databank repository.
 
-\subsection{Quality evaluation of x-ray scattering form factors}
+\subsection{Quality evaluation of X-ray scattering form factors}
 %While C-H bond order parameters relate to the conformational ensembles of individual lipids, the x-ray scattering form factors depend on membrane dimensions and density distribution~\cite{ollila16}.
-Because experiments give form factors only in relative intensity scale, they should scaled before comparing with the simulation data.
+Because experiments give form factors only in relative intensity scale, they should be scaled before comparing with the simulation data.
 Here we use the scaling coefficient for experimental intensities defined in the SIMtoEXP program~\cite{kucerka10}
@@
 
-Also a quality measure based on differences in simulated and experimental form factors accross available q-range is defined in the SIMtoEXP program~\cite{kucerka10}.
-However, the lobe heigths in simulated form factors depend on the simulation box size as shown in Fig.~\ref{fig:sizedependence}.
+Also a quality measure based on differences in simulated and experimental form factors across available $q$-range is defined in the SIMtoEXP program~\cite{kucerka10}.
+However, the lobe heights in simulated form factors depend on the simulation box size as shown in Fig.~\ref{fig:sizedependence}.
 Therefore, the quality measure defined in SIMtoEXP  would also depend on the simulation box size.
-Nevertheless, locations of form factor minima are independent on simulation box size in Fig.~\ref{fig:sizedependence}.
+In contrast, locations of form factor minima are independent on simulation box size in Fig.~\ref{fig:sizedependence}.
 Here we use the location of the first form factor minima for the quality evaluation because automatic detection of the location of second minima is inaccurate for some experimental data due to fluctuations, such as for the POPE data in Figs.~\ref{fig:quality} D and E.
@@
 The first minimum correlates well with the thickness of a membrane (Fig.~\ref{fig:quality} F), although the correlation of the second minima would be even stronger (Fig.~\ref{fig:QualityCorrelationsSI}).
-In practise, we first filter the fluctuations from the form factor data using Savitzky-Golay filter (window lenght 30 and polynomial order 1) and locate the first minima above 0.1\,\AA$^{-1}$ from both simulation and experimental data.
-The quality of a form factor is then defined as Eucledian distance between the minima in simulated and experimental form factors, $FF_q = |FF_{\rm{min}}^{\rm{sim}}-FF_{\rm{min}}^{\rm{exp}}|$.
+In practise, we first filter the fluctuations from the form factor data using Savitzky--Golay filter (window length 30 and polynomial order 1) and locate the first minima above 0.1\,\AA$^{-1}$ from both simulation and experimental data.
+The quality of a form factor is then defined as Euclidean distance between the minima in simulated and experimental form factors, $FF_q = |FF_\mathrm{min}^\mathrm{sim}-FF_\mathrm{min}^\mathrm{exp}|$.
 
-The quality evaluation of form factors is implemented in {\it /Scripts/BuildDatabank/QualityEvaluation.py} in the NMRlipids databank repository.
-The resulting form factor qualities are stored in files named {\it FormFactorQuality.json} at folders in {\it /Data/Simulations} in the NMRlipids databank repository.
+The quality evaluation of form factors is implemented in \texttt{/Scripts/BuildDatabank/QualityEvaluation.py} in the NMRlipids databank repository.
+The resulting form factor qualities are stored in files named \texttt{FormFactorQuality.json} at folders in \texttt{/Data/Simulations} in the NMRlipids databank repository.
 
-\subsection{Calculation of lipid flip-flops}
-Flip-flop rates were calculated using {\it AssignLeaflets} and {\it FlipFlop} tools from LiPyphilic package~\cite{LiPyphilic2021}.
+\subsection{Calculation of lipid flip--flops}
+Flip--flop rates were calculated using \texttt{AssignLeaflets} and \texttt{FlipFlop} tools from LiPyphilic package~\cite{LiPyphilic2021}.
 Headgroup atoms of each molecule as defined in the mapping file were used to determine in which leaflet the molecule locates.
 The midplane cut-off defining the region between leaflets was 1\,nm and frame cut-off was 100.
-This means that if the headgroup of a molecule entered within the distance of 1\,nm from the bilayer midplane and was found from the opposing leaflet after 100 steps, this event was considered as a successfull flip-flop event.
-The code that finds flip-flop events from all simulations in the NMRlipids databank is available at {\it scripts/FlipFlop.py} and the results at {\it Data/Flipflops/} in the repository at~\url{https://github.com/NMRLipids/DataBankManuscript/}.
+This means that if the headgroup of a molecule entered within the distance of 1\,nm from the bilayer midplane and was found from the opposing leaflet after 100 steps, this event was considered as a successful flip--flop event.
+The code that finds flip--flop events from all simulations in the NMRlipids databank is available at \texttt{scripts/FlipFlop.py} and the results at \texttt{Data/Flipflops/} in the repository at~\url{https://github.com/NMRLipids/DataBankManuscript/}.
 
@@
 Water permeability through membranes was calculated from equation $P=r/2c_w$, where $r$ is the rate of permeation events per time and area, and $c_w$=33.3679\,(nm)$^{-3}$ is the concentration of water in bulk~\cite{venable19}.
-The number of permeation events in each trajectory was calculated using the code by Camilo et al.~\cite{camilo2022}, available at~\url{https://github.com/crobertocamilo/MD-permeation}.
+The number of permeation events in each trajectory was calculated using the code by \citeauthor{camilo2022}~\cite{camilo2022}, available at~\url{https://github.com/crobertocamilo/MD-permeation}.
 %This code was used within the loop that goes over all simulations in the NMRlipids databank and extracts the required information using the information in README.yaml files. 
-The code that calculates permeabilities for all simulations in the NMRlipids databank is available at {\it /scripts/calcMD-PERMEATION.py} and the resulting permeabilities are stored at {\it /Data/MD-PERMEATION} in the repository containing all analyses specific for this publication at~\url{https://github.com/NMRLipids/DataBankManuscript/}.
+The code that calculates permeabilities for all simulations in the NMRlipids databank is available at \texttt{/scripts/calcMD-PERMEATION.py} and the resulting permeabilities are stored at \texttt{/Data/MD-PERMEATION} in the repository containing all analyses specific for this publication at~\url{https://github.com/NMRLipids/DataBankManuscript/}.
 %While the order parameters, form factors, area per lipid and thickness are stored within the NMRlipids databank (\url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations}), further analyses can be conventiently stored in separate repositories with the same folder structure based on hash identities of trajectory and topology files.
@@
 
-The lateral diffusion of water along the membrane surface, $D_{||}$, was calculated with Einstein's equation using {\it -lateral} option in {\it gmx msd} program within the Gromacs software package~\cite{gromacsMANUAL}.
-The code that calculates $D_{||}$ for water from all simulations in the NMRlipids databank is available at {\it /scripts/calcWATERdiffusion.py}, and the resulting diffusion coefficients are stored at {\it /Data/WATERdiffusion} in the repository at~\url{https://github.com/NMRLipids/DataBankManuscript/}.
+The lateral diffusion of water along the membrane surface, $D_{||}$, was calculated with Einstein's equation using \texttt{-lateral} option in \texttt{gmx msd} program within the Gromacs software package~\cite{gromacsMANUAL}.
+The code that calculates $D_{||}$ for water from all simulations in the NMRlipids databank is available at \texttt{/scripts/calcWATERdiffusion.py}, and the resulting diffusion coefficients are stored at \texttt{/Data/WATERdiffusion} in the repository at~\url{https://github.com/NMRLipids/DataBankManuscript/}.
 
-Water diffusion in the perpendicular direction of lipid bilayers in a multilamellar stack was estimated from the Tanner equation $D_\perp = \frac{D_{||} P z_w}{D_{||} + P z_w}$~\cite{tanner78,wasterby02}, where the water layer thickness, $z_w$, was estimated by subtracting bilayer thickness from the size of the simulation box in membrane normal direction.
+Water diffusion in the perpendicular direction of lipid bilayers in a multilamellar stack was estimated from the Tanner equation $D_\mathrm{perp} = \frac{D_{||} P z_w}{D_{||} + P z_w}$~\cite{tanner78,wasterby02}, where the water layer thickness, $z_w$, was estimated by subtracting bilayer thickness from the size of the simulation box in membrane normal direction.
 

### 2023-01-12 SI.tex
Update text and suthors

@@
 
-\title{SUPPLEMENTARY INFORMATION: NMRlipids Databank: Overlay Databank of Lipid Membrane Simulations Arising from Open Collaboration}
+\title{SUPPLEMENTARY INFORMATION: NMRlipids Databank makes data-driven analysis of biomembrane properties accessible for all}
 
-\author[1]{Anne Kiirikki}
-\author[2]{...}
-\author[1,*]{O. H. Samuli Ollila}
+\author[1]{Anne Kiirikki}         %ORCID: ?
+\author[2]{Hanne Antila}          %ORCID: 0000-0002-2474-5053
+\author[2]{Lara Bort}             %ORCID: ?
+\author[3]{Pavel Buslaev}         %ORCID: ?
+\author[4]{Fernando Favela}
+\author[5]{Tiago Mendes Ferreira}
+\author[6,7]{Patrick Fuchs}
+\author[8,9]{Rebeca Garcia-Fandino}
+\author[2]{Ivan Gushchin}
+%\author[2]{Matti Javanainen}      %ORCID: 0000-0003-4858-364X
+\author[2]{Batuhan Kav}           %ORCID: ?
+\author[2]{Milla Kurki}
+\author[2]{Alexander Kuzmin}
+\author[10,11]{Jesper Madsen}
+\author[2]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
+%\author[2]{...}                   %ORCID: ?
+\author[12]{Thomas Piggot}
+\author[2]{Suman Samantray}
+\author[2]{Fabi{\'a}n Suarez-Leston}
+\author[1,*]{O. H. Samuli Ollila} %ORCID: 0000-0002- 8728-1006
 %\author[1,2,+]{Christine Author}
 %\author[2,+]{Derek Author}
-\affil[1]{University of Helsinki, Institute of Biotechonology, Helsinki, Finland}
+
+\affil[1]{University of Helsinki, Institute of Biotechnology, Helsinki, Finland}
 \affil[2]{Affiliation, department, city, postcode, country}
+\affil[3]{Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, P.O. Box 35, Jyv{\"a}skyl{\"a}, 40014 , Finland}
+\affil[4]{Departamento de Ciencias B\'{a}sicas, Tecnol\'{o}gico Nacional de M\'{e}xico - ITS Zacatecas Occidente, Sombrerete, Zacatecas, 99102, M\'{e}xico}
+\affil[5]{NMR group - Institute for Physics, Martin Luther University Halle-Wittenberg,  Halle (Saale), 06120, Germany}
+\affil[6]{Sorbonne Universit{\'e}, Ecole Normale Sup{\'e}rieure, PSL University, CNRS, Laboratoire des Biomol{\'e}cules (LBM), Paris, 75005, France}
+\affil[7]{Universit{\'e} de Paris, UFR Sciences du Vivant, Paris, 75013, France}
+\affil[8]{Center for Research in Biological Chemistry and Molecular Materials (CiQUS), Universidade de Santiago de Compostela,  Santiago de Compostela, E-15782, Spain}
+\affil[9]{CIQUP, Centro de Investigação em Qu{\'i}mica, Departamento de Qu{\'i}mica e Bioqu{\'i}mica, Faculdade de Ci{\^e}ncias, Universidade do Porto, Porto, 4169-007, Portugal}
+\affil[10]{Department of Chemistry, The University of Chicago, Chicago, Illinois, 60637, United States of America}
+\affil[11]{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, 33612, United States of America}
+\affil[12]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}
 
@@
     \includegraphics[width = 180mm]{Figures/QualityCorrelationsSI.pdf}
-    \caption{Dependence of form factor, electron density profiles along membrane normal, and C-H bond orderparameters (from top to bottom) on simulation box size with different cholesterol concentration.
-}
+    \caption{Scatter plots and Pearson correlation coefficients, $r$, for the membrane area per lipid with X-ray scattering form factor minima (A and B), and for thickness with the average order parameter of the {\textit sn}-1 acyl chain (B) and with the second minimum from X-ray scattering form factors (D) extracted from the NMRlipids databank.
+All correlation coefficients have p-value below 0.001 as indicated by $^{***}$.}
     \label{fig:QualityCorrelationsSI}
@@
     \includegraphics[width = 180mm]{Figures/permeationSI.pdf}
-    \caption{a) Structure of an overlay databank. 
-    More detailed structure of the layer 2 in the NMRlipids databank is illustrated in Fig.~\ref{DatabankStructure} in the SI.
-    b) Distribution of the lengths of the trajectories, total number of trajectories and total lenght of the simulations in the NMRlipids databank.
-    c) Distribution of lipids present in the trajectories in the NMRlipids databank. Lipids occuring in five or less simulations ('others') are listed in the right. 
-    d) Currently available binary mixtures in the NMRlipids databank. 
-    e) Distribution of force fields in the simulations in the NMRlipids databank.
-    The figures and numbers are created on 9th of May 2022.}
+    \caption{Water permeation through membranes analyzed from the databank as a function of (A) hydration level, (B) fraction of cholesterol, (C) fraction of charged lipids, and (D) fraction of POPE in membrane.
+Values from simulations with non-zero permeation values are shown with blue dots.
+Histogrammed values are shown with black dots.
+    }
     \label{fig:permeationSI}
@@
     \includegraphics[width = 180mm]{Figures/LateralDiffusionSI.pdf}
-    \caption{a) Structure of an overlay databank. 
-    More detailed structure of the layer 2 in the NMRlipids databank is illustrated in Fig.~\ref{DatabankStructure} in the SI.
-    b) Distribution of the lengths of the trajectories, total number of trajectories and total lenght of the simulations in the NMRlipids databank.
-    c) Distribution of lipids present in the trajectories in the NMRlipids databank. Lipids occuring in five or less simulations ('others') are listed in the right. 
-    d) Currently available binary mixtures in the NMRlipids databank. 
-    e) Distribution of force fields in the simulations in the NMRlipids databank.
-    The figures and numbers are created on 9th of May 2022.}
+    \caption{Lateral diffusion of water as a function of (A) area per lipid, (B) temperature, (C) membrane thickness, and (D) fraction of charged lipids in a membrane.}
     \label{fig:diffusionSI}

### 2023-01-12 main.tex
Update text and suthors

@@
 \author[2]{Lara Bort}             %ORCID: ?
-\author[2]{Pavel Buslaev}         %ORCID: ?
+\author[3]{Pavel Buslaev}         %ORCID: ?
+\author[4]{Fernando Favela}
+\author[5]{Tiago Mendes Ferreira}
+\author[6,7]{Patrick Fuchs}
+\author[8,9]{Rebeca Garcia-Fandino}
+\author[2]{Ivan Gushchin}
 %\author[2]{Matti Javanainen}      %ORCID: 0000-0003-4858-364X
 \author[2]{Batuhan Kav}           %ORCID: ?
+\author[2]{Milla Kurki}
+\author[2]{Alexander Kuzmin}
+\author[10,11]{Jesper Madsen}
 \author[2]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
-\author[2]{...}                   %ORCID: ?
+%\author[2]{...}                   %ORCID: ?
+\author[12]{Thomas Piggot}
+\author[2]{Suman Samantray}
+\author[2]{Fabi{\'a}n Suarez-Leston}
 \author[1,*]{O. H. Samuli Ollila} %ORCID: 0000-0002- 8728-1006
@@
 %\author[2,+]{Derek Author}
+
 \affil[1]{University of Helsinki, Institute of Biotechnology, Helsinki, Finland}
 \affil[2]{Affiliation, department, city, postcode, country}
+\affil[3]{Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, P.O. Box 35, Jyv{\"a}skyl{\"a}, 40014 , Finland}
+\affil[4]{Departamento de Ciencias B\'{a}sicas, Tecnol\'{o}gico Nacional de M\'{e}xico - ITS Zacatecas Occidente, Sombrerete, Zacatecas, 99102, M\'{e}xico}
+\affil[5]{NMR group - Institute for Physics, Martin Luther University Halle-Wittenberg,  Halle (Saale), 06120, Germany}
+\affil[6]{Sorbonne Universit{\'e}, Ecole Normale Sup{\'e}rieure, PSL University, CNRS, Laboratoire des Biomol{\'e}cules (LBM), Paris, 75005, France}
+\affil[7]{Universit{\'e} de Paris, UFR Sciences du Vivant, Paris, 75013, France}
+\affil[8]{Center for Research in Biological Chemistry and Molecular Materials (CiQUS), Universidade de Santiago de Compostela,  Santiago de Compostela, E-15782, Spain}
+\affil[9]{CIQUP, Centro de Investigação em Qu{\'i}mica, Departamento de Qu{\'i}mica e Bioqu{\'i}mica, Faculdade de Ci{\^e}ncias, Universidade do Porto, Porto, 4169-007, Portugal}
+\affil[10]{Department of Chemistry, The University of Chicago, Chicago, Illinois, 60637, United States of America}
+\affil[11]{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, 33612, United States of America}
+\affil[12]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}
 
@@
 Membrane lipid composition is implicated in diseases, such as cancer and neurodegenerative disorders, and therapeutics that affect membrane compositions are emerging~\cite{torres21}.
-However, biomembranes are often difficult to study experimentally because at biological conditions they are in disordered fluid state, and sample preparation can be highly complicated due to the rich phase behaviour and complex interactions between lipids and other biomolecules.
-Consequently, the connections between complex lipid interactions and biological functions remain poorly understood.
-%why such variety of lipids is needed and how their. 
+However, biomembranes are often difficult to study experimentally because at biological conditions they are in disordered fluid state, and sample preparation can be arduous due to the complicated phase behaviour and rich interactions between lipids and other biomolecules.
+For those reasons, the detailed connections between complex lipid interactions and biological functions taking place in or around membranes remain poorly understood.
 Molecular dynamics (MD) simulations have been particularly useful in understanding membrane systems, although their accuracy has often been compromised by artefacts such as the quality of model parameters~\cite{antila22b,gupta22}.
-%While MD simulations of systems with few biomolecules continue having major contributions in our understanding on biology and drug design, 
-Presently, the accuracy of models is becoming increasingly important as researches are progressing from simulations of 
-%now searching new avenues by building models not only 
-individual molecules to simulating whole organelles or even cells using interdisciplinary approaches~\cite{johnson15,thornburg22,gupta22}. In such systems exhibiting exceedingly complex behaviour,
-model inaccuracies become increasingly difficult to detect and multiplication of small errors may have major impact on the conclusions.
+Presently, the accuracy of models is becoming increasingly important as researches are progressing from simulations of individual molecules to simulating whole organelles or even cells using interdisciplinary approaches~\cite{johnson15,thornburg22,gupta22}.
+Such systems exhibit complex emergent behaviour making inaccuracies more difficult to detect and accumulation of even modest errors may have major impact on the conclusions.
 
-In contrast to experimental structural biology, where standard protocols to share and quality-evaluate the resolved structures are firmly established \textit{via} the Protein Data Bank (PDB)~\cite{montelione13}, such best practices are yet to be defined for MD simulations.
-Importance of such approaches is widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} and data-sharing solutions are emerging for proteins in solution~\cite{meyer10,kamp10} and in membranes~\cite{newport19,espigares20,leston22}, as well as for nucleic acids~\cite{hospital16}, cyclodextrins~\cite{mixcoha16}, COVID-19--involved macromolecules~(\url{bioexcel-cv19.bsc.es}) and arbitrary biomolecules~\cite{bekker20}.
-However,
-%The relevance of quality evaluation of simulation trajectories in databanks regarding technical details of simulations and accuracy of the underlying physical description of the system (force field) has become evident~\cite{tai04,meyer10,hospital20} and such
-automatic quality evaluation of simulation data~\cite{meyer10,hospital16} and programmatic access are still rare. 
-%However, straightforward quality comparisons between individual simulations or force fields within these databanks remain challenging. 
+In contrast to experimental structural biology, where standard protocols to share and quality-evaluate the resolved structures are firmly established \textit{via} the Protein Data Bank (PDB)~\cite{montelione13}, the equivalent best practices are yet to be defined for MD simulations.
+the importance of such approaches is widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} and data-sharing solutions are emerging for proteins in solution~\cite{meyer10,kamp10}, proteins in membranes~\cite{newport19,espigares20,leston22}, as well as for nucleic acids~\cite{hospital16}, cyclodextrins~\cite{mixcoha16}, COVID-19--involved macromolecules~(\url{bioexcel-cv19.bsc.es}) and arbitrary biomolecules~\cite{bekker20}.
+However, automatic quality evaluation of simulation data~\cite{meyer10,hospital16} and programmatic access is still rare.
 In particular, tools for automatic quality evaluation of membrane simulations, or training sets for machine learning models of membrane-containing systems, are not yet available.
@@
 
-%Applications of MD simulations in understanding biomolecular behaviour are steadily increasing, but protocols for quality evaluation and sharing simulation data fall behind the ones in more elaborated fields, such as structural biology.
-Development of such protocols has been limited by practical challenges, such as lack of universally defined quality measures for MD simulations, incompatible naming conventions in different models and simulation softwares, and high technical requirements for hardware to share large trajectories from MD simulations, as well as incentives for scientist to evaluate and share their MD simulation data.
-
 Here we present the NMRlipids Databank---a community-driven, open-for-all database featuring programmatic access to atom-resolution MD simulations of lipid bilayers.
 The Databank performs automatic quality evaluation of membrane simulations, which facilitates the selection of best models for specific applications and accelerates the development of simulation parameters and methodology.
-The programmatic access enables users to apply data-driven approaches on the Databank, thus facilitating the creation of new tools for (and by) researchers in a wide range of fields, both in academia and industry, from cell membrane biology and lipid nanoparticle formulations to computational chemistry and machine learning.
-As two usage examples of what is already possible, we demonstrate here (i) how data-driven analysis of water anisotropic diffusion in all the membrane systems available in the Databank can extend the scope of MD simulations to magnetic resonance imaging (MRI) and pharmacokinetics, and (ii) how the Databank allows its users to analyse rare phenomena that are beyond the scope of standard MD simulation studies.
+The programmatic access enables users to apply data-driven approaches on the Databank, thus facilitating the creation of new tools for (and by) researchers in a wide range of fields covering academia and industry, from cell membrane biology and lipid nanoparticle formulations to computational chemistry and machine learning.
+As two usage examples of what is already possible, we demonstrate here (i) how data-driven analysis of water anisotropic diffusion in all the membrane systems available in the Databank can extend the scope of MD simulations to magnetic resonance imaging (MRI) and pharmacokinetics, and (ii) how the Databank allows its users to analyse rare phenomena that are beyond the scope of standard MD simulation investigations.
 Wider adaptation of the NMRlipids Databank will open even more possibities.
-%The practical relevance of the NMRlipids databank is exemplified by automatic quality evaluation and ranking of large amount of MD simulation data,
 
-While the NMRlipids Databank currently contains only lipid bilayer systems, it can be readily expanded to contain other molecules, such as disordered proteins or sugars. 
-%a solution for lipid bilayers based on overlay databank structure illustrated in Fig.~\ref{fig:overlay}.  
-The powerful combination of the overlay databank structure and open collaboration can be applied also when building databases in other fields,  
-%is developed here to solve the practical challenges in generating databanks of MD simulation data enabling flexible analyses over large sets of simulation data, but it can be  potentially used for wide range of situations.
-particularly when storage of raw data requires significant resources, best practices in the field are not defined, and incentives to share data do not exist.
-%overlay databank approach lowers the barrier to start without compromizing the long term stability or scalability.
+While the NMRlipids Databank currently contains only lipid bilayer systems, it can readily be expanded to contain other molecules, such as disordered proteins or sugars.
+The powerful combination of the overlay databank structure and a community-wide open collaboration can be applied also when building databases in other fields, particularly when storage of raw data requires significant resources, best practices in the field are not defined, and incentives to share data do not exist.
 Therefore, the approach introduced here opens up possibilities to create databases enabling data-driven and machine learning applications in new fields.
-
-%The Introduction section, of referenced text\cite{Figueredo:2009dg} expands on the background of the work (some overlap with the Abstract is acceptable).
-The introduction should not include subheadings.
 
@@
 
-%Up to three levels of \textbf{subheading} are permitted. Subheadings should not be numbered.
 \subsection{Programmatic access to MD simulation data of membranes composed of most common biologically abundant lipids}
 
-NMRlipids Databank is a community driven database containing atomistic MD simulations of biologically relevant lipid membranes emerging from the NMRlipids open collaboration~\cite{botan15,ollila16,catte16,antila19,bacle21}.
-Nearly thousand simulation trajectories with the total length approaching half a millisecond can be accessed programmatically using Python API or graphical user interface (\url{http://www.databank.nmrlipids.fi/}).
-%NMRlipids databank is the only databank of MD simulations with open programmatic access to its content enabling fully flexible analyses of its content. 
-Following the FAIR principles~\cite{wilkinson16} and NMRlipids project protocol, not only the content and computer programs related to the databank but also the whole construction process of the databank are open access~\cite{botan15}.
+NMRlipids Databank is a community-driven database containing atomistic MD simulations of biologically relevant lipid membranes emerging from the NMRlipids open collaboration~\cite{botan15,ollila16,catte16,antila19,bacle21}.
+Nearly thousand simulation trajectories with the total length approaching half a millisecond can be accessed programmatically using the Python application programming interface (API) or a graphical user interface (GUI) (\url{http://www.databank.nmrlipids.fi/}).
+Following the principles of improving data Findability, Accessibility, Interoperability, and Reuse (FAIR)~\cite{wilkinson16} and NMRlipids project protocol, not only the content and computer programs related to the databank, but also the entire construction process of the databank are open access~\cite{botan15}.
 %The NMRlipids databank will be always open for data and other contributions.
@@
 Currently available single component lipid membranes and binary mixtures in the NMRlipids databank are illustrated in Fig.~\ref{fig:overlay}D.
-While these compose the majority of available trajectories, also mixtures up to five lipids are available.
+While these comprise the majority of available trajectories, mixtures with up to five lipid types are available.
 The distribution of available simulations containing a specific lipid is shown in Fig.~\ref{fig:overlay}C.
-This roughly resembles the biological abundancies of different lipid types, PC being the most common followed by cholesterol, PE, PS, PG, PI, and other lipids depending on organism and organelle~\cite{vanmeer08}.
-The force fields used in the simulations cover all the essential parameters used in lipid simulations, ranging from highly popular CHARMM36 parameters to united atom and polarizable force fields.
+This roughly resembles the biological abundance of different lipid types with phosphatidylcholine (PC) being the most common followed by cholesterol, phosphatidylethanolamine (PE), phosphatidylserine (PS), phosphatidylglycerol (PG), phosphatidylinositol (PI), and other lipids depending on organism and organelle~\cite{vanmeer08}.
+The force fields used in the simulations cover all the essential parameters commonly used in lipid simulations, ranging from highly popular CHARMM36 parameters to united atom and polarizable force fields.
 Therefore, the averages calculated over the databank can be considered as mean predictions from available lipid models for an average cell membrane.
-%Currently the databank is composed of approximately 500 trajectories with the total length of approximately 231 microseconds of which most are contributed for the previous publications from the NMRlipids open collaboration~\cite{botan15,catte16,antila19,bacle21}.
-The distribution of lipids, force fields, length of the trjajectories and available binary mixtures are shown in Fig.~\ref{fig:overlay}.
 
@@
 The {\it application layer} contains outputs from the databank, such as graphical user interface (\url{http://www.databank.nmrlipids.fi/}) and results from analyses described in sections below.
-More detailed description on the NMRlipids databank structure is available the supplementary information.
-
-%are enabled by the 
-%The databank can be used in layer 3 by accessing the raw data and information stored in the databank by employing the universal naming conventions for atoms, molecules and simulation details.
-The applications can be linked to the core databank (layer 2) without actually including them, thereby enabling flexible extension of the data without compromising the simplicity and lightness of the core databank.
+A more detailed description on the NMRlipids databank structure is available the supplementary information.
 
@@
 %Quality of membrane simulations with different force fields have been evaluated against experimental data during parameterization and in separate comparison studies~\cite{botan15,ollila16,catte16,pluhackova16,perez17,leonard19,??}, but the lack of universal quality measure for membrane simulations complicates the estimations of reliability of simulations.
-Understanding the accuracy of the used force field parameters is crucial when estimating the reliability and significance of MD simulation results.
-However, the lack of universal quality measures for MD simulations and complicated landscape of force field quality for membranes hamper the quality of simulations which has resulted to many controversial results~\cite{antila22b}.
+Understanding the accuracy of the used force field parameters is crucial when assessing the reliability and significance of MD simulation results.
+However, the lack of universal quality measures for MD simulations and the complicated landscape of force field quality for membranes hamper the applicability of such simulations, resulting in substantial controversies~\cite{antila22b}.
 For example, previous quality evaluations of membrane simulations have concluded that CHARMM36 parameters give the best description for lipid headgroup conformational ensembles~\cite{bacle21}, GROMOS-CKP parameters best capture the membrane packing in POPS bilayers~\cite{antila22b}, and that OPLS3e parameters overcome CHARMM36 in structural quality for POPC but predicts overestimated ion binding~\cite{kurki22}.
-%, thereby being a major obstacle in many applications of membrane MD simulations. 
-To enable the navigation in this complex force field quality landscape, we define here quantitative quality measures that can be used to rapidly find the best available parameters for a specific application or to guide force field development.
+To enable the navigation in this complex force field milieu, we define here quantitative quality measures that can be used to rapidly find the best available parameters for a specific application or to guide the development of improved force fields.
 For detailed definitions of quality measures, see the supplementary information.
@@
     B) Best simulations ranked based on the overall order parameter quality of POPE lipid. 
-    C)--E) Evaluation against experimental (X-ray scattering  NMR order parameter) data exemplified for a simulation with the best overall order parameter quality (C), the best quality for POPE lipid (D), and the headgroup quality for POPE (E).
-    F) Scatter plots and Pearson correlation coefficients for the membrane area per lipid, thickness, first minima of x-ray scattering form factor and average order parameter of the {\textit sn}-1 acyl chain extracted from the NMRlipids databank.
-    All correlation coefficients have p-value below 0.001. For more correlations see Fig.~\ref{fig:QualityCorrelationsSI}.
+    C)--E) Evaluation against experimental (NMR order parameters and X-ray scattering ) data exemplified for a simulation with the best overall order parameter quality (C), the best quality for POPE lipid (D), and the headgroup quality for POPE (E).
+    F) Scatter plots and Pearson correlation coefficients, $r$, for the membrane area per lipid, thickness, first minimum of X-ray scattering form factor and average order parameter of the {\textit sn}-1 acyl chain extracted from the NMRlipids databank.
+All correlation coefficients have p-value below 0.001 as indicated by $^{***}$.
+For more correlations see Fig.~\ref{fig:QualityCorrelationsSI}.
     }
@@
 
-Ranking of all simulations with the highest scores based on evaluation against C--H bond order parameters from NMR experiments is shown in Fig.~\ref{fig:quality} A and ranking of best POPE lipid bilayer simulations in Fig.~\ref{fig:quality} B.
-Direct comparisons with experiments for selected simulations are shown in Figs.~\ref{fig:quality} C--D.
-C--H bond order parameters are sensitive to conformational ensembles of individual lipid molecules~\cite{ollila16}, but they correlate well also with the membrane thickness and lateral packing as shown in Fig.~\ref{fig:quality} F.
-This is complemented by comparing the absolute values of X-ray scattering form factors between simulations and experiments that are related to the overall membrane dimensions \textit{via} electron density profile.
-Particularly, the location of minima in these form factors are correlated with the membrane dimensions as shown in Figs.~\ref{fig:quality} F,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}.
-
-%best simulation, POPC with OPLS3e, is shown in
-%based on C-H bond order parameters from NMR experiments and x-ray scattering form factors that enable the rapid quality evaluation of membrane MD simulations in the NMRlipids databank.
-
-%Order parameters and scattering form factors are robust experimental measurables that can be directly connected to the simulation data~\cite{ollila16}.
-
-%NMRlipids databank contains also experimental order parameter and x-ray form factor data that is connected to corresponding  simulations in order to define the simulation qualities and rank them to select the suitable simulation models for particular applications.
-
-%Fig.~\ref{fig:quality} C, where discrepancies with experiments are observed only for the third carbon in the $\it{sn}$-1 chain, and carbons 8 and 9 close to the double bond in $\it{sn}$-2 chain.
-The best overall quality for POPE is found in GROMOS-CKP simulation, for which the direct comparison in Fig.~\ref{fig:quality} D show minor differences in acyl chains, but major discrepancies in glycerol backbone and in the first carbon of $\it{sn}$-2 chain.
-These parts are better described for POPE by CHARMM36 parameters (direct comparison shown in Fig.~\ref{fig:quality} E), but its overall quality suffers from the overestimated order in acyl chains.
+Ranking of all simulations with the highest scores based on evaluation against C--H bond order parameters from NMR experiments is shown in Fig.~\ref{fig:quality}A and ranking of best POPE lipid bilayer simulations in Fig.~\ref{fig:quality}B.
+Direct comparisons with experiments for selected simulations are shown in Figs.~\ref{fig:quality}C--E.
+C--H bond order parameters are sensitive to conformational ensembles of individual lipid molecules~\cite{ollila16} while also correlating well also with the membrane thickness and lateral packing as shown in Fig.~\ref{fig:quality}F.
+This is complemented by comparing the absolute values of X-ray scattering form factors between simulations and experiments that are related to the overall membrane dimensions \textit{via} the electron density profile.
+In particular, the location of minima in these form factors are correlated with the membrane dimensions as shown in Figs.~\ref{fig:quality}F,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}.
 
@@
 Predictions from different force fields deviate significantly in terms of absolute values and the slope of decrease upon addition of POPE.
-Because the area per lipid correlates with the average order parameter of $\textit{sn}$-1 chain (Fig.~\ref{fig:quality} F), simulations with the best predictions for area per lipids can be selected based on order parameter quality evaluation.
-Quality evaluation in Fig.~\ref{fig:quality} and direct comparison in Fig.~\ref{fig:POPC_POPE_dataSI} reveal that Slipids ranks 1\textsuperscript{st} for POPC with a clear difference to others, and 2\textsuperscript{nd} for POPE with only marginally lower quality than GROMOS-CKP, therefore being the best selection for studies where packing effects of PE lipids are relevant.
-%The direct comparison with experiments in Fig.~\ref{fig:POPC_POPE_dataSI} in the supplementary information reveals that 
-Simulations with CHARMM36 and GROMOS-CKP predict too packed bilayer for POPC (overestimated order in Fig.~\ref{fig:POPC_POPE_dataSI}), while area per lipid for POPC is overestimated in Lipid17.
-For POPE,
-%order parameters in  Slipids and GROMOS-CKP are close to experiments suggesting that their are per lipids are most reasonable, while
-CHARMM36 and Lipid17 predict too packed membranes.
+Because the area per lipid correlates with the average order parameter of the $\textit{sn}$-1 chain (Fig.~\ref{fig:quality} F), simulations with the best predictions for area per lipids can be selected based on order parameter quality evaluation.
+Quality evaluation in Fig.~\ref{fig:quality} and direct comparison in Fig.~\ref{fig:POPC_POPE_dataSI} reveal that Slipids ranks 1\textsuperscript{st} for POPC with a clear difference to others, and 2\textsuperscript{nd} for POPE with only marginally lower quality than GROMOS-CKP and is therefore the best available choice for studies where packing effects of PE lipids are relevant.
+Simulations with CHARMM36 and GROMOS-CKP predict overly packed bilayer for POPC (overestimated order in Fig.~\ref{fig:POPC_POPE_dataSI}), while the area per lipid for POPC is overestimated in Lipid17.
+For POPE, CHARMM36 and Lipid17 predict membranes that are too packed.
 
-In summary, 
-%the quality evaluation based on the NMRlipids databank suggests that the Slipids parameters would give the most reliable description for membrane packing of POPC:POPE mixtures.
-Slipids was a relatively good model also for POPC:POPS mixtures in similar comparisons utilizing the preliminary version of the NMRlipids databank~\cite{antila22b}, although charged membranes are more complicated as the counterion binding affinity plays a significant role in membrane packing~\cite{antila19}.
-On the other hand, CHARMM36 overcomes Slipids when the focus is on conformational ensembles of headgroups  in different types of lipids because the accuracy of glycerol backbone is low in Slipids~\cite{bacle21}.
-Altogether, these examples demonstrate how
-the automatic quality evaluation of the NMRlipids databank enables the rapid selection of the best simulation parameters to specific applications without extensive and tedious manual force field evaluation.
+In summary, the automatic quality evaluation of the NMRlipids databank enables the rapid exploration and selection of the best simulation parameters for specific applications without extensive and tedious manual force field evaluation.
 This possibility will promote more reliable simulation results for wide range of applications, support force field development, and streamline the parametrization of coarse grained force fields against the best atomistic MD simulations.
 The automatically extracted quantitative quality measures will be particularly useful to guide automatic parametrization procedures.
+
+%Understanding such complex picture of lipid bilayer MD simulation quality would not be possible without automatic quality evaluation of large sets of simulations enabled by the NMRlipids databank.
 
@@
 
-%On the other hand, the first form factor minima is correlated stronger with membrane thickness (Fig.~\ref{fig:quality} C) than with the area per lipid (Fig.~\ref{fig:QualityCorrelationsSI}). 
-%Scatter plots and Pearson correlation coefficients in Fig.~\ref{fig:quality} C reveal strong correlation between membrane area per lipid and thickness.
-
-%Understanding such complex picture of lipid bilayer MD simulation quality would not be possible without automatic quality evaluation of large sets of simulations enabled by the NMRlipids databank.
-
-%The power of NMRlipids quality metrics in selecting the best model for particular application is demontrated in Fig.~\ref{fig:quality} c) where the area per lipids of POPC:POPE mixtures from different force fields are shown and quality of POPE simulation in different force fields are evaluated.
-
- %This is demonstrated in Fig.~\ref{fig:quality} a) showing correlations between membrane lateral density (area per lipid), thickness, minima of form factor and acyl chain order analyzed from all apporimately 500 trajectories in the NMRlipids databank.
-Membrane area per lipid has negative correlations with membrane order and thickness with Pearson coefficient of -0.78 and -0.49, respectively, because decreased area leads to more ordered, tightly packed and stretched acyl chains leading to thicker membranes.
-On the other hand, locations of form factor minima have positive correlation with the membrane area and negative correlation with the thickness.
-In conclusion, both x-ray scattering form factors and acyl chain order parameters from NMR are on average good proxies for membrane packing and thickness, while NMR order parameters give segmental resolution information on the quality of conformational ensembles of lipids.
-
 \subsection{Detecting rare phenomena using NMRlipids databank: Cholesterol flip--flops}
 Lipid flip--flops from one bilayer leaflet to another play an important role in lipid trafficking and regulating membrane properties~\cite{vanmeer08}.
-Phospholipid flip--flops are slow, occurring with the timescales of hours or days, while cholesterol, diacylglycerol and ceramide flip-flops are faster, yet the reported timescales range from minutes to sub-millisecods~\cite{vanmeer08,steck12,parisio16,gu19}.
+Phospholipid flip--flops are slow when not facilitated by proteins, occurring spontaneously on the timescale of hours or days, while cholesterol, diacylglycerol and ceramide flip-flops are faster.
+Still, the reported timescales range from minutes to sub-millisecods~\cite{vanmeer08,steck12,parisio16,gu19}.
 These timescales were previously accessible only by coarse-grained simulations or free energy calculations~\cite{parisio16}, yet atomistic simulations reporting cholesterol flip--flop events have been published recently~\cite{gu19,javanainen19,baral20}.
 %However, systematic studies on how cholesterol flip-flop rate depends on membrane properties are still beyond the current computational resources available for a single research group. 
-These studies report increase in cholesterol flip--flop rates with increasing acyl chain unsaturation level and decreasing cholesterol concentration~\cite{gu19,javanainen19}, but correlations between cholesterol flip--flop rates and membrane properties have not been systematically studied.
-Here we demonstrate that the large amount of simulation data available in the NMRlipids databank, along with programmatic access providing the option to flexibly implement varied queries of the data, enables anybody to perform such analyses.
+These studies report an increase in cholesterol flip--flop rates with increasing acyl chain unsaturation level and decreasing cholesterol concentration~\cite{gu19,javanainen19}, but correlations between cholesterol flip--flop rates and membrane properties have not been systematically studied.
+Here, we demonstrate that the large amount of simulation data available in the NMRlipids databank, along with programmatic access providing the option to flexibly implement varied queries of the data, enables anybody to perform such analyses.
 
@@
 Flip--flops were not observed for other lipids, giving the upper limits for PC lipid flip--flop rate of 9\,s$^{-1}$ and for ceramide (\textit{N}-palmitoyl-\textsc{D}-\textit{erythro}-sphingosine) of 0.002\,\textmu{}s$^{-1}$.
-Therefore, data in the NMRlipids databank suggests that the lipid flip--flip rate decreases in the order of diacylglycerols $>$ cholesterol $>$ other lipids including ceramides.
+Thus, available data in the NMRlipids databank suggests that the lipid flip--flip rate decreases in the order of diacylglycerols $>$ cholesterol $>$ other lipids including ceramides.
 However, the amount of data for diacylgycerols (8 simulations with Lipid17 force field) and ceramide (3 simulations with CHARMM36 force field) is less than for cholesterol (83 simulations), thus we cannot fully exclude the effect of force field or composition on this comparison.
@@
 Flip--flop rates increase by an order of magnitude when membrane packing density decreases and a major jump is observed at low membrane packing.
-The order of magnitude changes in cholesterol flip.-flop rate with the membrane composition may have major implications in understanding lipid trafficking and membrane biochemistry~\cite{gu19,baral20}.
+The order of magnitude changes in cholesterol flip--flop rate with the membrane composition may have major implications in understanding lipid trafficking and membrane biochemistry~\cite{gu19,baral20}.
 Because the results from the NMRlipids databank are averaged over large range of membrane compositions and force fields, they show that the strong dependence of cholesterol flip--flop rate on membrane properties is not limited to certain lipid compositions or force fields used in previous studies~\cite{gu19,javanainen19,baral20}.
@@
 
-Cholesterol flip--flops are phenomena whose analyses from MD simulations has only recently became possible for research groups with substantial resources and special expertise~\cite{gu19,javanainen19,baral20}.
-Our analysis demonstrates how the NMRlipids databank makes analyses of such rare events accessible to wide range of scientists in different fields of science and industry who do not have the access to required resources and expertise to produce the large amounts of MD simulation data.
+Explorations of cholesterol flip--flops using MD simulations have only recently become possible for research groups with substantial resources and specialized expertise~\cite{gu19,javanainen19,baral20}.
+Our analysis demonstrates how the NMRlipids databank makes analyses of such rare events accessible to wide range of scientists in different fields of science and industry who do not have the access to required resources and expertise to produce the large amounts of MD simulation data required.
 
@@
 The anisotropy in water diffusion in parallel and perpendicular directions with respect to membranes plays a role in the drug translocation through biological material, particularly in skin \cite{hansen13,wen18,nitsche19,roberts21}, and in MRI imaging~\cite{topgaard20}.
-MD simulations are rarely used to analyze anistropic diffusion of water as only few membrane permeation events for water are typically observed in a single MD simulation trajectory~\cite{venable19,camilo2022}, thereby making the collection of sufficient amount of data challenging.
-Here we show that the NMRlipids databank can be used to analyze how the anisotropic water diffusion in multilamellar membrane systems depends on membrane properties.
+MD simulations are rarely used to analyze anistropic diffusion of water since only few membrane permeation events for water are typically observed in a single MD simulation trajectory~\cite{venable19,camilo2022}, thereby making the collection of sufficient amount of data challenging or impossible.
+Here, we show that the NMRlipids databank can be used to systemically analyze how the anisotropic water diffusion in multilamellar membrane systems depends on membrane properties.
 
@@
 
-To this end, we first calculated the water permeability through membranes from all simulations in the NMRlipids databank.
+To this end, we calculated the water permeability through membranes from all simulations in the NMRlipids databank.
 The resulting non-zero values range between 0.3 and 322\,\textmu{}m/s with the mean and median of 14\,\textmu{}m/s and 8\,\textmu{}m/s, respectively.
-These values agree with the previously reported simulation results~\cite{venable19,camilo2022} and have the same order of magnitude as experimentally determined diffusive permeability coefficients, but are on average larger than the values reported for PC lipids in liquid crystalline phase,~0.19--0.33\,\textmu{}m/s~\cite{jansen95}.
+These values agree with the previously reported simulation results~\cite{venable19,camilo2022} and have the same order of magnitude as experimentally determined diffusive permeability coefficients, but are on average larger than the values reported for PC lipids in the liquid crystalline phase,~0.19--0.33\,\textmu{}m/s~\cite{jansen95}.
 Observed permeabilities and their histogrammed values are shown in Figs.~\ref{fig:permeability} B--E as a function of temperature, membrane thickness, area per lipid, and acyl chain order.
-%on average, we calculated the average permeabilities over all systems with the value of temperature, thickness, area per lipid, or acyl chain order within a fixed range.
-These averaged permeabilities also shown in Figs.~\ref{fig:permeability} with black dots.
-As expected, the permeability increases with the temperature, giving 17\,$\pm$\,4\,$k_\mathrm{B}T$ for the average energy barrier for the water permeation from the Arrhenius plot in Fig.~\ref{fig:permeability} B.
+As expected, the permeability increases with the temperature, giving an average energy barrier of 17\,$\pm$\,4\,$k_\mathrm{B}T$ for the water permeation from the Arrhenius plot in Fig.~\ref{fig:permeability} B.
 On the other hand, the permeability of water decreases on average when membranes become more packed, \textit{i.e.}, with decreasing area per lipid and increasing thickness and acyl chain order (Figs.~\ref{fig:permeability} C--E).
-Permeation of water through bilayers depends on membrane properties also according previous studies, but it is not agreed whether the area per lipid~\cite{nagle08} or bilayer thickness~\cite{frallicciardi22} would be the main parameter determining the permeability.
+Permeation of water through bilayers also depends on membrane properties also according previous studies, but there is no established consensus on whether the area per lipid~\cite{nagle08} or bilayer thickness~\cite{frallicciardi22} is the main parameter determining the permeability.
 %is previously reported from simulations and experiments for some systems, but not for all~\cite{mathai07,venable19,frallicciardi22}. 
@@
 The parallel diffusion coefficient of water, $D_{||}$, decreases with reduced hydration and increases with the temperature, but dependencies on membrane area per lipid, thickness, or fraction of charged lipids were not observed in Figs.~\ref{fig:permeability} and~\ref{fig:diffusionSI}.
-%The experimental water diffusion increases with increasing hydration level toward the value for bulk water . 
-%As expected, water translational diffusion increase in the diffusion is observed with the increased temperature and hydration level (Figs.~\ref{fig:permeability} H and~\ref{fig:diffusionSI} B). 
-%The results at different hydration levels are shown in Fig.~\ref{fig:permeability} F together with the experimental data~\cite{rudakova04}. 
-Simulation results are close to the experimental values with low hydration levels in Fig.~\ref{fig:permeability} F, but increase to the values of approximately two times higher than experimental bulk water diffusion value (3.1$\cdot$\,10$^{-9}$\,m$^2$/s at 313\,K~\cite{khakimov08}) with high hydration levels.
+Simulation results are close to the experimental values with low hydration levels in Fig.~\ref{fig:permeability} F, but increase to approximately two times higher than the experimental value for bulk water diffusion value (3.1$\cdot$\,10$^{-9}$\,m$^2$/s at 313\,K~\cite{khakimov08}) with high hydration levels.
 This is not surprising as the most common water model used in membrane simulations, TIP3P, overestimates the bulk water diffusion~\cite{pathirannahalage21}.
 To estimate the diffusion anisotropy of water, $D_\mathrm{\perp}/D_{||}$, in multilamellar membrane system, the permeability coefficients of water through membranes were translated to perpendicular diffusion coefficients, $D_\mathrm{\perp}$, using the Tanner equation~\cite{tanner78,wasterby02}.
-The resulting perpendicular diffusion coefficients are approximately five orders of magnitude smaller than lateral diffusion coefficients of water (Figs.~\ref{fig:permeability} G and H), which is at the upper limit of the anisotropic estimated from the experimental data~\cite{nitsche19}.
+The resulting perpendicular diffusion coefficients are approximately five orders of magnitude smaller than lateral diffusion coefficients of water (Figs.~\ref{fig:permeability} G and H), which is at the upper limit of the anisotropy estimated from the experimental data~\cite{nitsche19}.
 %Large anisotropy values are understandable as simulations give slightly slower permeation rates and higher lateral diffusion rates than experiments. 
@@
 Several folds larger anisotropy in membranes with higher lateral density is expected to play a role in pharmacokinetic models not only for water but also for other hydrophilic molecules~\cite{nitsche19}.
-Furthermore, the understanding of this anisotropy may help in developing new (functional) MRI imaging methods~\cite{topgaard20}.
+Furthermore, the enhanced understanding of this anisotropy may help in developing new (functional) MRI imaging methods~\cite{topgaard20}.
 
@@
 The focus of biomolecular simulations is moving from studies of individual molecules to larger complexes and even whole cells and organelles~\cite{johnson15,thornburg22,gupta22}.
-On the other hand, machine learning based models predicting behaviour of biomolecules and automatic approaches to parametrize models are emerging~\cite{jumper21,antila22b}.
+Simultaneously, machine learning based models predicting behaviour of biomolecules and automatic approaches to parametrize models are emerging~\cite{jumper21,antila22b}.
 The NMRlipids databank will support the development in all these directions.
-The automatic quality evaluation and ranking of simulations in the NMRlipids databank guides the optimization of new force field parameters and the selection of best parameters for large biomolecular complexes where simulation quality becomes increasingly important due the multiplication of small errors in large simulations.
-On the other hand, the NMRlipids databank serves as a training set for diverse machine learning applications.
-%Most straightforward applications include models that would predict connections between membrane properties that are already stored in the databank. 
+The automatic quality evaluation and ranking of simulations in the NMRlipids databank guides the optimization of new force field parameters and the selection of best parameters for large biomolecular complexes where simulation quality becomes increasingly important due the accumulation of small errors in large simulations.
+Furthermore, the NMRlipids databank serves as a training set for diverse machine learning applications.
 For example, a machine learning model predicting electron density profiles from form factors would be highly useful in interpretation of scattering experiments.
-Furthermore, the data contained in the NMRlipids databank will enable training of more elaborated models to predict arbitrary membrane properties.
+In addition, the data contained in the NMRlipids databank will enable training of more elaborated models to predict arbitrary membrane properties.
 
 Providing programmatic access to large scale MD simulation data in the NMRlipids databank can lead to applications in unprecedented directions in fields where MD simulations are less commonly utilized.
-Here we exemplified this by analysing the anisotropic diffusion of water (Fig.~\ref{fig:permeability}), which is relevant in pharmacokinetic modeling and in MRI imaging where MD simulations are not yet commonly used~\cite{nitsche19,topgaard20}.
+Here, we exemplified this by analysing the anisotropic diffusion of water (Fig.~\ref{fig:permeability}), which is relevant in pharmacokinetic modeling and in MRI imaging, both areas where MD simulations are not yet commonly used~\cite{nitsche19,topgaard20}.
 %The NMRlipids databank will be of great benefit to scientists involved in MD simulations.
@@
 The selection of best models for a specific application is demonstrated for PC/PE lipid mixtures in Fig.~\ref{fig:quality}, and previously for PC/PS lipid mixtures~\cite{antila22b}, and lipid headgroups~\cite{bacle21}.
-On the other hand, programmatic access in the NMRlipids databank enables automatic analyses over larger sets of simulation data in terms of quantity (\textit{e.g.}, simulation length and number of conformations) and content (\textit{e.g.}, lipid compositions and ion concentrations) that is currently possible in a single research group.
-Added value of such access to simulation data is demonstrated here for the analysis of simulation qualities (Fig.~\ref{fig:quality}), water permeation through membranes (Fig.~\ref{fig:permeability}) and cholesterol flip--flop events (Fig.~\ref{fig:flip-flops}).
+The programmatic access in the NMRlipids databank enables automatic analyses over larger sets of simulation data in terms of quantity (\textit{e.g.}, simulation length and number of conformations) and content (\textit{e.g.}, lipid compositions and ion concentrations) than is currently possible by a single research group.
+The added value that access to this simulation data provides is demonstrated here for the analysis of simulation qualities (Fig.~\ref{fig:quality}), water permeation through membranes (Fig.~\ref{fig:permeability}) and cholesterol flip--flop events (Fig.~\ref{fig:flip-flops}).
 %Such models could be used to connect membrane composition to relevant membrane properties in wide range of fields from molecular biology to drug development and biomolecular imaging. 
-Different types of applications enabled by the NMRlipids databank in wide range of fields are exemplified in Table~\ref{tab:applications}.
-The increasing amount of data is expected to further increase the scope of potential applications of the NMRlipids databank in wide range of fields ranging from molecular biology and biotechnology to
-%the design of lipid nanoparticles for drug delivery or other applications, and 
-material science and biomolecular imaging. 
-cide
+Different types of applications enabled by the NMRlipids databank across a wide range of fields are exemplified in Table~\ref{tab:applications}.
+The ever growing amount of data is expected to further increase the scope of potential applications of the NMRlipids databank in biology, biotechnology,material science, biomolecular imaging and beyond.
+
 \begin{table}[t]
@@
 
-Main practical limitations in building open access databanks of molecular dynamics simulations have been the required commitment in long term support for hardware and software maintenance, and the lack of incentives for researchers to share the data.
+The main practical limitations in building open access databanks of molecular dynamics simulations have been the required commitment in long term support for hardware and software maintenance, and the lack of incentives for researchers to share the data.
 The NMRlipids databank circumvents these challenges with the overlay databank design and open collaboration approach developed in the NMRlipids project~\cite{botan15}.
 In this model, the file storage is distributed to publicly available stable repositories (\textit{Data layer} in Fig.~\ref{fig:overlay} A) and maintenance of the databank does not depend on individual scientists or groups because all its version controlled content is available with an open access licence (\textit{Databank layer} in Fig.~\ref{fig:overlay} A).
-%containing relevant information about the trajectories, including the location of the raw data. This architecture streamlines the construction and maintenance of the databank
-%
-%, the demand of hardware can be distributed and open collaboration model reduces the risk for ending software maintenance. 
-%Furthermore, the 
-%by distributing the long term commitment for hardware and software maintenance. 
 Incentive to share the data is created in the NMRlipids open collaboration by offering authorship in published articles to the contributors following the NMRlipids project protocol~\cite{botan15}.
-%, thereby creating an incentive for contributions. 
-%
-%Universal quality measures and naming conventions for molecules and atoms are defined in the NMRlipids databank to enable the automatic quality evaluation and other analyses.
-%
-%of lipid bilayer MD simulations introduced in the NMRlipids databank enables rapid ranking of available simulation models agains experimental NMR and x-ray scattering data. 
-%
-The current NMRlipids databank contains only lipid bilayer simulations, but the concept can be applied also to other biomolecules, such as disordered proteins and membrane-protein systems,
-%combining open collaboration and overlay databank structure can be applied also 
-or other fields where similar barriers to establish publicly accessible databanks exist. 
-%The overlay databank design can open such avenues also for other
-
-%
-%, thereby not requiring large resources to handle and maintain. 
-%This lowers the barrier for starting such databank as well as for long term storage. 
-%The NMRlipids databank  
-%In addition to all computers where the databank is developed and used, the NMRlipids databank git is stored to Zenodo server, thereby enabling a very cost effective long term storage for the databank.
+The current NMRlipids databank contains only lipid bilayer simulations, but the concept can also be applied to other biomolecules, such as disordered proteins and membrane-protein systems, or other fields where similar barriers to establish publicly accessible databanks exist.
 
@@
 The content of these files is described in detail in Table~\ref{tab:READMEkeys} in the supplementary information.
-%To keep the databank light, 
 Raw MD simulation data are stored in external publicly available and stable repositories (\textit{Data layer} in Fig.~\ref{fig:overlay}), such as Zenodo (\url{www.zenodo.org}), from where it can be downloaded whenever needed using the links in README.yaml files.
@@
 After the pull request is manually accepted, the rest of the information for the README.yaml file, listed in Table~\ref{tab:READMEkeys}, will be automatically extracted using the \texttt{/Scripts/BuildDatabank/AddData.py} script.
-Currently the NMRlipids databank is composed of simulations found from Zenodo repository with an appropriate licence.
+Currently, the NMRlipids databank is composed of simulations found from Zenodo repository with an appropriate licence.
 Most, but not all, of these trajectories originate from previous NMRlipids projects~\cite{botan15,catte16,antila19,bacle21}.
@@
 For the calculation of density profiles, atom coordinates were first centred around the centre of mass of lipid molecules for every time frame, and a histogram of these centred positions weighted with the number of electrons in each atom was then calculated with the bin width of $1/3$~\AA{}.
-%Similarly to order parameters, the script loops over all simulations, downloads the data on local computer, and uses the information in README.yaml file to calculate form factors. 
 The script to calculate form factors for all simulations in the NMRlipids databank is available at \texttt{Scripts/AnalyzeDatabank/calc\_FormFactors.py}.
@@
 \subsection{Quality evaluation of C--H bond order parameters}
-As the first step to evaluate simulation qualities against experimental data, each simulation is connected to an experimental data if molar concentrations of all molecules are within $\pm$3 percentage units, charged lipids have the same counterions, and temperatures are within $\pm$2 degrees.
+As the first step to evaluate simulation qualities against experimental data, each simulation is connected to an experimental data if molar concentrations of all molecules are within $\pm$3 percentage units, charged lipids have the same counterions, and temperatures are within $\pm$2\,K.
 For molar concentrations of water, the exact hydration level is considered only for systems with molar water to lipid ratio below 25, otherwise the systems are considered as fully hydrated.
@@
 
-%\subsubsection{Conformational ensembles using C-H bond order parameters}
-%
-%The quality of conformational ensembles of individual lipid molecules in simulations are evaluated using the C-H bond order parameters which can be directly compared with robust experimental data~\cite{ollila16}.
-
 The quality of each C--H bond order parameter is estimated by calculating the probability for a simulated value to locate within the error bars of the experimental value.
-Because conformational ensembles of individual lipids are independent in a fluid lipid bilayer, $\frac{S_\mathrm{CH}-\mu}{s/\sqrt{n}}$ has a Student's t-distribution with $n-1$ degrees of freedom, and the probability for an order parameter from simulation to locate within experimental error bars can be estimated from equation
+Because conformational ensembles of individual lipids are assumed to be independent in a fluid lipid bilayer, $\frac{S_\mathrm{CH}-\mu}{s/\sqrt{n}}$ has a Student's t-distribution with $n-1$ degrees of freedom, and the probability for an order parameter from simulation to locate within experimental error bars can be estimated from equation
 \begin{equation}\label{probability}
@@
 To avoid such numerical instabilities, we use the first order Student's t-distribution having slightly higher probabilities for values far away from the mean.
-%Because phospholipids sample their conformational ensemble within nanosecond timescale~\cite{ferreira15}, all simulations in the databank would be sufficiently long to sample the realistic conformational phase of individual lipids.
 On the other hand, some force fields exhibit too slow dynamics which leads to large error bars in order parameter values~\cite{antila21a}.
-Such large error bars widen the Student's t-distribution in Eq.~\ref{probability} thereby artificially increasing the probability to find the simulated value within experimental error bars.
+Such artificially slow dynamics widens the Student's t-distribution in Eq.~\ref{probability}, thereby increasing the probability to find the simulated value within experimental error bars.
 Therefore, the order parameters with simulation error bars above the experimental error 0.02 are not included in the quality evaluation.
-
-%We estimate the probability for the simulation order parameter to locate within experimental
-%error bars from equation
-%\begin{equation}\label{probability}
-%  P = T(S_{\rm exp}+\Delta S_{\rm exp}) - T(S_{\rm exp}-\Delta S_{\rm exp}),
-%\end{equation}
-%where T is the first order Student's t-distribution shifted to $\bar{X}$ and
-%scaled with $s$. 
-%where the probability is calculated from the normal distribution
-%\begin{equation}\label{gaussian}
-%    P = \int_{S_{\rm exp}-\Delta S_{\rm exp}}^{S_{\rm exp}+
-%    \Delta S_{\rm exp}}  \frac{1}{\sigma \sqrt{2\pi}} e^{-\frac{1}{2}(\frac{x-\mu}{\sigma})^2} {\rm d}x.
-%\end{equation}
-%$S_{\rm exp}$ and $\Delta S_{\rm exp}$ are experimental order parameter and its error, and $\mu$ and $\sigma$ are the mean order parameter and its standard deviation from simulations.
-The quality measure $S_q$ approaches to zero when probability for agreement between simulation and experimental results approach one, and increases when simulated and experimental values diverge.
 
@@
 
-Also a quality measure based on differences in simulated and experimental form factors across available $q$-range is defined in the SIMtoEXP program~\cite{kucerka10}.
+Also, a quality measure based on differences in simulated and experimental form factors across available $q$-range is defined in the SIMtoEXP program~\cite{kucerka10}.
 However, the lobe heights in simulated form factors depend on the simulation box size as shown in Fig.~\ref{fig:sizedependence}.
@@
 Here we use the location of the first form factor minima for the quality evaluation because automatic detection of the location of second minima is inaccurate for some experimental data due to fluctuations, such as for the POPE data in Figs.~\ref{fig:quality} D and E.
-%In order to define a quality measure which is independent on simulation box size, we measure the form factor quality based on the location of the first minima observed in experiments. 
 The first minimum correlates well with the thickness of a membrane (Fig.~\ref{fig:quality} F), although the correlation of the second minima would be even stronger (Fig.~\ref{fig:QualityCorrelationsSI}).
@@
 Headgroup atoms of each molecule as defined in the mapping file were used to determine in which leaflet the molecule locates.
-The midplane cut-off defining the region between leaflets was 1\,nm and frame cut-off was 100.
+The midplane cut-off defining the region between leaflets was 1\,nm and frame cutoff was 100.
 This means that if the headgroup of a molecule entered within the distance of 1\,nm from the bilayer midplane and was found from the opposing leaflet after 100 steps, this event was considered as a successful flip--flop event.
@@
 Water permeability through membranes was calculated from equation $P=r/2c_w$, where $r$ is the rate of permeation events per time and area, and $c_w$=33.3679\,(nm)$^{-3}$ is the concentration of water in bulk~\cite{venable19}.
-The number of permeation events in each trajectory was calculated using the code by \citeauthor{camilo2022}~\cite{camilo2022}, available at~\url{https://github.com/crobertocamilo/MD-permeation}.
+The number of permeation events in each trajectory was calculated using the code by Camilo et al.~\cite{camilo2022}, available at~\url{https://github.com/crobertocamilo/MD-permeation}.
 %This code was used within the loop that goes over all simulations in the NMRlipids databank and extracts the required information using the information in README.yaml files. 

### 2023-01-13 main.tex
Updates from Overleaf

@@
 In contrast to experimental structural biology, where standard protocols to share and quality-evaluate the resolved structures are firmly established \textit{via} the Protein Data Bank (PDB)~\cite{montelione13}, the equivalent best practices are yet to be defined for MD simulations.
-the importance of such approaches is widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} and data-sharing solutions are emerging for proteins in solution~\cite{meyer10,kamp10}, proteins in membranes~\cite{newport19,espigares20,leston22}, as well as for nucleic acids~\cite{hospital16}, cyclodextrins~\cite{mixcoha16}, COVID-19--involved macromolecules~(\url{bioexcel-cv19.bsc.es}) and arbitrary biomolecules~\cite{bekker20}.
+the importance of such approaches is widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} and data-sharing solutions are emerging for proteins in solution~\cite{meyer10,kamp10}, proteins in membranes~\cite{newport19,espigares20,leston22}, nucleic acids~\cite{hospital16}, nucleic acids and proteins~\cite{bekker20}, cyclodextrins~\cite{mixcoha16}, COVID-19--involved macromolecules~(\url{bioexcel-cv19.bsc.es}).
 However, automatic quality evaluation of simulation data~\cite{meyer10,hospital16} and programmatic access is still rare.
@@
 
-\subsection{Programmatic access to MD simulation data of membranes composed of most common biologically abundant lipids}
+\subsection{Programmatic access to MD simulations of membranes composed of most common biologically abundant lipids}
 
 NMRlipids Databank is a community-driven database containing atomistic MD simulations of biologically relevant lipid membranes emerging from the NMRlipids open collaboration~\cite{botan15,ollila16,catte16,antila19,bacle21}.
-Nearly thousand simulation trajectories with the total length approaching half a millisecond can be accessed programmatically using the Python application programming interface (API) or a graphical user interface (GUI) (\url{http://www.databank.nmrlipids.fi/}).
-Following the principles of improving data Findability, Accessibility, Interoperability, and Reuse (FAIR)~\cite{wilkinson16} and NMRlipids project protocol, not only the content and computer programs related to the databank, but also the entire construction process of the databank are open access~\cite{botan15}.
-%The NMRlipids databank will be always open for data and other contributions.
-
-Currently available single component lipid membranes and binary mixtures in the NMRlipids databank are illustrated in Fig.~\ref{fig:overlay}D.
+It has been designed to improve data Findability, Accessibility, Interoperability, and Reuse (FAIR)~\cite{wilkinson16} using the NMRlipids project protocol where all the content is openly accessible from the beginning of the project~\cite{botan15}.
+The NMRlipids databank contains currently nearly thousand simulation trajectories with the total length approaching half a millisecond.
+The distribution of available simulations containing a specific lipid in Fig.~\ref{fig:overlay}C roughly resembles the biological abundance of different lipid types with phosphatidylcholine (PC) being the most common followed by cholesterol, phosphatidylethanolamine (PE), phosphatidylserine (PS), phosphatidylglycerol (PG), phosphatidylinositol (PI), and other lipids depending on organism and organelle~\cite{vanmeer08}.Currently available single component lipid membranes and binary mixtures in the NMRlipids databank are illustrated in Fig.~\ref{fig:overlay}D.
 While these comprise the majority of available trajectories, mixtures with up to five lipid types are available.
-The distribution of available simulations containing a specific lipid is shown in Fig.~\ref{fig:overlay}C.
-This roughly resembles the biological abundance of different lipid types with phosphatidylcholine (PC) being the most common followed by cholesterol, phosphatidylethanolamine (PE), phosphatidylserine (PS), phosphatidylglycerol (PG), phosphatidylinositol (PI), and other lipids depending on organism and organelle~\cite{vanmeer08}.
-The force fields used in the simulations cover all the essential parameters commonly used in lipid simulations, ranging from highly popular CHARMM36 parameters to united atom and polarizable force fields.
+Force fields used in simulations cover all the essential parameters commonly used in lipid simulations, see Fig.~\ref{fig:overlay}E, including also united atom and polarizable force fields.
 Therefore, the averages calculated over the databank can be considered as mean predictions from available lipid models for an average cell membrane.
 
-The efficient upcycling of large MD simulation trajectories in the NMRlipids databank is enabled by the overlay structure illustrated in Fig.~\ref{fig:overlay}A.
-Raw simulation data in the {\it data layer} can be stored in any publicly available location with long term stability and permanent links to the data, such as digital object identifiers (DOIs) or uniform resource names (URNs).
+The overlay structure, illustrated in Fig.~\ref{fig:overlay}A, is designed to enable efficient upcycling of MD simulations for data-driven and machine learning applications using 
+The NMRlipids databank.
+Raw simulation data in the {\it data layer} can be stored in any publicly available location with long term stability, such as Zenodo (\url{zenodo.org}), and with permanent links to the data, such as digital object identifiers (DOIs).
+%or uniform resource names (URNs). 
 The {\it Databank layer} is the core of the databank containing all the relevant information about the simulations, including links to the raw data, relevant metadata describing the systems, universal naming conventions for lipids and their atoms, computer programs to create and analyse the entries, basic properties calculated from all simulations (area per lipid, C--H bond order parameters, X-ray scattering form factors and membrane thickness), and quality evaluation of simulations against experimental data.
-In practise, this information is stored in the git repository which is currently available at~\url{https://github.com/NMRLipids/Databank}.
-The {\it application layer} contains outputs from the databank, such as graphical user interface (\url{http://www.databank.nmrlipids.fi/}) and results from analyses described in sections below.
+In practise, this information is stored in the git repository which is available at~\url{https://github.com/NMRLipids/Databank}.
+The {\it application layer} contains outputs from the databank, such as graphical user interface (\url{http://www.databank.nmrlipids.fi/}) and results from analyses described in sections below which are available at \url{https://github.com/NMRLipids/DataBankManuscript}.
 A more detailed description on the NMRlipids databank structure is available the supplementary information.
@@
 \end{figure}
+
+\subsection{Accessing NMRlipids databank}
+The content of the NMRlipids databank is easily searchable and browsable via graphical user interface (GUI) (\url{http://www.databank.nmrlipids.fi/}), which can be also used to visualize the systems, compare basic properties between diffent systems and experiments, order the simulations based on their quality.
+
+%\subsection{Analyzing the databank}
+More flexible analyses, enabling wide range of novel applications can be performed by the Python application programming interface (API) available at \url{https://github.com/NMRLipids/Databank}. 
+Programs that analyze large data sets from the NMRlipids databank loop over README.yaml files of all simulations, download the raw simulation data to the local computer, and perform the analysis from the downloaded trajectories.
+When molecule or atom names are needed in the analysis, they will be read from the README.yaml file and mapping files defined therein.
 
@@
 
-\subsection{Analyzing the databank}
-Programs that analyze large data sets from the NMRlipids databank loop over README.yaml files of all simulations, download the raw simulation data to the local computer, and perform the analysis from the downloaded trajectories.
-When molecule or atom names are needed in the analysis, they will be read from the README.yaml file and mapping files defined therein.
-
 \subsection{Calculation of C--H bond order parameters}

### 2023-01-16 main.tex

@@
 \subsection{Accessing NMRlipids databank}
-The content of the NMRlipids databank is easily searchable and browsable via graphical user interface (GUI) (\url{http://www.databank.nmrlipids.fi/}), which can be also used to visualize the systems, compare basic properties between diffent systems and experiments, order the simulations based on their quality.
+The content of the NMRlipids databank is easily searchable and browsable via graphical user interface (GUI) (\url{http://www.databank.nmrlipids.fi/}), which can be also used to visualize the systems, compare basic properties between diffent systems and experiments, and rank simulations based on their quality evaluated against experimental data.
 
 %\subsection{Analyzing the databank}
-More flexible analyses, enabling wide range of novel applications can be performed by the Python application programming interface (API) available at \url{https://github.com/NMRLipids/Databank}. 
-Programs that analyze large data sets from the NMRlipids databank loop over README.yaml files of all simulations, download the raw simulation data to the local computer, and perform the analysis from the downloaded trajectories.
-When molecule or atom names are needed in the analysis, they will be read from the README.yaml file and mapping files defined therein.
+More flexible analyses that enable wide range of novel applications can be performed by the Python application programming interface (API) available at \url{https://github.com/NMRLipids/Databank}.
+The API can be used to automatically analyse virtually any property from all simulations in the NMRlipids databank following the flowchart demonstrated in Fig.~\ref{fig:POPCPOPEapls}A.
+Iteration through the README.yaml files in the {\it databank layer} enables the access to raw simulation data via the links stored in these files.
+Affiliation of universal molecule and atom names to the simulation specific names in README.yaml and mapping files enable automatic analyses independently on the specific naming convention used in a simulation.
+Analysis codes and results for membrane area per lipid, C–H bond order parameters, X-ray scattering form factors and thickness are included in the {\it Databank layer}.
+Any other analyses can be implement by users and the results can be stored to the application layer.
+
+These tools offer open programmatic access to unprecedented amount of simulation trajectories for all, which enables possibilities for wide range of novel analyses.
+This is particularly useful for scientist without access to resources to perform large scale MD simulations and for analyses of rare phenomena that require exceptional amounts of simulation data.
+On the other hand, universal naming convention enables systematic comparisons between simulations having different model parameters and thereby different naming conventions.
+This will guide the development of more accurate simulation parameters and usage of atomistic models to train coarse grained simulations.
+
+\begin{figure}[tbp]
+    \centering
+    \includegraphics[width=88mm]{Figures/flowchart2.pdf}
+    \caption{Differences in area per lipids of POPC and POPE bilayers between different force fields.  
+    }
+    \label{fig:POPCPOPEapls}
+\end{figure}
 
@@
 %Quality of membrane simulations with different force fields have been evaluated against experimental data during parameterization and in separate comparison studies~\cite{botan15,ollila16,catte16,pluhackova16,perez17,leonard19,??}, but the lack of universal quality measure for membrane simulations complicates the estimations of reliability of simulations.
-Understanding the accuracy of the used force field parameters is crucial when assessing the reliability and significance of MD simulation results.
-However, the lack of universal quality measures for MD simulations and the complicated landscape of force field quality for membranes hamper the applicability of such simulations, resulting in substantial controversies~\cite{antila22b}.
-For example, previous quality evaluations of membrane simulations have concluded that CHARMM36 parameters give the best description for lipid headgroup conformational ensembles~\cite{bacle21}, GROMOS-CKP parameters best capture the membrane packing in POPS bilayers~\cite{antila22b}, and that OPLS3e parameters overcome CHARMM36 in structural quality for POPC but predicts overestimated ion binding~\cite{kurki22}.
-To enable the navigation in this complex force field milieu, we define here quantitative quality measures that can be used to rapidly find the best available parameters for a specific application or to guide the development of improved force fields.
-For detailed definitions of quality measures, see the supplementary information.
+Understanding of how closely simulations resemble the reality is crucial when assessing the reliability and significance of MD simulation results, as well as when using atomistic simulations in parametrization of coarse grained models.
+However, universal quality measures for MD simulations are lacking, and a prevailing force field that would perform best in all situations has not been found for membranes.
+For example, OPLS3e parameters overcome CHARMM36 in structural quality for POPC but predicts overestimated ion binding~\cite{kurki22}.
+%and the complicated landscape of force field quality for membranes hamper the applicability of such simulations, resulting in substantial controversies~\cite{antila22b}. 
+%For example, previous quality evaluations of membrane simulations have concluded that 
+On the other hand, CHARMM36 gives the best description for lipid headgroup conformational ensembles~\cite{bacle21}, whileGROMOS-CKP parameters best capture the membrane packing in POPS bilayers~\cite{antila22b}.
+Furthermore, Fig.~\ref{fig:POPCPOPEapls}B illustrates divergent predictions for the difference in packing between POPC and POPE membranes from different force fields.
+To streamline the selection of best simulations for specific applications, and to guide the development of atomistic and coarse grained parameters, quantitative quality measures for lipid bilayer simulations are defined in the NMRlipids databank.
+%that can be used to rapidly find the best available parameters .
+
+The quality for each simulation in the NMRlipids databank with the corresponding experimental data available is currently done using C--H bond order parameters from NMRexperiments and form factors from X-ray scattering experiments.
+The C--H bond order parameters are sensitive to conformational ensembles of individual lipid molecules~\cite{ollila16}, while correlating well also with the membrane thickness and lateral packing as shown in Fig.~\ref{fig:quality}F.
+The X-ray scattering form factors are related to membrane dimensions \textit{via} the electron density profile.
+In particular, locations of minima in these form factors are correlated with the membrane lateral packing and thickness as shown in Figs.~\ref{fig:quality}F,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}.
+For quality measures, probabilities for each order parameter to locate within experimental error and distance of the first form factor minimum from the experimental value are calculated from a simulation.
+%For detailed definitions of quality measures, see the supplementary information.
+
+Ranking based on C--H bond order parameter quality for the best performing simulations is shown in Fig.~\ref{fig:quality}A, and for the best performing POPE lipid bilayer simulations in Fig.~\ref{fig:quality}B.
+Direct comparisons with experiments are shown for selected simulations in Figs.~\ref{fig:quality}C--E.
+%The power of the NMRlipids databank in selecting the best simulation parameters for a specific application is demonstrated for mixtures of PE and PC lipids in Fig.~\ref{fig:POPCPOPEapls} showing area per lipids from available POPC and POPE bilayer simulations in the databank from different force fields at 310\,K.
+Predictions from different force fields deviate significantly in terms of absolute values and the slope of decrease upon addition of POPE.
+This can be used to select the the simulations that best caputer area per lipids for POPE and POPC lipids in Fig.~\ref{fig:POPCPOPEapls}B because the area per lipid correlates with the average order parameter of the $\textit{sn}$-1 chain (Fig.~\ref{fig:quality} F).
+%, simulations with the best predictions for area per lipids can be selected based on order parameter quality evaluation. 
+Quality evaluation in Fig.~\ref{fig:quality} and direct comparison in Fig.~\ref{fig:POPC_POPE_dataSI} reveal that Slipids ranks 1\textsuperscript{st} for POPC with a clear difference to others, and 2\textsuperscript{nd} for POPE with only marginally lower quality than GROMOS-CKP.
+Simulations with CHARMM36 and GROMOS-CKP predict overly packed bilayer for POPC (overestimated order in Fig.~\ref{fig:POPC_POPE_dataSI}), while the area per lipid for POPC is overestimated in Lipid17.
+For POPE, CHARMM36 and Lipid17 predict membranes that are too packed.
+In conclusion, Slipids parameters is the best available choice for studies where packing effects of PE lipids are relevant
 
@@
 
-Ranking of all simulations with the highest scores based on evaluation against C--H bond order parameters from NMR experiments is shown in Fig.~\ref{fig:quality}A and ranking of best POPE lipid bilayer simulations in Fig.~\ref{fig:quality}B.
-Direct comparisons with experiments for selected simulations are shown in Figs.~\ref{fig:quality}C--E.
-C--H bond order parameters are sensitive to conformational ensembles of individual lipid molecules~\cite{ollila16} while also correlating well also with the membrane thickness and lateral packing as shown in Fig.~\ref{fig:quality}F.
-This is complemented by comparing the absolute values of X-ray scattering form factors between simulations and experiments that are related to the overall membrane dimensions \textit{via} the electron density profile.
-In particular, the location of minima in these form factors are correlated with the membrane dimensions as shown in Figs.~\ref{fig:quality}F,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}.
-
-%\subsection{Finding the best models for PC and PE lipid mixtures}
-The power of the NMRlipids databank in selecting the best simulation parameters for a specific application is demonstrated for mixtures of PE and PC lipids in Fig.~\ref{fig:POPCPOPEapls} showing area per lipids from available POPC and POPE bilayer simulations in the databank from different force fields at 310\,K.
-Predictions from different force fields deviate significantly in terms of absolute values and the slope of decrease upon addition of POPE.
-Because the area per lipid correlates with the average order parameter of the $\textit{sn}$-1 chain (Fig.~\ref{fig:quality} F), simulations with the best predictions for area per lipids can be selected based on order parameter quality evaluation.
-Quality evaluation in Fig.~\ref{fig:quality} and direct comparison in Fig.~\ref{fig:POPC_POPE_dataSI} reveal that Slipids ranks 1\textsuperscript{st} for POPC with a clear difference to others, and 2\textsuperscript{nd} for POPE with only marginally lower quality than GROMOS-CKP and is therefore the best available choice for studies where packing effects of PE lipids are relevant.
-Simulations with CHARMM36 and GROMOS-CKP predict overly packed bilayer for POPC (overestimated order in Fig.~\ref{fig:POPC_POPE_dataSI}), while the area per lipid for POPC is overestimated in Lipid17.
-For POPE, CHARMM36 and Lipid17 predict membranes that are too packed.
-
 In summary, the automatic quality evaluation of the NMRlipids databank enables the rapid exploration and selection of the best simulation parameters for specific applications without extensive and tedious manual force field evaluation.
@@
 %Understanding such complex picture of lipid bilayer MD simulation quality would not be possible without automatic quality evaluation of large sets of simulations enabled by the NMRlipids databank.
-
-\begin{figure}[tbp]
-    \centering
-    \includegraphics[width=88mm]{Figures/APL.pdf}
-    \caption{Differences in area per lipids of POPC and POPE bilayers between different force fields.  
-    }
-    \label{fig:POPCPOPEapls}
-\end{figure}
 

### 2023-01-16 main.tex

@@
 It has been designed to improve data Findability, Accessibility, Interoperability, and Reuse (FAIR)~\cite{wilkinson16} using the NMRlipids project protocol where all the content is openly accessible from the beginning of the project~\cite{botan15}.
-The NMRlipids databank contains currently nearly thousand simulation trajectories with the total length approaching half a millisecond.
-The distribution of available simulations containing a specific lipid in Fig.~\ref{fig:overlay}C roughly resembles the biological abundance of different lipid types with phosphatidylcholine (PC) being the most common followed by cholesterol, phosphatidylethanolamine (PE), phosphatidylserine (PS), phosphatidylglycerol (PG), phosphatidylinositol (PI), and other lipids depending on organism and organelle~\cite{vanmeer08}.Currently available single component lipid membranes and binary mixtures in the NMRlipids databank are illustrated in Fig.~\ref{fig:overlay}D.
+Currently the NMRlipids databank contains nearly thousand simulation trajectories with the total length approaching half a millisecond.
+The distribution of available simulations containing a specific lipid in Fig.~\ref{fig:overlay}C roughly resembles the biological abundance of different lipid types with phosphatidylcholine (PC) being the most common followed by cholesterol, phosphatidylethanolamine (PE), phosphatidylserine (PS), phosphatidylglycerol (PG), phosphatidylinositol (PI), and other lipids depending on organism and organelle~\cite{vanmeer08}.
+Currently available single component lipid membranes and binary mixtures in the NMRlipids databank are illustrated in Fig.~\ref{fig:overlay}D.
 While these comprise the majority of available trajectories, mixtures with up to five lipid types are available.
@@
 Analysis codes and results for membrane area per lipid, C–H bond order parameters, X-ray scattering form factors and thickness are included in the {\it Databank layer}.
-Any other analyses can be implement by users and the results can be stored to the application layer.
+Further analysis codes can be implemented by users and the results can be stored to the {\it application layer} in the same format that is used in the {\it Databank layer}.
+The results from analyses can be then programmatically accessed and furhter analyzed using the NMRlipids databank API following the flowchart demonstrated in Fig.~\ref{fig:POPCPOPEapls}B.
 
@@
 
-\begin{figure}[tbp]
+\begin{figure}[t]
     \centering
     \includegraphics[width=88mm]{Figures/flowchart2.pdf}
-    \caption{Differences in area per lipids of POPC and POPE bilayers between different force fields.  
+    \caption{ A) Flowchart for analysing properties from all MD simulations in the NMRlipids databank.
+    B) Flowchart for accessing the results analyzed from the NMRlipids databank.
+    C) Area per lipids of POPC and POPE lipid bilayers predicted by different force fields.  
     }
@@
 %For example, previous quality evaluations of membrane simulations have concluded that 
-On the other hand, CHARMM36 gives the best description for lipid headgroup conformational ensembles~\cite{bacle21}, whileGROMOS-CKP parameters best capture the membrane packing in POPS bilayers~\cite{antila22b}.
+On the other hand, CHARMM36 gives the best description for lipid headgroup conformational ensembles~\cite{bacle21}, while GROMOS-CKP parameters best capture the membrane packing in POPS bilayers~\cite{antila22b}.
 Furthermore, Fig.~\ref{fig:POPCPOPEapls}B illustrates divergent predictions for the difference in packing between POPC and POPE membranes from different force fields.
 To streamline the selection of best simulations for specific applications, and to guide the development of atomistic and coarse grained parameters, quantitative quality measures for lipid bilayer simulations are defined in the NMRlipids databank.
+Together with the API, these will enable rapid and automatic quality evaluation of simulations in the NMRlipids databank, thereby promoting the selection of best simulations for specific applications without laborious force field evaluation, and automatic parametrization procedures for atomistic and coarse grained simulations.
 %that can be used to rapidly find the best available parameters .
 
-The quality for each simulation in the NMRlipids databank with the corresponding experimental data available is currently done using C--H bond order parameters from NMRexperiments and form factors from X-ray scattering experiments.
+The quality for each simulation in the NMRlipids databank with the corresponding experimental data available is currently done using C--H bond order parameters from NMR experiments and form factors from X-ray scattering experiments.
 The C--H bond order parameters are sensitive to conformational ensembles of individual lipid molecules~\cite{ollila16}, while correlating well also with the membrane thickness and lateral packing as shown in Fig.~\ref{fig:quality}F.
@@
 In particular, locations of minima in these form factors are correlated with the membrane lateral packing and thickness as shown in Figs.~\ref{fig:quality}F,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}.
-For quality measures, probabilities for each order parameter to locate within experimental error and distance of the first form factor minimum from the experimental value are calculated from a simulation.
+For quantitative quality measures, probabilities for each order parameter to locate within experimental error, and the distance of the first form factor minimum from the experimental value are calculated from a simulation.
 %For detailed definitions of quality measures, see the supplementary information.
@@
 
-In summary, the automatic quality evaluation of the NMRlipids databank enables the rapid exploration and selection of the best simulation parameters for specific applications without extensive and tedious manual force field evaluation.
+%In summary, the automatic quality evaluation of the NMRlipids databank enables the rapid exploration and selection of the best simulation parameters for specific applications without extensive and tedious manual force field evaluation.
 This possibility will promote more reliable simulation results for wide range of applications, support force field development, and streamline the parametrization of coarse grained force fields against the best atomistic MD simulations.

### 2023-01-17 main.tex
Update text

@@
 In contrast to experimental structural biology, where standard protocols to share and quality-evaluate the resolved structures are firmly established \textit{via} the Protein Data Bank (PDB)~\cite{montelione13}, the equivalent best practices are yet to be defined for MD simulations.
-the importance of such approaches is widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} and data-sharing solutions are emerging for proteins in solution~\cite{meyer10,kamp10}, proteins in membranes~\cite{newport19,espigares20,leston22}, nucleic acids~\cite{hospital16}, nucleic acids and proteins~\cite{bekker20}, cyclodextrins~\cite{mixcoha16}, COVID-19--involved macromolecules~(\url{bioexcel-cv19.bsc.es}).
+The importance of such approaches is widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} and data-sharing solutions are emerging for proteins in solution~\cite{meyer10,kamp10}, proteins in membranes~\cite{newport19,espigares20,leston22}, nucleic acids~\cite{hospital16}, nucleic acids and proteins~\cite{bekker20}, cyclodextrins~\cite{mixcoha16}, COVID-19--involved macromolecules~(\url{bioexcel-cv19.bsc.es}).
 However, automatic quality evaluation of simulation data~\cite{meyer10,hospital16} and programmatic access is still rare.
 In particular, tools for automatic quality evaluation of membrane simulations, or training sets for machine learning models of membrane-containing systems, are not yet available.
-The profound, and sometimes unexpected, effect that such open data sharing and quality evaluation can have on the field cannot be understated.
+Recent advances using machine learning approaches utilizing publicly available databanks, for example to solve protein structures~\cite{jumper21}, emphasize the increasing importance of such resources.
+%The profound, and sometimes unexpected, effect that such open data sharing and quality evaluation can have on the field cannot be understated.
 At the very least, open access data will reduce the man and computational hours originating from double work, reducing the environmental impact.
@@
 Here we present the NMRlipids Databank---a community-driven, open-for-all database featuring programmatic access to atom-resolution MD simulations of lipid bilayers.
-The Databank performs automatic quality evaluation of membrane simulations, which facilitates the selection of best models for specific applications and accelerates the development of simulation parameters and methodology.
 The programmatic access enables users to apply data-driven approaches on the Databank, thus facilitating the creation of new tools for (and by) researchers in a wide range of fields covering academia and industry, from cell membrane biology and lipid nanoparticle formulations to computational chemistry and machine learning.
@@
 Wider adaptation of the NMRlipids Databank will open even more possibities.
+Furthermore, the Databank performs automatic quality evaluation of membrane simulations, which facilitates the selection of best models for specific applications and accelerates the development of simulation parameters and methodology.
 
@@
 
-\subsection{Accessing NMRlipids databank}
+\subsection{Accessing NMRlipids databank}\label{section:access}
 The content of the NMRlipids databank is easily searchable and browsable via graphical user interface (GUI) (\url{http://www.databank.nmrlipids.fi/}), which can be also used to visualize the systems, compare basic properties between diffent systems and experiments, and rank simulations based on their quality evaluated against experimental data.
@@
     B) Flowchart for accessing the results analyzed from the NMRlipids databank.
-    C) Area per lipids of POPC and POPE lipid bilayers predicted by different force fields.  
+    C) Area per lipids of POPC and POPE lipid bilayers predicted by different force fields at 310\,K.
+The points from the best performing simulations in the NMRlipids databank quality evaluation for each lipid (Figs.~\ref{fig:quality} and~\ref{fig:POPC_POPE_dataSI}) are surrounded by black circles.
     }
@@
 For example, OPLS3e parameters overcome CHARMM36 in structural quality for POPC but predicts overestimated ion binding~\cite{kurki22}.
-%and the complicated landscape of force field quality for membranes hamper the applicability of such simulations, resulting in substantial controversies~\cite{antila22b}. 
-%For example, previous quality evaluations of membrane simulations have concluded that 
 On the other hand, CHARMM36 gives the best description for lipid headgroup conformational ensembles~\cite{bacle21}, while GROMOS-CKP parameters best capture the membrane packing in POPS bilayers~\cite{antila22b}.
@@
 Predictions from different force fields deviate significantly in terms of absolute values and the slope of decrease upon addition of POPE.
-This can be used to select the the simulations that best caputer area per lipids for POPE and POPC lipids in Fig.~\ref{fig:POPCPOPEapls}B because the area per lipid correlates with the average order parameter of the $\textit{sn}$-1 chain (Fig.~\ref{fig:quality} F).
+This can be used to select the the simulations that best capture area per lipids for POPE and POPC lipids in Fig.~\ref{fig:POPCPOPEapls}B because the area per lipid correlates with the average order parameter of the $\textit{sn}$-1 chain (Fig.~\ref{fig:quality} F).
 %, simulations with the best predictions for area per lipids can be selected based on order parameter quality evaluation. 
@@
 These timescales were previously accessible only by coarse-grained simulations or free energy calculations~\cite{parisio16}, yet atomistic simulations reporting cholesterol flip--flop events have been published recently~\cite{gu19,javanainen19,baral20}.
-%However, systematic studies on how cholesterol flip-flop rate depends on membrane properties are still beyond the current computational resources available for a single research group. 
-These studies report an increase in cholesterol flip--flop rates with increasing acyl chain unsaturation level and decreasing cholesterol concentration~\cite{gu19,javanainen19}, but correlations between cholesterol flip--flop rates and membrane properties have not been systematically studied.
-Here, we demonstrate that the large amount of simulation data available in the NMRlipids databank, along with programmatic access providing the option to flexibly implement varied queries of the data, enables anybody to perform such analyses.
+These studies report an increase in cholesterol flip--flop rates with increasing acyl chain unsaturation level and decreasing cholesterol concentration~\cite{gu19,javanainen19}, but the amount of data in these individual studies was not sufficient to systematically assess correlations between cholesterol flip--flop rates and membrane properties.
+Here, we demonstrate that the analysis of such correlation is now accessible to all by using the API to access the large amount of MD simulation data in the NMRlipids databank as described in section~\ref{section:access}.
 
+Using workflow depicted in Fig.~\ref{fig:POPCPOPEapls}A, we first calculated the cholesterol flip--flop rates from all the simulations available in the NMRlipids databank.
 Flip--flops were observed for cholesterol, % (634 events in 77 different simulations),
@@
 The flip--flop rate of DCHOL, 0.2\,\textmu{}s$^{-1}$, was close to the average value of cholesterol, while average rates for diacylglycerols DOG and SDG were higher than for cholesterol, 0.4\,\textmu{}s$^{-1}$ and 0.5\,\textmu{}s$^{-1}$, respectively.
-Flip--flops were not observed for other lipids, giving the upper limits for PC lipid flip--flop rate of 9\,s$^{-1}$ and for ceramide (\textit{N}-palmitoyl-\textsc{D}-\textit{erythro}-sphingosine) of 0.002\,\textmu{}s$^{-1}$.
+Flip--flops were not observed for other lipids, giving the upper limits for PC lipid flip--flop rate of 9$\cdot$10$^{-6}$\,\textmu{}s$^{-1}$ and for ceramide (\textit{N}-palmitoyl-\textsc{D}-\textit{erythro}-sphingosine) of 0.002\,\textmu{}s$^{-1}$.
 Thus, available data in the NMRlipids databank suggests that the lipid flip--flip rate decreases in the order of diacylglycerols $>$ cholesterol $>$ other lipids including ceramides.
@@
 
-Nevertheless, we use the wide range of available simulations to analyse how cholesterol flip--flop rate depends on membrane properties.
+Nevertheless, we use the workflow depicted in Fig.~\ref{fig:POPCPOPEapls}B and the results for flip--flop rates from the NMRlipids databank to analyse how they depend on membrane properties.
 Cholesterol flip--flop rates and their histograms are plotted as a function of membrane thickness, lateral density and order in Figs.~\ref{fig:flip-flops} B--D.
@@
 \subsection{Extending the scope of MD simulations to new fields using the NMRlipids databank: Water diffusion anisotropy in membrane systems}
-The anisotropy in water diffusion in parallel and perpendicular directions with respect to membranes plays a role in the drug translocation through biological material, particularly in skin \cite{hansen13,wen18,nitsche19,roberts21}, and in MRI imaging~\cite{topgaard20}.
-MD simulations are rarely used to analyze anistropic diffusion of water since only few membrane permeation events for water are typically observed in a single MD simulation trajectory~\cite{venable19,camilo2022}, thereby making the collection of sufficient amount of data challenging or impossible.
-Here, we show that the NMRlipids databank can be used to systemically analyze how the anisotropic water diffusion in multilamellar membrane systems depends on membrane properties.
+The anisotropy in water diffusion in parallel and perpendicular directions with respect to membranes can be related to the translocation of drugs  through biological material, particularly in skin \cite{hansen13,wen18,nitsche19,roberts21}, and to the formation of signals in diffusion tensor MRI imaging~\cite{topgaard20}.
+MD simulations are rarely used to analyze the anisotropic diffusion of water since only few membrane permeation events for water are typically observed in a single MD simulation trajectory~\cite{venable19,camilo2022}, thereby making the collection of sufficient amount of data challenging.
+Here, we show that the API access to the data in NMRlipids databank enables systematic analyses on how the anisotropic diffusion of water depends on membrane properties in multilamellar membrane systems, thereby extending the applications of MD simulations to new fields.
 
@@
 
-To this end, we calculated the water permeability through membranes from all simulations in the NMRlipids databank.
+To this end, we first calculated the water permeability through membranes from all simulations in the NMRlipids databank using the workflow depicted in Fig.~\ref{fig:POPCPOPEapls}A.
 The resulting non-zero values range between 0.3 and 322\,\textmu{}m/s with the mean and median of 14\,\textmu{}m/s and 8\,\textmu{}m/s, respectively.
 These values agree with the previously reported simulation results~\cite{venable19,camilo2022} and have the same order of magnitude as experimentally determined diffusive permeability coefficients, but are on average larger than the values reported for PC lipids in the liquid crystalline phase,~0.19--0.33\,\textmu{}m/s~\cite{jansen95}.
-Observed permeabilities and their histogrammed values are shown in Figs.~\ref{fig:permeability} B--E as a function of temperature, membrane thickness, area per lipid, and acyl chain order.
+Using the workflow depicted in Fig.~\ref{fig:POPCPOPEapls}b, we then plotted the observed permeabilities and their histogrammed values in Figs.~\ref{fig:permeability} B--E as a function of temperature, membrane thickness, area per lipid, and acyl chain order.
 As expected, the permeability increases with the temperature, giving an average energy barrier of 17\,$\pm$\,4\,$k_\mathrm{B}T$ for the water permeation from the Arrhenius plot in Fig.~\ref{fig:permeability} B.
@@
 
-To analyze how water diffusion anisotropy depends on membrane properties in a multi-lamellar lipid bilayer system, we calculated the water diffusion parallel to the membrane surface from all simulations in the NMRlipids databank.
-The parallel diffusion coefficient of water, $D_{||}$, decreases with reduced hydration and increases with the temperature, but dependencies on membrane area per lipid, thickness, or fraction of charged lipids were not observed in Figs.~\ref{fig:permeability} and~\ref{fig:diffusionSI}.
+To analyze how water diffusion anisotropy depends on membrane properties in a multi-lamellar lipid bilayer system, we calculated the water diffusion parallel to the membrane surface from all simulations in the NMRlipids databank using the workflow depicted in Fig.~\ref{fig:POPCPOPEapls}A.
+The parallel diffusion coefficient of water, $D_{||}$, decreases with reduced hydration and increases with the temperature, but dependencies on membrane area per lipid, thickness, or fraction of charged lipids were not observed in Figs.~\ref{fig:permeability} and~\ref{fig:diffusionSI} when the workflow depicted in Fig.~\ref{fig:POPCPOPEapls}B was used.
 Simulation results are close to the experimental values with low hydration levels in Fig.~\ref{fig:permeability} F, but increase to approximately two times higher than the experimental value for bulk water diffusion value (3.1$\cdot$\,10$^{-9}$\,m$^2$/s at 313\,K~\cite{khakimov08}) with high hydration levels.
@@
 Several folds larger anisotropy in membranes with higher lateral density is expected to play a role in pharmacokinetic models not only for water but also for other hydrophilic molecules~\cite{nitsche19}.
-Furthermore, the enhanced understanding of this anisotropy may help in developing new (functional) MRI imaging methods~\cite{topgaard20}.
+Furthermore, the enhanced understanding of this anisotropy may help in developing new diffusion tensor based MRI imaging methods where signals originate from anisotropic diffusion of water in biological material~\cite{topgaard20}.
 
@@
 \subsection{Calculation of X-ray scattering form factors}
-X-ray scattering form factors were calculated with standard equation for symmetric lipid bilayers~\cite{ollila16}
+X-ray scattering form factors were calculated with the standard equation for symmetric lipid bilayers~\cite{ollila16}
 \begin{equation}\label{FFsimpl}
@@
 where $\Delta \rho_e(z)$ is the difference between total and solvent electron densities, and $D$ is the simulation box size in the $z$-direction (normal to the membrane).
-For the calculation of density profiles, atom coordinates were first centred around the centre of mass of lipid molecules for every time frame, and a histogram of these centred positions weighted with the number of electrons in each atom was then calculated with the bin width of $1/3$~\AA{}.
+For the calculation of density profiles, atom coordinates were first centred around the centre of mass of lipid molecules for every time frame, and a histogram of these centred positions, weighted with the number of electrons in each atom, was then calculated with the bin width of $1/3$~\AA{}.
+Electron density profiles were then calculated as an average of these histograms over the time frames in simulations.
 The script to calculate form factors for all simulations in the NMRlipids databank is available at \texttt{Scripts/AnalyzeDatabank/calc\_FormFactors.py}.
@@
 \subsection{Calculation of a bilayer area per lipid and thickness}
-Area per lipids of bilayers were calculated by dividing the time-averaged area of the simulation box with the total number of lipid molecules in the simulation.
+Area per lipids of bilayers were calculated by dividing the time-averaged area of the simulation box with the total number of lipid and surfactant molecules (see the list in table~\ref{tab:abbreviations}) in the simulation.
 The script that calculates area per lipids from all simulations in the NMRlipids databank repository is available at \texttt{Scripts/AnalyzeDatabank/calcAPL.py} in the NMRlipids databank repository.
@@
 The resulting thicknesses are stored in files named \texttt{thickness.json} at folders in \texttt{/Data/Simulations} in the NMRlipids databank repository.
+
+\subsection{PCA analysis of equilibration of simulations}
 
@@
 
-To streamline the comparison between simulations, we define the qualities of different fragments (headgroup, acyl chains or total lipid) within each lipid type in a simulation as
+To streamline the comparison between simulations, we define the qualities for different fragments (frag) within each lipid type in a simulation as
 \begin{equation}
-    P^\mathrm{frag}[\mathrm{lipid}] = \frac{\langle P[\mathrm{lipid}]\rangle_\mathrm{frag}}{p_\mathrm{frag}[\mathrm{lipid}]},
+    P^\mathrm{frag}[\mathrm{lipid}] = \langle P[\mathrm{lipid}]\rangle_\mathrm{frag} F_\mathrm{frag}[\mathrm{lipid}],
 \end{equation}
-where $\langle P[\mathrm{lipid}]\rangle_\mathrm{frag}$ is the average of individual C-H bond order parameters qualities within the fragment, $p_\mathrm{frag}[\mathrm{lipid}]$ is the percentage of order parameters for which the quality is available within the fragment, and frag can be \textit{sn}-1, \textit{sn}-2, headgroup or total (all order parameters within a molecule).
+where $\langle P[\mathrm{lipid}]\rangle_\mathrm{frag}$ is the average of individual C-H bond order parameters qualities within the fragment, $F_\mathrm{frag}[\mathrm{lipid}]$ is the percentage of order parameters for which the quality is available within the fragment, and frag can be \textit{sn}-1, \textit{sn}-2, headgroup or total (all order parameters within a molecule).
 The overall quality of different fragments in a simulation are then defined as a molar fraction weighted average over different lipid components

### 2023-01-17 SI.tex
Update text
Update pdfs

@@
     \hline
+    Lipids and surfactants & \\
+    \hline
     POPC &  1-palmitoyl-2-oleoyl-sn-glycero-3-phosphocholine  \\
@@
     \hline
+    Other molecules & \\
+    \hline
     POT & potassium ion  \\
@@
 \subsection{NMR experiments}
-Acyl chain order parameters of POPE and POPG were analyzed from the experiments that were used to determine headgroup order parameters in Ref.~\citenum{bacle21}.
+Acyl chain order parameters of POPE and POPG were analyzed from the experiments that were previosly used to determine headgroup order parameters~\cite{bacle21}.
 The analysis of the crowded spectral region at 29-31 ppm was based on the previous assignment reported for POPC membranes~\cite{ferreira13}.

### 2023-01-26 SI.tex
Update text

@@
 
-\begin{table}[!h]
+\begin{table}[p]
     \centering
@@
 
-\begin{table}[!h]
+\begin{table}[h]
+    \centering
+    \begin{tabular}{  p{5.0cm}  p{10.0cm}}
+    \toprule
+    key & description \\
+    \midrule
+    DOI & DOI of the publication related to the experimental data. \\
+    TEMPERATURE & Temperature of the experiment. \\
+    MOLAR\_FRACTIONS & Dictionary of molar fractions of bilayer components \\
+    ION\_CONCENTRATIONS & Dictionary of ion concentrations of the system \\
+    % (defined as ??) \\
+    TOTAL\_LIPID\_CONCENTRATION & Total concentration of lipid components. If exact concentration is not known, but experiments are performed in excess water, 'full hydration' can be given. \\
+    COUNTER\_IONS & Type of counter ions if present.
+\end{tabular}
+    \caption{Keys stored in the README.yaml files of experiments.}
+    \label{tab:READMEkeysEXP}
+\end{table}
+
+\begin{table}[p]
     \centering
@@
 
-\begin{table}[!h]
-    \centering
-    \begin{tabular}{  p{5.0cm}  p{10.0cm}}
-    \toprule
-    key & description \\
-    \midrule
-    DOI & DOI of the publication related to the experimental data. \\
-    TEMPERATURE & Temperature of the experiment. \\
-    MOLAR\_FRACTIONS & Dictionary of molar fractions of bilayer components \\
-    ION\_CONCENTRATIONS & Dictionary of ion concentrations of the system \\% (defined as ??) \\
-    TOTAL\_LIPID\_CONCENTRATION & Total concentration of lipid components. If exact concentration is not known, but experiments are performed in excess water, 'full hydration' can be given. \\
-    COUNTER\_IONS & Type of counter ions if present.
-\end{tabular}
-    \caption{Keys stored in the README.yaml files of experiments.}
-    \label{tab:READMEkeysEXP}
-\end{table}
-
-\newpage
-\section{Experimental data}
-\subsection{NMR experiments}
+\clearpage
+%\section{Experimental data}
+\section{NMR experiments}
 Acyl chain order parameters of POPE and POPG were analyzed from the experiments that were previosly used to determine headgroup order parameters~\cite{bacle21}.
@@
 
-\begin{figure}
+\begin{figure}[h]
   \includegraphics[width=\textwidth]{Figures/POPE_INEPT_contour_SCH.pdf}
@@
 (A) $^{13}$C rINEPT spectrum with peak assignment.
-The labels used are shown in the chemical structure of POPE depicted.
+The labels used are shown in the chemical structure of POPE.
 The chemical shift of the methyl groups was defined as 13.8 ppm.
@@
 (C) C-H order parameter profile for the acyl chains of POPE.
-The splittings used for calculating the order parameters are shown in the next figure.
+The splittings used for calculating the order parameters are shown in Fig.~\ref{POPEexp2}.
 The unassigned peaks belong to the headgroup and glycerol backbone carbons.
@@
 
-\begin{figure}
+\begin{figure}[p]
   \includegraphics[width=\textwidth]{Figures/POPE_slices.pdf}
-  \caption{Dipolar spectra obtained from the 2D R-PDLF spectrum described in the previous figure.
+  \caption{Dipolar spectra obtained from the 2D R-PDLF spectrum from POPE in Fig.~\ref{POPEexp1}.
 The number at the top left of each plot denotes the corresponding chemical shift.
 The carbon label for each splitting is displayed on the top right.
-The labels are the same as shown in the previous figure.
+The labels are the same as in Fig.~\ref{POPEexp1}.
 }
-  \label{POPEexp1}
+  \label{POPEexp2}
 \end{figure}
 
-\begin{figure}
+\begin{figure}[p]
   \includegraphics[width=\textwidth]{Figures/POPG_INEPT_contour_SCH.pdf}
@@
 (A) $^{13}$C rINEPT spectrum with peak assignment.
-The labels used are shown in the chemical structure of POPE depicted.
+The labels used are shown in the chemical structure of POPG.
 The chemical shift of the methyl groups was defined as 13.8 ppm.
@@
 The assignment was based on a previous assignment reported for POPC membranes~\cite{ferreira13}.
-(C) C-H order parameter profile for the acyl chains of POPE.
-The splittings used for calculating the order parameters are shown in the next figure.
+(C) C-H order parameter profile for the acyl chains of POPG.
+The splittings used for calculating the order parameters are shown in Fig.~\ref{POPGexp2}.
 The unassigned peaks belong to the headgroup and glycerol backbone carbons.
 A detailed assignment and order parameter analysis of these carbons was shown previously~\cite{bacle21}.}
-  \label{POPEexp1}
+  \label{POPGexp1}
 \end{figure}
 
-\begin{figure}
+\begin{figure}[p]
   \includegraphics[width=\textwidth]{Figures/POPG_slices.pdf}
-  \caption{Dipolar spectra obtained from the 2D R-PDLF spectrum described in the previous figure.
+  \caption{Dipolar spectra obtained from the 2D R-PDLF spectrum described in Fig.~\ref{POPGexp1}.
 The number at the top left of each plot denotes the corresponding chemical shift.
 The carbon label for each splitting is displayed on the top right.
-The labels are the same as shown in the previous figure.
+The labels are the same as in Fig.~\ref{POPGexp1}.
 }
-  \label{POPEexp1}
+  \label{POPGexp2}
 \end{figure}
 
-\pagebreak
-\subsection{X-ray scattering experiments}
-X-ray scattering form factors contributed to the NMRlipids III project are included in the databank (\url{http://nmrlipids.blogspot.com/2022/09/nmrlipids-iii-including-lipid-lateral.html}).
+%\pagebreak
+%\subsection{X-ray scattering experiments}
+%X-ray scattering form factors contributed to the NMRlipids III project are included in the databank (\url{http://nmrlipids.blogspot.com/2022/09/nmrlipids-iii-including-lipid-lateral.html}).
 

### 2023-01-26 main.tex
Update text

@@
 \begin{abstract}
-We present a databank of lipid bilayer simulations from the NMRlipids open collaboration project.
+Cellular membrane lipid composition is implicated in diseases and major biological functions in cells, but membranes are difficult to study experimentally due to their intrinsic disorder and complex phase behaviour.
+Molecular dynamics (MD) simulations have been useful in understanding membrane systems, but they require significant computational resources and often suffer inaccuracies in model parameters.
+Applications of data-driven and machine learning methods, currently revolutionising many fields, are limited to membrane systems due to the lack of suitable training sets.
+Here we present the NMRlipids Databank---a community-driven, open-for-all database featuring programmatic access to quality evaluated atom-resolution MD simulations of lipid bilayers.
+The NMRlipids databank will benefit scientists in different disciplines by provoding automatic ranking of simulations based on their quality against experiments, flexible implementations of data-driven and machine learning applications, and rapid access to simulation data via graphical user interface.
+To demonstrate the unlocked possibilities beyond current MD simulation studies, we analyzed how anisotropic diffusion of water and cholesterol flip–flop rates depend on membrane properties.
+
 \end{abstract}
@@
     \includegraphics[width=\linewidth]{Figures/overlay2.pdf}
-    \caption{a) Structure of an overlay databank. 
-    %Raw data is available from a publicly accessible repository (layer 1).
-    %The core of the databank contains information on the location of the raw data and indexing of molecular %compositions and simulation details described with universal naming conventions (layer 2).
-    %The content of databank can be viewed and analysed using external programs that automatically read the information and data from layers 1 and 2 (layer 3).
-Examples of such applications are the GitHub repository generating the results presented in this work and the graphical access to the databank.
-    A more detailed structure of the databank layer in the NMRlipids databank is illustrated in Fig.~\ref{DatabankStructure} in the SI.
+    \caption{a) Schematic presentation of the overlay structure used in the NMRlipids databank. A more detailed structure of the databank layer is shown in Fig.~\ref{DatabankStructure} in the SI.
     b) Distribution of the lengths of the trajectories, total number of trajectories and total length of the simulations in the NMRlipids databank.
@@
     \includegraphics[width=88mm]{Figures/flowchart2.pdf}
-    \caption{ A) Flowchart for analysing properties from all MD simulations in the NMRlipids databank.
-    B) Flowchart for accessing the results analyzed from the NMRlipids databank.
-    C) Area per lipids of POPC and POPE lipid bilayers predicted by different force fields at 310\,K.
+    \caption{ A) Flowchart for performing an analysis of properties through all MD simulations in the NMRlipids databank using the API.
+    B) Flowchart for accessing results calculated from the NMRlipids databank and stored to the application layer.
+    C) Area per lipids of POPC and POPE lipid bilayers predicted by different force fields at 310\,K in simulations that are available in the NMRlipids databank.
 The points from the best performing simulations in the NMRlipids databank quality evaluation for each lipid (Figs.~\ref{fig:quality} and~\ref{fig:POPC_POPE_dataSI}) are surrounded by black circles.
@@
 
-\subsection{Selection of best simulation parameters for applications using the NMRlipids databank}
+\subsection{Applications of the NMRlipids databank}
+\subsubsection{Selection of simulation parameters using the NMRlipids databank: Best models for most abundant neutral membrane lipids}
 %Quality of membrane simulations with different force fields have been evaluated against experimental data during parameterization and in separate comparison studies~\cite{botan15,ollila16,catte16,pluhackova16,perez17,leonard19,??}, but the lack of universal quality measure for membrane simulations complicates the estimations of reliability of simulations.
-Understanding of how closely simulations resemble the reality is crucial when assessing the reliability and significance of MD simulation results, as well as when using atomistic simulations in parametrization of coarse grained models.
-However, universal quality measures for MD simulations are lacking, and a prevailing force field that would perform best in all situations has not been found for membranes.
-For example, OPLS3e parameters overcome CHARMM36 in structural quality for POPC but predicts overestimated ion binding~\cite{kurki22}.
+Understanding of how closely simulations resemble the reality is crucial for selecting best models for specific applications, assessing the reliability and significance of MD simulation results, and using atomistic simulations in parametrization of coarse grained models.
+However, undefined universal quality measures for MD simulations and the lack of a prevailing force field that would perform best in all simulations has led to many controversial results in membrane simulations~\cite{antila22b}.
+%For example, OPLS3e parameters overcome CHARMM36 in structural quality for POPC but predicts overestimated ion binding~\cite{kurki22}.
 On the other hand, CHARMM36 gives the best description for lipid headgroup conformational ensembles~\cite{bacle21}, while GROMOS-CKP parameters best capture the membrane packing in POPS bilayers~\cite{antila22b}.
-Furthermore, Fig.~\ref{fig:POPCPOPEapls}B illustrates divergent predictions for the difference in packing between POPC and POPE membranes from different force fields.
-To streamline the selection of best simulations for specific applications, and to guide the development of atomistic and coarse grained parameters, quantitative quality measures for lipid bilayer simulations are defined in the NMRlipids databank.
-Together with the API, these will enable rapid and automatic quality evaluation of simulations in the NMRlipids databank, thereby promoting the selection of best simulations for specific applications without laborious force field evaluation, and automatic parametrization procedures for atomistic and coarse grained simulations.
-%that can be used to rapidly find the best available parameters .
-
-The quality for each simulation in the NMRlipids databank with the corresponding experimental data available is currently done using C--H bond order parameters from NMR experiments and form factors from X-ray scattering experiments.
+For example, Fig.~\ref{fig:POPCPOPEapls}B illustrates divergent predictions from different force fields for the lateral packing of POPC and POPE membranes that are the two most biologically abundant neutral membrane lipids~\cite{vanmeer08}.
+We have previously resolved such controversies for membrane structures and ion binding affinities for neutral~\cite{botan15,catte16} and charged lipids~\cite{antila19,bacle21} by comparing simulations with C-H bond order parameters from NMR, yet this requires manual comparisons between large number of simulations.
+To streamline this process,
+%the selection of best simulations for specific applications, and to guide the development of atomistic and coarse grained parameters,
+we have defined quantitative quality measures, based on C--H bond order parameters from NMR and X-ray scattering form factors, for lipid bilayer simulations in the NMRlipids databank. 
+%The quality for each simulation in the NMRlipids databank with the corresponding experimental data available is currently done using C--H bond order parameters from NMR experiments and form factors from X-ray scattering experiments.
 The C--H bond order parameters are sensitive to conformational ensembles of individual lipid molecules~\cite{ollila16}, while correlating well also with the membrane thickness and lateral packing as shown in Fig.~\ref{fig:quality}F.
@@
 %For detailed definitions of quality measures, see the supplementary information.
+Together with the API, these enable rapid and automatic quality evaluation of simulations in the NMRlipids databank, thereby promoting the selection of best simulations for specific applications without laborious force field evaluation, and automatic parametrization procedures for atomistic and coarse grained simulations.
 
-Ranking based on C--H bond order parameter quality for the best performing simulations is shown in Fig.~\ref{fig:quality}A, and for the best performing POPE lipid bilayer simulations in Fig.~\ref{fig:quality}B.
-Direct comparisons with experiments are shown for selected simulations in Figs.~\ref{fig:quality}C--E.
+To select the best models for POPC and POPE among simulations compared in Fig.~\ref{fig:POPCPOPEapls}C, we show the ranking based on C--H bond order parameter quality in Fig.~\ref{fig:quality}A for the best performing simulations in the NMRlipids databank, and for the best performing POPE lipid bilayer simulations in Fig.~\ref{fig:quality}B.
+Direct comparisons with experiments are shown for selected simulations in Figs.~\ref{fig:quality}C--E and~\ref{fig:POPC_POPE_dataSI}.
 %The power of the NMRlipids databank in selecting the best simulation parameters for a specific application is demonstrated for mixtures of PE and PC lipids in Fig.~\ref{fig:POPCPOPEapls} showing area per lipids from available POPC and POPE bilayer simulations in the databank from different force fields at 310\,K.
 Predictions from different force fields deviate significantly in terms of absolute values and the slope of decrease upon addition of POPE.
-This can be used to select the the simulations that best capture area per lipids for POPE and POPC lipids in Fig.~\ref{fig:POPCPOPEapls}B because the area per lipid correlates with the average order parameter of the $\textit{sn}$-1 chain (Fig.~\ref{fig:quality} F).
-%, simulations with the best predictions for area per lipids can be selected based on order parameter quality evaluation. 
-Quality evaluation in Fig.~\ref{fig:quality} and direct comparison in Fig.~\ref{fig:POPC_POPE_dataSI} reveal that Slipids ranks 1\textsuperscript{st} for POPC with a clear difference to others, and 2\textsuperscript{nd} for POPE with only marginally lower quality than GROMOS-CKP.
-Simulations with CHARMM36 and GROMOS-CKP predict overly packed bilayer for POPC (overestimated order in Fig.~\ref{fig:POPC_POPE_dataSI}), while the area per lipid for POPC is overestimated in Lipid17.
+Because the area per lipid correlates with the average order parameter of the $\textit{sn}$-1 chain (Fig.~\ref{fig:quality} F), quality evaluation in Fig.~\ref{fig:quality} and direct comparison in Fig.~\ref{fig:POPC_POPE_dataSI} can be used to select the simulations that best capture area per lipids for POPE and POPC lipids in Fig.~\ref{fig:POPCPOPEapls}C.
+In these results, simulations with Slipids parameters ranks 1\textsuperscript{st} for POPC with a clear difference to others, and 2\textsuperscript{nd} for POPE with only marginally lower quality than GROMOS-CKP.
+Simulations with CHARMM36 and GROMOS-CKP parameters predict overly packed bilayer for POPC (overestimated order in Fig.~\ref{fig:POPC_POPE_dataSI}), while the area per lipid for POPC is overestimated in Lipid17.
 For POPE, CHARMM36 and Lipid17 predict membranes that are too packed.
-In conclusion, Slipids parameters is the best available choice for studies where packing effects of PE lipids are relevant
+In conclusion, the quality evalution based on the NMRlipids databank suggests that Slipids parameters are the best available choice for simulations with PC and PE lipids, at least for the cases were membrane packing is relevant.
 
@@
 
-\subsection{Detecting rare phenomena using NMRlipids databank: Cholesterol flip--flops}
+\subsubsection{Detecting rare phenomena using NMRlipids databank: Cholesterol flip--flops}
 Lipid flip--flops from one bilayer leaflet to another play an important role in lipid trafficking and regulating membrane properties~\cite{vanmeer08}.
@@
 Still, the reported timescales range from minutes to sub-millisecods~\cite{vanmeer08,steck12,parisio16,gu19}.
-These timescales were previously accessible only by coarse-grained simulations or free energy calculations~\cite{parisio16}, yet atomistic simulations reporting cholesterol flip--flop events have been published recently~\cite{gu19,javanainen19,baral20}.
+These timescales were previously accessible only by coarse-grained simulations or free energy calculations~\cite{parisio16}, yet atomistic simulations reporting cholesterol flip--flop events have been published only recently~\cite{gu19,javanainen19,baral20}.
 These studies report an increase in cholesterol flip--flop rates with increasing acyl chain unsaturation level and decreasing cholesterol concentration~\cite{gu19,javanainen19}, but the amount of data in these individual studies was not sufficient to systematically assess correlations between cholesterol flip--flop rates and membrane properties.
-Here, we demonstrate that the analysis of such correlation is now accessible to all by using the API to access the large amount of MD simulation data in the NMRlipids databank as described in section~\ref{section:access}.
+Here, we demonstrate that the NMRlipids databank API makes analyses of such rare phenomena accessible to everyone by enabling the access to large amount of MD simulation data as described in section~\ref{section:access}.
+This is particularly useful for scientists in different fields of science and industry who do not have the access to required computational resources and expertise to produce the large amounts of MD simulation data.
 
@@
 DOG (1,2-dioleoyl-\textit{sn}-glycerol), %, 44 flip-flops in 4 simulations) 
-and SDG (1-stearoyl-2-docosahexaenoyl-\textit{sn}-glycerol) in the NMRlipids databank.
-Cholesterol flip--flop rates range between 0.001--1.6\,\textmu{}s$^{-1}$ with the mean of 0.16\,\textmu{}s$^{-1}$ and median of 0.07\,\textmu{}s$^{-1}$.
-These values are in line with the previously reported values from atomistic MD simulations~\cite{gu19,javanainen19,baral20}.
+and SDG (1-stearoyl-2-docosahexaenoyl-\textit{sn}-glycerol).
+The observed cholesterol flip--flop rates, ranging between 0.001--1.6\,\textmu{}s$^{-1}$ with the mean of 0.16\,\textmu{}s$^{-1}$ and median of 0.07\,\textmu{}s$^{-1}$, are in line with the previously reported values from atomistic MD simulations~\cite{gu19,javanainen19,baral20}.
 The flip--flop rate of DCHOL, 0.2\,\textmu{}s$^{-1}$, was close to the average value of cholesterol, while average rates for diacylglycerols DOG and SDG were higher than for cholesterol, 0.4\,\textmu{}s$^{-1}$ and 0.5\,\textmu{}s$^{-1}$, respectively.
@@
 
-Nevertheless, we use the workflow depicted in Fig.~\ref{fig:POPCPOPEapls}B and the results for flip--flop rates from the NMRlipids databank to analyse how they depend on membrane properties.
+Nevertheless, we used the workflow depicted in Fig.~\ref{fig:POPCPOPEapls}B to analyse how flip--flop rates calculated from the NMRlipids databank depend on membrane properties.
 Cholesterol flip--flop rates and their histograms are plotted as a function of membrane thickness, lateral density and order in Figs.~\ref{fig:flip-flops} B--D.
-The results reveal a non-linear correlation between cholesterol flip--flop rate and membrane packing.
+The results reveal a non-linear correlation between cholesterol flip--flop rate and membrane packing which is depicted as area per lipid.
 Flip--flop rates increase by an order of magnitude when membrane packing density decreases and a major jump is observed at low membrane packing.
@@
     \includegraphics[width=88mm]{Figures/CholFlipFlops.pdf}
-    \caption{\textbf{A} Illustration of cholesterol flip-flop.  
+    \caption{\textbf{A} Illustration of cholesterol flip--flop.  
       \textbf{B--D} Cholesterol flip--flops analyzed from the databank as a function of membrane thickness, area per lipid, and acyl chain order.
@@
 
-Explorations of cholesterol flip--flops using MD simulations have only recently become possible for research groups with substantial resources and specialized expertise~\cite{gu19,javanainen19,baral20}.
-Our analysis demonstrates how the NMRlipids databank makes analyses of such rare events accessible to wide range of scientists in different fields of science and industry who do not have the access to required resources and expertise to produce the large amounts of MD simulation data required.
-
-\subsection{Extending the scope of MD simulations to new fields using the NMRlipids databank: Water diffusion anisotropy in membrane systems}
+\subsubsection{Extending the scope of MD simulations to new fields using the NMRlipids databank: Water diffusion anisotropy in membrane systems}
 The anisotropy in water diffusion in parallel and perpendicular directions with respect to membranes can be related to the translocation of drugs  through biological material, particularly in skin \cite{hansen13,wen18,nitsche19,roberts21}, and to the formation of signals in diffusion tensor MRI imaging~\cite{topgaard20}.
@@
 The resulting non-zero values range between 0.3 and 322\,\textmu{}m/s with the mean and median of 14\,\textmu{}m/s and 8\,\textmu{}m/s, respectively.
-These values agree with the previously reported simulation results~\cite{venable19,camilo2022} and have the same order of magnitude as experimentally determined diffusive permeability coefficients, but are on average larger than the values reported for PC lipids in the liquid crystalline phase,~0.19--0.33\,\textmu{}m/s~\cite{jansen95}.
-Using the workflow depicted in Fig.~\ref{fig:POPCPOPEapls}b, we then plotted the observed permeabilities and their histogrammed values in Figs.~\ref{fig:permeability} B--E as a function of temperature, membrane thickness, area per lipid, and acyl chain order.
-As expected, the permeability increases with the temperature, giving an average energy barrier of 17\,$\pm$\,4\,$k_\mathrm{B}T$ for the water permeation from the Arrhenius plot in Fig.~\ref{fig:permeability} B.
+These values agree with the previously reported simulation results~\cite{venable19,camilo2022}, but are on average slightly larger than experimental values reported for PC lipids in the liquid crystalline phase,~0.19--0.33\,\textmu{}m/s~\cite{jansen95}.
+Using the workflow depicted in Fig.~\ref{fig:POPCPOPEapls}B, we then plotted the observed permeabilities and their histogrammed values in Figs.~\ref{fig:permeability}B--E as a function of temperature, membrane thickness, area per lipid, and acyl chain order.
+As expected, the permeability increases with the temperature, giving an average energy barrier of 17\,$\pm$\,4\,$k_\mathrm{B}T$ for the water permeation from the Arrhenius plot in Fig.~\ref{fig:permeability}B.
 On the other hand, the permeability of water decreases on average when membranes become more packed, \textit{i.e.}, with decreasing area per lipid and increasing thickness and acyl chain order (Figs.~\ref{fig:permeability} C--E).
-Permeation of water through bilayers also depends on membrane properties also according previous studies, but there is no established consensus on whether the area per lipid~\cite{nagle08} or bilayer thickness~\cite{frallicciardi22} is the main parameter determining the permeability.
+Permeation of water through bilayers depends on membrane properties also according to previous studies, but there is no established consensus on whether the area per lipid~\cite{nagle08} or bilayer thickness~\cite{frallicciardi22} is the main parameter determining the permeability.
 %is previously reported from simulations and experiments for some systems, but not for all~\cite{mathai07,venable19,frallicciardi22}. 
@@
 Simultaneously, machine learning based models predicting behaviour of biomolecules and automatic approaches to parametrize models are emerging~\cite{jumper21,antila22b}.
-The NMRlipids databank will support the development in all these directions.
+The NMRlipids databank will support the development in both of these directions.
 The automatic quality evaluation and ranking of simulations in the NMRlipids databank guides the optimization of new force field parameters and the selection of best parameters for large biomolecular complexes where simulation quality becomes increasingly important due the accumulation of small errors in large simulations.
@@
 
-Providing programmatic access to large scale MD simulation data in the NMRlipids databank can lead to applications in unprecedented directions in fields where MD simulations are less commonly utilized.
-Here, we exemplified this by analysing the anisotropic diffusion of water (Fig.~\ref{fig:permeability}), which is relevant in pharmacokinetic modeling and in MRI imaging, both areas where MD simulations are not yet commonly used~\cite{nitsche19,topgaard20}.
 %The NMRlipids databank will be of great benefit to scientists involved in MD simulations.
@@
 The selection of best models for a specific application is demonstrated for PC/PE lipid mixtures in Fig.~\ref{fig:quality}, and previously for PC/PS lipid mixtures~\cite{antila22b}, and lipid headgroups~\cite{bacle21}.
-The programmatic access in the NMRlipids databank enables automatic analyses over larger sets of simulation data in terms of quantity (\textit{e.g.}, simulation length and number of conformations) and content (\textit{e.g.}, lipid compositions and ion concentrations) than is currently possible by a single research group.
-The added value that access to this simulation data provides is demonstrated here for the analysis of simulation qualities (Fig.~\ref{fig:quality}), water permeation through membranes (Fig.~\ref{fig:permeability}) and cholesterol flip--flop events (Fig.~\ref{fig:flip-flops}).
-%Such models could be used to connect membrane composition to relevant membrane properties in wide range of fields from molecular biology to drug development and biomolecular imaging. 
+The programmatic access in the NMRlipids databank enables automatic analyses of structural and dynamic properties over larger sets of simulation data in terms of quantity (\textit{e.g.}, simulation length and number of conformations) and content (\textit{e.g.}, lipid compositions and ion concentrations) than is currently possible by a single research group.
+The added value of such analyses is demonstrated here by analysing the anisotropic diffusion of water in membrane systems (Fig.~\ref{fig:permeability}) and cholesterol flip--flop events (Fig.~\ref{fig:flip-flops}).
+By enabling these analyses for everyone, the NMRlipids databank can lead to applications in unprecedented directions in fields where MD simulations are less commonly utilized, such as pharmacokinetic modeling and in MRI imaging where anisotropic diffusion of water play important roles~\cite{nitsche19,topgaard20}.
 Different types of applications enabled by the NMRlipids databank across a wide range of fields are exemplified in Table~\ref{tab:applications}.
-The ever growing amount of data is expected to further increase the scope of potential applications of the NMRlipids databank in biology, biotechnology,material science, biomolecular imaging and beyond.
+The ever growing amount of data is expected to further increase the scope of potential applications of the NMRlipids databank in biology, biotechnology, material science, biomolecular imaging and beyond.
 
@@
     \hline
-    Analyses of rare phenomena               & Lipid flip--flops, water permeation & Membrane scientists \\
+    Analyses of rare phenomena               & Lipid flip--flops, water permeation (Figs.~\ref{fig:flip-flops} and~\ref{fig:permeability}) & Membrane scientists \\
     \\
@@
     Applications that are outside typical scope of MD simulations & 
-    Anisotropic water diffusion for pharmacokinetics and MRI imaging applications & 
+    Anisotropic water diffusion for pharmacokinetics and MRI imaging applications (Fig.~\ref{fig:permeability}) & 
     Scientist in fields where MD simulations are not usually applied \\
@@
     Training set for machine learning applicatons &
-    programmatic access to the data and results enables training of machine learning type of models for various applications, such as predictions of membrane properties from composition & Scientists building and using machine learning applications for biomolecules.
+    Programmatic access to the data and results enables training of machine learning type of models for various applications, such as predictions of membrane properties from composition & Scientists building and using machine learning applications for biomolecules.
     \end{tabular}
@@
 The quality of each C--H bond order parameter is estimated by calculating the probability for a simulated value to locate within the error bars of the experimental value.
-Because conformational ensembles of individual lipids are assumed to be independent in a fluid lipid bilayer, $\frac{S_\mathrm{CH}-\mu}{s/\sqrt{n}}$ has a Student's t-distribution with $n-1$ degrees of freedom, and the probability for an order parameter from simulation to locate within experimental error bars can be estimated from equation
+Because conformational ensembles of individual lipids are assumed to be independent in a fluid lipid bilayer, $\frac{S_\mathrm{CH}-\mu}{s/\sqrt{n}}$ has a Student's t-distribution with $n-1$ degrees of freedom and $\mu$ representing the real mean of the order parameter.
+The probability for an order parameter from simulation to locate within experimental error bars can be estimated from equation
 \begin{equation}\label{probability}
@@
 where $f(t)$ is the %first order 
-Student's t-distribution, $\mu$ is the real mean of the order parameter, $n$ is the number of independent sample points for each C--H bond which equals the number of lipids in a simulation, $S_\mathrm{CH}$ is the sample mean from Eq.~\ref{OP}, $s$ is the variance of $S_\mathrm {CH}$ calculated over individual lipids, $S_\mathrm{exp}$ is the experimental value, and $\Delta S_\mathrm{exp}$ its error.
+Student's t-distribution, $n$ is the number of independent sample points for each C--H bond which equals the number of lipids in a simulation, $S_\mathrm{CH}$ is the sample mean from Eq.~\ref{OP}, $s$ is the variance of $S_\mathrm {CH}$ calculated over individual lipids, $S_\mathrm{exp}$ is the experimental value, and $\Delta S_\mathrm{exp}$ its error.
 The error of $\Delta S_\mathrm{exp} = 0.02$ is currently assumed for all experimental order parameters~\cite{ollila16}, yet more accurate ones may be available in the future~\cite{wurl22}.
@@
 
-To streamline the comparison between simulations, we define the qualities for different fragments (frag) within each lipid type in a simulation as
+To streamline the comparison between simulations, we define the qualities for different fragments (frag = \textit{sn}-1, \textit{sn}-2, headgroup or total referring to all order parameters within a molecule) within each lipid type in a simulation as
 \begin{equation}
@@
 \end{equation}
-where $\langle P[\mathrm{lipid}]\rangle_\mathrm{frag}$ is the average of individual C-H bond order parameters qualities within the fragment, $F_\mathrm{frag}[\mathrm{lipid}]$ is the percentage of order parameters for which the quality is available within the fragment, and frag can be \textit{sn}-1, \textit{sn}-2, headgroup or total (all order parameters within a molecule).
+where $\langle P[\mathrm{lipid}]\rangle_\mathrm{frag}$ is the average of individual C-H bond order parameters qualities within the fragment, and $F_\mathrm{frag}[\mathrm{lipid}]$ is the percentage of order parameters for which the quality is available within the fragment.
+%, and frag can be \textit{sn}-1, \textit{sn}-2, headgroup or total (all order parameters within a molecule). 
 The overall quality of different fragments in a simulation are then defined as a molar fraction weighted average over different lipid components

### 2023-01-27 SI.tex
Updates from Overleaf

@@
 \section{NMR experiments}
-Acyl chain order parameters of POPE and POPG were analyzed from the experiments that were previosly used to determine headgroup order parameters~\cite{bacle21}.
+Acyl chain order parameters of POPE and POPG were analyzed from the experiments that were previously used to determine headgroup order parameters~\cite{bacle21}.
 The analysis of the crowded spectral region at 29-31 ppm was based on the previous assignment reported for POPC membranes~\cite{ferreira13}.

### 2023-01-30 main.tex
Updates from Overleaf

@@
 \author[2]{Hanne Antila}          %ORCID: 0000-0002-2474-5053
-\author[2]{Lara Bort}             %ORCID: ?
+\author[14]{Lara Bort}             %ORCID: ?
 \author[3]{Pavel Buslaev}         %ORCID: ?
@@
 \author[2]{Batuhan Kav}           %ORCID: ?
+\author[13]{Patrik Kula}
 \author[2]{Milla Kurki}
 \author[2]{Alexander Kuzmin}
-\author[10,11]{Jesper Madsen}
+\author[10,11]{Jesper J. Madsen}  %ORCID: 0000-0003-1411-9080
 \author[2]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
 %\author[2]{...}                   %ORCID: ?
+\author[1]{Ricky Nencini}
 \author[12]{Thomas Piggot}
+\author[2]{{\'A}ngel Pi{\~n}eiro}
 \author[2]{Suman Samantray}
@@
 \affil[9]{CIQUP, Centro de Investigação em Qu{\'i}mica, Departamento de Qu{\'i}mica e Bioqu{\'i}mica, Faculdade de Ci{\^e}ncias, Universidade do Porto, Porto, 4169-007, Portugal}
-\affil[10]{Department of Chemistry, The University of Chicago, Chicago, Illinois, 60637, United States of America}
-\affil[11]{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, 33612, United States of America}
+\affil[10]{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, 33612, United States of America}
+\affil[11]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida, 33612, United States of America}
 \affil[12]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}
+\affil[13]{Institute of Organic Chemistry and Biochemistry of the 
+Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2,  Prague, CZ-16610, Czech Republic}
+\affil[14]{University of Potsdam, Institute of Physics and Astronomy, Potsdam-Golm, 14476, Germany}
 

### 2023-01-30 main.tex
Updates from Overleaf

--- 2023-01-26 main.tex
@@
 
-\author[1]{Anne Kiirikki}         %ORCID: ?
-\author[2]{Hanne Antila}          %ORCID: 0000-0002-2474-5053
-\author[2]{Lara Bort}             %ORCID: ?
-\author[3]{Pavel Buslaev}         %ORCID: ?
+\author[1]{Anne M. Kiirikki}         %ORCID: ?
+\author[2]{Hanne S. Antila}          %ORCID: 0000-0002-2474-5053
+\author[14]{Lara Bort}             %ORCID: ?
+\author[3]{Pavel Buslaev}         %ORCID: 0000-0003-2031-4691
 \author[4]{Fernando Favela}
 \author[5]{Tiago Mendes Ferreira}
-\author[6,7]{Patrick Fuchs}
-\author[8,9]{Rebeca Garcia-Fandino}
+\author[6,7]{Patrick F.J. Fuchs}
+\author[8]{Rebeca Garcia-Fandino}
 \author[2]{Ivan Gushchin}
@@
 \author[2]{Batuhan Kav}           %ORCID: ?
+\author[13]{Patrik Kula}
 \author[2]{Milla Kurki}
@@
 %\author[2]{...}                   %ORCID: ?
+\author[1]{Ricky Nencini}
 \author[12]{Thomas Piggot}
+\author[9]{{\'A}ngel Pi{\~n}eiro}
 \author[2]{Suman Samantray}
-\author[2]{Fabi{\'a}n Suarez-Leston}
+\author[8,9,15]{Fabi{\'a}n Suarez-Leston}
 \author[1,*]{O. H. Samuli Ollila} %ORCID: 0000-0002- 8728-1006
@@
 \affil[6]{Sorbonne Universit{\'e}, Ecole Normale Sup{\'e}rieure, PSL University, CNRS, Laboratoire des Biomol{\'e}cules (LBM), Paris, 75005, France}
-\affil[7]{Universit{\'e} de Paris, UFR Sciences du Vivant, Paris, 75013, France}
+\affil[7]{Universit{\'e} Paris Cit{\'e}, UFR Sciences du Vivant, Paris, 75013, France}
 \affil[8]{Center for Research in Biological Chemistry and Molecular Materials (CiQUS), Universidade de Santiago de Compostela,  Santiago de Compostela, E-15782, Spain}
-\affil[9]{CIQUP, Centro de Investigação em Qu{\'i}mica, Departamento de Qu{\'i}mica e Bioqu{\'i}mica, Faculdade de Ci{\^e}ncias, Universidade do Porto, Porto, 4169-007, Portugal}
+\affil[9]{Department of Applied Physics, Faculty of Physics, University of Santiago de Compostela, Santiago de Compostela, E-15782, Spain}
 \affil[10]{Department of Chemistry, The University of Chicago, Chicago, Illinois, 60637, United States of America}
@@
 \affil[12]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}
+\affil[13]{Institute of Organic Chemistry and Biochemistry of the 
+Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2,  Prague, CZ-16610, Czech Republic}
+\affil[14]{University of Potsdam, Institute of Physics and Astronomy, Potsdam-Golm, 14476, Germany}
+\affil[15]{MD.USE Innovations S.L., Edificio Emprendia, 15782 Santiago de Compostela, Spain}
 
@@
 
-\subsection{PCA analysis of equilibration of simulations}
+\subsection{Principal Components Analysis of equilibration of simulations}
+To estimate how well the provided trajectories are converged, the Principal Components Analysis (PCA) was used \cite{buslaev16,buslaev2020principal}.
+Following the PCALipids \cite{buslaev2020principal} protocol, the lipids of a specific type were first selected, and trajectories of individual lipids were concatenated.
+Next, each lipid configuration was aligned to the average lipid structure.
+Afterwards, the PCA was applied on the cartesian coordinates of all heavy atoms of the lipid.
+Since it was shown before, that the motions along the first, major, principal component (PC) are the slowest ones \cite{buslaev16}, the equilibration of individual lipids was estimated based on the dynamics of the first PC.
+The suggested way to estimate the characteristic timescales needed to adequately sample the configurational space of individual lipids is to calculate the distribution convergence of the trajectories projected on the first PC\cite{buslaev16}.
+However, due to the linear dependence between autocorrelation decay times and distribution convergence times, faster calculation times and higher computational stability of autocorrelation decay times\cite{buslaev16, buslaev2020principal}, the distribution convergence timescales were calculated from autocorrelation decay times:
+\begin{equation}
+    \tau_{\text{convergence}} = 49*\tau_{\text{autocorrelation}}
+\end{equation}
+
+The empirical coefficient 49 was calculated based on the analysis of 8 trajectories from the Databank and included simulations of POPC, POPS, POPE, POPG, and DPPC with CHARMM36 force fields.
+All simulations, used for calculating the empirical coefficient, that relate convergence and autocorrelation, were at least 200 ns long.
+As was shown previously \cite{buslaev16}, the coefficient does not depend on the force field, thus only CHARMM36 simulations as the most frequent in the Databank were used.
+
+As a measure of equilibration quality for each trajectory, the ratio between the convergence time of first PC and the trajectory length was calculated for each lipid type individually.
+Sterols were excluded from the analysis.
+The script that calculates the equilibration of lipids is available at \texttt{Scripts/BuildDatabank/NMRPCA\_timerelax.py} in the NMRlipids databank repository.
+The resulting values are stored in files named \texttt{eq\_times.json} at folders in \texttt{/Data/Simulations} in the NMRlipids databank repository.
 
@@
 \section*{Acknowledgements}
-
 %Acknowledgements should be brief, and should not include thanks to anonymous referees and editors, or effusive comments. Grant or contribution numbers may be acknowledged.
 
+P.B. was supported by the Academy of Finland (Grants 311031 and 342908).
+R.G.-F, A.P. and F.S.-L. thank Centro de Supercomputación de Galicia for computational support.
+R.G.-F.
+thanks Ministerio de Ciencia, Innovaci{\'o}n y Universidades for a ''Ram{\'o}n y Cajal'' contract (RYC-2016-20335), and Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (RTI2018-098795-A-I00, PDC2022-133402-I00), Xunta de Galicia and the ERDF (ED431F 2020/05, 02\_IN606D\_2022\_2667887 and Centro singular de investigaci{\'o}n de Galicia accreditation 2016-2019, ED431G/09).
+AP thanks Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (PID2019-111327GB-I00, PDC2022-133402-I00), Xunta de Galicia and the ERDF (ED431B 2022/36, 02\_IN606D\_2022\_2667887).
+F. S.-L. thanks Axencia Galega de Innovaci{\'o}n for his predoctoral contract (02\_IN606D\_2022\_2667887), and Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (PID2019-111327GB-I00). 
+O.H.S.O, A.M.K, and R.N. acknowledge CSC − IT Center for Science for computational resources and Academy of Finland (grants 315596 and 319902) for ﬁnancial support.
 \section*{Author contributions statement}
+%Must include all authors, identified by initials, for example:
+%A.A. conceived the experiment(s),  A.A. and B.A. conducted the experiment(s), C.A. and D.A. analysed the results.  All authors reviewed the manuscript.
 
-Must include all authors, identified by initials, for example:
-A.A. conceived the experiment(s),  A.A. and B.A. conducted the experiment(s), C.A. and D.A. analysed the results.  All authors reviewed the manuscript.
+P.B. implemented the PCA-based equilibration metric, supervised the work of A.K., and participated in code refactoring.
+R.G.-F was responsible for the design, validation, supervision and funding of the back-end and front-end for the graphical user interface (GUI) (http://www.databank.nmrlipids.fi/).
+A.P. was responsible for the design, validation, supervision and funding of the back-end and front-end for the graphical user interface (GUI) (http://www.databank.nmrlipids.fi/).
+F.S.-L. developed the back-end and participated in the design and development and validation of the graphical user interface (GUI) (http://www.databank.nmrlipids.fi/).
+P.F.J.F. contributed to changes in the buildH software so that the Databank can handle united atom simulation data.
 

### 2023-01-30 main.tex

@@
 \author[2]{Alexander Kuzmin}
-\author[10,11]{Jesper Madsen}
+\author[10,11]{Jesper J. Madsen}  %ORCID: 0000-0003-1411-9080
 \author[2]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
@@
 \affil[9]{Department of Applied Physics, Faculty of Physics, University of Santiago de Compostela, Santiago de Compostela, E-15782, Spain}
-\affil[10]{Department of Chemistry, The University of Chicago, Chicago, Illinois, 60637, United States of America}
-\affil[11]{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, 33612, United States of America}
+\affil[10]{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, 33612, United States of America}
+\affil[11]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida, 33612, United States of America}
 \affil[12]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}

### 2023-01-31 SI.tex
Update authors
Update affiliations

@@
 
-\author[1]{Anne Kiirikki}         %ORCID: ?
-\author[2]{Hanne Antila}          %ORCID: 0000-0002-2474-5053
-\author[2]{Lara Bort}             %ORCID: ?
-\author[3]{Pavel Buslaev}         %ORCID: ?
-\author[4]{Fernando Favela}
-\author[5]{Tiago Mendes Ferreira}
-\author[6,7]{Patrick Fuchs}
-\author[8,9]{Rebeca Garcia-Fandino}
-\author[2]{Ivan Gushchin}
+\author[1]{Anne M. Kiirikki}         %ORCID: ?
+\author[2]{Hanne S. Antila}          %ORCID: 0000-0002-2474-5053
+\author[3]{Lara Bort}             %ORCID: ?
+\author[4]{Pavel Buslaev}         %ORCID: 0000-0003-2031-4691
+\author[5]{Fernando Favela}
+\author[6]{Tiago Mendes Ferreira}
+\author[7,8]{Patrick F.J. Fuchs}
+\author[9]{Rebeca Garcia-Fandino}
+\author[10]{Ivan Gushchin}
 %\author[2]{Matti Javanainen}      %ORCID: 0000-0003-4858-364X
-\author[2]{Batuhan Kav}           %ORCID: ?
-\author[2]{Milla Kurki}
-\author[2]{Alexander Kuzmin}
-\author[10,11]{Jesper Madsen}
-\author[2]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
+\author[11,12]{Batuhan Kav}           %ORCID: 0000-0003-4990-373X
+\author[13]{Patrik Kula}
+\author[14]{Milla Kurki}
+\author[10]{Alexander Kuzmin}
+\author[15,16]{Jesper J. Madsen}  %ORCID: 0000-0003-1411-9080
+\author[17,18]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
 %\author[2]{...}                   %ORCID: ?
-\author[12]{Thomas Piggot}
-\author[2]{Suman Samantray}
-\author[2]{Fabi{\'a}n Suarez-Leston}
+\author[1]{Ricky Nencini}
+\author[19]{Thomas Piggot}
+\author[20]{{\'A}ngel Pi{\~n}eiro}
+\author[10]{Suman Samantray}
+\author[9,20,21]{Fabi{\'a}n Suarez-Leston}
 \author[1,*]{O. H. Samuli Ollila} %ORCID: 0000-0002- 8728-1006
@@
 \affil[1]{University of Helsinki, Institute of Biotechnology, Helsinki, Finland}
-\affil[2]{Affiliation, department, city, postcode, country}
-\affil[3]{Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, P.O. Box 35, Jyv{\"a}skyl{\"a}, 40014 , Finland}
-\affil[4]{Departamento de Ciencias B\'{a}sicas, Tecnol\'{o}gico Nacional de M\'{e}xico - ITS Zacatecas Occidente, Sombrerete, Zacatecas, 99102, M\'{e}xico}
-\affil[5]{NMR group - Institute for Physics, Martin Luther University Halle-Wittenberg,  Halle (Saale), 06120, Germany}
-\affil[6]{Sorbonne Universit{\'e}, Ecole Normale Sup{\'e}rieure, PSL University, CNRS, Laboratoire des Biomol{\'e}cules (LBM), Paris, 75005, France}
-\affil[7]{Universit{\'e} de Paris, UFR Sciences du Vivant, Paris, 75013, France}
-\affil[8]{Center for Research in Biological Chemistry and Molecular Materials (CiQUS), Universidade de Santiago de Compostela,  Santiago de Compostela, E-15782, Spain}
-\affil[9]{CIQUP, Centro de Investigação em Qu{\'i}mica, Departamento de Qu{\'i}mica e Bioqu{\'i}mica, Faculdade de Ci{\^e}ncias, Universidade do Porto, Porto, 4169-007, Portugal}
-\affil[10]{Department of Chemistry, The University of Chicago, Chicago, Illinois, 60637, United States of America}
-\affil[11]{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, 33612, United States of America}
-\affil[12]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}
+\affil[2]{Department of Theory \& Bio-Systems and Department of Biomaterials, Max Planck Institute of Colloids and Interfaces, Germany}
+\affil[3]{University of Potsdam, Institute of Physics and Astronomy, Potsdam-Golm, 14476, Germany}
+%\affil[2]{Affiliation, department, city, postcode, country}
+\affil[4]{Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, P.O. Box 35, Jyv{\"a}skyl{\"a}, 40014 , Finland}
+\affil[5]{Departamento de Ciencias B\'{a}sicas, Tecnol\'{o}gico Nacional de M\'{e}xico - ITS Zacatecas Occidente, Sombrerete, Zacatecas, 99102, M\'{e}xico}
+\affil[6]{NMR group - Institute for Physics, Martin Luther University Halle-Wittenberg,  Halle (Saale), 06120, Germany}
+\affil[7]{Sorbonne Universit{\'e}, Ecole Normale Sup{\'e}rieure, PSL University, CNRS, Laboratoire des Biomol{\'e}cules (LBM), Paris, 75005, France}
+\affil[8]{Universit{\'e} Paris Cit{\'e}, UFR Sciences du Vivant, Paris, 75013, France}
+\affil[9]{Center for Research in Biological Chemistry and Molecular Materials (CiQUS), Universidade de Santiago de Compostela,  Santiago de Compostela, E-15782, Spain}
+\affil[10]{no affiliation}
+\affil[11]{Institute of Biological Information Processing: Structural Biochemistry (IBI-7), Forschungszentrum Jülich, Jülich 52428, Germany}
+\affil[12]{ariadne.ai GmbH (Germany), Häusserstra{\ss}e 3 Heidelberg 69115, Germany }
+\affil[13]{Institute of Organic Chemistry and Biochemistry of the Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2,  Prague, CZ-16610, Czech Republic}
+\affil[14]{School of Pharmacy, University of Eastern Finland, 70211 Kuopio, Finland}
+\affil[15]{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, 33612, United States of America}
+\affil[16]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida, 33612, United States of America}
+\affil[17]{Department of Chemistry, University of Bergen, Norway}
+\affil[18]{Computational Biology Unit, Department of Informatics, University of Bergen, Norway}
+\affil[19]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}
+\affil[20]{Department of Applied Physics, Faculty of Physics, University of Santiago de Compostela, Santiago de Compostela, E-15782, Spain}
+\affil[21]{MD.USE Innovations S.L., Edificio Emprendia, 15782 Santiago de Compostela, Spain}
 
@@
 
-\begin{abstract}
+%\begin{abstract}
 %We present a databank of lipid bilayer simulations from the NMRlipids open collaboration project.
-\end{abstract}
+%\end{abstract}
 

### 2023-01-31 main.tex
Update affiliations

@@
 \author[2]{Hanne S. Antila}          %ORCID: 0000-0002-2474-5053
-\author[14]{Lara Bort}             %ORCID: ?
-\author[3]{Pavel Buslaev}         %ORCID: 0000-0003-2031-4691
-\author[4]{Fernando Favela}
-\author[5]{Tiago Mendes Ferreira}
-\author[6,7]{Patrick F.J. Fuchs}
-\author[8]{Rebeca Garcia-Fandino}
-\author[2]{Ivan Gushchin}
+\author[3]{Lara Bort}             %ORCID: ?
+\author[4]{Pavel Buslaev}         %ORCID: 0000-0003-2031-4691
+\author[5]{Fernando Favela}
+\author[6]{Tiago Mendes Ferreira}
+\author[7,8]{Patrick F.J. Fuchs}
+\author[9]{Rebeca Garcia-Fandino}
+\author[10]{Ivan Gushchin}
 %\author[2]{Matti Javanainen}      %ORCID: 0000-0003-4858-364X
-\author[2]{Batuhan Kav}           %ORCID: ?
+\author[11,12]{Batuhan Kav}           %ORCID: 0000-0003-4990-373X
 \author[13]{Patrik Kula}
-\author[2]{Milla Kurki}
-\author[2]{Alexander Kuzmin}
-\author[10,11]{Jesper J. Madsen}  %ORCID: 0000-0003-1411-9080
-\author[2]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
+\author[14]{Milla Kurki}
+\author[10]{Alexander Kuzmin}
+\author[15,16]{Jesper J. Madsen}  %ORCID: 0000-0003-1411-9080
+\author[17,18]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
 %\author[2]{...}                   %ORCID: ?
 \author[1]{Ricky Nencini}
-\author[12]{Thomas Piggot}
-\author[9]{{\'A}ngel Pi{\~n}eiro}
-\author[2]{Suman Samantray}
-\author[8,9,15]{Fabi{\'a}n Suarez-Leston}
+\author[19]{Thomas Piggot}
+\author[20]{{\'A}ngel Pi{\~n}eiro}
+\author[10]{Suman Samantray}
+\author[9,20,21]{Fabi{\'a}n Suarez-Leston}
 \author[1,*]{O. H. Samuli Ollila} %ORCID: 0000-0002- 8728-1006
@@
 \affil[1]{University of Helsinki, Institute of Biotechnology, Helsinki, Finland}
-\affil[2]{Affiliation, department, city, postcode, country}
-\affil[3]{Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, P.O. Box 35, Jyv{\"a}skyl{\"a}, 40014 , Finland}
-\affil[4]{Departamento de Ciencias B\'{a}sicas, Tecnol\'{o}gico Nacional de M\'{e}xico - ITS Zacatecas Occidente, Sombrerete, Zacatecas, 99102, M\'{e}xico}
-\affil[5]{NMR group - Institute for Physics, Martin Luther University Halle-Wittenberg,  Halle (Saale), 06120, Germany}
-\affil[6]{Sorbonne Universit{\'e}, Ecole Normale Sup{\'e}rieure, PSL University, CNRS, Laboratoire des Biomol{\'e}cules (LBM), Paris, 75005, France}
-\affil[7]{Universit{\'e} Paris Cit{\'e}, UFR Sciences du Vivant, Paris, 75013, France}
-\affil[8]{Center for Research in Biological Chemistry and Molecular Materials (CiQUS), Universidade de Santiago de Compostela,  Santiago de Compostela, E-15782, Spain}
-\affil[9]{Department of Applied Physics, Faculty of Physics, University of Santiago de Compostela, Santiago de Compostela, E-15782, Spain}
-\affil[10]{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, 33612, United States of America}
-\affil[11]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida, 33612, United States of America}
-\affil[12]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}
-\affil[13]{Institute of Organic Chemistry and Biochemistry of the 
-Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2,  Prague, CZ-16610, Czech Republic}
-\affil[14]{University of Potsdam, Institute of Physics and Astronomy, Potsdam-Golm, 14476, Germany}
-\affil[15]{MD.USE Innovations S.L., Edificio Emprendia, 15782 Santiago de Compostela, Spain}
+\affil[2]{Department of Theory \& Bio-Systems and Department of Biomaterials, Max Planck Institute of Colloids and Interfaces, Germany}
+\affil[3]{University of Potsdam, Institute of Physics and Astronomy, Potsdam-Golm, 14476, Germany}
+%\affil[2]{Affiliation, department, city, postcode, country}
+\affil[4]{Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, P.O. Box 35, Jyv{\"a}skyl{\"a}, 40014 , Finland}
+\affil[5]{Departamento de Ciencias B\'{a}sicas, Tecnol\'{o}gico Nacional de M\'{e}xico - ITS Zacatecas Occidente, Sombrerete, Zacatecas, 99102, M\'{e}xico}
+\affil[6]{NMR group - Institute for Physics, Martin Luther University Halle-Wittenberg,  Halle (Saale), 06120, Germany}
+\affil[7]{Sorbonne Universit{\'e}, Ecole Normale Sup{\'e}rieure, PSL University, CNRS, Laboratoire des Biomol{\'e}cules (LBM), Paris, 75005, France}
+\affil[8]{Universit{\'e} Paris Cit{\'e}, UFR Sciences du Vivant, Paris, 75013, France}
+\affil[9]{Center for Research in Biological Chemistry and Molecular Materials (CiQUS), Universidade de Santiago de Compostela,  Santiago de Compostela, E-15782, Spain}
+\affil[10]{no affiliation}
+\affil[11]{Institute of Biological Information Processing: Structural Biochemistry (IBI-7), Forschungszentrum Jülich, Jülich 52428, Germany}
+\affil[12]{ariadne.ai GmbH (Germany), Häusserstra{\ss}e 3 Heidelberg 69115, Germany }
+\affil[13]{Institute of Organic Chemistry and Biochemistry of the Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2,  Prague, CZ-16610, Czech Republic}
+\affil[14]{School of Pharmacy, University of Eastern Finland, 70211 Kuopio, Finland}
+\affil[15]{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, 33612, United States of America}
+\affil[16]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida, 33612, United States of America}
+\affil[17]{Department of Chemistry, University of Bergen, Norway}
+\affil[18]{Computational Biology Unit, Department of Informatics, University of Bergen, Norway}
+\affil[19]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}
+\affil[20]{Department of Applied Physics, Faculty of Physics, University of Santiago de Compostela, Santiago de Compostela, E-15782, Spain}
+\affil[21]{MD.USE Innovations S.L., Edificio Emprendia, 15782 Santiago de Compostela, Spain}
 
@@
 Here we present the NMRlipids Databank---a community-driven, open-for-all database featuring programmatic access to quality evaluated atom-resolution MD simulations of lipid bilayers.
-The NMRlipids databank will benefit scientists in different disciplines by provoding automatic ranking of simulations based on their quality against experiments, flexible implementations of data-driven and machine learning applications, and rapid access to simulation data via graphical user interface.
+The NMRlipids databank will benefit scientists in different disciplines by providing automatic ranking of simulations based on their quality against experiments, flexible implementations of data-driven and machine learning applications, and rapid access to simulation data via graphical user interface.
 To demonstrate the unlocked possibilities beyond current MD simulation studies, we analyzed how anisotropic diffusion of water and cholesterol flip–flop rates depend on membrane properties.
@@
 On the other hand, CHARMM36 gives the best description for lipid headgroup conformational ensembles~\cite{bacle21}, while GROMOS-CKP parameters best capture the membrane packing in POPS bilayers~\cite{antila22b}.
-For example, Fig.~\ref{fig:POPCPOPEapls}B illustrates divergent predictions from different force fields for the lateral packing of POPC and POPE membranes that are the two most biologically abundant neutral membrane lipids~\cite{vanmeer08}.
+For example, Fig.~\ref{fig:POPCPOPEapls}C illustrates divergent predictions from different force fields for the lateral packing of POPC and POPE membranes that are the two most biologically abundant neutral membrane lipids~\cite{vanmeer08}.
 We have previously resolved such controversies for membrane structures and ion binding affinities for neutral~\cite{botan15,catte16} and charged lipids~\cite{antila19,bacle21} by comparing simulations with C-H bond order parameters from NMR, yet this requires manual comparisons between large number of simulations.
@@
 \begin{equation}
-    \tau_{\text{convergence}} = 49*\tau_{\text{autocorrelation}}
+    \tau_{\text{convergence}} = k*\tau_{\text{autocorrelation}}
 \end{equation}
 
-The empirical coefficient 49 was calculated based on the analysis of 8 trajectories from the Databank and included simulations of POPC, POPS, POPE, POPG, and DPPC with CHARMM36 force fields.
-All simulations, used for calculating the empirical coefficient, that relate convergence and autocorrelation, were at least 200 ns long.
-As was shown previously \cite{buslaev16}, the coefficient does not depend on the force field, thus only CHARMM36 simulations as the most frequent in the Databank were used.
+The empirical coefficient $k$ is equal to $k = 49$ and was calculated based on the analysis of 8 trajectories from the Databank and included simulations of POPC, POPS, POPE, POPG, and DPPC with CHARMM36 force fields.
+All simulations, used for $k$ calculation, were at least 200 ns long.
+As was shown previously \cite{buslaev16}, the coefficient does not depend on the force field, thus only CHARMM36 simulations, as the most frequent in the Databank, were used.
 
@@
 F. S.-L. thanks Axencia Galega de Innovaci{\'o}n for his predoctoral contract (02\_IN606D\_2022\_2667887), and Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (PID2019-111327GB-I00). 
+M.S.M. was supported by the Trond Mohn Foundation BFS2017TMT01.
 O.H.S.O, A.M.K, and R.N. acknowledge CSC − IT Center for Science for computational resources and Academy of Finland (grants 315596 and 319902) for ﬁnancial support.

### 2023-02-26 SI.tex
Update text

@@
 
-%\begin{abstract}
+\begin{abstract}
 %We present a databank of lipid bilayer simulations from the NMRlipids open collaboration project.
-%\end{abstract}
+\end{abstract}
 
@@
 %
-\thispagestyle{empty}
+%\thispagestyle{empty}
 
@@
 \end{figure}
+
+\begin{table}[!h]
+    \centering
+    %\begin{tabular}{c c c c}
+    \begin{tabular}{p{3.0cm}  p{5.0cm}  p{3.0cm}  p{4.0cm}}
+        Code & Function & Input & Output \\
+        \hline
+        \multicolumn{4}{c}{ {\bf Databank layer}}\\
+        \hline
+        \multicolumn{4}{c}{ {\bf Scripts/BuildDatabank/ }}\\
+         AddData.py & Create README.yaml from user given information in info.yaml  & info.yaml & README.yaml \\
+         searchDATABANK.py & Pair simulations with available experimental data  & README.yaml files & Updated README.yaml files \\
+         QualityEvaluation.py & Quality evaluate simulations that are paired with experimental data  & README.yaml files & [lipid\_name]\_OrderParameters\_quality.json, [lipid\_name]\_FragmentQuality.json, system\_quality.json, FormFactorQuality.json \\
+         \multicolumn{4}{c}{ {\bf Scripts/AnalyzeDatabank/ }}\\
+         calcOrderParameters.py & Calculate C-H bond order parameters for all lipids in the simulation & README.yaml files & [lipid\_name]\_OrderParameters.json \\
+         calcAPL.py & Calculate area per lipid as a function of time & README.yaml files & apl.json \\
+         calc\_FormFactors.py & Calculate x-ray scattering form factors & README.yaml files & FormFactor.json \\
+         calc\_thickness.py & Calculate membrane thickness & README.yaml files & thickness.json \\
+
+    \end{tabular}
+    \caption{Caption}
+    \label{tab:my_label}
+\end{table}
 
@@
 
+\begin{table}[p]
+    \centering
+    \begin{tabular}{l}
+    Force field name and references \\
+    \hline
+    CHARMM36~\cite{klauda10}\\
+    Slipids~\cite{jambeck12,jambeck12b,jambeck2012another,ermilova16,grote20}  \\
+    MacRog~\cite{Kulig15b}  \\
+    Amber Lipid14/17~\cite{dickson14,dickson22}  \\
+    Charmm-Drude~\cite{li2017drude}      \\
+    ECClipids~\cite{melcr18,melcr20,bacle21}  \\
+    GROMOS-CKP~\cite{Chandrasekhar03,kukol09,piggot12}  \\
+    Berger~\cite{berger97}  \\
+    ECC-CHARMM36~\cite{nencini2019development}  \\
+    Orange~\cite{catte16}  \\
+    Poger~\cite{poger10}  \\
+    GROMOS 43A1-S3~\cite{chiu09}  \\ 
+    GAFFlipid~\cite{dickson12}  \\
+    OPLS3e~\cite{roos19}  \\
+    Ulmschneider~\cite{Ulmschneider09}  \\
+    Chiu Gromos~\cite{chiu09}  \\
+    \end{tabular}
+    \caption{List of current force fields used in simulations in the databank with references.}
+    \label{tab:ForceFields}
+\end{table}
+
 \clearpage

### 2023-02-26 main.tex
Updates from Overleaf
Update text

@@
 
-\subsection{Programmatic access to MD simulations of membranes composed of most common biologically abundant lipids}
+\subsection{MD simulations of membranes composed of most common biologically abundant lipids}
 
 NMRlipids Databank is a community-driven database containing atomistic MD simulations of biologically relevant lipid membranes emerging from the NMRlipids open collaboration~\cite{botan15,ollila16,catte16,antila19,bacle21}.
-It has been designed to improve data Findability, Accessibility, Interoperability, and Reuse (FAIR)~\cite{wilkinson16} using the NMRlipids project protocol where all the content is openly accessible from the beginning of the project~\cite{botan15}.
-Currently the NMRlipids databank contains nearly thousand simulation trajectories with the total length approaching half a millisecond.
-The distribution of available simulations containing a specific lipid in Fig.~\ref{fig:overlay}C roughly resembles the biological abundance of different lipid types with phosphatidylcholine (PC) being the most common followed by cholesterol, phosphatidylethanolamine (PE), phosphatidylserine (PS), phosphatidylglycerol (PG), phosphatidylinositol (PI), and other lipids depending on organism and organelle~\cite{vanmeer08}.
-Currently available single component lipid membranes and binary mixtures in the NMRlipids databank are illustrated in Fig.~\ref{fig:overlay}D.
-While these comprise the majority of available trajectories, mixtures with up to five lipid types are available.
-Force fields used in simulations cover all the essential parameters commonly used in lipid simulations, see Fig.~\ref{fig:overlay}E, including also united atom and polarizable force fields.
+It has been designed to improve data Findability, Accessibility, Interoperability, and Reuse (FAIR)~\cite{wilkinson16} for MD simulation trajectories.
+The NMRlipids databank is constructed using the NMRlipids project protocol where all the content is openly accessible throughout the project~\cite{botan15}.
+Currently the NMRlipids databank contains 726 simulation trajectories with the total length of approximately 0.4\,ms.
+The distribution of available simulations containing a specific lipid, shown in Fig.~\ref{fig:overlay}B, roughly resembles the biological abundance of different lipid types with phosphatidylcholine (PC) being the most common followed by cholesterol, phosphatidylethanolamine (PE), phosphatidylserine (PS), phosphatidylglycerol (PG), phosphatidylinositol (PI), and other lipids depending on organism and organelle~\cite{vanmeer08}.
+Abbreviations and full names of all lipids present in the databank are listed in Table~\ref{tab:abbreviations}.
+Single component lipid membranes and binary mixtures, depicted in Fig.~\ref{fig:overlay}D, are currently most abundant in the NMRlipids databank, yet mixtures with up to five lipid types are available.
+Force fields used in simulations cover all the essential parameters commonly used in lipid simulations, see Fig.~\ref{fig:overlay}C and table~\ref{tab:ForceFields}, including also united atom and polarizable force fields.
 Therefore, the averages calculated over the databank can be considered as mean predictions from available lipid models for an average cell membrane.
 
-The overlay structure, illustrated in Fig.~\ref{fig:overlay}A, is designed to enable efficient upcycling of MD simulations for data-driven and machine learning applications using 
-The NMRlipids databank.
-Raw simulation data in the {\it data layer} can be stored in any publicly available location with long term stability, such as Zenodo (\url{zenodo.org}), and with permanent links to the data, such as digital object identifiers (DOIs).
+The overlay structure of the NMRlipids databank, illustrated in Fig.~\ref{fig:overlay}A, is designed to enable efficient upcycling of MD simulations for data-driven and machine learning applications with mimimal investment on new infrastructure.
+Raw simulation data in the {\it data layer} can be stored in any publicly available location with long term stability, such as Zenodo (\url{www.zenodo.org}), and with permanent links to the data, such as digital object identifiers (DOIs).
 %or uniform resource names (URNs). 
-The {\it Databank layer} is the core of the databank containing all the relevant information about the simulations, including links to the raw data, relevant metadata describing the systems, universal naming conventions for lipids and their atoms, computer programs to create and analyse the entries, basic properties calculated from all simulations (area per lipid, C--H bond order parameters, X-ray scattering form factors and membrane thickness), and quality evaluation of simulations against experimental data.
-In practise, this information is stored in the git repository which is available at~\url{https://github.com/NMRLipids/Databank}.
-The {\it application layer} contains outputs from the databank, such as graphical user interface (\url{http://www.databank.nmrlipids.fi/}) and results from analyses described in sections below which are available at \url{https://github.com/NMRLipids/DataBankManuscript}.
+The {\it Databank layer} (\url{www.github.com/NMRLipids/Databank}) is the core of the databank containing all the relevant information about the simulations, including links to the raw data, relevant metadata describing the systems, universal naming conventions for lipids and their atoms, quality evaluation of simulations against experimental data, and computer programs to create the entries and analyse basic properties calculated from all simulations (area per lipid, C--H bond order parameters, X-ray scattering form factors, membrane thickness and equilibration times of principal components).
+Also the results for these basic properties are stored in the {\it Databank layer}.
+
+The {\it application layer} is composed of repositories and tools that read information from the {\it databank layer} for further analyses, and their results.
+Because the tools in the {\it application layer} do not interfere with the {\it databank layer}, they can be freely extended by anyone for wide range of purposes.
+Here we exemplify this with two For example, the analyses described in below sections locate in an application layer repository at \url{www.github.com/NMRLipids/DataBankManuscript}, and content of the databank can be accessed with a graphical user interface tool at~\url{www.databank.nmrlipids.fi/}.
 A more detailed description on the NMRlipids databank structure is available the supplementary information.
@@
     \centering
-    \includegraphics[width=\linewidth]{Figures/overlay2.pdf}
-    \caption{a) Schematic presentation of the overlay structure used in the NMRlipids databank. A more detailed structure of the databank layer is shown in Fig.~\ref{DatabankStructure} in the SI.
-    b) Distribution of the lengths of the trajectories, total number of trajectories and total length of the simulations in the NMRlipids databank.
-    c) Distribution of lipids present in the trajectories in the NMRlipids databank. Lipids occuring in five or fewer simulations ('others') are listed on the right. 
-    d) Currently available binary mixtures in the NMRlipids databank. 
-    e) Distribution of force fields in the simulations in the NMRlipids databank.
+    \includegraphics[width=\linewidth]{Figures/overlay3.pdf}
+    \caption{A) Schematic presentation of the overlay structure used in the NMRlipids databank. A more detailed structure of the databank layer is shown in Fig.~\ref{DatabankStructure} in the SI.
+    %b) Distribution of the lengths of the trajectories, total number of trajectories and total length of the simulations in the NMRlipids databank.
+    B) Distribution of lipids present in the trajectories in the NMRlipids databank. Lipids occuring in five or fewer simulations ('others') are listed on the right. 
+    C) Distribution of force fields in the simulations in the NMRlipids databank. References for each force field are given in table~\ref{tab:ForceFields}.
+    D) Currently available binary mixtures in the NMRlipids databank. 
     }
@@
 
-\subsection{Accessing NMRlipids databank}\label{section:access}
-The content of the NMRlipids databank is easily searchable and browsable via graphical user interface (GUI) (\url{http://www.databank.nmrlipids.fi/}), which can be also used to visualize the systems, compare basic properties between diffent systems and experiments, and rank simulations based on their quality evaluated against experimental data.
+\subsection{NMRlipids databank-GUI: graphical access to the MD simulation data}\label{section:access}
+NMRlipids databank-GUI provides easy access to the content of the NMRlipids databank via graphical user interface (GUI) that is available at~\url{www.databank.nmrlipids.fi/}.
+The content of the NMRlipids databank can be easily browsed and adanced searched options based on systems content and properties are available.
+Systems and their basic properties can be visualized and compared between different simulations and experiments.
+Simulations can be also ranked simulations based on their quality evaluated against experimental data.
 
-%\subsection{Analyzing the databank}
-More flexible analyses that enable wide range of novel applications can be performed by the Python application programming interface (API) available at \url{https://github.com/NMRLipids/Databank}.
-The API can be used to automatically analyse virtually any property from all simulations in the NMRlipids databank following the flowchart demonstrated in Fig.~\ref{fig:POPCPOPEapls}A.
-Iteration through the README.yaml files in the {\it databank layer} enables the access to raw simulation data via the links stored in these files.
-Affiliation of universal molecule and atom names to the simulation specific names in README.yaml and mapping files enable automatic analyses independently on the specific naming convention used in a simulation.
-Analysis codes and results for membrane area per lipid, C–H bond order parameters, X-ray scattering form factors and thickness are included in the {\it Databank layer}.
-Further analysis codes can be implemented by users and the results can be stored to the {\it application layer} in the same format that is used in the {\it Databank layer}.
-The results from analyses can be then programmatically accessed and furhter analyzed using the NMRlipids databank API following the flowchart demonstrated in Fig.~\ref{fig:POPCPOPEapls}B.
+\subsection{NMRlipids databank-API: programmatic access to the MD simulation data}
+The NMRlipids databank-API, available at \url{www.github.com/NMRLipids/Databank}, enables automatic analysis of virtually any property from all simulations in the NMRlipids databank.
+The NMRlipids databank-API delivers functions that enable straightforward implementation of the flowchart demonstrated in Fig.~\ref{fig:POPCPOPEapls}A.
+Iteration through the README.yaml files in the {\it databank layer} enables the access to the raw data of each simulation within the databank via the links stored in these files.
+The affiliation of universal molecule and atom names to the simulation specific names are delivered for each simulation in README.yaml and mapping files, which enables automatic analyses over all systems independently on the specific naming convention used in a simulation.
+Finally, the analysis results are stored using the same structure as in the {\it databank layer}.
 
-These tools offer open programmatic access to unprecedented amount of simulation trajectories for all, which enables possibilities for wide range of novel analyses.
-This is particularly useful for scientist without access to resources to perform large scale MD simulations and for analyses of rare phenomena that require exceptional amounts of simulation data.
-On the other hand, universal naming convention enables systematic comparisons between simulations having different model parameters and thereby different naming conventions.
-This will guide the development of more accurate simulation parameters and usage of atomistic models to train coarse grained simulations.
+While analysis codes and results for basic membrane properties 
+%area per lipid, C–H bond order parameters, X-ray scattering form factors and thickness 
+are included in the {\it Databank layer}, unlimited amount of further analyses can be implemented by anyone in separate repositories in the {\it application layer}.
+{\it Application layer} repositories that are organized using the {\it Databank layer} structure can be straightforwardly accessed programmatically and fuhther analyzed by implementing the flowchart demonstrated in Fig.~\ref{fig:POPCPOPEapls}B by using the NMRlipids databank-API functions.
+An {\it application layer} repository used to produce the novel analyses from the NMRlipids databank exemplified in the next section is available at~\url{www.github.com/NMRLipids/DataBankManuscript}.
 
@@
 In addition, the data contained in the NMRlipids databank will enable training of more elaborated models to predict arbitrary membrane properties.
+
+These tools offer open programmatic access to unprecedented amount of simulation trajectories for all, which enables possibilities for wide range of novel analyses.
+This is particularly useful for scientist without access to resources to perform large scale MD simulations and for analyses of rare phenomena that require exceptional amounts of simulation data.
+On the other hand, universal naming convention enables systematic comparisons between simulations having different model parameters and thereby different naming conventions.
+This will guide the development of more accurate simulation parameters and usage of atomistic models to train coarse grained simulations.
 
@@
 M.S.M. was supported by the Trond Mohn Foundation BFS2017TMT01.
-O.H.S.O, A.M.K, and R.N. acknowledge CSC − IT Center for Science for computational resources and Academy of Finland (grants 315596 and 319902) for ﬁnancial support.
+O.H.S.O, A.M.K, and R.N. acknowledge CSC --- IT Center for Science for computational resources and Academy of Finland (grants 315596 and 319902) for financial support.
 \section*{Author contributions statement}

### 2023-02-26 main.tex
Update text and pdfs

@@
 
-The {\it application layer} is composed of repositories and tools that read information from the {\it databank layer} for further analyses, and their results.
-Because the tools in the {\it application layer} do not interfere with the {\it databank layer}, they can be freely extended by anyone for wide range of purposes.
-Here we exemplify this with two For example, the analyses described in below sections locate in an application layer repository at \url{www.github.com/NMRLipids/DataBankManuscript}, and content of the databank can be accessed with a graphical user interface tool at~\url{www.databank.nmrlipids.fi/}.
-A more detailed description on the NMRlipids databank structure is available the supplementary information.
+The {\it application layer} is composed of repositories and tools that read information from the {\it databank layer} for further analyses.
+Because {\it application layer} does not interfere with the {\it databank layer}, it can be freely extended by anyone for wide range of purposes.
+Here we demonstrate this with two examples: NMRlipids databank graphical user interface (NMRlipids databank-GUI) tool at~\url{www.databank.nmrlipids.fi/} and a repository exemplifying novel analyses utilizing NMRlipids databank as discussed below (\url{www.github.com/NMRLipids/DataBankManuscript}).
+More detailed description of the NMRlipids databank structure is available in the supplementary information.
 
@@
 NMRlipids databank-GUI provides easy access to the content of the NMRlipids databank via graphical user interface (GUI) that is available at~\url{www.databank.nmrlipids.fi/}.
-The content of the NMRlipids databank can be easily browsed and adanced searched options based on systems content and properties are available.
-Systems and their basic properties can be visualized and compared between different simulations and experiments.
-Simulations can be also ranked simulations based on their quality evaluated against experimental data.
+The NMRlipids databank-GUI has advanced searching options for system content and properties.
+The search results are ranked based on their quality evaluation against experimental data which enables rapid selection of best available simulations for specific systems.
+Systems can be also visualized and their properties can be compared with different simulations and experiments.
 
@@
 The NMRlipids databank-API, available at \url{www.github.com/NMRLipids/Databank}, enables automatic analysis of virtually any property from all simulations in the NMRlipids databank.
-The NMRlipids databank-API delivers functions that enable straightforward implementation of the flowchart demonstrated in Fig.~\ref{fig:POPCPOPEapls}A.
+This enables wide range of novel analyses that take the advantage on unprecendented amount MD simulation data with programmatic access.
+This is realized in the NMRlipids databank-API by delivering functions that enable straightforward implementation of the flowchart demonstrated in Fig.~\ref{fig:POPCPOPEapls}A.
 Iteration through the README.yaml files in the {\it databank layer} enables the access to the raw data of each simulation within the databank via the links stored in these files.
@@
 are included in the {\it Databank layer}, unlimited amount of further analyses can be implemented by anyone in separate repositories in the {\it application layer}.
-{\it Application layer} repositories that are organized using the {\it Databank layer} structure can be straightforwardly accessed programmatically and fuhther analyzed by implementing the flowchart demonstrated in Fig.~\ref{fig:POPCPOPEapls}B by using the NMRlipids databank-API functions.
-An {\it application layer} repository used to produce the novel analyses from the NMRlipids databank exemplified in the next section is available at~\url{www.github.com/NMRLipids/DataBankManuscript}.
+{\it Application layer} repositories that are organized using the {\it Databank layer} structure can be straightforwardly accessed programmatically and further analyzed by implementing the flowchart demonstrated in Fig.~\ref{fig:POPCPOPEapls}B by using the NMRlipids databank-API functions.
+Novel analyses demonstrating the power of the NMRlipids databank in selecting best simulation models, analyses of rare phenomena and extending MD simulations to new fields are implemented in an {\it application layer} repository located at~\url{www.github.com/NMRLipids/DataBankManuscript}.
 
@@
 %Quality of membrane simulations with different force fields have been evaluated against experimental data during parameterization and in separate comparison studies~\cite{botan15,ollila16,catte16,pluhackova16,perez17,leonard19,??}, but the lack of universal quality measure for membrane simulations complicates the estimations of reliability of simulations.
-Understanding of how closely simulations resemble the reality is crucial for selecting best models for specific applications, assessing the reliability and significance of MD simulation results, and using atomistic simulations in parametrization of coarse grained models.
+%
+%Understanding of how closely simulations resemble the reality is crucial for selecting best models for specific applications, assessing the reliability and significance of MD simulation results, and using atomistic simulations in parametrization of coarse grained models.
 However, undefined universal quality measures for MD simulations and the lack of a prevailing force field that would perform best in all simulations has led to many controversial results in membrane simulations~\cite{antila22b}.
+%
 %For example, OPLS3e parameters overcome CHARMM36 in structural quality for POPC but predicts overestimated ion binding~\cite{kurki22}.
 On the other hand, CHARMM36 gives the best description for lipid headgroup conformational ensembles~\cite{bacle21}, while GROMOS-CKP parameters best capture the membrane packing in POPS bilayers~\cite{antila22b}.
-For example, Fig.~\ref{fig:POPCPOPEapls}C illustrates divergent predictions from different force fields for the lateral packing of POPC and POPE membranes that are the two most biologically abundant neutral membrane lipids~\cite{vanmeer08}.
-We have previously resolved such controversies for membrane structures and ion binding affinities for neutral~\cite{botan15,catte16} and charged lipids~\cite{antila19,bacle21} by comparing simulations with C-H bond order parameters from NMR, yet this requires manual comparisons between large number of simulations.
-To streamline this process,
-%the selection of best simulations for specific applications, and to guide the development of atomistic and coarse grained parameters,
-we have defined quantitative quality measures, based on C--H bond order parameters from NMR and X-ray scattering form factors, for lipid bilayer simulations in the NMRlipids databank. 
-%The quality for each simulation in the NMRlipids databank with the corresponding experimental data available is currently done using C--H bond order parameters from NMR experiments and form factors from X-ray scattering experiments.
+%
+%Understanding of how closely simulations resemble the reality is crucial for selecting best models for specific applications,
+The quality of lipid bilayer MD simulations has to be carefully assessed in their applications to minize the detrimental consequences of artificial MD simulation results for lipid bilayers~\cite{antila22b}.
+This is possible, for example, by using C--H bond order parameters from NMR and X-ray scattering form factors \cite{ollila16}, although it requires comparisons between large number of simulations which is laborious even with collaborative approaches~\cite{botan15,catte16,antila19,bacle21}.
+The quantitative quality measures in the NMRlipids databank substantially streamline this process by enabling automatic ranking of lipid bilayer simulations based on their quality against experiments.
+Here we demonstrate this by selecting the best models for simulations of membranes with POPC and POPE lipids that are the two most biologically abundant neutral membrane lipids~\cite{vanmeer08}.
+
+Divergent predictions for the lateral packing of POPC and POPE membranes from different force fields is illustrated in Fig.~\ref{fig:POPCPOPEapls}C.
+To select the best models for these membranes, the ranking based on C--H bond order parameter quality for the best performing POPC and POPE simulations in the NMRlipids databank are shown in Figs.~\ref{fig:quality}A and B, respectively.
+Direct comparisons with the experimental are shown in Figs.~\ref{fig:quality}C--E and~\ref{fig:POPC_POPE_dataSI} for the most relevant simulations.
+Based on these rankings, simulations with Slipids parameters are the most realistic for POPC lipids with a clear difference to others, and 2\textsuperscript{nd} most realistic for POPE with only marginally lower quality than GROMOS-CKP simulations.
+Simulations with CHARMM36 and GROMOS-CKP parameters predict overly packed bilayer for POPC (overestimated order in Fig.~\ref{fig:POPC_POPE_dataSI}), while the area per lipid for POPC is overestimated in Lipid17.
+For POPE, CHARMM36 and Lipid17 predict membranes that are too packed.
+Because the area per lipid correlates with the average order parameter of the $\textit{sn}$-1 chain (Fig.~\ref{fig:quality} F), quality evaluation in Fig.~\ref{fig:quality} and direct comparison in Fig.~\ref{fig:POPC_POPE_dataSI} can be used to select the simulations that best capture area per lipids for POPE and POPC lipids in Fig.~\ref{fig:POPCPOPEapls}C.
+In conclusion, the quality evalution based on the NMRlipids databank suggests that Slipids parameters are the best available choice for simulations with PC and PE lipids, at least for the cases were membrane packing is relevant.
+
 The C--H bond order parameters are sensitive to conformational ensembles of individual lipid molecules~\cite{ollila16}, while correlating well also with the membrane thickness and lateral packing as shown in Fig.~\ref{fig:quality}F.
@@
 Together with the API, these enable rapid and automatic quality evaluation of simulations in the NMRlipids databank, thereby promoting the selection of best simulations for specific applications without laborious force field evaluation, and automatic parametrization procedures for atomistic and coarse grained simulations.
-
-To select the best models for POPC and POPE among simulations compared in Fig.~\ref{fig:POPCPOPEapls}C, we show the ranking based on C--H bond order parameter quality in Fig.~\ref{fig:quality}A for the best performing simulations in the NMRlipids databank, and for the best performing POPE lipid bilayer simulations in Fig.~\ref{fig:quality}B.
-Direct comparisons with experiments are shown for selected simulations in Figs.~\ref{fig:quality}C--E and~\ref{fig:POPC_POPE_dataSI}.
-%The power of the NMRlipids databank in selecting the best simulation parameters for a specific application is demonstrated for mixtures of PE and PC lipids in Fig.~\ref{fig:POPCPOPEapls} showing area per lipids from available POPC and POPE bilayer simulations in the databank from different force fields at 310\,K.
-Predictions from different force fields deviate significantly in terms of absolute values and the slope of decrease upon addition of POPE.
-Because the area per lipid correlates with the average order parameter of the $\textit{sn}$-1 chain (Fig.~\ref{fig:quality} F), quality evaluation in Fig.~\ref{fig:quality} and direct comparison in Fig.~\ref{fig:POPC_POPE_dataSI} can be used to select the simulations that best capture area per lipids for POPE and POPC lipids in Fig.~\ref{fig:POPCPOPEapls}C.
-In these results, simulations with Slipids parameters ranks 1\textsuperscript{st} for POPC with a clear difference to others, and 2\textsuperscript{nd} for POPE with only marginally lower quality than GROMOS-CKP.
-Simulations with CHARMM36 and GROMOS-CKP parameters predict overly packed bilayer for POPC (overestimated order in Fig.~\ref{fig:POPC_POPE_dataSI}), while the area per lipid for POPC is overestimated in Lipid17.
-For POPE, CHARMM36 and Lipid17 predict membranes that are too packed.
-In conclusion, the quality evalution based on the NMRlipids databank suggests that Slipids parameters are the best available choice for simulations with PC and PE lipids, at least for the cases were membrane packing is relevant.
 

### 2023-02-28 main.tex
UPdate text

@@
 Divergent predictions for the lateral packing of POPC and POPE membranes from different force fields is illustrated in Fig.~\ref{fig:POPCPOPEapls}C.
-To select the best models for these membranes, the ranking based on C--H bond order parameter quality for the best performing POPC and POPE simulations in the NMRlipids databank are shown in Figs.~\ref{fig:quality}A and B, respectively.
-Direct comparisons with the experimental are shown in Figs.~\ref{fig:quality}C--E and~\ref{fig:POPC_POPE_dataSI} for the most relevant simulations.
+Because the membrane packing in terms of area per lipid correlates with the average order parameter of the $\textit{sn}$-1 chain (Fig.~\ref{fig:quality} F), simulations with the best quality for this part of POPE and POPC lipids are considered to give the most realistic results in Fig.~\ref{fig:POPCPOPEapls}C.
+To this end, the ranking based on C--H bond order parameter quality for the best performing POPC and POPE simulations in the NMRlipids databank are shown in Figs.~\ref{fig:quality}A and B, respectively.
+Direct comparisons with the experimental data are shown in Figs.~\ref{fig:quality}C--E and~\ref{fig:POPC_POPE_dataSI} for the most relevant simulations.
 Based on these rankings, simulations with Slipids parameters are the most realistic for POPC lipids with a clear difference to others, and 2\textsuperscript{nd} most realistic for POPE with only marginally lower quality than GROMOS-CKP simulations.
@@
 For POPE, CHARMM36 and Lipid17 predict membranes that are too packed.
-Because the area per lipid correlates with the average order parameter of the $\textit{sn}$-1 chain (Fig.~\ref{fig:quality} F), quality evaluation in Fig.~\ref{fig:quality} and direct comparison in Fig.~\ref{fig:POPC_POPE_dataSI} can be used to select the simulations that best capture area per lipids for POPE and POPC lipids in Fig.~\ref{fig:POPCPOPEapls}C.
 In conclusion, the quality evalution based on the NMRlipids databank suggests that Slipids parameters are the best available choice for simulations with PC and PE lipids, at least for the cases were membrane packing is relevant.

### 2023-03-02 main.tex
Update text and pdf

@@
 Simultaneously, machine learning based models predicting behaviour of biomolecules and automatic approaches to parametrize models are emerging~\cite{jumper21,antila22b}.
-The NMRlipids databank will support the development in both of these directions.
-The automatic quality evaluation and ranking of simulations in the NMRlipids databank guides the optimization of new force field parameters and the selection of best parameters for large biomolecular complexes where simulation quality becomes increasingly important due the accumulation of small errors in large simulations.
-Furthermore, the NMRlipids databank serves as a training set for diverse machine learning applications.
-For example, a machine learning model predicting electron density profiles from form factors would be highly useful in interpretation of scattering experiments.
-In addition, the data contained in the NMRlipids databank will enable training of more elaborated models to predict arbitrary membrane properties.
+The new NMRlipids databank will support the development in both of these directions.
+The automatic quality evaluation and ranking of simulations against experimental data in the NMRlipids databank can be used to select the best parameters for simulations of large biomolecular complexes and to guide optimizations of new force field parameters for both atomistic and coarse grained simulations.
+These issues are becoming increasingly important due the accumulation of small errors when complexity and size of simulated systems are increasing.
 
-These tools offer open programmatic access to unprecedented amount of simulation trajectories for all, which enables possibilities for wide range of novel analyses.
-This is particularly useful for scientist without access to resources to perform large scale MD simulations and for analyses of rare phenomena that require exceptional amounts of simulation data.
-On the other hand, universal naming convention enables systematic comparisons between simulations having different model parameters and thereby different naming conventions.
-This will guide the development of more accurate simulation parameters and usage of atomistic models to train coarse grained simulations.
-
+The open programmatic access to unprecedented amount of MD simulation via NMRlipids databank-API enables wide range of novel data-driven analyses and machine learning applications.
+For example, large amount of accessible simulation data in terms of quantity (\textit{e.g.}, simulation length and number of conformations) and content (\textit{e.g.}, lipid compositions and ion concentrations) enables analyses of rare phenomena beyond current possibilities for a single research group as demonstrated for cholesterol flip--flop events (Fig.~\ref{fig:flip-flops}) and water permeation through membranes (Fig.~\ref{fig:permeability}).
+This will also pave the way of applications of MD simulations in new fields as demonstrated here by analysing the anisotropic diffusion of water in membrane systems (Fig.~\ref{fig:permeability}) that is essential parameter in pharmacokinetic modeling and MRI imaging~\cite{nitsche19,topgaard20}.
+Furthermore, the NMRlipids databank-API paves the way for diverse machine learning applications to
+%For example, a machine learning model predicting electron density profiles from form factors would be highly useful in interpretation of scattering experiments.
+%In addition, the data contained in the NMRlipids databank will enable training of more elaborated models to
+predict membrane properties using NMRlipids databank as a training set.
+Such applications are analogous to AlphaFold \cite{jumper21} and other tools that predict proteins structures from their sequence using artificial intelligence.
+These possibilities are particularly valuable for scientist who do not typically have access to large scale MD simulation data.
+%and for analyses of rare phenomena that require exceptional amounts of simulation data.
+%On the other hand, universal naming convention enables systematic comparisons between simulations having different model parameters and thereby different naming conventions.
+%
 %The NMRlipids databank will be of great benefit to scientists involved in MD simulations.
@@
 The selection of best models for a specific application is demonstrated for PC/PE lipid mixtures in Fig.~\ref{fig:quality}, and previously for PC/PS lipid mixtures~\cite{antila22b}, and lipid headgroups~\cite{bacle21}.
-The programmatic access in the NMRlipids databank enables automatic analyses of structural and dynamic properties over larger sets of simulation data in terms of quantity (\textit{e.g.}, simulation length and number of conformations) and content (\textit{e.g.}, lipid compositions and ion concentrations) than is currently possible by a single research group.
-The added value of such analyses is demonstrated here by analysing the anisotropic diffusion of water in membrane systems (Fig.~\ref{fig:permeability}) and cholesterol flip--flop events (Fig.~\ref{fig:flip-flops}).
-By enabling these analyses for everyone, the NMRlipids databank can lead to applications in unprecedented directions in fields where MD simulations are less commonly utilized, such as pharmacokinetic modeling and in MRI imaging where anisotropic diffusion of water play important roles~\cite{nitsche19,topgaard20}.
-Different types of applications enabled by the NMRlipids databank across a wide range of fields are exemplified in Table~\ref{tab:applications}.
-The ever growing amount of data is expected to further increase the scope of potential applications of the NMRlipids databank in biology, biotechnology, material science, biomolecular imaging and beyond.
+%
+Different types of expected applications of the NMRlipids databank in wide range of fields are listed in Table~\ref{tab:applications}, yet the scope of such applications
+%in biology, biotechnology, material science, biomolecular imaging and beyond
+is expected to further widen with increasing amount of data in the NMRlipids databank.
 

### 2023-03-09 main.tex
Update figures

@@
 
-\begin{figure}
+\begin{figure}[t]
     \centering
@@
 
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=88mm]{Figures/flowchart2.pdf}
-    \caption{ A) Flowchart for performing an analysis of properties through all MD simulations in the NMRlipids databank using the API.
-    B) Flowchart for accessing results calculated from the NMRlipids databank and stored to the application layer.
-    C) Area per lipids of POPC and POPE lipid bilayers predicted by different force fields at 310\,K in simulations that are available in the NMRlipids databank.
+%\begin{figure}[t]
+%    \centering
+%    \includegraphics[width=88mm]{Figures/flowchart2.pdf}
+%    \caption{ A) Flowchart for performing an analysis of properties through all MD simulations in the NMRlipids databank using the API.
+%    B) Flowchart for accessing results calculated from the NMRlipids databank and stored to the application layer.
+%    C) Area per lipids of POPC and POPE lipid bilayers predicted by different force fields at 310\,K in simulations that are available in the NMRlipids databank.
 The points from the best performing simulations in the NMRlipids databank quality evaluation for each lipid (Figs.~\ref{fig:quality} and~\ref{fig:POPC_POPE_dataSI}) are surrounded by black circles.
-    }
-    \label{fig:POPCPOPEapls}
-\end{figure}
+%    }
+%    \label{fig:POPCPOPEapls}
+%\end{figure}
 
@@
     \centering
-    \includegraphics[width=\linewidth]{Figures/quality3.pdf}
+    \includegraphics[width=\linewidth]{Figures/quality5.pdf}
     \caption{ A) Best simulations ranked based on overall order parameter quality.

### 2023-03-09 main.tex
Updates from Overleaf

--- 2023-03-02 main.tex
@@
 The importance of such approaches is widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} and data-sharing solutions are emerging for proteins in solution~\cite{meyer10,kamp10}, proteins in membranes~\cite{newport19,espigares20,leston22}, nucleic acids~\cite{hospital16}, nucleic acids and proteins~\cite{bekker20}, cyclodextrins~\cite{mixcoha16}, COVID-19--involved macromolecules~(\url{bioexcel-cv19.bsc.es}).
-However, automatic quality evaluation of simulation data~\cite{meyer10,hospital16} and programmatic access is still rare.
+However, automatic quality evaluation of simulation data~\cite{meyer10,hospital16} and programmatic access are still rare.
 In particular, tools for automatic quality evaluation of membrane simulations, or training sets for machine learning models of membrane-containing systems, are not yet available.
@@
 
-While the NMRlipids Databank currently contains only lipid bilayer systems, it can readily be expanded to contain other molecules, such as disordered proteins or sugars.
-The powerful combination of the overlay databank structure and a community-wide open collaboration can be applied also when building databases in other fields, particularly when storage of raw data requires significant resources, best practices in the field are not defined, and incentives to share data do not exist.
-Therefore, the approach introduced here opens up possibilities to create databases enabling data-driven and machine learning applications in new fields.
+While the NMRlipids Databank currently contains only lipid bilayer systems, its key elements that enable the programmatic access to large scale MD simulation data can be applied also for other molecules, such as disordered proteins or sugars.
+The powerful combination of the overlay databank structure and a community-wide open collaboration can be used to build databases that enable data-driven and machine learning applications also in other fields where storage of raw data requires significant resources, best practices in the field are not defined, and incentives to share data do not exist, such as assignment of NMR spectra~\cite{klukowski22}.
 
@@
 Simultaneously, machine learning based models predicting behaviour of biomolecules and automatic approaches to parametrize models are emerging~\cite{jumper21,antila22b}.
-The new NMRlipids databank will support the development in both of these directions.
-The automatic quality evaluation and ranking of simulations against experimental data in the NMRlipids databank can be used to select the best parameters for simulations of large biomolecular complexes and to guide optimizations of new force field parameters for both atomistic and coarse grained simulations.
-These issues are becoming increasingly important due the accumulation of small errors when complexity and size of simulated systems are increasing.
+The resources delivered by the NMRlipids databank will support the development in both of these directions.
+The automatic quality evaluation and ranking of simulations against experimental data can be used to select the best parameters for simulations of large biomolecular complexes and to guide optimizations of new force field parameters for both atomistic and coarse grained simulations.
+The accuracy of simulations is becoming increasingly important due the accumulation of small errors when complexity and size of simulated systems are increasing.
 
-The open programmatic access to unprecedented amount of MD simulation via NMRlipids databank-API enables wide range of novel data-driven analyses and machine learning applications.
-For example, large amount of accessible simulation data in terms of quantity (\textit{e.g.}, simulation length and number of conformations) and content (\textit{e.g.}, lipid compositions and ion concentrations) enables analyses of rare phenomena beyond current possibilities for a single research group as demonstrated for cholesterol flip--flop events (Fig.~\ref{fig:flip-flops}) and water permeation through membranes (Fig.~\ref{fig:permeability}).
-This will also pave the way of applications of MD simulations in new fields as demonstrated here by analysing the anisotropic diffusion of water in membrane systems (Fig.~\ref{fig:permeability}) that is essential parameter in pharmacokinetic modeling and MRI imaging~\cite{nitsche19,topgaard20}.
-Furthermore, the NMRlipids databank-API paves the way for diverse machine learning applications to
+The open programmatic access to unprecedented amount of MD simulation data via NMRlipids databank-API enables wide range of novel data-driven analyses and machine learning applications.
+The analyses of cholesterol flip--flop events (Fig.~\ref{fig:flip-flops}) and water permeation through membranes (Fig.~\ref{fig:permeability}) demonstrate how large amount of accessible simulation data in terms of quantity (\textit{e.g.}, simulation length and number of conformations) and content (\textit{e.g.}, lipid compositions and ion concentrations) enable analyses of rare phenomena that are beyond the current possibilities for a single research group.
+Such analyses also pave the way for applications of MD simulations in new fields as demonstrated here by analysing the anisotropic diffusion of water in membrane systems (Fig.~\ref{fig:permeability}) that is essential parameter in pharmacokinetic modeling and MRI imaging~\cite{nitsche19,topgaard20}.
+Furthermore, the NMRlipids databank-API delivers access to a training set data that paves the way for diverse machine learning applications to
 %For example, a machine learning model predicting electron density profiles from form factors would be highly useful in interpretation of scattering experiments.
 %In addition, the data contained in the NMRlipids databank will enable training of more elaborated models to
-predict membrane properties using NMRlipids databank as a training set.
+predict membrane properties.
 Such applications are analogous to AlphaFold \cite{jumper21} and other tools that predict proteins structures from their sequence using artificial intelligence.
@@
     Selection of the best simulation model for a specific application & 
-    Best model for POPC lipids (Fig.~\ref{fig:quality}), headgroup conformations~\cite{bacle21}, 
+    Best model for membranes with PC and PE lipids (Fig.~\ref{fig:quality}), lipid headgroup conformations~\cite{bacle21}, 
     packing of PS~\cite{antila22b} and PE (Fig.~\ref{fig:POPCPOPEapls}) containing membranes. &
@@
     \\
-    Training and target data for coarse grained models & 
+    Training and target data for the development of coarse grained models & 
     Optimizing parameters of coarse grained models against NMRlipids databank, extracting continuum parameters for membranes. &
@@
 
-The main practical limitations in building open access databanks of molecular dynamics simulations have been the required commitment in long term support for hardware and software maintenance, and the lack of incentives for researchers to share the data.
-The NMRlipids databank circumvents these challenges with the overlay databank design and open collaboration approach developed in the NMRlipids project~\cite{botan15}.
-In this model, the file storage is distributed to publicly available stable repositories (\textit{Data layer} in Fig.~\ref{fig:overlay} A) and maintenance of the databank does not depend on individual scientists or groups because all its version controlled content is available with an open access licence (\textit{Databank layer} in Fig.~\ref{fig:overlay} A).
-Incentive to share the data is created in the NMRlipids open collaboration by offering authorship in published articles to the contributors following the NMRlipids project protocol~\cite{botan15}.
-The current NMRlipids databank contains only lipid bilayer simulations, but the concept can also be applied to other biomolecules, such as disordered proteins and membrane-protein systems, or other fields where similar barriers to establish publicly accessible databanks exist.
+Potential benefits of sharing MD simulation data are widely recognized by different stakeholders from scientists~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} to funders and journal staff, but data sharing with efficient tools for upcycling have been rare.
+The main barriers for such solutions have been the required commitment for the long term support of hardware and software, and the lack of incentives for researchers to share the data.
+The first issue is solved in the NMRlipids databank using the overlay design where raw data is distributed to already publicly available decentralized locations while the core of the databank is composed only of metadata stored in a version controlled git repository with an open access licence.
+On the other hand, the open collaboration approach developed in the NMRlipids project~\cite{botan15} creates incentives for sharing the data by offering authorship in published articles to the contributors.
+Advantages of such approach are demonstrated here for membrane simulations, but the concept can be applied also not only to other biomolecules, such as disordered proteins and membrane-protein systems, but also in other fields where similar barriers hinder the machine learning revolution, such as assignment of NMR spectra~\cite{klukowski22}.
 

### 2023-03-12 main.tex
Update text

@@
 Abbreviations and full names of all lipids present in the databank are listed in Table~\ref{tab:abbreviations}.
-Single component lipid membranes and binary mixtures, depicted in Fig.~\ref{fig:overlay}D, are currently most abundant in the NMRlipids databank, yet mixtures with up to five lipid types are available.
+Single component lipid membranes and binary mixtures, depicted in Fig.~\ref{fig:overlay}F, are currently most abundant in the NMRlipids databank, yet mixtures with up to five lipid types are available.
 Force fields used in simulations cover all the essential parameters commonly used in lipid simulations, see Fig.~\ref{fig:overlay}C and table~\ref{tab:ForceFields}, including also united atom and polarizable force fields.
@@
 Because {\it application layer} does not interfere with the {\it databank layer}, it can be freely extended by anyone for wide range of purposes.
-Here we demonstrate this with two examples: NMRlipids databank graphical user interface (NMRlipids databank-GUI) tool at~\url{www.databank.nmrlipids.fi/} and a repository exemplifying novel analyses utilizing NMRlipids databank as discussed below (\url{www.github.com/NMRLipids/DataBankManuscript}).
+This is demonstrated here with two examples: NMRlipids databank graphical user interface (NMRlipids databank-GUI) tool at~\url{www.databank.nmrlipids.fi/} and a repository exemplifying novel analyses utilizing NMRlipids databank as discussed below (\url{www.github.com/NMRLipids/DataBankManuscript}).
 More detailed description of the NMRlipids databank structure is available in the supplementary information.
@@
     C) Distribution of force fields in the simulations in the NMRlipids databank. References for each force field are given in table~\ref{tab:ForceFields}.
-    D) Currently available binary mixtures in the NMRlipids databank. 
+    D) Flowchart for performing an analysis of properties through all MD simulations in the NMRlipids databank using the API.
+    E) Flowchart for accessing results calculated from the NMRlipids databank and stored to the application layer.
+    F) Currently available binary mixtures in the NMRlipids databank. 
     }
@@
 \subsection{NMRlipids databank-GUI: graphical access to the MD simulation data}\label{section:access}
-NMRlipids databank-GUI provides easy access to the content of the NMRlipids databank via graphical user interface (GUI) that is available at~\url{www.databank.nmrlipids.fi/}.
-The NMRlipids databank-GUI has advanced searching options for system content and properties.
-The search results are ranked based on their quality evaluation against experimental data which enables rapid selection of best available simulations for specific systems.
-Systems can be also visualized and their properties can be compared with different simulations and experiments.
+NMRlipids databank-GUI, available at~\url{www.databank.nmrlipids.fi/}, provides easy access to the NMRlipids databank content via graphical user interface (GUI).
+Simulations can be searched based on their composition, used force field, temperature, properties and quality.
+The search results are ranked based on simulation quality evaluated against experimental data.
+NMRlipids databank-GUI enables also visualization of simulations and comparison of properties between different simulations and experiments.
+The NMRlipids databank-GUI enables rapid surveys of available simulation data, selection of best available simulations for specific systems based on ranking lists, and comparisons of basic properties between different types of membranes.
+Notably, GUI enables these operations to be performed by scientists with wide ranges of backgrounds who do not necessarily have expertise in programming or other means to access to MD simulation data.
 
 \subsection{NMRlipids databank-API: programmatic access to the MD simulation data}
-The NMRlipids databank-API, available at \url{www.github.com/NMRLipids/Databank}, enables automatic analysis of virtually any property from all simulations in the NMRlipids databank.
-This enables wide range of novel analyses that take the advantage on unprecendented amount MD simulation data with programmatic access.
-This is realized in the NMRlipids databank-API by delivering functions that enable straightforward implementation of the flowchart demonstrated in Fig.~\ref{fig:POPCPOPEapls}A.
-Iteration through the README.yaml files in the {\it databank layer} enables the access to the raw data of each simulation within the databank via the links stored in these files.
-The affiliation of universal molecule and atom names to the simulation specific names are delivered for each simulation in README.yaml and mapping files, which enables automatic analyses over all systems independently on the specific naming convention used in a simulation.
+The NMRlipids databank-API, available at \url{www.github.com/NMRLipids/Databank}, provides programmatic access to all simulation data in the NMRlipids databank.
+This enables wide range of novel data-driven applications from construction of machine learning models that predict membrane properties to automatic analyses of virtually any property from all simulations in the databank.
+%Wide range of novel analyses taking the advantage of programmatic access to unprecendented amount MD simulation data can be performed with the tools delivered by the NMRlipids databank-API. 
+Practical implementation of such analyses is illustrated in the flowchart in Fig.~\ref{fig:quality}D.
+First, raw data of each simulation can be accessed by iterating through the README.yaml files in the {\it databank layer} that contain links pointing to the locations of simulation trajectories and other relevant files.
+Simulations can be then automatically analyzed with the help of README.yaml and mapping files that affiliate specific naming conventions in each simulation to the universal molecule and atom names.
 Finally, the analysis results are stored using the same structure as in the {\it databank layer}.
@@
 are included in the {\it Databank layer}, unlimited amount of further analyses can be implemented by anyone in separate repositories in the {\it application layer}.
-{\it Application layer} repositories that are organized using the {\it Databank layer} structure can be straightforwardly accessed programmatically and further analyzed by implementing the flowchart demonstrated in Fig.~\ref{fig:POPCPOPEapls}B by using the NMRlipids databank-API functions.
+When {\it Application layer} repositories are organized by mimicking the {\it Databank layer} structure, they can be accessed programmatically and further analyzed using the tools in NMRlipids databank-API by implementing the flowchart demonstrated in Fig.~\ref{fig:quality}E.
 Novel analyses demonstrating the power of the NMRlipids databank in selecting best simulation models, analyses of rare phenomena and extending MD simulations to new fields are implemented in an {\it application layer} repository located at~\url{www.github.com/NMRLipids/DataBankManuscript}.
@@
 
-Divergent predictions for the lateral packing of POPC and POPE membranes from different force fields is illustrated in Fig.~\ref{fig:POPCPOPEapls}C.
-Because the membrane packing in terms of area per lipid correlates with the average order parameter of the $\textit{sn}$-1 chain (Fig.~\ref{fig:quality} F), simulations with the best quality for this part of POPE and POPC lipids are considered to give the most realistic results in Fig.~\ref{fig:POPCPOPEapls}C.
+Predictions for the lateral packing of POPC and POPE membranes in terms of area per lipid, one of the most important parameter used to characterize cellular membranes, diverge between different force fields as illustrated in Fig.~\ref{fig:quality}A.
+Because the membrane packing  correlates with the average order parameter of the $\textit{sn}$-1 chain (Fig.~\ref{fig:quality} F), simulations with the best quality for this part of POPE and POPC lipids are considered to give the most realistic results in Fig.~\ref{fig:POPCPOPEapls}C.
 To this end, the ranking based on C--H bond order parameter quality for the best performing POPC and POPE simulations in the NMRlipids databank are shown in Figs.~\ref{fig:quality}A and B, respectively.

### 2023-03-20 main.tex

@@
 Membrane lipid composition is implicated in diseases, such as cancer and neurodegenerative disorders, and therapeutics that affect membrane compositions are emerging~\cite{torres21}.
-However, biomembranes are often difficult to study experimentally because at biological conditions they are in disordered fluid state, and sample preparation can be arduous due to the complicated phase behaviour and rich interactions between lipids and other biomolecules.
+However, biomembranes are often difficult to study experimentally because they are complex mitures of lipids and proteins that are in disordered fluid state with complex phase behaviour at biological conditions.
 For those reasons, the detailed connections between complex lipid interactions and biological functions taking place in or around membranes remain poorly understood.
@@
 This enables wide range of novel data-driven applications from construction of machine learning models that predict membrane properties to automatic analyses of virtually any property from all simulations in the databank.
-%Wide range of novel analyses taking the advantage of programmatic access to unprecendented amount MD simulation data can be performed with the tools delivered by the NMRlipids databank-API. 
 Practical implementation of such analyses is illustrated in the flowchart in Fig.~\ref{fig:quality}D.
@@
 %
-%Understanding of how closely simulations resemble the reality is crucial for selecting best models for specific applications, assessing the reliability and significance of MD simulation results, and using atomistic simulations in parametrization of coarse grained models.
-However, undefined universal quality measures for MD simulations and the lack of a prevailing force field that would perform best in all simulations has led to many controversial results in membrane simulations~\cite{antila22b}.
-%
 %For example, OPLS3e parameters overcome CHARMM36 in structural quality for POPC but predicts overestimated ion binding~\cite{kurki22}.
@@
 %
-%Understanding of how closely simulations resemble the reality is crucial for selecting best models for specific applications,
 The quality of lipid bilayer MD simulations has to be carefully assessed in their applications to minize the detrimental consequences of artificial MD simulation results for lipid bilayers~\cite{antila22b}.
 This is possible, for example, by using C--H bond order parameters from NMR and X-ray scattering form factors \cite{ollila16}, although it requires comparisons between large number of simulations which is laborious even with collaborative approaches~\cite{botan15,catte16,antila19,bacle21}.
-The quantitative quality measures in the NMRlipids databank substantially streamline this process by enabling automatic ranking of lipid bilayer simulations based on their quality against experiments.
-Here we demonstrate this by selecting the best models for simulations of membranes with POPC and POPE lipids that are the two most biologically abundant neutral membrane lipids~\cite{vanmeer08}.
+To streamline this process we have defined quantitative quality measures for lipid bilayer simulations that enable automatic ranking of lipid bilayer simulations based on their quality against experiments.
+Conformational ensembles of individual lipid molecules are evaluated in the NMRlipids databank by calculating the probabilities for each C-H bond order parameter to locate within experimental error.
+Furthermore, the quality against X-ray scattering experiments is estimated from the distance of the first form factor minimum between the experiment and simulation.
+These measures can be used to evaluate membrane dimensions as the form factor minima locations and acyl chain order parameters correlate with the membrane lateral packing and thickness as shown in Figs.~\ref{fig:quality}G,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}.
+%while correlating well also with the membrane thickness and lateral packing as shown in Fig.~\ref{fig:quality}F. 
+%The X-ray scattering form factors are related to membrane dimensions \textit{via} the electron density profile. 
+Here we demonstrate how automatic simulation quality evaluation and NMRlipids databank-API enable rapid selection of the best models for simulations of membranes with POPC and POPE lipids that are the two most biologically abundant neutral membrane lipids~\cite{vanmeer08}.
 
 Predictions for the lateral packing of POPC and POPE membranes in terms of area per lipid, one of the most important parameter used to characterize cellular membranes, diverge between different force fields as illustrated in Fig.~\ref{fig:quality}A.
-Because the membrane packing  correlates with the average order parameter of the $\textit{sn}$-1 chain (Fig.~\ref{fig:quality} F), simulations with the best quality for this part of POPE and POPC lipids are considered to give the most realistic results in Fig.~\ref{fig:POPCPOPEapls}C.
+To select the most realistic models among these, we ranked the simulations according to the quality of C-H bond order parameters evaluated against experimental data in Fig.~\ref{fig:top50simulations}.
+Because the membrane packing correlates with the average order parameter of the $\textit{sn}$-1 chain (Fig.~\ref{fig:quality}G), force fields occurring in Fig.~\ref{fig:quality}A are ranked according to the $\textit{sn}$-1 chain quality in Figs.~\ref{fig:quality}B and C.
+Also the highest ranked simulation (POPC bilayer with OPLS3e parameters) is included for the reference.
+%for this part of POPE and POPC lipids are considered to give the most realistic results in Fig.~\ref{fig:POPCPOPEapls}C.
 To this end, the ranking based on C--H bond order parameter quality for the best performing POPC and POPE simulations in the NMRlipids databank are shown in Figs.~\ref{fig:quality}A and B, respectively.
 Direct comparisons with the experimental data are shown in Figs.~\ref{fig:quality}C--E and~\ref{fig:POPC_POPE_dataSI} for the most relevant simulations.
-Based on these rankings, simulations with Slipids parameters are the most realistic for POPC lipids with a clear difference to others, and 2\textsuperscript{nd} most realistic for POPE with only marginally lower quality than GROMOS-CKP simulations.
+Based on these rankings, Slipids simulation gives the most realistic prediction for POPC membrane in Fig.~\ref{fig:quality}A with a clear difference to others, and 2\textsuperscript{nd} most realistic prediction for POPE with only marginally lower quality than GROMOS-CKP simulations.
 Simulations with CHARMM36 and GROMOS-CKP parameters predict overly packed bilayer for POPC (overestimated order in Fig.~\ref{fig:POPC_POPE_dataSI}), while the area per lipid for POPC is overestimated in Lipid17.
 For POPE, CHARMM36 and Lipid17 predict membranes that are too packed.
-In conclusion, the quality evalution based on the NMRlipids databank suggests that Slipids parameters are the best available choice for simulations with PC and PE lipids, at least for the cases were membrane packing is relevant.
-
-The C--H bond order parameters are sensitive to conformational ensembles of individual lipid molecules~\cite{ollila16}, while correlating well also with the membrane thickness and lateral packing as shown in Fig.~\ref{fig:quality}F.
-The X-ray scattering form factors are related to membrane dimensions \textit{via} the electron density profile.
-In particular, locations of minima in these form factors are correlated with the membrane lateral packing and thickness as shown in Figs.~\ref{fig:quality}F,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}.
-For quantitative quality measures, probabilities for each order parameter to locate within experimental error, and the distance of the first form factor minimum from the experimental value are calculated from a simulation.
-%For detailed definitions of quality measures, see the supplementary information.
-Together with the API, these enable rapid and automatic quality evaluation of simulations in the NMRlipids databank, thereby promoting the selection of best simulations for specific applications without laborious force field evaluation, and automatic parametrization procedures for atomistic and coarse grained simulations.
+In conclusion, the quality evaluation based on the NMRlipids databank suggests that Slipids parameters are the best available choice for simulations with PC and PE lipids, at least for applications where membrane packing is relevant.
 
@@
 The resources delivered by the NMRlipids databank will support the development in both of these directions.
-The automatic quality evaluation and ranking of simulations against experimental data can be used to select the best parameters for simulations of large biomolecular complexes and to guide optimizations of new force field parameters for both atomistic and coarse grained simulations.
-The accuracy of simulations is becoming increasingly important due the accumulation of small errors when complexity and size of simulated systems are increasing.
+Automatic quality evaluation and ranking of simulations against experimental data enables the selection of best simulations for specific applications without laborious manual force field evaluation.
+This also streamlines automatic parametrization procedures for atomistic and coarse grained simulations.
+%can be used to select the best parameters for simulations of large biomolecular complexes and to guide optimizations of new force field parameters for both atomistic and coarse grained simulations. 
+Such practises fostering the accuracy of simulations are becoming increasingly important due the accumulation of small errors when complexity and size of simulated systems are increasing.
 
@@
 
-As a measure of equilibration quality for each trajectory, the ratio between the convergence time of first PC and the trajectory length was calculated for each lipid type individually.
+As a measure of equilibration quality for each trajectory, the ratio between the convergence time of first PC and the trajectory length, $\tau_{\rm rel} = \tau_{\text{convergence}}/\tau_{\text{sim}}$, was calculated for each lipid type individually.
 Sterols were excluded from the analysis.

### 2023-03-24 SI.tex
Update text
Update figures

--- 2023-01-31 SI.tex
@@
 \section{Finding the best models for PC and PE mixtures}
+
+\begin{figure}
+    \centering
+    \includegraphics[width = 110mm]{Figures/totalrankingCOMPL.pdf}
+    \caption{Top 50 simulations ranked based on order parameter quality against experiments.}
+    \label{fig:top50simulations}
+\end{figure}
 
@@
 \end{figure}
+
+\begin{table}[!h]
+    \centering
+    %\begin{tabular}{c c c c}
+    \begin{tabular}{p{3.0cm}  p{5.0cm}  p{3.0cm}  p{4.0cm}}
+        Code & Function & Input & Output \\
+        \hline
+        \multicolumn{4}{c}{ {\bf Databank layer}}\\
+        \hline
+        \multicolumn{4}{c}{ {\bf Scripts/BuildDatabank/ }}\\
+         AddData.py & Create README.yaml from user given information in info.yaml  & info.yaml & README.yaml \\
+         searchDATABANK.py & Pair simulations with available experimental data  & README.yaml files & Updated README.yaml files \\
+         QualityEvaluation.py & Quality evaluate simulations that are paired with experimental data  & README.yaml files & [lipid\_name]\_OrderParameters\_quality.json, [lipid\_name]\_FragmentQuality.json, system\_quality.json, FormFactorQuality.json \\
+         \multicolumn{4}{c}{ {\bf Scripts/AnalyzeDatabank/ }}\\
+         calcOrderParameters.py & Calculate C-H bond order parameters for all lipids in the simulation & README.yaml files & [lipid\_name]\_OrderParameters.json \\
+         calcAPL.py & Calculate area per lipid as a function of time & README.yaml files & apl.json \\
+         calc\_FormFactors.py & Calculate x-ray scattering form factors & README.yaml files & FormFactor.json \\
+         calc\_thickness.py & Calculate membrane thickness & README.yaml files & thickness.json \\
+         template.py $\qquad$  template.ipynb & Minimal code to loop over simulations that can be used as a template for new analyses. & README.yaml files & \\
+
+    \end{tabular}
+    \caption{List of relevant codes used to build the databank and perform analyses in the {\it databank layer} available at~\url{https://github.com/NMRLipids/Databank/}.}
+    \label{tab:codes}
+\end{table}
+
+\begin{table}[!h]
+    \centering
+    %\begin{tabular}{c c c c}
+    \begin{tabular}{p{3.5cm}  p{4.5cm}  p{3.0cm}  p{4.0cm}}
+        Code & Function & Input & Output \\
+        \hline
+        \multicolumn{4}{c}{ {\bf Application layer}}\\
+        \hline
+        %\multicolumn{4}{c}{ {\bf Scripts/BuildDatabank/ }}\\
+        AreaPerLipidAnd ThicknessCorrelations.ipynb  & Analyze correlations between results in the {\it databank layer}  & Results in the {\it databank layer} & Figs.~\ref{fig:quality}G and~\ref{fig:QualityCorrelationsSI} \\
+        FlipFlop.py & Calculate flip-flop rates of lipids in all simulations & README.yaml & flipflop.dat files at {\it /Data/Flipflops/}. \\
+        plotFlipFlop.ipynb & Plots how flip-flop rates depend on membrane properties & README.yaml and flipflop.dat files at {\it /Data/Flipflops/}. & Fig.~\ref{fig:flip-flops} \\
+        calcMD-PERMEATION.py  & Calculate permeation rate of water molecules through bilayers from all simulations. & README.yaml & Counting\_events.txt files at {\it /Data/MD-PERMEATION/} \\
+        calcWATERdiffusion.py & Calculate water lateral diffusion along membrane surface. & README.yaml & WATERlateralMSD.xvg files at {\it /Data/WATERdiffusion/} \\
+        plotWaterPermeation.ipynb & Plots how water permeation and diffusion depends on membrane properties.
+& README.yaml, Counting\_events.txt\,at {\it /Data/MD-PERMEATION/} and\,WATERlateralMSD.xvg files\,at {\it /Data/WATERdiffusion/} & Figs.~\ref{fig:permeability},~\ref{fig:permeationSI}, and~\ref{fig:diffusionSI}.
+\\
+        plotSimulation.ipynb & Plot results from a simulation together with the experimental data if available.
+& README.yaml files and results in the {\it databank layer} & Figs~\ref{fig:quality}D-F and ~\ref{fig:POPC_POPE_dataSI}.
+    \end{tabular}
+    \caption{Examples of codes that analyze membrane properties from the databank in an {\it application layer} available at~\url{https://github.com/NMRLipids/DataBankManuscript/}.}
+    \label{tab:codesApplication}
+\end{table}
 
@@
 
+\begin{table}[p]
+    \centering
+    \begin{tabular}{l}
+    Force field name and references \\
+    \hline
+    CHARMM36~\cite{klauda10}\\
+    Slipids~\cite{jambeck12,jambeck12b,jambeck2012another,ermilova16,grote20}  \\
+    MacRog~\cite{Kulig15b}  \\
+    Amber Lipid14/17~\cite{dickson14,dickson22}  \\
+    Charmm-Drude~\cite{li2017drude}      \\
+    ECClipids~\cite{melcr18,melcr20,bacle21}  \\
+    GROMOS-CKP~\cite{Chandrasekhar03,kukol09,piggot12}  \\
+    Berger~\cite{berger97}  \\
+    ECC-CHARMM36~\cite{nencini2019development}  \\
+    Orange~\cite{catte16}  \\
+    Poger~\cite{poger10}  \\
+    GROMOS 43A1-S3~\cite{chiu09}  \\ 
+    GAFFlipid~\cite{dickson12}  \\
+    OPLS3e~\cite{roos19}  \\
+    Ulmschneider~\cite{Ulmschneider09}  \\
+    Chiu Gromos~\cite{chiu09}  \\
+    AMOEBA~\cite{chu2018polarizable} \\
+    \end{tabular}
+    \caption{List of current force fields used in simulations in the databank with references.}
+    \label{tab:ForceFields}
+\end{table}
+
 \clearpage
@@
 \section{NMR experiments}
-Acyl chain order parameters of POPE and POPG were analyzed from the experiments that were previously used to determine headgroup order parameters~\cite{bacle21}.
+Acyl chain order parameters of POPE (Figs.~\ref{POPEexp1} and~\ref{POPEexp2}) and POPG (Figs.~\ref{POPGexp1} and~\ref{POPGexp2}) were analyzed from the same data that were previously recorded to determine headgroup order parameters~\cite{bacle21}.
 The analysis of the crowded spectral region at 29-31 ppm was based on the previous assignment reported for POPC membranes~\cite{ferreira13}.
+To measure the order parameters for DOPC (Fig.~\ref{fig:DOPCexp}), the sample was prepared and experiments performed similarly to previous studies~\cite{bacle21}.
 
@@
   \label{POPGexp2}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{Figures/DOPCfig.pdf}
+    \caption{Determination of DOPC order parameters from a R-PDLF spectrum measured at a magic angle spinning frequency of 5.15 kHz.
+(A) $^{13}$C rINEPT spectrum with peak assignment.
+The labels used are shown in the chemical structure of DOPC.
+The chemical shift of the methyl groups was defined as 13.8 ppm.
+(B) Contour plot of the R-PDLF spectrum for the crowded spectral region.
+The assignment was based on a previous assignment reported for POPC membranes~\cite{ferreira13}.
+(C) C-H order parameters of DOPC.}
+    \label{fig:DOPCexp}
 \end{figure}

### 2023-03-24 main.tex
Update text

@@
 Finally, the analysis results are stored using the same structure as in the {\it databank layer}.
+Examples of codes that utilize NMRlipids databank-API and a template for new analyses can be found from locations listed in table~\ref{tab:codes}.
 
@@
 Novel analyses demonstrating the power of the NMRlipids databank in selecting best simulation models, analyses of rare phenomena and extending MD simulations to new fields are implemented in an {\it application layer} repository located at~\url{www.github.com/NMRLipids/DataBankManuscript}.
+The related codes are listed in table~\ref{tab:codesApplication}.
 
@@
 These measures can be used to evaluate membrane dimensions as the form factor minima locations and acyl chain order parameters correlate with the membrane lateral packing and thickness as shown in Figs.~\ref{fig:quality}G,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}.
-%while correlating well also with the membrane thickness and lateral packing as shown in Fig.~\ref{fig:quality}F. 
-%The X-ray scattering form factors are related to membrane dimensions \textit{via} the electron density profile. 
+In addition, the ergodicity of conformational sampling of lipids is estimated by calculating the convergence time of the slowest principal component divided by the length of a simulation.
 Here we demonstrate how automatic simulation quality evaluation and NMRlipids databank-API enable rapid selection of the best models for simulations of membranes with POPC and POPE lipids that are the two most biologically abundant neutral membrane lipids~\cite{vanmeer08}.
@@
 Predictions for the lateral packing of POPC and POPE membranes in terms of area per lipid, one of the most important parameter used to characterize cellular membranes, diverge between different force fields as illustrated in Fig.~\ref{fig:quality}A.
-To select the most realistic models among these, we ranked the simulations according to the quality of C-H bond order parameters evaluated against experimental data in Fig.~\ref{fig:top50simulations}.
-Because the membrane packing correlates with the average order parameter of the $\textit{sn}$-1 chain (Fig.~\ref{fig:quality}G), force fields occurring in Fig.~\ref{fig:quality}A are ranked according to the $\textit{sn}$-1 chain quality in Figs.~\ref{fig:quality}B and C.
+To select the most realistic models among these, we first ranked all simulations in the databank according to the quality of C-H bond order parameters evaluated against experimental data in Fig.~\ref{fig:top50simulations}.
+From these results, we then picked the force fields occurring in Fig.~\ref{fig:quality}A and ranked them according to the quality of $\textit{sn}$-1 chain in Figs.~\ref{fig:quality}B and C.
+Simulations with $\tau_{\rm rel}$ clearly above one were discarded.
 Also the highest ranked simulation (POPC bilayer with OPLS3e parameters) is included for the reference.
-%for this part of POPE and POPC lipids are considered to give the most realistic results in Fig.~\ref{fig:POPCPOPEapls}C.
-To this end, the ranking based on C--H bond order parameter quality for the best performing POPC and POPE simulations in the NMRlipids databank are shown in Figs.~\ref{fig:quality}A and B, respectively.
-Direct comparisons with the experimental data are shown in Figs.~\ref{fig:quality}C--E and~\ref{fig:POPC_POPE_dataSI} for the most relevant simulations.
-Based on these rankings, Slipids simulation gives the most realistic prediction for POPC membrane in Fig.~\ref{fig:quality}A with a clear difference to others, and 2\textsuperscript{nd} most realistic prediction for POPE with only marginally lower quality than GROMOS-CKP simulations.
-Simulations with CHARMM36 and GROMOS-CKP parameters predict overly packed bilayer for POPC (overestimated order in Fig.~\ref{fig:POPC_POPE_dataSI}), while the area per lipid for POPC is overestimated in Lipid17.
-For POPE, CHARMM36 and Lipid17 predict membranes that are too packed.
+Because the membrane packing correlates with the average order parameter of the $\textit{sn}$-1 chain (Fig.~\ref{fig:quality}G), rankings in Figs.~\ref{fig:quality}B and C were used to select the simulations giving most realistic results in~Fig.~\ref{fig:quality}A.
+Also direct comparisons with the experimental data for the most relevant simulations are shown in Figs.~\ref{fig:quality}C--E and~\ref{fig:POPC_POPE_dataSI}.
+Based on these rankings, Lipid17 and Slipids simulations give the most realistic predictions for POPC membrane, while simulations with CHARMM36 and GROMOS-CKP parameters predict overly packed bilayers (overestimated order in Fig.~\ref{fig:POPC_POPE_dataSI}).
+On the other hand, GROMOS-CKP and Slipids give the most realistic results for POPE, while CHARMM36 and Lipid17 predict membranes that are too packed.
 In conclusion, the quality evaluation based on the NMRlipids databank suggests that Slipids parameters are the best available choice for simulations with PC and PE lipids, at least for applications where membrane packing is relevant.
@@
     \includegraphics[width=\linewidth]{Figures/quality5.pdf}
-    \caption{ A) Best simulations ranked based on overall order parameter quality.
-    B) Best simulations ranked based on the overall order parameter quality of POPE lipid. 
-    C)--E) Evaluation against experimental (NMR order parameters and X-ray scattering ) data exemplified for a simulation with the best overall order parameter quality (C), the best quality for POPE lipid (D), and the headgroup quality for POPE (E).
-    F) Scatter plots and Pearson correlation coefficients, $r$, for the membrane area per lipid, thickness, first minimum of X-ray scattering form factor and average order parameter of the {\textit sn}-1 acyl chain extracted from the NMRlipids databank.
-All correlation coefficients have p-value below 0.001 as indicated by $^{***}$.
+    \caption{ A) Area per lipids of POPC and POPE lipid bilayers predicted by different force fields at 310\,K in simulations that are available in the NMRlipids databank.
+The points from the best performing simulations based on rankings in B and C are surrounded by black circles.
+    B) Best POPC simulations ranked based on {\it sn}-1 acyl chain order parameter quality.
+    C) Best POPE simulations ranked based on {\it sn}-1 acyl chain order parameter quality. 
+    D)--F) Direct comparison against experimental (NMR order parameters and X-ray scattering ) data exemplified for a simulation with the best overall order parameter quality (D), the best quality for POPE lipid (E), and the headgroup quality for POPE (F).
+    G) Scatter plots and Pearson correlation coefficients, $r$, for the membrane area per lipid, thickness, first minimum of X-ray scattering form factor and average order parameter of the {\textit sn}-1 acyl chain extracted from the NMRlipids databank.
+All correlation coefficients have p-value below 0.001 as indicated by~$^{***}$.
 For more correlations see Fig.~\ref{fig:QualityCorrelationsSI}.
@@
 
-Using workflow depicted in Fig.~\ref{fig:POPCPOPEapls}A, we first calculated the cholesterol flip--flop rates from all the simulations available in the NMRlipids databank.
+Using workflow depicted in Fig.~\ref{fig:POPCPOPEapls}D, we first calculated the cholesterol flip--flop rates from all the simulations available in the NMRlipids databank.
 Flip--flops were observed for cholesterol, % (634 events in 77 different simulations),
@@
 
-Nevertheless, we used the workflow depicted in Fig.~\ref{fig:POPCPOPEapls}B to analyse how flip--flop rates calculated from the NMRlipids databank depend on membrane properties.
+Nevertheless, we used the workflow depicted in Fig.~\ref{fig:POPCPOPEapls}E to analyse how flip--flop rates calculated from the NMRlipids databank depend on membrane properties.
 Cholesterol flip--flop rates and their histograms are plotted as a function of membrane thickness, lateral density and order in Figs.~\ref{fig:flip-flops} B--D.
@@
 Clear dependencies of permeability on charged lipids, cholesterol, POPE, or hydration level were not observed in Fig.~\ref{fig:permeationSI} in the supplementary information.
-%, although weak decrease for the latter two may be visible.
 %
-%
-%
-%%Several models on how water dynamics depends on membrane properties have been proposed based on different experimental and %%theoretical methods, yet consensus 
-%%%on how water permeability depends on membrane properties, such as area per lipid and thickness, or molecular composition, 
-%%has not been reached~\cite{mathai07,nitsche13,nitsche16,shinoda16,venable19,frallicciardi22}.
+%%Several models on how water dynamics depends on membrane properties have been proposed~\cite{mathai07,nitsche13,nitsche16,shinoda16,venable19,frallicciardi22}.
 
@@
 The resulting perpendicular diffusion coefficients are approximately five orders of magnitude smaller than lateral diffusion coefficients of water (Figs.~\ref{fig:permeability} G and H), which is at the upper limit of the anisotropy estimated from the experimental data~\cite{nitsche19}.
-%Large anisotropy values are understandable as simulations give slightly slower permeation rates and higher lateral diffusion rates than experiments. 
 Significant increase in the diffusion anisotropy with membrane packing is observed, as $D_\mathrm{\perp}/D_{||}$ drifts away from one with decreasing area per lipid and increasing thickness in Figs.~\ref{fig:permeability} G and H.
@@
 This also streamlines automatic parametrization procedures for atomistic and coarse grained simulations.
-%can be used to select the best parameters for simulations of large biomolecular complexes and to guide optimizations of new force field parameters for both atomistic and coarse grained simulations. 
 Such practises fostering the accuracy of simulations are becoming increasingly important due the accumulation of small errors when complexity and size of simulated systems are increasing.
@@
 These possibilities are particularly valuable for scientist who do not typically have access to large scale MD simulation data.
-%and for analyses of rare phenomena that require exceptional amounts of simulation data.
-%On the other hand, universal naming convention enables systematic comparisons between simulations having different model parameters and thereby different naming conventions.
-%
-%The NMRlipids databank will be of great benefit to scientists involved in MD simulations.
-They will be able to rapidly evaluate the quality of MD simulations in order to filter out potentially misleading results or facilitate force field parameter development~\cite{antila22b}.
-The selection of best models for a specific application is demonstrated for PC/PE lipid mixtures in Fig.~\ref{fig:quality}, and previously for PC/PS lipid mixtures~\cite{antila22b}, and lipid headgroups~\cite{bacle21}.
-%
 Different types of expected applications of the NMRlipids databank in wide range of fields are listed in Table~\ref{tab:applications}, yet the scope of such applications
@@
 \subsection{Structure of the databank}
-Structure of the NMRlipids databank is illustrated more detailed in Fig.~\ref{DatabankStructure} in the supplementary information.
-The core content of the databank (\textit{Databank layer} in Fig.~\ref{fig:overlay}) locates as a git repository at \url{https://github.com/NMRLipids/Databank/} and is permanently stored in Zenodo repository (\url{www.zenodo.org})~\cite{??}.
-Whenever a specific file is referred here, the file path within the NMRlipids databank repository is given.
+The overlay structure designed for the NMRlipids databank is composed of three layers illustrated in Fig.~\ref{fig:overlay}A.
+The {\it data layer} contains raw data that can be distributed to publicly available servers such as Zenodo (\url{www.zenodo.org}).
+The core content of the databank locates in the {\it databank layer} which is a git repository at \url{https://github.com/NMRLipids/Databank/} and is also permanently stored in Zenodo repository (\url{www.zenodo.org})~\cite{??}.
+The Essential information of each simulation is stored in a human and machine readable README.yaml file located at subfolders of \texttt{/Data/Simulations} folder in the {\it databank layer} repository.
+Each subfolder has a unique name constructed based on hash code of trajectory and topology files of each simulation.
+The README.yaml files in these folders contain access to all information that are needed for further analyses of simulations, such as links to the raw data and affiliations of universal molecule and atom names.
+The content of these files is described in detail in Table~\ref{tab:READMEkeys} in the supplementary information.
+Results from analyses of basic membrane properties (area per lipid, thickness, C-H bond order parameters, x-ray scattering form factors and relaxation of principal components) are stored in the same folders with README.yaml files.
+Experimental data used for ranking is stored in \texttt{/Data/experiments} folder, results from such ranking in \texttt{/Data/Ranking} and  relevant scripts in \texttt{/Scripts/} folder in the {\it databank layer} repository.
+%Raw MD simulation data are stored in external publicly available and stable repositories (\textit{Data layer} in Fig.~\ref{fig:overlay}), such as Zenodo (\url{www.zenodo.org}), from where it can be downloaded whenever needed using the links in README.yaml files.
 The scripts in the NMRlipids databank are mainly written in Python and many of them utilize the MDAnalysis module~\cite{gowers2019mdanalysis,michaud2011mdanalysis}.
-
-Essential information of each simulation in the NMRlipids databank is stored in a human and machine readable README.yaml file located at \texttt{/Data/Simulations} in the NMRlipids databank repository.
-These files contain access to all information that are needed for further analyses of simulations.
-The content of these files is described in detail in Table~\ref{tab:READMEkeys} in the supplementary information.
-Raw MD simulation data are stored in external publicly available and stable repositories (\textit{Data layer} in Fig.~\ref{fig:overlay}), such as Zenodo (\url{www.zenodo.org}), from where it can be downloaded whenever needed using the links in README.yaml files.
+The databank structure is illustrated more detailed in Fig.~\ref{DatabankStructure} in the supplementary information.
+Whenever specific files or folders are referred here, they locate at {\it datbank layer} repository unless stated otherwise.
 
@@
 The NMRlipids databank is open for additions of simulation data by anyone.
-The first step is to create an info.yaml file containing the information that needs to be manually entered as listed in Table~\ref{tab:READMEkeys}.
-This file can be then added into \texttt{/Scripts/BuildDatabank/info\_files} folder in the NMRlipids databank repository via pull requests.
-After the pull request is manually accepted, the rest of the information for the README.yaml file, listed in Table~\ref{tab:READMEkeys}, will be automatically extracted using the \texttt{/Scripts/BuildDatabank/AddData.py} script.
-Currently, the NMRlipids databank is composed of simulations found from Zenodo repository with an appropriate licence.
+The list information that a contributor has to deliver is given in Table~\ref{tab:READMEkeys}.
+The rest of the information to be stored in README.yaml files, also listed in Table~\ref{tab:READMEkeys}, will be automatically extracted using the \texttt{/Scripts/BuildDatabank/AddData.py} script.
+In practise, the manually entered data is first stored into an info.yaml file that is then added into \texttt{/Scripts/BuildDatabank/info\_files} folder via pull request.
+To avoid ineligible entries and minimize human errors, the pull requests are monitored before the acceptance and generation of README.yaml files.
+Currently, the NMRlipids databank is composed of simulations that are found from the Zenodo repository with an appropriate licence.
 Most, but not all, of these trajectories originate from previous NMRlipids projects~\cite{botan15,catte16,antila19,bacle21}.
@@
 The keys and their descriptions for the experimental data are given in Table~\ref{tab:READMEkeysEXP}.
-NMR data currently in the NMRlipids databank are taken from Refs.~\citenum{scherer87,ferreira13,antila19,melcr20,bacle21} and X-ray scattering data from Refs.~\citenum{Kucerka08a,kucerka11,pan12b,pan14,kucerka15}.
-In addition, some previously unpublished NMR and X-ray scattering data are used.
-These are measured with previously established methods as described in the supplementary information.
+NMR data currently in the NMRlipids databank are taken from Refs.~\citenum{scherer87,ferreira13,antila19,melcr20,bacle21} and X-ray scattering data from Refs.~\citenum{Kucerka08a,kucerka11,pan12b,pan14,kucerka15,NMRlipidsIII}.
+In addition, previously unpublished NMR data for POPE, POPG and DOPC was acquired as described in the supplementary information (Figs.~\ref{POPEexp1}-\ref{fig:DOPCexp}) and contributed to the databank.
+
+\subsection{Analysing simulations}
+In practise, simulation data in the NMRlipids databank can be currently analyzed by first downloading the data using the information in README.yaml and then performing the analysis on local computer utilising the universal naming conventions for molecules and atoms therein.
+This procedure is illustrated in Fig.~\ref{fig:overlay}D and tables~\ref{tab:codes} and~\ref{tab:codesApplication} list examples of codes that perform such analyses in practise.
 

### 2023-03-24 main.tex
Update pdfs

@@
 
-Using workflow depicted in Fig.~\ref{fig:POPCPOPEapls}D, we first calculated the cholesterol flip--flop rates from all the simulations available in the NMRlipids databank.
+Using workflow depicted in Fig.~\ref{fig:overlay}D, we first calculated the cholesterol flip--flop rates from all the simulations available in the NMRlipids databank.
 Flip--flops were observed for cholesterol, % (634 events in 77 different simulations),
@@
 
-Nevertheless, we used the workflow depicted in Fig.~\ref{fig:POPCPOPEapls}E to analyse how flip--flop rates calculated from the NMRlipids databank depend on membrane properties.
+Nevertheless, we used the workflow depicted in Fig.~\ref{fig:overlay}E to analyse how flip--flop rates calculated from the NMRlipids databank depend on membrane properties.
 Cholesterol flip--flop rates and their histograms are plotted as a function of membrane thickness, lateral density and order in Figs.~\ref{fig:flip-flops} B--D.
@@
 
-To this end, we first calculated the water permeability through membranes from all simulations in the NMRlipids databank using the workflow depicted in Fig.~\ref{fig:POPCPOPEapls}A.
+To this end, we first calculated the water permeability through membranes from all simulations in the NMRlipids databank using the workflow depicted in Fig.~\ref{fig:overlay}D.
 The resulting non-zero values range between 0.3 and 322\,\textmu{}m/s with the mean and median of 14\,\textmu{}m/s and 8\,\textmu{}m/s, respectively.
 These values agree with the previously reported simulation results~\cite{venable19,camilo2022}, but are on average slightly larger than experimental values reported for PC lipids in the liquid crystalline phase,~0.19--0.33\,\textmu{}m/s~\cite{jansen95}.
-Using the workflow depicted in Fig.~\ref{fig:POPCPOPEapls}B, we then plotted the observed permeabilities and their histogrammed values in Figs.~\ref{fig:permeability}B--E as a function of temperature, membrane thickness, area per lipid, and acyl chain order.
+Using the workflow depicted in Fig.~\ref{fig:overlay}E, we then plotted the observed permeabilities and their histogrammed values in Figs.~\ref{fig:permeability}B--E as a function of temperature, membrane thickness, area per lipid, and acyl chain order.
 As expected, the permeability increases with the temperature, giving an average energy barrier of 17\,$\pm$\,4\,$k_\mathrm{B}T$ for the water permeation from the Arrhenius plot in Fig.~\ref{fig:permeability}B.
@@
 
-To analyze how water diffusion anisotropy depends on membrane properties in a multi-lamellar lipid bilayer system, we calculated the water diffusion parallel to the membrane surface from all simulations in the NMRlipids databank using the workflow depicted in Fig.~\ref{fig:POPCPOPEapls}A.
-The parallel diffusion coefficient of water, $D_{||}$, decreases with reduced hydration and increases with the temperature, but dependencies on membrane area per lipid, thickness, or fraction of charged lipids were not observed in Figs.~\ref{fig:permeability} and~\ref{fig:diffusionSI} when the workflow depicted in Fig.~\ref{fig:POPCPOPEapls}B was used.
+To analyze how water diffusion anisotropy depends on membrane properties in a multi-lamellar lipid bilayer system, we calculated the water diffusion parallel to the membrane surface from all simulations in the NMRlipids databank using the workflow depicted in Fig.~\ref{fig:overlay}D.
+The parallel diffusion coefficient of water, $D_{||}$, decreases with reduced hydration and increases with the temperature, but dependencies on membrane area per lipid, thickness, or fraction of charged lipids were not observed in Figs.~\ref{fig:permeability} and~\ref{fig:diffusionSI} when the workflow depicted in Fig.~\ref{fig:overlay}E was used.
 Simulation results are close to the experimental values with low hydration levels in Fig.~\ref{fig:permeability} F, but increase to approximately two times higher than the experimental value for bulk water diffusion value (3.1$\cdot$\,10$^{-9}$\,m$^2$/s at 313\,K~\cite{khakimov08}) with high hydration levels.
@@
     Best model for membranes with PC and PE lipids (Fig.~\ref{fig:quality}), lipid headgroup conformations~\cite{bacle21}, 
-    packing of PS~\cite{antila22b} and PE (Fig.~\ref{fig:POPCPOPEapls}) containing membranes. &
+    packing of PS~\cite{antila22b} and PE (Fig.~\ref{fig:quality}) containing membranes. &
     Scientists using MD simulations \\

### 2023-03-25 main.tex
Update text

@@
 \subsection{Analysing simulations}
-In practise, simulation data in the NMRlipids databank can be currently analyzed by first downloading the data using the information in README.yaml and then performing the analysis on local computer utilising the universal naming conventions for molecules and atoms therein.
-This procedure is illustrated in Fig.~\ref{fig:overlay}D and tables~\ref{tab:codes} and~\ref{tab:codesApplication} list examples of codes that perform such analyses in practise.
+In practise, simulations in the NMRlipids databank can be analyzed by executing a program that loops over README.yaml files in the {\it databank layer}, downloads the data using the information in README.yaml files, and then performs the desired analysis on a local computer utilising the universal naming conventions for molecules and atoms defined in README.yaml and mapping files.
+This procedure is illustrated in Fig.~\ref{fig:overlay}D and examples of codes that perform such analyses in practise are listed in tables~\ref{tab:codes} and~\ref{tab:codesApplication}.
+
+\subsection{Principal Components Analysis of equilibration of simulations}
+To estimate how well conformational ensembles of lipids are converged in trajectories, the Principal Components Analysis (PCA) following the PCALipids protocol was used \cite{buslaev16,buslaev2020principal}.
+%, the lipids of a specific type were first selected, and trajectories of individual lipids were concatenated. 
+To this end, each lipid configuration was first aligned to the average structure of that lipid type, and PCA analysis was then applied on the cartesian coordinates of all heavy atoms of the lipid.
+Because the motions along the first, major, principal component are the slowest ones \cite{buslaev16}, the equilibration of each lipid type was estimated from the ratio between the distribution convergence of the trajectories projected on the first PC and the trajectory length, $\tau_{\rm rel} = \tau_{\text{convergence}}/\tau_{\text{sim}}$~\cite{buslaev16, buslaev2020principal}.
+If $\tau_{\rm rel} < 1$, simulations can be considered to be sufficiently long for lipid molecules to sample their conformationla ensembles, while in simulations with $\tau_{\rm rel} > 1$ individual molecules may not have fully sample their conformational ensembles.
+Rigid molecules that do not exhibit significant conformational fluctuations, such as sterols, were excluded from the analysis.
+%The suggested way to estimate the characteristic timescales needed to adequately sample the configurational space of individual lipids is to calculate the distribution convergence of the trajectories projected on the first PC\cite{buslaev16}.
+In practise, the distribution convergence times were calculated utilising its linear dependence on autocorrelation decay times,
+%\begin{equation}
+    $\tau_{\text{convergence}} = k*\tau_{\text{autocorrelation}}$,
+%\end{equation}
+because calculation of autocorrelation decay times is faster and computationally more stable than direct calculation of distribution convergence times~\cite{buslaev16, buslaev2020principal}.
+The empirical coefficient $k$ is equal to $k = 49$ and was calculated based on the analysis of 8 trajectories with the length of more than 200\,ns including simulations of POPC, POPS, POPE, POPG, and DPPC with CHARMM36 force fields.
+%All simulations, used for $k$ calculation, were at least 200 ns long.
+Because the coefficient does not depend on the force field~\cite{buslaev16}, the value determined from these CHARMM36 simulations ,can be used for all simualtions in the Databank.
+The script that calculates the equilibration of lipids is available at \texttt{Scripts/BuildDatabank/NMRPCA\_timerelax.py} in the NMRlipids databank repository.
+The resulting values are stored in files named \texttt{eq\_times.json} at folders in \texttt{/Data/Simulations} in the NMRlipids databank repository.
 
@@
 As in previous NMRlipids publications, the order parameters were first calculated separately for each lipid and the standard error of the mean over different lipids was used as the error estimate~\cite{botan15}.
+However, order parameters for simulations with $\tau_{\rm rel} > 1$ may be influenced by the starting structure and thereby their error bars may be underestimated.
 The script that calculates C--H bond order parameters from all simulations in the NMRlipids databank is available at \texttt{/Scripts/AnalyzeDatabank/calcOrderParameters.py} in the NMRlipids databank repository.
@@
 \subsection{Calculation of a bilayer area per lipid and thickness}
-Area per lipids of bilayers were calculated by dividing the time-averaged area of the simulation box with the total number of lipid and surfactant molecules (see the list in table~\ref{tab:abbreviations}) in the simulation.
+Area per lipids of bilayers were calculated by dividing the time-averaged area of the simulation box with the total number of lipid and surfactant molecules in the simulation (see table~\ref{tab:abbreviations} for the list of molecules considered as lipids or surfactants).
 The script that calculates area per lipids from all simulations in the NMRlipids databank repository is available at \texttt{Scripts/AnalyzeDatabank/calcAPL.py} in the NMRlipids databank repository.
@@
 Thicknesses of lipid bilayers were calculated from the intersections of lipid and water electron densities.
-The script that calculates thickness of all simulations in the NMRlipids databank is available at \texttt{Scripts/AnalyzeDatabank/calc\_thickness.py} in the NMRlipids databank repository.
+The script that calculates thicknesses of all simulations in the NMRlipids databank is available at \texttt{Scripts/AnalyzeDatabank/calc\_thickness.py} in the NMRlipids databank repository.
 The resulting thicknesses are stored in files named \texttt{thickness.json} at folders in \texttt{/Data/Simulations} in the NMRlipids databank repository.
-
-\subsection{Principal Components Analysis of equilibration of simulations}
-To estimate how well the provided trajectories are converged, the Principal Components Analysis (PCA) was used \cite{buslaev16,buslaev2020principal}.
-Following the PCALipids \cite{buslaev2020principal} protocol, the lipids of a specific type were first selected, and trajectories of individual lipids were concatenated.
-Next, each lipid configuration was aligned to the average lipid structure.
-Afterwards, the PCA was applied on the cartesian coordinates of all heavy atoms of the lipid.
-Since it was shown before, that the motions along the first, major, principal component (PC) are the slowest ones \cite{buslaev16}, the equilibration of individual lipids was estimated based on the dynamics of the first PC.
-The suggested way to estimate the characteristic timescales needed to adequately sample the configurational space of individual lipids is to calculate the distribution convergence of the trajectories projected on the first PC\cite{buslaev16}.
-However, due to the linear dependence between autocorrelation decay times and distribution convergence times, faster calculation times and higher computational stability of autocorrelation decay times\cite{buslaev16, buslaev2020principal}, the distribution convergence timescales were calculated from autocorrelation decay times:
-\begin{equation}
-    \tau_{\text{convergence}} = k*\tau_{\text{autocorrelation}}
-\end{equation}
-
-The empirical coefficient $k$ is equal to $k = 49$ and was calculated based on the analysis of 8 trajectories from the Databank and included simulations of POPC, POPS, POPE, POPG, and DPPC with CHARMM36 force fields.
-All simulations, used for $k$ calculation, were at least 200 ns long.
-As was shown previously \cite{buslaev16}, the coefficient does not depend on the force field, thus only CHARMM36 simulations, as the most frequent in the Databank, were used.
-
-As a measure of equilibration quality for each trajectory, the ratio between the convergence time of first PC and the trajectory length, $\tau_{\rm rel} = \tau_{\text{convergence}}/\tau_{\text{sim}}$, was calculated for each lipid type individually.
-Sterols were excluded from the analysis.
-The script that calculates the equilibration of lipids is available at \texttt{Scripts/BuildDatabank/NMRPCA\_timerelax.py} in the NMRlipids databank repository.
-The resulting values are stored in files named \texttt{eq\_times.json} at folders in \texttt{/Data/Simulations} in the NMRlipids databank repository.
 
@@
 In contrast, locations of form factor minima are independent on simulation box size in Fig.~\ref{fig:sizedependence}.
-Here we use the location of the first form factor minima for the quality evaluation because automatic detection of the location of second minima is inaccurate for some experimental data due to fluctuations, such as for the POPE data in Figs.~\ref{fig:quality} D and E.
+Here we use only the location of the first form factor minimum for the quality evaluation because location of the second minimum is difficult to detect automatically in some experimental data sets due to fluctuations, such as for the POPE data in Figs.~\ref{fig:quality} D and E.
 The first minimum correlates well with the thickness of a membrane (Fig.~\ref{fig:quality} F), although the correlation of the second minima would be even stronger (Fig.~\ref{fig:QualityCorrelationsSI}).

### 2023-03-27 main.tex
change the equation for X-ray scattering form factors into non-symertic form

--- 2023-03-24 main.tex
@@
 \makeatother
+
+\newcommand{\iu}{{i\mkern1mu}}
 
@@
 \subsection{Calculation of X-ray scattering form factors}
-X-ray scattering form factors were calculated with the standard equation for symmetric lipid bilayers~\cite{ollila16}
+X-ray scattering form factors were calculated with the standard equation for lipid bilayers~\cite{ollila16} that allows also for unsymetry in membranes,
 \begin{equation}\label{FFsimpl}
-F(q)=\int_{-D/2}^{D/2}\Delta \rho_e(z) \cos(zq_z) {\rm d}z,
+F(q)= \left\lvert\int_{-D/2}^{D/2}\Delta \rho_e(z) \exp(\iu zq_z) {\rm d}z \right\rvert,
 \end{equation}

### 2023-03-27 main.tex
Manually merge https://github.com/NMRLipids/DataBankManuscriptText/pull/7

--- 2023-03-25 main.tex
@@
 \subsection{Calculation of X-ray scattering form factors}
-X-ray scattering form factors were calculated with the standard equation for symmetric lipid bilayers~\cite{ollila16}
+X-ray scattering form factors were calculated using the standard equation for lipid bilayers that does not assume symmetric membranes~\cite{ollila16},
 \begin{equation}\label{FFsimpl}
-F(q)=\int_{-D/2}^{D/2}\Delta \rho_e(z) \cos(zq_z) {\rm d}z,
+F(q)= \left\lvert\int_{-D/2}^{D/2}\Delta \rho_e(z) \exp(\iu zq_z) {\rm d}z \right\rvert,
 \end{equation}

### 2023-03-28 main.tex
Update text

--- 2023-03-25 main.tex
@@
 This procedure is illustrated in Fig.~\ref{fig:overlay}D and examples of codes that perform such analyses in practise are listed in tables~\ref{tab:codes} and~\ref{tab:codesApplication}.
+The equilibration period given by the user (TIMELEFTOUT in table~\ref{tab:READMEkeys}) is discarded from the trajectories in all analysis codes.
 
@@
 %, and frag can be \textit{sn}-1, \textit{sn}-2, headgroup or total (all order parameters within a molecule). 
-The overall quality of different fragments in a simulation are then defined as a molar fraction weighted average over different lipid components
+The overall quality of different fragments in a simulation (frag = tails, headgroup or total) are then defined as a molar fraction weighted average over different lipid components
 \begin{equation}
@@
 \end{equation}
-where $\chi_\mathrm{lipid}$ is the molar fraction of a lipid in the bilayer.
+where $\chi_\mathrm{lipid}$ is the molar fraction of a lipid in the bilayer and "tails" refer to the average of all acyl chains.
 

### 2023-04-13 main.tex
Update text
Update pdfs
Update main.tex
Requested information from Fernando Favela.

@@
 \author[4]{Pavel Buslaev}         %ORCID: 0000-0003-2031-4691
-\author[5]{Fernando Favela}
+\author[5]{Fernando Favela-Rosales}
 \author[6]{Tiago Mendes Ferreira}
@@
 \author[11,12]{Batuhan Kav}           %ORCID: 0000-0003-4990-373X
+\author[22]{Norbert Ku{\v c}erka}
 \author[13]{Patrik Kula}
@@
 \affil[21]{MD.USE Innovations S.L., Edificio Emprendia, 15782 Santiago de Compostela, Spain}
+\affil[22]{Department of Physical Chemistry of Drugs and Faculty of Pharmacy, Comenius University Bratislava, 832 32 Bratislava, Slovakia}
 
@@
 F. S.-L. thanks Axencia Galega de Innovaci{\'o}n for his predoctoral contract (02\_IN606D\_2022\_2667887), and Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (PID2019-111327GB-I00).
+F.
+F.-R.
+acknowledges Tecnol{\'o}gico Nacional de M{\'e}xico, Direcci{\'o}n General de Asuntos del Personal Acad{\'e}mico (DGAPA), Programa de Apoyo a Proyectos de Investigaci{\'o}n e Innovaci{\'o}n Tecnol{\'o}gica (PAPIIT) 101923, CONACyT Ciencia de Frontera 74884, for financial support and Miztli-Direcci{\'o}n de C{\'o}mputo y de Tecnolog{\'i}as de Informaci{\'o}n y Comunicaci{\'o}n (DGTIC) - Universidad Nacional Aut{\'o}noma de M{\'e}xico (UNAM) (Project LANCAD-UNAM-DGTIC-057) facilities for computing time allocation.
 M.S.M. was supported by the Trond Mohn Foundation BFS2017TMT01.
@@
 %A.A. conceived the experiment(s),  A.A. and B.A. conducted the experiment(s), C.A. and D.A. analysed the results.  All authors reviewed the manuscript. 
-
+F. F.-R. set up and performed simulations with Slipis Force Field: ID 617, ID 207 and ID 82. 
 P.B. implemented the PCA-based equilibration metric, supervised the work of A.K., and participated in code refactoring.

### 2023-04-13 main.tex
Update text

--- 2023-03-28 main.tex
@@
 \author[2]{Hanne S. Antila}          %ORCID: 0000-0002-2474-5053
-\author[3]{Lara Bort}             %ORCID: ?
+\author[3]{Lara S. Bort}             %ORCID: ?
 \author[4]{Pavel Buslaev}         %ORCID: 0000-0003-2031-4691
@@
 \begin{abstract}
-Cellular membrane lipid composition is implicated in diseases and major biological functions in cells, but membranes are difficult to study experimentally due to their intrinsic disorder and complex phase behaviour.
-Molecular dynamics (MD) simulations have been useful in understanding membrane systems, but they require significant computational resources and often suffer inaccuracies in model parameters.
-Applications of data-driven and machine learning methods, currently revolutionising many fields, are limited to membrane systems due to the lack of suitable training sets.
-Here we present the NMRlipids Databank---a community-driven, open-for-all database featuring programmatic access to quality evaluated atom-resolution MD simulations of lipid bilayers.
-The NMRlipids databank will benefit scientists in different disciplines by providing automatic ranking of simulations based on their quality against experiments, flexible implementations of data-driven and machine learning applications, and rapid access to simulation data via graphical user interface.
-To demonstrate the unlocked possibilities beyond current MD simulation studies, we analyzed how anisotropic diffusion of water and cholesterol flip–flop rates depend on membrane properties.
+Cellular membrane lipid composition is implicated in diseases and controls major biological functions, but membranes are difficult to study experimentally due to their intrinsic disorder and complex phase behaviour.
+Molecular dynamics (MD) simulations have been useful in understanding membrane systems, but they require significant computational resources and often suffer from inaccuracies in model parameters.
+Applications of data-driven and machine learning methods, currently revolutionising many fields, remain of limited use for membrane systems due to the lack of suitable training sets.
+Here we present the NMRlipids Databank---a community-driven, open-for-all database featuring programmatic access to quality-evaluated atom-resolution MD simulations of lipid bilayers.
+The NMRlipids Databank will benefit scientists in different disciplines by providing automatic ranking of simulations based on their quality against experiments, programmable interface for flexible implementation of data-driven and machine learning applications, and rapid access to simulation data through a graphical user interface.
+To demonstrate how it unlocks possibilities beyond current MD simulation studies, we analyzed the NMRlipids Databank to reveal how anisotropic diffusion of water and cholesterol flip-flop rates depend on membrane properties.
 
@@
 Membrane lipid composition is implicated in diseases, such as cancer and neurodegenerative disorders, and therapeutics that affect membrane compositions are emerging~\cite{torres21}.
-However, biomembranes are often difficult to study experimentally because they are complex mitures of lipids and proteins that are in disordered fluid state with complex phase behaviour at biological conditions.
+However, biomembranes are often difficult to study experimentally, because they are complex mixtures of proteins and lipids in disordered fluid state with complicated phase behaviour at biological conditions.
 For those reasons, the detailed connections between complex lipid interactions and biological functions taking place in or around membranes remain poorly understood.
@@
 Presently, the accuracy of models is becoming increasingly important as researches are progressing from simulations of individual molecules to simulating whole organelles or even cells using interdisciplinary approaches~\cite{johnson15,thornburg22,gupta22}.
-Such systems exhibit complex emergent behaviour making inaccuracies more difficult to detect and accumulation of even modest errors may have major impact on the conclusions.
+Such systems exhibit intricate emergent behaviour making inaccuracies more difficult to detect, and accumulation of even modest errors may have a dramatic impact on the conclusions drawn.
 
-In contrast to experimental structural biology, where standard protocols to share and quality-evaluate the resolved structures are firmly established \textit{via} the Protein Data Bank (PDB)~\cite{montelione13}, the equivalent best practices are yet to be defined for MD simulations.
-The importance of such approaches is widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} and data-sharing solutions are emerging for proteins in solution~\cite{meyer10,kamp10}, proteins in membranes~\cite{newport19,espigares20,leston22}, nucleic acids~\cite{hospital16}, nucleic acids and proteins~\cite{bekker20}, cyclodextrins~\cite{mixcoha16}, COVID-19--involved macromolecules~(\url{bioexcel-cv19.bsc.es}).
+In contrast to experimental structural biology, where standard protocols to share and quality-evaluate resolved biomolecular structures are firmly established by the Protein Data Bank (PDB)~\cite{montelione13}, equivalent best practices are yet to be defined for MD simulations.
+The importance of such approaches is widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} and data-sharing solutions are emerging for proteins in solution~\cite{meyer10,kamp10}, proteins in membranes~\cite{newport19,espigares20,leston22}, nucleic acids~\cite{hospital16}, nucleic acids and proteins~\cite{bekker20}, cyclodextrins~\cite{mixcoha16}, COVID-19--involved macromolecules~(\href{https://bioexcel-cv19.bsc.es}{bioexcel-cv19.bsc.es}).
 However, automatic quality evaluation of simulation data~\cite{meyer10,hospital16} and programmatic access are still rare.
@@
 Here we present the NMRlipids Databank---a community-driven, open-for-all database featuring programmatic access to atom-resolution MD simulations of lipid bilayers.
-The programmatic access enables users to apply data-driven approaches on the Databank, thus facilitating the creation of new tools for (and by) researchers in a wide range of fields covering academia and industry, from cell membrane biology and lipid nanoparticle formulations to computational chemistry and machine learning.
-As two usage examples of what is already possible, we demonstrate here (i) how data-driven analysis of water anisotropic diffusion in all the membrane systems available in the Databank can extend the scope of MD simulations to magnetic resonance imaging (MRI) and pharmacokinetics, and (ii) how the Databank allows its users to analyse rare phenomena that are beyond the scope of standard MD simulation investigations.
-Wider adaptation of the NMRlipids Databank will open even more possibities.
-Furthermore, the Databank performs automatic quality evaluation of membrane simulations, which facilitates the selection of best models for specific applications and accelerates the development of simulation parameters and methodology.
+The programmatic access enables users to apply their data-driven approaches on the Databank, thus facilitating the creation of new tools for (and by) researchers in a wide range of fields covering academia and industry, from cell membrane biology and lipid nanoparticle formulations to computational chemistry and machine learning.
+As two usage examples of what is already possible, we demonstrate here (i) how data-driven analysis of water anisotropic diffusion in all the membrane systems available in the Databank can extend the scope of MD simulations to magnetic resonance imaging (MRI) and pharmacokinetics and (ii) how the Databank allows its users to analyse rare phenomena that are beyond the scope of standard MD simulation investigations.
+Wider adaptation of the NMRlipids Databank will open even more possibilities.
+Furthermore, the Databank performs automatic quality evaluation of membrane simulations, which facilitates the selection of best-performing models for each given application and accelerates the development of simulation parameters and methodology.
 
-While the NMRlipids Databank currently contains only lipid bilayer systems, its key elements that enable the programmatic access to large scale MD simulation data can be applied also for other molecules, such as disordered proteins or sugars.
-The powerful combination of the overlay databank structure and a community-wide open collaboration can be used to build databases that enable data-driven and machine learning applications also in other fields where storage of raw data requires significant resources, best practices in the field are not defined, and incentives to share data do not exist, such as assignment of NMR spectra~\cite{klukowski22}.
+While the NMRlipids Databank currently contains only lipid bilayer systems, its key elements that enable programmatic access to large-scale MD simulation data can be applied also for other molecules, such as disordered proteins or sugars.
+The powerful combination of an overlay databank structure and a community-wide open collaboration can be used to build databases that enable data-driven and machine learning applications also in other fields, where storage of raw data requires significant resources, best practices in the field are not yet defined, and incentives to share data do not exist---such as assignment of NMR spectra~\cite{klukowski22}.
 
@@
 
-\subsection{MD simulations of membranes composed of most common biologically abundant lipids}
+\subsection{MD simulations of membranes composed of the biologically most  abundant lipids}
 
 NMRlipids Databank is a community-driven database containing atomistic MD simulations of biologically relevant lipid membranes emerging from the NMRlipids open collaboration~\cite{botan15,ollila16,catte16,antila19,bacle21}.
-It has been designed to improve data Findability, Accessibility, Interoperability, and Reuse (FAIR)~\cite{wilkinson16} for MD simulation trajectories.
-The NMRlipids databank is constructed using the NMRlipids project protocol where all the content is openly accessible throughout the project~\cite{botan15}.
-Currently the NMRlipids databank contains 726 simulation trajectories with the total length of approximately 0.4\,ms.
-The distribution of available simulations containing a specific lipid, shown in Fig.~\ref{fig:overlay}B, roughly resembles the biological abundance of different lipid types with phosphatidylcholine (PC) being the most common followed by cholesterol, phosphatidylethanolamine (PE), phosphatidylserine (PS), phosphatidylglycerol (PG), phosphatidylinositol (PI), and other lipids depending on organism and organelle~\cite{vanmeer08}.
-Abbreviations and full names of all lipids present in the databank are listed in Table~\ref{tab:abbreviations}.
-Single component lipid membranes and binary mixtures, depicted in Fig.~\ref{fig:overlay}F, are currently most abundant in the NMRlipids databank, yet mixtures with up to five lipid types are available.
-Force fields used in simulations cover all the essential parameters commonly used in lipid simulations, see Fig.~\ref{fig:overlay}C and table~\ref{tab:ForceFields}, including also united atom and polarizable force fields.
-Therefore, the averages calculated over the databank can be considered as mean predictions from available lipid models for an average cell membrane.
+It has been designed to improve the Findability, Accessibility, Interoperability, and Reuse (FAIR)~\cite{wilkinson16} of MD simulation data, most importantly the output trajectories.
+The NMRlipids Databank is constructed using the NMRlipids project protocol, in which all the content is openly accessible throughout the project~\cite{botan15}.
+Currently, the NMRlipids Databank contains 726 simulation trajectories with the total length of approximately 0.4\,ms.
+Single-component lipid membranes and binary mixtures, depicted in Fig.~\ref{fig:overlay}F, are currently most abundant in the NMRlipids Databank, yet mixtures with up to five lipid types are available.
+The distribution of lipids among the available simulations, shown in Fig.~\ref{fig:overlay}B, roughly resembles the biological relative abundance of different lipid types, with phosphatidylcholine (PC) being the most common followed by cholesterol, phosphatidylethanolamine (PE), phosphatidylserine (PS), phosphatidylglycerol (PG), phosphatidylinositol (PI), and other lipids from different organisms and organelles~\cite{vanmeer08}.
+Abbreviations and full names of all lipids present in the Databank are listed in Table~\ref{tab:abbreviations}.
+Force fields used in simulations cover all the essential parameter sets commonly used in lipid simulations, see Fig.~\ref{fig:overlay}C and Table~\ref{tab:ForceFields}, including also united atom and polarizable force fields.
+Therefore, the averages calculated over the Databank can be considered as mean predictions from available lipid models for an average cell membrane.
 
-The overlay structure of the NMRlipids databank, illustrated in Fig.~\ref{fig:overlay}A, is designed to enable efficient upcycling of MD simulations for data-driven and machine learning applications with mimimal investment on new infrastructure.
-Raw simulation data in the {\it data layer} can be stored in any publicly available location with long term stability, such as Zenodo (\url{www.zenodo.org}), and with permanent links to the data, such as digital object identifiers (DOIs).
+The overlay structure of the NMRlipids Databank, illustrated in Fig.~\ref{fig:overlay}A, is designed to enable efficient upcycling of MD simulations for data-driven and machine learning applications with minimal investment on new infrastructure.
+Raw simulation data in the {\it Data layer} can be stored in any publicly available location with long term stability---such as Zenodo (\href{https://zenodo.org}{zenodo.org}), figshare (\href{https://figshare.com}{figshare.com}), or NIRD research data archive (\href{https://archive.norstore.no}{archive.norstore.no})---and with permanent links to the data, such as digital object identifiers (DOIs).
 %or uniform resource names (URNs). 
-The {\it Databank layer} (\url{www.github.com/NMRLipids/Databank}) is the core of the databank containing all the relevant information about the simulations, including links to the raw data, relevant metadata describing the systems, universal naming conventions for lipids and their atoms, quality evaluation of simulations against experimental data, and computer programs to create the entries and analyse basic properties calculated from all simulations (area per lipid, C--H bond order parameters, X-ray scattering form factors, membrane thickness and equilibration times of principal components).
-Also the results for these basic properties are stored in the {\it Databank layer}.
+The {\it Databank layer} (\href{https://github.com/NMRlipids/Databank}{github.com/NMRlipids/Databank}) is the core of the Databank containing all the relevant information about the simulations: links to the raw data, relevant metadata describing the systems, universal naming conventions for lipids and their atoms, quality evaluation of simulations against experimental data, and the computer programs to create the entries and to analyse the five basic properties extracted from all simulations (area per lipid, C--H bond order parameters, X-ray scattering form factors, membrane thickness, and equilibration times of principal components).
+Also the values for these five basic properties are stored in the {\it Databank layer}.
 
-The {\it application layer} is composed of repositories and tools that read information from the {\it databank layer} for further analyses.
-Because {\it application layer} does not interfere with the {\it databank layer}, it can be freely extended by anyone for wide range of purposes.
-This is demonstrated here with two examples: NMRlipids databank graphical user interface (NMRlipids databank-GUI) tool at~\url{www.databank.nmrlipids.fi/} and a repository exemplifying novel analyses utilizing NMRlipids databank as discussed below (\url{www.github.com/NMRLipids/DataBankManuscript}).
-More detailed description of the NMRlipids databank structure is available in the supplementary information.
+The {\it Application layer} is composed of repositories and tools that read information from the {\it Databank layer} for further analyses.
+Because the {\it Application layer} does not interfere with the {\it Databank layer}, it can be freely extended by anyone for a wide range of purposes.
+This is demonstrated here with two examples: the NMRlipids Databank graphical user interface (NMRlipids Databank-GUI) at~\href{https://www.databank.nmrlipids.fi}{databank.nmrlipids.fi} and a repository exemplifying novel analyses utilizing NMRlipids Databank as discussed below (\href{https://github.com/NMRlipids/DataBankManuscript}{github.com/NMRlipids/DataBankManuscript}).
+A more detailed description of the NMRlipids Databank structure is available in the supplementary information.
 
@@
     \includegraphics[width=\linewidth]{Figures/overlay3.pdf}
-    \caption{A) Schematic presentation of the overlay structure used in the NMRlipids databank. A more detailed structure of the databank layer is shown in Fig.~\ref{DatabankStructure} in the SI.
+    \caption{A) Schematic presentation of the overlay structure used in the NMRlipids Databank.
+A more detailed structure of the \textit{Databank layer} is shown in Fig.~\ref{DatabankStructure} in the SI.
     %b) Distribution of the lengths of the trajectories, total number of trajectories and total length of the simulations in the NMRlipids databank.
-    B) Distribution of lipids present in the trajectories in the NMRlipids databank. Lipids occuring in five or fewer simulations ('others') are listed on the right. 
-    C) Distribution of force fields in the simulations in the NMRlipids databank. References for each force field are given in table~\ref{tab:ForceFields}.
-    D) Flowchart for performing an analysis of properties through all MD simulations in the NMRlipids databank using the API.
-    E) Flowchart for accessing results calculated from the NMRlipids databank and stored to the application layer.
-    F) Currently available binary mixtures in the NMRlipids databank. 
+    B) Distribution of lipids present in the trajectories of the Databank. `Others' lists lipids occurring in five or fewer simulations. 
+    C) Distribution of force fields in the simulations in the Databank. References for each force field are given in Table~\ref{tab:ForceFields}.
+    D) Flowchart for performing an analysis of properties through all MD simulations in the NMRlipids Databank using the API.
+    E) Flowchart for accessing results calculated from the NMRlipids Databank and stored to the \textit{Application layer}.
+    F) Currently available binary mixtures in the NMRlipids Databank. 
     }
@@
 
-\subsection{NMRlipids databank-GUI: graphical access to the MD simulation data}\label{section:access}
-NMRlipids databank-GUI, available at~\url{www.databank.nmrlipids.fi/}, provides easy access to the NMRlipids databank content via graphical user interface (GUI).
-Simulations can be searched based on their composition, used force field, temperature, properties and quality.
-The search results are ranked based on simulation quality evaluated against experimental data.
-NMRlipids databank-GUI enables also visualization of simulations and comparison of properties between different simulations and experiments.
-The NMRlipids databank-GUI enables rapid surveys of available simulation data, selection of best available simulations for specific systems based on ranking lists, and comparisons of basic properties between different types of membranes.
-Notably, GUI enables these operations to be performed by scientists with wide ranges of backgrounds who do not necessarily have expertise in programming or other means to access to MD simulation data.
+\subsection{NMRlipids Databank-GUI: graphical access to the MD simulation data}\label{section:access}
+NMRlipids Databank-GUI, available at~\href{https://www.databank.nmrlipids.fi}{databank.nmrlipids.fi}, provides easy access to the NMRlipids Databank content through a graphical user interface (GUI).
+Simulations can be searched based on their molecular composition, force field, temperature, membrane properties, and quality; the search results are ranked based on the simulation quality as evaluated against experimental data.
+Membranes can be visualized, and properties between different simulations and experiments compared.
+The NMRlipids Databank-GUI enables rapid surveying of what simulation data is available, selection of the best available simulations for specific systems based on ranking lists, and comparisons of basic properties between different types of membranes.
+Notably, the GUI enables these operations to be performed by scientists with a wide range of backgrounds---including those who do not necessarily have programming expertise or other means to access MD simulation data.
 
-\subsection{NMRlipids databank-API: programmatic access to the MD simulation data}
-The NMRlipids databank-API, available at \url{www.github.com/NMRLipids/Databank}, provides programmatic access to all simulation data in the NMRlipids databank.
-This enables wide range of novel data-driven applications from construction of machine learning models that predict membrane properties to automatic analyses of virtually any property from all simulations in the databank.
-Practical implementation of such analyses is illustrated in the flowchart in Fig.~\ref{fig:quality}D.
-First, raw data of each simulation can be accessed by iterating through the README.yaml files in the {\it databank layer} that contain links pointing to the locations of simulation trajectories and other relevant files.
-Simulations can be then automatically analyzed with the help of README.yaml and mapping files that affiliate specific naming conventions in each simulation to the universal molecule and atom names.
-Finally, the analysis results are stored using the same structure as in the {\it databank layer}.
-Examples of codes that utilize NMRlipids databank-API and a template for new analyses can be found from locations listed in table~\ref{tab:codes}.
+\subsection{NMRlipids Databank-API: programmatic access to the MD simulation data}
+The NMRlipids Databank-API, available at \href{https://github.com/NMRlipids/Databank}{github.com/NMRlipids/Databank}, provides programmatic access to all simulation data in the NMRlipids Databank.
+This enables wide range of novel data-driven applications---from construction of machine learning models that predict membrane properties, to automatic analysis of virtually any property across all simulations in the Databank.
+The flowchart in Fig.~\ref{fig:overlay}D illustrates the practical implementation of such an analysis.
+First, raw data of each simulation can be accessed by iterating through the README.yaml files in the {\it Databank layer} that contain links pointing to the locations of simulation trajectories and other relevant files.
+Simulations can then be automatically analyzed with the help of the README.yaml and the mapping files that associate the specific naming conventions in each simulation with the universal molecule and atom names used by the Databank.
+Finally, the analysis results are stored using the same structure as in the {\it Databank layer}.
+Examples of codes that utilize the NMRlipids Databank-API, as well as a template for a new analysis, can be found from locations listed in Table~\ref{tab:codes}.
 
@@
 %area per lipid, C–H bond order parameters, X-ray scattering form factors and thickness 
-are included in the {\it Databank layer}, unlimited amount of further analyses can be implemented by anyone in separate repositories in the {\it application layer}.
-When {\it Application layer} repositories are organized by mimicking the {\it Databank layer} structure, they can be accessed programmatically and further analyzed using the tools in NMRlipids databank-API by implementing the flowchart demonstrated in Fig.~\ref{fig:quality}E.
-Novel analyses demonstrating the power of the NMRlipids databank in selecting best simulation models, analyses of rare phenomena and extending MD simulations to new fields are implemented in an {\it application layer} repository located at~\url{www.github.com/NMRLipids/DataBankManuscript}.
-The related codes are listed in table~\ref{tab:codesApplication}.
+are included in the {\it Databank layer}, unlimited further analyses can be implemented by anyone in separate repositories in the {\it Application layer}.
+When {\it Application layer} repositories are organized by mimicking the {\it Databank layer} structure, they can be accessed programmatically and further analyzed using the tools in the NMRlipids Databank-API by implementing the flowchart demonstrated in Fig.~\ref{fig:overlay}E.
+Novel analyses that demonstrate the power of NMRlipids Databank in selecting the best simulation models, analysing rare phenomena, and extending MD simulations to new fields are implemented in an {\it Application-layer} repository located at~\href{https://github.com/NMRlipids/DataBankManuscript}{github.com/NMRlipids/DataBankManuscript}.
+The related codes are listed in Table~\ref{tab:codesApplication}.
 
@@
 
-\subsection{Applications of the NMRlipids databank}
-\subsubsection{Selection of simulation parameters using the NMRlipids databank: Best models for most abundant neutral membrane lipids}
+\subsection{Applications of the NMRlipids Databank}
+\subsubsection{Selecting simulation parameters using NMRlipids Databank: Best models for most abundant neutral membrane lipids}
 %Quality of membrane simulations with different force fields have been evaluated against experimental data during parameterization and in separate comparison studies~\cite{botan15,ollila16,catte16,pluhackova16,perez17,leonard19,??}, but the lack of universal quality measure for membrane simulations complicates the estimations of reliability of simulations.
@@
 %
-The quality of lipid bilayer MD simulations has to be carefully assessed in their applications to minize the detrimental consequences of artificial MD simulation results for lipid bilayers~\cite{antila22b}.
-This is possible, for example, by using C--H bond order parameters from NMR and X-ray scattering form factors \cite{ollila16}, although it requires comparisons between large number of simulations which is laborious even with collaborative approaches~\cite{botan15,catte16,antila19,bacle21}.
-To streamline this process we have defined quantitative quality measures for lipid bilayer simulations that enable automatic ranking of lipid bilayer simulations based on their quality against experiments.
-Conformational ensembles of individual lipid molecules are evaluated in the NMRlipids databank by calculating the probabilities for each C-H bond order parameter to locate within experimental error.
-Furthermore, the quality against X-ray scattering experiments is estimated from the distance of the first form factor minimum between the experiment and simulation.
-These measures can be used to evaluate membrane dimensions as the form factor minima locations and acyl chain order parameters correlate with the membrane lateral packing and thickness as shown in Figs.~\ref{fig:quality}G,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}.
-In addition, the ergodicity of conformational sampling of lipids is estimated by calculating the convergence time of the slowest principal component divided by the length of a simulation.
-Here we demonstrate how automatic simulation quality evaluation and NMRlipids databank-API enable rapid selection of the best models for simulations of membranes with POPC and POPE lipids that are the two most biologically abundant neutral membrane lipids~\cite{vanmeer08}.
+To minimize the detrimental consequences of artificial MD simulation results for their applications, the quality of lipid bilayer MD simulations has to be carefully assessed~\cite{antila22b}.
+This can be done, for example, against the C--H bond order parameters from NMR spectroscopy and the form factors from X-ray scattering~\cite{ollila16}, although it requires comparisons between large number of simulations, which is laborious even with collaborative approaches~\cite{botan15,catte16,antila19,bacle21}.
+To streamline this process, we have defined quantitative quality measures that enable automatic ranking of lipid bilayer simulations based on their quality against experiments.
+Conformational ensembles of individual lipid molecules are evaluated in the NMRlipids Databank by calculating the probabilities for each C--H bond order parameter to locate within experimental error ($P^{\mathrm{hg}}$, $P^{\mathrm{sn1}}$,$P^{\mathrm{sn2}}$, and $P^{\mathrm{total}}$).
+Furthermore, the quality against X-ray scattering experiments ($FF_q$) is estimated as the difference in the experimental and simulated locations of the first form factor minimum.
+These measures can be used to evaluate membrane dimensions, as the form factor minima locations and acyl chain C--H bond order parameters correlate with the membrane lateral packing and thickness (Figs.~\ref{fig:quality}G,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}).
+In addition, the ergodicity of conformational sampling of lipids is estimated by calculating $\tau_\mathrm{rel}$, the convergence time of the slowest principal component divided by the simulation length.
+Let us demonstrate how the automatic simulation-quality evaluation and the NMRlipids Databank-API enable rapid selection of the best models for simulations of membranes with the two most biologically-abundant neutral membrane lipids~\cite{vanmeer08}: POPC and POPE.
 
-Predictions for the lateral packing of POPC and POPE membranes in terms of area per lipid, one of the most important parameter used to characterize cellular membranes, diverge between different force fields as illustrated in Fig.~\ref{fig:quality}A.
-To select the most realistic models among these, we first ranked all simulations in the databank according to the quality of C-H bond order parameters evaluated against experimental data in Fig.~\ref{fig:top50simulations}.
-From these results, we then picked the force fields occurring in Fig.~\ref{fig:quality}A and ranked them according to the quality of $\textit{sn}$-1 chain in Figs.~\ref{fig:quality}B and C.
-Simulations with $\tau_{\rm rel}$ clearly above one were discarded.
-Also the highest ranked simulation (POPC bilayer with OPLS3e parameters) is included for the reference.
-Because the membrane packing correlates with the average order parameter of the $\textit{sn}$-1 chain (Fig.~\ref{fig:quality}G), rankings in Figs.~\ref{fig:quality}B and C were used to select the simulations giving most realistic results in~Fig.~\ref{fig:quality}A.
-Also direct comparisons with the experimental data for the most relevant simulations are shown in Figs.~\ref{fig:quality}C--E and~\ref{fig:POPC_POPE_dataSI}.
-Based on these rankings, Lipid17 and Slipids simulations give the most realistic predictions for POPC membrane, while simulations with CHARMM36 and GROMOS-CKP parameters predict overly packed bilayers (overestimated order in Fig.~\ref{fig:POPC_POPE_dataSI}).
-On the other hand, GROMOS-CKP and Slipids give the most realistic results for POPE, while CHARMM36 and Lipid17 predict membranes that are too packed.
-In conclusion, the quality evaluation based on the NMRlipids databank suggests that Slipids parameters are the best available choice for simulations with PC and PE lipids, at least for applications where membrane packing is relevant.
+As Fig.~\ref{fig:quality}A illustrates, predictions for the lateral packing of POPC and POPE membranes in terms of area per lipid, one of the most important parameters used to characterize cellular membranes, diverge between different force fields.
+To select the most realistic currently available force fields to describe POPC--POPE membranes, we first rank all simulations in the Databank according to the quality of their C--H bond order parameters  (Fig.~\ref{fig:top50simulations}).
+From these results, we then pick the force fields occurring in Fig.~\ref{fig:quality}A (that is: force fields for which both POPC and POPE simulations exist in the Databank) and rank them according to the quality of the \textit{sn}-1 chain of POPC (Fig.~\ref{fig:quality}B) and of POPE (Fig.~\ref{fig:quality}C).
+Simulations with $\tau_\mathrm{rel}$ clearly above one are discarded.
+Because the membrane packing correlates with the average order parameter of the \textit{sn}-1 chain (Fig.~\ref{fig:quality}G), rankings in Figs.~\ref{fig:quality}B and C can be used to select the simulations giving the most realistic results in Fig.~\ref{fig:quality}A.
+(Figs.~\ref{fig:quality}E--F and~\ref{fig:POPC_POPE_dataSI} also show direct comparisons with the experimental data for the most relevant simulations; for reference, Fig.~\ref{fig:quality}D shows the overall highest-ranked simulation: POPC bilayer with OPLS3e parameters.) Based on these rankings, Lipid17 and Slipids simulations give the most realistic predictions for a POPC membrane, while simulations with CHARMM36 and GROMOS-CKP parameters predict overly packed bilayers (overestimated order in Fig.~\ref{fig:POPC_POPE_dataSI}).
+For POPE, on the other hand, GROMOS-CKP and Slipids give the most realistic results, while CHARMM36 and Lipid17 predict membranes that are too packed.
+In conclusion, the quality evaluation based on the NMRlipids Databank suggests that the Slipids parameters are the best currently available choice for simulations with PC and PE lipids, at least for applications where membrane packing is relevant.
 
@@
     \includegraphics[width=\linewidth]{Figures/quality5.pdf}
-    \caption{ A) Area per lipids of POPC and POPE lipid bilayers predicted by different force fields at 310\,K in simulations that are available in the NMRlipids databank.
-The points from the best performing simulations based on rankings in B and C are surrounded by black circles.
-    B) Best POPC simulations ranked based on {\it sn}-1 acyl chain order parameter quality.
-    C) Best POPE simulations ranked based on {\it sn}-1 acyl chain order parameter quality. 
+    \caption{ A) Area per lipid of POPC and POPE lipid bilayers predicted by different force fields at 310\,K in simulations that are available in the NMRlipids Databank.
+The data points from the best-performing simulations, based on rankings in panels B and C, are surrounded by black circles.
+    B) Best POPC simulations ranked based on the \textit{sn}-1 acyl chain order parameter quality.
+    C) Best POPE simulations ranked based on the \textit{sn}-1 acyl chain order parameter quality. 
     D)--F) Direct comparison against experimental (NMR order parameters and X-ray scattering ) data exemplified for a simulation with the best overall order parameter quality (D), the best quality for POPE lipid (E), and the headgroup quality for POPE (F).
-    G) Scatter plots and Pearson correlation coefficients, $r$, for the membrane area per lipid, thickness, first minimum of X-ray scattering form factor and average order parameter of the {\textit sn}-1 acyl chain extracted from the NMRlipids databank.
+    G) Scatter plots and Pearson correlation coefficients, $r$, for the membrane area per lipid, thickness, first minimum of X-ray scattering form factor and average order parameter of the \textit{sn}-1 acyl chain extracted from the NMRlipids Databank.
 All correlation coefficients have p-value below 0.001 as indicated by~$^{***}$.
@@
 
-\subsubsection{Detecting rare phenomena using NMRlipids databank: Cholesterol flip--flops}
-Lipid flip--flops from one bilayer leaflet to another play an important role in lipid trafficking and regulating membrane properties~\cite{vanmeer08}.
-Phospholipid flip--flops are slow when not facilitated by proteins, occurring spontaneously on the timescale of hours or days, while cholesterol, diacylglycerol and ceramide flip-flops are faster.
+\subsubsection{Detecting rare phenomena using NMRlipids Databank: Cholesterol flip-flops}
+Lipid flip-flops from one bilayer leaflet to another play an important role in lipid trafficking and regulating membrane properties~\cite{vanmeer08}.
+Phospholipid flip-flop events are rare when not facilitated by proteins, occurring spontaneously on the timescale of hours or days, while cholesterol, diacylglycerol, and ceramide flip-flop much more often.
 Still, the reported timescales range from minutes to sub-millisecods~\cite{vanmeer08,steck12,parisio16,gu19}.
-These timescales were previously accessible only by coarse-grained simulations or free energy calculations~\cite{parisio16}, yet atomistic simulations reporting cholesterol flip--flop events have been published only recently~\cite{gu19,javanainen19,baral20}.
-These studies report an increase in cholesterol flip--flop rates with increasing acyl chain unsaturation level and decreasing cholesterol concentration~\cite{gu19,javanainen19}, but the amount of data in these individual studies was not sufficient to systematically assess correlations between cholesterol flip--flop rates and membrane properties.
-Here, we demonstrate that the NMRlipids databank API makes analyses of such rare phenomena accessible to everyone by enabling the access to large amount of MD simulation data as described in section~\ref{section:access}.
-This is particularly useful for scientists in different fields of science and industry who do not have the access to required computational resources and expertise to produce the large amounts of MD simulation data.
+These timescales were previously accessible only by coarse-grained simulations or free energy calculations~\cite{parisio16}, and atomistic simulations reporting cholesterol flip-flop events have been published only recently~\cite{gu19,javanainen19,baral20}.
+The atomistic studies report an increase in cholesterol flip-flop rates with increasing acyl chain unsaturation level and decreasing cholesterol concentration~\cite{gu19,javanainen19}, but the amount of data in these individual studies was not sufficient to systematically assess correlations between cholesterol flip-flop rates and membrane properties.
+Here, we demonstrate that the NMRlipids Databank-API makes analyses of such rare phenomena accessible for all by enabling access to a large amount of MD simulation data as described in section~\ref{section:access}.
+This is particularly useful for scientists in various fields of science and industry who lack access to the computational resources or the expertise to produce the large amounts of MD simulation data required for such analyses.
 
-Using workflow depicted in Fig.~\ref{fig:overlay}D, we first calculated the cholesterol flip--flop rates from all the simulations available in the NMRlipids databank.
-Flip--flops were observed for cholesterol, % (634 events in 77 different simulations),
+Using the general workflow depicted in Fig.~\ref{fig:overlay}D, we first calculated the flip-flop rates from all the simulations available in the NMRlipids Databank.
+Flip-flops were observed for cholesterol, % (634 events in 77 different simulations),
 DCHOL (18,19-di-nor-cholesterol), %, 16 flip-flops in 1 simulation) 
@@
 and SDG (1-stearoyl-2-docosahexaenoyl-\textit{sn}-glycerol).
-The observed cholesterol flip--flop rates, ranging between 0.001--1.6\,\textmu{}s$^{-1}$ with the mean of 0.16\,\textmu{}s$^{-1}$ and median of 0.07\,\textmu{}s$^{-1}$, are in line with the previously reported values from atomistic MD simulations~\cite{gu19,javanainen19,baral20}.
-The flip--flop rate of DCHOL, 0.2\,\textmu{}s$^{-1}$, was close to the average value of cholesterol, while average rates for diacylglycerols DOG and SDG were higher than for cholesterol, 0.4\,\textmu{}s$^{-1}$ and 0.5\,\textmu{}s$^{-1}$, respectively.
-Flip--flops were not observed for other lipids, giving the upper limits for PC lipid flip--flop rate of 9$\cdot$10$^{-6}$\,\textmu{}s$^{-1}$ and for ceramide (\textit{N}-palmitoyl-\textsc{D}-\textit{erythro}-sphingosine) of 0.002\,\textmu{}s$^{-1}$.
-Thus, available data in the NMRlipids databank suggests that the lipid flip--flip rate decreases in the order of diacylglycerols $>$ cholesterol $>$ other lipids including ceramides.
-However, the amount of data for diacylgycerols (8 simulations with Lipid17 force field) and ceramide (3 simulations with CHARMM36 force field) is less than for cholesterol (83 simulations), thus we cannot fully exclude the effect of force field or composition on this comparison.
+The observed cholesterol flip-flop rates, ranging between 0.001--1.6\,\textmu{}s$^{-1}$ with the mean of 0.16\,\textmu{}s$^{-1}$ and median of 0.07\,\textmu{}s$^{-1}$, are in line with the previously reported values from atomistic MD simulations~\cite{gu19,javanainen19,baral20}.
+The flip-flop rate of DCHOL, 0.2\,\textmu{}s$^{-1}$, was close to the average value of cholesterol, while
+the average rates for diacylglycerols DOG (0.4\,\textmu{}s$^{-1}$) and SDG (0.5\,\textmu{}s$^{-1}$) were higher than for cholesterol.
+Flip-flops were not observed for other lipids, giving the upper limits for
+PC-lipid flip-flop rate as 9$\times$10$^{-6}$\,\textmu{}s$^{-1}$ and for
+ceramide (\textit{N}-palmitoyl-\textsc{D}-\textit{erythro}-sphingosine) as 0.002\,\textmu{}s$^{-1}$.
+Thus, the available data in the NMRlipids Databank suggest that the lipid flip-flop rate decreases in the order: diacylglycerols $>$ cholesterol $>$ other lipids including ceramides.
+However, the amount of data for diacylgycerols (8 simulations with the Lipid17 force field) and ceramide (3 simulations with CHARMM36) is less than that for cholesterol (83 simulations); thus we cannot fully exclude the effect of force field or composition on this comparison.
 
-Nevertheless, we used the workflow depicted in Fig.~\ref{fig:overlay}E to analyse how flip--flop rates calculated from the NMRlipids databank depend on membrane properties.
-Cholesterol flip--flop rates and their histograms are plotted as a function of membrane thickness, lateral density and order in Figs.~\ref{fig:flip-flops} B--D.
-The results reveal a non-linear correlation between cholesterol flip--flop rate and membrane packing which is depicted as area per lipid.
-Flip--flop rates increase by an order of magnitude when membrane packing density decreases and a major jump is observed at low membrane packing.
-The order of magnitude changes in cholesterol flip--flop rate with the membrane composition may have major implications in understanding lipid trafficking and membrane biochemistry~\cite{gu19,baral20}.
-Because the results from the NMRlipids databank are averaged over large range of membrane compositions and force fields, they show that the strong dependence of cholesterol flip--flop rate on membrane properties is not limited to certain lipid compositions or force fields used in previous studies~\cite{gu19,javanainen19,baral20}.
+Nevertheless, we used the general workflow depicted in Fig.~\ref{fig:overlay}E to analyse how the flip-flop rates calculated from the NMRlipids Databank depend on membrane properties.
+Figures~\ref{fig:flip-flops}B--D show cholesterol flip-flop rates and their histograms as a function of membrane thickness, lateral density, and acyl chain order.
+The results reveal a non-linear correlation between cholesterol flip-flop rate and membrane packing (depicted as area per lipid): Flip-flop rates increase by an order of magnitude when membrane packing density decreases, and a major jump is observed at low membrane packing.
+Such order-of-magnitude changes in cholesterol flip-flop rate with the membrane composition may have major implications in understanding lipid trafficking and membrane biochemistry~\cite{gu19,baral20}.
+Because the results from the NMRlipids Databank are averaged over a large range of membrane compositions and force fields, they show that the strong dependence of cholesterol flip-flop rate on membrane properties is not limited to the particular lipid compositions or force fields used in the previous studies~\cite{gu19,javanainen19,baral20}.
 
@@
     \includegraphics[width=88mm]{Figures/CholFlipFlops.pdf}
-    \caption{\textbf{A} Illustration of cholesterol flip--flop.  
-      \textbf{B--D} Cholesterol flip--flops analyzed from the databank as a function of membrane thickness, area per lipid, and acyl chain order.
-Values from simulations with non-zero flip--flop rates are shown with blue dots.
+    \caption{\textbf{A} Illustration of cholesterol flip-flop.  
+      \textbf{B--D} Cholesterol flip-flops analyzed from the Databank as a function of membrane thickness, area per lipid, and acyl chain order.
+Values from simulations with non-zero flip-flop rates are shown with blue dots.
 Histogrammed values are shown with black dots.
@@
 
-\subsubsection{Extending the scope of MD simulations to new fields using the NMRlipids databank: Water diffusion anisotropy in membrane systems}
-The anisotropy in water diffusion in parallel and perpendicular directions with respect to membranes can be related to the translocation of drugs  through biological material, particularly in skin \cite{hansen13,wen18,nitsche19,roberts21}, and to the formation of signals in diffusion tensor MRI imaging~\cite{topgaard20}.
-MD simulations are rarely used to analyze the anisotropic diffusion of water since only few membrane permeation events for water are typically observed in a single MD simulation trajectory~\cite{venable19,camilo2022}, thereby making the collection of sufficient amount of data challenging.
-Here, we show that the API access to the data in NMRlipids databank enables systematic analyses on how the anisotropic diffusion of water depends on membrane properties in multilamellar membrane systems, thereby extending the applications of MD simulations to new fields.
+\subsubsection{Extending the scope of MD simulations to new fields using NMRlipids Databank: Water diffusion anisotropy in membrane systems}
+The anisotropy of water diffusion in directions parallel and perpendicular to membranes can be related to translocation of drugs through biological material, particularly in the skin \cite{hansen13,wen18,nitsche19,roberts21}, and to signal formation in diffusion tensor MRI imaging~\cite{topgaard20}.
+MD simulations are rarely used to analyze the anisotropic diffusion of water, since only a few membrane permeation events of water are typically observed in a single MD simulation trajectory~\cite{venable19,camilo2022}, thereby making the collection of a sufficient amount of data challenging.
+Here, we show that the API access to the data in NMRlipids Databank enables systematic analysis on how the anisotropic diffusion of water depends on membrane properties in multilamellar membrane systems, thereby extending the application of MD simulations to new fields.
 
@@
     \includegraphics[width=\linewidth]{Figures/permeation2.pdf}
-    \caption{\textbf{A} Water diffusion, $D_\mathrm{perp}$, and permeability, $P$, through membranes, and lateral diffusion along the membrane, $D_{||}$, illustrated in a multilamellar stack of lipid bilayers.
+    \caption{\textbf{A} Water diffusion, $D_\perp$, and permeability, $P$, through membranes, and lateral diffusion along the membrane, $D_\parallel$, illustrated in a multilamellar stack of lipid bilayers.
     \textbf{B--E} Water permeation through membranes analyzed from the databank as a function of temperature, thickness, area per lipid, and acyl chain order.
@@
 
-To this end, we first calculated the water permeability through membranes from all simulations in the NMRlipids databank using the workflow depicted in Fig.~\ref{fig:overlay}D.
-The resulting non-zero values range between 0.3 and 322\,\textmu{}m/s with the mean and median of 14\,\textmu{}m/s and 8\,\textmu{}m/s, respectively.
-These values agree with the previously reported simulation results~\cite{venable19,camilo2022}, but are on average slightly larger than experimental values reported for PC lipids in the liquid crystalline phase,~0.19--0.33\,\textmu{}m/s~\cite{jansen95}.
+To this end, we first calculated the water permeability through membranes from all simulations in the NMRlipids Databank using the general workflow depicted in Fig.~\ref{fig:overlay}D.
+The resulting non-zero values range between 0.3--322\,\textmu{}m/s with the mean of 14\,\textmu{}m/s and median of 8\,\textmu{}m/s.
+These values agree with the previously reported simulation results~\cite{venable19,camilo2022}, but are on average slightly larger than experimental values reported for PC-lipids in the liquid crystalline phase,~0.19--0.33\,\textmu{}m/s~\cite{jansen95}.
 Using the workflow depicted in Fig.~\ref{fig:overlay}E, we then plotted the observed permeabilities and their histogrammed values in Figs.~\ref{fig:permeability}B--E as a function of temperature, membrane thickness, area per lipid, and acyl chain order.
@@
 To analyze how water diffusion anisotropy depends on membrane properties in a multi-lamellar lipid bilayer system, we calculated the water diffusion parallel to the membrane surface from all simulations in the NMRlipids databank using the workflow depicted in Fig.~\ref{fig:overlay}D.
-The parallel diffusion coefficient of water, $D_{||}$, decreases with reduced hydration and increases with the temperature, but dependencies on membrane area per lipid, thickness, or fraction of charged lipids were not observed in Figs.~\ref{fig:permeability} and~\ref{fig:diffusionSI} when the workflow depicted in Fig.~\ref{fig:overlay}E was used.
-Simulation results are close to the experimental values with low hydration levels in Fig.~\ref{fig:permeability} F, but increase to approximately two times higher than the experimental value for bulk water diffusion value (3.1$\cdot$\,10$^{-9}$\,m$^2$/s at 313\,K~\cite{khakimov08}) with high hydration levels.
+The parallel diffusion coefficient of water, $D_\parallel$, decreases with reduced hydration and increases with the temperature, but dependencies on membrane area per lipid, thickness, or fraction of charged lipids were not observed in Figs.~\ref{fig:permeability} and~\ref{fig:diffusionSI} when the workflow depicted in Fig.~\ref{fig:overlay}E was used.
+Simulation results are close to the experimental values with low hydration levels in Fig.~\ref{fig:permeability}F, but increase to approximately two times higher than the experimental value for bulk water diffusion value (3.1$\cdot$\,10$^{-9}$\,m$^2$/s at 313\,K~\cite{khakimov08}) with high hydration levels.
 This is not surprising as the most common water model used in membrane simulations, TIP3P, overestimates the bulk water diffusion~\cite{pathirannahalage21}.
-To estimate the diffusion anisotropy of water, $D_\mathrm{\perp}/D_{||}$, in multilamellar membrane system, the permeability coefficients of water through membranes were translated to perpendicular diffusion coefficients, $D_\mathrm{\perp}$, using the Tanner equation~\cite{tanner78,wasterby02}.
+To estimate the diffusion anisotropy of water, $D_\perp/D_\parallel$, in multilamellar membrane system, the permeability coefficients of water through membranes were translated to perpendicular diffusion coefficients, $D_\perp$, using the Tanner equation~\cite{tanner78,wasterby02}.
 The resulting perpendicular diffusion coefficients are approximately five orders of magnitude smaller than lateral diffusion coefficients of water (Figs.~\ref{fig:permeability} G and H), which is at the upper limit of the anisotropy estimated from the experimental data~\cite{nitsche19}.
-Significant increase in the diffusion anisotropy with membrane packing is observed, as $D_\mathrm{\perp}/D_{||}$ drifts away from one with decreasing area per lipid and increasing thickness in Figs.~\ref{fig:permeability} G and H.
+Significant increase in the diffusion anisotropy with membrane packing is observed, as $D_\perp/D_\parallel$ drifts away from one with decreasing area per lipid and increasing thickness in Figs.~\ref{fig:permeability} G and H.
 This follows from decreasing water permeability with membrane packing (Figs.~\ref{fig:permeability} C and D), while lateral diffusion remains approximately constant (Figs.~\ref{fig:diffusionSI} A and C).
@@
 The focus of biomolecular simulations is moving from studies of individual molecules to larger complexes and even whole cells and organelles~\cite{johnson15,thornburg22,gupta22}.
-Simultaneously, machine learning based models predicting behaviour of biomolecules and automatic approaches to parametrize models are emerging~\cite{jumper21,antila22b}.
+Simultaneously, machine learning -based models predicting behaviour of biomolecules and automatic approaches to parametrize models are emerging~\cite{jumper21,antila22b}.
 The resources delivered by the NMRlipids databank will support the development in both of these directions.
@@
 
-The open programmatic access to unprecedented amount of MD simulation data via NMRlipids databank-API enables wide range of novel data-driven analyses and machine learning applications.
+The open programmatic access to unprecedented amount of MD simulation data \textit{via} NMRlipids databank-API enables wide range of novel data-driven analyses and machine learning applications.
 The analyses of cholesterol flip--flop events (Fig.~\ref{fig:flip-flops}) and water permeation through membranes (Fig.~\ref{fig:permeability}) demonstrate how large amount of accessible simulation data in terms of quantity (\textit{e.g.}, simulation length and number of conformations) and content (\textit{e.g.}, lipid compositions and ion concentrations) enable analyses of rare phenomena that are beyond the current possibilities for a single research group.
@@
     Anisotropic water diffusion for pharmacokinetics and MRI imaging applications (Fig.~\ref{fig:permeability}) & 
-    Scientist in fields where MD simulations are not usually applied \\
+    Scientists in fields where MD simulations are not usually applied \\
     \\
@@
     Training set for machine learning applicatons &
-    Programmatic access to the data and results enables training of machine learning type of models for various applications, such as predictions of membrane properties from composition & Scientists building and using machine learning applications for biomolecules.
+    Programmatic access to the data and results enables training of machine learning models for various applications, such as predictions of membrane properties from composition & Scientists building and using machine learning applications for biomolecules.
     \end{tabular}
@@
 On the other hand, the open collaboration approach developed in the NMRlipids project~\cite{botan15} creates incentives for sharing the data by offering authorship in published articles to the contributors.
-Advantages of such approach are demonstrated here for membrane simulations, but the concept can be applied also not only to other biomolecules, such as disordered proteins and membrane-protein systems, but also in other fields where similar barriers hinder the machine learning revolution, such as assignment of NMR spectra~\cite{klukowski22}.
+Advantages of such approach are demonstrated here for membrane simulations, but the concept can be applied also not only to other biomolecules, such as disordered proteins and membrane--protein systems, but also in other fields where similar barriers hinder the machine learning revolution, such as assignment of NMR spectra~\cite{klukowski22}.
 
@@
 The databank structure is illustrated more detailed in Fig.~\ref{DatabankStructure} in the supplementary information.
-Whenever specific files or folders are referred here, they locate at {\it datbank layer} repository unless stated otherwise.
+Whenever specific files or folders are referred here, they locate at {\it databank layer} repository unless stated otherwise.
 
@@
 To enable automatic analyses over simulations with different atom and molecule names in the NMRlipids databank, we have defined unique naming conventions for molecules and atoms therein.
-Unique abbreviations used in the NMRlipids databank for each molecule are listed in table~\ref{tab:abbreviations} in the supplementary information.
+Unique abbreviations used in the NMRlipids databank for each molecule are listed in Table~\ref{tab:abbreviations} in the supplementary information.
 Atom names used in simulation trajectories are connected to unique atom names using mapping files that are defined in the NMRlipids project (\url{https://nmrlipids.blogspot.com/2022/04/new-yaml-format-of-mapping-files.html}).
@@
 In practise, simulations in the NMRlipids databank can be analyzed by executing a program that loops over README.yaml files in the {\it databank layer}, downloads the data using the information in README.yaml files, and then performs the desired analysis on a local computer utilising the universal naming conventions for molecules and atoms defined in README.yaml and mapping files.
-This procedure is illustrated in Fig.~\ref{fig:overlay}D and examples of codes that perform such analyses in practise are listed in tables~\ref{tab:codes} and~\ref{tab:codesApplication}.
-The equilibration period given by the user (TIMELEFTOUT in table~\ref{tab:READMEkeys}) is discarded from the trajectories in all analysis codes.
+This procedure is illustrated in Fig.~\ref{fig:overlay}D and examples of codes that perform such analyses in practise are listed in Tables~\ref{tab:codes} and~\ref{tab:codesApplication}.
+The equilibration period given by the user (TIMELEFTOUT in Table~\ref{tab:READMEkeys}) is discarded from the trajectories in all analysis codes.
 
@@
 Because the motions along the first, major, principal component are the slowest ones \cite{buslaev16}, the equilibration of each lipid type was estimated from the ratio between the distribution convergence of the trajectories projected on the first PC and the trajectory length, $\tau_{\rm rel} = \tau_{\text{convergence}}/\tau_{\text{sim}}$~\cite{buslaev16, buslaev2020principal}.
-If $\tau_{\rm rel} < 1$, simulations can be considered to be sufficiently long for lipid molecules to sample their conformationla ensembles, while in simulations with $\tau_{\rm rel} > 1$ individual molecules may not have fully sample their conformational ensembles.
+If $\tau_{\rm rel} < 1$, simulations can be considered to be sufficiently long for lipid molecules to sample their conformational ensembles, while in simulations with $\tau_{\rm rel} > 1$ individual molecules may not have fully sampled their conformational ensembles.
 Rigid molecules that do not exhibit significant conformational fluctuations, such as sterols, were excluded from the analysis.
@@
 \subsection{Calculation of a bilayer area per lipid and thickness}
-Area per lipids of bilayers were calculated by dividing the time-averaged area of the simulation box with the total number of lipid and surfactant molecules in the simulation (see table~\ref{tab:abbreviations} for the list of molecules considered as lipids or surfactants).
+Area per lipids of bilayers were calculated by dividing the time-averaged area of the simulation box with the total number of lipid and surfactant molecules in the simulation (see Table~\ref{tab:abbreviations} for the list of molecules considered as lipids or surfactants).
 The script that calculates area per lipids from all simulations in the NMRlipids databank repository is available at \texttt{Scripts/AnalyzeDatabank/calcAPL.py} in the NMRlipids databank repository.
@@
 In practise, we first filter the fluctuations from the form factor data using Savitzky--Golay filter (window length 30 and polynomial order 1) and locate the first minima above 0.1\,\AA$^{-1}$ from both simulation and experimental data.
-The quality of a form factor is then defined as Euclidean distance between the minima in simulated and experimental form factors, $FF_q = |FF_\mathrm{min}^\mathrm{sim}-FF_\mathrm{min}^\mathrm{exp}|$.
+The quality of a form factor is then defined as Euclidean distance between the minima in simulated and experimental form factors, $FF_q = |FF_\mathrm{min}^\mathrm{sim}-FF_\mathrm{min}^\mathrm{exp}| \times 100$.
 
@@
 
-Water diffusion in the perpendicular direction of lipid bilayers in a multilamellar stack was estimated from the Tanner equation $D_\mathrm{perp} = \frac{D_{||} P z_w}{D_{||} + P z_w}$~\cite{tanner78,wasterby02}, where the water layer thickness, $z_w$, was estimated by subtracting bilayer thickness from the size of the simulation box in membrane normal direction.
+Water diffusion in the perpendicular direction of lipid bilayers in a multilamellar stack was estimated from the Tanner equation $D_\mathrm{\perp} = \frac{D_{||} P z_w}{D_{||} + P z_w}$~\cite{tanner78,wasterby02}, where the water layer thickness, $z_w$, was estimated by subtracting bilayer thickness from the size of the simulation box in membrane normal direction.
 
@@
 O.H.S.O, A.M.K, and R.N. acknowledge CSC --- IT Center for Science for computational resources and Academy of Finland (grants 315596 and 319902) for financial support.
+
 \section*{Author contributions statement}
@@
 
+H. S. A. contributed to the quality evaluation metrics based on NMR and SAXS, and was involved in testing the GUI and polishing the manuscript.
+Lara S. Bort implemented about 630 pre-existing lipid double layer simulations into the databank layer repository and created corresponding mapping files.
 P.B. implemented the PCA-based equilibration metric, supervised the work of A.K., and participated in code refactoring.
+B.K. contributed to the initial design and implementation of the databank's Python back-end and was involved with the writing the manuscript.
 R.G.-F was responsible for the design, validation, supervision and funding of the back-end and front-end for the graphical user interface (GUI) (http://www.databank.nmrlipids.fi/).
@@
 P.F.J.F. contributed to changes in the buildH software so that the Databank can handle united atom simulation data.
+A.K. implemented the PCA-based equilibration metric. 
+I.G. advised on implementation of the PCA-based equilibration metric and supervised the work of A.K.
+N.K. contributed experimental scattering data and SIMtoEXP algorithm.
+M.K. provided simulations with OPLS3e force field
 

### 2023-04-14 main.tex
Updates from Overleaf
Update main.tex

@@
 O.H.S.O, A.M.K, and R.N. acknowledge CSC --- IT Center for Science for computational resources and Academy of Finland (grants 315596 and 319902) for financial support.
+M.K acknowledge CSC – IT Center for Science, Finland, for computational resources and thanks Finnish Cultural Foundation and the UEF Doctoral Programme for the financial support
 
@@
 M.K. provided simulations with OPLS3e force field
+J.J.M. provided simulations with Amber force field, general discussion, and manuscript editing
 

### 2023-04-14 main.tex

@@
 To minimize the detrimental consequences of artificial MD simulation results for their applications, the quality of lipid bilayer MD simulations has to be carefully assessed~\cite{antila22b}.
-This can be done, for example, against the C--H bond order parameters from NMR spectroscopy and the form factors from X-ray scattering~\cite{ollila16}, although it requires comparisons between large number of simulations, which is laborious even with collaborative approaches~\cite{botan15,catte16,antila19,bacle21}.
+This can be done, for example, against the C--H bond order parameters from NMR spectroscopy~\cite{bacle21,Wur23} and the form factors from X-ray scattering~\cite{ollila16}, although it requires comparisons between large number of simulations, which is laborious even with collaborative approaches~\cite{botan15,catte16,antila19,bacle21}.
 To streamline this process, we have defined quantitative quality measures that enable automatic ranking of lipid bilayer simulations based on their quality against experiments.
@@
 N.K. contributed experimental scattering data and SIMtoEXP algorithm.
-M.K. provided simulations with OPLS3e force field
-J.J.M. provided simulations with Amber force field, general discussion, and manuscript editing
+M.K. provided simulations with OPLS3e force field. 
+T.M.F. performed and analysed the solid-state NMR experiments.
 

### 2023-04-17 main.tex
Updates from Overleaf

@@
 
-Here we present the NMRlipids Databank---a community-driven, open-for-all database featuring programmatic access to atom-resolution MD simulations of lipid bilayers.
+Here we present the NMRlipids Databank --- a community-driven, open-for-all database featuring programmatic access to atom-resolution MD simulations of lipid bilayers.
 The programmatic access enables users to apply their data-driven approaches on the Databank, thus facilitating the creation of new tools for (and by) researchers in a wide range of fields covering academia and industry, from cell membrane biology and lipid nanoparticle formulations to computational chemistry and machine learning.
-As two usage examples of what is already possible, we demonstrate here (i) how data-driven analysis of water anisotropic diffusion in all the membrane systems available in the Databank can extend the scope of MD simulations to magnetic resonance imaging (MRI) and pharmacokinetics and (ii) how the Databank allows its users to analyse rare phenomena that are beyond the scope of standard MD simulation investigations.
+As two usage examples of what is already possible, we demonstrate here (i) how data-driven analysis of water anisotropic diffusion in all the membrane systems available in the Databank can extend the scope of MD simulations to magnetic resonance imaging (MRI) and pharmacokinetics, and (ii) how the Databank allows its users to analyse rare phenomena that are beyond the scope of standard MD simulation investigations.
 Wider adaptation of the NMRlipids Databank will open even more possibilities.
@@
     E) Flowchart for accessing results calculated from the NMRlipids Databank and stored to the \textit{Application layer}.
-    F) Currently available binary mixtures in the NMRlipids Databank. 
+    F) Currently available single lipid component and binary mixtures in the NMRlipids Databank. Colorbar shows the number of available simulations with the darkest green indicating three or more.  
     }
@@
 The data points from the best-performing simulations, based on rankings in panels B and C, are surrounded by black circles.
-    B) Best POPC simulations ranked based on the \textit{sn}-1 acyl chain order parameter quality.
+    B) Best POPC simulations ranked based on the \textit{sn}-1 acyl chain order parameter quality ($P^\mathrm{sn1}$).
+Also \textit{sn}-2 acyl chain ($P^\mathrm{sn2}$), headgroup ($P^\mathrm{hg}$) and total ($P^\mathrm{total}$) order parameter qualities, form factor quality ($FF_q$), and relative equilibration time for conformations ($\tau_\mathrm{rel}$) are shown.
     C) Best POPE simulations ranked based on the \textit{sn}-1 acyl chain order parameter quality. 
@@
     G) Scatter plots and Pearson correlation coefficients, $r$, for the membrane area per lipid, thickness, first minimum of X-ray scattering form factor and average order parameter of the \textit{sn}-1 acyl chain extracted from the NMRlipids Databank.
-All correlation coefficients have p-value below 0.001 as indicated by~$^{***}$.
+All correlation coefficients have p-value below 0.001.
 For more correlations see Fig.~\ref{fig:QualityCorrelationsSI}.
@@
 Histogrammed values are shown with black dots.
+Error bars show the standard error of the mean within each bin.
     }
@@
     \centering
-    \includegraphics[width=\linewidth]{Figures/permeation2.pdf}
+    \includegraphics[width=\linewidth]{Figures/permeation3.pdf}
     \caption{\textbf{A} Water diffusion, $D_\perp$, and permeability, $P$, through membranes, and lateral diffusion along the membrane, $D_\parallel$, illustrated in a multilamellar stack of lipid bilayers.
-    \textbf{B--E} Water permeation through membranes analyzed from the databank as a function of temperature, thickness, area per lipid, and acyl chain order.
-Values from simulations with non-zero permeation values are shown with blue dots.
-Histogrammed values are shown with black dots.
+    \textbf{B--E} Water permeation through membranes analyzed from the Databank as a function of temperature, thickness, area per lipid, and acyl chain order.
 Inset in B) shows the Arrhenius plot of permeation ($\ln(P)$ vs.
@@
     \textbf{F} Lateral diffusion of water as a function of hydration level. Experimental points for DMPC bilayers at 313\,K at different hydration levels are shown~\cite{rudakova04}.
-    \textbf{G--H} Diffusion anistoropy of water as a function of thickness and area per lipid. }
+    \textbf{G--H} Diffusion anistoropy of water as a function of thickness and area per lipid. 
+    Non-zero permeation and diffusion values from simulations are shown with blue dots.
+Histogrammed values are shown with black dots.
+Error bars show the standard error of the mean within each bin.
+Only bins with more than one microsecond of data in total were used for water permeation.}
     \label{fig:permeability}
@@
 The resulting non-zero values range between 0.3--322\,\textmu{}m/s with the mean of 14\,\textmu{}m/s and median of 8\,\textmu{}m/s.
-These values agree with the previously reported simulation results~\cite{venable19,camilo2022}, but are on average slightly larger than experimental values reported for PC-lipids in the liquid crystalline phase,~0.19--0.33\,\textmu{}m/s~\cite{jansen95}.
+These values agree with the previously reported simulation results~\cite{venable19,camilo2022}, but are on average slightly larger than experimental values reported for PC lipids in the liquid crystalline phase,~0.19--0.33\,\textmu{}m/s~\cite{jansen95}.
 Using the workflow depicted in Fig.~\ref{fig:overlay}E, we then plotted the observed permeabilities and their histogrammed values in Figs.~\ref{fig:permeability}B--E as a function of temperature, membrane thickness, area per lipid, and acyl chain order.
 As expected, the permeability increases with the temperature, giving an average energy barrier of 17\,$\pm$\,4\,$k_\mathrm{B}T$ for the water permeation from the Arrhenius plot in Fig.~\ref{fig:permeability}B.
-On the other hand, the permeability of water decreases on average when membranes become more packed, \textit{i.e.}, with decreasing area per lipid and increasing thickness and acyl chain order (Figs.~\ref{fig:permeability} C--E).
+On the other hand, the water permeability on average decreases when membranes become more packed, that is, with decreasing area per lipid and increasing thickness and acyl chain order (Figs.~\ref{fig:permeability}C--E).
 Permeation of water through bilayers depends on membrane properties also according to previous studies, but there is no established consensus on whether the area per lipid~\cite{nagle08} or bilayer thickness~\cite{frallicciardi22} is the main parameter determining the permeability.
 %is previously reported from simulations and experiments for some systems, but not for all~\cite{mathai07,venable19,frallicciardi22}. 
-Our analysis over the NMRlipids databank, containing significantly more data than what was available in previous studies, suggest non-linear dependencies on both of these parameters, yet the linear correlation would be a good approximation for thicknesses above $\sim$\,3.9\,nm and area per lipids below $\sim$69\,\AA{}$^2$ (insets in Figs.~\ref{fig:permeability} C--D).
+Our analysis over the NMRlipids Databank, containing significantly more data than what was available in previous studies, suggest non-linear dependencies on both of these parameters, yet the linear correlation would be a good approximation for thicknesses $\gtrsim$\,3.9\,nm and area per lipids $\lesssim$\,69\,\AA{}$^2$ (insets in Figs.~\ref{fig:permeability}C--D).
 %within the range of experimentally reported values  permeability which is 1-2 orders of magnitude slower than osmotic permeability \cite{jansen95,??}. 
-Clear dependencies of permeability on charged lipids, cholesterol, POPE, or hydration level were not observed in Fig.~\ref{fig:permeationSI} in the supplementary information.
+Clear dependencies of permeability on hydration level or the fraction of charged lipids, cholesterol, or POPE in the membrane were not observed (Fig.~\ref{fig:permeationSI}).
 %
@@
 
-To analyze how water diffusion anisotropy depends on membrane properties in a multi-lamellar lipid bilayer system, we calculated the water diffusion parallel to the membrane surface from all simulations in the NMRlipids databank using the workflow depicted in Fig.~\ref{fig:overlay}D.
-The parallel diffusion coefficient of water, $D_\parallel$, decreases with reduced hydration and increases with the temperature, but dependencies on membrane area per lipid, thickness, or fraction of charged lipids were not observed in Figs.~\ref{fig:permeability} and~\ref{fig:diffusionSI} when the workflow depicted in Fig.~\ref{fig:overlay}E was used.
-Simulation results are close to the experimental values with low hydration levels in Fig.~\ref{fig:permeability}F, but increase to approximately two times higher than the experimental value for bulk water diffusion value (3.1$\cdot$\,10$^{-9}$\,m$^2$/s at 313\,K~\cite{khakimov08}) with high hydration levels.
+To analyze how water diffusion anisotropy depends on membrane properties in a multi-lamellar lipid bilayer system, we calculated the water diffusion parallel to the membrane surface from all simulations in the NMRlipids Databank using the general workflow depicted in Fig.~\ref{fig:overlay}D.
+The parallel diffusion coefficient of water, $D_\parallel$, decreases with reduced hydration and increases with the temperature, but dependencies on the membrane area per lipid, thickness, or fraction of charged lipids were not observed in Figs.~\ref{fig:permeability} and~\ref{fig:diffusionSI} when the workflow depicted in Fig.~\ref{fig:overlay}E was used.
+Simulation results are close to the experimental values with low hydration levels in Fig.~\ref{fig:permeability}F, but increase to approximately two times higher than the experimental value for bulk water diffusion value (3.1$\times$\,10$^{-9}$\,m$^2$/s at 313\,K~\cite{khakimov08}) with high hydration levels.
 This is not surprising as the most common water model used in membrane simulations, TIP3P, overestimates the bulk water diffusion~\cite{pathirannahalage21}.
 To estimate the diffusion anisotropy of water, $D_\perp/D_\parallel$, in multilamellar membrane system, the permeability coefficients of water through membranes were translated to perpendicular diffusion coefficients, $D_\perp$, using the Tanner equation~\cite{tanner78,wasterby02}.
-The resulting perpendicular diffusion coefficients are approximately five orders of magnitude smaller than lateral diffusion coefficients of water (Figs.~\ref{fig:permeability} G and H), which is at the upper limit of the anisotropy estimated from the experimental data~\cite{nitsche19}.
-Significant increase in the diffusion anisotropy with membrane packing is observed, as $D_\perp/D_\parallel$ drifts away from one with decreasing area per lipid and increasing thickness in Figs.~\ref{fig:permeability} G and H.
-This follows from decreasing water permeability with membrane packing (Figs.~\ref{fig:permeability} C and D), while lateral diffusion remains approximately constant (Figs.~\ref{fig:diffusionSI} A and C).
+The resulting perpendicular diffusion coefficients are approximately five orders of magnitude smaller than the lateral diffusion coefficients of water (Figs.~\ref{fig:permeability}G--H), which is at the upper limit of anisotropy estimated from experimental data~\cite{nitsche19}.
+A significant increase in the diffusion anisotropy with membrane packing is observed, as $D_\perp/D_\parallel$ deviates further from unity with decreasing area per lipid and increasing thickness in Figs.~\ref{fig:permeability}G--H.
+This follows from decreasing water permeability with membrane packing (Figs.~\ref{fig:permeability}C--D), while lateral diffusion remains approximately constant (Figs.~\ref{fig:diffusionSI}A,~C).
 
 In summary, our results suggest that the bilayer packing has a substantial effect on anisotropic water diffusion in multi-membrane lipid systems.
-Several folds larger anisotropy in membranes with higher lateral density is expected to play a role in pharmacokinetic models not only for water but also for other hydrophilic molecules~\cite{nitsche19}.
-Furthermore, the enhanced understanding of this anisotropy may help in developing new diffusion tensor based MRI imaging methods where signals originate from anisotropic diffusion of water in biological material~\cite{topgaard20}.
+The several-fold larger anisotropy in membranes with higher lateral density is expected to play a role in pharmacokinetic models not only for water but also for other hydrophilic molecules~\cite{nitsche19}.
+Furthermore, the enhanced understanding of this anisotropy may help in developing new diffusion-tensor-based MRI imaging methods where signals originate from the anisotropic diffusion of water in biological matter~\cite{topgaard20}.
 
@@
 The focus of biomolecular simulations is moving from studies of individual molecules to larger complexes and even whole cells and organelles~\cite{johnson15,thornburg22,gupta22}.
-Simultaneously, machine learning -based models predicting behaviour of biomolecules and automatic approaches to parametrize models are emerging~\cite{jumper21,antila22b}.
-The resources delivered by the NMRlipids databank will support the development in both of these directions.
-Automatic quality evaluation and ranking of simulations against experimental data enables the selection of best simulations for specific applications without laborious manual force field evaluation.
+Simultaneously, machine-learning-based models for predicting the behaviour of biomolecules and automatic approaches to parametrize models are emerging~\cite{jumper21,antila22b}.
+The resources delivered by the NMRlipids Databank will support the development in both of these directions.
+Automatic quality evaluation and ranking of simulations against experimental data enable the selection of best simulations for specific applications without laborious manual force field evaluation.
 This also streamlines automatic parametrization procedures for atomistic and coarse grained simulations.
-Such practises fostering the accuracy of simulations are becoming increasingly important due the accumulation of small errors when complexity and size of simulated systems are increasing.
+Such practises for fostering the accuracy of simulations are becoming increasingly important due the accumulation of small errors when complexity and size of simulated systems are increasing.
 
-The open programmatic access to unprecedented amount of MD simulation data \textit{via} NMRlipids databank-API enables wide range of novel data-driven analyses and machine learning applications.
-The analyses of cholesterol flip--flop events (Fig.~\ref{fig:flip-flops}) and water permeation through membranes (Fig.~\ref{fig:permeability}) demonstrate how large amount of accessible simulation data in terms of quantity (\textit{e.g.}, simulation length and number of conformations) and content (\textit{e.g.}, lipid compositions and ion concentrations) enable analyses of rare phenomena that are beyond the current possibilities for a single research group.
-Such analyses also pave the way for applications of MD simulations in new fields as demonstrated here by analysing the anisotropic diffusion of water in membrane systems (Fig.~\ref{fig:permeability}) that is essential parameter in pharmacokinetic modeling and MRI imaging~\cite{nitsche19,topgaard20}.
-Furthermore, the NMRlipids databank-API delivers access to a training set data that paves the way for diverse machine learning applications to
+The open programmatic access to unprecedented amounts of MD simulation data through the NMRlipids Databank-API enables a wide range of novel data-driven analyses and machine learning applications.
+The analyses of cholesterol flip-flop events (Fig.~\ref{fig:flip-flops}) and water permeation through membranes (Fig.~\ref{fig:permeability}) demonstrate how a large amount of accessible simulation data in terms of quantity (\textit{e.g.}, simulation length and number of conformations) and content (\textit{e.g.}, lipid compositions and ion concentrations) enable analyses of rare phenomena that are beyond the current possibilities for a single research group.
+Such analyses also pave the way for applications of MD simulations in new fields, as demonstrated here by analysing an essential parameter in pharmacokinetic modeling and MRI imaging:\cite{nitsche19,topgaard20} the anisotropic diffusion of water in membrane systems (Fig.~\ref{fig:permeability}).
+Furthermore, the NMRlipids Databank-API delivers access to a training set data that paves the way for diverse machine learning applications to
 %For example, a machine learning model predicting electron density profiles from form factors would be highly useful in interpretation of scattering experiments.
@@
 predict membrane properties.
-Such applications are analogous to AlphaFold \cite{jumper21} and other tools that predict proteins structures from their sequence using artificial intelligence.
-These possibilities are particularly valuable for scientist who do not typically have access to large scale MD simulation data.
-Different types of expected applications of the NMRlipids databank in wide range of fields are listed in Table~\ref{tab:applications}, yet the scope of such applications
+Such applications are analogous to AlphaFold \cite{jumper21} and other tools that predict protein structures from their sequence using artificial intelligence.
+These possibilities are particularly valuable for scientists who do not typically have access to large scale MD simulation data.
+Expected applications of the NMRlipids Databank in a wide range of fields are listed in Table~\ref{tab:applications}, yet the scope of such applications
 %in biology, biotechnology, material science, biomolecular imaging and beyond
-is expected to further widen with increasing amount of data in the NMRlipids databank.
+is expected to further widen with increasing amount of data in the Databank.
 
@@
     \hline
-    Analyses of rare phenomena               & Lipid flip--flops, water permeation (Figs.~\ref{fig:flip-flops} and~\ref{fig:permeability}) & Membrane scientists \\
+    Analyses of rare phenomena               & Lipid flip-flops, water permeation (Figs.~\ref{fig:flip-flops} and~\ref{fig:permeability}) & Membrane scientists \\
     \\
@@
     \\
-    Applications that are outside typical scope of MD simulations & 
+    Applications that are outside the typical scope of MD simulations & 
     Anisotropic water diffusion for pharmacokinetics and MRI imaging applications (Fig.~\ref{fig:permeability}) & 
@@
     Best model for membranes with PC and PE lipids (Fig.~\ref{fig:quality}), lipid headgroup conformations~\cite{bacle21}, 
-    packing of PS~\cite{antila22b} and PE (Fig.~\ref{fig:quality}) containing membranes. &
+    packing of PS~\cite{antila22b} and PE (Fig.~\ref{fig:quality}) containing membranes &
     Scientists using MD simulations \\
@@
     Training and target data for the development of coarse grained models & 
-    Optimizing parameters of coarse grained models against NMRlipids databank, extracting continuum parameters for membranes. &
+    Optimizing parameters of coarse grained models against NMRlipids Databank, extracting continuum parameters for membranes &
     Scientists developing and using coarse grained MD simulations \\
@@
     Training set for machine learning applicatons &
-    Programmatic access to the data and results enables training of machine learning models for various applications, such as predictions of membrane properties from composition & Scientists building and using machine learning applications for biomolecules.
+    Programmatic access to the data and results enables training of machine learning models for various applications, such as predictions of membrane properties from composition & Scientists building and using machine learning applications for biomolecules
     \end{tabular}
-    \caption{Examples on applications of the NMRlipids databank.}
+    \caption{Examples on applications of the NMRlipids Databank.}
     \label{tab:applications}
@@
 The main barriers for such solutions have been the required commitment for the long term support of hardware and software, and the lack of incentives for researchers to share the data.
-The first issue is solved in the NMRlipids databank using the overlay design where raw data is distributed to already publicly available decentralized locations while the core of the databank is composed only of metadata stored in a version controlled git repository with an open access licence.
-On the other hand, the open collaboration approach developed in the NMRlipids project~\cite{botan15} creates incentives for sharing the data by offering authorship in published articles to the contributors.
-Advantages of such approach are demonstrated here for membrane simulations, but the concept can be applied also not only to other biomolecules, such as disordered proteins and membrane--protein systems, but also in other fields where similar barriers hinder the machine learning revolution, such as assignment of NMR spectra~\cite{klukowski22}.
+The first issue is solved in the NMRlipids Databank by the overlay design, where the raw data is distributed to already publicly available decentralized locations, while the core of the Databank is composed only of the metadata stored in a version-controlled git repository with an open-access license.
+On the other hand, the open collaboration approach developed in the NMRlipids Project~\cite{botan15} creates incentives for sharing the data by offering authorship in published articles to the contributors.
+Advantages of such an approach are demonstrated here for membrane simulations, but the concept can be applied not only to other biomolecules, such as disordered proteins and membrane--protein systems, but also in other fields where similar barriers hinder the machine learning revolution, such as the assignment of NMR spectra~\cite{klukowski22}.
 
@@
 
-\subsection{Structure of the databank}
-The overlay structure designed for the NMRlipids databank is composed of three layers illustrated in Fig.~\ref{fig:overlay}A.
-The {\it data layer} contains raw data that can be distributed to publicly available servers such as Zenodo (\url{www.zenodo.org}).
-The core content of the databank locates in the {\it databank layer} which is a git repository at \url{https://github.com/NMRLipids/Databank/} and is also permanently stored in Zenodo repository (\url{www.zenodo.org})~\cite{??}.
-The Essential information of each simulation is stored in a human and machine readable README.yaml file located at subfolders of \texttt{/Data/Simulations} folder in the {\it databank layer} repository.
-Each subfolder has a unique name constructed based on hash code of trajectory and topology files of each simulation.
-The README.yaml files in these folders contain access to all information that are needed for further analyses of simulations, such as links to the raw data and affiliations of universal molecule and atom names.
+\subsection{Structure of the Databank}
+The overlay structure designed for the NMRlipids Databank is composed of three layers (Fig.~\ref{fig:overlay}A).
+The {\it Data layer} contains raw data that can be distributed to publicly available servers such as Zenodo (\href{https://www.zenodo.org/}{zenodo.org}).
+The core content of the Databank locates in the {\it Databank layer}, which is a git repository at \href{https://github.com/NMRlipids/Databank/}{github.com/NMRlipids/Databank} and is also permanently stored in a Zenodo repository (\href{https://www.zenodo.org/}{zenodo.org})~\cite{??}.
+The essential information of each simulation is stored in a human-and-machine-readable README.yaml file located in a subfolder of the \texttt{/Data/Simulations} folder in the {\it Databank layer} repository; each subfolder has a unique name constructed based on a hash code of the trajectory and topology files of each simulation.
+The README.yaml files in these folders contain access to all information that is needed for further analysis of simulations, such as links to the raw data and associations with the universal molecule and atom names.
 The content of these files is described in detail in Table~\ref{tab:READMEkeys} in the supplementary information.
-Results from analyses of basic membrane properties (area per lipid, thickness, C-H bond order parameters, x-ray scattering form factors and relaxation of principal components) are stored in the same folders with README.yaml files.
-Experimental data used for ranking is stored in \texttt{/Data/experiments} folder, results from such ranking in \texttt{/Data/Ranking} and  relevant scripts in \texttt{/Scripts/} folder in the {\it databank layer} repository.
+Results from analyses of basic membrane properties (area per lipid, thickness, C--H bond order parameters, X-ray scattering form factors, and relaxation of principal components) are stored in the same folders as the README.yaml files.
+Experimental data used for ranking is stored in the \texttt{/Data/experiments} folder,
+the ranking results in \texttt{/Data/Ranking},
+and the relevant scripts in the \texttt{/Scripts/} folder in the {\it Databank layer} repository. 
 %Raw MD simulation data are stored in external publicly available and stable repositories (\textit{Data layer} in Fig.~\ref{fig:overlay}), such as Zenodo (\url{www.zenodo.org}), from where it can be downloaded whenever needed using the links in README.yaml files.
-The scripts in the NMRlipids databank are mainly written in Python and many of them utilize the MDAnalysis module~\cite{gowers2019mdanalysis,michaud2011mdanalysis}.
-The databank structure is illustrated more detailed in Fig.~\ref{DatabankStructure} in the supplementary information.
-Whenever specific files or folders are referred here, they locate at {\it databank layer} repository unless stated otherwise.
+The scripts in the NMRlipids Databank are mainly written in Python and many of them use the MDAnalysis module~\cite{gowers2019mdanalysis,michaud2011mdanalysis}.
+The Databank structure is illustrated in more detail in Fig.~\ref{DatabankStructure} in the supplementary information.
+Whenever specific files or folders are referred here, they locate at the {\it Databank layer} repository unless stated otherwise.
 
-\subsection{Molecule and atom naming convention} \label{naming}
-When analysing simulations, atoms and molecules needs to be often called by their names defined in the simulation trajectory.
-However, these names typically vary between force fields because the universal naming convention has not been defined for lipids.
-To enable automatic analyses over simulations with different atom and molecule names in the NMRlipids databank, we have defined unique naming conventions for molecules and atoms therein.
-Unique abbreviations used in the NMRlipids databank for each molecule are listed in Table~\ref{tab:abbreviations} in the supplementary information.
-Atom names used in simulation trajectories are connected to unique atom names using mapping files that are defined in the NMRlipids project (\url{https://nmrlipids.blogspot.com/2022/04/new-yaml-format-of-mapping-files.html}).
-These files are located at \texttt{/Scripts/BuildDatabank/mapping\_files} in the NMRlipids databank repository.
-These files also define whether an atom belongs to headgroup, glycerol backbone, or acyl chain region in a lipid.
-In practise, force field specific molecule names and mapping files
-names are given in the COMPOSITION dictionary in README.yaml files for each molecule in each simulation in the NMRlipids databank.
+\subsection{Universal naming convention for molecules and atoms} \label{naming}
+When analysing simulation trajectories, atoms and molecules often need to be  called by the names used in the trajectory.
+However, these names typically vary between force fields, as a universal naming convention has not been defined for lipids.
+To enable automatic analyses over all the simulations in the NMRlipids Databank, we have defined universal naming conventions for the molecules and atoms therein.
+The universal abbreviations used in the NMRlipids Databank for each molecule are listed in Table~\ref{tab:abbreviations} in the supplementary information.
+The atom names used in simulation trajectories are connected to the universal atom names using the mapping files defined in the NMRlipids Project (\url{https://nmrlipids.blogspot.com/2022/04/new-yaml-format-of-mapping-files.html}).
+These files are located at \texttt{/Scripts/BuildDatabank/mapping\_files} in the NMRlipids Databank repository.
+These files also define whether an atom belongs to the headgroup, glycerol backbone, or acyl chain region in a lipid.
+In practise, the force-field-specific molecule names and mapping file
+names are given in the COMPOSITION dictionary in the README.yaml files for each molecule in each simulation in the NMRlipids Databank.
 
-\subsection{Adding data into the NMRlipids databank}
-The NMRlipids databank is open for additions of simulation data by anyone.
-The list information that a contributor has to deliver is given in Table~\ref{tab:READMEkeys}.
-The rest of the information to be stored in README.yaml files, also listed in Table~\ref{tab:READMEkeys}, will be automatically extracted using the \texttt{/Scripts/BuildDatabank/AddData.py} script.
-In practise, the manually entered data is first stored into an info.yaml file that is then added into \texttt{/Scripts/BuildDatabank/info\_files} folder via pull request.
-To avoid ineligible entries and minimize human errors, the pull requests are monitored before the acceptance and generation of README.yaml files.
-Currently, the NMRlipids databank is composed of simulations that are found from the Zenodo repository with an appropriate licence.
-Most, but not all, of these trajectories originate from previous NMRlipids projects~\cite{botan15,catte16,antila19,bacle21}.
+\subsection{Adding data to NMRlipids Databank}
+The NMRlipids Databank is open for additions of simulation data by anyone.
+The list of information that a contributor has to deliver is given in Table~\ref{tab:READMEkeys}.
+The rest of the information to be stored in the README.yaml files, also listed in Table~\ref{tab:READMEkeys}, will be automatically extracted using the \texttt{/Scripts/BuildDatabank/AddData.py} script.
+In practise, the manually entered data is first stored into an info.yaml file that is then added to the \texttt{/Scripts/BuildDatabank/info\_files} folder trough a git pull request.
+To avoid ineligible entries and minimize human errors, the pull requests are monitored before the acceptance and generation of the README.yaml files.
+Currently, the NMRlipids Databank is composed of simulations that are found from the Zenodo repository with an appropriate license; most, but not all, of these trajectories originate from previous NMRlipids projects~\cite{botan15,catte16,antila19,bacle21}.
 
 \subsection{Experimental data}
-Experimental data used in the quality evaluation, currently composed of C--H bond order parameters and X-ray scattering form factors, are stored in \texttt{/Data/experiments} in the NMRlipids databank repository.
+Experimental data used in the quality evaluation, currently composed of C--H bond order parameters and X-ray scattering form factors, are stored in \texttt{/Data/experiments} in the NMRlipids Databank repository.
 Similarly to simulations, each experimental data set has a README.yaml file containing all the relevant information about the experiment.
 The keys and their descriptions for the experimental data are given in Table~\ref{tab:READMEkeysEXP}.
-NMR data currently in the NMRlipids databank are taken from Refs.~\citenum{scherer87,ferreira13,antila19,melcr20,bacle21} and X-ray scattering data from Refs.~\citenum{Kucerka08a,kucerka11,pan12b,pan14,kucerka15,NMRlipidsIII}.
-In addition, previously unpublished NMR data for POPE, POPG and DOPC was acquired as described in the supplementary information (Figs.~\ref{POPEexp1}-\ref{fig:DOPCexp}) and contributed to the databank.
+The NMR data currently in the NMRlipids Databank are taken from Refs.~\citenum{scherer87,ferreira13,antila19,melcr20,bacle21} and the X-ray scattering data from Refs.~\citenum{Kucerka08a,kucerka11,pan12b,pan14,kucerka15,NMRlipidsIII}.
+In addition, previously unpublished NMR data for POPE, POPG, and DOPC was acquired as described in the supplementary information (Figs.~\ref{POPEexp1}--\ref{fig:DOPCexp}) and contributed to the Databank.
 
 \subsection{Analysing simulations}
-In practise, simulations in the NMRlipids databank can be analyzed by executing a program that loops over README.yaml files in the {\it databank layer}, downloads the data using the information in README.yaml files, and then performs the desired analysis on a local computer utilising the universal naming conventions for molecules and atoms defined in README.yaml and mapping files.
-This procedure is illustrated in Fig.~\ref{fig:overlay}D and examples of codes that perform such analyses in practise are listed in Tables~\ref{tab:codes} and~\ref{tab:codesApplication}.
+In practise, simulations in the NMRlipids Databank can be analyzed by executing a program that
+1) loops over the README.yaml files in the {\it Databank layer},
+2) downloads the data using the information in the README.yaml files, and then
+3) performs the desired analysis on a local computer utilising the universal naming conventions for molecules and atoms defined in the README.yaml and mapping files.
+This general procedure is illustrated in Fig.~\ref{fig:overlay}D, and practical examples of codes that perform such analyses are listed in Tables~\ref{tab:codes} and~\ref{tab:codesApplication}.
 The equilibration period given by the user (TIMELEFTOUT in Table~\ref{tab:READMEkeys}) is discarded from the trajectories in all analysis codes.
@@
 \subsection{Principal Components Analysis of equilibration of simulations}
-To estimate how well conformational ensembles of lipids are converged in trajectories, the Principal Components Analysis (PCA) following the PCALipids protocol was used \cite{buslaev16,buslaev2020principal}.
+To estimate how well conformational ensembles of lipids are converged in trajectories, the Principal Component Analysis (PCA) following the PCALipids protocol was used \cite{buslaev16,buslaev2020principal}.
 %, the lipids of a specific type were first selected, and trajectories of individual lipids were concatenated. 
-To this end, each lipid configuration was first aligned to the average structure of that lipid type, and PCA analysis was then applied on the cartesian coordinates of all heavy atoms of the lipid.
+To this end, each lipid configuration was first aligned to the average structure of that lipid type, and PCA analysis was then applied on the Cartesian coordinates of all heavy atoms of the lipid.
 Because the motions along the first, major, principal component are the slowest ones \cite{buslaev16}, the equilibration of each lipid type was estimated from the ratio between the distribution convergence of the trajectories projected on the first PC and the trajectory length, $\tau_{\rm rel} = \tau_{\text{convergence}}/\tau_{\text{sim}}$~\cite{buslaev16, buslaev2020principal}.
-If $\tau_{\rm rel} < 1$, simulations can be considered to be sufficiently long for lipid molecules to sample their conformational ensembles, while in simulations with $\tau_{\rm rel} > 1$ individual molecules may not have fully sampled their conformational ensembles.
+If $\tau_{\rm rel} < 1$, simulations can be considered to be sufficiently long for the lipid molecules to have sampled their conformational ensembles, while in simulations with $\tau_{\rm rel} > 1$ individual molecules may not have fully sampled their conformational ensembles.
 Rigid molecules that do not exhibit significant conformational fluctuations, such as sterols, were excluded from the analysis.
@@
 %\begin{equation}
-    $\tau_{\text{convergence}} = k*\tau_{\text{autocorrelation}}$,
+    $\tau_{\text{convergence}} = k\tau_{\text{autocorrelation}}$,
 %\end{equation}
 because calculation of autocorrelation decay times is faster and computationally more stable than direct calculation of distribution convergence times~\cite{buslaev16, buslaev2020principal}.
-The empirical coefficient $k$ is equal to $k = 49$ and was calculated based on the analysis of 8 trajectories with the length of more than 200\,ns including simulations of POPC, POPS, POPE, POPG, and DPPC with CHARMM36 force fields.
+The empirical coefficient $k = 49$ was calculated based on the analysis of 8 trajectories with the length of more than 200\,ns, including simulations of POPC, POPS, POPE, POPG, and DPPC with the CHARMM36 force field.
 %All simulations, used for $k$ calculation, were at least 200 ns long.
-Because the coefficient does not depend on the force field~\cite{buslaev16}, the value determined from these CHARMM36 simulations ,can be used for all simualtions in the Databank.
-The script that calculates the equilibration of lipids is available at \texttt{Scripts/BuildDatabank/NMRPCA\_timerelax.py} in the NMRlipids databank repository.
-The resulting values are stored in files named \texttt{eq\_times.json} at folders in \texttt{/Data/Simulations} in the NMRlipids databank repository.
+Because the coefficient $k$ does not depend on the force field~\cite{buslaev16}, the value determined from these CHARMM36 simulations can be used for all simulations in the Databank.
+The script that calculates the equilibration of lipids is available at \texttt{Scripts/BuildDatabank/NMRPCA\_timerelax.py} in the NMRlipids Databank repository.
+The resulting values are stored in files named \texttt{eq\_times.json} at folders in \texttt{/Data/Simulations} in the NMRlipids Databank repository.
 
@@
 However, order parameters for simulations with $\tau_{\rm rel} > 1$ may be influenced by the starting structure and thereby their error bars may be underestimated.
-The script that calculates C--H bond order parameters from all simulations in the NMRlipids databank is available at \texttt{/Scripts/AnalyzeDatabank/calcOrderParameters.py} in the NMRlipids databank repository.
+The script that calculates C--H bond order parameters from all simulations in the NMRlipids Databank is available at \texttt{/Scripts/AnalyzeDatabank/calcOrderParameters.py} in the NMRlipids Databank repository.
 The resulting order parameters are stored for all simulations in files named %\newline
-[\texttt{lipid\_name}]\texttt{OrderParameters.json} at folders in \texttt{/Data/Simulations} in the NMRlipids databank repository.
+[\texttt{lipid\_name}]\texttt{OrderParameters.json} at folders in \texttt{/Data/Simulations} in the NMRlipids Databank repository.
 
@@
 \end{equation}
-where $\Delta \rho_e(z)$ is the difference between total and solvent electron densities, and $D$ is the simulation box size in the $z$-direction (normal to the membrane).
+where $\Delta \rho_e(z)$ is the difference between the total and solvent electron densities, and $D$ is the simulation box size in the $z$-direction (normal to the membrane).
 For the calculation of density profiles, atom coordinates were first centred around the centre of mass of lipid molecules for every time frame, and a histogram of these centred positions, weighted with the number of electrons in each atom, was then calculated with the bin width of $1/3$~\AA{}.
 Electron density profiles were then calculated as an average of these histograms over the time frames in simulations.
-The script to calculate form factors for all simulations in the NMRlipids databank is available at \texttt{Scripts/AnalyzeDatabank/calc\_FormFactors.py}.
-The resulting form factors are stored for all simulations in files named \texttt{FormFactor.json} at folders in \texttt{/Data/Simulations} in the NMRlipids databank repository.
+The script to calculate form factors for all simulations in the NMRlipids Databank is available at \texttt{Scripts/AnalyzeDatabank/calc\_FormFactors.py}.
+The resulting form factors are stored for all simulations in files named \texttt{FormFactor.json} at folders in \texttt{/Data/Simulations} in the NMRlipids Databank repository.
 
-\subsection{Calculation of a bilayer area per lipid and thickness}
-Area per lipids of bilayers were calculated by dividing the time-averaged area of the simulation box with the total number of lipid and surfactant molecules in the simulation (see Table~\ref{tab:abbreviations} for the list of molecules considered as lipids or surfactants).
-The script that calculates area per lipids from all simulations in the NMRlipids databank repository is available at \texttt{Scripts/AnalyzeDatabank/calcAPL.py} in the NMRlipids databank repository.
+\subsection{Calculation of area per lipid and bilayer thickness}
+Area per lipids of bilayers were calculated by dividing the time-averaged area of the simulation box with the total number of lipids and surfactant molecules in the simulation (see Table~\ref{tab:abbreviations} for the list of molecules considered as lipids or surfactants).
+The script that calculates the area per lipid from all simulations in the NMRlipids Databank repository is available at \texttt{Scripts/AnalyzeDatabank/calcAPL.py} in the NMRlipids Databank repository.
 The resulting area per lipids are stored for all simulations in files named \texttt{apl.json} at folders in \texttt{/Data/Simulations}.
 
-Thicknesses of lipid bilayers were calculated from the intersections of lipid and water electron densities.
-The script that calculates thicknesses of all simulations in the NMRlipids databank is available at \texttt{Scripts/AnalyzeDatabank/calc\_thickness.py} in the NMRlipids databank repository.
-The resulting thicknesses are stored in files named \texttt{thickness.json} at folders in \texttt{/Data/Simulations} in the NMRlipids databank repository.
+Thicknesses of lipid bilayers were calculated from the intersection points of lipid and water electron densities.
+The script that calculates the thicknesses of all simulations in the NMRlipids Databank is available at \texttt{Scripts/AnalyzeDatabank/calc\_thickness.py} in the NMRlipids Databank repository.
+The resulting thicknesses are stored in files named \texttt{thickness.json} at folders in \texttt{/Data/Simulations} in the NMRlipids Databank repository.
 
 \subsection{Quality evaluation of C--H bond order parameters}
-As the first step to evaluate simulation qualities against experimental data, each simulation is connected to an experimental data if molar concentrations of all molecules are within $\pm$3 percentage units, charged lipids have the same counterions, and temperatures are within $\pm$2\,K.
-For molar concentrations of water, the exact hydration level is considered only for systems with molar water to lipid ratio below 25, otherwise the systems are considered as fully hydrated.
-In practise, this is implemented by adding the path of the experimental data into the simulation README.yaml file using the \texttt{/Scripts/BuildDatabank/searchDATABANK.py} script in the NMRlipids databank repository.
+As the first step to evaluate simulation qualities against experimental data, a simulation is connected to an experimental data set if the molar concentrations of all molecules are within $\pm$3 percentage units, charged lipids have the same counterions, and temperatures are within $\pm$2\,K.
+For molar concentrations of water, the exact hydration level is considered only for systems with molar water-to-lipid ratio below 25, otherwise the systems are considered as fully hydrated.
+In practise, the connection is implemented by adding the experimental data path into the simulation README.yaml file using the \texttt{/Scripts/BuildDatabank/searchDATABANK.py} script in the NMRlipids Databank repository.
 
@@
 where $f(t)$ is the %first order 
-Student's t-distribution, $n$ is the number of independent sample points for each C--H bond which equals the number of lipids in a simulation, $S_\mathrm{CH}$ is the sample mean from Eq.~\ref{OP}, $s$ is the variance of $S_\mathrm {CH}$ calculated over individual lipids, $S_\mathrm{exp}$ is the experimental value, and $\Delta S_\mathrm{exp}$ its error.
+Student's t-distribution, $n$ is the number of independent sample points for each C--H bond (which equals the number of lipids in a simulation), $S_\mathrm{CH}$ is the sample mean from Eq.~\eqref{OP}, $s$ is the variance of $S_\mathrm {CH}$ calculated over individual lipids, $S_\mathrm{exp}$ is the experimental value, and $\Delta S_\mathrm{exp}$ its error.
 The error of $\Delta S_\mathrm{exp} = 0.02$ is currently assumed for all experimental order parameters~\cite{ollila16}, yet more accurate ones may be available in the future~\cite{wurl22}.
-Because a lipid bilayer simulation contains at least dozens of lipids, the Student's t-distribution could be safely approximated with a normal distribution.
-However, the normal distribution gives probability values that are below the numerical accuracy of computers when simulation values are far from experiments.
-To avoid such numerical instabilities, we use the first order Student's t-distribution having slightly higher probabilities for values far away from the mean.
-On the other hand, some force fields exhibit too slow dynamics which leads to large error bars in order parameter values~\cite{antila21a}.
+Because a lipid bilayer simulation contains at least dozens of lipids, the Student's t-distribution could be safely approximated with a normal distribution. However,
+with the quality of currently available force fields, the simulation values can be so far from experiments that a
+normal distribution leads to probability values below the numerical accuracy of computers.
+To avoid such numerical instabilities, we opted to use the first order Student's t-distribution that has slightly higher probabilities for values far away from the mean.
+On the other hand, some force fields exhibit too slow dynamics, which leads to large error bars in the $S_\mathrm{CH}$ values~\cite{antila21a}.
 Such artificially slow dynamics widens the Student's t-distribution in Eq.~\ref{probability}, thereby increasing the probability to find the simulated value within experimental error bars.
-Therefore, the order parameters with simulation error bars above the experimental error 0.02 are not included in the quality evaluation.
+Therefore, the $S_\mathrm{CH}$ with simulation error bars above the experimental error 0.02 are not included in the quality evaluation.
 
-To streamline the comparison between simulations, we define the qualities for different fragments (frag = \textit{sn}-1, \textit{sn}-2, headgroup or total referring to all order parameters within a molecule) within each lipid type in a simulation as
+To streamline the comparison between simulations, we define the average qualities for different fragments (frag = '\textit{sn}-1', '\textit{sn}-2', 'headgroup', or 'total', with the last referring to all order parameters within a molecule) within each lipid type in a simulation as
 \begin{equation}
@@
 \end{equation}
-where $\langle P[\mathrm{lipid}]\rangle_\mathrm{frag}$ is the average of individual C-H bond order parameters qualities within the fragment, and $F_\mathrm{frag}[\mathrm{lipid}]$ is the percentage of order parameters for which the quality is available within the fragment.
+where $\langle P[\mathrm{lipid}]\rangle_\mathrm{frag}$ is the average of the individual $S_\mathrm{CH}$ qualities within the fragment, and $F_\mathrm{frag}[\mathrm{lipid}]$ is the percentage of order parameters for which the quality is available within the fragment.
 %, and frag can be \textit{sn}-1, \textit{sn}-2, headgroup or total (all order parameters within a molecule). 
-The overall quality of different fragments in a simulation (frag = tails, headgroup or total) are then defined as a molar fraction weighted average over different lipid components
+The overall quality of different fragments in a simulation (frag = 'tails', 'headgroup', or 'total') are then defined as a molar-fraction-weighted average over different lipid components
 \begin{equation}
@@
 \end{equation}
-where $\chi_\mathrm{lipid}$ is the molar fraction of a lipid in the bilayer and "tails" refer to the average of all acyl chains.
+where $\chi_\mathrm{lipid}$ is the molar fraction of a lipid in the bilayer and 'tails' refer to the average of all acyl chains.
 
-The quality evaluation of order parameters is implemented in \texttt{/Scripts/BuildDatabank/QualityEvaluation.py} in the NMRlipids databank repository.
-The resulting qualities for each order parameters are stored in files named  [\texttt{lipid\_name}]\texttt{ \_OrderParameters\_quality.json}, for individual lipids in files named [\texttt{lipid\_name}]\texttt{\_FragmentQuality.json}, and for overall quality for fragments in files named \texttt{system\_quality.json} at folders in \texttt{/Data/Simulations} in the NMRlipids databank repository.
+The quality evaluation of order parameters is implemented in \texttt{/Scripts/BuildDatabank/QualityEvaluation.py} in the NMRlipids Databank repository.
+The resulting qualities for each $S_\mathrm{CH}$ are stored in files named  [\texttt{lipid\_name}]\texttt{ \_OrderParameters\_quality.json}, for individual lipids in files named [\texttt{lipid\_name}]\texttt{\_FragmentQuality.json}, and for overall quality for fragments in files named \texttt{system\_quality.json} at folders in \texttt{/Data/Simulations} in the NMRlipids Databank repository.
 
@@
 %While C-H bond order parameters relate to the conformational ensembles of individual lipids, the x-ray scattering form factors depend on membrane dimensions and density distribution~\cite{ollila16}.
-Because experiments give form factors only in relative intensity scale, they should be scaled before comparing with the simulation data.
+Because experiments give form factors only on a relative intensity scale, they should be scaled before comparing with the simulation data.
 Here we use the scaling coefficient for experimental intensities defined in the SIMtoEXP program~\cite{kucerka10}
@@
 
-Also, a quality measure based on differences in simulated and experimental form factors across available $q$-range is defined in the SIMtoEXP program~\cite{kucerka10}.
-However, the lobe heights in simulated form factors depend on the simulation box size as shown in Fig.~\ref{fig:sizedependence}.
-Therefore, the quality measure defined in SIMtoEXP  would also depend on the simulation box size.
-In contrast, locations of form factor minima are independent on simulation box size in Fig.~\ref{fig:sizedependence}.
-Here we use only the location of the first form factor minimum for the quality evaluation because location of the second minimum is difficult to detect automatically in some experimental data sets due to fluctuations, such as for the POPE data in Figs.~\ref{fig:quality} D and E.
-The first minimum correlates well with the thickness of a membrane (Fig.~\ref{fig:quality} F), although the correlation of the second minima would be even stronger (Fig.~\ref{fig:QualityCorrelationsSI}).
-In practise, we first filter the fluctuations from the form factor data using Savitzky--Golay filter (window length 30 and polynomial order 1) and locate the first minima above 0.1\,\AA$^{-1}$ from both simulation and experimental data.
-The quality of a form factor is then defined as Euclidean distance between the minima in simulated and experimental form factors, $FF_q = |FF_\mathrm{min}^\mathrm{sim}-FF_\mathrm{min}^\mathrm{exp}| \times 100$.
+Also, a quality measure based on differences in the simulated and experimental form factors across the available $q$-range is defined in the SIMtoEXP program~\cite{kucerka10}.
+However, the lobe heights in the simulated form factors depend on the simulation box size, as shown in Fig.~\ref{fig:sizedependence}; consequently, the quality measure defined in SIMtoEXP  would also depend on the simulation box size.
+In contrast, locations of the form factor minima (or, in precise terms: the minima of the absolute value of the form factor) are independent of the simulation box size (Fig.~\ref{fig:sizedependence}).
+Here we use only the location of the first form factor minimum for quality evaluation,
+because (due to fluctuations) the location of the second minimum is difficult to detect automatically in some experimental data sets, such as the POPE data in Figs.~\ref{fig:quality}D, E.
+The first minimum correlates well with the thickness of a membrane (Fig.~\ref{fig:quality}F), although the correlation of the second minima would be even stronger (Fig.~\ref{fig:QualityCorrelationsSI}).
+In practise, we first filter the fluctuations from the form factor data using the Savitzky--Golay filter (window length 30 and polynomial order 1) and locate the first minimum at  $q>$\,0.1\,\AA$^{-1}$ from both simulation ($F\!F_\mathrm{min}^\mathrm{sim}$) and experiment ($F\!F_\mathrm{min}^\mathrm{exp}$).
+The quality of a form factor is then defined as the Euclidean distance between the minima locations: $F\!F_q = |F\!F_\mathrm{min}^\mathrm{sim}-F\!F_\mathrm{min}^\mathrm{exp}| \times 100$.
 
-The quality evaluation of form factors is implemented in \texttt{/Scripts/BuildDatabank/QualityEvaluation.py} in the NMRlipids databank repository.
-The resulting form factor qualities are stored in files named \texttt{FormFactorQuality.json} at folders in \texttt{/Data/Simulations} in the NMRlipids databank repository.
+The quality evaluation of form factors is implemented in \texttt{/Scripts/BuildDatabank/QualityEvaluation.py} in the NMRlipids Databank repository.
+The resulting form factor qualities are stored in files named \texttt{FormFactorQuality.json} at folders in \texttt{/Data/Simulations} in the NMRlipids Databank repository.
 
-\subsection{Calculation of lipid flip--flops}
-Flip--flop rates were calculated using \texttt{AssignLeaflets} and \texttt{FlipFlop} tools from LiPyphilic package~\cite{LiPyphilic2021}.
-Headgroup atoms of each molecule as defined in the mapping file were used to determine in which leaflet the molecule locates.
-The midplane cut-off defining the region between leaflets was 1\,nm and frame cutoff was 100.
-This means that if the headgroup of a molecule entered within the distance of 1\,nm from the bilayer midplane and was found from the opposing leaflet after 100 steps, this event was considered as a successful flip--flop event.
-The code that finds flip--flop events from all simulations in the NMRlipids databank is available at \texttt{scripts/FlipFlop.py} and the results at \texttt{Data/Flipflops/} in the repository at~\url{https://github.com/NMRLipids/DataBankManuscript/}.
+\subsection{Calculation of lipid flip-flops}
+Flip-flop rates were calculated using the \texttt{AssignLeaflets} and \texttt{FlipFlop} tools from the LiPyphilic package~\cite{LiPyphilic2021}.
+Headgroup atoms of each molecule, as defined in the mapping file, were used to determine in which leaflet the molecule locates.
+The midplane cut-off, defining the region between leaflets, was 1\,nm and the frame cutoff was 100.
+This means that if the headgroup of a molecule entered within the distance of 1\,nm from the bilayer midplane and was found in the opposing leaflet after 100 steps, this event was considered as a successful flip-flop event.
+The code that finds the flip-flop events from all simulations in the NMRlipids Databank is available at \texttt{scripts/FlipFlop.py}, and the results at \texttt{Data/Flipflops/} in the repository at~\url{https://github.com/NMRLipids/DataBankManuscript/}.
 
-\subsection{Analysing anisotropic diffusion of water in membrane environment from the NMRlipids databank}
+\subsection{Analysing anisotropic diffusion of water in a membrane environment from NMRlipids Databank}
 
@@
 
-Water permeability through membranes was calculated from equation $P=r/2c_w$, where $r$ is the rate of permeation events per time and area, and $c_w$=33.3679\,(nm)$^{-3}$ is the concentration of water in bulk~\cite{venable19}.
+Water permeability through membranes was calculated from equation $P=r/2c_w$, where $r$ is the rate of permeation events per time and area, and $c_w=$\,33.3679\,nm$^{-3}$ is the concentration of water in bulk~\cite{venable19}.
 The number of permeation events in each trajectory was calculated using the code by Camilo et al.~\cite{camilo2022}, available at~\url{https://github.com/crobertocamilo/MD-permeation}.
 %This code was used within the loop that goes over all simulations in the NMRlipids databank and extracts the required information using the information in README.yaml files. 
-The code that calculates permeabilities for all simulations in the NMRlipids databank is available at \texttt{/scripts/calcMD-PERMEATION.py} and the resulting permeabilities are stored at \texttt{/Data/MD-PERMEATION} in the repository containing all analyses specific for this publication at~\url{https://github.com/NMRLipids/DataBankManuscript/}.
+The code that calculates permeabilities for all simulations in the NMRlipids Databank is available at \texttt{/scripts/calcMD-PERMEATION.py}, and the resulting permeabilities are stored at \texttt{/Data/MD-PERMEATION} in the repository containing all analyses specific for this publication at~\url{https://github.com/NMRLipids/DataBankManuscript/}.
 %While the order parameters, form factors, area per lipid and thickness are stored within the NMRlipids databank (\url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations}), further analyses can be conventiently stored in separate repositories with the same folder structure based on hash identities of trajectory and topology files.
 For example, results from further analyses performed here are stored in folders at \url{https://github.com/NMRLipids/DataBankManuscript/tree/main/Data}.
-This repository is organized similarly to the NMRlipids databank repository, enabling the upcycling of also the analyzed data without overloading the main NMRlipids databank repository.
+This repository is organized similarly to the NMRlipids Databank repository, enabling the upcycling of also the analyzed data without overloading the main NMRlipids Databank repository.
 
-The lateral diffusion of water along the membrane surface, $D_{||}$, was calculated with Einstein's equation using \texttt{-lateral} option in \texttt{gmx msd} program within the Gromacs software package~\cite{gromacsMANUAL}.
-The code that calculates $D_{||}$ for water from all simulations in the NMRlipids databank is available at \texttt{/scripts/calcWATERdiffusion.py}, and the resulting diffusion coefficients are stored at \texttt{/Data/WATERdiffusion} in the repository at~\url{https://github.com/NMRLipids/DataBankManuscript/}.
+The lateral diffusion of water along the membrane surface, $D_\parallel$, was calculated with the Einstein's equation using the \texttt{-lateral} option in the \texttt{gmx\,msd} program within the Gromacs software package~\cite{gromacsMANUAL}.
+The code that calculates $D_\parallel$ for water from all simulations in the NMRlipids Databank is available at \texttt{/scripts/calcWATERdiffusion.py}, and the resulting diffusion coefficients are stored at \texttt{/Data/WATERdiffusion} in the repository at~\url{https://github.com/NMRLipids/DataBankManuscript/}.
 
-Water diffusion in the perpendicular direction of lipid bilayers in a multilamellar stack was estimated from the Tanner equation $D_\mathrm{\perp} = \frac{D_{||} P z_w}{D_{||} + P z_w}$~\cite{tanner78,wasterby02}, where the water layer thickness, $z_w$, was estimated by subtracting bilayer thickness from the size of the simulation box in membrane normal direction.
+Water diffusion along the perpendicular direction of lipid bilayers in a multilamellar stack was estimated from the Tanner equation $D_\mathrm{\perp} = \frac{D_\parallel P z_w}{D_\parallel + P z_w}$~\cite{tanner78,wasterby02}, where the water layer thickness, $z_w$, was estimated by subtracting the bilayer thickness from the size of the simulation box in the membrane normal direction.
 

### 2023-04-18 SI.tex
Update text
Update pdfs

--- 2023-01-31 SI.tex
@@
 \pagebreak
-\section{Correlations between area per lipid and thickness with order parameters and form factors}
+\section{Correlations of area per lipid and thickness with order parameters and form factors}
 
@@
     \caption{Scatter plots and Pearson correlation coefficients, $r$, for the membrane area per lipid with X-ray scattering form factor minima (A and B), and for thickness with the average order parameter of the {\textit sn}-1 acyl chain (B) and with the second minimum from X-ray scattering form factors (D) extracted from the NMRlipids databank.
-All correlation coefficients have p-value below 0.001 as indicated by $^{***}$.}
+All correlation coefficients have p-value below 0.001.}
     \label{fig:QualityCorrelationsSI}
@@
 \pagebreak
-\section{Dependence of form factor and order parameters in the size of simulation}
+\section{Dependence of form factor and order parameters on the simulation box size}
 
@@
     \includegraphics[width = 240mm]{Figures/SizeDependence.pdf}
-    \caption{Dependence of form factor, electron density profiles along membrane normal, and C-H bond orderparameters (from top to bottom) on simulation box size with different cholesterol concentration.
-Simulations with 64, 256 or 1024 POPC lipids taken from Ref.~\citenum{javanainen_matti_2021_7035350}.
+    \caption{Dependence of the form factor $F(q_z)$, the electron density profiles along membrane normal, and the C--H bond order parameters $S_\mathrm{CH}$ (from top to bottom) on the simulation box size (with different columns showing different cholesterol concentrations).
+Simulations with 64, 256, and 1024 POPC lipids are from Ref.~\citenum{javanainen_matti_2021_7035350}.
 }
@@
     \centering
-    \includegraphics[width = 140mm]{Figures/POPC_POPE_dataSI.pdf}
+    \includegraphics[height = 0.9\textheight]{Figures/totalrankingCOMPL.pdf}
+    \caption{Top 50 simulations in the NMRlipids Databank ranked based on the C--H bond order parameter quality against experiments.}
+    \label{fig:top50simulations}
+\end{figure}
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[height = 0.9\textheight]{Figures/POPC_POPE_dataSI.pdf}
     \caption{Simulations with the data for both POPC (top) and POPE (bottom) directly compared with the experimental data.
-      The area per lipid in simulations increase from left to right.
-      Simulations with the best overall quality for POPC and POPE order parameters are circled by a square.
+      The area per lipid increases from left to right.
+      Simulations with the best overall quality for POPC and POPE order parameters are highlighted with a solid border.
   }
@@
     \centering
-    \includegraphics[width = 180mm]{Figures/permeationSI.pdf}
-    \caption{Water permeation through membranes analyzed from the databank as a function of (A) hydration level, (B) fraction of cholesterol, (C) fraction of charged lipids, and (D) fraction of POPE in membrane.
+    \includegraphics[width = \textwidth]{Figures/permeationSI.pdf}
+    \caption{Water permeation through membranes analyzed from the Databank as a function of (A) hydration level, (B) fraction of cholesterol, (C) fraction of charged lipids, and (D) fraction of POPE in membrane.
 Values from simulations with non-zero permeation values are shown with blue dots.
@@
     \centering
-    \includegraphics[width = 180mm]{Figures/LateralDiffusionSI.pdf}
+    \includegraphics[width = \textwidth]{Figures/LateralDiffusionSI.pdf}
     \caption{Lateral diffusion of water as a function of (A) area per lipid, (B) temperature, (C) membrane thickness, and (D) fraction of charged lipids in a membrane.}
@@
   \includegraphics[width=\textwidth]{Figures/DataBankChart.pdf}
-  \caption{Structure of the NMRlipids databank.
-Manually added input data (blue boxes) includes basic information on the simulation, permanent links to the raw data, and experimental data if available.
-The databank entries (red box) and analysis results (green boxes), locating at \url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations} are automatically generated by the computer programs included in the NMRlipids databank (yellow boxes).
-Because raw data are not permanently stored but can be accessed based on the information in the databank, this connection is marked with the dashed line.}
+  \caption{Structure of the NMRlipids Databank.
+Manually added input data (blue boxes) include basic information on the simulation, permanent links to the raw data, and experimental data if available.
+The databank entries (red box) and analysis results (green boxes), at \url{https://github.com/NMRlipids/Databank/tree/main/Data/Simulations} are automatically generated by the computer programs included in the NMRlipids Databank (yellow boxes).
+Because the raw data are not permanently stored but can be accessed based on the information in the Databank, this connection is marked with a dashed line.}
   \label{DatabankStructure}
 \end{figure}
+
+\begin{table}[!h]
+    \centering
+    %\begin{tabular}{c c c c}
+    \begin{tabular}{p{3.0cm}  p{5.0cm}  p{3.0cm}  p{4.0cm}}
+        Code & Function & Input & Output \\
+        \hline
+        \multicolumn{4}{c}{ {\bf Databank layer}}\\
+        \hline
+        \multicolumn{4}{c}{ {\bf Scripts/BuildDatabank/ }}\\
+         AddData.py & Create README.yaml from user given information in info.yaml  & info.yaml & README.yaml \\
+         searchDATABANK.py & Pair simulations with available experimental data  & README.yaml files & Updated README.yaml files \\
+         QualityEvaluation.py & Quality-evaluate simulations that are paired with experimental data  & README.yaml files & [lipid\_name]\_OrderParameters\_quality.json, [lipid\_name]\_FragmentQuality.json, system\_quality.json, FormFactorQuality.json \\
+         \multicolumn{4}{c}{ {\bf Scripts/AnalyzeDatabank/ }}\\
+         calcOrderParameters.py & Calculate C-H bond order parameters for all lipids in the simulation & README.yaml files & [lipid\_name]\_OrderParameters.json \\
+         calcAPL.py & Calculate area per lipid as a function of time & README.yaml files & apl.json \\
+         calc\_FormFactors.py & Calculate x-ray scattering form factors & README.yaml files & FormFactor.json \\
+         calc\_thickness.py & Calculate membrane thickness & README.yaml files & thickness.json \\
+         template.py $\qquad$  template.ipynb & Minimal code to loop over simulations that can be used as a template for new analyses & README.yaml files & \\
+
+    \end{tabular}
+    \caption{List of relevant codes used to build the Databank and perform analyses in the {\it Databank layer} available at~\url{https://github.com/NMRLipids/Databank/}.}
+    \label{tab:codes}
+\end{table}
+
+\begin{table}[!h]
+    \centering
+    %\begin{tabular}{c c c c}
+    \begin{tabular}{p{3.5cm}  p{4.5cm}  p{3.0cm}  p{4.0cm}}
+        Code & Function & Input & Output \\
+        \hline
+        \multicolumn{4}{c}{ {\bf Application layer}}\\
+        \hline
+        %\multicolumn{4}{c}{ {\bf Scripts/BuildDatabank/ }}\\
+        AreaPerLipidAnd ThicknessCorrelations.ipynb  & Analyze correlations between results in the {\it Databank layer}  & Results in the {\it Databank layer} & Figs.~\ref{fig:quality}G and~\ref{fig:QualityCorrelationsSI} \\
+        FlipFlop.py & Calculate flip-flop rates of lipids in all simulations & README.yaml & flipflop.dat files at {\it /Data/Flipflops/} \\
+        plotFlipFlop.ipynb & Plot how flip-flop rates depend on membrane properties & README.yaml and flipflop.dat files at {\it /Data/Flipflops/} & Fig.~\ref{fig:flip-flops} \\
+        calcMD-PERMEATION.py  & Calculate permeation rate of water molecules through bilayers from all simulations & README.yaml & Counting\_events.txt files at {\it /Data/MD-PERMEATION/} \\
+        calcWATERdiffusion.py & Calculate water lateral diffusion along membrane surface & README.yaml & WATERlateralMSD.xvg files at {\it /Data/WATERdiffusion/} \\
+        plotWaterPermeation.ipynb & Plot how water permeation and diffusion depend on membrane properties & README.yaml, Counting\_events.txt\,at {\it /Data/MD-PERMEATION/} and\,WATERlateralMSD.xvg files\,at {\it /Data/WATERdiffusion/} & Figs.~\ref{fig:permeability},~\ref{fig:permeationSI}, and~\ref{fig:diffusionSI}.
+\\
+        plotSimulation.ipynb & Plot results from a simulation together with the experimental data if available & README.yaml files and results in the {\it databank layer} & Figs~\ref{fig:quality}D-F and ~\ref{fig:POPC_POPE_dataSI}
+    \end{tabular}
+    \caption{Examples of codes that analyze membrane properties from the Databank in an {\it Application layer} available at~\url{https://github.com/NMRLipids/DataBankManuscript/}.}
+    \label{tab:codesApplication}
+\end{table}
 
@@
     DOI & DOI from where the raw data is found & user given (compulsory) \\
-    SOFTWARE & Software used to run the simulation (e.g. Gromacs, Amber, NAMD, etc.) & \\
+    SOFTWARE & Software used to run the simulation (Gromacs, Amber, NAMD, etc.) & \\
     TRJ & Name of the trajectory file found from DOI (trr or xtc for Gromacs, dcd for OpenMM) & \\
@@
     \hline
-    PUBLICATION & Give reference to a publication(s) related to the data. & User given (optional)\\
-    AUTHORS\_CONTACT & Name and email of the main author(s) of the data. & \\
+    PUBLICATION & Give reference to a publication(s) related to the data & user given (optional)\\
+    AUTHORS\_CONTACT & Name and email of the main author(s) of the data & \\
     SYSTEM & System description on free text format & \\
@@
     FF & Name of the used force field & \\
-    FF\_SOURCE & Source of the force field parameters, e.g, CHARMM-GUI, webpage, citation to a publication, etc. & \\
-    FF\_DATE &  Date when force field parameters were accessed on the gives source (day/month/year). & \\
-    FF{molename} & Molecule specific force field information, e.g., water model with FFSOL and sodium parameters with FFSOD. & \\
-    CPT & Name of the Gromacs checkpoint file. & \\
-    LOG & Name of the Gromacs log file. & \\
-    GRO & Name of the Gromacs gro file. & \\
-    TOP & Name of top file for Gromacs or psf file for OpenMM. & \\
-    CRD & Name of crd file for OpenMM. & \\
-    WARNINGS & Dictionary containing information about unusual features in the trajectory, such as ambiguous atom names, membrane normal not oriented in z-direction, old Gromacs version used, etc.
-& \\
+    FF\_SOURCE & Source of the force field parameters, e.g, CHARMM-GUI, webpage, citation to a publication & \\
+    FF\_DATE &  Date when force field parameters were accessed on the given source (day/month/year) & \\
+    FF{molename} & Molecule specific force field information, e.g., water model with FFSOL and sodium parameters with FFSOD & \\
+    CPT & Name of the Gromacs checkpoint file & \\
+    LOG & Name of the Gromacs log file & \\
+    GRO & Name of the Gromacs gro file & \\
+    TOP & Name of top file for Gromacs or psf file for OpenMM & \\
+    CRD & Name of crd file for OpenMM & \\
+    WARNINGS & Dictionary containing information about unusual features in the trajectory, such as ambiguous atom names, membrane normal not oriented in $z$-direction, old Gromacs version used & \\
     \hline
-    TRAJECTORY\_SIZE & Size of the trajectory file in bytes & automatically extracted data. \\
-    TRJLENGTH & Lenght of the trajectory (ps). & \\
-    TEMPERATURE & Temperature of the simulation. & \\
-    NUMBER\_OF\_ATOMS & Number of atoms in the simulation. & \\
-    DATEOFRUNNIG & Date when added into the databank & \\
+    TRAJECTORY\_SIZE & Size of the trajectory file in bytes & automatically extracted data \\
+    TRJLENGTH & Lenght of the trajectory (ps) & \\
+    TEMPERATURE & Temperature of the simulation & \\
+    NUMBER\_OF\_ATOMS & Number of atoms in the simulation & \\
+    DATEOFRUNNIG & Date when added into the Databank & \\
     EXPERIMENT & Potentially connected experimental data & \\
-    COMPOSITION & Numbers of lipid molecules in both leaflets and numbers of other molecules are added to the dictionary. & \\
-    ID & Unique ID number to ease the analyses. & \\
+    COMPOSITION & Numbers of lipid molecules in both leaflets and numbers of other molecules are added to the dictionary & \\
+    ID & Unique ID number to ease the analyses & \\
     \end{tabular}
@@
     \midrule
-    DOI & DOI of the publication related to the experimental data. \\
-    TEMPERATURE & Temperature of the experiment. \\
+    DOI & DOI of the publication related to the experimental data \\
+    TEMPERATURE & Temperature of the experiment \\
     MOLAR\_FRACTIONS & Dictionary of molar fractions of bilayer components \\
@@
     % (defined as ??) \\
-    TOTAL\_LIPID\_CONCENTRATION & Total concentration of lipid components. If exact concentration is not known, but experiments are performed in excess water, 'full hydration' can be given. \\
-    COUNTER\_IONS & Type of counter ions if present.
+    TOTAL\_LIPID\_CONCENTRATION & Total concentration of lipid components; if exact concentration is not known, but experiments are performed in excess water, 'full hydration' can be given \\
+    COUNTER\_IONS & Type of counter ions if present
 \end{tabular}
@@
     \end{tabular}
-    \caption{Abbreviations for molecules used in the databank}
+    \caption{Abbreviations for molecule names used in the Databank.}
     \label{tab:abbreviations}
+\end{table}
+
+\begin{table}[p]
+    \centering
+    \begin{tabular}{l}
+    Force field name and references \\
+    \hline
+    CHARMM36~\cite{klauda10}\\
+    Slipids~\cite{jambeck12,jambeck12b,jambeck2012another,ermilova16,grote20}  \\
+    MacRog~\cite{Kulig15b}  \\
+    Amber Lipid14/17~\cite{dickson14,dickson22}  \\
+    Charmm-Drude~\cite{li2017drude}      \\
+    ECClipids~\cite{melcr18,melcr20,bacle21}  \\
+    GROMOS-CKP~\cite{Chandrasekhar03,kukol09,piggot12}  \\
+    Berger~\cite{berger97}  \\
+    ECC-CHARMM36~\cite{nencini2019development}  \\
+    Orange~\cite{catte16}  \\
+    Poger~\cite{poger10}  \\
+    GROMOS 43A1-S3~\cite{chiu09}  \\ 
+    GAFFlipid~\cite{dickson12}  \\
+    OPLS3e~\cite{roos19}  \\
+    Ulmschneider~\cite{Ulmschneider09}  \\
+    Chiu Gromos~\cite{chiu09}  \\
+    AMOEBA~\cite{chu2018polarizable} \\
+    \end{tabular}
+    \caption{List of current force fields used in simulations in the Databank, with references.}
+    \label{tab:ForceFields}
 \end{table}
@@
 \section{NMR experiments}
-Acyl chain order parameters of POPE and POPG were analyzed from the experiments that were previously used to determine headgroup order parameters~\cite{bacle21}.
-The analysis of the crowded spectral region at 29-31 ppm was based on the previous assignment reported for POPC membranes~\cite{ferreira13}.
+Acyl chain order parameters of POPE (Figs.~\ref{POPEexp1} and~\ref{POPEexp2}) and POPG (Figs.~\ref{POPGexp1} and~\ref{POPGexp2}) were analyzed from the same data that were previously recorded to determine headgroup order parameters~\cite{bacle21}.
+The analysis of the crowded spectral region at 29--31 ppm was based on the previous assignment reported for POPC membranes~\cite{ferreira13}.
+To measure the order parameters for DOPC (Fig.~\ref{fig:DOPCexp}), the sample was prepared and experiments performed similarly to previous studies~\cite{bacle21}.
 
@@
 The assignment was based on a previous assignment reported for POPC membranes~\cite{ferreira13}.
-(C) C-H order parameter profile for the acyl chains of POPE.
+(C) C--H bond order parameter profile for the acyl chains of POPE.
 The splittings used for calculating the order parameters are shown in Fig.~\ref{POPEexp2}.
@@
   \caption{Dipolar spectra obtained from the 2D R-PDLF spectrum from POPE in Fig.~\ref{POPEexp1}.
-The number at the top left of each plot denotes the corresponding chemical shift.
-The carbon label for each splitting is displayed on the top right.
+The number at the top left corner of each panel denotes the corresponding chemical shift.
+The carbon label for each splitting is displayed on the top right corner.
 The labels are the same as in Fig.~\ref{POPEexp1}.
@@
 The assignment was based on a previous assignment reported for POPC membranes~\cite{ferreira13}.
-(C) C-H order parameter profile for the acyl chains of POPG.
+(C) C--H bond order parameter profile for the acyl chains of POPG.
 The splittings used for calculating the order parameters are shown in Fig.~\ref{POPGexp2}.
@@
   \caption{Dipolar spectra obtained from the 2D R-PDLF spectrum described in Fig.~\ref{POPGexp1}.
-The number at the top left of each plot denotes the corresponding chemical shift.
-The carbon label for each splitting is displayed on the top right.
+The number at the top left corner of each panel denotes the corresponding chemical shift.
+The carbon label for each splitting is displayed on the top right corner.
 The labels are the same as in Fig.~\ref{POPGexp1}.
@@
   \label{POPGexp2}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{Figures/DOPCfig.pdf}
+    \caption{Determination of DOPC order parameters from a R-PDLF spectrum measured at a magic angle spinning frequency of 5.15 kHz.
+(A) $^{13}$C rINEPT spectrum with peak assignment.
+The labels used are shown in the chemical structure of DOPC.
+The chemical shift of the methyl groups was defined as 13.8 ppm.
+(B) Contour plot of the R-PDLF spectrum for the crowded spectral region.
+The assignment was based on a previous assignment reported for POPC membranes~\cite{ferreira13}.
+(C) C--H bond order parameters of DOPC.}
+    \label{fig:DOPCexp}
+\end{figure}
+
+\begin{figure}[p]
+  \includegraphics[width=\textwidth]{Figures/DOPC_slices.pdf}
+  \caption{Dipolar spectra obtained from the 2D R-PDLF spectrum described in Fig.~\ref{fig:DOPCexp}.
+The number at the top left corner of each panel denotes the corresponding chemical shift.
+The carbon label for each splitting is displayed on the top right corner.
+The labels are the same as in Fig.~\ref{fig:DOPCexp}.
+}
+  \label{DOPCexp2}
 \end{figure}

### 2023-04-18 main.tex
Updates from Overleaf

@@
 Histogrammed values are shown with black dots.
-Error bars show the standard error of the mean within each bin.
+For the mean value in each bin, average weighted with the simulation lengths was used, and error bars show the standard error of the mean.
     }
@@
 Inset in B) shows the Arrhenius plot of permeation ($\ln(P)$ vs.
-$1/T$) that gives 17\,$\pm$\,4\,$k_\mathrm{B}T$ for the average activation energy for water permeation through lipid bilayer.
-Insets in C) and D) show the region where the dependence could be considered approximately linear.
+$1/T$) that gives 14\,$\pm$\,3\,$k_\mathrm{B}T$ for the average activation energy for water permeation through lipid bilayer.
     \textbf{F} Lateral diffusion of water as a function of hydration level. Experimental points for DMPC bilayers at 313\,K at different hydration levels are shown~\cite{rudakova04}.
@@
 Histogrammed values are shown with black dots.
-Error bars show the standard error of the mean within each bin.
+For the mean value in each bin, average weighted with the simulation lengths was used, and error bars show the standard error of the mean.
 Only bins with more than one microsecond of data in total were used for water permeation.}
@@
 Using the workflow depicted in Fig.~\ref{fig:overlay}E, we then plotted the observed permeabilities and their histogrammed values in Figs.~\ref{fig:permeability}B--E as a function of temperature, membrane thickness, area per lipid, and acyl chain order.
-As expected, the permeability increases with the temperature, giving an average energy barrier of 17\,$\pm$\,4\,$k_\mathrm{B}T$ for the water permeation from the Arrhenius plot in Fig.~\ref{fig:permeability}B.
+As expected, the permeability increases with the temperature, giving an average energy barrier of 14\,$\pm$\,3\,$k_\mathrm{B}T$ for the water permeation from the Arrhenius plot in Fig.~\ref{fig:permeability}B.
 On the other hand, the water permeability on average decreases when membranes become more packed, that is, with decreasing area per lipid and increasing thickness and acyl chain order (Figs.~\ref{fig:permeability}C--E).
@@
 %is previously reported from simulations and experiments for some systems, but not for all~\cite{mathai07,venable19,frallicciardi22}. 
-Our analysis over the NMRlipids Databank, containing significantly more data than what was available in previous studies, suggest non-linear dependencies on both of these parameters, yet the linear correlation would be a good approximation for thicknesses $\gtrsim$\,3.9\,nm and area per lipids $\lesssim$\,69\,\AA{}$^2$ (insets in Figs.~\ref{fig:permeability}C--D).
+Our analysis over the NMRlipids Databank, containing significantly more data than what was available in previous studies, suggest non-linear dependencies on both of these parameters.
+%, yet the linear correlation would be a good approximation for thicknesses $\gtrsim$\,3.9\,nm and area per lipids $\lesssim$\,69\,\AA{}$^2$ (insets in Figs.~\ref{fig:permeability}C--D).
 %within the range of experimentally reported values  permeability which is 1-2 orders of magnitude slower than osmotic permeability \cite{jansen95,??}. 

### 2023-04-18 SI.tex

@@
 Values from simulations with non-zero permeation values are shown with blue dots.
-Histogrammed values are shown with black dots.
+Histogrammed values are shown with black dots For the mean value in each bin, average weighted with the simulation lengths was used, and error bars show the standard error of the mean.
+Only bins with more than one microsecond of data were used.
+Only simulations with the temperatures between 300-315\,K were used.
     }

### 2023-04-18 SI.tex
Update text

@@
     \includegraphics[width = \textwidth]{Figures/LateralDiffusionSI.pdf}
-    \caption{Lateral diffusion of water as a function of (A) area per lipid, (B) temperature, (C) membrane thickness, and (D) fraction of charged lipids in a membrane.}
+    \caption{Lateral diffusion of water as a function of (A) area per lipid, (B) temperature, (C) membrane thickness, and (D) fraction of charged lipids in a membrane.
+Non-zero permeation and diffusion values from simulations are shown with blue dots.
+Histogrammed values are shown with black dots.
+For the mean value in each bin, average weighted with the simulation lengths was used, and error bars show the standard error of the mean.
+Only bins with more than one microsecond of data in total were used for water permeation.
+Only simulations with the temperatures between 300-315\,K were used in D.}
     \label{fig:diffusionSI}

### 2023-04-18 main.tex
Update text

@@
 Membrane lipid composition is implicated in diseases, such as cancer and neurodegenerative disorders, and therapeutics that affect membrane compositions are emerging~\cite{torres21}.
-However, biomembranes are often difficult to study experimentally, because they are complex mixtures of proteins and lipids in disordered fluid state with complicated phase behaviour at biological conditions.
+However, biomembranes are often difficult to study experimentally, because they are complex mixtures of proteins and lipids in disordered fluid state with complicated phase behaviour at physiologicalbiological conditions.
 For those reasons, the detailed connections between complex lipid interactions and biological functions taking place in or around membranes remain poorly understood.
@@
 Recent advances using machine learning approaches utilizing publicly available databanks, for example to solve protein structures~\cite{jumper21}, emphasize the increasing importance of such resources.
-%The profound, and sometimes unexpected, effect that such open data sharing and quality evaluation can have on the field cannot be understated.
-At the very least, open access data will reduce the man and computational hours originating from double work, reducing the environmental impact.
-Beyond that, by providing the training data, the PDB laid the foundation for the machine learning algorithms now closing into solving the protein folding problem and the strides those machine learning tools have facilitated in the biomedical sciences since.
 
@@
 
-NMRlipids Databank is a community-driven database containing atomistic MD simulations of biologically relevant lipid membranes emerging from the NMRlipids open collaboration~\cite{botan15,ollila16,catte16,antila19,bacle21}.
+NMRlipids Databank is a community-driven catalogue containing atomistic MD simulations of biologically relevant lipid membranes emerging from the NMRlipids open collaboration~\cite{botan15,ollila16,catte16,antila19,bacle21}.
 It has been designed to improve the Findability, Accessibility, Interoperability, and Reuse (FAIR)~\cite{wilkinson16} of MD simulation data, most importantly the output trajectories.
@@
 Single-component lipid membranes and binary mixtures, depicted in Fig.~\ref{fig:overlay}F, are currently most abundant in the NMRlipids Databank, yet mixtures with up to five lipid types are available.
-The distribution of lipids among the available simulations, shown in Fig.~\ref{fig:overlay}B, roughly resembles the biological relative abundance of different lipid types, with phosphatidylcholine (PC) being the most common followed by cholesterol, phosphatidylethanolamine (PE), phosphatidylserine (PS), phosphatidylglycerol (PG), phosphatidylinositol (PI), and other lipids from different organisms and organelles~\cite{vanmeer08}.
+The distribution of lipids among the available simulations, shown in Fig.~\ref{fig:overlay}B, roughly resembles the biological relative abundance of different lipid types, with phosphatidylcholine (PC) being the most common followed by cholesterol, phosphatidylethanolamine (PE), phosphatidylserine (PS), phosphatidylglycerol (PG), phosphatidylinositol (PI), and other lipids, depending on organism and organelle~\cite{vanmeer08}.
 Abbreviations and full names of all lipids present in the Databank are listed in Table~\ref{tab:abbreviations}.
 Force fields used in simulations cover all the essential parameter sets commonly used in lipid simulations, see Fig.~\ref{fig:overlay}C and Table~\ref{tab:ForceFields}, including also united atom and polarizable force fields.
-Therefore, the averages calculated over the Databank can be considered as mean predictions from available lipid models for an average cell membrane.
+Therefore, the averages calculated over the Databank can be considered as mean predictions from available lipid models (average over force field parameters) for an average cell membrane (average over lipid compositions).
 
 The overlay structure of the NMRlipids Databank, illustrated in Fig.~\ref{fig:overlay}A, is designed to enable efficient upcycling of MD simulations for data-driven and machine learning applications with minimal investment on new infrastructure.
-Raw simulation data in the {\it Data layer} can be stored in any publicly available location with long term stability---such as Zenodo (\href{https://zenodo.org}{zenodo.org}), figshare (\href{https://figshare.com}{figshare.com}), or NIRD research data archive (\href{https://archive.norstore.no}{archive.norstore.no})---and with permanent links to the data, such as digital object identifiers (DOIs).
+Raw simulation data in the {\it Data layer} can be stored in any publicly available location with long term stability
+%figshare (\href{https://figshare.com}{figshare.com}), or NIRD research data archive (\href{https://archive.norstore.no}{archive.norstore.no})---
+and with permanent links to the data, such as digital object identifiers (DOIs), such as Zenodo (\href{https://zenodo.org}{zenodo.org}). 
 %or uniform resource names (URNs). 
@@
 A more detailed structure of the \textit{Databank layer} is shown in Fig.~\ref{DatabankStructure} in the SI.
-    %b) Distribution of the lengths of the trajectories, total number of trajectories and total length of the simulations in the NMRlipids databank.
     B) Distribution of lipids present in the trajectories of the Databank. `Others' lists lipids occurring in five or fewer simulations. 
@@
 
-\subsection{NMRlipids Databank-GUI: graphical access to the MD simulation data}\label{section:access}
+\subsection{NMRlipids Databank-GUI: graphical access to the MD simulation data}
 NMRlipids Databank-GUI, available at~\href{https://www.databank.nmrlipids.fi}{databank.nmrlipids.fi}, provides easy access to the NMRlipids Databank content through a graphical user interface (GUI).
@@
 
-\subsection{NMRlipids Databank-API: programmatic access to the MD simulation data}
+\subsection{NMRlipids Databank-API: programmatic access to the MD simulation data}\label{section:access}
 The NMRlipids Databank-API, available at \href{https://github.com/NMRlipids/Databank}{github.com/NMRlipids/Databank}, provides programmatic access to all simulation data in the NMRlipids Databank.
@@
 
-While analysis codes and results for basic membrane properties 
-%area per lipid, C–H bond order parameters, X-ray scattering form factors and thickness 
-are included in the {\it Databank layer}, unlimited further analyses can be implemented by anyone in separate repositories in the {\it Application layer}.
+While analysis codes and results for basic membrane properties are included in the {\it Databank layer}, unlimited further analyses can be implemented by anyone in separate repositories in the {\it Application layer}.
 When {\it Application layer} repositories are organized by mimicking the {\it Databank layer} structure, they can be accessed programmatically and further analyzed using the tools in the NMRlipids Databank-API by implementing the flowchart demonstrated in Fig.~\ref{fig:overlay}E.
@@
 The related codes are listed in Table~\ref{tab:codesApplication}.
-
-%\begin{figure}[t]
-%    \centering
-%    \includegraphics[width=88mm]{Figures/flowchart2.pdf}
-%    \caption{ A) Flowchart for performing an analysis of properties through all MD simulations in the NMRlipids databank using the API.
-%    B) Flowchart for accessing results calculated from the NMRlipids databank and stored to the application layer.
-%    C) Area per lipids of POPC and POPE lipid bilayers predicted by different force fields at 310\,K in simulations that are available in the NMRlipids databank.
-The points from the best performing simulations in the NMRlipids databank quality evaluation for each lipid (Figs.~\ref{fig:quality} and~\ref{fig:POPC_POPE_dataSI}) are surrounded by black circles.
-%    }
-%    \label{fig:POPCPOPEapls}
-%\end{figure}
 
@@
 To streamline this process, we have defined quantitative quality measures that enable automatic ranking of lipid bilayer simulations based on their quality against experiments.
-Conformational ensembles of individual lipid molecules are evaluated in the NMRlipids Databank by calculating the probabilities for each C--H bond order parameter to locate within experimental error ($P^{\mathrm{hg}}$, $P^{\mathrm{sn1}}$,$P^{\mathrm{sn2}}$, and $P^{\mathrm{total}}$).
+Conformational ensembles of individual lipid molecules are evaluated in the NMRlipids Databank by first calculating the probabilities for each C--H bond order parameter to locate within experimental error and then averaging the possibilities over different lipid segments ($P^{\mathrm{hg}}$, $P^{\mathrm{sn1}}$,$P^{\mathrm{sn2}}$, and $P^{\mathrm{total}}$).
 Furthermore, the quality against X-ray scattering experiments ($FF_q$) is estimated as the difference in the experimental and simulated locations of the first form factor minimum.
@@
 In addition, the ergodicity of conformational sampling of lipids is estimated by calculating $\tau_\mathrm{rel}$, the convergence time of the slowest principal component divided by the simulation length.
-Let us demonstrate how the automatic simulation-quality evaluation and the NMRlipids Databank-API enable rapid selection of the best models for simulations of membranes with the two most biologically-abundant neutral membrane lipids~\cite{vanmeer08}: POPC and POPE.
+Here we demonstrate how the automatic simulation-quality evaluation and the NMRlipids Databank-API enable rapid selection of the best models for simulations of membranes with the two most biologically-abundant neutral membrane lipids~\cite{vanmeer08}: POPC and POPE.
 
@@
 From these results, we then pick the force fields occurring in Fig.~\ref{fig:quality}A (that is: force fields for which both POPC and POPE simulations exist in the Databank) and rank them according to the quality of the \textit{sn}-1 chain of POPC (Fig.~\ref{fig:quality}B) and of POPE (Fig.~\ref{fig:quality}C).
-Simulations with $\tau_\mathrm{rel}$ clearly above one are discarded.
+Simulations with $\tau_\mathrm{rel}$ clearly above one (larger than 1.3) are discarded.
 Because the membrane packing correlates with the average order parameter of the \textit{sn}-1 chain (Fig.~\ref{fig:quality}G), rankings in Figs.~\ref{fig:quality}B and C can be used to select the simulations giving the most realistic results in Fig.~\ref{fig:quality}A.
@@
 \end{figure}
-
-%In summary, the automatic quality evaluation of the NMRlipids databank enables the rapid exploration and selection of the best simulation parameters for specific applications without extensive and tedious manual force field evaluation.
-This possibility will promote more reliable simulation results for wide range of applications, support force field development, and streamline the parametrization of coarse grained force fields against the best atomistic MD simulations.
-The automatically extracted quantitative quality measures will be particularly useful to guide automatic parametrization procedures.
-
-%Understanding such complex picture of lipid bilayer MD simulation quality would not be possible without automatic quality evaluation of large sets of simulations enabled by the NMRlipids databank.
 

### 2023-04-19 SI.tex
Updates from Overleaf

@@
     \centering
-    \includegraphics[height = 0.9\textheight]{Figures/totalrankingCOMPL.pdf}
-    \caption{Top 50 simulations in the NMRlipids Databank ranked based on the C--H bond order parameter quality against experiments.}
+    \includegraphics[height = 0.85\textheight]{Figures/totalrankingCOMPL.pdf}
+    \caption{Top 50 simulations in the NMRlipids Databank ranked based on the C--H bond order parameter quality against experiments.
+The columns 2-4 show qualities for acyl chain order parameters ($P^\mathrm{tails}$), headgroup order parameters ($P^\mathrm{hg}$), all order parameters ($P^\mathrm{total}$), and for X-ray scattering form factors ($FF_q$).
+Column 5 shows relative equilibration times for conformations ($\tau_\mathrm{rel}$).
+Note that the best possible order parameter quality is one, while the best possible form factor quality is zero.
+ID values in the last column can be used to identify each simulation in the databank.}
     \label{fig:top50simulations}

### 2023-04-19 main.tex

@@
 NMRlipids Databank is a community-driven catalogue containing atomistic MD simulations of biologically relevant lipid membranes emerging from the NMRlipids open collaboration~\cite{botan15,ollila16,catte16,antila19,bacle21}.
-It has been designed to improve the Findability, Accessibility, Interoperability, and Reuse (FAIR)~\cite{wilkinson16} of MD simulation data, most importantly the output trajectories.
+It has been designed to improve the Findability, Accessibility, Interoperability, and Reuse (FAIR)~\cite{wilkinson16} of MD simulation data, most importantly the output trajectories and necessary information to their reuse.
 The NMRlipids Databank is constructed using the NMRlipids project protocol, in which all the content is openly accessible throughout the project~\cite{botan15}.
@@
     \includegraphics[width=\linewidth]{Figures/overlay3.pdf}
-    \caption{A) Schematic presentation of the overlay structure used in the NMRlipids Databank.
+    \caption{Overview of the NMRlipids Databank. 
+    \textbf{A} Schematic presentation of the overlay structure used in the NMRlipids Databank.
 A more detailed structure of the \textit{Databank layer} is shown in Fig.~\ref{DatabankStructure} in the SI.
-    B) Distribution of lipids present in the trajectories of the Databank. `Others' lists lipids occurring in five or fewer simulations. 
-    C) Distribution of force fields in the simulations in the Databank. References for each force field are given in Table~\ref{tab:ForceFields}.
-    D) Flowchart for performing an analysis of properties through all MD simulations in the NMRlipids Databank using the API.
-    E) Flowchart for accessing results calculated from the NMRlipids Databank and stored to the \textit{Application layer}.
-    F) Currently available single lipid component and binary mixtures in the NMRlipids Databank. Colorbar shows the number of available simulations with the darkest green indicating three or more.  
+    \textbf{B} Distribution of lipids present in the trajectories of the Databank. `Others' lists lipids occurring in five or fewer simulations. 
+    \textbf{C} Distribution of force fields in the simulations in the Databank. References for each force field are given in Table~\ref{tab:ForceFields}.
+    \textbf{D} Flowchart for performing an analysis of properties through all MD simulations in the NMRlipids Databank using the API.
+    \textbf{E} Flowchart for accessing results calculated from the NMRlipids Databank and stored to the \textit{Application layer}.
+    \textbf{F} Currently available single lipid component and binary mixtures in the NMRlipids Databank.
+Colorbar shows the number of available simulations with the darkest green indicating three or more.
     }
@@
 NMRlipids Databank-GUI, available at~\href{https://www.databank.nmrlipids.fi}{databank.nmrlipids.fi}, provides easy access to the NMRlipids Databank content through a graphical user interface (GUI).
-Simulations can be searched based on their molecular composition, force field, temperature, membrane properties, and quality; the search results are ranked based on the simulation quality as evaluated against experimental data.
+Simulations can be searched based on their molecular composition, force field, temperature, membrane properties, and quality; the search results are ranked based on the simulation quality as evaluated against experimental data when available.
 Membranes can be visualized, and properties between different simulations and experiments compared.
@@
 The flowchart in Fig.~\ref{fig:overlay}D illustrates the practical implementation of such an analysis.
-First, raw data of each simulation can be accessed by iterating through the README.yaml files in the {\it Databank layer} that contain links pointing to the locations of simulation trajectories and other relevant files.
+After cloning the databank repository to a local computer, raw data of each simulation can be accessed by iterating through the README.yaml files in the {\it Databank layer} that contain links pointing to the locations of simulation trajectories and other relevant files.
 Simulations can then be automatically analyzed with the help of the README.yaml and the mapping files that associate the specific naming conventions in each simulation with the universal molecule and atom names used by the Databank.
 Finally, the analysis results are stored using the same structure as in the {\it Databank layer}.
-Examples of codes that utilize the NMRlipids Databank-API, as well as a template for a new analysis, can be found from locations listed in Table~\ref{tab:codes}.
+Practical examples of codes that utilize the NMRlipids Databank-API, as well as a template for a new analysis, can be found from locations listed in Table~\ref{tab:codes}.
 
@@
     \includegraphics[width=\linewidth]{Figures/quality5.pdf}
-    \caption{ A) Area per lipid of POPC and POPE lipid bilayers predicted by different force fields at 310\,K in simulations that are available in the NMRlipids Databank.
+    \caption{Examples of data obtained using NMRlipids Databank. 
+    \textbf{A} Area per lipid of POPC and POPE lipid bilayers predicted by different force fields at 310\,K in simulations that are available in the NMRlipids Databank.
 The data points from the best-performing simulations, based on rankings in panels B and C, are surrounded by black circles.
-    B) Best POPC simulations ranked based on the \textit{sn}-1 acyl chain order parameter quality ($P^\mathrm{sn1}$).
+    \textbf{B} Best POPC simulations ranked based on the \textit{sn}-1 acyl chain order parameter quality ($P^\mathrm{sn1}$).
 Also \textit{sn}-2 acyl chain ($P^\mathrm{sn2}$), headgroup ($P^\mathrm{hg}$) and total ($P^\mathrm{total}$) order parameter qualities, form factor quality ($FF_q$), and relative equilibration time for conformations ($\tau_\mathrm{rel}$) are shown.
-    C) Best POPE simulations ranked based on the \textit{sn}-1 acyl chain order parameter quality. 
-    D)--F) Direct comparison against experimental (NMR order parameters and X-ray scattering ) data exemplified for a simulation with the best overall order parameter quality (D), the best quality for POPE lipid (E), and the headgroup quality for POPE (F).
-    G) Scatter plots and Pearson correlation coefficients, $r$, for the membrane area per lipid, thickness, first minimum of X-ray scattering form factor and average order parameter of the \textit{sn}-1 acyl chain extracted from the NMRlipids Databank.
+Note that the best possible order parameter quality is one, while the best possible form factor quality is zero.
+    \textbf{C} Best POPE simulations ranked based on the \textit{sn}-1 acyl chain order parameter quality. 
+    \textbf{D--F} Direct comparison against experimental (NMR order parameters and X-ray scattering ) data exemplified for a simulation with the best overall order parameter quality (D), the best quality for POPE lipid (E), and the headgroup quality for POPE (F).
+    \textbf{G} Scatter plots and Pearson correlation coefficients, $r$, for the membrane area per lipid, thickness, first minimum of X-ray scattering form factor and average order parameter of the \textit{sn}-1 acyl chain extracted from the NMRlipids Databank.
 All correlation coefficients have p-value below 0.001.
@@
 The atomistic studies report an increase in cholesterol flip-flop rates with increasing acyl chain unsaturation level and decreasing cholesterol concentration~\cite{gu19,javanainen19}, but the amount of data in these individual studies was not sufficient to systematically assess correlations between cholesterol flip-flop rates and membrane properties.
-Here, we demonstrate that the NMRlipids Databank-API makes analyses of such rare phenomena accessible for all by enabling access to a large amount of MD simulation data as described in section~\ref{section:access}.
+Here, we demonstrate that the NMRlipids Databank-API makes analyses of such rare phenomena accessible for all by enabling access to a large amount of MD simulation data as illustrated in Fig.~\ref{fig:overlay}.
 This is particularly useful for scientists in various fields of science and industry who lack access to the computational resources or the expertise to produce the large amounts of MD simulation data required for such analyses.
@@
 Using the general workflow depicted in Fig.~\ref{fig:overlay}D, we first calculated the flip-flop rates from all the simulations available in the NMRlipids Databank.
-Flip-flops were observed for cholesterol, % (634 events in 77 different simulations),
-DCHOL (18,19-di-nor-cholesterol), %, 16 flip-flops in 1 simulation) 
-DOG (1,2-dioleoyl-\textit{sn}-glycerol), %, 44 flip-flops in 4 simulations) 
-and SDG (1-stearoyl-2-docosahexaenoyl-\textit{sn}-glycerol).
+Flip-flops were observed for cholesterol, DCHOL (18,19-di-nor-cholesterol), DOG (1,2-dioleoyl-\textit{sn}-glycerol), and SDG (1-stearoyl-2-docosahexaenoyl-\textit{sn}-glycerol).
 The observed cholesterol flip-flop rates, ranging between 0.001--1.6\,\textmu{}s$^{-1}$ with the mean of 0.16\,\textmu{}s$^{-1}$ and median of 0.07\,\textmu{}s$^{-1}$, are in line with the previously reported values from atomistic MD simulations~\cite{gu19,javanainen19,baral20}.
-The flip-flop rate of DCHOL, 0.2\,\textmu{}s$^{-1}$, was close to the average value of cholesterol, while
-the average rates for diacylglycerols DOG (0.4\,\textmu{}s$^{-1}$) and SDG (0.5\,\textmu{}s$^{-1}$) were higher than for cholesterol.
-Flip-flops were not observed for other lipids, giving the upper limits for
-PC-lipid flip-flop rate as 9$\times$10$^{-6}$\,\textmu{}s$^{-1}$ and for
-ceramide (\textit{N}-palmitoyl-\textsc{D}-\textit{erythro}-sphingosine) as 0.002\,\textmu{}s$^{-1}$.
+The flip-flop rate of DCHOL, 0.2\,\textmu{}s$^{-1}$, was close to the average value of cholesterol, while the average rates for diacylglycerols DOG (0.4\,\textmu{}s$^{-1}$) and SDG (0.5\,\textmu{}s$^{-1}$) were higher than for cholesterol.
+Flip-flops were not observed for other lipids, giving the upper limits for PC-lipid flip-flop rate as 9$\times$10$^{-6}$\,\textmu{}s$^{-1}$ and for ceramide (\textit{N}-palmitoyl-\textsc{D}-\textit{erythro}-sphingosine) as 0.002\,\textmu{}s$^{-1}$.
 Thus, the available data in the NMRlipids Databank suggest that the lipid flip-flop rate decreases in the order: diacylglycerols $>$ cholesterol $>$ other lipids including ceramides.
@@
     \includegraphics[width=88mm]{Figures/CholFlipFlops.pdf}
-    \caption{\textbf{A} Illustration of cholesterol flip-flop.  
+    \caption{Quantification of cholesterol flip-flop events in NMRlipids Databank simulations. 
+    \textbf{A} Illustration of cholesterol flip-flop.  
       \textbf{B--D} Cholesterol flip-flops analyzed from the Databank as a function of membrane thickness, area per lipid, and acyl chain order.
@@
     \includegraphics[width=\linewidth]{Figures/permeation3.pdf}
-    \caption{\textbf{A} Water diffusion, $D_\perp$, and permeability, $P$, through membranes, and lateral diffusion along the membrane, $D_\parallel$, illustrated in a multilamellar stack of lipid bilayers.
+    \caption{Quantification of water diffusion in NMRlipids Databank simulations.
+    \textbf{A} Water diffusion, $D_\perp$, and permeability, $P$, through membranes, and lateral diffusion along the membrane, $D_\parallel$, illustrated in a multilamellar stack of lipid bilayers. 
     \textbf{B--E} Water permeation through membranes analyzed from the Databank as a function of temperature, thickness, area per lipid, and acyl chain order.
@@
 The resulting non-zero values range between 0.3--322\,\textmu{}m/s with the mean of 14\,\textmu{}m/s and median of 8\,\textmu{}m/s.
-These values agree with the previously reported simulation results~\cite{venable19,camilo2022}, but are on average slightly larger than experimental values reported for PC lipids in the liquid crystalline phase,~0.19--0.33\,\textmu{}m/s~\cite{jansen95}.
+These values agree with the previously reported simulation results~\cite{venable19,camilo2022}, but are on average larger than experimental values reported for PC lipids in the liquid crystalline phase,~0.19--0.33\,\textmu{}m/s~\cite{jansen95}.
 Using the workflow depicted in Fig.~\ref{fig:overlay}E, we then plotted the observed permeabilities and their histogrammed values in Figs.~\ref{fig:permeability}B--E as a function of temperature, membrane thickness, area per lipid, and acyl chain order.
@@
 
-To analyze how water diffusion anisotropy depends on membrane properties in a multi-lamellar lipid bilayer system, we calculated the water diffusion parallel to the membrane surface from all simulations in the NMRlipids Databank using the general workflow depicted in Fig.~\ref{fig:overlay}D.
-The parallel diffusion coefficient of water, $D_\parallel$, decreases with reduced hydration and increases with the temperature, but dependencies on the membrane area per lipid, thickness, or fraction of charged lipids were not observed in Figs.~\ref{fig:permeability} and~\ref{fig:diffusionSI} when the workflow depicted in Fig.~\ref{fig:overlay}E was used.
-Simulation results are close to the experimental values with low hydration levels in Fig.~\ref{fig:permeability}F, but increase to approximately two times higher than the experimental value for bulk water diffusion value (3.1$\times$\,10$^{-9}$\,m$^2$/s at 313\,K~\cite{khakimov08}) with high hydration levels.
+To examine how water diffusion anisotropy depends on membrane properties in a multi-lamellar lipid bilayer system, we analyzed the water diffusion parallel to the membrane surface from all simulations in the NMRlipids Databank using the general workflows depicted in Figs.~\ref{fig:overlay}D and E.
+The parallel diffusion coefficient of water, $D_\parallel$, decreases with reduced hydration and increases with the temperature, but dependencies on the membrane area per lipid, thickness, or fraction of charged lipids were not observed in Figs.~\ref{fig:permeability} and~\ref{fig:diffusionSI}.
+Simulation results are close to the experimental values with low hydration levels in Fig.~\ref{fig:permeability}F, but increase to approximately 50\% higher than the experimental value for bulk water diffusion value (3.1$\times$\,10$^{-9}$\,m$^2$/s at 313\,K~\cite{khakimov08}) with high hydration levels.
 This is not surprising as the most common water model used in membrane simulations, TIP3P, overestimates the bulk water diffusion~\cite{pathirannahalage21}.
@@
 Such analyses also pave the way for applications of MD simulations in new fields, as demonstrated here by analysing an essential parameter in pharmacokinetic modeling and MRI imaging:\cite{nitsche19,topgaard20} the anisotropic diffusion of water in membrane systems (Fig.~\ref{fig:permeability}).
-Furthermore, the NMRlipids Databank-API delivers access to a training set data that paves the way for diverse machine learning applications to
-%For example, a machine learning model predicting electron density profiles from form factors would be highly useful in interpretation of scattering experiments.
-%In addition, the data contained in the NMRlipids databank will enable training of more elaborated models to
-predict membrane properties.
-Such applications are analogous to AlphaFold \cite{jumper21} and other tools that predict protein structures from their sequence using artificial intelligence.
+Furthermore, the NMRlipids Databank-API delivers access to a training set data that paves the way for diverse machine learning applications to predict membrane properties.
+Such applications could be analogous to AlphaFold \cite{jumper21} and other tools that predict protein structures from their sequence using artificial intelligence.
 These possibilities are particularly valuable for scientists who do not typically have access to large scale MD simulation data.
@@
 Results from analyses of basic membrane properties (area per lipid, thickness, C--H bond order parameters, X-ray scattering form factors, and relaxation of principal components) are stored in the same folders as the README.yaml files.
-Experimental data used for ranking is stored in the \texttt{/Data/experiments} folder,
-the ranking results in \texttt{/Data/Ranking},
-and the relevant scripts in the \texttt{/Scripts/} folder in the {\it Databank layer} repository. 
-%Raw MD simulation data are stored in external publicly available and stable repositories (\textit{Data layer} in Fig.~\ref{fig:overlay}), such as Zenodo (\url{www.zenodo.org}), from where it can be downloaded whenever needed using the links in README.yaml files.
+Experimental data used for ranking is stored in the \texttt{/Data/experiments} folder, the ranking results in \texttt{/Data/Ranking}, and the relevant scripts in the \texttt{/Scripts/} folder in the {\it Databank layer} repository.
 The scripts in the NMRlipids Databank are mainly written in Python and many of them use the MDAnalysis module~\cite{gowers2019mdanalysis,michaud2011mdanalysis}.
@@
 The C--H bond order parameters were calculated directly from the carbon and hydrogen positions using the definition
-%The order parameters are defined as
 \begin{equation}\label{OP}
@@
 The error of $\Delta S_\mathrm{exp} = 0.02$ is currently assumed for all experimental order parameters~\cite{ollila16}, yet more accurate ones may be available in the future~\cite{wurl22}.
-Because a lipid bilayer simulation contains at least dozens of lipids, the Student's t-distribution could be safely approximated with a normal distribution. However,
-with the quality of currently available force fields, the simulation values can be so far from experiments that a
-normal distribution leads to probability values below the numerical accuracy of computers.
+Because a lipid bilayer simulation contains at least dozens of lipids, the Student's t-distribution could be safely approximated with a normal distribution.
+However, with the quality of currently available force fields, the simulation values can be so far from experiments that a normal distribution leads to probability values below the numerical accuracy of computers.
 To avoid such numerical instabilities, we opted to use the first order Student's t-distribution that has slightly higher probabilities for values far away from the mean.
@@
 where $\langle P[\mathrm{lipid}]\rangle_\mathrm{frag}$ is the average of the individual $S_\mathrm{CH}$ qualities within the fragment, and $F_\mathrm{frag}[\mathrm{lipid}]$ is the percentage of order parameters for which the quality is available within the fragment.
-%, and frag can be \textit{sn}-1, \textit{sn}-2, headgroup or total (all order parameters within a molecule). 
 The overall quality of different fragments in a simulation (frag = 'tails', 'headgroup', or 'total') are then defined as a molar-fraction-weighted average over different lipid components
@@
 \subsection{Quality evaluation of X-ray scattering form factors}
-%While C-H bond order parameters relate to the conformational ensembles of individual lipids, the x-ray scattering form factors depend on membrane dimensions and density distribution~\cite{ollila16}.
 Because experiments give form factors only on a relative intensity scale, they should be scaled before comparing with the simulation data.
@@
 
-%The README.yaml files contain all the essential information to perform arbitrary analyses of simulations in the NMRlipids databank, i.e., the permanent location of the original data and naming convention for all atoms and molecules in each system.
-%In practise, the analyse codes contains a loop over all README.yaml files (i.e., simulations in the NMRlipids databank) which first downloads the raw simulation to a local computer and then uses the information about the atom and molecule naming conventions in README.yaml and mapping files to perform the desired analyses.
-
-%A minimal example of a analysis code is available at \url{https://github.com/NMRLipids/Databank/blob/main/Scripts/AnalyzeDatabank/template.ipynb}.
-
 Water permeability through membranes was calculated from equation $P=r/2c_w$, where $r$ is the rate of permeation events per time and area, and $c_w=$\,33.3679\,nm$^{-3}$ is the concentration of water in bulk~\cite{venable19}.
 The number of permeation events in each trajectory was calculated using the code by Camilo et al.~\cite{camilo2022}, available at~\url{https://github.com/crobertocamilo/MD-permeation}.
-%This code was used within the loop that goes over all simulations in the NMRlipids databank and extracts the required information using the information in README.yaml files. 
 The code that calculates permeabilities for all simulations in the NMRlipids Databank is available at \texttt{/scripts/calcMD-PERMEATION.py}, and the resulting permeabilities are stored at \texttt{/Data/MD-PERMEATION} in the repository containing all analyses specific for this publication at~\url{https://github.com/NMRLipids/DataBankManuscript/}.
-%While the order parameters, form factors, area per lipid and thickness are stored within the NMRlipids databank (\url{https://github.com/NMRLipids/Databank/tree/main/Data/Simulations}), further analyses can be conventiently stored in separate repositories with the same folder structure based on hash identities of trajectory and topology files.
-For example, results from further analyses performed here are stored in folders at \url{https://github.com/NMRLipids/DataBankManuscript/tree/main/Data}.
 This repository is organized similarly to the NMRlipids Databank repository, enabling the upcycling of also the analyzed data without overloading the main NMRlipids Databank repository.
@@
 
-H. S. A. contributed to the quality evaluation metrics based on NMR and SAXS, and was involved in testing the GUI and polishing the manuscript.
-Lara S. Bort implemented about 630 pre-existing lipid double layer simulations into the databank layer repository and created corresponding mapping files.
+A.M.K implemented the {\it Databank layer} and NMRlipids Databank-API.
+H.S.A. contributed to the quality evaluation metrics based on NMR and SAXS, and was involved in testing the GUI and polishing the manuscript.
+L.S.B. implemented about 630 pre-existing lipid bilayer simulations into the databank and created corresponding mapping files.
 P.B. implemented the PCA-based equilibration metric, supervised the work of A.K., and participated in code refactoring.
-B.K. contributed to the initial design and implementation of the databank's Python back-end and was involved with the writing the manuscript.
+F.F.-R. set up and performed simulations with Slipids Force Field: ID 617, ID 207 and ID 82. 
+T.M.F. performed and analysed the solid-state NMR experiments.
+P.F.J.F. contributed to changes in the buildH software so that the Databank can handle united atom simulation data.
 R.G.-F was responsible for the design, validation, supervision and funding of the back-end and front-end for the graphical user interface (GUI) (http://www.databank.nmrlipids.fi/).
+I.G. advised on implementation of the PCA-based equilibration metric and supervised the work of A.K.
+B.K. contributed to the initial design and implementation of the databank and was involved with the writing the manuscript.
+N.K. contributed experimental scattering data. % and SIMtoEXP algorithm.
+P.K. Implemented the flip-flop analysis.
+M.K. provided simulations with OPLS3e force field. 
+A.K. implemented the PCA-based equilibration metric.
+J.J.M. provided simulations with Amber force field, general discussion, and manuscript editing.
+R.N. implemented the code for the calculations of X-ray scattering form factors.
 A.P. was responsible for the design, validation, supervision and funding of the back-end and front-end for the graphical user interface (GUI) (http://www.databank.nmrlipids.fi/).
 F.S.-L. developed the back-end and participated in the design and development and validation of the graphical user interface (GUI) (http://www.databank.nmrlipids.fi/).
-P.F.J.F. contributed to changes in the buildH software so that the Databank can handle united atom simulation data.
-A.K. implemented the PCA-based equilibration metric. 
-I.G. advised on implementation of the PCA-based equilibration metric and supervised the work of A.K.
-N.K. contributed experimental scattering data and SIMtoEXP algorithm.
-M.K. provided simulations with OPLS3e force field. 
-T.M.F. performed and analysed the solid-state NMR experiments.
+O.H.S.O. Designed and supervised the work, implemented water permeation and diffusion analyses, and wrote the manuscript.
 

### 2023-04-20 main.tex
Update

@@
 \author[9,20,21]{Fabi{\'a}n Suarez-Leston}
-\author[1,*]{O. H. Samuli Ollila} %ORCID: 0000-0002- 8728-1006
+\author[1,23,*]{O. H. Samuli Ollila} %ORCID: 0000-0002- 8728-1006
 %\author[1,2,+]{Christine Author}
@@
 \affil[22]{Department of Physical Chemistry of Drugs and Faculty of Pharmacy, Comenius University Bratislava, 832 32 Bratislava, Slovakia}
+\affil[23]{VTT Technical Research Centre of Finland, Espoo, Finland}
 
@@
 M.K acknowledge CSC – IT Center for Science, Finland, for computational resources and thanks Finnish Cultural Foundation and the UEF Doctoral Programme for the financial support
+We acknowledge all the NMRlipids project contributors for making the NMRlipids databank development possible.
 
@@
 R.N. implemented the code for the calculations of X-ray scattering form factors.
+T.P. Performed many of the all-atom and united-atom simulations and contributed to discussions on the project.
 A.P. was responsible for the design, validation, supervision and funding of the back-end and front-end for the graphical user interface (GUI) (http://www.databank.nmrlipids.fi/).

### 2023-04-27 SI.tex
Update
Update text.

@@
-\documentclass[fleqn,10pt]{wlscirep}
+\documentclass[fleqn,10pt]{wlscirepSI}
 \usepackage[utf8]{inputenc}
@@
 \author[1]{Anne M. Kiirikki}         %ORCID: ?
-\author[2]{Hanne S. Antila}          %ORCID: 0000-0002-2474-5053
-\author[3]{Lara S. Bort}             %ORCID: ?
-\author[4]{Pavel Buslaev}         %ORCID: 0000-0003-2031-4691
-\author[5]{Fernando Favela-Rosales}
-\author[6]{Tiago Mendes Ferreira}
-\author[7,8]{Patrick F.J. Fuchs}
-\author[9]{Rebeca Garcia-Fandino}
-\author[10]{Ivan Gushchin}
+\author[2,3]{Hanne S. Antila}          %ORCID: 0000-0002-2474-5053
+\author[2,4]{Lara S. Bort}             %ORCID: ?
+\author[5]{Pavel Buslaev}         %ORCID: 0000-0003-2031-4691
+\author[6]{Fernando Favela-Rosales}
+\author[7]{Tiago Mendes Ferreira}
+\author[8,9]{Patrick F.J. Fuchs}
+\author[10]{Rebeca Garcia-Fandino}
+\author[11]{Ivan Gushchin}
 %\author[2]{Matti Javanainen}      %ORCID: 0000-0003-4858-364X
-\author[11,12]{Batuhan Kav}           %ORCID: 0000-0003-4990-373X
-\author[22]{Norbert Ku{\v c}erka}
-\author[13]{Patrik Kula}
-\author[14]{Milla Kurki}
-\author[10]{Alexander Kuzmin}
-\author[15,16]{Jesper J. Madsen}  %ORCID: 0000-0003-1411-9080
-\author[17,18]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
+\author[12,13]{Batuhan Kav}           %ORCID: 0000-0003-4990-373X
+\author[14]{Norbert Ku{\v c}erka}
+\author[15]{Patrik Kula}
+\author[16]{Milla Kurki}
+\author[11]{Alexander Kuzmin}
+\author[17,18]{Jesper J. Madsen}  %ORCID: 0000-0003-1411-9080
+\author[2,19,20]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
 %\author[2]{...}                   %ORCID: ?
 \author[1]{Ricky Nencini}
-\author[19]{Thomas Piggot}
-\author[20]{{\'A}ngel Pi{\~n}eiro}
-\author[10]{Suman Samantray}
-\author[9,20,21]{Fabi{\'a}n Suarez-Leston}
-\author[1,*]{O. H. Samuli Ollila} %ORCID: 0000-0002- 8728-1006
+\author[21]{Thomas J. Piggot}
+\author[22]{{\'A}ngel Pi{\~n}eiro}
+%\author[11]{Suman Samantray} %ORCID: 0000-0003-3361-9582
+\author[10,22,23]{Fabi{\'a}n Su{\'a}rez-Lest{\'o}n}
+\author[1,24,*]{O. H. Samuli Ollila} %ORCID: 0000-0002- 8728-1006
 %\author[1,2,+]{Christine Author}
@@
 \affil[1]{University of Helsinki, Institute of Biotechnology, Helsinki, Finland}
-\affil[2]{Department of Theory \& Bio-Systems and Department of Biomaterials, Max Planck Institute of Colloids and Interfaces, Germany}
-\affil[3]{University of Potsdam, Institute of Physics and Astronomy, Potsdam-Golm, 14476, Germany}
+\affil[2]{Department of Theory and Bio-Systems, Max Planck Institute of Colloids and Interfaces, 14424 Potsdam, Germany}
+\affil[3]{Department of Biomaterials, Max Planck Institute of Colloids and Interfaces, 14424 Potsdam, Germany}
+\affil[4]{University of Potsdam, Institute of Physics and Astronomy, Potsdam-Golm, 14476, Germany}
 %\affil[2]{Affiliation, department, city, postcode, country}
-\affil[4]{Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, P.O. Box 35, Jyv{\"a}skyl{\"a}, 40014 , Finland}
-\affil[5]{Departamento de Ciencias B\'{a}sicas, Tecnol\'{o}gico Nacional de M\'{e}xico - ITS Zacatecas Occidente, Sombrerete, Zacatecas, 99102, M\'{e}xico}
-\affil[6]{NMR group - Institute for Physics, Martin Luther University Halle-Wittenberg,  Halle (Saale), 06120, Germany}
-\affil[7]{Sorbonne Universit{\'e}, Ecole Normale Sup{\'e}rieure, PSL University, CNRS, Laboratoire des Biomol{\'e}cules (LBM), Paris, 75005, France}
-\affil[8]{Universit{\'e} Paris Cit{\'e}, UFR Sciences du Vivant, Paris, 75013, France}
-\affil[9]{Center for Research in Biological Chemistry and Molecular Materials (CiQUS), Universidade de Santiago de Compostela,  Santiago de Compostela, E-15782, Spain}
-\affil[10]{no affiliation}
-\affil[11]{Institute of Biological Information Processing: Structural Biochemistry (IBI-7), Forschungszentrum Jülich, Jülich 52428, Germany}
-\affil[12]{ariadne.ai GmbH (Germany), Häusserstra{\ss}e 3 Heidelberg 69115, Germany }
-\affil[13]{Institute of Organic Chemistry and Biochemistry of the Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2,  Prague, CZ-16610, Czech Republic}
-\affil[14]{School of Pharmacy, University of Eastern Finland, 70211 Kuopio, Finland}
-\affil[15]{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, 33612, United States of America}
-\affil[16]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida, 33612, United States of America}
-\affil[17]{Department of Chemistry, University of Bergen, Norway}
-\affil[18]{Computational Biology Unit, Department of Informatics, University of Bergen, Norway}
-\affil[19]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}
-\affil[20]{Department of Applied Physics, Faculty of Physics, University of Santiago de Compostela, Santiago de Compostela, E-15782, Spain}
-\affil[21]{MD.USE Innovations S.L., Edificio Emprendia, 15782 Santiago de Compostela, Spain}
-\affil[22]{Department of Physical Chemistry of Drugs and Faculty of Pharmacy, Comenius University Bratislava, 832 32 Bratislava, Slovakia}
+\affil[5]{Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, P.O. Box 35, Jyv{\"a}skyl{\"a}, 40014 , Finland}
+\affil[6]{Departamento de Ciencias B\'{a}sicas, Tecnol\'{o}gico Nacional de M\'{e}xico - ITS Zacatecas Occidente, Sombrerete, Zacatecas, 99102, M\'{e}xico}
+\affil[7]{NMR group - Institute for Physics, Martin Luther University Halle-Wittenberg,  Halle (Saale), 06120, Germany}
+\affil[8]{Sorbonne Universit{\'e}, Ecole Normale Sup{\'e}rieure, PSL University, CNRS, Laboratoire des Biomol{\'e}cules (LBM), Paris, 75005, France}
+\affil[9]{Universit{\'e} Paris Cit{\'e}, UFR Sciences du Vivant, Paris, 75013, France}
+\affil[10]{Center for Research in Biological Chemistry and Molecular Materials (CiQUS), Universidade de Santiago de Compostela,  Santiago de Compostela, E-15782, Spain}
+\affil[11]{no affiliation}
+%Physical and Computational Sciences Division, Pacific Northwest National Laboratory, Richland, Washington 99352, United States
+\affil[12]{Institute of Biological Information Processing: Structural Biochemistry (IBI-7), Forschungszentrum Jülich, Jülich 52428, Germany}
+\affil[13]{ariadne.ai GmbH (Germany), Häusserstra{\ss}e 3 Heidelberg 69115, Germany }
+\affil[14]{Department of Physical Chemistry of Drugs and Faculty of Pharmacy, Comenius University Bratislava, 832 32 Bratislava, Slovakia}
+\affil[15]{Institute of Organic Chemistry and Biochemistry of the Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2,  Prague, CZ-16610, Czech Republic}
+\affil[16]{School of Pharmacy, University of Eastern Finland, 70211 Kuopio, Finland}
+\affil[17]{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, 33612, United States of America}
+\affil[18]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida, 33612, United States of America}
+\affil[19]{Department of Chemistry, University of Bergen, 5020 Bergen, Norway}
+\affil[20]{Computational Biology Unit, Department of Informatics, University of Bergen, 5020 Bergen, Norway}
+\affil[21]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}
+\affil[22]{Department of Applied Physics, Faculty of Physics, University of Santiago de Compostela, Santiago de Compostela, E-15782, Spain}
+\affil[23]{MD.USE Innovations S.L., Edificio Emprendia, 15782 Santiago de Compostela, Spain}
+\affil[24]{VTT Technical Research Centre of Finland, Espoo, Finland}
 
@@
     \centering
-    \includegraphics[height = 0.85\textheight]{Figures/totalrankingCOMPL.pdf}
+    \includegraphics[height = 0.8\textheight]{Figures/totalrankingCOMPL.pdf}
     \caption{Top 50 simulations in the NMRlipids Databank ranked based on the C--H bond order parameter quality against experiments.

### 2023-04-27 CoverLetter.tex

\documentclass[11pt]{letter}
\usepackage{graphicx,baskervald,microtype}
\usepackage{hyperref,amsmath,xcolor}
\hypersetup{
  colorlinks=true,
  linkcolor=black,
  urlcolor=black
}
\usepackage{pgfplots,marginnote}
\usepackage[top=0.5in,bottom=0.5in,left=1in,right=1in]{geometry}
%\newgeometry{margin=2.5cm}
\newcommand{\omamargin}[1]{\marginnote{\textbf{#1}}[7pt]}
\begin{document}
\reversemarginpar
\pagestyle{empty}
\noindent Dr. O. H. Samuli Ollila \\
\noindent Institute of Biotechnology \\
\noindent University of Helsinki, Finland \\
\noindent samuli.ollila@helsinki.fi\\
\noindent +358\,503\,746\,963 \\

Dear Editor,

Please find enclosed our manuscript titled
{\it NMRlipids Databank makes data-driven analysis of biomembrane properties accessible for all}, which we would like to submit for your consideration to be published as a Resource in \textit{Nature Methods}.

In the manuscript, we present the NMRlipids Databank --- the first community-driven, open-for-all database featuring programmatic access to atom-resolution molecular dynamics (MD) simulations of cellular membranes. The Databank has been designed to improve the Findability, Accessibility, Interoperability, and Reuse (FAIR) of MD simulation data. Nearly thousand simulations with the total length approaching half a microsecond are made accessible to a wide interdisciplinary audience of academic and industry researchers through a graphical user interface (GUI) at \href{https://databank.nmrlipids.fi/}{\tt databank.nmrlipids.fi}, and for imaginative data-driven analyses through an application programming interface (API) at \href{https://github.com/NMRlipids/Databank}{\tt github.com/NMRlipids/Databank}.

By providing rapid open access to simulation data and automaticly  assessing
%ranking of simulations based on 
their quality against experiments, the NMRlipids Databank holds immediate practical relevance for the broad basic research community studying membranes in life sciences --- ranging from molecular cell biology to biomedical applications such as lipid nanoparticles. Even more importantly, the programmable interface for flexible implementation of data-driven and machine learning applications will advance new biological applications of MD simulation data.
We validate the practical relevance of the NMRlipids Databank here by successfully analysing how anisotropic diffusion of water and cholesterol flip-flop depend on membrane properties. Both of these processes are highly relevant for magnetic resonance imaging (MRI), pharmacokinetics, lipid trafficking, and regulating membrane properties of cellular membranes --- yet their timescales are beyond the scope of standard MD simulation investigations.

The NMRlipids Databank emerges from the NMRlipids open collaboration project started by us in 2013.
%with 314 citations in total) 
%where all the content is openly accessible throughout the project. 
The current manuscript demonstrates how MD simulation data can be shared by combining the open collaboration approach with an overlay databank structure, and describes how we take major steps in FAIRness by standardizing file formats, developing community guidelines, and creating incentives for sharing MD simulation data. Crucially, our success paves the way for corresponding smart guidelines and community-consensus best-practices for data sharing also in other fields than membrane MD simulations.
Such developments will, we believe, strongly support the reproducibility and reuse of data that is becoming increasingly important in the age of big data and artificial intelligence --- as also pointed out by a recent editorial in \textit{Nature Methods} (\href{https://doi.org/10.1038/s41592-023-01865-4}{\tt DOI: 10.1038/s41592-023-01865-4}).

We see that the NMRlipids Databank has both immediate and long term impact for a broad interdisciplinary audience of academic and industry researchers across
%at different fields in
life sciences: from biophysics and biochemistry to biology, biotechnology and biomedicine. Consequently, we trust our manuscript to fulfill the high standards for publication in %the \textit{Resource} section of
\textit{Nature Methods}.\\

%We now leave it for your own consideration.

Sincerely yours,

on behalf of the authors,\\

Dr. O. H. Samuli Ollila

\end{document}

### 2023-04-27 main.tex

--- 2023-04-19 main.tex
@@
 \subsection{NMRlipids Databank-API: programmatic access to the MD simulation data}\label{section:access}
-The NMRlipids Databank-API, available at \href{https://github.com/NMRlipids/Databank}{github.com/NMRlipids/Databank}, provides programmatic access to all simulation data in the NMRlipids Databank.
+The NMRlipids Databank-API, available at \href{https://github.com/NMRlipids/Databank}{github.com/NMRlipids/Databank}, provides programmatic access to all simulation data in the NMRlipids Databank through application programming interface (API).
 This enables wide range of novel data-driven applications---from construction of machine learning models that predict membrane properties, to automatic analysis of virtually any property across all simulations in the Databank.
@@
 As Fig.~\ref{fig:quality}A illustrates, predictions for the lateral packing of POPC and POPE membranes in terms of area per lipid, one of the most important parameters used to characterize cellular membranes, diverge between different force fields.
-To select the most realistic currently available force fields to describe POPC--POPE membranes, we first rank all simulations in the Databank according to the quality of their C--H bond order parameters  (Fig.~\ref{fig:top50simulations}).
+To select the most realistic currently available force fields to describe POPC--POPE membranes, we first order simulations in the Databank according to their C--H bond order parameter quality ranking (Fig.~\ref{fig:top50simulations}).
 From these results, we then pick the force fields occurring in Fig.~\ref{fig:quality}A (that is: force fields for which both POPC and POPE simulations exist in the Databank) and rank them according to the quality of the \textit{sn}-1 chain of POPC (Fig.~\ref{fig:quality}B) and of POPE (Fig.~\ref{fig:quality}C).
@@
 Automatic quality evaluation and ranking of simulations against experimental data enable the selection of best simulations for specific applications without laborious manual force field evaluation.
-This also streamlines automatic parametrization procedures for atomistic and coarse grained simulations.
+This also streamlines automatic parametrization procedures for atomistic and coarse grained simulations by, for example, pinpointing typical failures of force fields and highlighting points of improvement.
 Such practises for fostering the accuracy of simulations are becoming increasingly important due the accumulation of small errors when complexity and size of simulated systems are increasing.
@@
 Furthermore, the NMRlipids Databank-API delivers access to a training set data that paves the way for diverse machine learning applications to predict membrane properties.
-Such applications could be analogous to AlphaFold \cite{jumper21} and other tools that predict protein structures from their sequence using artificial intelligence.
+Such applications could be analogous to AlphaFold \cite{jumper21} and other tools~\cite{baek21,lin23} that predict protein structures from their sequence using artificial intelligence.
 These possibilities are particularly valuable for scientists who do not typically have access to large scale MD simulation data.
@@
 thanks Ministerio de Ciencia, Innovaci{\'o}n y Universidades for a ''Ram{\'o}n y Cajal'' contract (RYC-2016-20335), and Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (RTI2018-098795-A-I00, PDC2022-133402-I00), Xunta de Galicia and the ERDF (ED431F 2020/05, 02\_IN606D\_2022\_2667887 and Centro singular de investigaci{\'o}n de Galicia accreditation 2016-2019, ED431G/09).
-AP thanks Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (PID2019-111327GB-I00, PDC2022-133402-I00), Xunta de Galicia and the ERDF (ED431B 2022/36, 02\_IN606D\_2022\_2667887).
-F. S.-L. thanks Axencia Galega de Innovaci{\'o}n for his predoctoral contract (02\_IN606D\_2022\_2667887), and Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (PID2019-111327GB-I00).
-F.
-F.-R.
+A.P. thanks Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (PID2019-111327GB-I00, PDC2022-133402-I00), Xunta de Galicia and the ERDF (ED431B 2022/36, 02\_IN606D\_2022\_2667887).
+F.S.-L. thanks Axencia Galega de Innovaci{\'o}n for his predoctoral contract (02\_IN606D\_2022\_2667887), and Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (PID2019-111327GB-I00).
+F.F.-R.
 acknowledges Tecnol{\'o}gico Nacional de M{\'e}xico, Direcci{\'o}n General de Asuntos del Personal Acad{\'e}mico (DGAPA), Programa de Apoyo a Proyectos de Investigaci{\'o}n e Innovaci{\'o}n Tecnol{\'o}gica (PAPIIT) 101923, CONACyT Ciencia de Frontera 74884, for financial support and Miztli-Direcci{\'o}n de C{\'o}mputo y de Tecnolog{\'i}as de Informaci{\'o}n y Comunicaci{\'o}n (DGTIC) - Universidad Nacional Aut{\'o}noma de M{\'e}xico (UNAM) (Project LANCAD-UNAM-DGTIC-057) facilities for computing time allocation.
 M.S.M. was supported by the Trond Mohn Foundation BFS2017TMT01.
-O.H.S.O, A.M.K, and R.N. acknowledge CSC --- IT Center for Science for computational resources and Academy of Finland (grants 315596 and 319902) for financial support.
-M.K acknowledge CSC – IT Center for Science, Finland, for computational resources and thanks Finnish Cultural Foundation and the UEF Doctoral Programme for the financial support
+O.H.S.O, A.M.K, R.N. and L. S. B. acknowledge CSC --- IT Center for Science for computational resources and Academy of Finland (grants 315596 and 319902) for financial support.
+M.K. acknowledge CSC –-- IT Center for Science for computational resources and thanks Finnish Cultural Foundation and the UEF Doctoral Programme for financial support.
+We acknowledge all the NMRlipids Project contributors for making development of the NMRlipids Databank possible.
 
@@
 
-A.M.K implemented the {\it Databank layer} and NMRlipids Databank-API.
-H.S.A. contributed to the quality evaluation metrics based on NMR and SAXS, and was involved in testing the GUI and polishing the manuscript.
-L.S.B. implemented about 630 pre-existing lipid bilayer simulations into the databank and created corresponding mapping files.
+A.M.K. implemented the {\it Databank layer} and NMRlipids Databank-API.
+H.S.A. contributed to the quality evaluation metrics based on NMR and X-ray scattering form factors, and was involved in testing the GUI and polishing the manuscript.
+L.S.B. implemented roughly 630 pre-existing lipid bilayer simulations into the Databank and created the corresponding mapping files.
 P.B. implemented the PCA-based equilibration metric, supervised the work of A.K., and participated in code refactoring.
@@
 P.F.J.F. contributed to changes in the buildH software so that the Databank can handle united atom simulation data.
-R.G.-F was responsible for the design, validation, supervision and funding of the back-end and front-end for the graphical user interface (GUI) (http://www.databank.nmrlipids.fi/).
+R.G.-F. was responsible for the design, validation and supervision of the back-end and front-end for the graphical user interface (GUI) development (databank.nmrlipids.fi).
 I.G. advised on implementation of the PCA-based equilibration metric and supervised the work of A.K.
-B.K. contributed to the initial design and implementation of the databank and was involved with the writing the manuscript.
+B.K. contributed to the initial design and implementation of the Databank and was involved with writing the manuscript.
 N.K. contributed experimental scattering data. % and SIMtoEXP algorithm.
-P.K. Implemented the flip-flop analysis.
-M.K. provided simulations with OPLS3e force field. 
+P.K. Implemented the flip-flop analysis and supported the integration of PCA equilibration metric into the databank.
+M.K. provided simulations with the OPLS3e force field. 
 A.K. implemented the PCA-based equilibration metric.
-J.J.M. provided simulations with Amber force field, general discussion, and manuscript editing.
+J.J.M. provided simulations with the Amber force field, general discussion, and manuscript editing.
+M.S.M supported O.H.S.O.
+in conceptualising the project, contributed to the quality evaluation metrics based on NMR and X-ray scattering form factors, assisted in refining the manuscript and the GUI, and guided the work of L.S.B.
+and M.K.
 R.N. implemented the code for the calculations of X-ray scattering form factors.
-A.P. was responsible for the design, validation, supervision and funding of the back-end and front-end for the graphical user interface (GUI) (http://www.databank.nmrlipids.fi/).
-F.S.-L. developed the back-end and participated in the design and development and validation of the graphical user interface (GUI) (http://www.databank.nmrlipids.fi/).
-O.H.S.O. Designed and supervised the work, implemented water permeation and diffusion analyses, and wrote the manuscript.
+T.J.P. performed many of the all-atom and united-atom simulations and contributed to discussions on the project.
+A.P. was responsible for the design, validation and supervision of the back-end and front-end for the graphical user interface (GUI) development (databank.nmrlipids.fi).
+%S.S. performed some of the simulations with CHARMM36 force field.
+F.S.-L. developed the back-end and participated in the design and development and validation of the graphical user interface (GUI) (databank.nmrlipids.fi).
+O.H.S.O. designed and supervised the work, implemented the water permeation and diffusion analyses, and wrote the manuscript.
 

### 2023-04-28 main.tex
Small updates

@@
 While the NMRlipids Databank currently contains only lipid bilayer systems, its key elements that enable programmatic access to large-scale MD simulation data can be applied also for other molecules, such as disordered proteins or sugars.
-The powerful combination of an overlay databank structure and a community-wide open collaboration can be used to build databases that enable data-driven and machine learning applications also in other fields, where storage of raw data requires significant resources, best practices in the field are not yet defined, and incentives to share data do not exist---such as assignment of NMR spectra~\cite{klukowski22}.
+%The powerful combination of an overlay databank structure and a community-wide open collaboration can be used to build databases that enable data-driven and machine learning applications also in other fields, where storage of raw data requires significant resources, best practices in the field are not yet defined, and incentives to share data do not exist---such as assignment of NMR spectra~\cite{klukowski22}.
+The powerful combination of an overlay databank structure and the community-wide open collaboration  described here is potentially useful for building databases that enable data-driven and machine learning applications also in other fields.
+This is especially true for research activities where storage of raw data requires significant resources, best practices in the field are not yet defined, and incentives to share data do not exist---such as assignment of NMR spectra~\cite{klukowski22}.
 
@@
 O.H.S.O, A.M.K, R.N. and L. S. B. acknowledge CSC --- IT Center for Science for computational resources and Academy of Finland (grants 315596 and 319902) for financial support.
-M.K. acknowledge CSC –-- IT Center for Science for computational resources and thanks Finnish Cultural Foundation and the UEF Doctoral Programme for financial support.
+M.K.
+acknowledge CSC –-- IT Center for Science for computational resources and thanks Finnish Cultural Foundation and the UEF Doctoral Programme for financial support.
+T.M.F.
+greatly acknowledges financial support by the Ministry of Economics, Science and Digitalisation of the State of Saxony-Anhalt.
 We acknowledge all the NMRlipids Project contributors for making development of the NMRlipids Databank possible.

### 2023-04-28 main.tex
Update

@@
 The {\it Data layer} contains raw data that can be distributed to publicly available servers such as Zenodo (\href{https://www.zenodo.org/}{zenodo.org}).
-The core content of the Databank locates in the {\it Databank layer}, which is a git repository at \href{https://github.com/NMRlipids/Databank/}{github.com/NMRlipids/Databank} and is also permanently stored in a Zenodo repository (\href{https://www.zenodo.org/}{zenodo.org})~\cite{??}.
+The core content of the Databank locates in the {\it Databank layer}, which is a git repository at \href{https://github.com/NMRlipids/Databank/}{github.com/NMRlipids/Databank} and is also permanently stored in a Zenodo repository (\href{https://doi.org/10.5281/zenodo.7875567}{DOI: 10.5281/zenodo.7875567}).
 The essential information of each simulation is stored in a human-and-machine-readable README.yaml file located in a subfolder of the \texttt{/Data/Simulations} folder in the {\it Databank layer} repository; each subfolder has a unique name constructed based on a hash code of the trajectory and topology files of each simulation.

### 2023-05-10 main.tex
Update title, introduction and discussion based on feedback from a journal

@@
 
-\title{NMRlipids Databank makes data-driven analysis of biomembrane properties accessible for all}
+\title{Overlay databanks unlock data-driven analyses of biomolecules for all}
+%NMRlipids Databank makes data-driven analysis of biomembrane properties accessible for all}
 %NMRlipids Databank: Making data-driven analyses of membrane properties accessible for all^
@@
 \begin{abstract}
+Tools based on artificial intelligence (AI) are currently revolutionising many fields, yet their applications are often limited by the lack of suitable training data in programmatically accessible format.
+Here we propose an effective solution to make data scattered in various locations and formats accessible for data-driven and machine learning applications using the overlay databank format.
+To demonstrate the practical relevance of such approach, we present the NMRlipids Databank---a community-driven, open-for-all database featuring programmatic access to quality-evaluated atom-resolution molecular dynamics simulations of cellular membranes.
 Cellular membrane lipid composition is implicated in diseases and controls major biological functions, but membranes are difficult to study experimentally due to their intrinsic disorder and complex phase behaviour.
+While MD simulations have been useful in understanding membrane systems, they require significant computational resources and often suffer from inaccuracies in model parameters.
+Here, we demonstrate how programmable interface for flexible implementation of data-driven and machine learning applications, and rapid access to simulation data through a graphical user interface unlocks possibilities beyond current MD simulation and experimental studies to understand cellular membranes.
+The proposed overlay databank concept can be further applied to other biomolecules, as well as in other fields where similar barriers hinder the AI revolution.
+
+%Cellular membrane lipid composition is implicated in diseases and controls major biological functions, but membranes are difficult to study experimentally due to their intrinsic disorder and complex phase behaviour.
 Molecular dynamics (MD) simulations have been useful in understanding membrane systems, but they require significant computational resources and often suffer from inaccuracies in model parameters.
@@
 
+Tools based on artificial intelligence (AI) are currently revolutionising many fields, yet their success relies on training data that is available in programmatically accessible and standardized format~\cite{sever23}.
+For example in structural biology, Protein Data Bank (PDB)~\cite{montelione13} has enabled revolutionary tools that can predict protein structures with unprecedented accuracy~\cite{jumper21}.
+However, the lack of smart guidelines and community-consensus best-practices for data sharing are limiting the development of such databanks and AI applications in many fields~\cite{sever23}.
+This is the case in biomolecular modeling where vast amount of data is already available for proteins, lipids, nucleic acids and carbohydrates, but tools that enable applications of this data for data-driven applications are not yet available~\cite{tiemann23}.
+Here we propose a cost-effective solution to make data scattered in various locations and formats accessible for data-driven and machine learning applications using the overlay databank format.
+We demonstrate the practical relevance of such approach for understanding cellular membranes by incorporating available lipid bilayer simulations into the NMRlipids databank.
+Nevertheless, the basic principles of overlay databanks can be applied also to simulation data of other biomolecules that are becoming available in increasing amounts~\cite{tiemann23}, as well as to other fields where development of AI based tools are limited by the data access and best best community practices --- such as assignment of NMR spectra~\cite{klukowski22}.
+Incentives to share data can be further accelerated by combining overlay databanks with open collaboration approach developed by us~\cite{botan15}.
+
 Cellular membranes contain hundreds of different types of lipid molecules that regulate the membrane properties, morphology, and biological functions~\cite{vanmeer08,Lorent:2020a,Slatter:2016a}.
 Membrane lipid composition is implicated in diseases, such as cancer and neurodegenerative disorders, and therapeutics that affect membrane compositions are emerging~\cite{torres21}.
-However, biomembranes are often difficult to study experimentally, because they are complex mixtures of proteins and lipids in disordered fluid state with complicated phase behaviour at physiologicalbiological conditions.
-For those reasons, the detailed connections between complex lipid interactions and biological functions taking place in or around membranes remain poorly understood.
-Molecular dynamics (MD) simulations have been particularly useful in understanding membrane systems, although their accuracy has often been compromised by artefacts such as the quality of model parameters~\cite{antila22b,gupta22}.
+However, biomembranes are often difficult to study experimentally, because they are complex mixtures of proteins and lipids in disordered fluid state with complicated phase behaviour at biological conditions.
+Molecular dynamics (MD) simulations can be used to model biomembranes in detail, but computational cost for simulating all possible biological membrane compositions would be substantial.
+For those reasons, data-driven and machine learning based models that predict biomembrane properties will benefit wide range of fields covering academia and industry, from cell membrane biology to lipid nanoparticle formulations.
+%For those reasons, the detailed connections between complex lipid interactions and biological functions taking place in or around membranes remain poorly understood. 
+%Molecular dynamics (MD) simulations have been particularly useful in understanding membrane systems, although their accuracy has often been compromised by artefacts such as the quality of model parameters~\cite{antila22b,gupta22}.
 Presently, the accuracy of models is becoming increasingly important as researches are progressing from simulations of individual molecules to simulating whole organelles or even cells using interdisciplinary approaches~\cite{johnson15,thornburg22,gupta22}.
@@
 
-In contrast to experimental structural biology, where standard protocols to share and quality-evaluate resolved biomolecular structures are firmly established by the Protein Data Bank (PDB)~\cite{montelione13}, equivalent best practices are yet to be defined for MD simulations.
+%In contrast to experimental structural biology, where standard protocols to share and quality-evaluate resolved biomolecular structures are firmly established by the Protein Data Bank (PDB)~\cite{montelione13}, equivalent best practices are yet to be defined for MD simulations.
 The importance of such approaches is widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} and data-sharing solutions are emerging for proteins in solution~\cite{meyer10,kamp10}, proteins in membranes~\cite{newport19,espigares20,leston22}, nucleic acids~\cite{hospital16}, nucleic acids and proteins~\cite{bekker20}, cyclodextrins~\cite{mixcoha16}, COVID-19--involved macromolecules~(\href{https://bioexcel-cv19.bsc.es}{bioexcel-cv19.bsc.es}).
@@
 Here we present the NMRlipids Databank --- a community-driven, open-for-all database featuring programmatic access to atom-resolution MD simulations of lipid bilayers.
-The programmatic access enables users to apply their data-driven approaches on the Databank, thus facilitating the creation of new tools for (and by) researchers in a wide range of fields covering academia and industry, from cell membrane biology and lipid nanoparticle formulations to computational chemistry and machine learning.
-As two usage examples of what is already possible, we demonstrate here (i) how data-driven analysis of water anisotropic diffusion in all the membrane systems available in the Databank can extend the scope of MD simulations to magnetic resonance imaging (MRI) and pharmacokinetics, and (ii) how the Databank allows its users to analyse rare phenomena that are beyond the scope of standard MD simulation investigations.
-Wider adaptation of the NMRlipids Databank will open even more possibilities.
+To demonstrate its advances to available approaches, we show how rare phenomena that are beyond the scope of standard MD simulation investigations can be analyzed from the Databank.
+To demonstrate the immediate relevance of NMRlipids databank in extending the scope of MD simulations to new fields, we  use data-driven approach to analyse how anisotropic diffusion of water depends on membrane properties.
+This benefits understanding of magnetic resonance imaging (MRI)~\cite{topgaard20} and pharmacokinetics~\cite{nitsche19} where MD simulations are rarely applied.
+%Wider adaptation of the NMRlipids Databank will open even more possibilities. 
 Furthermore, the Databank performs automatic quality evaluation of membrane simulations, which facilitates the selection of best-performing models for each given application and accelerates the development of simulation parameters and methodology.
+The overlay databank and open collaboration approaches facilitate programmatic access and collection of community contributed data also in other fields than membrane simulations.
+For example, substantial biomolecular MD simulation data is already available but in scattered locations and formats~\cite{tiemann23}.
+The technical advances presented here will immediately benefit in making this data programmatically accessible.
+Furthermore, the overlay databank configuration will benefit broad range of fields where access to training data is bottle neck for building AI based tools.
+%NMRlipids databanksolutions presented as well as broader communities that 
+%enables users to apply their data-driven approaches on the Databank, thus facilitating the creation of new tools for (and by) researchers in a .
+% to computational chemistry and machine learning.
 
-While the NMRlipids Databank currently contains only lipid bilayer systems, its key elements that enable programmatic access to large-scale MD simulation data can be applied also for other molecules, such as disordered proteins or sugars.
-%The powerful combination of an overlay databank structure and a community-wide open collaboration can be used to build databases that enable data-driven and machine learning applications also in other fields, where storage of raw data requires significant resources, best practices in the field are not yet defined, and incentives to share data do not exist---such as assignment of NMR spectra~\cite{klukowski22}.
-The powerful combination of an overlay databank structure and the community-wide open collaboration  described here is potentially useful for building databases that enable data-driven and machine learning applications also in other fields.
+%While the NMRlipids Databank currently contains only lipid bilayer systems, its key elements that enable programmatic access to large-scale MD simulation data can be applied also for other molecules, such as disordered proteins or sugars.
+%%The powerful combination of an overlay databank structure and a community-wide open collaboration can be used to build databases that enable data-driven and machine learning applications also in other fields, where storage of raw data requires significant resources, best practices in the field are not yet defined, and incentives to share data do not exist---such as assignment of NMR spectra~\cite{klukowski22}.
+%The powerful combination of an overlay databank structure and the community-wide open collaboration  described here is potentially useful for building databases that enable data-driven and machine learning applications also in other fields.
 This is especially true for research activities where storage of raw data requires significant resources, best practices in the field are not yet defined, and incentives to share data do not exist---such as assignment of NMR spectra~\cite{klukowski22}.
@@
 %
+Molecular dynamics (MD) simulations have been particularly useful in understanding membrane systems, although their accuracy has often been compromised by artefacts such as the quality of model parameters~\cite{antila22b,gupta22}.
+Presently, the accuracy of models is becoming increasingly important as researches are progressing from simulations of individual molecules to simulating whole organelles or even cells using interdisciplinary approaches~\cite{johnson15,thornburg22,gupta22}.
+Such systems exhibit intricate emergent behaviour making inaccuracies more difficult to detect, and accumulation of even modest errors may have a dramatic impact on the conclusions drawn.
 To minimize the detrimental consequences of artificial MD simulation results for their applications, the quality of lipid bilayer MD simulations has to be carefully assessed~\cite{antila22b}.
@@
 
+Sharing of biomolecular MD simulation and other data is becoming increasingly important in the age of big data and artificial intelligence~\cite{sever23,tiemann23}. 
+%The main barriers for such solutions have been the required commitment for the long term support of hardware and software, and the lack of incentives for researchers to share the data. 
+Besides the data itself, also programmatic access is a necessary requirement for data-driven and machine learning applications.
+This is particularly challenging when field specific smart guidelines and community-consensus best-practices have not yet been defined, which is the case in biomolecular simulations~\cite{tiemann23}.
+The NMRlipids Databank demonstrates how these issues can be solved by the overlay databank design where the raw data is distributed to already publicly available decentralized locations, while the core of the Databank is composed only of the metadata stored in a version-controlled git repository with an open-access license.
+On the other hand, the open collaboration approach developed in the NMRlipids Project~\cite{botan15} creates incentives for sharing the data by offering authorship in published articles to the contributors.
+Advantages of such an approach are demonstrated here for membrane simulations, yet the concept can be applied to other biomolecules, %such as disordered proteins and membrane--protein systems,
+as well as in other fields where similar barriers hinder the AI revolution, such as the assignment of NMR spectra~\cite{klukowski22}.
+
+The NMRlipids Databank-API delivers programmatic access to MD simulation data that can be used as training set for diverse data-driven and machine learning applications that predict membrane properties.
+Such applications could be analogous to AlphaFold \cite{jumper21} and other tools~\cite{baek21,lin23} that predict protein structures from their sequence using artificial intelligence.
+%The open programmatic access to unprecedented amounts of MD simulation data through the NMRlipids Databank-API enables a wide range of novel data-driven analyses and machine learning applications. 
+The analyses of cholesterol flip-flop events (Fig.~\ref{fig:flip-flops}) and water permeation through membranes (Fig.~\ref{fig:permeability}) demonstrate how a large amount of accessible simulation data in terms of quantity (\textit{e.g.}, simulation length and number of conformations) and content (\textit{e.g.}, lipid compositions and ion concentrations) enable analyses of rare phenomena that are beyond the current possibilities for a single research group.
+Such analyses also pave the way for applications of MD simulations in new fields, as demonstrated here by analysing an essential parameter in pharmacokinetic modeling and MRI imaging:\cite{nitsche19,topgaard20} the anisotropic diffusion of water in membrane systems (Fig.~\ref{fig:permeability}).
+These possibilities are particularly valuable for scientists who do not typically have access to large scale MD simulation data.
+
 The focus of biomolecular simulations is moving from studies of individual molecules to larger complexes and even whole cells and organelles~\cite{johnson15,thornburg22,gupta22}.
@@
 Such practises for fostering the accuracy of simulations are becoming increasingly important due the accumulation of small errors when complexity and size of simulated systems are increasing.
-
-The open programmatic access to unprecedented amounts of MD simulation data through the NMRlipids Databank-API enables a wide range of novel data-driven analyses and machine learning applications.
-The analyses of cholesterol flip-flop events (Fig.~\ref{fig:flip-flops}) and water permeation through membranes (Fig.~\ref{fig:permeability}) demonstrate how a large amount of accessible simulation data in terms of quantity (\textit{e.g.}, simulation length and number of conformations) and content (\textit{e.g.}, lipid compositions and ion concentrations) enable analyses of rare phenomena that are beyond the current possibilities for a single research group.
-Such analyses also pave the way for applications of MD simulations in new fields, as demonstrated here by analysing an essential parameter in pharmacokinetic modeling and MRI imaging:\cite{nitsche19,topgaard20} the anisotropic diffusion of water in membrane systems (Fig.~\ref{fig:permeability}).
-Furthermore, the NMRlipids Databank-API delivers access to a training set data that paves the way for diverse machine learning applications to predict membrane properties.
-Such applications could be analogous to AlphaFold \cite{jumper21} and other tools~\cite{baek21,lin23} that predict protein structures from their sequence using artificial intelligence.
-These possibilities are particularly valuable for scientists who do not typically have access to large scale MD simulation data.
-Expected applications of the NMRlipids Databank in a wide range of fields are listed in Table~\ref{tab:applications}, yet the scope of such applications
-%in biology, biotechnology, material science, biomolecular imaging and beyond
-is expected to further widen with increasing amount of data in the Databank.
+Examples of impact of NMRlipids and other overlay Databanks in different disciplines are listed in Table~\ref{tab:applications}, yet the scope of applications is expected to further widen with increasing amount of publicly shared data.
 
@@
     \centering
-    \begin{tabular}{p{5.0cm}  p{5.0cm}  p{4.0cm}}
-    Type of application     & Practical examples & Target group \\
+    \begin{tabular}{p{3.0cm}  p{4.0cm}  p{7.0cm}}
+     Target group  &  Outcomes     & Practical examples \\
     \hline
-    Analyses of rare phenomena               & Lipid flip-flops, water permeation (Figs.~\ref{fig:flip-flops} and~\ref{fig:permeability}) & Membrane scientists \\
-    \\
-    Correlations between membrane properties & 
-    Membrane structural properties, water dynamics (Figs.~\ref{fig:quality} and~\ref{fig:permeability}) & 
-    Membrane scientists \\
-    \\
-    Applications that are outside the typical scope of MD simulations & 
-    Anisotropic water diffusion for pharmacokinetics and MRI imaging applications (Fig.~\ref{fig:permeability}) & 
-    Scientists in fields where MD simulations are not usually applied \\
-    \\
-    Selection of the best simulation model for a specific application & 
-    Best model for membranes with PC and PE lipids (Fig.~\ref{fig:quality}), lipid headgroup conformations~\cite{bacle21}, 
-    packing of PS~\cite{antila22b} and PE (Fig.~\ref{fig:quality}) containing membranes &
-    Scientists using MD simulations \\
-    \\
-    Guidance for force field development & 
-    Improvements in ion binding to lipids~\cite{melcr18,melcr20} and lipid headgroup conformational ensembles~\cite{yu21,dickson22,grote20} &
-    Scientists developing parameters for MD simulations \\
-    \\
-    Training and target data for the development of coarse grained models & 
-    Optimizing parameters of coarse grained models against NMRlipids Databank, extracting continuum parameters for membranes &
-    Scientists developing and using coarse grained MD simulations \\
-    \\
-    Training set for machine learning applicatons &
-    Programmatic access to the data and results enables training of machine learning models for various applications, such as predictions of membrane properties from composition & Scientists building and using machine learning applications for biomolecules
+    Computational and data scientists & Databanks with programmatic access, training sets for machine learning applications, automatic optimization of simulation models & Programmatic access to all publicly available MD simulation trajectories, machine learning models predicting properties of complex biomolecular assemblies, optimizing force field and model parameters from atomistic to continuum scale \\
+
+    Biomedical scientists & Applications of biomolecular modeling in new fields, coupling omics to biomolecular structure and dynamics & Anisotropic water diffusion for pharmacokinetics and MRI imaging applications (Fig.~\ref{fig:permeability}), properties of complex cellular structures based on composition analyses from omics \\
+
+    Material scientists & Predictions of complex biomolecular material properties from data-driven and machine learning models & Optimising composition of lipid nanoparticle formulations for desired properties, predictions of bioinspired material properties \\
+
+    Biophysicists &  Novel analyses and predictions for properties of biomolecular assemblies and their compositions  & Analyses of rare phenomena (for example, Lipid flip-flops and water permeation in Figs.~\ref{fig:flip-flops} and~\ref{fig:permeability}), data-driven analyzes and machine learning models  predicting correlations between properties and compositions of biomolecular assemblies (Figs.~\ref{fig:quality} and~\ref{fig:permeability}) \\
+
+    Students and teachers & Graphical and programmatic access to biomolecular structure and dynamics & Illustration of disorder and dynamics in biomolecules, example data for bioinformatics and data-science courses \\
+
+    Reviewers and publishers & Tool to facilitate FAIRness of data, transparency and reproducibility & All publicly available data can be included in an overlay databank irrespectively on its location \\
+
     \end{tabular}
-    \caption{Examples on applications of the NMRlipids Databank.}
+    \caption{Examples of impact of NMRlipids and other overlay databanks in different disciplines.}
     \label{tab:applications}
@@
 
-Potential benefits of sharing MD simulation data are widely recognized by different stakeholders from scientists~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} to funders and journal staff, but data sharing with efficient tools for upcycling have been rare.
+%\begin{table}[t]
+%    \centering
+%    \begin{tabular}{p{5.0cm}  p{5.0cm}  p{4.0cm}}
+%    Type of application     & Practical examples & Target group \\
+%    \hline
+%    Analyses of rare phenomena               & Lipid flip-flops, water permeation (Figs.~\ref{fig:flip-flops} and~\ref{fig:permeability}) & Membrane scientists \\
+%    \\
+%    Correlations between membrane properties & 
+%    Membrane structural properties, water dynamics (Figs.~\ref{fig:quality} and~\ref{fig:permeability}) %& 
+%    Membrane scientists \\
+%    \\
+%    Applications that are outside the typical scope of MD simulations & 
+%    Anisotropic water diffusion for pharmacokinetics and MRI imaging applications %(Fig.~\ref{fig:permeability}) & 
+%    Scientists in fields where MD simulations are not usually applied \\
+%    \\
+%    Selection of the best simulation model for a specific application & 
+%    Best model for membranes with PC and PE lipids (Fig.~\ref{fig:quality}), lipid headgroup %conformations~\cite{bacle21}, 
+%    packing of PS~\cite{antila22b} and PE (Fig.~\ref{fig:quality}) containing membranes &
+%    Scientists using MD simulations \\
+%    \\
+%    Guidance for force field development & 
+%    Improvements in ion binding to lipids~\cite{melcr18,melcr20} and lipid headgroup conformational %ensembles~\cite{yu21,dickson22,grote20} &
+%    Scientists developing parameters for MD simulations \\
+%    \\
+%    Training and target data for the development of coarse grained models & 
+%    Optimizing parameters of coarse grained models against NMRlipids Databank, extracting continuum %parameters for membranes &
+%    Scientists developing and using coarse grained MD simulations \\
+%   \\
+%    Training set for machine learning applicatons &
+%    Programmatic access to the data and results enables training of machine learning models for various %applications, such as predictions of membrane properties from composition & Scientists building and %using machine learning applications for biomolecules
+%    \end{tabular}
+%    \caption{Examples on applications of the NMRlipids Databank.}
+%    \label{tab:applications}
+%\end{table}
+
+%Potential benefits of sharing MD simulation data are widely recognized by different stakeholders from scientists~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} to funders and journal staff, but data sharing with efficient tools for upcycling have been rare.
 The main barriers for such solutions have been the required commitment for the long term support of hardware and software, and the lack of incentives for researchers to share the data.

### 2023-05-12 main.tex
Update

@@
 
-\title{Overlay databanks unlock data-driven analyses of biomolecules for all}
+\title{Overlay databank unlocks data-driven analyses of biomolecules for all}
 %NMRlipids Databank makes data-driven analysis of biomembrane properties accessible for all}
@@
 While MD simulations have been useful in understanding membrane systems, they require significant computational resources and often suffer from inaccuracies in model parameters.
-Here, we demonstrate how programmable interface for flexible implementation of data-driven and machine learning applications, and rapid access to simulation data through a graphical user interface unlocks possibilities beyond current MD simulation and experimental studies to understand cellular membranes.
+Here, we demonstrate how programmable interface for flexible implementation of data-driven and machine learning applications, and rapid access to simulation data through a graphical user interface, unlock possibilities beyond current MD simulation and experimental studies to understand cellular membranes.
 The proposed overlay databank concept can be further applied to other biomolecules, as well as in other fields where similar barriers hinder the AI revolution.
@@
 However, the lack of smart guidelines and community-consensus best-practices for data sharing are limiting the development of such databanks and AI applications in many fields~\cite{sever23}.
-This is the case in biomolecular modeling where vast amount of data is already available for proteins, lipids, nucleic acids and carbohydrates, but tools that enable applications of this data for data-driven applications are not yet available~\cite{tiemann23}.
-Here we propose a cost-effective solution to make data scattered in various locations and formats accessible for data-driven and machine learning applications using the overlay databank format.
-We demonstrate the practical relevance of such approach for understanding cellular membranes by incorporating available lipid bilayer simulations into the NMRlipids databank.
-Nevertheless, the basic principles of overlay databanks can be applied also to simulation data of other biomolecules that are becoming available in increasing amounts~\cite{tiemann23}, as well as to other fields where development of AI based tools are limited by the data access and best best community practices --- such as assignment of NMR spectra~\cite{klukowski22}.
+This is the case in biomolecular modeling, where vast amount of data is already available for proteins, lipids, nucleic acids, and carbohydrates, but tools that enable applications of these data for data-driven applications are not yet available~\cite{tiemann23}.
+Here we propose a cost-effective solution to make data scattered in various locations and formats accessible for data-driven and machine learning applications using an overlay databank.
+We demonstrate the practical relevance of such approach for understanding cellular membranes by incorporating available lipid bilayer simulations into the NMRlipids Databank.
+Importantly, the basic principles of an overlay databank can be applied also to simulation data of other biomolecules that are becoming available in increasing amounts~\cite{tiemann23}, as well as to other any field where the development of AI-based tools is limited by the lack of data access and community best-practices --- such as assignment of NMR spectra~\cite{klukowski22}.
 Incentives to share data can be further accelerated by combining overlay databanks with open collaboration approach developed by us~\cite{botan15}.
@@
 
-\subsection{MD simulations of membranes composed of the biologically most abundant lipids}
+\subsection{NMRlipids overlay databank delivers access to MD simulations of membranes composed of the biologically most abundant lipids}
 

### 2023-05-12 CoverLetter2.tex

--- 2023-04-27 CoverLetter.tex
@@
 Please find enclosed our manuscript titled
-{\it NMRlipids Databank makes data-driven analysis of biomembrane properties accessible for all}, which we would like to submit for your consideration to be published as a Resource in \textit{Nature Methods}.
+{\it NMRlipids Databank makes data-driven analysis of biomembrane properties accessible for all}, which we would like to submit for your consideration to be published as a Resource in \textit{ Nature Structural \& Molecular Biology}.
 
@@
 
-By providing rapid open access to simulation data and automaticly  assessing
-%ranking of simulations based on 
-their quality against experiments, the NMRlipids Databank holds immediate practical relevance for the broad basic research community studying membranes in life sciences --- ranging from molecular cell biology to biomedical applications such as lipid nanoparticles.
+By providing rapid open access to simulation data and automaticly assessing their quality against experiments, the NMRlipids Databank holds immediate practical relevance for the broad basic research community studying membranes in life sciences --- ranging from molecular cell biology to biomedical applications such as lipid nanoparticles.
 Even more importantly, the programmable interface for flexible implementation of data-driven and machine learning applications will advance new biological applications of MD simulation data.
-We validate the practical relevance of the NMRlipids Databank here by successfully analysing how anisotropic diffusion of water and cholesterol flip-flop depend on membrane properties.
+We validate the practical relevance of the NMRlipids Databank on functional and mechanistic understanding of biological process
+%how molecular components in a work together
+by successfully analysing how anisotropic diffusion of water and cholesterol flip-flop depend on membrane properties.
 Both of these processes are highly relevant for magnetic resonance imaging (MRI), pharmacokinetics, lipid trafficking, and regulating membrane properties of cellular membranes --- yet their timescales are beyond the scope of standard MD simulation investigations.
@@
 
-We see that the NMRlipids Databank has both immediate and long term impact for a broad interdisciplinary audience of academic and industry researchers across
-%at different fields in
-life sciences: from biophysics and biochemistry to biology, biotechnology and biomedicine.
-Consequently, we trust our manuscript to fulfill the high standards for publication in %the \textit{Resource} section of
-\textit{Nature Methods}.\\
+We see that the NMRlipids Databank has both immediate and long term impact for a broad interdisciplinary audience of academic and industry researchers across life sciences: from biophysics and biochemistry to biology, biotechnology and biomedicine.
+Feedback on our first submission to \textit{Nature Methods} states that the manuscript is of interest on this field of research.
+Therefore, we trust our manuscript to fulfill the high standards for publication in \textit{ Nature Structural \& Molecular Biology}.
+.\\
 

### 2023-05-12 CoverLetter3.tex

\documentclass[11pt]{letter}
\usepackage{graphicx,baskervald,microtype}
\usepackage{hyperref,amsmath,xcolor}
\hypersetup{
  colorlinks=true,
  linkcolor=black,
  urlcolor=black
}
\usepackage{pgfplots,marginnote}
\usepackage[top=0.2in,bottom=0.5in,left=1in,right=1in]{geometry}
%\newgeometry{margin=2.5cm}
\newcommand{\omamargin}[1]{\marginnote{\textbf{#1}}[7pt]}
\begin{document}
\reversemarginpar
\pagestyle{empty}
\noindent Dr. O. H. Samuli Ollila \\
\noindent Institute of Biotechnology, University of Helsinki, Finland \\
\noindent samuli.ollila@helsinki.fi, +358\,503\,746\,963

Dear Editor,

Please find enclosed our manuscript titled {\it Overlay databank unlocks data-driven analyses of biomolecules for all}. Previous version of the manuscript, titled {\it NMRlipids Databank makes data-driven analysis of biomembrane properties accessible for all}, was rejected from \textit{Nature Methods} by Dr. Arunima Singh (NMETH-RS52135) because it was not considered to have a sufficiently significant and immediate impact on a broader readership. After receiving this feedback, we re-evaluated our manuscript and concluded that it was indeed probably formulated for too specialized audience. That said, we do believe that the presented overlay databank concept and its practical implementation have significant and immediate impact on broad readership much beyond the membrane biophysics field (see below for more detailed justification). Thus, inspired by the previous feedback, we have rewritten the manuscript to better convey its importance to a wide audience. We want thank you for the feedback on the previous version of our manuscript, and kindly ask you to consider the reworked manuscript to be published as a Resource in \textit{Nature Methods}.

Importance and potential of making data accessible in the age of big data and artificial intelligence (AI) has been recently recognized in correspondence (\href{https://doi.org/10.1038/s41592-023-01817-y}{\tt DOI: 10.1038/s41592-023-01817-y}) and editorial articles (\href{https://doi.org/10.1038/s41592-023-01865-4}{\tt DOI: 10.1038/s41592-023-01865-4}) in \textit{Nature Methods}. In the NMRlipids open collaboration (\href{http://nmrlipids.blogspot.com/}{\tt nmrlipids.blogspot.com}), we have been promoting Findability, Accessibility, Interoperability, and Reuse (FAIR) of molecular dynamics (MD) simulation data for biomolecules almost a decade now. Other scientists have started to recently follow our example, and the amount of publicly accessible MD simulation data is increasing, although a significant fraction still originates from the NMRlipids Project. This is summarized by an authorative group of eminent scientists from different disciplines of MD simulations in a very recent preprint article (\href{http://dx.doi.org/10.1101/2023.05.02.538537}{\tt DOI: 10.1101/2023.05.02.538537}, published on May 2$^\mathrm{nd}$ 2023), which also %appropriately credits NMRlipids project and 
cites the preprint of the manuscript submitted here. However, for big-data and AI applications the data need to be made available also in a programmatically accessible format. This is typically an issue in fields where smart guidelines and community-consensus best-practices for data sharing have not been defined, such as biomolecular modeling.

Here we show for the first time how data in scattered locations can be made programmatically accessible for data-driven and machine learning applications 
%programmable interfaces for flexible implementation of  can be build using publicly available  and 
using the overlay databank concept. 
%Here we propose a cost-effective solution to make data scattered in various locations and formats accessible for data-driven and machine learning applications using the overlay databank format. 
The practical relevance of this approach is demonstrated by introducing the NMRlipids Databank --- the first community-driven, open-for-all database featuring programmatic access to atom-resolution molecular dynamics simulations of biomolecules. Crucially, as the overlay databank concept can be applied beyond applications in biomolecular simulations, we are convinced that our manuscript has significant and immediate impact on the broad readership of \textit{Nature Methods}.

The NMRlipids Databank makes nearly thousand simulations --- with the total length approaching half a microsecond --- accessible to a wide interdisciplinary audience of academic and industry researchers through a graphical user interface (GUI) at \href{https://databank.nmrlipids.fi/}{\tt databank.nmrlipids.fi}, and for imaginative data-driven analyses through an application programming interface (API) at \href{https://github.com/NMRlipids/Databank}{\tt github.com/NMRlipids/Databank}. By providing rapid open access to biomolecular simulation data and automaticly assessing their quality against experiments, the NMRlipids Databank holds immediate practical relevance for broad audience in life sciences --- ranging from molecular cell biology to biomedical applications such as lipid nanoparticles. Furthermore, it enables novel analyses that can advance new biological applications of MD simulation data in unprecedented directions: We were already approached by medical scientists who, after reading the preprint of our article, wish to use the NMRlipids Databank to predict pharmacokinetics of drugs in skin. Our preprint has gained broad interest also according to statistics: Research interest score in ResearchGate is higher than 96\% of research items published in 2023.

%We validate the practical relevance of the NMRlipids Databank here by successfully analysing how anisotropic diffusion of water and cholesterol flip-flop depend on membrane properties. Both of these processes are highly relevant for magnetic resonance imaging (MRI), pharmacokinetics, lipid trafficking, and regulating membrane properties of cellular membranes --- yet their timescales are beyond the scope of standard MD simulation investigations.

%The NMRlipids Databank emerges from the NMRlipids open collaboration project started by us in 2013.
%%with 314 citations in total) 
%%where all the content is openly accessible throughout the project. 
%The current manuscript demonstrates how MD simulation data can be shared by combining the open collaboration approach with an overlay databank structure, and describes how we take major steps in FAIRness by standardizing file formats, developing community guidelines, and creating incentives for sharing MD simulation data. Crucially, our success paves the way for corresponding smart guidelines and community-consensus best-practices for data sharing also in other fields than membrane MD simulations.
%Such developments will, we believe, strongly support the reproducibility and reuse of data that is becoming increasingly important in the age of big data and artificial intelligence --- as also pointed out by a recent editorial in \textit{Nature Methods} (\href{https://doi.org/10.1038/s41592-023-01865-4}{\tt DOI: 10.1038/s41592-023-01865-4}).

%We see that the NMRlipids Databank has both immediate and long term impact for a broad interdisciplinary audience of academic and industry researchers across
%%at different fields in
%life sciences: from biophysics and biochemistry to biology, biotechnology and biomedicine. Consequently, we trust our manuscript to fulfill the high standards for publication in %the \textit{Resource} section of
%\textit{Nature Methods}.\\

%We now leave it for your own consideration.

Sincerely yours, on behalf of the authors,

Dr. O. H. Samuli Ollila

\end{document}

### 2023-05-12 main.tex
Update

@@
 This is the case in biomolecular modeling, where vast amount of data is already available for proteins, lipids, nucleic acids, and carbohydrates, but tools that enable applications of these data for data-driven applications are not yet available~\cite{tiemann23}.
-Here we propose a cost-effective solution to make data scattered in various locations and formats accessible for data-driven and machine learning applications using an overlay databank.
-We demonstrate the practical relevance of such approach for understanding cellular membranes by incorporating available lipid bilayer simulations into the NMRlipids Databank.
+Here we propose a cost-effective solution to make data scattered in various locations and formats accessible for data-driven and machine learning applications: an overlay databank.
+We demonstrate the practical relevance of such approach for understanding cellular membranes by incorporating openly available lipid bilayer simulations into the NMRlipids Databank.
 Importantly, the basic principles of an overlay databank can be applied also to simulation data of other biomolecules that are becoming available in increasing amounts~\cite{tiemann23}, as well as to other any field where the development of AI-based tools is limited by the lack of data access and community best-practices --- such as assignment of NMR spectra~\cite{klukowski22}.
-Incentives to share data can be further accelerated by combining overlay databanks with open collaboration approach developed by us~\cite{botan15}.
+Incentives to share data can be further accelerated by combining overlay databanks with an open collaboration approach (see Ref.~\citenum{botan15}).
 
@@
 
+\newpage
+
 \section*{Acknowledgements}

### 2023-05-12 main.tex
Update

@@
 Here we propose a cost-effective solution to make data scattered in various locations and formats accessible for data-driven and machine learning applications: an overlay databank.
-We demonstrate the practical relevance of such approach for understanding cellular membranes by incorporating openly available lipid bilayer simulations into the NMRlipids Databank.
-Importantly, the basic principles of an overlay databank can be applied also to simulation data of other biomolecules that are becoming available in increasing amounts~\cite{tiemann23}, as well as to other any field where the development of AI-based tools is limited by the lack of data access and community best-practices --- such as assignment of NMR spectra~\cite{klukowski22}.
+We demonstrate the practical relevance of such approach for understanding cellular membranes by incorporating available lipid bilayer simulations into the NMRlipids Databank.
+Importantly, the basic principles of an overlay databank can be applied to simulation data of any other biomolecules (that are becoming available in increasing amounts~\cite{tiemann23}), as well as to any other field where the development of AI-based tools is limited by the lack of data access and community best-practices --- such as the assignment of NMR spectra~\cite{klukowski22}.
 Incentives to share data can be further accelerated by combining overlay databanks with an open collaboration approach (see Ref.~\citenum{botan15}).
@@
 However, biomembranes are often difficult to study experimentally, because they are complex mixtures of proteins and lipids in disordered fluid state with complicated phase behaviour at biological conditions.
-Molecular dynamics (MD) simulations can be used to model biomembranes in detail, but computational cost for simulating all possible biological membrane compositions would be substantial.
-For those reasons, data-driven and machine learning based models that predict biomembrane properties will benefit wide range of fields covering academia and industry, from cell membrane biology to lipid nanoparticle formulations.
+Molecular dynamics (MD) simulations can be used to model biomembranes in detail, but the computational cost for simulating all possible biological membrane compositions would be formidable.
+For those reasons, data-driven and machine-learning-based models that predict biomembrane properties will benefit wide range of fields covering academia and industry, from cell membrane biology to lipid nanoparticle formulations.
 %For those reasons, the detailed connections between complex lipid interactions and biological functions taking place in or around membranes remain poorly understood. 
@@
 Here we present the NMRlipids Databank --- a community-driven, open-for-all database featuring programmatic access to atom-resolution MD simulations of lipid bilayers.
-To demonstrate its advances to available approaches, we show how rare phenomena that are beyond the scope of standard MD simulation investigations can be analyzed from the Databank.
-To demonstrate the immediate relevance of NMRlipids databank in extending the scope of MD simulations to new fields, we  use data-driven approach to analyse how anisotropic diffusion of water depends on membrane properties.
-This benefits understanding of magnetic resonance imaging (MRI)~\cite{topgaard20} and pharmacokinetics~\cite{nitsche19} where MD simulations are rarely applied.
+To demonstrate its advantages over existing approaches, we show how information on rare phenomena that are beyond the scope of standard MD simulation investigations can be gleaned from the Databank.
+In addition, we demonstrate the immediate relevance of NMRlipids Databank in extending the scope of MD simulations to new fields: Using a data-driven approach, we are able to analyse how anisotropic diffusion of water depends on membrane properties; this benefits understanding in magnetic resonance imaging (MRI)~\cite{topgaard20} and pharmacokinetics~\cite{nitsche19}, where MD simulations have, until now, been rarely applied.
 %Wider adaptation of the NMRlipids Databank will open even more possibilities. 
 Furthermore, the Databank performs automatic quality evaluation of membrane simulations, which facilitates the selection of best-performing models for each given application and accelerates the development of simulation parameters and methodology.
-The overlay databank and open collaboration approaches facilitate programmatic access and collection of community contributed data also in other fields than membrane simulations.
-For example, substantial biomolecular MD simulation data is already available but in scattered locations and formats~\cite{tiemann23}.
-The technical advances presented here will immediately benefit in making this data programmatically accessible.
-Furthermore, the overlay databank configuration will benefit broad range of fields where access to training data is bottle neck for building AI based tools.
+Notably, the overlay-databank and open-collaboration approaches would facilitate the collecting of community-contributed data and providing programmatic access to them also in fields other than membrane simulations.
+For example, substantial biomolecular MD simulation data are already available but in scattered locations and formats~\cite{tiemann23}.
+The technical advances presented here will immediately benefit making these data programmatically accessible.
+Furthermore, the overlay databank configuration will benefit a broad range of fields where access to training data is a bottleneck for building AI-based tools.
 %NMRlipids databanksolutions presented as well as broader communities that 
@@
 Besides the data itself, also programmatic access is a necessary requirement for data-driven and machine learning applications.
-This is particularly challenging when field specific smart guidelines and community-consensus best-practices have not yet been defined, which is the case in biomolecular simulations~\cite{tiemann23}.
-The NMRlipids Databank demonstrates how these issues can be solved by the overlay databank design where the raw data is distributed to already publicly available decentralized locations, while the core of the Databank is composed only of the metadata stored in a version-controlled git repository with an open-access license.
-On the other hand, the open collaboration approach developed in the NMRlipids Project~\cite{botan15} creates incentives for sharing the data by offering authorship in published articles to the contributors.
-Advantages of such an approach are demonstrated here for membrane simulations, yet the concept can be applied to other biomolecules, %such as disordered proteins and membrane--protein systems,
-as well as in other fields where similar barriers hinder the AI revolution, such as the assignment of NMR spectra~\cite{klukowski22}.
+This is particularly challenging when field-specific smart guidelines and community-consensus best-practices have not yet been defined, which is the case in biomolecular simulations~\cite{tiemann23}.
+The NMRlipids Databank demonstrates how these issues can be solved by the overlay databank design, where the raw data are distributed to already publicly available decentralized locations, while the core of the databank is composed only of the metadata stored in a version-controlled git repository with an open-access license.
+On the other hand, the open-collaboration approach developed in the NMRlipids Project~\cite{botan15} creates incentives for sharing the data by offering authorship in published articles to the contributors.
+Advantages of such an approach are demonstrated here for membrane simulations, yet the concept can be applied to any other biomolecules, %such as disordered proteins and membrane--protein systems,
+as well as in any other field where similar barriers hinder the AI revolution, such as the assignment of NMR spectra~\cite{klukowski22}.
 
@@
 %The open programmatic access to unprecedented amounts of MD simulation data through the NMRlipids Databank-API enables a wide range of novel data-driven analyses and machine learning applications. 
-The analyses of cholesterol flip-flop events (Fig.~\ref{fig:flip-flops}) and water permeation through membranes (Fig.~\ref{fig:permeability}) demonstrate how a large amount of accessible simulation data in terms of quantity (\textit{e.g.}, simulation length and number of conformations) and content (\textit{e.g.}, lipid compositions and ion concentrations) enable analyses of rare phenomena that are beyond the current possibilities for a single research group.
+The analysis of cholesterol flip-flop events (Fig.~\ref{fig:flip-flops}) and water permeation through membranes (Fig.~\ref{fig:permeability}) demonstrate how a large amount of accessible simulation data in terms of quantity (\textit{e.g.}, simulation length and number of conformations) and content (\textit{e.g.}, lipid compositions and ion concentrations) enable analyses of rare phenomena that are beyond the current possibilities for a single research group.
 Such analyses also pave the way for applications of MD simulations in new fields, as demonstrated here by analysing an essential parameter in pharmacokinetic modeling and MRI imaging:\cite{nitsche19,topgaard20} the anisotropic diffusion of water in membrane systems (Fig.~\ref{fig:permeability}).
-These possibilities are particularly valuable for scientists who do not typically have access to large scale MD simulation data.
+These possibilities are particularly valuable for scientists who do not typically have access to large-scale MD simulation data.
 
@@
 Simultaneously, machine-learning-based models for predicting the behaviour of biomolecules and automatic approaches to parametrize models are emerging~\cite{jumper21,antila22b}.
-The resources delivered by the NMRlipids Databank will support the development in both of these directions.
+The resources delivered by the NMRlipids Databank will support developments in both of these directions.
 Automatic quality evaluation and ranking of simulations against experimental data enable the selection of best simulations for specific applications without laborious manual force field evaluation.
 This also streamlines automatic parametrization procedures for atomistic and coarse grained simulations by, for example, pinpointing typical failures of force fields and highlighting points of improvement.
-Such practises for fostering the accuracy of simulations are becoming increasingly important due the accumulation of small errors when complexity and size of simulated systems are increasing.
-Examples of impact of NMRlipids and other overlay Databanks in different disciplines are listed in Table~\ref{tab:applications}, yet the scope of applications is expected to further widen with increasing amount of publicly shared data.
+Such practises for fostering the accuracy of simulations are becoming increasingly important as small errors accumulate when complexity and size of simulated systems are increasing.
+Examples of impact of NMRlipids and other overlay databanks in different disciplines are listed in Table~\ref{tab:applications}, yet the scope of applications is expected to further widen with increasing amount of publicly shared data.
 

### 2023-10-09 main.tex
Updates from Overleaf

@@
 \begin{abstract}
+%TC:ignore
 Tools based on artificial intelligence (AI) are currently revolutionising many fields, yet their applications are often limited by the lack of suitable training data in programmatically accessible format.
@@
 To demonstrate how it unlocks possibilities beyond current MD simulation studies, we analyzed the NMRlipids Databank to reveal how anisotropic diffusion of water and cholesterol flip-flop rates depend on membrane properties.
-
-\end{abstract}
-\begin{document}
+%TC:endignore
+\end{abstract}\begin{document}
 
 \flushbottom
+%TC:ignore
 \maketitle
+%TC:endignore
 % * <john.hammersley@gmail.com> 2015-02-09T12:07:31.197Z:
@@
 For those reasons, data-driven and machine-learning-based models that predict biomembrane properties will benefit wide range of fields covering academia and industry, from cell membrane biology to lipid nanoparticle formulations.
-%For those reasons, the detailed connections between complex lipid interactions and biological functions taking place in or around membranes remain poorly understood. 
-%Molecular dynamics (MD) simulations have been particularly useful in understanding membrane systems, although their accuracy has often been compromised by artefacts such as the quality of model parameters~\cite{antila22b,gupta22}.
-Presently, the accuracy of models is becoming increasingly important as researches are progressing from simulations of individual molecules to simulating whole organelles or even cells using interdisciplinary approaches~\cite{johnson15,thornburg22,gupta22}.
-Such systems exhibit intricate emergent behaviour making inaccuracies more difficult to detect, and accumulation of even modest errors may have a dramatic impact on the conclusions drawn.
-
-%In contrast to experimental structural biology, where standard protocols to share and quality-evaluate resolved biomolecular structures are firmly established by the Protein Data Bank (PDB)~\cite{montelione13}, equivalent best practices are yet to be defined for MD simulations.
-The importance of such approaches is widely recognized~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} and data-sharing solutions are emerging for proteins in solution~\cite{meyer10,kamp10}, proteins in membranes~\cite{newport19,espigares20,leston22}, nucleic acids~\cite{hospital16}, nucleic acids and proteins~\cite{bekker20}, cyclodextrins~\cite{mixcoha16}, COVID-19--involved macromolecules~(\href{https://bioexcel-cv19.bsc.es}{bioexcel-cv19.bsc.es}).
-However, automatic quality evaluation of simulation data~\cite{meyer10,hospital16} and programmatic access are still rare.
-In particular, tools for automatic quality evaluation of membrane simulations, or training sets for machine learning models of membrane-containing systems, are not yet available.
-Recent advances using machine learning approaches utilizing publicly available databanks, for example to solve protein structures~\cite{jumper21}, emphasize the increasing importance of such resources.
 
 Here we present the NMRlipids Databank --- a community-driven, open-for-all database featuring programmatic access to atom-resolution MD simulations of lipid bilayers.
-To demonstrate its advantages over existing approaches, we show how information on rare phenomena that are beyond the scope of standard MD simulation investigations can be gleaned from the Databank.
+To demonstrate its advantages over existing approaches, we build a machine learning model that predicts membrane properties from its lipid composition and show how information on rare phenomena that are beyond the scope of standard MD simulation investigations can be gleaned from the Databank.
 In addition, we demonstrate the immediate relevance of NMRlipids Databank in extending the scope of MD simulations to new fields: Using a data-driven approach, we are able to analyse how anisotropic diffusion of water depends on membrane properties; this benefits understanding in magnetic resonance imaging (MRI)~\cite{topgaard20} and pharmacokinetics~\cite{nitsche19}, where MD simulations have, until now, been rarely applied.
-%Wider adaptation of the NMRlipids Databank will open even more possibilities. 
 Furthermore, the Databank performs automatic quality evaluation of membrane simulations, which facilitates the selection of best-performing models for each given application and accelerates the development of simulation parameters and methodology.
@@
 Furthermore, the overlay databank configuration will benefit a broad range of fields where access to training data is a bottleneck for building AI-based tools.
-%NMRlipids databanksolutions presented as well as broader communities that 
-%enables users to apply their data-driven approaches on the Databank, thus facilitating the creation of new tools for (and by) researchers in a .
-% to computational chemistry and machine learning.
-
-%While the NMRlipids Databank currently contains only lipid bilayer systems, its key elements that enable programmatic access to large-scale MD simulation data can be applied also for other molecules, such as disordered proteins or sugars.
-%%The powerful combination of an overlay databank structure and a community-wide open collaboration can be used to build databases that enable data-driven and machine learning applications also in other fields, where storage of raw data requires significant resources, best practices in the field are not yet defined, and incentives to share data do not exist---such as assignment of NMR spectra~\cite{klukowski22}.
-%The powerful combination of an overlay databank structure and the community-wide open collaboration  described here is potentially useful for building databases that enable data-driven and machine learning applications also in other fields.
-This is especially true for research activities where storage of raw data requires significant resources, best practices in the field are not yet defined, and incentives to share data do not exist---such as assignment of NMR spectra~\cite{klukowski22}.
 
@@
 The overlay structure of the NMRlipids Databank, illustrated in Fig.~\ref{fig:overlay}A, is designed to enable efficient upcycling of MD simulations for data-driven and machine learning applications with minimal investment on new infrastructure.
-Raw simulation data in the {\it Data layer} can be stored in any publicly available location with long term stability
-%figshare (\href{https://figshare.com}{figshare.com}), or NIRD research data archive (\href{https://archive.norstore.no}{archive.norstore.no})---
-and with permanent links to the data, such as digital object identifiers (DOIs), such as Zenodo (\href{https://zenodo.org}{zenodo.org}). 
-%or uniform resource names (URNs). 
+Raw simulation data in the {\it Data layer} can be stored in any publicly available location with long term stability and with permanent links to the data, such as digital object identifiers (DOIs), such as Zenodo (\href{https://zenodo.org}{zenodo.org}).
 The {\it Databank layer} (\href{https://github.com/NMRlipids/Databank}{github.com/NMRlipids/Databank}) is the core of the Databank containing all the relevant information about the simulations: links to the raw data, relevant metadata describing the systems, universal naming conventions for lipids and their atoms, quality evaluation of simulations against experimental data, and the computer programs to create the entries and to analyse the five basic properties extracted from all simulations (area per lipid, C--H bond order parameters, X-ray scattering form factors, membrane thickness, and equilibration times of principal components).
@@
     \centering
-    \includegraphics[width=\linewidth]{Figures/overlay3.pdf}
+    \includegraphics[width=\linewidth]{Figures/overlay4.pdf}
     \caption{Overview of the NMRlipids Databank. 
@@
 \subsection{NMRlipids Databank-API: programmatic access to the MD simulation data}\label{section:access}
-The NMRlipids Databank-API, available at \href{https://github.com/NMRlipids/Databank}{github.com/NMRlipids/Databank}, provides programmatic access to all simulation data in the NMRlipids Databank through application programming interface (API).
+The NMRlipids Databank-API provides programmatic access to all simulation data in the NMRlipids Databank through application programming interface (API).
 This enables wide range of novel data-driven applications---from construction of machine learning models that predict membrane properties, to automatic analysis of virtually any property across all simulations in the Databank.
 The flowchart in Fig.~\ref{fig:overlay}D illustrates the practical implementation of such an analysis.
-After cloning the databank repository to a local computer, raw data of each simulation can be accessed by iterating through the README.yaml files in the {\it Databank layer} that contain links pointing to the locations of simulation trajectories and other relevant files.
-Simulations can then be automatically analyzed with the help of the README.yaml and the mapping files that associate the specific naming conventions in each simulation with the universal molecule and atom names used by the Databank.
-Finally, the analysis results are stored using the same structure as in the {\it Databank layer}.
-Practical examples of codes that utilize the NMRlipids Databank-API, as well as a template for a new analysis, can be found from locations listed in Table~\ref{tab:codes}.
+After cloning the databank repository to a local computer, raw data of each simulation can be accessed and analyzed with the help of functions delivered by the NMRlipids Databank-API.
+%by iterating through the README.yaml files in the {\it Databank layer} that contain links pointing to the locations of simulation trajectories and other relevant files.
+Simulations can then be automatically analyzed with the help of the README.yaml and the
+Analyses over simulations with different naming convetions can be automatically performed with the help of mapping files that associate the specific naming conventions in each simulation with the universal molecule and atom names used by the Databank.
+Finally, the analysis results can be stored using the same structure as in the {\it Databank layer}.
+Documentation of the NMRlipids Databank-API and further instructions are available at \href{https://nmrlipids.github.io/}{nmrlipids.github.io}, and template for new user-defined analyses at \href{https://github.com/NMRLipids/databank-template}{github.com/NMRLipids/databank-template}.
+%Practical examples of codes that utilize the NMRlipids Databank-API, as well as a template for a new analysis, can be found from locations listed in Table~\ref{tab:codes}.
 
@@
 \subsubsection{Selecting simulation parameters using NMRlipids Databank: Best models for most abundant neutral membrane lipids}
-%Quality of membrane simulations with different force fields have been evaluated against experimental data during parameterization and in separate comparison studies~\cite{botan15,ollila16,catte16,pluhackova16,perez17,leonard19,??}, but the lack of universal quality measure for membrane simulations complicates the estimations of reliability of simulations.
-%
-%For example, OPLS3e parameters overcome CHARMM36 in structural quality for POPC but predicts overestimated ion binding~\cite{kurki22}.
-On the other hand, CHARMM36 gives the best description for lipid headgroup conformational ensembles~\cite{bacle21}, while GROMOS-CKP parameters best capture the membrane packing in POPS bilayers~\cite{antila22b}.
-%
 Molecular dynamics (MD) simulations have been particularly useful in understanding membrane systems, although their accuracy has often been compromised by artefacts such as the quality of model parameters~\cite{antila22b,gupta22}.
@@
 This can be done, for example, against the C--H bond order parameters from NMR spectroscopy~\cite{bacle21,Wur23} and the form factors from X-ray scattering~\cite{ollila16}, although it requires comparisons between large number of simulations, which is laborious even with collaborative approaches~\cite{botan15,catte16,antila19,bacle21}.
-To streamline this process, we have defined quantitative quality measures that enable automatic ranking of lipid bilayer simulations based on their quality against experiments.
-Conformational ensembles of individual lipid molecules are evaluated in the NMRlipids Databank by first calculating the probabilities for each C--H bond order parameter to locate within experimental error and then averaging the possibilities over different lipid segments ($P^{\mathrm{hg}}$, $P^{\mathrm{sn1}}$,$P^{\mathrm{sn2}}$, and $P^{\mathrm{total}}$).
-Furthermore, the quality against X-ray scattering experiments ($FF_q$) is estimated as the difference in the experimental and simulated locations of the first form factor minimum.
-These measures can be used to evaluate membrane dimensions, as the form factor minima locations and acyl chain C--H bond order parameters correlate with the membrane lateral packing and thickness (Figs.~\ref{fig:quality}G,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}).
-In addition, the ergodicity of conformational sampling of lipids is estimated by calculating $\tau_\mathrm{rel}$, the convergence time of the slowest principal component divided by the simulation length.
-Here we demonstrate how the automatic simulation-quality evaluation and the NMRlipids Databank-API enable rapid selection of the best models for simulations of membranes with the two most biologically-abundant neutral membrane lipids~\cite{vanmeer08}: POPC and POPE.
 
-As Fig.~\ref{fig:quality}A illustrates, predictions for the lateral packing of POPC and POPE membranes in terms of area per lipid, one of the most important parameters used to characterize cellular membranes, diverge between different force fields.
-To select the most realistic currently available force fields to describe POPC--POPE membranes, we first order simulations in the Databank according to their C--H bond order parameter quality ranking (Fig.~\ref{fig:top50simulations}).
-From these results, we then pick the force fields occurring in Fig.~\ref{fig:quality}A (that is: force fields for which both POPC and POPE simulations exist in the Databank) and rank them according to the quality of the \textit{sn}-1 chain of POPC (Fig.~\ref{fig:quality}B) and of POPE (Fig.~\ref{fig:quality}C).
-Simulations with $\tau_\mathrm{rel}$ clearly above one (larger than 1.3) are discarded.
-Because the membrane packing correlates with the average order parameter of the \textit{sn}-1 chain (Fig.~\ref{fig:quality}G), rankings in Figs.~\ref{fig:quality}B and C can be used to select the simulations giving the most realistic results in Fig.~\ref{fig:quality}A.
-(Figs.~\ref{fig:quality}E--F and~\ref{fig:POPC_POPE_dataSI} also show direct comparisons with the experimental data for the most relevant simulations; for reference, Fig.~\ref{fig:quality}D shows the overall highest-ranked simulation: POPC bilayer with OPLS3e parameters.) Based on these rankings, Lipid17 and Slipids simulations give the most realistic predictions for a POPC membrane, while simulations with CHARMM36 and GROMOS-CKP parameters predict overly packed bilayers (overestimated order in Fig.~\ref{fig:POPC_POPE_dataSI}).
-For POPE, on the other hand, GROMOS-CKP and Slipids give the most realistic results, while CHARMM36 and Lipid17 predict membranes that are too packed.
+Here we streamlined this process by defining quantitative quality measures 
+for conformational ensembles of individual lipid molecules and membrane dimensions using C--H bond order parameters from NMR and X-ray scattering form factors~\cite{ollila16}.
+These measures enable automatic ranking of lipid bilayer simulations based on their quality against experiments.
+Qualities of order parameters were evaluated by first calculating the probabilities for each C--H bond order parameter to locate within experimental error, and then averaging the possibilities over different lipid segments ($P^{\mathrm{hg}}$, $P^{\mathrm{sn1}}$,$P^{\mathrm{sn2}}$, and $P^{\mathrm{total}}$).
+Qualities against X-ray scattering experiments ($FF_q$) were estimated as the difference in the experimental and simulated locations of the first form factor minimum.
+These measures are good proxies for membrane properties
+%can be used to evaluate membrane dimensions, as the form factor minima locations and acyl chain C--H bond order parameters 
+because they correlate with the membrane lateral packing and thickness (Figs.~\ref{fig:quality}G,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}).
+Ergodicity of conformational sampling of lipids was estimated by calculating $\tau_\mathrm{rel}$, the convergence time of the slowest principal component divided by the simulation length.
+
+Fig.~\ref{fig:quality} demonstrates how the automatic simulation-quality evaluation and the NMRlipids Databank-API enable rapid selection of the best models for membrane simulations.
+Fig.~\ref{fig:quality}A illustrates that predictions for the lateral packing of membranes composed of two most biologically-abundant neutral membrane lipids, POPC and POPE~\cite{vanmeer08},
+%for area per lipid, one of the most important parameters used to characterize cellular membranes, 
+diverge between different force fields.
+To find the most realistic parameters to simulate membranes with these lipids, we first ranked all simulations based on order parameter quality (Fig.~\ref{fig:top50simulations}), then picked force fields that occurr in Fig.~\ref{fig:quality}A (that is: force fields for both POPC and POPE in the Databank), and then ranked them according to the quality of the \textit{sn}-1 chain of POPC (Fig.~\ref{fig:quality}B) and of POPE (Fig.~\ref{fig:quality}C).
+Simulations with $\tau_\mathrm{rel}$ clearly above one (larger than 1.3) were discarded in this analysis.
+Because the average \textit{sn}-1 chain order parameter is a good proxy for the membrane packing (Fig.~\ref{fig:quality}G), rankings in Figs.~\ref{fig:quality}B and C can be used to select the simulations giving the most realistic results in Fig.~\ref{fig:quality}A.
+Based on this, Lipid17 and Slipids simulations are most realistic for a POPC membrane, while CHARMM36 and GROMOS-CKP simulations predict overly packed bilayers (overestimated order in Fig.~\ref{fig:POPC_POPE_dataSI}).
+For POPE, on the other hand, GROMOS-CKP and Slipids are most realistic, while CHARMM36 and Lipid17 predict too packed membranes.
 In conclusion, the quality evaluation based on the NMRlipids Databank suggests that the Slipids parameters are the best currently available choice for simulations with PC and PE lipids, at least for applications where membrane packing is relevant.
+Also direct comparisons with the experimental data for the most relevant simulations are shown in Figs.~\ref{fig:quality}E--F and~\ref{fig:POPC_POPE_dataSI} a.
+Fig.~\ref{fig:quality}D shows the overall highest-ranked simulation, POPC bilayer with OPLS3e parameters, for the reference.
 
@@
 \end{figure}
+
+\subsubsection{Using NMRlipids databank as a training set for machine learning applications: Predicting multi-component membrane properties}
+
+To demonstrate the usage of the NMRlipids databank to construct machine learning (ML) models that predict membrane properties, we trained a model that predicts area per lipids and thicknesses of membranes with diverse compositions.
+First we used randomly selected 80\% of the databank simulations to choose the hyperparameters for, and optimize, a set of machine learning models with the goal of predicting the area per lipid from the membrane composition.
+
+ After the parameter optimization, we tested predictions both against the remaining 20\% of the databank and against data from different ML models against area per lipids reported from simulations of membranes containing mixtures of POPC, POPE, POPS, PI, sphingomyelin lipids and cholesterol~\cite{shahane19,kumar21,oliveira22} not included in the databank.
+Linear regression and Ridge models gave the best correlation with the literature data in this test (predictions  from the linear regression model are shown in Fig.~\ref{fig:LinearRegressionTest}).
+The linear regression model was selected for further studies due to its simplicity.
+\begin{figure}
+    \centering
+    \includegraphics[width=88mm]{Figures/LinearRegressionTest.pdf}
+    \caption{Predictions of area per lipids of POPC, POPE, POPS, PI, sphingomyelin lipids and cholesterol from linear regression model against literature data from simulations (green~\cite{shahane19}, blue~\cite{kumar21}, and red~\cite{oliveira22}).}
+    \label{fig:LinearRegressionTest}
+\end{figure}
+
+To demonstrate the usefulness of the constructed models for understanding multi-component membrane properties, we predicted later packing (ares per lipids) and membrane thicknesses of common biological membranes based on their lipid compositions reported in the literature, see table~\ref{tab:MLpredictions}.
+The model predicts substantial 50\% difference in area per lipid and 0.5\,nm in thickness between most densely (influenza virus) and loosely (mitochondria) packed membranes.
+Because these parameters correlate with many biologically relevant functions of membranes, such as permeation, cholesterol flip-flops (see next sections), and interactions with proteins~\cite{phillips09}, our results demonstrate that NMRlipids databank can be used to give valuable insights on biologically relevant properties of complex biological membranes.
+Most importantly, the delivered programmatic access to increasing amount of MD simulation data enables training of machine learning models that predict various membrane properties for all.
+
+    \begin{table}[htb]
+    \caption{Area per lipids (APL) and membrane thicknesses predicted by the linear regression model trained using the NMRlipids databank for membrane compositions corresponding different biological membranes.
+Systems are sorted in the order of increasing membrane packing.
+Compositions of different membranes are estimated based on references given in composition column.}
+        \centering
+        \begin{tabular}{cccc}
+             Membrane&  Composition [lipid(\%-fraction)]&  APL[\AA$^2$] & Thickness [nm$^2$]\\
+             \hline
+             Mitochondria & POPC(37) : POPE(31) : PI(6) : CL(22) : 
+             CHOL(4)~\cite{vanmeer08,escriba15,casares19} & 74 (66)$^*$ & 4.3 \\
+             Bacterial & POPC(20) : POPE(35) : POPG(35) : CL(5) : CHOL(5)~\cite{chwastek20}  & 62 & 4.2 \\
+             ER &  POPC(54) : POPE(20) : POPS(4) : PI(11) : SM(4) : CHOL(8)~\cite{vanmeer08,escriba15,casares19} & 56 & 4.5 \\
+             Golgi & POPC(36) : POPE(21) : POPS(6) : PI(12) : SM(7) : CHOL(18)~\cite{vanmeer08,escriba15,casares19} & 51 & 4.7 \\
+             Plasma & POPC(23) : POPE(11) : POPS(8) : PI(7) : SM(17) : CHOL(34)~\cite{vanmeer08,escriba15,casares19} & 45 & 4.9 \\
+             Synaptic & POPC(28) : POPE(20) : POPS(9) : PI(2) : SM(4) : CHOL(37)~\cite{binotti21}  &  45 & 4.8 \\
+             Influenza & POPC(5) : POPE(32) : POPS(15) : SM(9) : CHOL(40)~\cite{gerl12,ivanova15} & 42 & 4.8 \\
+        \end{tabular}
+        \label{tab:MLpredictions}
+
+        $^*$ Value in parenthesis is the area per acyl chain taking into account that cardiolipin has four chains per molecule while other lipids have two.
+    \end{table}
 
@@
 \subsubsection{Extending the scope of MD simulations to new fields using NMRlipids Databank: Water diffusion anisotropy in membrane systems}
-The anisotropy of water diffusion in directions parallel and perpendicular to membranes can be related to translocation of drugs through biological material, particularly in the skin \cite{hansen13,wen18,nitsche19,roberts21}, and to signal formation in diffusion tensor MRI imaging~\cite{topgaard20}.
+The anisotropic diffusion of water and hydrophilic molecules in directions parallel and perpendicular to membranes is important parameter in models describing the translocation of drugs through biological material, particularly in the skin \cite{hansen13,wen18,nitsche19,roberts21}.
+Water anisotropic diffusion plays a role also in the signal formation in diffusion tensor MRI imaging~\cite{topgaard20}.
 MD simulations are rarely used to analyze the anisotropic diffusion of water, since only a few membrane permeation events of water are typically observed in a single MD simulation trajectory~\cite{venable19,camilo2022}, thereby making the collection of a sufficient amount of data challenging.
@@
 Permeation of water through bilayers depends on membrane properties also according to previous studies, but there is no established consensus on whether the area per lipid~\cite{nagle08} or bilayer thickness~\cite{frallicciardi22} is the main parameter determining the permeability.
-%is previously reported from simulations and experiments for some systems, but not for all~\cite{mathai07,venable19,frallicciardi22}. 
 Our analysis over the NMRlipids Databank, containing significantly more data than what was available in previous studies, suggest non-linear dependencies on both of these parameters.
-%, yet the linear correlation would be a good approximation for thicknesses $\gtrsim$\,3.9\,nm and area per lipids $\lesssim$\,69\,\AA{}$^2$ (insets in Figs.~\ref{fig:permeability}C--D).
-%within the range of experimentally reported values  permeability which is 1-2 orders of magnitude slower than osmotic permeability \cite{jansen95,??}. 
 Clear dependencies of permeability on hydration level or the fraction of charged lipids, cholesterol, or POPE in the membrane were not observed (Fig.~\ref{fig:permeationSI}).
@@
 
-%The Discussion should be succinct and must not contain subheadings.
-
 Sharing of biomolecular MD simulation and other data is becoming increasingly important in the age of big data and artificial intelligence~\cite{sever23,tiemann23}.
-%The main barriers for such solutions have been the required commitment for the long term support of hardware and software, and the lack of incentives for researchers to share the data. 
 Besides the data itself, also programmatic access is a necessary requirement for data-driven and machine learning applications.
@@
 On the other hand, the open-collaboration approach developed in the NMRlipids Project~\cite{botan15} creates incentives for sharing the data by offering authorship in published articles to the contributors.
-Advantages of such an approach are demonstrated here for membrane simulations, yet the concept can be applied to any other biomolecules, %such as disordered proteins and membrane--protein systems,
-as well as in any other field where similar barriers hinder the AI revolution, such as the assignment of NMR spectra~\cite{klukowski22}.
+Advantages of such an approach are demonstrated here for membrane simulations, yet the concept can be applied to any other biomolecules, as well as in any other field where similar barriers hinder the AI revolution, such as the assignment of NMR spectra~\cite{klukowski22}.
 
@@
 Such applications could be analogous to AlphaFold \cite{jumper21} and other tools~\cite{baek21,lin23} that predict protein structures from their sequence using artificial intelligence.
-%The open programmatic access to unprecedented amounts of MD simulation data through the NMRlipids Databank-API enables a wide range of novel data-driven analyses and machine learning applications. 
 The analysis of cholesterol flip-flop events (Fig.~\ref{fig:flip-flops}) and water permeation through membranes (Fig.~\ref{fig:permeability}) demonstrate how a large amount of accessible simulation data in terms of quantity (\textit{e.g.}, simulation length and number of conformations) and content (\textit{e.g.}, lipid compositions and ion concentrations) enable analyses of rare phenomena that are beyond the current possibilities for a single research group.
@@
 
-%\begin{table}[t]
-%    \centering
-%    \begin{tabular}{p{5.0cm}  p{5.0cm}  p{4.0cm}}
-%    Type of application     & Practical examples & Target group \\
-%    \hline
-%    Analyses of rare phenomena               & Lipid flip-flops, water permeation (Figs.~\ref{fig:flip-flops} and~\ref{fig:permeability}) & Membrane scientists \\
-%    \\
-%    Correlations between membrane properties & 
-%    Membrane structural properties, water dynamics (Figs.~\ref{fig:quality} and~\ref{fig:permeability}) %& 
-%    Membrane scientists \\
-%    \\
-%    Applications that are outside the typical scope of MD simulations & 
-%    Anisotropic water diffusion for pharmacokinetics and MRI imaging applications %(Fig.~\ref{fig:permeability}) & 
-%    Scientists in fields where MD simulations are not usually applied \\
-%    \\
-%    Selection of the best simulation model for a specific application & 
-%    Best model for membranes with PC and PE lipids (Fig.~\ref{fig:quality}), lipid headgroup %conformations~\cite{bacle21}, 
-%    packing of PS~\cite{antila22b} and PE (Fig.~\ref{fig:quality}) containing membranes &
-%    Scientists using MD simulations \\
-%    \\
-%    Guidance for force field development & 
-%    Improvements in ion binding to lipids~\cite{melcr18,melcr20} and lipid headgroup conformational %ensembles~\cite{yu21,dickson22,grote20} &
-%    Scientists developing parameters for MD simulations \\
-%    \\
-%    Training and target data for the development of coarse grained models & 
-%    Optimizing parameters of coarse grained models against NMRlipids Databank, extracting continuum %parameters for membranes &
-%    Scientists developing and using coarse grained MD simulations \\
-%   \\
-%    Training set for machine learning applicatons &
-%    Programmatic access to the data and results enables training of machine learning models for various %applications, such as predictions of membrane properties from composition & Scientists building and %using machine learning applications for biomolecules
-%    \end{tabular}
-%    \caption{Examples on applications of the NMRlipids Databank.}
-%    \label{tab:applications}
-%\end{table}
-
-%Potential benefits of sharing MD simulation data are widely recognized by different stakeholders from scientists~\cite{feig99,tai04,silva06,abraham19,hildebrand19,hospital20,abriata20,espigares20} to funders and journal staff, but data sharing with efficient tools for upcycling have been rare.
-The main barriers for such solutions have been the required commitment for the long term support of hardware and software, and the lack of incentives for researchers to share the data.
-The first issue is solved in the NMRlipids Databank by the overlay design, where the raw data is distributed to already publicly available decentralized locations, while the core of the Databank is composed only of the metadata stored in a version-controlled git repository with an open-access license.
-On the other hand, the open collaboration approach developed in the NMRlipids Project~\cite{botan15} creates incentives for sharing the data by offering authorship in published articles to the contributors.
-Advantages of such an approach are demonstrated here for membrane simulations, but the concept can be applied not only to other biomolecules, such as disordered proteins and membrane--protein systems, but also in other fields where similar barriers hinder the machine learning revolution, such as the assignment of NMR spectra~\cite{klukowski22}.
-
 \newpage
 
+%TC:ignore
 \section{Methods}
-
-%Topical subheadings are allowed. Authors must ensure that their Methods section includes adequate experimental and characterization data necessary for others in the field to reproduce their work.
 
@@
 The README.yaml files in these folders contain access to all information that is needed for further analysis of simulations, such as links to the raw data and associations with the universal molecule and atom names.
-The content of these files is described in detail in Table~\ref{tab:READMEkeys} in the supplementary information.
+The content of these files is described in detail in Table~\ref{tab:READMEkeys} in the supplementary information and in the NMRlipids databank documentation \href{https://nmrlipids.github.io/}{nmrlipids.github.io/}.
 Results from analyses of basic membrane properties (area per lipid, thickness, C--H bond order parameters, X-ray scattering form factors, and relaxation of principal components) are stored in the same folders as the README.yaml files.
@@
 To enable automatic analyses over all the simulations in the NMRlipids Databank, we have defined universal naming conventions for the molecules and atoms therein.
-The universal abbreviations used in the NMRlipids Databank for each molecule are listed in Table~\ref{tab:abbreviations} in the supplementary information.
-The atom names used in simulation trajectories are connected to the universal atom names using the mapping files defined in the NMRlipids Project (\url{https://nmrlipids.blogspot.com/2022/04/new-yaml-format-of-mapping-files.html}).
+The universal abbreviations used in the NMRlipids Databank for each molecule are listed in the NMRlipids databank documentation \href{https://nmrlipids.github.io/moleculesAndMapping.html}{nmrlipids.github.io/moleculesAndMapping.html}.
+%Table~\ref{tab:abbreviations} in the supplementary information. 
+The atom names used in simulation trajectories are connected to the universal atom names using the mapping files defined in the NMRlipids Project (\url{https://nmrlipids.github.io/moleculesAndMapping.html}).
 These files are located at \texttt{/Scripts/BuildDatabank/mapping\_files} in the NMRlipids Databank repository.
@@
 The NMRlipids Databank is open for additions of simulation data by anyone.
-The list of information that a contributor has to deliver is given in Table~\ref{tab:READMEkeys}.
-The rest of the information to be stored in the README.yaml files, also listed in Table~\ref{tab:READMEkeys}, will be automatically extracted using the \texttt{/Scripts/BuildDatabank/AddData.py} script.
-In practise, the manually entered data is first stored into an info.yaml file that is then added to the \texttt{/Scripts/BuildDatabank/info\_files} folder trough a git pull request.
+In practise, the required information is first manually entered into an info.yaml file that is then added into the \texttt{/Scripts/BuildDatabank/info\_files} folder trough a git pull request.
+Rest of the information to be stored in the README.yaml files will be then automatically extracted using the \texttt{/Scripts/BuildDatabank/AddData.py} script.
+The required manually entered and automatically extracted information are described in detail in Table~\ref{tab:READMEkeys} in the supplementary information and in the NMRlipids databank documentation \href{https://nmrlipids.github.io/}{nmrlipids.github.io/}.
+The documentation includes also detailed instructions to add data.
 To avoid ineligible entries and minimize human errors, the pull requests are monitored before the acceptance and generation of the README.yaml files.
@@
 Similarly to simulations, each experimental data set has a README.yaml file containing all the relevant information about the experiment.
-The keys and their descriptions for the experimental data are given in Table~\ref{tab:READMEkeysEXP}.
-The NMR data currently in the NMRlipids Databank are taken from Refs.~\citenum{scherer87,ferreira13,antila19,melcr20,bacle21} and the X-ray scattering data from Refs.~\citenum{Kucerka08a,kucerka11,pan12b,pan14,kucerka15,NMRlipidsIII}.
+The keys and their descriptions for the experimental data, as well as detailed instructions to add the date, are given in Table~\ref{tab:READMEkeysEXP} and in the NMRlipids databank documentation \href{https://nmrlipids.github.io/}{nmrlipids.github.io}.
+The NMR data currently in the NMRlipids Databank are taken from Refs.~\citenum{scherer87,ferreira13,antila19,melcr20,bacle21,hanashima19,doktorova20} and the X-ray scattering data from Refs.~\citenum{Kucerka08a,kucerka11,pan12b,pan14,kucerka15,NMRlipidsIII,doktorova20}.
 In addition, previously unpublished NMR data for POPE, POPG, and DOPC was acquired as described in the supplementary information (Figs.~\ref{POPEexp1}--\ref{fig:DOPCexp}) and contributed to the Databank.
@@
 3) performs the desired analysis on a local computer utilising the universal naming conventions for molecules and atoms defined in the README.yaml and mapping files.
-This general procedure is illustrated in Fig.~\ref{fig:overlay}D, and practical examples of codes that perform such analyses are listed in Tables~\ref{tab:codes} and~\ref{tab:codesApplication}.
+This general procedure is illustrated in Fig.~\ref{fig:overlay}D, and a template for user-defined analyses is available at at \href{https://github.com/NMRLipids/databank-template}{github.com/NMRLipids/databank-template}.
+Further practical examples of codes  performing such analyses are listed in Tables~\ref{tab:codes} and~\ref{tab:codesApplication}.
 The equilibration period given by the user (TIMELEFTOUT in Table~\ref{tab:READMEkeys}) is discarded from the trajectories in all analysis codes.
+For further details, see the NMRlipids databank documentation \href{https://nmrlipids.github.io/}{nmrlipids.github.io}.
 
@@
 To estimate how well conformational ensembles of lipids are converged in trajectories, the Principal Component Analysis (PCA) following the PCALipids protocol was used \cite{buslaev16,buslaev2020principal}.
-%, the lipids of a specific type were first selected, and trajectories of individual lipids were concatenated. 
 To this end, each lipid configuration was first aligned to the average structure of that lipid type, and PCA analysis was then applied on the Cartesian coordinates of all heavy atoms of the lipid.
@@
 Rigid molecules that do not exhibit significant conformational fluctuations, such as sterols, were excluded from the analysis.
-%The suggested way to estimate the characteristic timescales needed to adequately sample the configurational space of individual lipids is to calculate the distribution convergence of the trajectories projected on the first PC\cite{buslaev16}.
 In practise, the distribution convergence times were calculated utilising its linear dependence on autocorrelation decay times,
@@
 The empirical coefficient $k = 49$ was calculated based on the analysis of 8 trajectories with the length of more than 200\,ns, including simulations of POPC, POPS, POPE, POPG, and DPPC with the CHARMM36 force field.
-%All simulations, used for $k$ calculation, were at least 200 ns long.
 Because the coefficient $k$ does not depend on the force field~\cite{buslaev16}, the value determined from these CHARMM36 simulations can be used for all simulations in the Databank.
@@
 
+\subsection{Training machine learning model to predict properties of multi-component membranes}
+    Different ML approaches were tested to find the best model to predict the area per lipids of multi-component membranes  from their lipid composition.
+To his end, we tested linear, Lasso, Ridge, elastic net, decision tree, random forest, multi-layer perceptron, k-nearest neighbour, gradient boost, AdaBoost, and XGBoost regressors.
+The models were trained to predict the area per lipid for a membrane with the given molar fractions of membrane lipids.
+Most abundant lipids in the databank, POPC, POPE, POPG, POPS and cholesterol, were considered explicitly, while SM, cardiolipins and PI lipids with  more sparse data were all grouped together based on their headgroups independently on acyl chain content.
+Salt concentration, temperature and other conditions were not considered as parameters at this stage.
+
+To find the optimal model for this task, we first randomly divided simulations into training (80\% of the simulations) and test (20\% of the simulations) sets.
+Then, we used the training set with 5-fold cross validation to conduct a grid search for hyperparameters for a set of ML regression models.
+The best performing hyperparameters were then used to fit each model to the training set.
+The performance of the models was then tested on the remaining 20\% of the data extracted from the databank, and additionally,
+%selected optimal parameters for each model based on their performance in predicting area per lipids for simulations in the test set. 
+%After finding optimal parameters for each ML model, each model was trained again using all simulations in the databank. 
+%The performances of newly trained models were then tested 
+by predicting area per lipids from multi-component membrane simulations reported in the literature but not included in the databank~\cite{shahane19,kumar21,oliveira22}.
+Due to its  simplicity and good performance in these tests, we selected to use linear regression model for further studies.
+The  linear regression model predicting membrane thicknesses was trained in similarly.
+The ML models and analysis were implemented with Python using scikit-learn which was supplemented with the xgboost library to include extreme gradient boosted decision tree regressor to the set of tested model.
+
 \subsection{Calculation of lipid flip-flops}
@@
 If the lipid consists of several residues which is the case in some AMBER force fields, then a third column is needed which contains the name of the residue to which each atom belongs to.
+
+%TC:endignore
 

### 2023-10-09 SI.tex
Update

@@
 
-\title{SUPPLEMENTARY INFORMATION: NMRlipids Databank makes data-driven analysis of biomembrane properties accessible for all}
+\title{SUPPLEMENTARY INFORMATION: Overlay databank unlocks data-driven analyses of biomolecules for all}
 
@@
          calc\_thickness.py & Calculate membrane thickness & README.yaml files & thickness.json \\
-         template.py $\qquad$  template.ipynb & Minimal code to loop over simulations that can be used as a template for new analyses & README.yaml files & \\
+         %template.py $\qquad$  template.ipynb & Minimal code to loop over simulations that can be used as a template for new analyses & README.yaml files & \\
 
@@
 
+\begin{comment}
+
 \begin{table}[p]
@@
 \end{table}
+\end{comment}
 

### 2023-10-10 CoverLetter4.tex

--- 2023-04-27 CoverLetter.tex
@@
 Please find enclosed our manuscript titled
-{\it NMRlipids Databank makes data-driven analysis of biomembrane properties accessible for all}, which we would like to submit for your consideration to be published as a Resource in \textit{Nature Methods}.
+{\it Overlay databank unlocks data-driven analyses of biomolecules for all}, which we would like to submit for your consideration to be published as a Resource in \textit{ Nature Computational Science}.
 
@@
 
-By providing rapid open access to simulation data and automaticly  assessing
-%ranking of simulations based on 
-their quality against experiments, the NMRlipids Databank holds immediate practical relevance for the broad basic research community studying membranes in life sciences --- ranging from molecular cell biology to biomedical applications such as lipid nanoparticles.
-Even more importantly, the programmable interface for flexible implementation of data-driven and machine learning applications will advance new biological applications of MD simulation data.
-We validate the practical relevance of the NMRlipids Databank here by successfully analysing how anisotropic diffusion of water and cholesterol flip-flop depend on membrane properties.
+NMRlipids Databank is a groundbreaking tool that provides rapid open access to simulation data and automatic assessment of their quality against experiments.
+The programmable interface for flexible implementation of data-driven and machine learning applications will advance scientific research in wide range of fields
+%new biological applications of MD simulation data. 
+%, the  holds immediate practical relevance for the broad basic research community studying membranes in life sciences 
+--- ranging from molecular cell biology to biomedical applications such as lipid nanoparticles. 
+%Even more importantly, 
+%We validate the practical relevance of the NMRlipids Databank on functional and mechanistic understanding of biological process 
+%how molecular components in a work together
+We demonstrate the power of NMRlipids databank to find novel insights and solve challenging real-world problems
+by successfully analysing how anisotropic diffusion of water and cholesterol flip-flop depend on membrane properties.
 Both of these processes are highly relevant for magnetic resonance imaging (MRI), pharmacokinetics, lipid trafficking, and regulating membrane properties of cellular membranes --- yet their timescales are beyond the scope of standard MD simulation investigations.
+
+%The journal publishes both fundamental and applied research, from groundbreaking algorithms, tools and frameworks that notably help to advance scientific research, to methodologies that use computing capabilities in novel ways to find new insights and solve challenging real-world problems.
 
@@
 
-We see that the NMRlipids Databank has both immediate and long term impact for a broad interdisciplinary audience of academic and industry researchers across
-%at different fields in
-life sciences: from biophysics and biochemistry to biology, biotechnology and biomedicine.
-Consequently, we trust our manuscript to fulfill the high standards for publication in %the \textit{Resource} section of
-\textit{Nature Methods}.\\
+We see that the NMRlipids Databank has both immediate and long term impact for a broad interdisciplinary audience of academic and industry researchers across different disciplines: from biophysics and biochemistry to biology, biotechnology and biomedicine.
+Feedback from our first submissions to \textit{Nature Methods} and \textit{ Nature Structural \& Molecular Biology} states that the manuscript is of interest on these fields of research, yet the level of biological insight was not sufficient for these journals.
+Therefore, we believe that \textit{ Nature Computational Science} is the best forum for our manuscript and trust our manuscript to fulfill its high standards for publication.
+\\
 

### 2023-10-10 CoverLetter5.tex

--- 2023-10-10 CoverLetter4.tex
@@
 Please find enclosed our manuscript titled
-{\it Overlay databank unlocks data-driven analyses of biomolecules for all}, which we would like to submit for your consideration to be published as a Resource in \textit{ Nature Computational Science}.
+{\it Overlay databank unlocks data-driven analyses of biomolecules for all}, which we would like to submit for your consideration to be published as a Resource in \textit{ Nature Communications}.
 
@@
 We see that the NMRlipids Databank has both immediate and long term impact for a broad interdisciplinary audience of academic and industry researchers across different disciplines: from biophysics and biochemistry to biology, biotechnology and biomedicine.
-Feedback from our first submissions to \textit{Nature Methods} and \textit{ Nature Structural \& Molecular Biology} states that the manuscript is of interest on these fields of research, yet the level of biological insight was not sufficient for these journals.
-Therefore, we believe that \textit{ Nature Computational Science} is the best forum for our manuscript and trust our manuscript to fulfill its high standards for publication.
+Feedback from our first submissions to \textit{Nature Methods}, \textit{Nature Structural \& Molecular Biology}, and \textit{Nature Computational Science} states that the manuscript is of significant interest on these fields of research, yet the level of biological insights or broader interest were not sufficient for these journals.
+Nevertheless, we believe that our manuscript represents important advances of significance to specialists within these fields, and therefore fulfills the high standards for a publication in \textit{ Nature Communications}.
 \\

### 2023-10-10 main.tex

@@
 
+\pagebreak
 \begin{abstract}
@@
 The NMRlipids Databank is constructed using the NMRlipids project protocol, in which all the content is openly accessible throughout the project~\cite{botan15}.
-Currently, the NMRlipids Databank contains 726 simulation trajectories with the total length of approximately 0.4\,ms.
-Single-component lipid membranes and binary mixtures, depicted in Fig.~\ref{fig:overlay}F, are currently most abundant in the NMRlipids Databank, yet mixtures with up to five lipid types are available.
+Currently, the NMRlipids Databank contains 765 simulation trajectories with the total length of approximately 0.4\,ms.
+Single-component lipid membranes and binary mixtures are currently most abundant in the NMRlipids Databank, yet mixtures with up to five lipid types are available.
+For available mixtures, see Fig.~\ref{fig:overlay}E.
 The distribution of lipids among the available simulations, shown in Fig.~\ref{fig:overlay}B, roughly resembles the biological relative abundance of different lipid types, with phosphatidylcholine (PC) being the most common followed by cholesterol, phosphatidylethanolamine (PE), phosphatidylserine (PS), phosphatidylglycerol (PG), phosphatidylinositol (PI), and other lipids, depending on organism and organelle~\cite{vanmeer08}.
-Abbreviations and full names of all lipids present in the Databank are listed in Table~\ref{tab:abbreviations}.
+Abbreviations and full names of all lipids present in the Databank are listed in the NMRlipids databank documentation~\cite{NMRlipidsDocsMolAndMapping}.
+%\url{nmrlipids.github.io/moleculesAndMapping.html}.
 Force fields used in simulations cover all the essential parameter sets commonly used in lipid simulations, see Fig.~\ref{fig:overlay}C and Table~\ref{tab:ForceFields}, including also united atom and polarizable force fields.
@@
 
-\begin{figure}[t]
+\begin{figure}[tp]
     \centering
@@
 A more detailed structure of the \textit{Databank layer} is shown in Fig.~\ref{DatabankStructure} in the SI.
-    \textbf{B} Distribution of lipids present in the trajectories of the Databank. `Others' lists lipids occurring in five or fewer simulations. 
+    \textbf{B} Distribution of lipids present in the trajectories of the Databank. `Others' lists lipids occurring in six or fewer simulations. 
     \textbf{C} Distribution of force fields in the simulations in the Databank. References for each force field are given in Table~\ref{tab:ForceFields}.
     \textbf{D} Flowchart for performing an analysis of properties through all MD simulations in the NMRlipids Databank using the API.
-    \textbf{E} Flowchart for accessing results calculated from the NMRlipids Databank and stored to the \textit{Application layer}.
-    \textbf{F} Currently available single lipid component and binary mixtures in the NMRlipids Databank.
-Colorbar shows the number of available simulations with the darkest green indicating three or more.
+    %\textbf{E} Flowchart for accessing results calculated from the NMRlipids Databank and stored to the \textit{Application layer}.
+    \textbf{E} Currently available lipid mixtures in the NMRlipids Databank. Colorbar shows the number of available simulations with the darkest green indicating three or more.  
     }

### 2023-10-10 SI.tex
Updates from Overleaf

@@
 \usepackage{xr}
+\usepackage{comment}
 
@@
     AMOEBA~\cite{chu2018polarizable} \\
+    OpenFF
     \end{tabular}

### 2023-10-10 main.tex
Updates from Overleaf

@@
 Finally, the analysis results can be stored using the same structure as in the {\it Databank layer}.
-Documentation of the NMRlipids Databank-API and further instructions are available at \href{https://nmrlipids.github.io/}{nmrlipids.github.io}, and template for new user-defined analyses at \href{https://github.com/NMRLipids/databank-template}{github.com/NMRLipids/databank-template}.
+Documentation of the NMRlipids Databank-API and template for new user-defined analyses with further instructions are available at the NMRlipids databank documentation~\cite{NMRlipidsExampleAndTutorials}.
+%\href{https://nmrlipids.github.io/}{nmrlipids.github.io},  at \href{https://github.com/NMRLipids/databank-template}{github.com/NMRLipids/databank-template}.
 %Practical examples of codes that utilize the NMRlipids Databank-API, as well as a template for a new analysis, can be found from locations listed in Table~\ref{tab:codes}.

### 2023-10-10 SI.tex
Update text

@@
 %\noindent Please note: Abbreviations should be introduced at the first mention in the main text – no abbreviations lists. Suggested structure of main text (not enforced) is provided below.
+
+\pagebreak
+\section{Processing results in the application layer}
+
+\begin{figure}[!h]
+    \centerfloat
+    \includegraphics[width = 90mm]{Figures/resultProcessing.pdf}
+    \caption{Flowchart for accessing results calculated from the NMRlipids Databank and stored to the \textit{Application layer}.}
+    \label{fig:resultsProcessingSI}
+\end{figure}
 
@@
 \\
-        plotSimulation.ipynb & Plot results from a simulation together with the experimental data if available & README.yaml files and results in the {\it databank layer} & Figs~\ref{fig:quality}D-F and ~\ref{fig:POPC_POPE_dataSI}
+        %plotSimulation.ipynb & Plot results from a simulation together with the experimental data if available & README.yaml files and results in the {\it databank layer} & Figs~\ref{fig:quality}D-F and ~\ref{fig:POPC_POPE_dataSI}
     \end{tabular}

### 2023-10-10 main.tex
Update text

@@
     \textbf{D} Flowchart for performing an analysis of properties through all MD simulations in the NMRlipids Databank using the API.
-    %\textbf{E} Flowchart for accessing results calculated from the NMRlipids Databank and stored to the \textit{Application layer}.
     \textbf{E} Currently available lipid mixtures in the NMRlipids Databank. Colorbar shows the number of available simulations with the darkest green indicating three or more.  
@@
 While analysis codes and results for basic membrane properties are included in the {\it Databank layer}, unlimited further analyses can be implemented by anyone in separate repositories in the {\it Application layer}.
-When {\it Application layer} repositories are organized by mimicking the {\it Databank layer} structure, they can be accessed programmatically and further analyzed using the tools in the NMRlipids Databank-API by implementing the flowchart demonstrated in Fig.~\ref{fig:overlay}E.
+When {\it Application layer} repositories are organized by mimicking the {\it Databank layer} structure, they can be accessed programmatically and further analyzed using the tools in the NMRlipids Databank-API by implementing the flowchart demonstrated in Fig.~\ref{fig:resultsProcessingSI}.
 Novel analyses that demonstrate the power of NMRlipids Databank in selecting the best simulation models, analysing rare phenomena, and extending MD simulations to new fields are implemented in an {\it Application-layer} repository located at~\href{https://github.com/NMRlipids/DataBankManuscript}{github.com/NMRlipids/DataBankManuscript}.

### 2023-10-11 SI.tex
Updates from Overleaf

@@
 \pagebreak
+\section{Testing machine learning models against literature data}
+\begin{figure}
+    \centering
+    \includegraphics[width=88mm]{Figures/MLtest/Ridgetest.pdf}
+    \includegraphics[width=88mm]{Figures/MLtest/AdaBoostRegressortest.pdf}
+    \caption{Predictions of area per lipids of POPC, POPE, POPS, PI, sphingomyelin lipids and cholesterol from linear regression model against literature data from simulations (green~\cite{shahane19}, blue~\cite{kumar21}, and red~\cite{oliveira22}).}
+    \label{fig:LinearRegressionTest}
+\end{figure}
+
+\pagebreak
 \section{Water permeation through membranes}
@@
 \\
+        APLpredictor.ipynb & Predict area per lipid of a membrane based in its lipid composition from machine learning models trained using the NMRlipids databank & README.yaml and apl.dat files & Machine learning models at /Data/APLpredictor/ \\
         %plotSimulation.ipynb & Plot results from a simulation together with the experimental data if available & README.yaml files and results in the {\it databank layer} & Figs~\ref{fig:quality}D-F and ~\ref{fig:POPC_POPE_dataSI}

### 2023-10-11 SI.tex

@@
 \section{Testing machine learning models against literature data}
-\begin{figure}
+\begin{figure}[hb]
     \centering
-    \includegraphics[width=88mm]{Figures/MLtest/Ridgetest.pdf}
-    \includegraphics[width=88mm]{Figures/MLtest/AdaBoostRegressortest.pdf}
+    \includegraphics[width=50mm]{Figures/MLtest/Ridgetest.pdf}
+    \includegraphics[width=50mm]{Figures/MLtest/Lassotest.pdf}
+    \includegraphics[width=50mm]{Figures/MLtest/ElasticNettest.pdf}
+    \includegraphics[width=50mm]{Figures/MLtest/DecisionTreeRegressortest.pdf}
+    \includegraphics[width=50mm]{Figures/MLtest/RandomForestRegressortest.pdf}
+    \includegraphics[width=50mm]{Figures/MLtest/GradientBoostingRegressortest.pdf}
+    \includegraphics[width=50mm]{Figures/MLtest/AdaBoostRegressortest.pdf}
+    \includegraphics[width=50mm]{Figures/MLtest/KNeighborsRegressortest.pdf}
+    \includegraphics[width=50mm]{Figures/MLtest/MLPRegressortest.pdf}
+    \includegraphics[width=50mm]{Figures/MLtest/XGBRegressortest.pdf}
     \caption{Predictions of area per lipids of POPC, POPE, POPS, PI, sphingomyelin lipids and cholesterol from linear regression model against literature data from simulations (green~\cite{shahane19}, blue~\cite{kumar21}, and red~\cite{oliveira22}).}

### 2023-10-11 SI.tex
Update text

@@
     \includegraphics[width=50mm]{Figures/MLtest/XGBRegressortest.pdf}
-    \caption{Predictions of area per lipids of POPC, POPE, POPS, PI, sphingomyelin lipids and cholesterol from linear regression model against literature data from simulations (green~\cite{shahane19}, blue~\cite{kumar21}, and red~\cite{oliveira22}).}
-    \label{fig:LinearRegressionTest}
+    \caption{Predictions of area per lipids of POPC, POPE, POPS, PI, sphingomyelin lipids and cholesterol from the tested machine learning models against literature data from simulations (green~\cite{shahane19}, blue~\cite{kumar21}, and red~\cite{oliveira22}).}
+    \label{fig:MLTest}
 \end{figure}

### 2023-10-11 main.tex
Updates from Overleaf
Update text

@@
  After the parameter optimization, we tested predictions both against the remaining 20\% of the databank and against data from different ML models against area per lipids reported from simulations of membranes containing mixtures of POPC, POPE, POPS, PI, sphingomyelin lipids and cholesterol~\cite{shahane19,kumar21,oliveira22} not included in the databank.
-Linear regression and Ridge models gave the best correlation with the literature data in this test (predictions  from the linear regression model are shown in Fig.~\ref{fig:LinearRegressionTest}).
+Linear regression and Ridge models gave the best correlation with the literature data in this test (predictions  from the linear regression model are shown in Fig.~\ref{fig:LinearRegressionTest} and from other models in Fig.~\ref{fig:MLTest} in the supplementary information).
 The linear regression model was selected for further studies due to its simplicity.
@@
     \centering
-    \includegraphics[width=88mm]{Figures/LinearRegressionTest.pdf}
+    \includegraphics[width=88mm]{Figures/MLtest/LinearRegressiontest.pdf}
     \caption{Predictions of area per lipids of POPC, POPE, POPS, PI, sphingomyelin lipids and cholesterol from linear regression model against literature data from simulations (green~\cite{shahane19}, blue~\cite{kumar21}, and red~\cite{oliveira22}).}
@@
 To demonstrate the usefulness of the constructed models for understanding multi-component membrane properties, we predicted later packing (ares per lipids) and membrane thicknesses of common biological membranes based on their lipid compositions reported in the literature, see table~\ref{tab:MLpredictions}.
-The model predicts substantial 50\% difference in area per lipid and 0.5\,nm in thickness between most densely (influenza virus) and loosely (mitochondria) packed membranes.
-Because these parameters correlate with many biologically relevant functions of membranes, such as permeation, cholesterol flip-flops (see next sections), and interactions with proteins~\cite{phillips09}, our results demonstrate that NMRlipids databank can be used to give valuable insights on biologically relevant properties of complex biological membranes.
+The model predicts substantial 50\% difference in area per lipid between most densely (influenza virus) and loosely (mitochondria) packed membranes.
+Difference of 0.8\,nm in thickness is predicted between thinnest (bacterial) and thicknest (plasma) membranes.
+Such differences are expected to effect on many biologically relevant functions of membranes, such as permeation, cholesterol flip-flops (see next sections), and interactions with proteins~\cite{phillips09}, demonstrating that NMRlipids databank can be used to give valuable insights on biologically relevant properties of complex biological membranes.
 Most importantly, the delivered programmatic access to increasing amount of MD simulation data enables training of machine learning models that predict various membrane properties for all.
@@
              Mitochondria & POPC(37) : POPE(31) : PI(6) : CL(22) : 
-             CHOL(4)~\cite{vanmeer08,escriba15,casares19} & 74 (66)$^*$ & 4.3 \\
-             Bacterial & POPC(20) : POPE(35) : POPG(35) : CL(5) : CHOL(5)~\cite{chwastek20}  & 62 & 4.2 \\
-             ER &  POPC(54) : POPE(20) : POPS(4) : PI(11) : SM(4) : CHOL(8)~\cite{vanmeer08,escriba15,casares19} & 56 & 4.5 \\
-             Golgi & POPC(36) : POPE(21) : POPS(6) : PI(12) : SM(7) : CHOL(18)~\cite{vanmeer08,escriba15,casares19} & 51 & 4.7 \\
-             Plasma & POPC(23) : POPE(11) : POPS(8) : PI(7) : SM(17) : CHOL(34)~\cite{vanmeer08,escriba15,casares19} & 45 & 4.9 \\
+             CHOL(4)~\cite{vanmeer08,escriba15,casares19} & 74 (66)$^*$ & 4.4 \\
+             Bacterial & POPC(20) : POPE(35) : POPG(35) : CL(5) : CHOL(5)~\cite{chwastek20}  & 61 (60)$^*$ & 4.2 \\
+             ER &  POPC(54) : POPE(20) : POPS(4) : PI(11) : SM(4) : CHOL(8)~\cite{vanmeer08,escriba15,casares19} & 57 & 4.6 \\
+             Golgi & POPC(36) : POPE(21) : POPS(6) : PI(12) : SM(7) : CHOL(18)~\cite{vanmeer08,escriba15,casares19} & 52 & 4.8 \\
+             Plasma & POPC(23) : POPE(11) : POPS(8) : PI(7) : SM(17) : CHOL(34)~\cite{vanmeer08,escriba15,casares19} & 45 & 5.0 \\
              Synaptic & POPC(28) : POPE(20) : POPS(9) : PI(2) : SM(4) : CHOL(37)~\cite{binotti21}  &  45 & 4.8 \\
@@
     \hline
-    Computational and data scientists & Databanks with programmatic access, training sets for machine learning applications, automatic optimization of simulation models & Programmatic access to all publicly available MD simulation trajectories, machine learning models predicting properties of complex biomolecular assemblies, optimizing force field and model parameters from atomistic to continuum scale \\
+    Computational and data scientists & Databanks with programmatic access, training sets for machine learning applications, automatic optimization of simulation models & Programmatic access to all publicly available MD simulation trajectories, machine learning models predicting properties of complex biomolecular assemblies (Table~\ref{tab:MLpredictions}), optimizing force field and model parameters from atomistic to continuum scale \\
 
@@
 To enable automatic analyses over all the simulations in the NMRlipids Databank, we have defined universal naming conventions for the molecules and atoms therein.
-The universal abbreviations used in the NMRlipids Databank for each molecule are listed in the NMRlipids databank documentation \href{https://nmrlipids.github.io/moleculesAndMapping.html}{nmrlipids.github.io/moleculesAndMapping.html}.
+The universal abbreviations used in the NMRlipids Databank for each molecule are listed in the NMRlipids Databank documentation~\cite{NMRlipidsDocsMolAndMapping}.
 %Table~\ref{tab:abbreviations} in the supplementary information. 
-The atom names used in simulation trajectories are connected to the universal atom names using the mapping files defined in the NMRlipids Project (\url{https://nmrlipids.github.io/moleculesAndMapping.html}).
+The atom names used in simulation trajectories are connected to the universal atom names using the mapping files defined in the NMRlipids Project~\cite{NMRlipidsDocsMolAndMapping}.
 These files are located at \texttt{/Scripts/BuildDatabank/mapping\_files} in the NMRlipids Databank repository.
@@
 In practise, the force-field-specific molecule names and mapping file
-names are given in the COMPOSITION dictionary in the README.yaml files for each molecule in each simulation in the NMRlipids Databank.
+names are defined in the README.yaml files for each molecule in each simulation as described in the NMRlipids Databank documentation~\cite{NMRlipidsREADMEfiles}.
 
@@
 Rest of the information to be stored in the README.yaml files will be then automatically extracted using the \texttt{/Scripts/BuildDatabank/AddData.py} script.
-The required manually entered and automatically extracted information are described in detail in Table~\ref{tab:READMEkeys} in the supplementary information and in the NMRlipids databank documentation \href{https://nmrlipids.github.io/}{nmrlipids.github.io/}.
+The required manually entered and automatically extracted information are described in detail in Table~\ref{tab:READMEkeys} in the supplementary information and in the NMRlipids databank documentation~\cite{NMRlipidsREADMEfiles}.
 The documentation includes also detailed instructions to add data.
@@
 Similarly to simulations, each experimental data set has a README.yaml file containing all the relevant information about the experiment.
-The keys and their descriptions for the experimental data, as well as detailed instructions to add the date, are given in Table~\ref{tab:READMEkeysEXP} and in the NMRlipids databank documentation \href{https://nmrlipids.github.io/}{nmrlipids.github.io}.
+The keys and their descriptions for the experimental data, as well as detailed instructions to add the date, are given in Table~\ref{tab:READMEkeysEXP} and in the NMRlipids databank documentation~\cite{NMRlipidsExperimentalData}.
 The NMR data currently in the NMRlipids Databank are taken from Refs.~\citenum{scherer87,ferreira13,antila19,melcr20,bacle21,hanashima19,doktorova20} and the X-ray scattering data from Refs.~\citenum{Kucerka08a,kucerka11,pan12b,pan14,kucerka15,NMRlipidsIII,doktorova20}.
@@
 3) performs the desired analysis on a local computer utilising the universal naming conventions for molecules and atoms defined in the README.yaml and mapping files.
-This general procedure is illustrated in Fig.~\ref{fig:overlay}D, and a template for user-defined analyses is available at at \href{https://github.com/NMRLipids/databank-template}{github.com/NMRLipids/databank-template}.
+This general procedure is illustrated in Fig.~\ref{fig:overlay}D, and a templates for user-defined analyses are available via NMRlipids Databank documentation~\cite{NMRlipidsExampleAndTutorials}.
 Further practical examples of codes performing such analyses are listed in Tables~\ref{tab:codes} and~\ref{tab:codesApplication}.
@@
 \subsection{Calculation of area per lipid and bilayer thickness}
-Area per lipids of bilayers were calculated by dividing the time-averaged area of the simulation box with the total number of lipids and surfactant molecules in the simulation (see Table~\ref{tab:abbreviations} for the list of molecules considered as lipids or surfactants).
+Area per lipids of bilayers were calculated by dividing the time-averaged area of the simulation box with the total number of lipids and surfactant molecules in the simulation. 
+%(see Table~\ref{tab:abbreviations} for the list of molecules considered as lipids or surfactants). 
 The script that calculates the area per lipid from all simulations in the NMRlipids Databank repository is available at \texttt{Scripts/AnalyzeDatabank/calcAPL.py} in the NMRlipids Databank repository.
@@
 \subsection{Training machine learning model to predict properties of multi-component membranes}
-    Different ML approaches were tested to find the best model to predict the area per lipids of multi-component membranes  from their lipid composition.
+Different machine learning approaches were tested to find the best model to predict the area per lipids of multi-component membranes  from their lipid composition.
 To his end, we tested linear, Lasso, Ridge, elastic net, decision tree, random forest, multi-layer perceptron, k-nearest neighbour, gradient boost, AdaBoost, and XGBoost regressors.
@@
 The best performing hyperparameters were then used to fit each model to the training set.
-The performance of the models was then tested on the remaining 20\% of the data extracted from the databank, and additionally,
-%selected optimal parameters for each model based on their performance in predicting area per lipids for simulations in the test set. 
-%After finding optimal parameters for each ML model, each model was trained again using all simulations in the databank. 
-%The performances of newly trained models were then tested 
-by predicting area per lipids from multi-component membrane simulations reported in the literature but not included in the databank~\cite{shahane19,kumar21,oliveira22}.
+The performance of the models was then tested on the remaining 20\% of the data extracted from the databank, and additionally, by predicting area per lipids from multi-component membrane simulations reported in the literature but not included in the databank~\cite{shahane19,kumar21,oliveira22}.
 Due to its  simplicity and good performance in these tests, we selected to use linear regression model for further studies.
@@
 The ML models and analysis were implemented with Python using scikit-learn which was supplemented with the xgboost library to include extreme gradient boosted decision tree regressor to the set of tested model.
+Jyputer notebook that trains the models and predicts are per lipids is available at \texttt{scripts/APLpredictor.ipynb} in the repository at~\url{https://github.com/NMRLipids/DataBankManuscript/}.
+In further applications, it is important to consider potential limitations arising from the currently limited amount of data for certain types of systems.
+For example, the current models have been trained and tested only for mixtures with palmitoyl and oleyl acyl chains because the databank does not yet contain enough data with varying number of double bonds.
+Such limitations are expected to decrease with increasing amount of data in the future.
 

### 2023-10-21 SI.tex
Update pdfs
Updates from Overleaf

@@
 \author[11]{Alexander Kuzmin}
+\author[26]{Anusha Lalitha} %ORCID: 0000-0002-2817-661X
+\author[27,28]{Fabio Lolicato}
 \author[17,18]{Jesper J. Madsen}  %ORCID: 0000-0003-1411-9080
@@
 %\author[2]{...}                   %ORCID: ?
+\author[25]{Cedric Mingham}
+\author[29]{Luca Monticelli}
 \author[1]{Ricky Nencini}
+\author[19,20]{Alexey M. Nesterenko}
 \author[21]{Thomas J. Piggot}
 \author[22]{{\'A}ngel Pi{\~n}eiro}
-%\author[11]{Suman Samantray} %ORCID: 0000-0003-3361-9582
+\author[12]{Suman Samantray} %ORCID: 0000-0003-3361-9582
+\author[19,20]{Nathalie Reuter}
 \author[10,22,23]{Fabi{\'a}n Su{\'a}rez-Lest{\'o}n}
+\author[19,20]{Reza Talandashti}
 \author[1,24,*]{O. H. Samuli Ollila} %ORCID: 0000-0002- 8728-1006
@@
 \affil[11]{no affiliation}
-%Physical and Computational Sciences Division, Pacific Northwest National Laboratory, Richland, Washington 99352, United States
-\affil[12]{Institute of Biological Information Processing: Structural Biochemistry (IBI-7), Forschungszentrum Jülich, Jülich 52428, Germany}
+\affil[12]{Institute of Biological Information Processing: Structural Biochemistry (IBI-7), Forschungszentrum Jülich, 52428 Jülich, Germany}
 \affil[13]{ariadne.ai GmbH (Germany), Häusserstra{\ss}e 3 Heidelberg 69115, Germany }
@@
 \affil[24]{VTT Technical Research Centre of Finland, Espoo, Finland}
+\affil[25]{Hochschule Mannheim, University of Applied Sciences, Mannheim, Germany}
+\affil[26]{Institut Charles Gerhardt Montpellier (UMR CNRS 5253), Universit{\'e} Montpellier, Place Eug{\`e}ne Bataillon, 34095 Montpellier, Cedex 05, France}
+\affil[27]{Heidelberg University Biochemistry Center, 69120 Heidelberg, Germany}
+\affil[28]{Department of Physics, University of Helsinki, FI-00014 Helsinki, Finland}
+\affil[29]{IBCP CNRS, UMR 5086 7, Lyon, France}
 
@@
     \includegraphics[width=50mm]{Figures/MLtest/XGBRegressortest.pdf}
-    \caption{Predictions of area per lipids of POPC, POPE, POPS, PI, sphingomyelin lipids and cholesterol from the tested machine learning models against literature data from simulations (green~\cite{shahane19}, blue~\cite{kumar21}, and red~\cite{oliveira22}).}
+    \caption{Predictions of area per lipids of POPC, POPE, POPS, PI, sphingomyelin lipids and cholesterol from the tested machine learning models against literature data from simulations
+      (green~\cite{shahane19}, blue~\cite{kumar21}, and red~\cite{oliveira22}). The black line depicts x=y where the prediction equals the litterature value.}
     \label{fig:MLTest}

### 2023-10-21 CoverLetterREV.tex

\documentclass[11pt]{letter}
\usepackage{graphicx,baskervald,microtype}
\usepackage{hyperref,amsmath,xcolor}
\hypersetup{
  colorlinks=true,
  linkcolor=black,
  urlcolor=black
}
\usepackage{pgfplots,marginnote}
\usepackage[top=0.5in,bottom=0.5in,left=1in,right=1in]{geometry}
%\newgeometry{margin=2.5cm}
\newcommand{\omamargin}[1]{\marginnote{\textbf{#1}}[7pt]}
\begin{document}
\reversemarginpar
\pagestyle{empty}
\noindent Dr. O. H. Samuli Ollila \\
\noindent Institute of Biotechnology \\
\noindent University of Helsinki, Finland \\
\noindent samuli.ollila@helsinki.fi\\
\noindent +358\,503\,746\,963 \\

Dear Editor,

thank you for considering our manuscript titled {\it Overlay databank unlocks data-driven analyses of biomolecules for all} for Nature Communications. We are grateful for the highly useful feedback that we received from our manuscript. Based on this feedback, we have now substantially revised the manuscript and the related online material. Particularly, we have improved the documentation and we now present a machine learning model that predicts multi-component membrane properties. Because we deliver new data and have major revisions in the manuscript that address the concerns of reviewers', we hope that you can reconsider our manuscript for Nature Communications. Overview on how the updates reply to reviewers' concerns are given in this letter, and detailed reply to reviewers' and the revised manuscript are enclosed.

The main concern of reviewer 1 can be summarized with these two quotations from the report: "\textit{This reviewer finds the database is important to the field of membrane biophysics and certainly could have more general applications. However, there lacks any novelty in this work that would capture the interest beyond those focused on simple model membranes that contain 1 or 2 lipids.}" and "\textit{If this manuscript presented an algorithm that learned how to predict membrane structure of bilayer with mixed lipids >3 components, then this would be more at the level for this high impact journal.}" Following the suggestion by the reviewer, we now present an algorithm that predicts structures of multi-component membranes, and apply this algorithm to membranes with lipid mixtures that are typical for various cellular membranes. Therefore, we believe that the main concern of reviewer 1 is well addressed in the revised version of our manuscript.

The main concern of the reviewer 3 is summarized in this quotation: "\textit{The paper itself proposes exciting ideas on how to use many MD data sets in conjunction with experimental data to obtain dynamics information on lipids that is not obtainable based on one or a few simulations alone. On the other hand, the tools provided do not allow an external user to easily do what the paper claims (“data-driven analysis […] for all”).}" To support this view, reviewer 3 provides extremely valuable technical feedback on the implementation of our API. We now provide new documentation for the NMRlipids databank at \url{https://nmrlipids.github.io/} and have implemented almost all suggestions by the reviewer 3. We totally agree with the reviewer about the importance of usability and documentation of the API, and understand its importance for the impact of our databank. However, we see this as rather technical feedback pointed on documentation and implementation of API. There are no critical points presented regarding the overlay structure of the databank (which is the core innovation in this work enabling the implementation of APIs and other applications), or the actual content of the manuscript. In contrary, all comments on the actual scientific content of the manuscript are positive, while the criticism considers technical implementation and documentation of the API.

Reviewer 2 seems to interpret the lack of author names and other information as a sign of negligence, although this was actually due to the double-blind review process. We believe that this may have biased his/her evaluation, and hope that comments of this reviewer would have less weight for the decision.

In conclusion, we believe that our revised manuscript and related content addresses all the concerns presented by reviewers' and hope that you can reconsider our manuscript for Nature Communications.

Sincerely yours,

on behalf of the authors,\\

Dr. O. H. Samuli Ollila

\end{document}

### 2023-10-21 main.tex

@@
 
- After the parameter optimization, we tested predictions both against the remaining 20\% of the databank and against data from different ML models against area per lipids reported from simulations of membranes containing mixtures of POPC, POPE, POPS, PI, sphingomyelin lipids and cholesterol~\cite{shahane19,kumar21,oliveira22} not included in the databank.
-Linear regression and Ridge models gave the best correlation with the literature data in this test (predictions  from the linear regression model are shown in Fig.~\ref{fig:LinearRegressionTest} and from other models in Fig.~\ref{fig:MLTest} in the supplementary information).
+ After the parameter optimization, we tested predictions from different ML models for the area per lipid both against the remaining 20\% of the databank and against area per lipids reported from simulations of membranes containing mixtures of POPC, POPE, POPS, PI, sphingomyelin lipids and cholesterol~\cite{shahane19,kumar21,oliveira22} that are not included in the databank.
+Essential differences between models were not observed when predicting area per lipids of 20\% of simulations selected as the test set, but linear regression and Ridge models gave the best correlations with the literature data (predictions  from the linear regression model are shown in Fig.~\ref{fig:LinearRegressionTest} and from other models in Fig.~\ref{fig:MLTest} in the supplementary information).
 The linear regression model was selected for further studies due to its simplicity.
@@
     \includegraphics[width=88mm]{Figures/MLtest/LinearRegressiontest.pdf}
-    \caption{Predictions of area per lipids of POPC, POPE, POPS, PI, sphingomyelin lipids and cholesterol from linear regression model against literature data from simulations (green~\cite{shahane19}, blue~\cite{kumar21}, and red~\cite{oliveira22}).}
+    \caption{Predictions of area per lipids of multi-component membranes composed of POPC, POPE, POPS, PI, sphingomyelin lipids and cholesterol from linear regression model against literature data from simulations (green~\cite{shahane19}, blue~\cite{kumar21}, and red~\cite{oliveira22}).
+Black line indicates x=y.}
     \label{fig:LinearRegressionTest}
@@
 The model predicts substantial 50\% difference in area per lipid between most densely (influenza virus) and loosely (mitochondria) packed membranes.
-Difference of 0.8\,nm in thickness is predicted between thinnest (bacterial) and thicknest (plasma) membranes.
+Difference of 0.8\,nm in thickness is predicted between thinnest (bacterial) and thickest (plasma) membranes.
 Such differences are expected to effect on many biologically relevant functions of membranes, such as permeation, cholesterol flip-flops (see next sections), and interactions with proteins~\cite{phillips09}, demonstrating that NMRlipids databank can be used to give valuable insights on biologically relevant properties of complex biological membranes.
@@
 Such applications could be analogous to AlphaFold \cite{jumper21} and other tools~\cite{baek21,lin23} that predict protein structures from their sequence using artificial intelligence.
-The analysis of cholesterol flip-flop events (Fig.~\ref{fig:flip-flops}) and water permeation through membranes (Fig.~\ref{fig:permeability}) demonstrate how a large amount of accessible simulation data in terms of quantity (\textit{e.g.}, simulation length and number of conformations) and content (\textit{e.g.}, lipid compositions and ion concentrations) enable analyses of rare phenomena that are beyond the current possibilities for a single research group.
+This is demonstrated here by building ML models to predict multi-component membrane properties.
+Furthermore, the analysis of cholesterol flip-flop events (Fig.~\ref{fig:flip-flops}) and water permeation through membranes (Fig.~\ref{fig:permeability}) demonstrate how a large amount of accessible simulation data in terms of quantity (\textit{e.g.}, simulation length and number of conformations) and content (\textit{e.g.}, lipid compositions and ion concentrations) enable analyses of rare phenomena that are beyond the current possibilities for a single research group.
 Such analyses also pave the way for applications of MD simulations in new fields, as demonstrated here by analysing an essential parameter in pharmacokinetic modeling and MRI imaging:\cite{nitsche19,topgaard20} the anisotropic diffusion of water in membrane systems (Fig.~\ref{fig:permeability}).
@@
 Different machine learning approaches were tested to find the best model to predict the area per lipids of multi-component membranes  from their lipid composition.
-To his end, we tested linear, Lasso, Ridge, elastic net, decision tree, random forest, multi-layer perceptron, k-nearest neighbour, gradient boost, AdaBoost, and XGBoost regressors.
+To his end, we tested linear, Lasso, Ridge, elastic net, decision tree, random forest, multi-layer perceptron, k-nearest neighbour, gradient boosting, AdaBoost, and XGBoost regressors.
 The models were trained to predict the area per lipid for a membrane with the given molar fractions of membrane lipids.
 Most abundant lipids in the databank, POPC, POPE, POPG, POPS and cholesterol, were considered explicitly, while SM, cardiolipins and PI lipids with  more sparse data were all grouped together based on their headgroups independently on acyl chain content.
-Salt concentration, temperature and other conditions were not considered as parameters at this stage.
+Salt concentration, temperature and other conditions were not considered.
 
 To find the optimal model for this task, we first randomly divided simulations into training (80\% of the simulations) and test (20\% of the simulations) sets.
-Then, we used the training set with 5-fold cross validation to conduct a grid search for hyperparameters for a set of ML regression models.
+Then, we used the training set with 5-fold cross validation to conduct a grid search for hyperparameters for the set of ML regression models.
 The best performing hyperparameters were then used to fit each model to the training set.
@@
 Due to its  simplicity and good performance in these tests, we selected to use linear regression model for further studies.
-The  linear regression model predicting membrane thicknesses was trained in similarly.
+The linear regression model predicting membrane thicknesses was trained similarly.
 The ML models and analysis were implemented with Python using scikit-learn which was supplemented with the xgboost library to include extreme gradient boosted decision tree regressor to the set of tested model.
-Jyputer notebook that trains the models and predicts are per lipids is available at \texttt{scripts/APLpredictor.ipynb} in the repository at~\url{https://github.com/NMRLipids/DataBankManuscript/}.
+Jupyter notebook that trains the models and predicts are per lipids is available at \texttt{scripts/APLpredictor.ipynb} in the repository at~\url{https://github.com/NMRLipids/DataBankManuscript/}.
 In further applications, it is important to consider potential limitations arising from the currently limited amount of data for certain types of systems.
 For example, the current models have been trained and tested only for mixtures with palmitoyl and oleyl acyl chains because the databank does not yet contain enough data with varying number of double bonds.
-Such limitations are expected to decrease with increasing amount of data in the future.
+Such limitations are expected to alleviate with increasing amount of data in the future.
 
@@
 greatly acknowledges financial support by the Ministry of Economics, Science and Digitalisation of the State of Saxony-Anhalt.
+F.L. was supported by the Deutsche Forschungsgemeinschaft (DFG LO 2821/1-1).
 We acknowledge all the NMRlipids Project contributors for making development of the NMRlipids Databank possible.
@@
 A.M.K. implemented the {\it Databank layer} and NMRlipids Databank-API.
-H.S.A. contributed to the quality evaluation metrics based on NMR and X-ray scattering form factors, and was involved in testing the GUI and polishing the manuscript.
+H.S.A.
+contributed to the quality evaluation metrics based on NMR and X-ray scattering form factors, wrote the machine learning code, and was involved in testing the GUI, polishing the manuscript, and in community outreach to source additional data.
 L.S.B. implemented roughly 630 pre-existing lipid bilayer simulations into the Databank and created the corresponding mapping files.
@@
 A.K. implemented the PCA-based equilibration metric.
+A.L. provided simulations of heterogeneous lipid membranes including sphingolipids and glycosphingolipids with CHARMM36 force field.
+F.S.-L. developed the back-end and participated in the design and development and validation of the graphical user interface (GUI) (databank.nmrlipids.fi).
+F.L. provided simulations using the CHARMM36m forcefield, incorporating cholesterol POPC and SAPI25 lipids.
 J.J.M. provided simulations with the Amber force field, general discussion, and manuscript editing.
@@
 and M.K.
+C.M. worked on the databank implementation and contributed code improvements.
+L.M. performed simulations (for sphingomyelin, CHARMM36 force field, and Orange force field)
 R.N. implemented the code for the calculations of X-ray scattering form factors.
+A.N. provided simulations with TLCL lipid, simulations with Reaction-Field instead of PME, and improved the implementation.
 T.J.P. performed many of the all-atom and united-atom simulations and contributed to discussions on the project.
 A.P. was responsible for the design, validation and supervision of the back-end and front-end for the graphical user interface (GUI) development (databank.nmrlipids.fi).
-%S.S. performed some of the simulations with CHARMM36 force field.
-F.S.-L. developed the back-end and participated in the design and development and validation of the graphical user interface (GUI) (databank.nmrlipids.fi).
-O.H.S.O. designed and supervised the work, implemented the water permeation and diffusion analyses, and wrote the manuscript.
+S.S. contributed simulations of heterogeneous lipid membranes including sphingomyelin and ceramides with CHARMM36 force field.
+R.T. Provided simulations of complex lipid-bilayers with charmm36m force field.
+N.R. contributed simulations of complex lipid-bilayers obtained with NAMD and the charmm36 force field
+O.H.S.O. designed and supervised the work, implemented the water permeation, diffusion and machine learning analyses, and wrote the manuscript.
 

### 2023-10-21 SI.tex
Updates from Overleaf
Update text

@@
     \centering
+    \includegraphics[width=50mm]{Figures/MLtest/LinearRegressiontest.pdf}
     \includegraphics[width=50mm]{Figures/MLtest/Ridgetest.pdf}
+    \includegraphics[width=50mm]{Figures/MLtest/MLPRegressortest.pdf}
     \includegraphics[width=50mm]{Figures/MLtest/Lassotest.pdf}
@@
     \includegraphics[width=50mm]{Figures/MLtest/KNeighborsRegressortest.pdf}
-    \includegraphics[width=50mm]{Figures/MLtest/MLPRegressortest.pdf}
     \includegraphics[width=50mm]{Figures/MLtest/XGBRegressortest.pdf}
-    \caption{Predictions of area per lipids of POPC, POPE, POPS, PI, sphingomyelin lipids and cholesterol from the tested machine learning models against literature data from simulations
-      (green~\cite{shahane19}, blue~\cite{kumar21}, and red~\cite{oliveira22}). The black line depicts x=y where the prediction equals the litterature value.}
+    \caption{Predictions of area per lipids of multi-component membranes composed of POPC, POPE, POPS, PI, sphingomyelin lipids and cholesterol from linear regression model against literature data from simulations
+      (green~\cite{shahane19}, blue~\cite{kumar21}, and red~\cite{oliveira22}).
+Mean squared error (RMSE) and Pearson correlation coefficient are shown for each model.
+The black line depicts x=y where the prediction equals the literature value.}
     \label{fig:MLTest}

### 2023-10-21 SI.tex
Update text

@@
 \author[22]{{\'A}ngel Pi{\~n}eiro}
+\author[19,20]{Nathalie Reuter}
 \author[12]{Suman Samantray} %ORCID: 0000-0003-3361-9582
-\author[19,20]{Nathalie Reuter}
 \author[10,22,23]{Fabi{\'a}n Su{\'a}rez-Lest{\'o}n}
@@
     AMOEBA~\cite{chu2018polarizable} \\
-    OpenFF
+    OpenFF~\cite{mobley18,boothroyd23,OpenFFgit}
     \end{tabular}

### 2023-10-23 main.tex
Update text
Update main.tex

@@
 F.S.-L. developed the back-end and participated in the design and development and validation of the graphical user interface (GUI) (databank.nmrlipids.fi).
-F.L. provided simulations using the CHARMM36m forcefield, incorporating cholesterol POPC and SAPI25 lipids.
-J.J.M. provided simulations with the Amber force field, general discussion, and manuscript editing.
+F.L. provided simulations using the CHARMM36m force field, incorporating cholesterol POPC and SAPI25 lipids.
+J.J.M. provided simulations with the Amber and CHARMM36m force fields, general discussion, and manuscript editing.
 M.S.M supported O.H.S.O.

### 2023-10-23 main.tex
Update main.tex

@@
 \affil[16]{School of Pharmacy, University of Eastern Finland, 70211 Kuopio, Finland}
-\affil[17]{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, 33612, United States of America}
-\affil[18]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida, 33612, United States of America}
+\affil[17]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida, 33612, United States of America}
+\affil[18]{Center for Global Health and Infectious Diseases Research, Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida 33612, United States of America}
 \affil[19]{Department of Chemistry, University of Bergen, 5020 Bergen, Norway}

### 2023-10-23 main.tex
Update main.tex

@@
 \affil[17]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida, 33612, United States of America}
-\affil[18]{Center for Global Health and Infectious Diseases Research, Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida 33612, United States of America}
+\affil[18]{Center for Global Health and Infectious Diseases Research, Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, 33612, United States of America}
 \affil[19]{Department of Chemistry, University of Bergen, 5020 Bergen, Norway}

### 2023-10-23 main.tex
Update main.tex

--- 2023-10-23 main.tex
@@
 \affil[16]{School of Pharmacy, University of Eastern Finland, 70211 Kuopio, Finland}
-\affil[17]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida, 33612, United States of America}
+\affil[17]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida 33612, United States of America}
 \affil[18]{Center for Global Health and Infectious Diseases Research, Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida 33612, United States of America}

### 2023-10-23 main.tex
Update

@@
 \affil[16]{School of Pharmacy, University of Eastern Finland, 70211 Kuopio, Finland}
-\affil[17]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida 33612, United States of America}
-\affil[18]{Center for Global Health and Infectious Diseases Research, Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida 33612, United States of America}
+\affil[17]{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, 33612, United States of America}
+\affil[18]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida, 33612, United States of America}
 \affil[19]{Department of Chemistry, University of Bergen, 5020 Bergen, Norway}
@@
 
-\subsubsection{Using NMRlipids databank as a training set for machine learning applications: Predicting multi-component membrane properties}
+\subsubsection{Using NMRlipids Databank as a training set for machine learning applications: Predicting multi-component membrane properties}
 
@@
 \subsubsection{Extending the scope of MD simulations to new fields using NMRlipids Databank: Water diffusion anisotropy in membrane systems}
-The anisotropic diffusion of water and hydrophilic molecules in directions parallel and perpendicular to membranes is important parameter in models describing the translocation of drugs through biological material, particularly in the skin \cite{hansen13,wen18,nitsche19,roberts21}.
-Water anisotropic diffusion plays a role also in the signal formation in diffusion tensor MRI imaging~\cite{topgaard20}.
+The anisotropic diffusion of water and hydrophilic molecules in directions parallel and perpendicular to membranes is an important parameter in models describing the translocation of drugs through biological material, particularly in the skin \cite{hansen13,wen18,nitsche19,roberts21}.
+Water anisotropic diffusion plays a role also in the signal formation in diffusion-tensor MRI imaging~\cite{topgaard20}.
 MD simulations are rarely used to analyze the anisotropic diffusion of water, since only a few membrane permeation events of water are typically observed in a single MD simulation trajectory~\cite{venable19,camilo2022}, thereby making the collection of a sufficient amount of data challenging.
@@
 F.S.-L. developed the back-end and participated in the design and development and validation of the graphical user interface (GUI) (databank.nmrlipids.fi).
-F.L. provided simulations using the CHARMM36m force field, incorporating cholesterol POPC and SAPI25 lipids.
-J.J.M. provided simulations with the Amber and CHARMM36m force fields, general discussion, and manuscript editing.
+F.L. provided simulations using the CHARMM36m forcefield, incorporating cholesterol POPC and SAPI25 lipids.
+J.J.M. provided simulations with the Amber force field, general discussion, and manuscript editing.
 M.S.M supported O.H.S.O.

### 2023-10-23 main.tex

@@
 \affil[16]{School of Pharmacy, University of Eastern Finland, 70211 Kuopio, Finland}
-\affil[17]{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, 33612, United States of America}
-\affil[18]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida, 33612, United States of America}
+\affil[17]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida 33612, United States of America}
+\affil[18]{Center for Global Health and Infectious Diseases Research, Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida 33612, United States of America}
 \affil[19]{Department of Chemistry, University of Bergen, 5020 Bergen, Norway}

### 2023-12-04 main.tex
Update text

--- 2023-10-23 main.tex
@@
 \author[29]{Luca Monticelli}
-\author[1]{Ricky Nencini}
+\author[1,30]{Ricky Nencini}
 \author[19,20]{Alexey M. Nesterenko}
@@
 \affil[29]{IBCP CNRS, UMR 5086 7, Lyon, France}
+\affil[30]{Division of Pharmaceutical Biosciences, Faculty of Pharmacy, University of Helsinki, Helsinki 00014, Finland}
 
@@
 
-%Cellular membrane lipid composition is implicated in diseases and controls major biological functions, but membranes are difficult to study experimentally due to their intrinsic disorder and complex phase behaviour.
-Molecular dynamics (MD) simulations have been useful in understanding membrane systems, but they require significant computational resources and often suffer from inaccuracies in model parameters.
-Applications of data-driven and machine learning methods, currently revolutionising many fields, remain of limited use for membrane systems due to the lack of suitable training sets.
-Here we present the NMRlipids Databank---a community-driven, open-for-all database featuring programmatic access to quality-evaluated atom-resolution MD simulations of lipid bilayers.
-The NMRlipids Databank will benefit scientists in different disciplines by providing automatic ranking of simulations based on their quality against experiments, programmable interface for flexible implementation of data-driven and machine learning applications, and rapid access to simulation data through a graphical user interface.
-To demonstrate how it unlocks possibilities beyond current MD simulation studies, we analyzed the NMRlipids Databank to reveal how anisotropic diffusion of water and cholesterol flip-flop rates depend on membrane properties.
 %TC:endignore
@@
 After cloning the databank repository to a local computer, raw data of each simulation can be accessed and analyzed with the help of functions delivered by the NMRlipids Databank-API.
-%by iterating through the README.yaml files in the {\it Databank layer} that contain links pointing to the locations of simulation trajectories and other relevant files.
-Simulations can then be automatically analyzed with the help of the README.yaml and the
 Analyses over simulations with different naming convetions can be automatically performed with the help of mapping files that associate the specific naming conventions in each simulation with the universal molecule and atom names used by the Databank.
@@
 Documentation of the NMRlipids Databank-API and template for new user-defined analyses with further instructions are available at the NMRlipids databank documentation~\cite{NMRlipidsExampleAndTutorials}.
-%\href{https://nmrlipids.github.io/}{nmrlipids.github.io},  at \href{https://github.com/NMRLipids/databank-template}{github.com/NMRLipids/databank-template}.
-%Practical examples of codes that utilize the NMRlipids Databank-API, as well as a template for a new analysis, can be found from locations listed in Table~\ref{tab:codes}.
 
@@
 Qualities against X-ray scattering experiments ($FF_q$) were estimated as the difference in the experimental and simulated locations of the first form factor minimum.
-These measures are good proxies for membrane properties
-%can be used to evaluate membrane dimensions, as the form factor minima locations and acyl chain C--H bond order parameters 
-because they correlate with the membrane lateral packing and thickness (Figs.~\ref{fig:quality}G,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}).
+These measures are good proxies for membrane properties because they correlate with the membrane lateral packing and thickness (Figs.~\ref{fig:quality}G,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}).
 Ergodicity of conformational sampling of lipids was estimated by calculating $\tau_\mathrm{rel}$, the convergence time of the slowest principal component divided by the simulation length.
@@
 Fig.~\ref{fig:quality} demonstrates how the automatic simulation-quality evaluation and the NMRlipids Databank-API enable rapid selection of the best models for membrane simulations.
-Fig.~\ref{fig:quality}A illustrates that predictions for the lateral packing of membranes composed of two most biologically-abundant neutral membrane lipids, POPC and POPE~\cite{vanmeer08},
-%for area per lipid, one of the most important parameters used to characterize cellular membranes, 
-diverge between different force fields.
+Fig.~\ref{fig:quality}A illustrates that predictions for the lateral packing of membranes composed of two most biologically-abundant neutral membrane lipids, POPC and POPE~\cite{vanmeer08}, diverge between different force fields.
 To find the most realistic parameters to simulate membranes with these lipids, we first ranked all simulations based on order parameter quality (Fig.~\ref{fig:top50simulations}), then picked force fields that occurr in Fig.~\ref{fig:quality}A (that is: force fields for both POPC and POPE in the Databank), and then ranked them according to the quality of the \textit{sn}-1 chain of POPC (Fig.~\ref{fig:quality}B) and of POPE (Fig.~\ref{fig:quality}C).
@@
 The universal abbreviations used in the NMRlipids Databank for each molecule are listed in the NMRlipids Databank documentation~\cite{NMRlipidsDocsMolAndMapping}.
-%Table~\ref{tab:abbreviations} in the supplementary information. 
 The atom names used in simulation trajectories are connected to the universal atom names using the mapping files defined in the NMRlipids Project~\cite{NMRlipidsDocsMolAndMapping}.
@@
 These files also define whether an atom belongs to the headgroup, glycerol backbone, or acyl chain region in a lipid.
-In practise, the force-field-specific molecule names and mapping file
-names are defined in the README.yaml files for each molecule in each simulation as described in the NMRlipids Databank documentation~\cite{NMRlipidsREADMEfiles}.
+In practise, the force-field-specific molecule names and mapping file names are defined in the README.yaml files for each molecule in each simulation as described in the NMRlipids Databank documentation~\cite{NMRlipidsREADMEfiles}.
 
@@
 \subsection{Calculation of area per lipid and bilayer thickness}
-Area per lipids of bilayers were calculated by dividing the time-averaged area of the simulation box with the total number of lipids and surfactant molecules in the simulation. 
-%(see Table~\ref{tab:abbreviations} for the list of molecules considered as lipids or surfactants). 
-The script that calculates the area per lipid from all simulations in the NMRlipids Databank repository is available at \texttt{Scripts/AnalyzeDatabank/calcAPL.py} in the NMRlipids Databank repository.
+Area per lipids of bilayers were calculated by dividing the time-averaged area of the simulation box with the total number of lipids and surfactant molecules in the simulation.The script that calculates the area per lipid from all simulations in the NMRlipids Databank repository is available at \texttt{Scripts/AnalyzeDatabank/calcAPL.py} in the NMRlipids Databank repository.
 The resulting area per lipids are stored for all simulations in files named \texttt{apl.json} at folders in \texttt{/Data/Simulations}.
@@
 
-%\newline \\
-%The purpose of a mapping file is to circumvent the problem caused by different atom naming conventions used by different force fields.
-The first column of a mapping file contains general atom names.
-The second column contains the name of the atom as it is in the force field.
-If the lipid consists of several residues which is the case in some AMBER force fields, then a third column is needed which contains the name of the residue to which each atom belongs to.
+\subsection{Data availability}
+All data related to this study are available via git repositories \url{https://github.com/NMRlipids/Databank/} (DOI: 10.5281/zenodo.7875567) and \url{https://github.com/NMRLipids/DataBankManuscript/} (DOI: 10.5281/zenodo.10252380).
+
+%name e.g “figshare”] with the identifier(s) [data DOI(s) e.g. “doi:10.6084/m9.figshare.1499292_D8”][Reference number]
+
+\subsection{Code availability}
+All codes used in this study are available via git repositories \url{https://github.com/NMRlipids/Databank/} (DOI: 10.5281/zenodo.7875567) and \url{https://github.com/NMRLipids/DataBankManuscript/} (DOI: 10.5281/zenodo.10252380).
 
@@
 M.S.M. was supported by the Trond Mohn Foundation BFS2017TMT01.
-O.H.S.O, A.M.K, R.N. and L. S. B. acknowledge CSC --- IT Center for Science for computational resources and Academy of Finland (grants 315596 and 319902) for financial support.
+O.H.S.O, A.M.K, R.N. and L. S. B. acknowledge CSC --- IT Center for Science for computational resources and Academy of Finland (grant nos. 315596, 319902 \& 345631) for financial support.
 M.K.

### 2023-12-04 SI.tex

@@
     \centering
-    \includegraphics[width=50mm]{Figures/MLtest/LinearRegressiontest.pdf}
-    \includegraphics[width=50mm]{Figures/MLtest/Ridgetest.pdf}
-    \includegraphics[width=50mm]{Figures/MLtest/MLPRegressortest.pdf}
-    \includegraphics[width=50mm]{Figures/MLtest/Lassotest.pdf}
-    \includegraphics[width=50mm]{Figures/MLtest/ElasticNettest.pdf}
-    \includegraphics[width=50mm]{Figures/MLtest/DecisionTreeRegressortest.pdf}
-    \includegraphics[width=50mm]{Figures/MLtest/RandomForestRegressortest.pdf}
-    \includegraphics[width=50mm]{Figures/MLtest/GradientBoostingRegressortest.pdf}
-    \includegraphics[width=50mm]{Figures/MLtest/AdaBoostRegressortest.pdf}
-    \includegraphics[width=50mm]{Figures/MLtest/KNeighborsRegressortest.pdf}
-    \includegraphics[width=50mm]{Figures/MLtest/XGBRegressortest.pdf}
-    \caption{Predictions of area per lipids of multi-component membranes composed of POPC, POPE, POPS, PI, sphingomyelin lipids and cholesterol from linear regression model against literature data from simulations
-      (green~\cite{shahane19}, blue~\cite{kumar21}, and red~\cite{oliveira22}).
+    \includegraphics[width=45mm]{Figures/MLtest/LinearRegressiontest.pdf}
+    \includegraphics[width=45mm]{Figures/MLtest/Ridgetest.pdf}
+    \includegraphics[width=45mm]{Figures/MLtest/MLPRegressortest.pdf}
+    \includegraphics[width=45mm]{Figures/MLtest/Lassotest.pdf}
+    \includegraphics[width=45mm]{Figures/MLtest/ElasticNettest.pdf}
+    \includegraphics[width=45mm]{Figures/MLtest/DecisionTreeRegressortest.pdf}
+    \includegraphics[width=45mm]{Figures/MLtest/RandomForestRegressortest.pdf}
+    \includegraphics[width=45mm]{Figures/MLtest/GradientBoostingRegressortest.pdf}
+    \includegraphics[width=45mm]{Figures/MLtest/AdaBoostRegressortest.pdf}
+    \includegraphics[width=45mm]{Figures/MLtest/KNeighborsRegressortest.pdf}
+    \includegraphics[width=45mm]{Figures/MLtest/XGBRegressortest.pdf}
+    \caption{Predictions of area per lipids of multi-component membranes composed of POPC, POPE, POPS, PI, sphingomyelin lipids and cholesterol from linear regression model against literature data from simulations using CHARMM36 parameters (green~\cite{shahane19}, blue~\cite{kumar21}, and red~\cite{oliveira22}).
 Mean squared error (RMSE) and Pearson correlation coefficient are shown for each model.
-The black line depicts x=y where the prediction equals the literature value.}
+The black line depicts x=y where the prediction equals the literature value.
+Our machine learning models that are trained using simulations with wide range of force field parameters  have general tendency to predict larger area per lipid values than reported from CHARMM36 simulations in the literature.
+One explanation could be lower area per lipids predicted by the CHARMM36 parameters when compared with many other force fields in the training set~\cite{javanainen23}, yet other explanations cannot be ruled out.}
     \label{fig:MLTest}

### 2023-12-04 main.tex
Update affiliations

--- 2023-10-23 main.tex
@@
 \author[25]{Cedric Mingham}
-\author[29]{Luca Monticelli}
+\author[29,30]{Luca Monticelli}
 \author[1]{Ricky Nencini}
@@
 %\affil[2]{Affiliation, department, city, postcode, country}
-\affil[5]{Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, P.O. Box 35, Jyv{\"a}skyl{\"a}, 40014 , Finland}
+\affil[5]{Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, Jyv{\"a}skyl{\"a}, 40014, Finland}
 \affil[6]{Departamento de Ciencias B\'{a}sicas, Tecnol\'{o}gico Nacional de M\'{e}xico - ITS Zacatecas Occidente, Sombrerete, Zacatecas, 99102, M\'{e}xico}
@@
 \affil[28]{Department of Physics, University of Helsinki, FI-00014 Helsinki, Finland}
-\affil[29]{IBCP CNRS, UMR 5086 7, Lyon, France}
+\affil[29]{University of Lyon, CNRS, Molecular Microbiology and Structural Biochemistry (MMSB, UMR 5086), F-69007, Lyon, France}
+\affil[30]{Institut National de la Sant{\'e} et de la Recherche M{\'e}dicale (INSERM), Lyon, France}
 
@@
 We acknowledge all the NMRlipids Project contributors for making development of the NMRlipids Databank possible.
+L.M. acknowledges funding by the Institut National de la Santé et de la Recherche Médicale (INSERM).
 

### 2023-12-04 main.tex

--- 2023-12-04 main.tex
@@
 \author[25]{Cedric Mingham}
-\author[29]{Luca Monticelli}
-\author[1,30]{Ricky Nencini}
+\author[29,30]{Luca Monticelli}
+\author[1,31]{Ricky Nencini}
 \author[19,20]{Alexey M. Nesterenko}
@@
 \affil[28]{Department of Physics, University of Helsinki, FI-00014 Helsinki, Finland}
-\affil[29]{IBCP CNRS, UMR 5086 7, Lyon, France}
-\affil[30]{Division of Pharmaceutical Biosciences, Faculty of Pharmacy, University of Helsinki, Helsinki 00014, Finland}
+\affil[29]{University of Lyon, CNRS, Molecular Microbiology and Structural Biochemistry (MMSB, UMR 5086), F-69007, Lyon, France}
+\affil[30]{Institut National de la Sant{\'e} et de la Recherche M{\'e}dicale (INSERM), Lyon, France}
+\affil[31]{Division of Pharmaceutical Biosciences, Faculty of Pharmacy, University of Helsinki, Helsinki 00014, Finland}
 

### 2023-12-04 SI.tex
Update pdfs

@@
 \author[11]{Ivan Gushchin}
-%\author[2]{Matti Javanainen}      %ORCID: 0000-0003-4858-364X
 \author[12,13]{Batuhan Kav}           %ORCID: 0000-0003-4990-373X
@@
 \author[2,19,20]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
-%\author[2]{...}                   %ORCID: ?
 \author[25]{Cedric Mingham}
-\author[29]{Luca Monticelli}
-\author[1,30]{Ricky Nencini}
+\author[29,30]{Luca Monticelli}
+\author[1,31]{Ricky Nencini}
 \author[19,20]{Alexey M. Nesterenko}
@@
 \author[1,24,*]{O. H. Samuli Ollila} %ORCID: 0000-0002- 8728-1006
-%\author[1,2,+]{Christine Author}
-%\author[2,+]{Derek Author}
 
@@
 \affil[4]{University of Potsdam, Institute of Physics and Astronomy, Potsdam-Golm, 14476, Germany}
-%\affil[2]{Affiliation, department, city, postcode, country}
-\affil[5]{Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, P.O. Box 35, Jyv{\"a}skyl{\"a}, 40014 , Finland}
+\affil[5]{Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, Jyv{\"a}skyl{\"a}, 40014, Finland}
 \affil[6]{Departamento de Ciencias B\'{a}sicas, Tecnol\'{o}gico Nacional de M\'{e}xico - ITS Zacatecas Occidente, Sombrerete, Zacatecas, 99102, M\'{e}xico}

### 2023-12-15 main.tex
Update pdfs
Update main.tex

@@
 \author[2,19,20]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
-%\author[2]{...}                   %ORCID: ?
 \author[25]{Cedric Mingham}
@@
 \affil[4]{University of Potsdam, Institute of Physics and Astronomy, Potsdam-Golm, 14476, Germany}
-%\affil[2]{Affiliation, department, city, postcode, country}
 \affil[5]{Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, Jyv{\"a}skyl{\"a}, 40014, Finland}
@@
 \affil[13]{ariadne.ai GmbH (Germany), Häusserstra{\ss}e 3 Heidelberg 69115, Germany }
-\affil[14]{Department of Physical Chemistry of Drugs and Faculty of Pharmacy, Comenius University Bratislava, 832 32 Bratislava, Slovakia}
+\affil[14]{Department of Physical Chemistry of Drugs, Faculty of Pharmacy, Comenius University Bratislava, 832 32 Bratislava, Slovakia}
 \affil[15]{Institute of Organic Chemistry and Biochemistry of the Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2,  Prague, CZ-16610, Czech Republic}
@@
 O.H.S.O, A.M.K, R.N. and L. S. B. acknowledge CSC --- IT Center for Science for computational resources and Academy of Finland (grant nos. 315596, 319902 \& 345631) for financial support.
+N.K. was supported by the Slovak Scientific Grant Agency (VEGA 1/0223/20).
 M.K.

### 2023-12-19 SI.tex
Editorial updates

@@
 
-\renewcommand{\thetable}{S\arabic{table}}%
-\renewcommand{\thefigure}{S\arabic{figure}}%
-\renewcommand{\thesection}{S\arabic{section}}%
-\renewcommand{\thepage}{S\arabic{page}}%
+%\renewcommand{\thetable}{S\arabic{table}}%
+%\renewcommand{\thefigure}{S\arabic{figure}}%
+%\renewcommand{\thesection}{S\arabic{section}}%
+%\renewcommand{\thepage}{S\arabic{page}}%
 
@@
 \author[11]{Alexander Kuzmin}
-\author[26]{Anusha Lalitha} %ORCID: 0000-0002-2817-661X
-\author[27,28]{Fabio Lolicato}
-\author[17,18]{Jesper J. Madsen}  %ORCID: 0000-0003-1411-9080
-\author[2,19,20]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
-\author[25]{Cedric Mingham}
-\author[29,30]{Luca Monticelli}
-\author[1,31]{Ricky Nencini}
-\author[19,20]{Alexey M. Nesterenko}
-\author[21]{Thomas J. Piggot}
-\author[22]{{\'A}ngel Pi{\~n}eiro}
-\author[19,20]{Nathalie Reuter}
-\author[12]{Suman Samantray} %ORCID: 0000-0003-3361-9582
-\author[10,22,23]{Fabi{\'a}n Su{\'a}rez-Lest{\'o}n}
-\author[19,20]{Reza Talandashti}
-\author[1,24,*]{O. H. Samuli Ollila} %ORCID: 0000-0002- 8728-1006
+\author[17]{Anusha Lalitha} %ORCID: 0000-0002-2817-661X
+\author[18,19]{Fabio Lolicato}
+\author[20,21]{Jesper J. Madsen}  %ORCID: 0000-0003-1411-9080
+\author[2,22,23]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
+\author[24]{Cedric Mingham}
+\author[25,26]{Luca Monticelli}
+\author[1,27]{Ricky Nencini}
+\author[22,23]{Alexey M. Nesterenko}
+\author[28]{Thomas J. Piggot}
+\author[29]{{\'A}ngel Pi{\~n}eiro}
+\author[22,23]{Nathalie Reuter}
+\author[12,30]{Suman Samantray} %ORCID: 0000-0003-3361-9582
+\author[10,29,31]{Fabi{\'a}n Su{\'a}rez-Lest{\'o}n}
+\author[22,23]{Reza Talandashti}
+\author[1,32,*]{O. H. Samuli Ollila} %ORCID: 0000-0002- 8728-1006
 
@@
 \affil[2]{Department of Theory and Bio-Systems, Max Planck Institute of Colloids and Interfaces, 14424 Potsdam, Germany}
-\affil[3]{Department of Biomaterials, Max Planck Institute of Colloids and Interfaces, 14424 Potsdam, Germany}
+\affil[3]{Department of Biomedicine, University of Bergen, 5020 Bergen, Norway}
 \affil[4]{University of Potsdam, Institute of Physics and Astronomy, Potsdam-Golm, 14476, Germany}
@@
 \affil[7]{NMR group - Institute for Physics, Martin Luther University Halle-Wittenberg,  Halle (Saale), 06120, Germany}
-\affil[8]{Sorbonne Universit{\'e}, Ecole Normale Sup{\'e}rieure, PSL University, CNRS, Laboratoire des Biomol{\'e}cules (LBM), Paris, 75005, France}
-\affil[9]{Universit{\'e} Paris Cit{\'e}, UFR Sciences du Vivant, Paris, 75013, France}
+\affil[8]{Sorbonne Universit{\'e}, Ecole Normale Sup{\'e}rieure, PSL University, CNRS, Laboratoire des Biomol{\'e}cules (LBM), F-75005 Paris, France}
+\affil[9]{Universit{\'e} Paris Cit{\'e}, F-75006 Paris, France}
 \affil[10]{Center for Research in Biological Chemistry and Molecular Materials (CiQUS), Universidade de Santiago de Compostela,  Santiago de Compostela, E-15782, Spain}
@@
 \affil[13]{ariadne.ai GmbH (Germany), Häusserstra{\ss}e 3 Heidelberg 69115, Germany }
-\affil[14]{Department of Physical Chemistry of Drugs and Faculty of Pharmacy, Comenius University Bratislava, 832 32 Bratislava, Slovakia}
+\affil[14]{Department of Physical Chemistry of Drugs, Faculty of Pharmacy, Comenius University Bratislava, 832 32 Bratislava, Slovakia}
 \affil[15]{Institute of Organic Chemistry and Biochemistry of the Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2,  Prague, CZ-16610, Czech Republic}
 \affil[16]{School of Pharmacy, University of Eastern Finland, 70211 Kuopio, Finland}
-\affil[17]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida 33612, United States of America}
-\affil[18]{Center for Global Health and Infectious Diseases Research, Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida 33612, United States of America}
-\affil[19]{Department of Chemistry, University of Bergen, 5020 Bergen, Norway}
-\affil[20]{Computational Biology Unit, Department of Informatics, University of Bergen, 5020 Bergen, Norway}
-\affil[21]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}
-\affil[22]{Department of Applied Physics, Faculty of Physics, University of Santiago de Compostela, Santiago de Compostela, E-15782, Spain}
-\affil[23]{MD.USE Innovations S.L., Edificio Emprendia, 15782 Santiago de Compostela, Spain}
-\affil[24]{VTT Technical Research Centre of Finland, Espoo, Finland}
-\affil[25]{Hochschule Mannheim, University of Applied Sciences, Mannheim, Germany}
-\affil[26]{Institut Charles Gerhardt Montpellier (UMR CNRS 5253), Universit{\'e} Montpellier, Place Eug{\`e}ne Bataillon, 34095 Montpellier, Cedex 05, France}
-\affil[27]{Heidelberg University Biochemistry Center, 69120 Heidelberg, Germany}
-\affil[28]{Department of Physics, University of Helsinki, FI-00014 Helsinki, Finland}
-\affil[29]{University of Lyon, CNRS, Molecular Microbiology and Structural Biochemistry (MMSB, UMR 5086), F-69007, Lyon, France}
-\affil[30]{Institut National de la Sant{\'e} et de la Recherche M{\'e}dicale (INSERM), Lyon, France}
-\affil[31]{Division of Pharmaceutical Biosciences, Faculty of Pharmacy, University of Helsinki, Helsinki 00014, Finland}
+\affil[17]{Institut Charles Gerhardt Montpellier (UMR CNRS 5253), Universit{\'e} Montpellier, Place Eug{\`e}ne Bataillon, 34095 Montpellier, Cedex 05, France}
+\affil[18]{Heidelberg University Biochemistry Center, 69120 Heidelberg, Germany}
+\affil[19]{Department of Physics, University of Helsinki, FI-00014 Helsinki, Finland}
+\affil[20]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida 33612, United States of America}
+\affil[21]{Center for Global Health and Infectious Diseases Research, Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida 33612, United States of America}
+\affil[22]{Department of Chemistry, University of Bergen, 5007 Bergen, Norway}
+\affil[23]{Computational Biology Unit, Department of Informatics, University of Bergen, 5008 Bergen, Norway}
+\affil[24]{Hochschule Mannheim, University of Applied Sciences, Mannheim, Germany}
+\affil[25]{University of Lyon, CNRS, Molecular Microbiology and Structural Biochemistry (MMSB, UMR 5086), F-69007, Lyon, France}
+\affil[26]{Institut National de la Sant{\'e} et de la Recherche M{\'e}dicale (INSERM), Lyon, France}
+\affil[27]{Division of Pharmaceutical Biosciences, Faculty of Pharmacy, University of Helsinki, Helsinki 00014, Finland}
+
+\affil[28]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}
+
+\affil[29]{Department of Applied Physics, Faculty of Physics, University of Santiago de Compostela, Santiago de Compostela, E-15782, Spain}
+\affil[30]{Institute of Biotechnology, RWTH Aachen University, Worringerweg 3, 52074 Aachen, Germany}
+\affil[31]{MD.USE Innovations S.L., Edificio Emprendia, 15782 Santiago de Compostela, Spain}
+\affil[32]{VTT Technical Research Centre of Finland, Espoo, Finland}
 
@@
 
-\pagebreak
-\section{Processing results in the application layer}
+%\pagebreak
+\section*{Processing results in the application layer}
 
@@
 \pagebreak
-\section{Correlations of area per lipid and thickness with order parameters and form factors}
+\section*{Correlations of area per lipid and thickness with order parameters and form factors}
 
@@
 \pagebreak
-\section{Dependence of form factor and order parameters on the simulation box size}
+\section*{Dependence of form factor and order parameters on the simulation box size}
 
@@
 \pagebreak
-\section{Finding the best models for PC and PE mixtures}
+\section*{Finding the best models for PC and PE mixtures}
 
@@
 \pagebreak
-\section{Testing machine learning models against literature data}
+\section*{Testing machine learning models against literature data}
 \begin{figure}[hb]
@@
 \pagebreak
-\section{Water permeation through membranes}
+\section*{Water permeation through membranes}
 
@@
 \pagebreak
-\section{Water diffusion along membranes}
+\section*{Water diffusion along membranes}
 
@@
 \pagebreak
-\section{Databank content}
+\section*{Databank content}
 
@@
 %\section{Experimental data}
-\section{NMR experiments}
+\section*{NMR experiments}
 Acyl chain order parameters of POPE (Figs.~\ref{POPEexp1} and~\ref{POPEexp2}) and POPG (Figs.~\ref{POPGexp1} and~\ref{POPGexp2}) were analyzed from the same data that were previously recorded to determine headgroup order parameters~\cite{bacle21}.

### 2023-12-19 main.tex
Editorial updates

@@
 \affil[16]{School of Pharmacy, University of Eastern Finland, 70211 Kuopio, Finland}
-\affil[17]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida 33612, United States of America}
-\affil[18]{Center for Global Health and Infectious Diseases Research, Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida 33612, United States of America}
-\affil[19]{Department of Chemistry, University of Bergen, 5020 Bergen, Norway}
-\affil[20]{Computational Biology Unit, Department of Informatics, University of Bergen, 5020 Bergen, Norway}
-\affil[21]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}
-\affil[22]{Department of Applied Physics, Faculty of Physics, University of Santiago de Compostela, Santiago de Compostela, E-15782, Spain}
-\affil[23]{MD.USE Innovations S.L., Edificio Emprendia, 15782 Santiago de Compostela, Spain}
-\affil[24]{VTT Technical Research Centre of Finland, Espoo, Finland}
-\affil[25]{Hochschule Mannheim, University of Applied Sciences, Mannheim, Germany}
-\affil[26]{Institut Charles Gerhardt Montpellier (UMR CNRS 5253), Universit{\'e} Montpellier, Place Eug{\`e}ne Bataillon, 34095 Montpellier, Cedex 05, France}
-\affil[27]{Heidelberg University Biochemistry Center, 69120 Heidelberg, Germany}
-\affil[28]{Department of Physics, University of Helsinki, FI-00014 Helsinki, Finland}
-\affil[29]{University of Lyon, CNRS, Molecular Microbiology and Structural Biochemistry (MMSB, UMR 5086), F-69007, Lyon, France}
-\affil[30]{Institut National de la Sant{\'e} et de la Recherche M{\'e}dicale (INSERM), Lyon, France}
-\affil[31]{Division of Pharmaceutical Biosciences, Faculty of Pharmacy, University of Helsinki, Helsinki 00014, Finland}
+\affil[17]{Institut Charles Gerhardt Montpellier (UMR CNRS 5253), Universit{\'e} Montpellier, Place Eug{\`e}ne Bataillon, 34095 Montpellier, Cedex 05, France}
+\affil[18]{Heidelberg University Biochemistry Center, 69120 Heidelberg, Germany}
+\affil[19]{Department of Physics, University of Helsinki, FI-00014 Helsinki, Finland}
+\affil[20]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida 33612, United States of America}
+\affil[21]{Center for Global Health and Infectious Diseases Research, Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida 33612, United States of America}
+\affil[22]{Department of Chemistry, University of Bergen, 5007 Bergen, Norway}
+\affil[23]{Computational Biology Unit, Department of Informatics, University of Bergen, 5008 Bergen, Norway}
+\affil[24]{Hochschule Mannheim, University of Applied Sciences, Mannheim, Germany}
+\affil[25]{University of Lyon, CNRS, Molecular Microbiology and Structural Biochemistry (MMSB, UMR 5086), F-69007, Lyon, France}
+\affil[26]{Institut National de la Sant{\'e} et de la Recherche M{\'e}dicale (INSERM), Lyon, France}
+\affil[27]{Division of Pharmaceutical Biosciences, Faculty of Pharmacy, University of Helsinki, Helsinki 00014, Finland}
+
+\affil[28]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}
+
+\affil[29]{Department of Applied Physics, Faculty of Physics, University of Santiago de Compostela, Santiago de Compostela, E-15782, Spain}
+\affil[30]{Institute of Biotechnology, RWTH Aachen University, Worringerweg 3, 52074 Aachen, Germany}
+\affil[31]{MD.USE Innovations S.L., Edificio Emprendia, 15782 Santiago de Compostela, Spain}
+\affil[32]{VTT Technical Research Centre of Finland, Espoo, Finland}
 
@@
 
-\section{Introduction}
+\section*{Introduction}
 
@@
 
-\section{Results}
+\section*{Results}
 
-\subsection{NMRlipids overlay databank delivers access to MD simulations of membranes composed of the biologically most abundant lipids}
+\subsection*{NMRlipids overlay databank delivers access to MD simulations of membranes composed of the biologically most abundant lipids}
 
@@
 The distribution of lipids among the available simulations, shown in Fig.~\ref{fig:overlay}B, roughly resembles the biological relative abundance of different lipid types, with phosphatidylcholine (PC) being the most common followed by cholesterol, phosphatidylethanolamine (PE), phosphatidylserine (PS), phosphatidylglycerol (PG), phosphatidylinositol (PI), and other lipids, depending on organism and organelle~\cite{vanmeer08}.
-Abbreviations and full names of all lipids present in the Databank are listed in the NMRlipids databank documentation~\cite{NMRlipidsDocsMolAndMapping}.
+Abbreviations and full names of all lipids present in the Databank are listed in the NMRlipids Databank documentation~\cite{NMRlipidsDocsMolAndMapping}.
 %\url{nmrlipids.github.io/moleculesAndMapping.html}.
-Force fields used in simulations cover all the essential parameter sets commonly used in lipid simulations, see Fig.~\ref{fig:overlay}C and Table~\ref{tab:ForceFields}, including also united atom and polarizable force fields.
+Force fields used in simulations cover all the essential parameter sets commonly used in lipid simulations, see Fig.~\ref{fig:overlay}C and Supplementary Table~\ref{tab:ForceFields}, including also united atom and polarizable force fields.
 Therefore, the averages calculated over the Databank can be considered as mean predictions from available lipid models (average over force field parameters) for an average cell membrane (average over lipid compositions).
@@
 This is demonstrated here with two examples: the NMRlipids Databank graphical user interface (NMRlipids Databank-GUI) at~\href{https://www.databank.nmrlipids.fi}{databank.nmrlipids.fi} and a repository exemplifying novel analyses utilizing NMRlipids Databank as discussed below (\href{https://github.com/NMRlipids/DataBankManuscript}{github.com/NMRlipids/DataBankManuscript}).
-A more detailed description of the NMRlipids Databank structure is available in the supplementary information.
+A more detailed description of the NMRlipids Databank structure is available in the Supplementary Information.
 
@@
     \textbf{A} Schematic presentation of the overlay structure used in the NMRlipids Databank.
-A more detailed structure of the \textit{Databank layer} is shown in Fig.~\ref{DatabankStructure} in the SI.
+A more detailed structure of the \textit{Databank layer} is shown in Supplementary Figure~\ref{DatabankStructure}.
     \textbf{B} Distribution of lipids present in the trajectories of the Databank. `Others' lists lipids occurring in six or fewer simulations. 
-    \textbf{C} Distribution of force fields in the simulations in the Databank. References for each force field are given in Table~\ref{tab:ForceFields}.
+    \textbf{C} Distribution of force fields in the simulations in the Databank. References for each force field are given in Supplementary Table~\ref{tab:ForceFields}.
     \textbf{D} Flowchart for performing an analysis of properties through all MD simulations in the NMRlipids Databank using the API.
@@
 
-\subsection{NMRlipids Databank-GUI: graphical access to the MD simulation data}
+\subsection*{NMRlipids Databank-GUI: graphical access to the MD simulation data}
 NMRlipids Databank-GUI, available at~\href{https://www.databank.nmrlipids.fi}{databank.nmrlipids.fi}, provides easy access to the NMRlipids Databank content through a graphical user interface (GUI).
@@
 
-\subsection{NMRlipids Databank-API: programmatic access to the MD simulation data}\label{section:access}
+\subsection*{NMRlipids Databank-API: programmatic access to the MD simulation data}\label{section:access}
 The NMRlipids Databank-API provides programmatic access to all simulation data in the NMRlipids Databank through application programming interface (API).
@@
 The flowchart in Fig.~\ref{fig:overlay}D illustrates the practical implementation of such an analysis.
-After cloning the databank repository to a local computer, raw data of each simulation can be accessed and analyzed with the help of functions delivered by the NMRlipids Databank-API.
-Analyses over simulations with different naming convetions can be automatically performed with the help of mapping files that associate the specific naming conventions in each simulation with the universal molecule and atom names used by the Databank.
+After cloning the Databank repository to a local computer, raw data of each simulation can be accessed and analyzed with the help of functions delivered by the NMRlipids Databank-API.
+Analyses over simulations with different naming conventions can be automatically performed with the help of mapping files that associate the specific naming conventions in each simulation with the universal molecule and atom names used by the Databank.
 Finally, the analysis results can be stored using the same structure as in the {\it Databank layer}.
-Documentation of the NMRlipids Databank-API and template for new user-defined analyses with further instructions are available at the NMRlipids databank documentation~\cite{NMRlipidsExampleAndTutorials}.
+Documentation of the NMRlipids Databank-API and template for new user-defined analyses with further instructions are available at the NMRlipids Databank documentation~\cite{NMRlipidsExampleAndTutorials}.
 
 While analysis codes and results for basic membrane properties are included in the {\it Databank layer}, unlimited further analyses can be implemented by anyone in separate repositories in the {\it Application layer}.
-When {\it Application layer} repositories are organized by mimicking the {\it Databank layer} structure, they can be accessed programmatically and further analyzed using the tools in the NMRlipids Databank-API by implementing the flowchart demonstrated in Fig.~\ref{fig:resultsProcessingSI}.
+When {\it Application layer} repositories are organized by mimicking the {\it Databank layer} structure, they can be accessed programmatically and further analyzed using the tools in the NMRlipids Databank-API by implementing the flowchart demonstrated in Supplementary Figure~\ref{fig:resultsProcessingSI}.
 Novel analyses that demonstrate the power of NMRlipids Databank in selecting the best simulation models, analysing rare phenomena, and extending MD simulations to new fields are implemented in an {\it Application-layer} repository located at~\href{https://github.com/NMRlipids/DataBankManuscript}{github.com/NMRlipids/DataBankManuscript}.
-The related codes are listed in Table~\ref{tab:codesApplication}.
+The related codes are listed in Supplementary Table~\ref{tab:codesApplication}.
 
-\subsection{Applications of the NMRlipids Databank}
-\subsubsection{Selecting simulation parameters using NMRlipids Databank: Best models for most abundant neutral membrane lipids}
+%\subsection*{Applications of the NMRlipids Databank}
+\subsection*{Selecting simulation parameters using NMRlipids Databank: Best models for most abundant neutral membrane lipids}
 Molecular dynamics (MD) simulations have been particularly useful in understanding membrane systems, although their accuracy has often been compromised by artefacts such as the quality of model parameters~\cite{antila22b,gupta22}.
@@
 
-Here we streamlined this process by defining quantitative quality measures 
-for conformational ensembles of individual lipid molecules and membrane dimensions using C--H bond order parameters from NMR and X-ray scattering form factors~\cite{ollila16}.
+Here we streamlined this process by defining quantitative quality measures for conformational ensembles of individual lipid molecules and membrane dimensions using C--H bond order parameters from NMR and X-ray scattering form factors~\cite{ollila16}.
 These measures enable automatic ranking of lipid bilayer simulations based on their quality against experiments.
@@
 Qualities against X-ray scattering experiments ($FF_q$) were estimated as the difference in the experimental and simulated locations of the first form factor minimum.
-These measures are good proxies for membrane properties because they correlate with the membrane lateral packing and thickness (Figs.~\ref{fig:quality}G,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}).
+These measures are good proxies for membrane properties because they correlate with the membrane lateral packing and thickness (Fig.~\ref{fig:quality}G, Supplementary Figure~\ref{fig:QualityCorrelationsSI}, and Supplementary Figure~\ref{fig:sizedependence}).
 Ergodicity of conformational sampling of lipids was estimated by calculating $\tau_\mathrm{rel}$, the convergence time of the slowest principal component divided by the simulation length.
@@
 Fig.~\ref{fig:quality}A illustrates that predictions for the lateral packing of membranes composed of two most biologically-abundant neutral membrane lipids, POPC and POPE~\cite{vanmeer08}, diverge between different force fields.
-To find the most realistic parameters to simulate membranes with these lipids, we first ranked all simulations based on order parameter quality (Fig.~\ref{fig:top50simulations}), then picked force fields that occurr in Fig.~\ref{fig:quality}A (that is: force fields for both POPC and POPE in the Databank), and then ranked them according to the quality of the \textit{sn}-1 chain of POPC (Fig.~\ref{fig:quality}B) and of POPE (Fig.~\ref{fig:quality}C).
+To find the most realistic parameters to simulate membranes with these lipids, we first ranked all simulations based on order parameter quality (Supplementary Figure~\ref{fig:top50simulations}), then picked force fields that occur in Fig.~\ref{fig:quality}A (that is: force fields for both POPC and POPE in the Databank), and then ranked them according to the quality of the \textit{sn}-1 chain of POPC (Fig.~\ref{fig:quality}B) and of POPE (Fig.~\ref{fig:quality}C).
 Simulations with $\tau_\mathrm{rel}$ clearly above one (larger than 1.3) were discarded in this analysis.
 Because the average \textit{sn}-1 chain order parameter is a good proxy for the membrane packing (Fig.~\ref{fig:quality}G), rankings in Figs.~\ref{fig:quality}B and C can be used to select the simulations giving the most realistic results in Fig.~\ref{fig:quality}A.
-Based on this, Lipid17 and Slipids simulations are most realistic for a POPC membrane, while CHARMM36 and GROMOS-CKP simulations predict overly packed bilayers (overestimated order in Fig.~\ref{fig:POPC_POPE_dataSI}).
+Based on this, Lipid17 and Slipids simulations are most realistic for a POPC membrane, while CHARMM36 and GROMOS-CKP simulations predict overly packed bilayers (overestimated order in Supplementary Figure~\ref{fig:POPC_POPE_dataSI}).
 For POPE, on the other hand, GROMOS-CKP and Slipids are most realistic, while CHARMM36 and Lipid17 predict too packed membranes.
 In conclusion, the quality evaluation based on the NMRlipids Databank suggests that the Slipids parameters are the best currently available choice for simulations with PC and PE lipids, at least for applications where membrane packing is relevant.
-Also direct comparisons with the experimental data for the most relevant simulations are shown in Figs.~\ref{fig:quality}E--F and~\ref{fig:POPC_POPE_dataSI} a.
+Also direct comparisons with the experimental data for the most relevant simulations are shown in Figs.~\ref{fig:quality}E--F and Supplementary Figure~\ref{fig:POPC_POPE_dataSI} a.
 Fig.~\ref{fig:quality}D shows the overall highest-ranked simulation, POPC bilayer with OPLS3e parameters, for the reference.
@@
     \textbf{D--F} Direct comparison against experimental (NMR order parameters and X-ray scattering ) data exemplified for a simulation with the best overall order parameter quality (D), the best quality for POPE lipid (E), and the headgroup quality for POPE (F).
+Error bars for simulations are standard error of the mean over different lipids (n=number of lipids in a simulation shown in B).
+Error bars for experiments are 0.02~\cite{ollila16}.
     \textbf{G} Scatter plots and Pearson correlation coefficients, $r$, for the membrane area per lipid, thickness, first minimum of X-ray scattering form factor and average order parameter of the \textit{sn}-1 acyl chain extracted from the NMRlipids Databank.
-All correlation coefficients have p-value below 0.001.
-For more correlations see Fig.~\ref{fig:QualityCorrelationsSI}.
+All correlation coefficients have p-value below 0.001 with two-sided test.
+For more correlations see Supplementary Figure~\ref{fig:QualityCorrelationsSI}.
     }
@@
 
-\subsubsection{Using NMRlipids Databank as a training set for machine learning applications: Predicting multi-component membrane properties}
+\subsection*{Using NMRlipids Databank as a training set for machine learning applications: Predicting multi-component membrane properties}
 
-To demonstrate the usage of the NMRlipids databank to construct machine learning (ML) models that predict membrane properties, we trained a model that predicts area per lipids and thicknesses of membranes with diverse compositions.
-First we used randomly selected 80\% of the databank simulations to choose the hyperparameters for, and optimize, a set of machine learning models with the goal of predicting the area per lipid from the membrane composition.
+To demonstrate the usage of the NMRlipids Databank to construct machine learning (ML) models that predict membrane properties, we trained a model that predicts area per lipids and thicknesses of membranes with diverse compositions.
+First we used randomly selected 80\% of the Databank simulations to choose the hyperparameters for, and optimize, a set of machine learning models with the goal of predicting the area per lipid from the membrane composition.
 
- After the parameter optimization, we tested predictions from different ML models for the area per lipid both against the remaining 20\% of the databank and against area per lipids reported from simulations of membranes containing mixtures of POPC, POPE, POPS, PI, sphingomyelin lipids and cholesterol~\cite{shahane19,kumar21,oliveira22} that are not included in the databank.
-Essential differences between models were not observed when predicting area per lipids of 20\% of simulations selected as the test set, but linear regression and Ridge models gave the best correlations with the literature data (predictions  from the linear regression model are shown in Fig.~\ref{fig:LinearRegressionTest} and from other models in Fig.~\ref{fig:MLTest} in the supplementary information).
+ After the parameter optimization, we tested predictions from different ML models for the area per lipid both against the remaining 20\% of the Databank and against areas per lipid reported from simulations of membranes containing mixtures of POPC, POPE, POPS, PI, sphingomyelin lipids, and cholesterol~\cite{shahane19,kumar21,oliveira22} that are not included in the Databank.
+Essential differences between models were not observed when predicting area per lipids of 20\% of simulations selected as the test set, but linear regression and Ridge models gave the best correlations with the literature data (predictions  from the linear regression model are shown in Fig.~\ref{fig:LinearRegressionTest} and from other models in Supplementary Figure~\ref{fig:MLTest}).
 The linear regression model was selected for further studies due to its simplicity.
@@
     \includegraphics[width=88mm]{Figures/MLtest/LinearRegressiontest.pdf}
-    \caption{Predictions of area per lipids of multi-component membranes composed of POPC, POPE, POPS, PI, sphingomyelin lipids and cholesterol from linear regression model against literature data from simulations (green~\cite{shahane19}, blue~\cite{kumar21}, and red~\cite{oliveira22}).
-Black line indicates x=y.}
+    \caption{Predictions of areas per lipid (APL) of multi-component membranes composed of POPC, POPE, POPS, PI, sphingomyelin lipids, and cholesterol from linear regression model against literature data from simulations (green~\cite{shahane19}, blue~\cite{kumar21}, and red~\cite{oliveira22}).
+Error bars are from the publications as the values.
+Black line indicates $x=y$.}
     \label{fig:LinearRegressionTest}
@@
 
-To demonstrate the usefulness of the constructed models for understanding multi-component membrane properties, we predicted later packing (ares per lipids) and membrane thicknesses of common biological membranes based on their lipid compositions reported in the literature, see table~\ref{tab:MLpredictions}.
+To demonstrate the usefulness of the constructed models for understanding multi-component membrane properties, we predicted later packing (areas per lipid) and membrane thicknesses of common biological membranes based on their lipid compositions reported in the literature, see table~\ref{tab:MLpredictions}.
 The model predicts substantial 50\% difference in area per lipid between most densely (influenza virus) and loosely (mitochondria) packed membranes.
-Difference of 0.8\,nm in thickness is predicted between thinnest (bacterial) and thickest (plasma) membranes.
-Such differences are expected to effect on many biologically relevant functions of membranes, such as permeation, cholesterol flip-flops (see next sections), and interactions with proteins~\cite{phillips09}, demonstrating that NMRlipids databank can be used to give valuable insights on biologically relevant properties of complex biological membranes.
+Difference of 0.8\,nm in thickness is predicted between the thinnest (bacterial) and thickest (plasma) membranes.
+Such differences are expected to effect on many biologically relevant functions of membranes, such as permeation, cholesterol flip-flops (see next sections), and interactions with proteins~\cite{phillips09}, demonstrating that NMRlipids Databank can be used to give valuable insights on biologically relevant properties of complex biological membranes.
 Most importantly, the delivered programmatic access to increasing amount of MD simulation data enables training of machine learning models that predict various membrane properties for all.
@@
     \begin{table}[htb]
-    \caption{Area per lipids (APL) and membrane thicknesses predicted by the linear regression model trained using the NMRlipids databank for membrane compositions corresponding different biological membranes.
+    \caption{Areas per lipid (APL) and membrane thicknesses predicted by the linear regression model trained using the NMRlipids Databank for membrane compositions corresponding different biological membranes.
 Systems are sorted in the order of increasing membrane packing.
-Compositions of different membranes are estimated based on references given in composition column.}
+Compositions of different membranes are estimated based on references given in the composition column.}
         \centering
@@
 
-\subsubsection{Detecting rare phenomena using NMRlipids Databank: Cholesterol flip-flops}
+\subsection*{Detecting rare phenomena using NMRlipids Databank: Cholesterol flip-flops}
 Lipid flip-flops from one bilayer leaflet to another play an important role in lipid trafficking and regulating membrane properties~\cite{vanmeer08}.
@@
 
-\subsubsection{Extending the scope of MD simulations to new fields using NMRlipids Databank: Water diffusion anisotropy in membrane systems}
+\subsection*{Extending the scope of MD simulations to new fields using NMRlipids Databank: Water diffusion anisotropy in membrane systems}
 The anisotropic diffusion of water and hydrophilic molecules in directions parallel and perpendicular to membranes is an important parameter in models describing the translocation of drugs through biological material, particularly in the skin \cite{hansen13,wen18,nitsche19,roberts21}.
@@
     \textbf{F} Lateral diffusion of water as a function of hydration level. Experimental points for DMPC bilayers at 313\,K at different hydration levels are shown~\cite{rudakova04}.
-    \textbf{G--H} Diffusion anistoropy of water as a function of thickness and area per lipid. 
+    \textbf{G--H} Diffusion anisotropy of water as a function of thickness and area per lipid. 
     Non-zero permeation and diffusion values from simulations are shown with blue dots.
@@
 
-\section{Discussion}
+\section*{Discussion}
 
@@
 
-    Reviewers and publishers & Tool to facilitate FAIRness of data, transparency and reproducibility & All publicly available data can be included in an overlay databank irrespectively on its location \\
+    Reviewers and publishers & Tool to facilitate FAIRness of data, transparency and reproducibility & All publicly available data can be included in an overlay databank irrespectively of its location \\
 
@@
 %TC:ignore
-\section{Methods}
+\section*{Methods}
 
-\subsection{Structure of the Databank}
+\subsection*{Structure of the Databank}
 The overlay structure designed for the NMRlipids Databank is composed of three layers (Fig.~\ref{fig:overlay}A).
@@
 The README.yaml files in these folders contain access to all information that is needed for further analysis of simulations, such as links to the raw data and associations with the universal molecule and atom names.
-The content of these files is described in detail in Table~\ref{tab:READMEkeys} in the supplementary information and in the NMRlipids databank documentation \href{https://nmrlipids.github.io/}{nmrlipids.github.io/}.
+The content of these files is described in detail in Supplementary Table~\ref{tab:READMEkeys} and in the NMRlipids Databank documentation \href{https://nmrlipids.github.io/}{nmrlipids.github.io}.
 Results from analyses of basic membrane properties (area per lipid, thickness, C--H bond order parameters, X-ray scattering form factors, and relaxation of principal components) are stored in the same folders as the README.yaml files.
@@
 The scripts in the NMRlipids Databank are mainly written in Python and many of them use the MDAnalysis module~\cite{gowers2019mdanalysis,michaud2011mdanalysis}.
-The Databank structure is illustrated in more detail in Fig.~\ref{DatabankStructure} in the supplementary information.
+The Databank structure is illustrated in more detail in Supplementary Figure~\ref{DatabankStructure}.
 Whenever specific files or folders are referred here, they locate at the {\it Databank layer} repository unless stated otherwise.
 
-\subsection{Universal naming convention for molecules and atoms} \label{naming}
+\subsection*{Universal naming convention for molecules and atoms} \label{naming}
 When analysing simulation trajectories, atoms and molecules often need to be  called by the names used in the trajectory.
@@
 
-\subsection{Adding data to NMRlipids Databank}
+\subsection*{Adding data to NMRlipids Databank}
 The NMRlipids Databank is open for additions of simulation data by anyone.
@@
 Rest of the information to be stored in the README.yaml files will be then automatically extracted using the \texttt{/Scripts/BuildDatabank/AddData.py} script.
-The required manually entered and automatically extracted information are described in detail in Table~\ref{tab:READMEkeys} in the supplementary information and in the NMRlipids databank documentation~\cite{NMRlipidsREADMEfiles}.
+The required manually entered and automatically extracted information are described in detail in Supplementary Table~\ref{tab:READMEkeys} and in the NMRlipids Databank documentation~\cite{NMRlipidsREADMEfiles}.
 The documentation includes also detailed instructions to add data.
@@
 
-\subsection{Experimental data}
+\subsection*{Experimental data}
 Experimental data used in the quality evaluation, currently composed of C--H bond order parameters and X-ray scattering form factors, are stored in \texttt{/Data/experiments} in the NMRlipids Databank repository.
 Similarly to simulations, each experimental data set has a README.yaml file containing all the relevant information about the experiment.
-The keys and their descriptions for the experimental data, as well as detailed instructions to add the date, are given in Table~\ref{tab:READMEkeysEXP} and in the NMRlipids databank documentation~\cite{NMRlipidsExperimentalData}.
+The keys and their descriptions for the experimental data, as well as detailed instructions to add the date, are given in Table~\ref{tab:READMEkeysEXP} and in the NMRlipids Databank documentation~\cite{NMRlipidsExperimentalData}.
 The NMR data currently in the NMRlipids Databank are taken from Refs.~\citenum{scherer87,ferreira13,antila19,melcr20,bacle21,hanashima19,doktorova20} and the X-ray scattering data from Refs.~\citenum{Kucerka08a,kucerka11,pan12b,pan14,kucerka15,NMRlipidsIII,doktorova20}.
-In addition, previously unpublished NMR data for POPE, POPG, and DOPC was acquired as described in the supplementary information (Figs.~\ref{POPEexp1}--\ref{fig:DOPCexp}) and contributed to the Databank.
+In addition, previously unpublished NMR data for POPE, POPG, and DOPC was acquired as described in  Supplementary Figures~\ref{POPEexp1}--\ref{fig:DOPCexp}) and contributed to the Databank.
 
-\subsection{Analysing simulations}
+\subsection*{Analysing simulations}
 In practise, simulations in the NMRlipids Databank can be analyzed by executing a program that
@@
 The equilibration period given by the user (TIMELEFTOUT in Table~\ref{tab:READMEkeys}) is discarded from the trajectories in all analysis codes.
-For further details, see the NMRlipids databank documentation \href{https://nmrlipids.github.io/}{nmrlipids.github.io}.
+For further details, see the NMRlipids Databank documentation \href{https://nmrlipids.github.io/}{nmrlipids.github.io}.
 
-\subsection{Principal Components Analysis of equilibration of simulations}
+\subsection*{Principal Components Analysis of equilibration of simulations}
 To estimate how well conformational ensembles of lipids are converged in trajectories, the Principal Component Analysis (PCA) following the PCALipids protocol was used \cite{buslaev16,buslaev2020principal}.
@@
 
-\subsection{Calculation of C--H bond order parameters}
+\subsection*{Calculation of C--H bond order parameters}
 The C--H bond order parameters were calculated directly from the carbon and hydrogen positions using the definition
@@
 
-\subsection{Calculation of X-ray scattering form factors}
+\subsection*{Calculation of X-ray scattering form factors}
 X-ray scattering form factors were calculated using the standard equation for lipid bilayers that does not assume symmetric membranes~\cite{ollila16},
@@
 
-\subsection{Calculation of area per lipid and bilayer thickness}
+\subsection*{Calculation of area per lipid and bilayer thickness}
 Area per lipids of bilayers were calculated by dividing the time-averaged area of the simulation box with the total number of lipids and surfactant molecules in the simulation.The script that calculates the area per lipid from all simulations in the NMRlipids Databank repository is available at \texttt{Scripts/AnalyzeDatabank/calcAPL.py} in the NMRlipids Databank repository.
@@
 
-\subsection{Quality evaluation of C--H bond order parameters}
+\subsection*{Quality evaluation of C--H bond order parameters}
 As the first step to evaluate simulation qualities against experimental data, a simulation is connected to an experimental data set if the molar concentrations of all molecules are within $\pm$3 percentage units, charged lipids have the same counterions, and temperatures are within $\pm$2\,K.
@@
 
-\subsection{Quality evaluation of X-ray scattering form factors}
+\subsection*{Quality evaluation of X-ray scattering form factors}
 Because experiments give form factors only on a relative intensity scale, they should be scaled before comparing with the simulation data.
@@
 
-\subsection{Training machine learning model to predict properties of multi-component membranes}
+\subsection*{Training machine learning model to predict properties of multi-component membranes}
 Different machine learning approaches were tested to find the best model to predict the area per lipids of multi-component membranes  from their lipid composition.
@@
 The models were trained to predict the area per lipid for a membrane with the given molar fractions of membrane lipids.
-Most abundant lipids in the databank, POPC, POPE, POPG, POPS and cholesterol, were considered explicitly, while SM, cardiolipins and PI lipids with  more sparse data were all grouped together based on their headgroups independently on acyl chain content.
+Most abundant lipids in the Databank, POPC, POPE, POPG, POPS, and cholesterol, were considered explicitly, while SM, cardiolipins, and PI lipids with  more sparse data were all grouped together based on their headgroups independently on acyl chain content.
 Salt concentration, temperature and other conditions were not considered.
@@
 The best performing hyperparameters were then used to fit each model to the training set.
-The performance of the models was then tested on the remaining 20\% of the data extracted from the databank, and additionally, by predicting area per lipids from multi-component membrane simulations reported in the literature but not included in the databank~\cite{shahane19,kumar21,oliveira22}.
+The performance of the models was then tested on the remaining 20\% of the data extracted from the Databank, and additionally, by predicting area per lipids from multi-component membrane simulations reported in the literature but not included in the Databank~\cite{shahane19,kumar21,oliveira22}.
 Due to its  simplicity and good performance in these tests, we selected to use linear regression model for further studies.
@@
 In further applications, it is important to consider potential limitations arising from the currently limited amount of data for certain types of systems.
-For example, the current models have been trained and tested only for mixtures with palmitoyl and oleyl acyl chains because the databank does not yet contain enough data with varying number of double bonds.
+For example, the current models have been trained and tested only for mixtures with palmitoyl and oleoyl acyl chains because the Databank does not yet contain enough data with varying number of double bonds.
 Such limitations are expected to alleviate with increasing amount of data in the future.
 
-\subsection{Calculation of lipid flip-flops}
+\subsection*{Calculation of lipid flip-flops}
 Flip-flop rates were calculated using the \texttt{AssignLeaflets} and \texttt{FlipFlop} tools from the LiPyphilic package~\cite{LiPyphilic2021}.
@@
 
-\subsection{Analysing anisotropic diffusion of water in a membrane environment from NMRlipids Databank}
+\subsection*{Analysing anisotropic diffusion of water in a membrane environment from NMRlipids Databank}
 
@@
 
-\subsection{Data availability}
+\subsection*{Data availability}
 All data related to this study are available via git repositories \url{https://github.com/NMRlipids/Databank/} (DOI: 10.5281/zenodo.7875567) and \url{https://github.com/NMRLipids/DataBankManuscript/} (DOI: 10.5281/zenodo.10252380).
+These repositories contain also source data files for all figures.
+Figures can be regenerated with codes mentioned in Methods section separately for each analysis.
+For instructions to access the data programmatically, see \url{https://nmrlipids.github.io/}.
 
@@
 
-\subsection{Code availability}
+\subsection*{Code availability}
 All codes used in this study are available via git repositories \url{https://github.com/NMRlipids/Databank/} (DOI: 10.5281/zenodo.7875567) and \url{https://github.com/NMRLipids/DataBankManuscript/} (DOI: 10.5281/zenodo.10252380).
@@
 P.B. was supported by the Academy of Finland (Grants 311031 and 342908).
-R.G.-F, A.P. and F.S.-L. thank Centro de Supercomputación de Galicia for computational support.
-R.G.-F.
-thanks Ministerio de Ciencia, Innovaci{\'o}n y Universidades for a ''Ram{\'o}n y Cajal'' contract (RYC-2016-20335), and Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (RTI2018-098795-A-I00, PDC2022-133402-I00), Xunta de Galicia and the ERDF (ED431F 2020/05, 02\_IN606D\_2022\_2667887 and Centro singular de investigaci{\'o}n de Galicia accreditation 2016-2019, ED431G/09).
-A.P. thanks Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (PID2019-111327GB-I00, PDC2022-133402-I00), Xunta de Galicia and the ERDF (ED431B 2022/36, 02\_IN606D\_2022\_2667887).
-F.S.-L. thanks Axencia Galega de Innovaci{\'o}n for his predoctoral contract (02\_IN606D\_2022\_2667887), and Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (PID2019-111327GB-I00).
 F.F.-R.
 acknowledges Tecnol{\'o}gico Nacional de M{\'e}xico, Direcci{\'o}n General de Asuntos del Personal Acad{\'e}mico (DGAPA), Programa de Apoyo a Proyectos de Investigaci{\'o}n e Innovaci{\'o}n Tecnol{\'o}gica (PAPIIT) 101923, CONACyT Ciencia de Frontera 74884, for financial support and Miztli-Direcci{\'o}n de C{\'o}mputo y de Tecnolog{\'i}as de Informaci{\'o}n y Comunicaci{\'o}n (DGTIC) - Universidad Nacional Aut{\'o}noma de M{\'e}xico (UNAM) (Project LANCAD-UNAM-DGTIC-057) facilities for computing time allocation.
-M.S.M. was supported by the Trond Mohn Foundation BFS2017TMT01.
-O.H.S.O, A.M.K, R.N. and L. S. B. acknowledge CSC --- IT Center for Science for computational resources and Academy of Finland (grant nos. 315596, 319902 \& 345631) for financial support.
+T.M.F. greatly acknowledges financial support by the Ministry of Economics, Science and Digitalisation of the State of Saxony-Anhalt.
+R.G.-F, A.P., and F.S.-L. thank Centro de Supercomputación de Galicia for computational support;
+R.G.-F.
+thanks Ministerio de Ciencia, Innovaci{\'o}n y Universidades for a ''Ram{\'o}n y Cajal'' contract (RYC-2016-20335), and Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (RTI2018-098795-A-I00, PDC2022-133402-I00), Xunta de Galicia and the ERDF (ED431F 2020/05, 02\_IN606D\_2022\_2667887 and Centro singular de investigaci{\'o}n de Galicia accreditation 2016-2019, ED431G/09);
+A.P. thanks Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (PID2019-111327GB-I00, PDC2022-133402-I00), Xunta de Galicia and the ERDF (ED431B 2022/36, 02\_IN606D\_2022\_2667887);
+F.S.-L. thanks Axencia Galega de Innovaci{\'o}n for his predoctoral contract (02\_IN606D\_2022\_2667887), and Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (PID2019-111327GB-I00).
 N.K. was supported by the Slovak Scientific Grant Agency (VEGA 1/0223/20).
-M.K.
-acknowledge CSC –-- IT Center for Science for computational resources and thanks Finnish Cultural Foundation and the UEF Doctoral Programme for financial support.
-T.M.F.
-greatly acknowledges financial support by the Ministry of Economics, Science and Digitalisation of the State of Saxony-Anhalt.
+M.K. acknowledges CSC --- IT Center for Science for computational resources and thanks Finnish Cultural Foundation and the UEF Doctoral Programme for financial support.
 F.L. was supported by the Deutsche Forschungsgemeinschaft (DFG LO 2821/1-1).
+M.S.M. was supported by the Trond Mohn Foundation (BFS2017TMT01).
+L.M. acknowledges funding by the Institut National de la Sant\'e et de la Recherche M\'edicale (INSERM).
+O.H.S.O, A.M.K, R.N., and L. S. B. acknowledge CSC --- IT Center for Science for computational resources and Academy of Finland (grant nos. 315596, 319902 \& 345631) for financial support. 
 We acknowledge all the NMRlipids Project contributors for making development of the NMRlipids Databank possible.
-L.M. acknowledges funding by the Institut National de la Santé et de la Recherche Médicale (INSERM).
 
@@
 H.S.A.
-contributed to the quality evaluation metrics based on NMR and X-ray scattering form factors, wrote the machine learning code, and was involved in testing the GUI, polishing the manuscript, and in community outreach to source additional data.
+contributed to the quality evaluation metrics based on NMR and X-ray scattering form factors, wrote the machine learning code, and was involved in testing the graphical user interface (GUI), polishing the manuscript, and in community outreach to source additional data.
 L.S.B. implemented roughly 630 pre-existing lipid bilayer simulations into the Databank and created the corresponding mapping files.
@@
 P.F.J.F. contributed to changes in the buildH software so that the Databank can handle united atom simulation data.
-R.G.-F. was responsible for the design, validation and supervision of the back-end and front-end for the graphical user interface (GUI) development (databank.nmrlipids.fi).
+R.G.-F. was responsible for the design, validation and supervision of the back-end and front-end for the GUI development (\href{https://www.databank.nmrlipids.fi}{databank.nmrlipids.fi}).
 I.G. advised on implementation of the PCA-based equilibration metric and supervised the work of A.K.
@@
 N.K. contributed experimental scattering data. % and SIMtoEXP algorithm.
-P.K. Implemented the flip-flop analysis and supported the integration of PCA equilibration metric into the databank.
+P.K. implemented the flip-flop analysis and supported the integration of PCA equilibration metric into the Databank.
 M.K. provided simulations with the OPLS3e force field. 
 A.K. implemented the PCA-based equilibration metric.
-A.L. provided simulations of heterogeneous lipid membranes including sphingolipids and glycosphingolipids with CHARMM36 force field.
-F.S.-L. developed the back-end and participated in the design and development and validation of the graphical user interface (GUI) (databank.nmrlipids.fi).
+A.L. provided simulations of heterogeneous lipid membranes including sphingolipids and glycosphingolipids with the CHARMM36 force field.
 F.L. provided simulations using the CHARMM36m force field, incorporating cholesterol POPC and SAPI25 lipids.
@@
 and M.K.
-C.M. worked on the databank implementation and contributed code improvements.
-L.M. performed simulations (for sphingomyelin, CHARMM36 force field, and Orange force field)
+C.M. worked on the Databank implementation and contributed code improvements.
+L.M. performed simulations (for sphingomyelin, CHARMM36 force field, and Orange force field).
 R.N. implemented the code for the calculations of X-ray scattering form factors.
-A.N. provided simulations with TLCL lipid, simulations with Reaction-Field instead of PME, and improved the implementation.
+A.M.N. provided simulations with TLCL lipid, simulations with Reaction-Field instead of PME, and improved the implementation.
 T.J.P. performed many of the all-atom and united-atom simulations and contributed to discussions on the project.
-A.P. was responsible for the design, validation and supervision of the back-end and front-end for the graphical user interface (GUI) development (databank.nmrlipids.fi).
-S.S. contributed simulations of heterogeneous lipid membranes including sphingomyelin and ceramides with CHARMM36 force field.
-R.T. Provided simulations of complex lipid-bilayers with charmm36m force field.
-N.R. contributed simulations of complex lipid-bilayers obtained with NAMD and the charmm36 force field
+A.P. was responsible for the design, validation and supervision of the back-end and front-end for the GUI development (\href{https://www.databank.nmrlipids.fi}{databank.nmrlipids.fi}).
+N.R. contributed simulations of complex lipid-bilayers obtained with NAMD and the CHARMM36 force field.
+S.S. contributed simulations of heterogeneous lipid membranes including sphingomyelin and ceramides with the CHARMM36 force field.
+F.S.-L. developed the back-end and participated in the design and development and validation of the GUI (\href{https://www.databank.nmrlipids.fi}{databank.nmrlipids.fi}).
+R.T. provided simulations of complex lipid-bilayers with the CHARMM36m force field.
 O.H.S.O. designed and supervised the work, implemented the water permeation, diffusion and machine learning analyses, and wrote the manuscript.
 
-\section*{Additional information}
+\section*{Competing interests}
+The authors declare no competing interests.
 
+\section*{Tables}
+
+  \noindent\begin{minipage}{\textwidth}
+    \textbf{Table 1.} Areas per lipid (APL) and membrane thicknesses predicted by the linear regression model trained using the NMRlipids Databank for membrane compositions corresponding different biological membranes.
+Systems are sorted in the order of increasing membrane packing.
+Compositions of different membranes are estimated based on references given in the composition column.
+        \centering
+        \begin{tabular}{cccc}
+             Membrane&  Composition [lipid(\%-fraction)]&  APL[\AA$^2$] & Thickness [nm$^2$]\\
+             \hline
+             Mitochondria & POPC(37) : POPE(31) : PI(6) : CL(22) : 
+             CHOL(4)~\cite{vanmeer08,escriba15,casares19} & 74 (66)$^*$ & 4.4 \\
+             Bacterial & POPC(20) : POPE(35) : POPG(35) : CL(5) : CHOL(5)~\cite{chwastek20}  & 61 (60)$^*$ & 4.2 \\
+             ER &  POPC(54) : POPE(20) : POPS(4) : PI(11) : SM(4) : CHOL(8)~\cite{vanmeer08,escriba15,casares19} & 57 & 4.6 \\
+             Golgi & POPC(36) : POPE(21) : POPS(6) : PI(12) : SM(7) : CHOL(18)~\cite{vanmeer08,escriba15,casares19} & 52 & 4.8 \\
+             Plasma & POPC(23) : POPE(11) : POPS(8) : PI(7) : SM(17) : CHOL(34)~\cite{vanmeer08,escriba15,casares19} & 45 & 5.0 \\
+             Synaptic & POPC(28) : POPE(20) : POPS(9) : PI(2) : SM(4) : CHOL(37)~\cite{binotti21}  &  45 & 4.8 \\
+             Influenza & POPC(5) : POPE(32) : POPS(15) : SM(9) : CHOL(40)~\cite{gerl12,ivanova15} & 42 & 4.8 \\
+        \end{tabular}
+        \label{tab:MLpredictions}
+
+        $^*$ Value in parenthesis is the area per acyl chain taking into account that cardiolipin has four chains per molecule while other lipids have two. 
+    \end{minipage}
+
+\vspace{1cm}
+
+  \noindent\begin{minipage}{\textwidth}
+  \textbf{Table 2.} Examples of impact of NMRlipids and other overlay databanks in different disciplines.
+    \centering
+    \begin{tabular}{p{3.0cm}  p{4.0cm}  p{7.0cm}}
+     Target group  &  Outcomes     & Practical examples \\
+    \hline
+    Computational and data scientists & Databanks with programmatic access, training sets for machine learning applications, automatic optimization of simulation models & Programmatic access to all publicly available MD simulation trajectories, machine learning models predicting properties of complex biomolecular assemblies (Table~\ref{tab:MLpredictions}), optimizing force field and model parameters from atomistic to continuum scale \\
+
+    Biomedical scientists & Applications of biomolecular modeling in new fields, coupling omics to biomolecular structure and dynamics & Anisotropic water diffusion for pharmacokinetics and MRI imaging applications (Fig.~\ref{fig:permeability}), properties of complex cellular structures based on composition analyses from omics \\
+
+    Material scientists & Predictions of complex biomolecular material properties from data-driven and machine learning models & Optimising composition of lipid nanoparticle formulations for desired properties, predictions of bioinspired material properties \\
+
+    Biophysicists &  Novel analyses and predictions for properties of biomolecular assemblies and their compositions  & Analyses of rare phenomena (for example, Lipid flip-flops and water permeation in Figs.~\ref{fig:flip-flops} and~\ref{fig:permeability}), data-driven analyzes and machine learning models  predicting correlations between properties and compositions of biomolecular assemblies (Figs.~\ref{fig:quality} and~\ref{fig:permeability}) \\
+
+    Students and teachers & Graphical and programmatic access to biomolecular structure and dynamics & Illustration of disorder and dynamics in biomolecules, example data for bioinformatics and data-science courses \\
+
+    Reviewers and publishers & Tool to facilitate FAIRness of data, transparency and reproducibility & All publicly available data can be included in an overlay databank irrespectively of its location \\
+
+    \end{tabular}
+\end{minipage}
+
+\section*{Figure captions}
+\noindent
+\textbf{Figure 1.} Overview of the NMRlipids Databank. 
+    \textbf{A} Schematic presentation of the overlay structure used in the NMRlipids Databank.
+A more detailed structure of the \textit{Databank layer} is shown in Fig.~\ref{DatabankStructure} in the SI.
+    \textbf{B} Distribution of lipids present in the trajectories of the Databank. `Others' lists lipids occurring in six or fewer simulations. 
+    \textbf{C} Distribution of force fields in the simulations in the Databank. References for each force field are given in Table~\ref{tab:ForceFields}.
+    \textbf{D} Flowchart for performing an analysis of properties through all MD simulations in the NMRlipids Databank using the API.
+    \textbf{E} Currently available lipid mixtures in the NMRlipids Databank. Colorbar shows the number of available simulations with the darkest green indicating three or more.
+
+\vspace{1cm}
+    \noindent
+\textbf{Figure 2.} Examples of data obtained using NMRlipids Databank. 
+    \textbf{A} Area per lipid of POPC and POPE lipid bilayers predicted by different force fields at 310\,K in simulations that are available in the NMRlipids Databank.
+The data points from the best-performing simulations, based on rankings in panels B and C, are surrounded by black circles.
+    \textbf{B} Best POPC simulations ranked based on the \textit{sn}-1 acyl chain order parameter quality ($P^\mathrm{sn1}$).
+Also \textit{sn}-2 acyl chain ($P^\mathrm{sn2}$), headgroup ($P^\mathrm{hg}$) and total ($P^\mathrm{total}$) order parameter qualities, form factor quality ($FF_q$), and relative equilibration time for conformations ($\tau_\mathrm{rel}$) are shown.
+Note that the best possible order parameter quality is one, while the best possible form factor quality is zero.
+    \textbf{C} Best POPE simulations ranked based on the \textit{sn}-1 acyl chain order parameter quality. 
+    \textbf{D--F} Direct comparison against experimental (NMR order parameters and X-ray scattering ) data exemplified for a simulation with the best overall order parameter quality (D), the best quality for POPE lipid (E), and the headgroup quality for POPE (F).
+    \textbf{G} Scatter plots and Pearson correlation coefficients, $r$, for the membrane area per lipid, thickness, first minimum of X-ray scattering form factor and average order parameter of the \textit{sn}-1 acyl chain extracted from the NMRlipids Databank.
+All correlation coefficients have p-value below 0.001.
+For more correlations see Fig.~\ref{fig:QualityCorrelationsSI}.
+
+    \vspace{1cm}
+    \noindent
+\textbf{Figure 3.}
+Predictions of areas per lipid of multi-component membranes composed of POPC, POPE, POPS, PI, sphingomyelin lipids, and cholesterol from linear regression model against literature data from simulations (green~\cite{shahane19}, blue~\cite{kumar21}, and red~\cite{oliveira22}).
+Black line indicates $x=y$.
+
+\vspace{1cm}
+    \noindent
+\textbf{Figure 4.}
+Quantification of cholesterol flip-flop events in NMRlipids Databank simulations. 
+    \textbf{A} Illustration of cholesterol flip-flop.  
+      \textbf{B--D} Cholesterol flip-flops analyzed from the Databank as a function of membrane thickness, area per lipid, and acyl chain order.
+Values from simulations with non-zero flip-flop rates are shown with blue dots.
+Histogrammed values are shown with black dots.
+For the mean value in each bin, average weighted with the simulation lengths was used, and error bars show the standard error of the mean.
+
+\vspace{1cm}
+    \noindent
+\textbf{Figure 5.}
+Quantification of water diffusion in NMRlipids Databank simulations.
+    \textbf{A} Water diffusion, $D_\perp$, and permeability, $P$, through membranes, and lateral diffusion along the membrane, $D_\parallel$, illustrated in a multilamellar stack of lipid bilayers. 
+    \textbf{B--E} Water permeation through membranes analyzed from the Databank as a function of temperature, thickness, area per lipid, and acyl chain order.
+Inset in B) shows the Arrhenius plot of permeation ($\ln(P)$ vs.
+$1/T$) that gives 14\,$\pm$\,3\,$k_\mathrm{B}T$ for the average activation energy for water permeation through lipid bilayer.
+    \textbf{F} Lateral diffusion of water as a function of hydration level. Experimental points for DMPC bilayers at 313\,K at different hydration levels are shown~\cite{rudakova04}.
+    \textbf{G--H} Diffusion anisotropy of water as a function of thickness and area per lipid. 
+    Non-zero permeation and diffusion values from simulations are shown with blue dots.
+Histogrammed values are shown with black dots.
+For the mean value in each bin, average weighted with the simulation lengths was used, and error bars show the standard error of the mean.
+Only bins with more than one microsecond of data in total were used for water permeation.
 %To include, in this order: \textbf{Accession codes} (where applicable); \textbf{Competing interests} (mandatory statement).

### 2023-12-19 main.tex
Updates

@@
 All codes used in this study are available via git repositories \url{https://github.com/NMRlipids/Databank/} (DOI: 10.5281/zenodo.7875567) and \url{https://github.com/NMRLipids/DataBankManuscript/} (DOI: 10.5281/zenodo.10252380).
+For instructions to use the codes, see \url{https://nmrlipids.github.io/}.
 
@@
 \noindent
-\textbf{Figure 1.} Overview of the NMRlipids Databank. 
+\textbf{Figure 1.} 
+Overview of the NMRlipids Databank. 
     \textbf{A} Schematic presentation of the overlay structure used in the NMRlipids Databank.
-A more detailed structure of the \textit{Databank layer} is shown in Fig.~\ref{DatabankStructure} in the SI.
+A more detailed structure of the \textit{Databank layer} is shown in Supplementary Figure~\ref{DatabankStructure}.
     \textbf{B} Distribution of lipids present in the trajectories of the Databank. `Others' lists lipids occurring in six or fewer simulations. 
-    \textbf{C} Distribution of force fields in the simulations in the Databank. References for each force field are given in Table~\ref{tab:ForceFields}.
+    \textbf{C} Distribution of force fields in the simulations in the Databank. References for each force field are given in Supplementary Table~\ref{tab:ForceFields}.
     \textbf{D} Flowchart for performing an analysis of properties through all MD simulations in the NMRlipids Databank using the API.
@@
     \textbf{D--F} Direct comparison against experimental (NMR order parameters and X-ray scattering ) data exemplified for a simulation with the best overall order parameter quality (D), the best quality for POPE lipid (E), and the headgroup quality for POPE (F).
+ Error bars for simulations are standard error of the mean over different lipids (n=number of lipids in a simulation shown in B). Error bars for experiments are 0.02~\cite{ollila16}.
     \textbf{G} Scatter plots and Pearson correlation coefficients, $r$, for the membrane area per lipid, thickness, first minimum of X-ray scattering form factor and average order parameter of the \textit{sn}-1 acyl chain extracted from the NMRlipids Databank.
-All correlation coefficients have p-value below 0.001.
-For more correlations see Fig.~\ref{fig:QualityCorrelationsSI}.
+All correlation coefficients have p-value below 0.001 with two-sided test.
+For more correlations see Supplementary Figure~\ref{fig:QualityCorrelationsSI}.
 
@@
 \textbf{Figure 3.}
-Predictions of areas per lipid of multi-component membranes composed of POPC, POPE, POPS, PI, sphingomyelin lipids, and cholesterol from linear regression model against literature data from simulations (green~\cite{shahane19}, blue~\cite{kumar21}, and red~\cite{oliveira22}).
+Predictions of areas per lipid (APL) of multi-component membranes composed of POPC, POPE, POPS, PI, sphingomyelin lipids, and cholesterol from linear regression model against literature data from simulations (green~\cite{shahane19}, blue~\cite{kumar21}, and red~\cite{oliveira22}).
+Error bars are from the publications as the values.
 Black line indicates $x=y$.

### 2024-01-12 SI.tex
Editorial updates

@@
 %\renewcommand{\thepage}{S\arabic{page}}%
+%\renewcommand{\figurename}{Supplementary Figure}
+
+\usepackage[figurename=Supplementary Figure]{caption}
+\usepackage[tablename=Supplementary Table]{caption}
+\usepackage[refname=Supplementary Table]{caption}
 
@@
 \author[10]{Rebeca Garcia-Fandino}
-\author[11]{Ivan Gushchin}
-\author[12,13]{Batuhan Kav}           %ORCID: 0000-0003-4990-373X
-\author[14]{Norbert Ku{\v c}erka}
-\author[15]{Patrik Kula}
-\author[16]{Milla Kurki}
-\author[11]{Alexander Kuzmin}
-\author[17]{Anusha Lalitha} %ORCID: 0000-0002-2817-661X
-\author[18,19]{Fabio Lolicato}
-\author[20,21]{Jesper J. Madsen}  %ORCID: 0000-0003-1411-9080
-\author[2,22,23]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
-\author[24]{Cedric Mingham}
-\author[25,26]{Luca Monticelli}
-\author[1,27]{Ricky Nencini}
-\author[22,23]{Alexey M. Nesterenko}
-\author[28]{Thomas J. Piggot}
-\author[29]{{\'A}ngel Pi{\~n}eiro}
-\author[22,23]{Nathalie Reuter}
-\author[12,30]{Suman Samantray} %ORCID: 0000-0003-3361-9582
-\author[10,29,31]{Fabi{\'a}n Su{\'a}rez-Lest{\'o}n}
-\author[22,23]{Reza Talandashti}
-\author[1,32,*]{O. H. Samuli Ollila} %ORCID: 0000-0002- 8728-1006
+\author[ ]{Ivan Gushchin}
+\author[11,12]{Batuhan Kav}           %ORCID: 0000-0003-4990-373X
+\author[13]{Norbert Ku{\v c}erka}
+\author[14]{Patrik Kula}
+\author[15]{Milla Kurki}
+\author[ ]{Alexander Kuzmin}
+\author[16]{Anusha Lalitha} %ORCID: 0000-0002-2817-661X
+\author[17,18]{Fabio Lolicato}
+\author[19,20]{Jesper J. Madsen}  %ORCID: 0000-0003-1411-9080
+\author[2,21,22]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
+\author[23]{Cedric Mingham}
+\author[24,25]{Luca Monticelli}
+\author[1,26]{Ricky Nencini}
+\author[21,22]{Alexey M. Nesterenko}
+\author[27]{Thomas J. Piggot}
+\author[28]{{\'A}ngel Pi{\~n}eiro}
+\author[21,22]{Nathalie Reuter}
+\author[12,29]{Suman Samantray} %ORCID: 0000-0003-3361-9582
+\author[10,28,30]{Fabi{\'a}n Su{\'a}rez-Lest{\'o}n}
+\author[21,22]{Reza Talandashti}
+\author[1,31,*]{O. H. Samuli Ollila} %ORCID: 0000-0002- 8728-1006
 
@@
 \affil[8]{Sorbonne Universit{\'e}, Ecole Normale Sup{\'e}rieure, PSL University, CNRS, Laboratoire des Biomol{\'e}cules (LBM), F-75005 Paris, France}
-\affil[9]{Universit{\'e} Paris Cit{\'e}, F-75006 Paris, France}
+\affil[9]{Universit{\'e} Paris Cit{\'e}, UFR Sciences du Vivant, Paris, 75013, France}
 \affil[10]{Center for Research in Biological Chemistry and Molecular Materials (CiQUS), Universidade de Santiago de Compostela,  Santiago de Compostela, E-15782, Spain}
-\affil[11]{no affiliation}
-\affil[12]{Institute of Biological Information Processing: Structural Biochemistry (IBI-7), Forschungszentrum Jülich, 52428 Jülich, Germany}
-\affil[13]{ariadne.ai GmbH (Germany), Häusserstra{\ss}e 3 Heidelberg 69115, Germany }
-\affil[14]{Department of Physical Chemistry of Drugs, Faculty of Pharmacy, Comenius University Bratislava, 832 32 Bratislava, Slovakia}
-\affil[15]{Institute of Organic Chemistry and Biochemistry of the Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2,  Prague, CZ-16610, Czech Republic}
-\affil[16]{School of Pharmacy, University of Eastern Finland, 70211 Kuopio, Finland}
-\affil[17]{Institut Charles Gerhardt Montpellier (UMR CNRS 5253), Universit{\'e} Montpellier, Place Eug{\`e}ne Bataillon, 34095 Montpellier, Cedex 05, France}
-\affil[18]{Heidelberg University Biochemistry Center, 69120 Heidelberg, Germany}
-\affil[19]{Department of Physics, University of Helsinki, FI-00014 Helsinki, Finland}
-\affil[20]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida 33612, United States of America}
-\affil[21]{Center for Global Health and Infectious Diseases Research, Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida 33612, United States of America}
-\affil[22]{Department of Chemistry, University of Bergen, 5007 Bergen, Norway}
-\affil[23]{Computational Biology Unit, Department of Informatics, University of Bergen, 5008 Bergen, Norway}
-\affil[24]{Hochschule Mannheim, University of Applied Sciences, Mannheim, Germany}
-\affil[25]{University of Lyon, CNRS, Molecular Microbiology and Structural Biochemistry (MMSB, UMR 5086), F-69007, Lyon, France}
-\affil[26]{Institut National de la Sant{\'e} et de la Recherche M{\'e}dicale (INSERM), Lyon, France}
-\affil[27]{Division of Pharmaceutical Biosciences, Faculty of Pharmacy, University of Helsinki, Helsinki 00014, Finland}
-
-\affil[28]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}
-
-\affil[29]{Department of Applied Physics, Faculty of Physics, University of Santiago de Compostela, Santiago de Compostela, E-15782, Spain}
-\affil[30]{Institute of Biotechnology, RWTH Aachen University, Worringerweg 3, 52074 Aachen, Germany}
-\affil[31]{MD.USE Innovations S.L., Edificio Emprendia, 15782 Santiago de Compostela, Spain}
-\affil[32]{VTT Technical Research Centre of Finland, Espoo, Finland}
+\affil[11]{Institute of Biological Information Processing: Structural Biochemistry (IBI-7), Forschungszentrum Jülich, 52428 Jülich, Germany}
+\affil[12]{ariadne.ai GmbH (Germany), Häusserstra{\ss}e 3 Heidelberg 69115, Germany }
+\affil[13]{Department of Physical Chemistry of Drugs, Faculty of Pharmacy, Comenius University Bratislava, 832 32 Bratislava, Slovakia}
+\affil[14]{Institute of Organic Chemistry and Biochemistry of the Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2,  Prague, CZ-16610, Czech Republic}
+\affil[15]{School of Pharmacy, University of Eastern Finland, 70211 Kuopio, Finland}
+\affil[16]{Institut Charles Gerhardt Montpellier (UMR CNRS 5253), Universit{\'e} Montpellier, Place Eug{\`e}ne Bataillon, 34095 Montpellier, Cedex 05, France}
+\affil[17]{Heidelberg University Biochemistry Center, 69120 Heidelberg, Germany}
+\affil[18]{Department of Physics, University of Helsinki, FI-00014 Helsinki, Finland}
+\affil[19]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida 33612, United States of America}
+\affil[20]{Center for Global Health and Infectious Diseases Research, Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida 33612, United States of America}
+\affil[21]{Department of Chemistry, University of Bergen, 5007 Bergen, Norway}
+\affil[22]{Computational Biology Unit, Department of Informatics, University of Bergen, 5008 Bergen, Norway}
+\affil[23]{Hochschule Mannheim, University of Applied Sciences, Mannheim, Germany}
+\affil[24]{University of Lyon, CNRS, Molecular Microbiology and Structural Biochemistry (MMSB, UMR 5086), F-69007, Lyon, France}
+\affil[25]{Institut National de la Sant{\'e} et de la Recherche M{\'e}dicale (INSERM), Lyon, France}
+\affil[26]{Division of Pharmaceutical Biosciences, Faculty of Pharmacy, University of Helsinki, Helsinki 00014, Finland}
+\affil[27]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}
+\affil[28]{Department of Applied Physics, Faculty of Physics, University of Santiago de Compostela, Santiago de Compostela, E-15782, Spain}
+\affil[29]{Institute of Biotechnology, RWTH Aachen University, Worringerweg 3, 52074 Aachen, Germany}
+\affil[30]{MD.USE Innovations S.L., Edificio Emprendia, 15782 Santiago de Compostela, Spain}
+\affil[31]{VTT Technical Research Centre of Finland, Espoo, Finland}
 
@@
         %\multicolumn{4}{c}{ {\bf Scripts/BuildDatabank/ }}\\
-        AreaPerLipidAnd ThicknessCorrelations.ipynb  & Analyze correlations between results in the {\it Databank layer}  & Results in the {\it Databank layer} & Figs.~\ref{fig:quality}G and~\ref{fig:QualityCorrelationsSI} \\
+        AreaPerLipidAnd ThicknessCorrelations.ipynb  & Analyze correlations between results in the {\it Databank layer}  & Results in the {\it Databank layer} & Figure~\ref{fig:quality}G in the main text and Supplementary Figure~\ref{fig:QualityCorrelationsSI} \\
+        \\
         FlipFlop.py & Calculate flip-flop rates of lipids in all simulations & README.yaml & flipflop.dat files at {\it /Data/Flipflops/} \\
-        plotFlipFlop.ipynb & Plot how flip-flop rates depend on membrane properties & README.yaml and flipflop.dat files at {\it /Data/Flipflops/} & Fig.~\ref{fig:flip-flops} \\
+        \\
+        plotFlipFlop.ipynb & Plot how flip-flop rates depend on membrane properties & README.yaml and flipflop.dat files at {\it /Data/Flipflops/} & Figure~\ref{fig:flip-flops} in the main text \\
+        \\
         calcMD-PERMEATION.py  & Calculate permeation rate of water molecules through bilayers from all simulations & README.yaml & Counting\_events.txt files at {\it /Data/MD-PERMEATION/} \\
+        \\
         calcWATERdiffusion.py & Calculate water lateral diffusion along membrane surface & README.yaml & WATERlateralMSD.xvg files at {\it /Data/WATERdiffusion/} \\
-        plotWaterPermeation.ipynb & Plot how water permeation and diffusion depend on membrane properties & README.yaml, Counting\_events.txt\,at {\it /Data/MD-PERMEATION/} and\,WATERlateralMSD.xvg files\,at {\it /Data/WATERdiffusion/} & Figs.~\ref{fig:permeability},~\ref{fig:permeationSI}, and~\ref{fig:diffusionSI}.
+        \\
+        plotWaterPermeation.ipynb & Plot how water permeation and diffusion depend on membrane properties & README.yaml, Counting\_events.txt\,at {\it /Data/MD-PERMEATION/} and\,WATERlateralMSD.xvg files\,at {\it /Data/WATERdiffusion/} & Figure~\ref{fig:permeability} in the main text, and Supplementary Figures~\ref{fig:permeationSI}, and~\ref{fig:diffusionSI}.
+\\
         \\
         APLpredictor.ipynb & Predict area per lipid of a membrane based in its lipid composition from machine learning models trained using the NMRlipids databank & README.yaml and apl.dat files & Machine learning models at /Data/APLpredictor/ \\
-        %plotSimulation.ipynb & Plot results from a simulation together with the experimental data if available & README.yaml files and results in the {\it databank layer} & Figs~\ref{fig:quality}D-F and ~\ref{fig:POPC_POPE_dataSI}
     \end{tabular}
@@
 \section*{NMR experiments}
-Acyl chain order parameters of POPE (Figs.~\ref{POPEexp1} and~\ref{POPEexp2}) and POPG (Figs.~\ref{POPGexp1} and~\ref{POPGexp2}) were analyzed from the same data that were previously recorded to determine headgroup order parameters~\cite{bacle21}.
+Acyl chain order parameters of POPE (Supplementary Figures~\ref{POPEexp1} and~\ref{POPEexp2}) and POPG (Supplementary Figures~\ref{POPGexp1} and~\ref{POPGexp2}) were analyzed from the same data that were previously recorded to determine headgroup order parameters~\cite{bacle21}.
 The analysis of the crowded spectral region at 29--31 ppm was based on the previous assignment reported for POPC membranes~\cite{ferreira13}.
-To measure the order parameters for DOPC (Fig.~\ref{fig:DOPCexp}), the sample was prepared and experiments performed similarly to previous studies~\cite{bacle21}.
+To measure the order parameters for DOPC (Supplementary Figure~\ref{fig:DOPCexp}), the sample was prepared and experiments performed similarly to previous studies~\cite{bacle21}.
 
@@
 (C) C--H bond order parameter profile for the acyl chains of POPE.
-The splittings used for calculating the order parameters are shown in Fig.~\ref{POPEexp2}.
+The splittings used for calculating the order parameters are shown in Supplementary Figure~\ref{POPEexp2}.
 The unassigned peaks belong to the headgroup and glycerol backbone carbons.
@@
   \includegraphics[width=\textwidth]{Figures/POPE_slices.pdf}
-  \caption{Dipolar spectra obtained from the 2D R-PDLF spectrum from POPE in Fig.~\ref{POPEexp1}.
+  \caption{Dipolar spectra obtained from the 2D R-PDLF spectrum from POPE in Supplementary Figure~\ref{POPEexp1}.
 The number at the top left corner of each panel denotes the corresponding chemical shift.
 The carbon label for each splitting is displayed on the top right corner.
-The labels are the same as in Fig.~\ref{POPEexp1}.
+The labels are the same as in Supplementary Figure~\ref{POPEexp1}.
 }
@@
 (C) C--H bond order parameter profile for the acyl chains of POPG.
-The splittings used for calculating the order parameters are shown in Fig.~\ref{POPGexp2}.
+The splittings used for calculating the order parameters are shown in Supplementary Figure~\ref{POPGexp2}.
 The unassigned peaks belong to the headgroup and glycerol backbone carbons.
@@
   \includegraphics[width=\textwidth]{Figures/POPG_slices.pdf}
-  \caption{Dipolar spectra obtained from the 2D R-PDLF spectrum described in Fig.~\ref{POPGexp1}.
+  \caption{Dipolar spectra obtained from the 2D R-PDLF spectrum described in Supplementary Figure~\ref{POPGexp1}.
 The number at the top left corner of each panel denotes the corresponding chemical shift.
 The carbon label for each splitting is displayed on the top right corner.
-The labels are the same as in Fig.~\ref{POPGexp1}.
+The labels are the same as in Supplementary Figure~\ref{POPGexp1}.
 }
@@
   \includegraphics[width=\textwidth]{Figures/DOPC_slices.pdf}
-  \caption{Dipolar spectra obtained from the 2D R-PDLF spectrum described in Fig.~\ref{fig:DOPCexp}.
+  \caption{Dipolar spectra obtained from the 2D R-PDLF spectrum described in Supplementary Figure~\ref{fig:DOPCexp}.
 The number at the top left corner of each panel denotes the corresponding chemical shift.
 The carbon label for each splitting is displayed on the top right corner.
-The labels are the same as in Fig.~\ref{fig:DOPCexp}.
+The labels are the same as in Supplementary Figure~\ref{fig:DOPCexp}.
 }
@@
 %X-ray scattering form factors contributed to the NMRlipids III project are included in the databank (\url{http://nmrlipids.blogspot.com/2022/09/nmrlipids-iii-including-lipid-lateral.html}).
+
+\renewcommand\refname{Supplementary References}
 

### 2024-01-12 SI.tex
Editorial changes

@@
 
+\pagebreak
+
+\section*{Databank content}
+
+\begin{figure}[!h]
+  \includegraphics[width=\textwidth]{Figures/DataBankChart.pdf}
+  \caption{Structure of the NMRlipids Databank.
+Manually added input data (blue boxes) include basic information on the simulation, permanent links to the raw data, and experimental data if available.
+The databank entries (red box) and analysis results (green boxes), at \url{https://github.com/NMRlipids/Databank/tree/main/Data/Simulations} are automatically generated by the computer programs included in the NMRlipids Databank (yellow boxes).
+Because the raw data are not permanently stored but can be accessed based on the information in the Databank, this connection is marked with a dashed line.}
+  \label{DatabankStructure}
+\end{figure}
+
+\begin{table}[!h]
+    \centering
+    \begin{tabular}{l}
+    Force field name and references \\
+    \hline
+    CHARMM36~\cite{klauda10}\\Slipids~\cite{jambeck12,jambeck12b,jambeck2012another,ermilova16,grote20}  \\
+    MacRog~\cite{Kulig15b}  \\
+    Amber Lipid14/17~\cite{dickson14,dickson22}  \\
+    Charmm-Drude~\cite{li2017drude}      \\
+    ECClipids~\cite{melcr18,melcr20,bacle21}  \\
+    GROMOS-CKP~\cite{Chandrasekhar03,kukol09,piggot12}  \\
+    Berger~\cite{berger97}  \\
+    ECC-CHARMM36~\cite{nencini2019development}  \\
+    Orange~\cite{catte16}  \\
+    Poger~\cite{poger10}  \\
+    GROMOS 43A1-S3~\cite{chiu09}  \\ 
+    GAFFlipid~\cite{dickson12}  \\
+    OPLS3e~\cite{roos19}  \\
+    Ulmschneider~\cite{Ulmschneider09}  \\
+    Chiu Gromos~\cite{chiu09}  \\
+    AMOEBA~\cite{chu2018polarizable} \\
+    OpenFF~\cite{mobley18,boothroyd23,OpenFFgit}
+    \end{tabular}
+    \caption{List of current force fields used in simulations in the Databank, with references.}
+    \label{tab:ForceFields}
+\end{table}
+
+\begin{table}[!h]
+    \centering
+    %\begin{tabular}{c c c c}
+    \begin{tabular}{p{3.0cm}  p{5.0cm}  p{3.0cm}  p{4.0cm}}
+        Code & Function & Input & Output \\
+        \hline
+        \multicolumn{4}{c}{ {\bf Databank layer}}\\
+        \hline
+        \multicolumn{4}{c}{ {\bf Scripts/BuildDatabank/ }}\\
+         AddData.py & Create README.yaml from user given information in info.yaml  & info.yaml & README.yaml \\
+         searchDATABANK.py & Pair simulations with available experimental data  & README.yaml files & Updated README.yaml files \\
+         QualityEvaluation.py & Quality-evaluate simulations that are paired with experimental data  & README.yaml files & [lipid\_name]\_OrderParameters\_quality.json, [lipid\_name]\_FragmentQuality.json, system\_quality.json, FormFactorQuality.json \\
+         \multicolumn{4}{c}{ {\bf Scripts/AnalyzeDatabank/ }}\\
+         calcOrderParameters.py & Calculate C-H bond order parameters for all lipids in the simulation & README.yaml files & [lipid\_name]\_OrderParameters.json \\
+         calcAPL.py & Calculate area per lipid as a function of time & README.yaml files & apl.json \\
+         calc\_FormFactors.py & Calculate x-ray scattering form factors & README.yaml files & FormFactor.json \\
+         calc\_thickness.py & Calculate membrane thickness & README.yaml files & thickness.json \\
+         %template.py $\qquad$  template.ipynb & Minimal code to loop over simulations that can be used as a template for new analyses & README.yaml files & \\
+
+    \end{tabular}
+    \caption{List of relevant codes used to build the Databank and perform analyses in the {\it Databank layer} available at~\url{https://github.com/NMRLipids/Databank/}.}
+    \label{tab:codes}
+\end{table}
+
+\begin{table}[!h]
+    \centering
+    %\begin{tabular}{c c c c}
+    \begin{tabular}{p{3.5cm}  p{4.5cm}  p{3.0cm}  p{4.0cm}}
+        Code & Function & Input & Output \\
+        \hline
+        \multicolumn{4}{c}{ {\bf Application layer}}\\
+        \hline
+        %\multicolumn{4}{c}{ {\bf Scripts/BuildDatabank/ }}\\
+        AreaPerLipidAnd ThicknessCorrelations.ipynb  & Analyze correlations between results in the {\it Databank layer}  & Results in the {\it Databank layer} & Figure~\ref{fig:quality}G in the main text and Supplementary Figure~\ref{fig:QualityCorrelationsSI} \\
+        \\
+        FlipFlop.py & Calculate flip-flop rates of lipids in all simulations & README.yaml & flipflop.dat files at {\it /Data/Flipflops/} \\
+        \\
+        plotFlipFlop.ipynb & Plot how flip-flop rates depend on membrane properties & README.yaml and flipflop.dat files at {\it /Data/Flipflops/} & Figure~\ref{fig:flip-flops} in the main text \\
+        \\
+        calcMD-PERMEATION.py  & Calculate permeation rate of water molecules through bilayers from all simulations & README.yaml & Counting\_events.txt files at {\it /Data/MD-PERMEATION/} \\
+        \\
+        calcWATERdiffusion.py & Calculate water lateral diffusion along membrane surface & README.yaml & WATERlateralMSD.xvg files at {\it /Data/WATERdiffusion/} \\
+        \\
+        plotWaterPermeation.ipynb & Plot how water permeation and diffusion depend on membrane properties & README.yaml, Counting\_events.txt\,at {\it /Data/MD-PERMEATION/} and\,WATERlateralMSD.xvg files\,at {\it /Data/WATERdiffusion/} & Figure~\ref{fig:permeability} in the main text, and Supplementary Figures~\ref{fig:permeationSI}, and~\ref{fig:diffusionSI}.
+\\
+        \\
+        APLpredictor.ipynb & Predict area per lipid of a membrane based in its lipid composition from machine learning models trained using the NMRlipids databank & README.yaml and apl.dat files & Machine learning models at /Data/APLpredictor/ \\
+    \end{tabular}
+    \caption{Examples of codes that analyze membrane properties from the Databank in an {\it Application layer} available at~\url{https://github.com/NMRLipids/DataBankManuscript/}.}
+    \label{tab:codesApplication}
+\end{table}
+
+\begin{table}[!h]
+    \centering
+    \begin{tabular}{  p{3.5cm}  p{9.5cm}  p{4.0cm} }
+    \toprule
+    key & description & type  \\
+    \midrule
+        %\hline
+    DOI & DOI from where the raw data is found & user given (compulsory) \\
+    SOFTWARE & Software used to run the simulation (Gromacs, Amber, NAMD, etc.) & \\
+    TRJ & Name of the trajectory file found from DOI (trr or xtc for Gromacs, dcd for OpenMM) & \\
+    TPR (Gromacs) & Name of the tpr topology file found from DOI for Gromacs simulations & \\
+    PDB (OpenMM) & Name of the pdb file found from DOI for OpenMM simulations & \\
+    PREEQTIME & Pre-equilibrate time simulated before the uploaded trajectory in nanoseconds. 
+    %For example, if you upload 100-200 ns part of total 200 ns simulation, this should value should be 100. 
+    & \\
+    TIMELEFTOUT & Equilibration period in the uploaded trajectory that should be discarded in analyses. 
+    %For example, if you upload 0-200 ns part of total 200 ns simulation where the first 100 ns should be considered as an equilibration, this value should be 100. 
+    & \\
+    COMPOSITION & Dictionary connecting universal molecule and atom names to the ones used in simulation & \\
+    DIR\_WRK & Temporary working directory in your local computer. \\
+    UNITEDATOM\_DICT & Information for constucting hydrogens for united atom simulations using buildH program~\cite{santuz21}. Empty for all atom simulations. & \\
+    TYPEOFSYSTEM & Lipid bilayer or something else & \\
+    \hline
+    PUBLICATION & Give reference to a publication(s) related to the data & user given (optional)\\
+    AUTHORS\_CONTACT & Name and email of the main author(s) of the data & \\
+    SYSTEM & System description on free text format & \\
+    SOFTWARE\_VERSION & Version of the used software & \\
+    FF & Name of the used force field & \\
+    FF\_SOURCE & Source of the force field parameters, e.g, CHARMM-GUI, webpage, citation to a publication & \\
+    FF\_DATE &  Date when force field parameters were accessed on the given source (day/month/year) & \\
+    FF{molename} & Molecule specific force field information, e.g., water model with FFSOL and sodium parameters with FFSOD & \\
+    CPT & Name of the Gromacs checkpoint file & \\
+    LOG & Name of the Gromacs log file & \\
+    GRO & Name of the Gromacs gro file & \\
+    TOP & Name of top file for Gromacs or psf file for OpenMM & \\
+    CRD & Name of crd file for OpenMM & \\
+    WARNINGS & Dictionary containing information about unusual features in the trajectory, such as ambiguous atom names, membrane normal not oriented in $z$-direction, old Gromacs version used & \\
+    \hline
+    TRAJECTORY\_SIZE & Size of the trajectory file in bytes & automatically extracted data \\
+    TRJLENGTH & Lenght of the trajectory (ps) & \\
+    TEMPERATURE & Temperature of the simulation & \\
+    NUMBER\_OF\_ATOMS & Number of atoms in the simulation & \\
+    DATEOFRUNNIG & Date when added into the Databank & \\
+    EXPERIMENT & Potentially connected experimental data & \\
+    COMPOSITION & Numbers of lipid molecules in both leaflets and numbers of other molecules are added to the dictionary & \\
+    ID & Unique ID number to ease the analyses & \\
+    \end{tabular}
+    \caption{Keys stored in the README.yaml files of simulations.}
+    \label{tab:READMEkeys}
+\end{table}
+
+\begin{table}[!h]
+    \centering
+    \begin{tabular}{  p{5.0cm}  p{10.0cm}}
+    \toprule
+    key & description \\
+    \midrule
+    DOI & DOI of the publication related to the experimental data \\
+    TEMPERATURE & Temperature of the experiment \\
+    MOLAR\_FRACTIONS & Dictionary of molar fractions of bilayer components \\
+    ION\_CONCENTRATIONS & Dictionary of ion concentrations of the system \\
+    % (defined as ??) \\
+    TOTAL\_LIPID\_CONCENTRATION & Total concentration of lipid components; if exact concentration is not known, but experiments are performed in excess water, 'full hydration' can be given \\
+    COUNTER\_IONS & Type of counter ions if present
+\end{tabular}
+    \caption{Keys stored in the README.yaml files of experiments.}
+    \label{tab:READMEkeysEXP}
+\end{table}
+
+\newpage
 %\pagebreak
@@
 \pagebreak
-\section*{Databank content}
-
-\begin{figure}[!h]
-  \includegraphics[width=\textwidth]{Figures/DataBankChart.pdf}
-  \caption{Structure of the NMRlipids Databank.
-Manually added input data (blue boxes) include basic information on the simulation, permanent links to the raw data, and experimental data if available.
-The databank entries (red box) and analysis results (green boxes), at \url{https://github.com/NMRlipids/Databank/tree/main/Data/Simulations} are automatically generated by the computer programs included in the NMRlipids Databank (yellow boxes).
-Because the raw data are not permanently stored but can be accessed based on the information in the Databank, this connection is marked with a dashed line.}
-  \label{DatabankStructure}
-\end{figure}
-
-\begin{table}[!h]
-    \centering
-    %\begin{tabular}{c c c c}
-    \begin{tabular}{p{3.0cm}  p{5.0cm}  p{3.0cm}  p{4.0cm}}
-        Code & Function & Input & Output \\
-        \hline
-        \multicolumn{4}{c}{ {\bf Databank layer}}\\
-        \hline
-        \multicolumn{4}{c}{ {\bf Scripts/BuildDatabank/ }}\\
-         AddData.py & Create README.yaml from user given information in info.yaml  & info.yaml & README.yaml \\
-         searchDATABANK.py & Pair simulations with available experimental data  & README.yaml files & Updated README.yaml files \\
-         QualityEvaluation.py & Quality-evaluate simulations that are paired with experimental data  & README.yaml files & [lipid\_name]\_OrderParameters\_quality.json, [lipid\_name]\_FragmentQuality.json, system\_quality.json, FormFactorQuality.json \\
-         \multicolumn{4}{c}{ {\bf Scripts/AnalyzeDatabank/ }}\\
-         calcOrderParameters.py & Calculate C-H bond order parameters for all lipids in the simulation & README.yaml files & [lipid\_name]\_OrderParameters.json \\
-         calcAPL.py & Calculate area per lipid as a function of time & README.yaml files & apl.json \\
-         calc\_FormFactors.py & Calculate x-ray scattering form factors & README.yaml files & FormFactor.json \\
-         calc\_thickness.py & Calculate membrane thickness & README.yaml files & thickness.json \\
-         %template.py $\qquad$  template.ipynb & Minimal code to loop over simulations that can be used as a template for new analyses & README.yaml files & \\
-
-    \end{tabular}
-    \caption{List of relevant codes used to build the Databank and perform analyses in the {\it Databank layer} available at~\url{https://github.com/NMRLipids/Databank/}.}
-    \label{tab:codes}
-\end{table}
-
-\begin{table}[!h]
-    \centering
-    %\begin{tabular}{c c c c}
-    \begin{tabular}{p{3.5cm}  p{4.5cm}  p{3.0cm}  p{4.0cm}}
-        Code & Function & Input & Output \\
-        \hline
-        \multicolumn{4}{c}{ {\bf Application layer}}\\
-        \hline
-        %\multicolumn{4}{c}{ {\bf Scripts/BuildDatabank/ }}\\
-        AreaPerLipidAnd ThicknessCorrelations.ipynb  & Analyze correlations between results in the {\it Databank layer}  & Results in the {\it Databank layer} & Figure~\ref{fig:quality}G in the main text and Supplementary Figure~\ref{fig:QualityCorrelationsSI} \\
-        \\
-        FlipFlop.py & Calculate flip-flop rates of lipids in all simulations & README.yaml & flipflop.dat files at {\it /Data/Flipflops/} \\
-        \\
-        plotFlipFlop.ipynb & Plot how flip-flop rates depend on membrane properties & README.yaml and flipflop.dat files at {\it /Data/Flipflops/} & Figure~\ref{fig:flip-flops} in the main text \\
-        \\
-        calcMD-PERMEATION.py  & Calculate permeation rate of water molecules through bilayers from all simulations & README.yaml & Counting\_events.txt files at {\it /Data/MD-PERMEATION/} \\
-        \\
-        calcWATERdiffusion.py & Calculate water lateral diffusion along membrane surface & README.yaml & WATERlateralMSD.xvg files at {\it /Data/WATERdiffusion/} \\
-        \\
-        plotWaterPermeation.ipynb & Plot how water permeation and diffusion depend on membrane properties & README.yaml, Counting\_events.txt\,at {\it /Data/MD-PERMEATION/} and\,WATERlateralMSD.xvg files\,at {\it /Data/WATERdiffusion/} & Figure~\ref{fig:permeability} in the main text, and Supplementary Figures~\ref{fig:permeationSI}, and~\ref{fig:diffusionSI}.
-\\
-        \\
-        APLpredictor.ipynb & Predict area per lipid of a membrane based in its lipid composition from machine learning models trained using the NMRlipids databank & README.yaml and apl.dat files & Machine learning models at /Data/APLpredictor/ \\
-    \end{tabular}
-    \caption{Examples of codes that analyze membrane properties from the Databank in an {\it Application layer} available at~\url{https://github.com/NMRLipids/DataBankManuscript/}.}
-    \label{tab:codesApplication}
-\end{table}
-
-\begin{table}[p]
-    \centering
-    \begin{tabular}{  p{3.5cm}  p{9.5cm}  p{4.0cm} }
-    \toprule
-    key & description & type  \\
-    \midrule
-        %\hline
-    DOI & DOI from where the raw data is found & user given (compulsory) \\
-    SOFTWARE & Software used to run the simulation (Gromacs, Amber, NAMD, etc.) & \\
-    TRJ & Name of the trajectory file found from DOI (trr or xtc for Gromacs, dcd for OpenMM) & \\
-    TPR (Gromacs) & Name of the tpr topology file found from DOI for Gromacs simulations & \\
-    PDB (OpenMM) & Name of the pdb file found from DOI for OpenMM simulations & \\
-    PREEQTIME & Pre-equilibrate time simulated before the uploaded trajectory in nanoseconds. 
-    %For example, if you upload 100-200 ns part of total 200 ns simulation, this should value should be 100. 
-    & \\
-    TIMELEFTOUT & Equilibration period in the uploaded trajectory that should be discarded in analyses. 
-    %For example, if you upload 0-200 ns part of total 200 ns simulation where the first 100 ns should be considered as an equilibration, this value should be 100. 
-    & \\
-    COMPOSITION & Dictionary connecting universal molecule and atom names to the ones used in simulation & \\
-    DIR\_WRK & Temporary working directory in your local computer. \\
-    UNITEDATOM\_DICT & Information for constucting hydrogens for united atom simulations using buildH program~\cite{santuz21}. Empty for all atom simulations. & \\
-    TYPEOFSYSTEM & Lipid bilayer or something else & \\
-    \hline
-    PUBLICATION & Give reference to a publication(s) related to the data & user given (optional)\\
-    AUTHORS\_CONTACT & Name and email of the main author(s) of the data & \\
-    SYSTEM & System description on free text format & \\
-    SOFTWARE\_VERSION & Version of the used software & \\
-    FF & Name of the used force field & \\
-    FF\_SOURCE & Source of the force field parameters, e.g, CHARMM-GUI, webpage, citation to a publication & \\
-    FF\_DATE &  Date when force field parameters were accessed on the given source (day/month/year) & \\
-    FF{molename} & Molecule specific force field information, e.g., water model with FFSOL and sodium parameters with FFSOD & \\
-    CPT & Name of the Gromacs checkpoint file & \\
-    LOG & Name of the Gromacs log file & \\
-    GRO & Name of the Gromacs gro file & \\
-    TOP & Name of top file for Gromacs or psf file for OpenMM & \\
-    CRD & Name of crd file for OpenMM & \\
-    WARNINGS & Dictionary containing information about unusual features in the trajectory, such as ambiguous atom names, membrane normal not oriented in $z$-direction, old Gromacs version used & \\
-    \hline
-    TRAJECTORY\_SIZE & Size of the trajectory file in bytes & automatically extracted data \\
-    TRJLENGTH & Lenght of the trajectory (ps) & \\
-    TEMPERATURE & Temperature of the simulation & \\
-    NUMBER\_OF\_ATOMS & Number of atoms in the simulation & \\
-    DATEOFRUNNIG & Date when added into the Databank & \\
-    EXPERIMENT & Potentially connected experimental data & \\
-    COMPOSITION & Numbers of lipid molecules in both leaflets and numbers of other molecules are added to the dictionary & \\
-    ID & Unique ID number to ease the analyses & \\
-    \end{tabular}
-    \caption{Keys stored in the README.yaml files of simulations.}
-    \label{tab:READMEkeys}
-\end{table}
-
-%\begin{table}[h]
-%    \centering
-%    \begin{tabular}{c|c}
-%        Abbreviation & Molecule name \\
-%        \hline
-%        POPC &  1-palmitoyl-2-oleoyl-sn-glycero-3-phosphocholine\\
-%        POPG &  1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoglycerol \\
-%        POPS & 1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-L-serine \\
-%        POPE & 1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoethanolamine \\
-%        CHOL & cholesterol \\
-%        DHMDMAB & dihexadecyldimethylammonium \\
-%        \hline
-%        POT & potassium ion \\
-%%        SOD & sodium ion \\
- %       CLA & chloride ion \\
-%        CAL & calcium ion \\
-%%    \end{tabular}
- %   \caption{Abbreviations used in the databank}
-%    \label{tab:abbreviations}
-%\end{table}
-
-\begin{table}[h]
-    \centering
-    \begin{tabular}{  p{5.0cm}  p{10.0cm}}
-    \toprule
-    key & description \\
-    \midrule
-    DOI & DOI of the publication related to the experimental data \\
-    TEMPERATURE & Temperature of the experiment \\
-    MOLAR\_FRACTIONS & Dictionary of molar fractions of bilayer components \\
-    ION\_CONCENTRATIONS & Dictionary of ion concentrations of the system \\
-    % (defined as ??) \\
-    TOTAL\_LIPID\_CONCENTRATION & Total concentration of lipid components; if exact concentration is not known, but experiments are performed in excess water, 'full hydration' can be given \\
-    COUNTER\_IONS & Type of counter ions if present
-\end{tabular}
-    \caption{Keys stored in the README.yaml files of experiments.}
-    \label{tab:READMEkeysEXP}
-\end{table}
-
-\begin{comment}
-
-\begin{table}[p]
-    \centering
-    \begin{tabular}{c c}
-    Abbreviation & Molecule name \\
-    \hline
-    Lipids and surfactants & \\
-    \hline
-    POPC &  1-palmitoyl-2-oleoyl-sn-glycero-3-phosphocholine  \\
-    POPG &  1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoglycerol \\
-    POPS & 1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-L-serine \\
-    POPE & 1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoethanolamine \\
-    PYPC & 1-(16:0)-2-(16:1$^\Delta9$)-sn-glycero-3-phosphocholine \\
-    PAzePCprot & 1-palmitoyl-2-azelaoyl-sn-glycero-3-phosphocholine protonated \\
-    PAzePCdeprot & 1-palmitoyl-2-azelaoyl-sn-glycero-3-phosphocholine deprotonated \\
-    DMPC & 1,2-dimyristoyl-sn-glycero-3-phosphocholine \\
-    DPPC & 1,2-dipalmitoyl-sn-glycero-3-phosphocholine \\
-    DPPE & 1,2-dipalmitoyl-sn-glycero-3-phosphoethanolamine \\
-    DPPG & 1,2-dipalmitoyl-sn-glycero-3-phospho-(1'-rac-glycerol) (sodium salt) \\
-    DEPC & 1,2-dierucoyl-sn-glycero-3-phosphocholine \\
-    DRPC & 1,2-(14:1$^\Delta9$)-sn-glycero-3-phosphocholine \\
-    DYPC & 1,2-(16:1$^\Delta9$)-sn-glycero-3-phosphocholine \\
-    DLPC & 1,2-dilauroyl-sn-glycero-3-phosphocholine \\
-    DLIPC& 1,2-dilinoleoyl-sn-glycero-3-phosphocholine \\
-    DOG  & 1,2-dioleoyl-sn-glycerol \\
-    DOPC & 1,2-dioleoyl-sn-glycero-3-phosphocholine \\
-    DOPE & 1,2-dioleoyl-sn-glycero-3-phosphoethanolamine \\
-    DDOPC& 1,2-didocosahexaenoyl-sn-glycero-3-phosphocholine \\
-    DOPS & 1,2-dioleoyl-sn-glycero-3-phospho-L-serine \\
-    DSPC & 1,2-distearoyl-sn-glycero-3-phosphocholine \\
-    DAPC & 1,2-diarachidonoyl-sn-glycero-3-phosphocholine \\
-    SLiPC & 1-(18:0)-2-(18:2 $^{\Delta9,12}$)-sn-glycero-3-phosphocholine  \\
-    DMTAP & 1,2-dimyristoyl-3-trimethylammonium-propane \\
-    SOPC & 1-stearoyl-2-oleoyl-sn-glycero-3-phosphocholine \\
-    POPI & \\ 
-    SAPI & \\
-    SLPI &  \\
-    SDG & 1-stearoyl-2-docosahexaenoyl-sn-glycerol \\
-    SDPE & 1-stearoyl-2-docosahexaenoyl-sn-glycero-3-phosphoethanolamine \\
-    CER  & N-palmitoyl-D-erythro-sphingosine \\
-    CHOL & cholesterol  \\
-    DCHOL & 18,19-di-nor-cholesterol \\
-    DHMDMAB & dihexadecyldimethylammonium  \\
-    \hline
-    Other molecules & \\
-    \hline
-    POT & potassium ion  \\
-    SOD & sodium ion  \\
-    CLA & chloride ion \\
-    CAL & calcium ion  \\
-    CES & caesium ion \\
-    SOL & water  \\
-    \end{tabular}
-    \caption{Abbreviations for molecule names used in the Databank.}
-    \label{tab:abbreviations}
-\end{table}
-\end{comment}
-
-\begin{table}[p]
-    \centering
-    \begin{tabular}{l}
-    Force field name and references \\
-    \hline
-    CHARMM36~\cite{klauda10}\\
-    Slipids~\cite{jambeck12,jambeck12b,jambeck2012another,ermilova16,grote20}  \\
-    MacRog~\cite{Kulig15b}  \\
-    Amber Lipid14/17~\cite{dickson14,dickson22}  \\
-    Charmm-Drude~\cite{li2017drude}      \\
-    ECClipids~\cite{melcr18,melcr20,bacle21}  \\
-    GROMOS-CKP~\cite{Chandrasekhar03,kukol09,piggot12}  \\
-    Berger~\cite{berger97}  \\
-    ECC-CHARMM36~\cite{nencini2019development}  \\
-    Orange~\cite{catte16}  \\
-    Poger~\cite{poger10}  \\
-    GROMOS 43A1-S3~\cite{chiu09}  \\ 
-    GAFFlipid~\cite{dickson12}  \\
-    OPLS3e~\cite{roos19}  \\
-    Ulmschneider~\cite{Ulmschneider09}  \\
-    Chiu Gromos~\cite{chiu09}  \\
-    AMOEBA~\cite{chu2018polarizable} \\
-    OpenFF~\cite{mobley18,boothroyd23,OpenFFgit}
-    \end{tabular}
-    \caption{List of current force fields used in simulations in the Databank, with references.}
-    \label{tab:ForceFields}
-\end{table}
 

### 2024-01-12 main.tex
Editorial changes

@@
 In conclusion, the quality evaluation based on the NMRlipids Databank suggests that the Slipids parameters are the best currently available choice for simulations with PC and PE lipids, at least for applications where membrane packing is relevant.
-Also direct comparisons with the experimental data for the most relevant simulations are shown in Figs.~\ref{fig:quality}E--F and Supplementary Figure~\ref{fig:POPC_POPE_dataSI} a.
+Also direct comparisons with the experimental data for the most relevant simulations are shown in Figs.~\ref{fig:quality}D--F and Supplementary Figure~\ref{fig:POPC_POPE_dataSI} a.
 Fig.~\ref{fig:quality}D shows the overall highest-ranked simulation, POPC bilayer with OPLS3e parameters, for the reference.
@@
 
-To demonstrate the usefulness of the constructed models for understanding multi-component membrane properties, we predicted later packing (areas per lipid) and membrane thicknesses of common biological membranes based on their lipid compositions reported in the literature, see table~\ref{tab:MLpredictions}.
+To demonstrate the usefulness of the constructed models for understanding multi-component membrane properties, we predicted later packing (areas per lipid) and membrane thicknesses of common biological membranes based on their lipid compositions reported in the literature, see table~1.
 The model predicts substantial 50\% difference in area per lipid between most densely (influenza virus) and loosely (mitochondria) packed membranes.
@@
 Systems are sorted in the order of increasing membrane packing.
-Compositions of different membranes are estimated based on references given in the composition column.}
+Compositions of different membranes are estimated based on references given in the composition column.}\label{tab:MLpredictions}
         \centering
@@
         \end{tabular}
-        \label{tab:MLpredictions}
 
@@
 Our analysis over the NMRlipids Databank, containing significantly more data than what was available in previous studies, suggest non-linear dependencies on both of these parameters.
-Clear dependencies of permeability on hydration level or the fraction of charged lipids, cholesterol, or POPE in the membrane were not observed (Fig.~\ref{fig:permeationSI}).
+Clear dependencies of permeability on hydration level or the fraction of charged lipids, cholesterol, or POPE in the membrane were not observed (Supplementary Figure~\ref{fig:permeationSI}).
 %
@@
 
-To examine how water diffusion anisotropy depends on membrane properties in a multi-lamellar lipid bilayer system, we analyzed the water diffusion parallel to the membrane surface from all simulations in the NMRlipids Databank using the general workflows depicted in Figs.~\ref{fig:overlay}D and E.
-The parallel diffusion coefficient of water, $D_\parallel$, decreases with reduced hydration and increases with the temperature, but dependencies on the membrane area per lipid, thickness, or fraction of charged lipids were not observed in Figs.~\ref{fig:permeability} and~\ref{fig:diffusionSI}.
+To examine how water diffusion anisotropy depends on membrane properties in a multi-lamellar lipid bilayer system, we analyzed the water diffusion parallel to the membrane surface from all simulations in the NMRlipids Databank using the general workflows depicted in Figs.~\ref{fig:overlay}D and~\ref{fig:resultsProcessingSI}.
+The parallel diffusion coefficient of water, $D_\parallel$, decreases with reduced hydration and increases with the temperature, but dependencies on the membrane area per lipid, thickness, or fraction of charged lipids were not observed in Figs.~\ref{fig:permeability} and Supplementary Figure~\ref{fig:diffusionSI}.
 Simulation results are close to the experimental values with low hydration levels in Fig.~\ref{fig:permeability}F, but increase to approximately 50\% higher than the experimental value for bulk water diffusion value (3.1$\times$\,10$^{-9}$\,m$^2$/s at 313\,K~\cite{khakimov08}) with high hydration levels.
@@
 A significant increase in the diffusion anisotropy with membrane packing is observed, as $D_\perp/D_\parallel$ deviates further from unity with decreasing area per lipid and increasing thickness in Figs.~\ref{fig:permeability}G--H.
-This follows from decreasing water permeability with membrane packing (Figs.~\ref{fig:permeability}C--D), while lateral diffusion remains approximately constant (Figs.~\ref{fig:diffusionSI}A,~C).
+This follows from decreasing water permeability with membrane packing (Figs.~\ref{fig:permeability}C--D), while lateral diffusion remains approximately constant (Supplementary Figures.~\ref{fig:diffusionSI}A,~C).
 
@@
     \hline
-    Computational and data scientists & Databanks with programmatic access, training sets for machine learning applications, automatic optimization of simulation models & Programmatic access to all publicly available MD simulation trajectories, machine learning models predicting properties of complex biomolecular assemblies (Table~\ref{tab:MLpredictions}), optimizing force field and model parameters from atomistic to continuum scale \\
+    Computational and data scientists & Databanks with programmatic access, training sets for machine learning applications, automatic optimization of simulation models & Programmatic access to all publicly available MD simulation trajectories, machine learning models predicting properties of complex biomolecular assemblies (Table~1), optimizing force field and model parameters from atomistic to continuum scale \\
 
@@
 Similarly to simulations, each experimental data set has a README.yaml file containing all the relevant information about the experiment.
-The keys and their descriptions for the experimental data, as well as detailed instructions to add the date, are given in Table~\ref{tab:READMEkeysEXP} and in the NMRlipids Databank documentation~\cite{NMRlipidsExperimentalData}.
+The keys and their descriptions for the experimental data, as well as detailed instructions to add the date, are given in Supplementary Table~\ref{tab:READMEkeysEXP} and in the NMRlipids Databank documentation~\cite{NMRlipidsExperimentalData}.
 The NMR data currently in the NMRlipids Databank are taken from Refs.~\citenum{scherer87,ferreira13,antila19,melcr20,bacle21,hanashima19,doktorova20} and the X-ray scattering data from Refs.~\citenum{Kucerka08a,kucerka11,pan12b,pan14,kucerka15,NMRlipidsIII,doktorova20}.
@@
 This general procedure is illustrated in Fig.~\ref{fig:overlay}D, and a templates for user-defined analyses are available via NMRlipids Databank documentation~\cite{NMRlipidsExampleAndTutorials}.
-Further practical examples of codes performing such analyses are listed in Tables~\ref{tab:codes} and~\ref{tab:codesApplication}.
-The equilibration period given by the user (TIMELEFTOUT in Table~\ref{tab:READMEkeys}) is discarded from the trajectories in all analysis codes.
+Further practical examples of codes performing such analyses are listed in Supplementary Tables~\ref{tab:codes} and~\ref{tab:codesApplication}.
+The equilibration period given by the user (TIMELEFTOUT in Supplementary Table~\ref{tab:READMEkeys}) is discarded from the trajectories in all analysis codes.
 For further details, see the NMRlipids Databank documentation \href{https://nmrlipids.github.io/}{nmrlipids.github.io}.
@@
 Also, a quality measure based on differences in the simulated and experimental form factors across the available $q$-range is defined in the SIMtoEXP program~\cite{kucerka10}.
-However, the lobe heights in the simulated form factors depend on the simulation box size, as shown in Fig.~\ref{fig:sizedependence}; consequently, the quality measure defined in SIMtoEXP  would also depend on the simulation box size.
-In contrast, locations of the form factor minima (or, in precise terms: the minima of the absolute value of the form factor) are independent of the simulation box size (Fig.~\ref{fig:sizedependence}).
+However, the lobe heights in the simulated form factors depend on the simulation box size, as shown in Supplementary Figure~\ref{fig:sizedependence}; consequently, the quality measure defined in SIMtoEXP  would also depend on the simulation box size.
+In contrast, locations of the form factor minima (or, in precise terms: the minima of the absolute value of the form factor) are independent of the simulation box size (Supplementary Figure~\ref{fig:sizedependence}).
 Here we use only the location of the first form factor minimum for quality evaluation,
-because (due to fluctuations) the location of the second minimum is difficult to detect automatically in some experimental data sets, such as the POPE data in Figs.~\ref{fig:quality}D, E.
-The first minimum correlates well with the thickness of a membrane (Fig.~\ref{fig:quality}F), although the correlation of the second minima would be even stronger (Fig.~\ref{fig:QualityCorrelationsSI}).
+because (due to fluctuations) the location of the second minimum is difficult to detect automatically in some experimental data sets, such as the POPE data in Figs.~\ref{fig:quality}E, F.
+The first minimum correlates well with the thickness of a membrane (Fig.~\ref{fig:quality}G), although the correlation of the second minima would be even stronger (Supplementary Figure~\ref{fig:QualityCorrelationsSI}).
 In practise, we first filter the fluctuations from the form factor data using the Savitzky--Golay filter (window length 30 and polynomial order 1) and locate the first minimum at  $q>$\,0.1\,\AA$^{-1}$ from both simulation ($F\!F_\mathrm{min}^\mathrm{sim}$) and experiment ($F\!F_\mathrm{min}^\mathrm{exp}$).
